### 2013-06-25 2013_06_26_groupmeeting.tex
Début group meeting 06 Juin 2013

\documentclass{beamer}
\usetheme{Boadilla}
\usecolortheme{crane}
\usefonttheme[stillsansseriflarge, stillsansserifsmall]{serif}
%\usefonttheme[stillsansseriflarge, stillsansserifsmall]{structurebold}
%\usefonttheme[onlymath]{serif}
\useinnertheme{default}

\usepackage{fontspec}
\usepackage{kpfonts}
%\setmainfont[Numbers=OldStyle]{Tex Gyre Pagella}
\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineLight}

\usepackage[frenchb]{babel}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{multirow}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, positioning, intersections, decorations.markings, mindmap, shapes.geometric, matrix, shapes.callouts}

\usepackage{pgfplots}
%\usepgfplotslibrary{units}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\tikzset{external/system call={lualatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"}}

%\tikzexternalize

%bibliography
\usepackage{natbib}
%\usepackage{bibentry}
\def\newblock{\hskip .11em plus .33em minus .07em}

\institute{ENS Lyon}
\title{Réunion hebdomadaire}

\author[M. Leocmach]{Mathieu Leocmach}
\date{26 June 2013}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={small}}
\tikzset{external/force remake=false}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}[plain]
	\titlepage
\end{frame}

\begin{frame}{Relier deux phénomènes}
\begin{columns}
\column{0.5\textwidth}
Fractures sous contrainte

\includegraphics[width=\columnwidth]{gap1mm_Y191_SSC300.jpg}

\column{0.5\textwidth}
Motif sans attache

\includegraphics[width=\columnwidth]{pattern_100um.png}

\end{columns}
\end{frame}

\begin{frame}{Fractures sous contrainte: Fluage}
\begin{columns}
\column{0.6\textwidth}
\begin{tikzpicture}
\begin{loglogaxis}[%
	xlabel={$t\,(s)$},
	ylabel={Taux de déformation $(s^{-1})$},
	]
	\addplot[no marks] file {Y34_strain_rate_log.txt};
	\draw[<->] (axis cs:1e-1,1) -- (axis cs:1.3e3,5e-4) node[midway,above,rotate=-21] {régime d'Andrade};
	\node at (axis cs:1.3e3,1e-4) (det){};
	\node at (axis cs:2.3e3,1e4) (touch){};
\end{loglogaxis}
\draw[<-, red, ultra thick] (det) -- ++(0,-4em) node[left] {Apparition des fractures};
\draw[<-, red, ultra thick] (touch) -- ++(0,3em) node[left] {Les fractures se rejoignent};
\end{tikzpicture}
\column{0.4\textwidth}
\includegraphics[width=\columnwidth, trim=0 0 0 1cm,clip=true]{Y210_SSC300_avant_div.jpg}

\begin{tikzpicture}[inner sep=0]
\node[anchor=north east]{\includegraphics[width=\columnwidth, trim=0 0 0 1cm,clip=true]{Y210_SSC300_premiere_frac.jpg}};
\node[anchor=north east]{\includegraphics[width=0.5\columnwidth,% trim=1cm 6mm 15cm 12cm,clip=true]{Y210_SSC300_premiere_frac.jpg}};
]{detail_premiere_frac.jpg}};
\draw[red,radius=0.05\columnwidth] (-0.775\columnwidth,-0.625\columnwidth) circle;
\draw[red,radius=0.06\columnwidth] (-0.29\columnwidth,-0.25\columnwidth) circle;
\end{tikzpicture}
\end{columns}
\end{frame}

\begin{frame}{Fractures sous contrainte: Décollement}
\begin{tikzpicture}
\begin{axis}[%
	scale only axis,
	xmin=0,xmax=2322,
	ymin=0, ymax=1100,
	ylabel={$r_\text{gap}\,(\mu m)$},
	axis on top,
	xtick={0,500,1000,1500,2000},
	]
	\addplot graphics[
		xmin=0,xmax=2322,
		ymin=0,ymax=1100,
		]{spatioY34};
\end{axis}
\begin{semilogyaxis}[%
	scale only axis,
	xmin=0,xmax=2322,
	%ymin=0, ymax=1100,
	axis y line*=right,
	axis x line=none,
	ylabel={Taux de déformation $(s^{-1})$},
	ylabel near ticks,
	]
	\addplot[no marks] file {Y34_strain_rate_log.txt};
	\node at (axis cs:1.3e3,2e-4) (a) {};
	\draw[green!50!black, dashed] (axis cs:2.18e3,1e-4) -- (axis cs:2.18e3,1e4) node (b) {};
\end{semilogyaxis}
\draw[<-, green!50!black, ultra thick] (a) -- ++(0,-3.5em) node[left] {Fracture au bord};
\draw[<-, green!50!black, ultra thick] (b) -- ++(0,2em) node[left] {Une fracture atteint le niveau du transducteur};
\end{tikzpicture}
\end{frame}

\begin{frame}{Fractures sous contrainte: exp ou pow ?}
\begin{columns}
\column{0.6\textwidth}
\begin{tikzpicture}
\begin{axis}[%
	name=phasediag,
	xlabel near ticks,
	ylabel near ticks,
	xmin=0,xmax=5,xlabel={\% caseinate},
	ymin=0,ymax=5,ylabel={\% GDL},
	legend style={legend pos=north west,},
	]
\addplot[mark=triangle*, green!50!black, only marks] coordinates {(4,3) (4,2) (4,1)};
\addplot[mark=square*, red, only marks] coordinates {(1,1)};
\addplot[mark=diamond*, orange, only marks] coordinates {(4,4)};\legend{puissance, exponentielle, coexistence};
\addplot[no marks, dotted, black] coordinates {(0,0) (5,5) (10,5) (0,0) (8,2)};
\node at (axis cs:4.5,4.5) {1h};
\node at (axis cs:4.5,1.125) {8h};
\end{axis}
\draw[->] ($(phasediag.north east)+(1em,0)$) -- ($(phasediag.south east)+(1em,0)$) node[midway,right] {$G^{\prime}\nearrow$};
\draw[->] ($(phasediag.north west)+(0,1em)$) -- ($(phasediag.north east)+(0,1em)$) node[midway,above] {$G^{\prime}\nearrow$};
\end{tikzpicture}
\column{0.4\textwidth}
$\tau_\text{casse}(\sigma)$
\begin{itemize}
\item \textcolor{green!50!black}{fractures $\Leftrightarrow$ loi de puissance}
\item \textcolor{red}{pas de fracture $\Leftrightarrow$ exponentielle}
\end{itemize}
Pour le 4\% 4\%, fractures et loi de puissance à bas stress, pas de fractures et exponentielle à haut stress.
\end{columns}
\end{frame}

\begin{frame}{Fractures sous contrainte: exp ou pow ?}
\begin{tikzpicture}
\begin{groupplot}[%
	group style={
		group size=2 by 1, 
		y descriptions at=edge left,
		horizontal sep=1em,
		},
	ymode=log,
	ylabel={$\tau_\text{casse}$ (s)},ylabel near ticks,
	xlabel={$\sigma$ (Pa)},
	]
	\nextgroupplot[
		xmode=normal,
		xmin=0,
		legend style={legend pos=north east,},
		]
	\draw (axis cs:200,1e3) -- (axis cs:700,5);
	\draw (axis cs:2,1.7e4) -- (axis cs:40,7.85e-1);
	\addplot+[only marks,mark=triangle, green!50!black] table[x index=0, y index=1]{cas4_GDL1_t_cassure_sigma.txt};
	\addplot+[only marks,mark=diamond, orange] table[x index=0, y index=2]{cas4_GDL4_t_cassure_sigma.txt};
	\addplot+[only marks,mark=square, red] table[x index=0, y index=2]{cas1_GDL1_t_cassure_sigma.txt};
	\legend{4\% 1\%, 4\% 4\%, 1\% 1\%};
	%

	\nextgroupplot[xmode=log]
	\draw (axis cs:150,5e5) -- (axis cs:1000,25);
	\draw (axis cs:30,1e5) -- (axis cs:200,2e3);
	\addplot+[only marks,mark=triangle, green!50!black] table[x index=0, y index=1]{cas4_GDL1_t_cassure_sigma.txt};
	\addplot+[only marks,mark=diamond, orange] table[x index=0, y index=2]{cas4_GDL4_t_cassure_sigma.txt};
	\addplot+[only marks,mark=square, red] table[x index=0, y index=2]{cas1_GDL1_t_cassure_sigma.txt};
\end{groupplot}
\end{tikzpicture}
Note: On voit la transition dans la colonne 4\% de caséine.
\end{frame}

\begin{frame}{Motifs: Ingrédients}

\end{frame}

\begin{frame}{Motifs: Diagramme de phase}
\begin{tikzpicture}
\begin{axis}[%
	name=phasediag,
	xmin=0,xmax=5,xlabel={\% caseinate},
	ymin=0,ymax=5,ylabel={\% GDL},
	legend style={legend pos=north west,}
	]
\addplot[mark=triangle*, green!50!black, only marks,forget plot] coordinates {(4,4) (4,3) (4,2) (4,1)};
\addplot[mark=triangle, green!50!black, only marks] coordinates {(3,3)};
\addplot[mark=square*, red, only marks,forget plot] coordinates {(1,1)};
\addplot[mark=square, red, only marks] coordinates {(2,2)};
\addplot[mark=diamond, orange, only marks] coordinates {(2,1) (3,2) (3,1)};
\legend{Motif, Pas motif, Coexistence};
\addplot[no marks, dotted, black] coordinates {(0,0) (5,5) (10,5) (0,0) (8,2)};
\node at (axis cs:4.5,4.5) {1h};
\node at (axis cs:4.5,1.125) {8h};
\end{axis}
\draw[->] ($(phasediag.north east)+(1em,0)$) -- ($(phasediag.south east)+(1em,0)$) node[midway,right] {$\begin{aligned}
%\tau_\text{prise}&\nearrow\\
G^{\prime}&\nearrow\\
\lambda&\nearrow
\end{aligned}$};
\draw[->] ($(phasediag.north west)+(0,1em)$) -- ($(phasediag.north east)+(0,1em)$) node[midway,above] {$G^{\prime}\nearrow$};
\end{tikzpicture}
\end{frame}

\end{document}

### 2013-06-25 2013_06_26_groupmeeting.tex
Suite group meeting 26 Juin 2013

@@
 %\setmainfont[Numbers=OldStyle]{Tex Gyre Pagella}
-\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineLight}
+\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}
 
@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,shapes,backgrounds, positioning, intersections, decorations.markings, mindmap, shapes.geometric, matrix, shapes.callouts}
+\usetikzlibrary{arrows,shapes,backgrounds, positioning, intersections, decorations.markings, mindmap, shapes.geometric, matrix, shapes.callouts,patterns}
 
@@
 \end{semilogyaxis}
-\draw[<-, green!50!black, ultra thick] (a) -- ++(0,-3.5em) node[left] {Fracture au bord};
+\draw[<-, green!50!black, ultra thick] (a) -- ++(0,-3.5em) node[left] {Première fracture au bord};
 \draw[<-, green!50!black, ultra thick] (b) -- ++(0,2em) node[left] {Une fracture atteint le niveau du transducteur};
@@
 \end{tikzpicture}
-Note: On voit la transition dans la colonne 4\% de caséine.
+Note: Dans le 4\% 4\% le préfacteur dépend du gap.
+\end{frame}
+
+\begin{frame}{Fractures sous contrainte: exp ou pow ?}
+On voit la transition dans la colonne 4\% de caséine.
+\begin{tikzpicture}
+\begin{groupplot}[%
+	group style={
+		group size=2 by 1, 
+		y descriptions at=edge left,
+		horizontal sep=1em,
+		},
+	ymode=log,
+	ylabel={$\tau_\text{casse}$ (s)},ylabel near ticks,
+	xlabel={$\sigma$ (Pa)},
+	]
+	\nextgroupplot[
+		xmode=normal,
+		xmin=0,
+		legend style={legend pos=north east,},
+		]
+	%\draw (axis cs:200,1e3) -- (axis cs:700,5);
+	%\draw (axis cs:2,1.7e4) -- (axis cs:40,7.85e-1);
+	\addplot+[only marks,mark=triangle, green!50!black] table[x index=0, y index=1]{cas4_GDL1_t_cassure_sigma.txt};
+	\addplot+[only marks,mark=*, blue,mark options={fill=blue}] table[x index=1, y index=2]{cassure_cas4_GDL2.csv};
+	\addplot+[only marks,mark=pentagon*, purple,mark options={fill=purple}] table[x index=1, y index=2]{cassure_cas4_GDL3.csv};
+	\addplot+[only marks,mark=diamond, orange] table[x index=0, y index=2]{cas4_GDL4_t_cassure_sigma.txt};
+	\legend{4\% 1\%, 4\% 2\%, 4\% 2\%, 4\% 1\%};
+	%
+
+	\nextgroupplot[xmode=log]
+	\draw (axis cs:150,5e5) -- (axis cs:1000,25);
+	\draw (axis cs:30,1e5) -- (axis cs:200,2e3);
+	\draw (axis cs:100,131182) -- (axis cs:200,16251);
+	\draw (axis cs:100,2e4) -- (axis cs:300,1.7e3);
+	\addplot+[only marks,mark=triangle, green!50!black] table[x index=0, y index=1]{cas4_GDL1_t_cassure_sigma.txt};
+	\addplot+[only marks,mark=*, blue,mark options={fill=blue}] table[x index=1, y index=2]{cassure_cas4_GDL2.csv};
+	\addplot+[only marks,mark=pentagon*, purple,mark options={fill=purple}] table[x index=1, y index=2]{cassure_cas4_GDL3.csv};
+	\addplot+[only marks,mark=diamond, orange] table[x index=0, y index=2]{cas4_GDL4_t_cassure_sigma.txt};
+\end{groupplot}
+\end{tikzpicture}
 \end{frame}
@@
 \begin{frame}{Motifs: Ingrédients}
+\begin{center}
+\begin{tikzpicture}
+\fill[pattern=north east lines,pattern color=blue] (0,0) rectangle (12,0.5) node[midway,fill=white,inner sep=1pt] {verre};
+\fill[pattern=north east lines,pattern color=blue] (0,-1) rectangle (12,-1.5) node[midway,fill=white,inner sep=1pt] {verre};
+\draw[blue,line width=2pt,green!80!black] (0,-1) -- (12,-1) (0,-1pt) -- (12,-1pt) node[below,pos=0.30] {brosse acrylamide ($10\sim 100$ nm)};
+\fill[gray] (0,0) rectangle (0.5,-1) (12,0) rectangle (11.5,-1) node[pos=0.75, left] {spacer};
+\draw[<->] (9,-2pt) -- (9,-1) node[midway,left] {$e\sim 100\,\mu m$};
+\draw[<->] (0.5,0.75) -- (11.5,0.75) node[midway,above] {$L\sim 2.5cm$};
+\end{tikzpicture}
+\end{center}
 
+\begin{itemize}
+\item Pas d'accroche sur les plus grandes surfaces grâce aux brosses.
+\item Accroche sur les extrémités (spacers solides, pas d'interface libre)
+\item Contraintes auto-générées par la séparation de phase/gélification
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Motifs: Formation}
+%\movie{\includegraphics[width=.9\textwidth]{une_image.jpg}}{videos/une_video.avi}
+\end{frame}
+
+\begin{frame}{Motifs: Structure}
+\begin{columns}
+\column{0.6\textwidth}
+\begin{tabular}{ccc}
+\includegraphics[width=0.3\textwidth]{fracture2_z=0um.jpg}&%
+\includegraphics[width=0.3\textwidth]{fracture2_z=15um.jpg}&%
+\includegraphics[width=0.3\textwidth]{fracture2_z=30um.jpg}\\
+\includegraphics[width=0.3\textwidth]{fracture2_z=45um.jpg}&%
+\includegraphics[width=0.3\textwidth]{fracture2_z=60um.jpg}&%
+\includegraphics[width=0.3\textwidth]{fracture2_z=75um.jpg}\\
+\includegraphics[width=0.3\textwidth]{fracture2_z=90um.jpg}&%
+\includegraphics[width=0.3\textwidth]{fracture2_z=105um.jpg}&%
+\includegraphics[width=0.3\textwidth]{fracture2_z=120um.jpg}\\
+\end{tabular} 
+\column{0.3\textwidth}
+\begin{tikzpicture}[inner sep=0]
+\node{\includegraphics[width=\textwidth]{fracture2_x10.jpg}};
+\draw[red] (-0.05\textwidth, -0.05\textwidth) rectangle (0.05\textwidth, 0.05\textwidth);
+\end{tikzpicture}
+
+\bigskip
+
+\begin{tikzpicture}[inner sep=0]
+\node{\includegraphics[width=\textwidth]{coupe_motif.pdf}};
+\node[anchor=west] at (-0.45\textwidth, 0) {gel};
+\draw[<-] (0.1\textwidth, -0.06\textwidth) -- (0.3\textwidth, -0.75) node[anchor=north west] {eau};
+\end{tikzpicture}
+Litterature des rides, crevasses\ldots
+
+Confocal demain.
+\end{columns}
+\end{frame}
+
+\begin{frame}{Motifs: Formation et structure}
+\begin{columns}
+\column{0.6\textwidth}
+\includegraphics[width=\textwidth]{intensities_by_zones.pdf}
+\column{0.4\textwidth}
+\begin{itemize}
+\item Les zones claires deviennent claires (décollement).
+\item Les lignes sombres restent sombres (matière de haut en bas).
+\item L'ensemble s’éclaircit.
+\end{itemize}
+\end{columns}
+L'expulsion d'eau est confirmée par confocal après plusieurs semaines (plus qu'une fine couche de gel).
 \end{frame}

### 2013-06-25 2013_06_26_groupmeeting.tex
Pourquoi on compare

@@
 
+\begin{frame}{Motifs et fractures : longueurs d'onde}
+Dans les compositions où on a une loi de puissance à bas $\sigma$ (et donc des fractures) on observe des motifs.
+Inversement, si la loi est exponentielle (pas de fracture), on n'observe pas de motifs.
+Est-ce que les motifs sont les fractures à $\sigma=0$ ?
+
+Est-ce que $\lambda_f$ et $\lambda_m$ évoluent de la même façon en fonction des paramètres de contrôles : gap, courbure, composition du gel, G', temps de prise, etc. ?
+\end{frame}
 \end{document}

### 2013-06-26 2013_06_26_groupmeeting.tex
Final state

--- 2013-06-25 2013_06_26_groupmeeting.tex
@@
 \usepackage{siunitx}
+\sisetup{
+  mathsmu = \text{µ},
+  textmu  = µ
+}
 \usepackage{booktabs}
@@
 %\tikzexternalize
+
+\usepackage{multimedia}
 
@@
 
-\includegraphics[width=\columnwidth]{pattern_100um.png}
+\includegraphics[width=\columnwidth]{pattern_100um.jpg}
 
@@
 \begin{frame}{Motifs: Formation}
-%\movie{\includegraphics[width=.9\textwidth]{une_image.jpg}}{videos/une_video.avi}
+\begin{columns}
+\column{0.5\textwidth}
+\movie[externalviewer]{\includegraphics[width=\textwidth]{cas4_GDL4_motif_final.jpg}}{cas4_GDL_4_100um_coat_1_1_comp_4.avi}\\
+\begin{tikzpicture}
+\draw(0,0)-- (0.757\textwidth,0) node[midway, below] {1 mm};
+\end{tikzpicture}
+\column{0.5\textwidth}
+\movie[externalviewer]{\includegraphics[width=\textwidth]{pattern_100um.jpg}}{cas4_GDL_4_100um_coat_macroscope_2.avi}\\
+\begin{tikzpicture}
+\draw(0,0)-- (0.444\textwidth,0) node[midway, below] {1 cm};
+\end{tikzpicture}
+\end{columns}
 \end{frame}
@@
 \end{tikzpicture}
-Litterature des rides, crevasses\ldots
+Littérature des rides, crevasses, cusps\ldots
 
@@
 \begin{frame}{Motifs: Diagramme de phase}
+\begin{columns}
+\column{0.6\textwidth}
 \begin{tikzpicture}
@@
 	name=phasediag,
+	width=0.9\textwidth,
 	xmin=0,xmax=5,xlabel={\% caseinate},
 	ymin=0,ymax=5,ylabel={\% GDL},
-	legend style={legend pos=north west,}
+	legend style={legend pos=north west,},
+	xlabel near ticks,
+	ylabel near ticks,
 	]
@@
 \end{tikzpicture}
+\column{0.4\textwidth}
+cas 2\% GDL 1\%\\
+\includegraphics[width=\textwidth]{motif_cas2_GDL1_zoom1.png}
+
+\bigskip
+
+cas 2\% GDL 2\%\\
+\includegraphics[width=\textwidth]{motif_cas3_GDL2_zoom1.png}
+\end{columns}
+
 \end{frame}
 
+\begin{frame}{Motifs et fractures : longueurs d'onde}
+\begin{itemize}
+\item Dans les compositions où on a une loi de puissance à bas $\sigma$ (et donc des fractures) on observe des motifs. 
+\item Inversement, si la loi est exponentielle (pas de fracture), on n'observe pas de motifs. 
+\item[$\Rightarrow$] Est-ce que les motifs sont les fractures à $\sigma=0$ ?
+\end{itemize}
+
+Est-ce que $\lambda_\text{fracture}$ et $\lambda_\text{motif}$ évoluent de la même façon en fonction des paramètres de contrôles : gap, courbure, composition du gel, G', temps de prise, etc. ?
+\end{frame}
+
+\begin{frame}{Longueurs d'onde en fonction du gap}
+\begin{columns}\small
+\column{0.5\textwidth}
+\begin{tabular}{cc}
+\SI{50}{\micro\metre} & \SI{100}{\micro\metre}\\
+\includegraphics[width=0.4\textwidth]{pattern_50um.jpg} & \includegraphics[width=0.4\textwidth]{pattern_100um.jpg} \\ 
+\SI{250}{\micro\metre} & \SI{450}{\micro\metre}\\
+\includegraphics[width=0.4\textwidth]{pattern_250um.jpg} & \includegraphics[width=0.4\textwidth]{pattern_450um.jpg} \\
+\SI{800}{\micro\metre} & \begin{tikzpicture}
+\draw[ultra thick](0,0)-- (0.177\textwidth,0) node[midway, above, inner sep=0] {1 cm};
+\end{tikzpicture}\\
+\multicolumn{2}{c}{\includegraphics[width=0.8\textwidth]{pattern_800um_x1.jpg}} \\ 
+\end{tabular}
+4\% 4\%
+
+\column{0.5\textwidth}
+\begin{tabular}{cc}
+\SI{0.5}{\milli\metre} & \SI{1}{\milli\metre}\\
+\includegraphics[width=0.4\textwidth, trim=2cm 5mm 3cm 4cm,clip=true]{gap500um_Y195_SSC300.jpg}&
+\includegraphics[width=0.4\textwidth, trim=2cm 5mm 3cm 4cm,clip=true]{gap1mm_Y191_SSC300.jpg}\\
+\SI{1.5}{\milli\metre} & \SI{3}{\milli\metre}\\
+\includegraphics[width=0.4\textwidth, trim=2cm 5mm 3cm 4cm,clip=true]{gap1500um_Y194_SSC300.jpg}&
+\includegraphics[width=0.4\textwidth, trim=2cm 5mm 3cm 4cm,clip=true]{gap3mm_Y202_SSC40.jpg}\\
+\end{tabular}
+\begin{tikzpicture}
+\begin{axis}[%
+	width=0.8\textwidth,
+	xmin=0,xmax=3,xlabel={gap (\si{\milli\metre})},
+	ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+	xlabel near ticks,
+	ylabel near ticks,
+	legend style={legend pos=north west,}
+	]
+	\addplot+[only marks] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+	\addplot+[domain=0:3, no marks, blue, forget plot] {4.23*x};
+	\addplot+[only marks,error bars/.cd, y dir=both,y explicit] coordinates {
+	(0.5,1.25) +- (0,0.14)
+	(1,2.19) +- (0,0.2)
+	(1.5,3) +- (0,0.2)
+	(2,3.3) +- (0,0.2)
+	(3,4.3) +- (0,0.3)};
+	\legend{motif, fractures};
+\end{axis}
+\end{tikzpicture}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Longueurs d'onde et composition}
+\begin{columns}[T]\small
+\column{0.5\textwidth}
+4\% 4\%\\
+\includegraphics[width=0.8\textwidth]{pattern_100um.jpg}\\
+Motifs très reproductibles
+
+\bigskip
+
+\begin{tikzpicture}
+\begin{axis}[%
+	width=0.9\textwidth,
+	xmin=0,xmax=3,xlabel={gap (\si{\milli\metre})},
+	ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+	xlabel near ticks,
+	ylabel near ticks,
+	legend style={legend pos=north west,}
+	]
+	\addplot+[only marks] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+	\addplot+[domain=0:3, no marks, blue, forget plot] {4.23*x};
+	\addplot+[only marks,error bars/.cd, y dir=both,y explicit] coordinates {
+	(0.5,1.25) +- (0,0.14)
+	(1,2.19) +- (0,0.2)
+	(1.5,3) +- (0,0.2)
+	(2,3.3) +- (0,0.2)
+	(3,4.3) +- (0,0.3)};
+	\addplot+[only marks,error bars/.cd, y dir=both,y explicit] coordinates {
+	(0.5,1.25) +- (0,0.1)
+	(1,1.8) +- (0,0.2)
+	(1.5,2.55) +- (0,0.05)
+	(2,3.6) +- (0,0.5)};
+	\legend{motif 4\% 4\%, frac. 4\% 4\%, frac. 4\% 1\%};
+\end{axis}
+\end{tikzpicture}
+
+\column{0.5\textwidth}
+4\% 1\%\\
+\includegraphics[width=0.8\textwidth]{motif_cas4_GDL1.png}
+\begin{itemize}
+\item Difficile d'obtenir des motifs à \SI{100}{\micro\metre}, 
+\item Pas de motif à plus petit gap.
+\end{itemize}
+\end{columns}
+\end{frame}
 \end{document}

### 2013-09-04 poster_Rome.tex
Poster for Rome. Color theme and intro

% We switch to portrait mode. This works as advertised.
\documentclass[a0,portrait]{a0poster}
% You might find the 'draft' option to a0 poster useful if you have
% lots of graphics, because they can take some time to process and
% display. (\documentclass[a0,draft]{a0poster})

%\usepackage[utf8]{inputenc}

% Switch off page numbers on a poster, obviously, and section numbers too.
\pagestyle{empty}
\setcounter{secnumdepth}{0}

%fonts
\usepackage[T1]{fontenc}
\usepackage{fontspec}
%\usepackage[oldstylenums, largesmallcaps]{kpfonts}
\setmainfont[Numbers=OldStyle]{Tex Gyre Pagella}
\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}
%\renewcommand*\sfdefault{ugq}

\usepackage{hyperref}
\hypersetup{%
	pdftitle={Effects of a real world size distribution on the heterogeneous crystallisation of hard sphere colloids},%the title
	pdfauthor={Mathieu Leocmach},%your name
}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{siunitx}

\usepackage{multirow}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, positioning, intersections, decorations.markings, decorations.shapes, mindmap, shapes.geometric, matrix, patterns}

\usepackage{pgfplots}
%\usepgfplotslibrary{units}
\usepgfplotslibrary{groupplots}
\pgfplotsset{every axis/.append style={xlabel near ticks,ylabel near ticks}}
\pgfplotsset{every axis plot post/.append style={very thick}}

\usepgfplotslibrary{external}
%\tikzexternalize
%\tikzsetexternalprefix{fig_Rome/}
\tikzset{external/system call={lualatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"}}

\usepackage{ragged2e}
\RaggedRight

\definecolor{Main}{rgb}{1, 0.66, 0}
\definecolor{Accent1}{rgb}{1,0.45,0}
\definecolor{Accent2}{rgb}{0.88,0.66,0}

% see documentation for a0poster class for the size options here
\let\Textsize\normalsize
\def\Head#1{\noindent\hbox to \hsize{\hfil{\LARGE\color{Main}\raggedright\textsf{#1}}}\bigskip}
\def\LHead#1{\noindent{\LARGE #1}\smallskip}
\def\Subhead#1{\noindent{\large\color{Accent1}\textsc{#1}}}
\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{#1}}}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.
\usepackage[absolute,overlay,showboxes
]{textpos}
% Set up the grid
%
% Note that [40mm,40mm] is the margin round the edge of the page --
% it is _not_ the grid size. That is always defined as 
% PAGE_WIDTH/HGRID and PAGE_HEIGHT/VGRID. In this case we use
% 15 x 25. This gives us a wide central column for text (7 grid
% spacings) and two narrow columns (3 each) at each side for 
% pictures, separated by 1 grid spacing.
%
% Note however that texblocks can be positioned fractionally as well,
% so really any convenient grid size can be used.
%
\TPGrid[40mm,40mm]{15}{25}  % 3 - 1 - 7 - 1 - 3 Columns

% Mess with these as you like
\parindent=0pt
%\parindent=1cm
\parskip=0.5\baselineskip

\usepackage{paralist}

%bibliography
\usepackage{natbib}
\usepackage{bibentry}
\def\newblock{\hskip .11em plus .33em minus .07em}

%\includeonly{}

\begin{document}
%\tikzset{every mark/.append style={scale=0.8}}
%\pgfplotsset{every axis/.append style={small}}

\bibliographystyle{notitle}
%\nobibliography{sift}

% Understanding textblocks is the key to being able to do a poster in
% LaTeX. In
%
%    \begin{textblock}{wid}(x,y)
%    ...
%    \end{textblock}
%
% the first argument gives the block width in units of the grid
% cells specified above in \TPGrid; the second gives the (x,y)
% position on the grid, with the y axis pointing down.

% You will have to do a lot of previewing to get everything in the 
% right place.

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.
\begin{textblock}{15}(0,0)
\centering
\Title{
Wrinkling and yielding of protein gel
}

\LHead{Mathieu Leocmach}\qquad\texttt{\color{Accent1}mathieu.leocmach@polytechnique.org}\\
\LHead{Mathieu Nespoulous, Christophe Perge, Nicolas Taberlet, Thomas Gibaud,
Thibaut Divoux, Sébastien Manneville}

\LHead{\textsc{Laboratoire de Physique -- UMR 5672, Ecole Normale Supérieure de Lyon}}

\end{textblock}

\begin{textblock}{3}(0,2.5)
	\Head{Casein gel}

	In millipore water
	\begin{itemize}
	\item[4\%] Sodium caseinate (milk protein)\\
	Isoelectric point $\approx 4.6$
	\item[4\%] $\delta$-gluconolactone (\textsc{gdl})\\
	Slow hydrolysis into gluconic acid
	\end{itemize}

	\begin{tikzpicture}
	\begin{axis}[%
		xmin=0,
		ymin=0, ymax=7,
		xlabel={time (h)}, ylabel={\textcolor{Accent1}{pH}},
		extra tick style={grid=major},%
		extra y ticks={4.6}, extra y tick labels={},%
		axis y line*=left,
		width=0.7\textwidth,
		]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
	\end{axis}
	\begin{axis}[%
		xmin=0,
		ymin=0,
		ylabel=\textcolor{Accent2}{$G^\prime$},
		axis y line*=right,
		axis x line=none,
		width=0.7\textwidth,
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
	\end{axis}
	\end{tikzpicture}
\end{textblock}

\begin{textblock}{7}(4,2.5)
	\Head{Sealed cell}

	\tikzsetnextfilename{cell_brushes}
	\tikzset{external/force remake}
	\begin{tikzpicture}
		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\textwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\textwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
		\draw[line width=2pt,Accent1] (0.05\textwidth,-2.5em) -- (0.95\textwidth,-2.5em) (0.05\textwidth,-1pt) -- (0.95\textwidth,-1pt) node[below,pos=0.30] {acrylamide brush ($10\sim 100$ nm)};
		\fill[gray] (0,0) rectangle (0.05\textwidth,-2.5em) (\textwidth,0) rectangle (0.95\textwidth,-2.5em) node[pos=1, above left] {spacer};
		\draw[<->] (0.75\textwidth,-2pt) -- (0.75\textwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] {$L\sim 2.5\,cm$};
	\end{tikzpicture}

	\begin{itemize}
	\item No adhesion on top and bottom (polymer brushes).
	\item Adhesion at the edge (solid spacers, no free interface)
	\end{itemize}

	\begin{tikzpicture}[inner sep=0, anchor=north west, thick]
	\node (a) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2}};
	\node at (a.north east) (b) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}};
	\node at (b.north east) (c) {\includegraphics[height=65mm]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
	\draw[Accent2] (36.4mm, -25mm) rectangle +(28.9mm,-21.7mm);
	\draw[Main] (b.north west) ++(34mm,-34mm) rectangle +(27mm,-19.7mm);
	\draw[Accent1] (c.south west) ++(29mm,29mm) rectangle +(36mm,36mm);
	%\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}
	%\includegraphics[height=65mm]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}
	\end{tikzpicture}

\end{textblock}

\begin{textblock}{3}(8,4.5)
	%\Head{Size distribution}

\end{textblock}

\begin{textblock}{7}(0,8)
	\Head{Single-scale tracking\ldots of polydisperse particles}

\end{textblock}

\begin{textblock}{3}(8,8)
	\Head{Scale space}

\end{textblock}

\begin{textblock}{3}(12,8)
	\Head{Multiscale tracking}

\end{textblock}
\begin{textblock}{4}(0,15.85)
\end{textblock}

\begin{textblock}{3.75}(11.25,15.85)

\end{textblock}

\begin{textblock}{3}(0,12.1)
	\Head{Buoyancy control}
	\Subhead{by temperature control}
\end{textblock}

\begin{textblock}{7}(4,12.1)
	\Head{Size distributions within local structures}

\end{textblock}

\begin{textblock}{15}(0,21.5)
	\Head{Heterogeneous nucleation, Exclusion of small particles \& Formation of grain boundaries}

\end{textblock}

\textblockcolour{lightgray!50!white}
\TPMargin*{ 0.125\TPHorizModule }
\begin{textblock}{2.875}(12,2.625)%real width of 3
	\Head{Take away}
	\Subhead{A new kind of wrinkling}
	\begin{itemize}
		\item nested
		\item self generated
	\end{itemize}
	\Subhead{A link with fractures}
\end{textblock}%Conclusion

\end{document}

### 2013-09-05 poster_Rome.tex
Kinetic

@@
 
-\definecolor{Main}{rgb}{1, 0.66, 0}
-\definecolor{Accent1}{rgb}{1,0.45,0}
-\definecolor{Accent2}{rgb}{0.88,0.66,0}
+\definecolor{Main}{rgb}{1, 0.57, 0}
+\definecolor{Accent1}{rgb}{1,0.28,0}
+\definecolor{Accent2}{rgb}{1,0.74,0}
 
@@
 		extra y ticks={4.6}, extra y tick labels={},%
+		extra x ticks={0.55, 0.64, 0.72, 0.80, 1, 1.27, 2.5}, extra x tick labels={},%
 		axis y line*=left,
@@
 \begin{textblock}{7}(4,2.5)
-	\Head{Sealed cell}
+	\Head{Sealed cell, spontaneous pattern}
 
@@
 
-	\begin{tikzpicture}[inner sep=0, anchor=north west, thick]
-	\node (a) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2}};
-	\node at (a.north east) (b) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}};
-	\node at (b.north east) (c) {\includegraphics[height=65mm]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
-	\draw[Accent2] (36.4mm, -25mm) rectangle +(28.9mm,-21.7mm);
-	\draw[Main] (b.north west) ++(34mm,-34mm) rectangle +(27mm,-19.7mm);
+	\begin{tikzpicture}[inner sep=0, very thick]
+	\node[anchor=west] (a) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2}};
+	\node at (3.5\TPHorizModule,0) (b) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}};
+	\node[anchor=east] at (7\TPHorizModule,0) (c) {\includegraphics[height=65mm]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
+	\draw[Accent2] (a.north west) ++(36.4mm, -25mm) rectangle +(28.9mm,-21.7mm) -- (b.south west) (a.north west) ++(65.3mm, -25mm) -- (b.north west);
+	\draw[Main] (b.north west) ++(34mm,-34mm) rectangle +(27mm,-19.7mm) -- (c.south west) (b.north west) ++(61mm,-34mm) --(c.north west);
 	\draw[Accent1] (c.south west) ++(29mm,29mm) rectangle +(36mm,36mm);
@@
 
-\begin{textblock}{3}(8,4.5)
-	%\Head{Size distribution}
-
+\begin{textblock}{15}(0,8)
+	\Head{Kinetic by transmission macroscope}
+	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccccccc}
+	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
+	\includegraphics[width=2\TPHorizModule]{prise_0100_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{prise_0130_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{prise_0160_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{prise_0190_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{prise_0250_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{prise_0360_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{prise_0799_resized.jpg}\\
+	\end{tabular*}
 \end{textblock}
 
-\begin{textblock}{7}(0,8)
-	\Head{Single-scale tracking\ldots of polydisperse particles}
-
+\begin{textblock}{3}(0,10)
+	\Subhead{Length scales}
+	\begin{itemize}
+	\item A fundamental that does not evolve
+	\item Nested structures as ``harmonics''
+	\end{itemize}
+	\Subhead{Qualitatively different from}
+	\begin{itemize}
+	\item phase separation patterns 
+	\item Turing patterns
+	\item usual wrinkles
+	\end{itemize}	 
 \end{textblock}
 
-\begin{textblock}{3}(8,8)
-	\Head{Scale space}
-
+\begin{textblock}{11}(4,11)
+	\Head{Confocal Microscopy}
 \end{textblock}
 
-\begin{textblock}{3}(12,8)
-	\Head{Multiscale tracking}
-
-\end{textblock}
-\begin{textblock}{4}(0,15.85)
+\begin{textblock}{3}(4,11)
+	\addvspace{2em}
+	\Subhead{Procedure}
+	\begin{itemize}
+		\item Fluorescently label 20\% of the casein
+		\item A confocal slice every \SI{5}{\micro\metre}
+		\item $(\SI{1.3}{\milli\metre})^2 \times \SI{120}{\micro\metre}$
+		\item[$\Rightarrow$] Protein density field
+		\item Detect top and bottom surfaces of the gel phase
+	\end{itemize}
 \end{textblock}
 
-\begin{textblock}{3.75}(11.25,15.85)
+\begin{textblock}{2}(8,11)
+	\addvspace{2em}
+	\Subhead{Final state}
 
+	\includegraphics[width=\textwidth]{collage_cut}
+
+	Tilled, enhanced relief
 \end{textblock}
 
-\begin{textblock}{3}(0,12.1)
-	\Head{Buoyancy control}
-	\Subhead{by temperature control}
+\begin{textblock}{4}(11,11)
+	\addvspace{2em}
+	\Subhead{Volume}
+
+	\begin{tikzpicture}
+	\begin{axis}[%
+		xmin=0, xtick={20,40,...,160},
+		ymin=0.5, ymax=1, 
+		xlabel={time (min)}, ylabel={$V_\text{gel}/V_\text{cell}$},
+		extra tick style={grid=major},%
+		extra x ticks={33, 38, 43, 48, 58, 77, 150}, extra x tick labels={},%
+		%width=2.65\TPHorizModule,
+		width=3.65\TPHorizModule,
+		%width=0.8\textwidth, 
+		height=1.5\TPVertModule,
+		]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60.+3}]{cas4_GDL4_fluo0p4_2.volume};
+	\end{axis}
+	\end{tikzpicture}
 \end{textblock}
 
-\begin{textblock}{7}(4,12.1)
-	\Head{Size distributions within local structures}
-
+\begin{textblock}{15}(0,14)
+	\Head{Kinetic by confocal microscope}
+	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccccccc}
+	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
+	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
+\end{tabular*}
 \end{textblock}
 
-\begin{textblock}{15}(0,21.5)
+\begin{textblock}{15}(0,16)
 	\Head{Heterogeneous nucleation, Exclusion of small particles \& Formation of grain boundaries}

### 2013-09-06 poster_Rome.tex
Composition and thickness Creep, fractures, USV2D

@@
 %\usepackage[oldstylenums, largesmallcaps]{kpfonts}
-\setmainfont[Numbers=OldStyle]{Tex Gyre Pagella}
+%\setmainfont[Numbers=OldStyle]{Tex Gyre Pagella}
+\setmainfont{Tex Gyre Pagella}
 \setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}
@@
 
-\begin{textblock}{15}(0,16)
-	\Head{Heterogeneous nucleation, Exclusion of small particles \& Formation of grain boundaries}
+\begin{textblock}{4}(0,16)
+	\Head{Composition}
+	\Subhead{Less gdl}
+	\begin{itemize}
+		\item slower kinetics
+		\item stronger gel
+		\item weaker overshoot
+		\item same fundamental \linebreak wavelength
+		\item less harmonics
+	\end{itemize}
+	\Subhead{More gdl}
+	\begin{itemize}
+		\item plasticity
+	\end{itemize}
+\end{textblock}
+\begin{textblock}{2}(2,16)
+	\addvspace{4.4em}
+	\hfill bottom surface, \textsc{gdl} 5\%\\
+	\includegraphics[width=2\TPHorizModule]{plastic_confocal}
+\end{textblock}
 
+\begin{textblock}{4}(0,20)
+	\Head{Thickness}
+	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccb{0.45\textwidth}}
+	\SI{50}{\micro\metre} & \SI{100}{\micro\metre}&
+	\multirow{4}{*}{\begin{tikzpicture}
+	\node[inner sep=0] (a) {\includegraphics[width=0.45\textwidth]{pattern_800um_x1.jpg}};
+	\node[above=0 of a] {\SI{800}{\micro\metre}};
+	\draw[ultra thick](a.north west) ++(0,2.25em) -- ++(0.0797\textwidth,0) node[midway, above, inner sep=0] {\SI{1}{\centi\metre}};
+	\end{tikzpicture}}\\
+
+	\includegraphics[width=0.225\textwidth]{pattern_50um.jpg} & \includegraphics[width=0.225\textwidth]{pattern_100um.jpg}&
+	\\
+
+	\SI{250}{\micro\metre} & \SI{450}{\micro\metre} &\\
+
+	\includegraphics[width=0.225\textwidth]{pattern_250um.jpg} & \includegraphics[width=0.225\textwidth]{pattern_450um.jpg}& \\
+	\end{tabular*}
+
+	\begin{tikzpicture}
+	\begin{axis}[%
+		width=0.95\textwidth,
+		height=0.4\textwidth,
+		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
+		xtick={0.1,0.3,...,0.9},
+		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		xlabel near ticks,
+		ylabel near ticks,
+		mark size={0.01\textwidth},
+		%legend style={legend pos=north west,}
+		]
+		\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+		\addplot+[domain=0:3, no marks, Accent1, forget plot] {4.23*x};
+	\end{axis}
+	\end{tikzpicture}
+\end{textblock}
+
+\begin{textblock}{4}(5,16)
+	\Head{Creep and fractures}
+	Couette cell, no brush, the gel sticks
+
+	\begin{tikzpicture}
+	\begin{loglogaxis}[%
+		name=a,
+		xlabel={$t\,(s)$},
+		ylabel={Shear rate $(s^{-1})$},
+		]
+		\addplot+[no marks, Accent2] file {Y34_strain_rate_log.txt};
+		\draw[<->] (axis cs:1e-1,1) -- (axis cs:1.3e3,5e-4) node[midway,above,rotate=-21] {Andrade};
+		\node at (axis cs:1.3e3,1e-4) (det){};
+		\node at (axis cs:2.3e3,1e4) (touch){};
+	\end{loglogaxis}
+	\draw[<-, Accent1, ultra thick] (det) -- ++(0,-4em) node[left] {Fractures appear};
+	\draw[<-, Accent1, ultra thick] (touch) -- ++(0,3em) node[left] {Fractures meet};
+	\end{tikzpicture}\hfill
+	\begin{tikzpicture}[inner sep=0]
+		\node[above right] at (0,1em) {\includegraphics[width=1.5\TPHorizModule, trim=0 6mm 2cm 1cm,clip=true]{Y210_SSC300_avant_div.jpg}};
+		\node[below right] (b){\includegraphics[width=1.5\TPHorizModule, trim=0 6mm 2cm 1cm,clip=true]{Y210_SSC300_premiere_frac.jpg}};
+		\node[below left=0 of b.north east]{\includegraphics[width=0.75\TPHorizModule]{detail_premiere_frac.jpg}};
+		\draw[Accent2,radius=0.07\TPHorizModule] ++(b.north east) +(-1.14\TPHorizModule,-1.05\TPHorizModule) circle;
+		\draw[Accent2,radius=0.09\TPHorizModule] ++(b.north east) +(-0.435\TPHorizModule,-0.375\TPHorizModule) circle;
+	\end{tikzpicture}
+\end{textblock}
+
+\begin{textblock}{5}(10,16)
+	\Head{Ultrasound speckles}
+	\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group size=2 by 1, 
+			horizontal sep=\TPHorizModule,
+		},
+		width=2\TPHorizModule,
+		xmin=0,xmax=200,
+		xlabel={t},
+		axis on top,
+		xtick={28, 100, 172},
+		xticklabels={49h30, 50h, 50h30},
+		]
+		\nextgroupplot[ymin=0, ymax=32,	ylabel={$z (\si{\milli\metre})$}, title={$r=\SI{200}{\micro\metre}$}]
+		\addplot graphics[
+			xmin=0,xmax=200,
+			ymin=0,ymax=32,
+			]{Y137_speckleBF_tz.png};
+		\plot[no marks,Accent2]  {32-1/(x-187)};
+		\draw[Accent1] (axis cs:187,0)--(axis cs:187,32);
+
+		\nextgroupplot[ymin=0, ymax=2,	ylabel={$r (\si{\milli\metre})$}, title={$z=\SI{8}{\milli\metre}$}]
+		\addplot graphics[
+			xmin=0,xmax=200,
+			ymin=0,ymax=2,
+			]{Y137_speckleBF_tr.png};
+			\draw[Accent2] (axis cs:172,0)--(axis cs:172,2) node[pos=0.66, left, font=\small] {detach from the rotor};
+			\draw[Accent1] (axis cs:187,0)--(axis cs:187,2) node[pos=0.33, left, font=\small] {global yield};
+	\end{groupplot}
+	%\draw[<-,Accent2, very thick] (decol) -- +(0,1em);
+	%\draw[<-,Accent1, very thick] (casse) -- +(0,1em);
+	\end{tikzpicture}
 \end{textblock}
@@
 	\Head{Take away}
-	\Subhead{A new kind of wrinkling}
+	\Subhead{A new kind of pattern}
 	\begin{itemize}
@@
 		\item self generated
+		\item boundary conditions
 	\end{itemize}

### 2013-09-09 poster_Rome.tex
parabolic fit of tz speckle decodehence
Link wrinkles -> fractures

@@
 \begin{textblock}{15}(0,8)
-	\Head{Kinetic by transmission macroscope}
+	\Head{Pattern kinetics by transmission macroscope}
 	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccccccc}
@@
 \begin{textblock}{3}(0,10)
+	Corresponds with the fall of $G^\prime$.
+
 	\Subhead{Length scales}
@@
 	\addvspace{2em}
-	\Subhead{Volume}
+	\Subhead{Synæresis and swelling}
 
@@
 \begin{textblock}{15}(0,14)
-	\Head{Kinetic by confocal microscope}
+	\Head{Wrinkling kinetics by confocal microscope}
 	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccccccc}
@@
 	\Head{Creep and fractures}
-	Couette cell, no brush, the gel sticks
+	Couette cell, no brush, the gel sticks, no pattern.\\
+	Wait \SI{8}{\hour} for equilibration, then apply constant stress.
 
@@
 	\end{tikzpicture}
+
+	\begin{tabular}{b{0.55\textwidth}b{0.45\textwidth}}
+	\begin{tikzpicture}
+	\begin{loglogaxis}[%
+		ylabel={Breaking time (s)},ylabel near ticks,
+		xlabel={Stress $\sigma$ (Pa)},
+		xmin=10,
+		legend style={font=\small\sc}, legend pos=south west,
+		mark size={0.2em},
+		]
+		\addplot+[only marks, Accent1, mark options={fill=Accent1}] coordinates {(40, 67225) (50,29860) (100,398) (200,4)};
+		\addplot+[only marks, Accent2, mark options={fill=Accent2}] table[x index=0, y index=1]{cas4_GDL1_t_cassure_sigma.txt};
+		\legend{gdl 4\%, gdl 1\%};
+		\addplot+[Accent1, no marks, forget plot, domain=40:200] gnuplot {6.32e+14*x^{-6.14}};
+		\addplot+[Accent2, no marks, forget plot, domain=150:1000] gnuplot {3.93e+17*x^{-5.44}};
+	\end{loglogaxis}
+	\end{tikzpicture}&
+	Irreversible fractures\linebreak
+	\Subhead{Fracture\linebreak wavelength}
+	\begin{itemize}
+		\item stress-independent
+		\item composition-indep.
+		\item $\propto$ gap
+	\end{itemize}\\
+	\end{tabular}
+
+	\begin{tikzpicture}
+	\begin{axis}[%
+		width=0.95\textwidth,
+		height=0.4\textwidth,
+		xmin=0,xmax=3.5,xlabel={gap (\si{\milli\metre})},
+		xtick={0, 0.5,...,3},
+		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		xlabel near ticks,
+		ylabel near ticks,
+		%mark size={0.01\textwidth},
+		legend cell align=left,
+		legend style={font=\small}, legend pos=north west,
+		]
+		\addplot+[%
+			only marks, Accent1, mark options={fill=Accent1},
+			error bars/.cd,
+			y dir=both,y explicit,
+			x dir=none,x fixed=0.05,
+			]
+			coordinates {
+				(0.5,0.6) +-(0,0.12)
+				(1, 1.33) +-(0,0.14)
+				(1.5,1.5) +-(0,0.14)
+				(1.5, 1.25) +- (0,0.77)
+				(3,3.75) +- (0,0.4)
+				(3, 4.2) +-(0,0.4)};
+		\addplot+[%
+			only marks, Main, mark options={fill=Main},
+			error bars/.cd,
+			y dir=both,y explicit,
+			x dir=none,x fixed=0.05,
+			]
+			coordinates
+			{(2.5,3.2) +- (0,0.84)};
+		\addplot+[%
+			only marks, Accent2, mark options={fill=Accent2}, mark=triangle,
+			error bars/.cd,
+			y dir=both,y explicit,
+			x dir=none,x fixed=0.05,
+			]
+			coordinates
+			{(2,2.29) +- (0,0.13)};
+		\legend{$R_\text{out}=\SI{25}{\milli\metre}$, $R_\text{out}=\SI{37.5}{\milli\metre}$, $R_\text{out}=\SI{40}{\milli\metre}$};
+		\addplot+[domain=0:3.5, no marks, Accent1, forget plot] {1.25*x};
+	\end{axis}
+	\end{tikzpicture}
 \end{textblock}
@@
 	\Head{Ultrasound speckles}
+	\Subhead{Probe deformation inside the gap}
+
+	\begin{tikzpicture}[ultra thick, font=\small]
+	\node[above right, inner sep=0] (sch) {\includegraphics[width=0.4\textwidth]{usv_3D_image}};
+	\draw[<-] (0.3\textwidth,0.11\textwidth) -- +(1em,0) node[right] {water in};
+	\draw[->] (0.3\textwidth,0.07\textwidth) -- +(1em,0) node[right] {water out};
+	\node[right] at (0.2\textwidth, 0.3\textwidth) (rh) {rheometer};
+	\node[above right] at (0.2\textwidth, 0) (arraylabel) {array of transducers};
+	\draw[->,dotted]	 (arraylabel.west) -- (0.1\textwidth, 0.1\textwidth);
+	\draw[->] (0.06\textwidth, 0.15\textwidth) -- (0,0.15\textwidth) node[draw, left, text width=0.18\textwidth] (mce) {multi-channel electronics};
+	\node[above=of mce, draw] (pc) {PC};
+	\draw[->] (mce)-- (pc);
+	\draw[->] (rh) -- +(0,1em) -| (pc.north);
+	\node[below right=1em and 0 of mce.south west, text width=0.18\textwidth] {Gallot et al. RSI (2013).};
+	\node[below right =of sch.north east, text width=0.32\textwidth] {3\% of \SI{30}{\micro\metre} acrylamide beads as contrast agent};
+	\end{tikzpicture}
+
 	\begin{tikzpicture}
@@
 		group style={
+			group name=grp,
 			group size=2 by 1, 
@@
 			]{Y137_speckleBF_tz.png};
-		\plot[no marks,Accent2]  {32-1/(x-187)};
+		\addplot[no marks,Accent2, domain=0:32]  ({187-0.272*(x-17.6)^2}, x);
 		\draw[Accent1] (axis cs:187,0)--(axis cs:187,32);
@@
 			]{Y137_speckleBF_tr.png};
-			\draw[Accent2] (axis cs:172,0)--(axis cs:172,2) node[pos=0.66, left, font=\small] {detach from the rotor};
+			\draw[Accent2] (axis cs:172,0)--(axis cs:172,2) node[pos=0.66, left, font=\small] {free from the rotor};
 			\draw[Accent1] (axis cs:187,0)--(axis cs:187,2) node[pos=0.33, left, font=\small] {global yield};
 	\end{groupplot}
-	%\draw[<-,Accent2, very thick] (decol) -- +(0,1em);
-	%\draw[<-,Accent1, very thick] (casse) -- +(0,1em);
+	\node[left=of grp c2r1.south west] {rotor};
+	\node[left=of grp c2r1.north west] {stator};
 	\end{tikzpicture}
+
+	Fractures appear when the gel detaches from the rotor. Creates a free surface that allows synæresis, relaxation of the gelation-induced \linebreak stress and possibly wrinkling.
+\end{textblock}
+
+\begin{textblock}{5}(10,22.5)
+	\Head{Non equilibrium Creep}
+	\Subhead{Fractures need complete overshoot}
+	\begin{itemize}
+		\item Wait only \SI{40}{\minute}. $G^\prime$ at maximum.
+		\item Start creep with $\sigma > \SI{200}{\pascal}$.
+		\item[$\Rightarrow$] Break before $G^\prime$ falls.
+		\item[$\Rightarrow$] No fracture
+	\end{itemize}
 \end{textblock}
@@
 	\end{itemize}
+	\Subhead{A good model for solids}
+	\begin{itemize}
+		\item Andrade creep over 6 decades in time
+		\item Irreversible fractures
+	\end{itemize}
 	\Subhead{A link with fractures}
+	\begin{itemize}
+		\item Wrinkles can induce plasticity
+		\item Wavelengths have the same dependence
+		\item Detachment creates a free surface
+		\item Fractures need complete overshoot
+	\end{itemize}
 \end{textblock}%Conclusion

### 2013-09-10 poster_Rome.tex
Point @ 300Pa \dotsfill to see more clearly paragraphs

@@
 \let\Textsize\normalsize
-\def\Head#1{\noindent\hbox to \hsize{\hfil{\LARGE\color{Main}\raggedright\textsf{#1}}}\bigskip}
+\def\Norulehead#1{\noindent\hbox to \hsize{\hfil\LARGE\textcolor{Main}{\textsf{#1}}}\bigskip}
+\def\Head#1{\Norulehead{\dotfill #1}}
 \def\LHead#1{\noindent{\LARGE #1}\smallskip}
@@
 % places on the page.
-\usepackage[absolute,overlay,showboxes
+\usepackage[absolute,overlay,%showboxes
 ]{textpos}
@@
 
+	\vspace{1em}
+
 	\begin{tikzpicture}
@@
 		]
-		\addplot+[only marks, Accent1, mark options={fill=Accent1}] coordinates {(40, 67225) (50,29860) (100,398) (200,4)};
+		\addplot+[only marks, Accent1, mark options={fill=Accent1}] coordinates {(40, 67225) (50,29860) (100,398) (200,4) (300,0.06)};
 		\addplot+[only marks, Accent2, mark options={fill=Accent2}] table[x index=0, y index=1]{cas4_GDL1_t_cassure_sigma.txt};
 		\legend{gdl 4\%, gdl 1\%};
-		\addplot+[Accent1, no marks, forget plot, domain=40:200] gnuplot {6.32e+14*x^{-6.14}};
+		\addplot+[Accent1, no marks, forget plot, domain=40:300] gnuplot {9.2e+15*x^{-6.8}};
 		\addplot+[Accent2, no marks, forget plot, domain=150:1000] gnuplot {3.93e+17*x^{-5.44}};
@@
 \begin{textblock}{2.875}(12,2.625)%real width of 3
-	\Head{Take away}
+	\Norulehead{Take away}
 	\Subhead{A new kind of pattern}

### 2013-09-10 poster_Rome.tex
Scale bars

@@
 	\draw[Accent1] (c.south west) ++(29mm,29mm) rectangle +(36mm,36mm);
-	%\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}
-	%\includegraphics[height=65mm]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}
+	\draw[ultra thick] (a.south west) ++(0,-0.25em) -- ++(30.588mm,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\centi\metre}};
+	\draw[ultra thick] (b.south west) ++(0,-0.25em) -- ++(45.774mm,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
+	\draw[ultra thick] (c.south west) ++(0,-0.25em) -- ++(30.415mm,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
 	\end{tikzpicture}
@@
 	\Head{Pattern kinetics by transmission macroscope}
-	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccccccc}
-	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
+	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}p{2\TPHorizModule}llllll}
+	33 min\hfill \begin{tikzpicture}[inner sep=0]
+		\draw[ultra thick] (0,0) -- ++(0.355\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{1}{\centi\metre}};
+	\end{tikzpicture}&
+	 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
 	\includegraphics[width=2\TPHorizModule]{prise_0100_resized.jpg}&
@@
 
-	\includegraphics[width=\textwidth]{collage_cut}
-
+	\begin{tikzpicture}[inner sep=0]
+		\node[above right]{\includegraphics[width=\textwidth]{collage_cut}};
+		\draw[ultra thick] ++(0,0.5em) -- ++(0.33\textwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
+	\end{tikzpicture}\\
 	Tilled, enhanced relief
@@
 	\Head{Wrinkling kinetics by confocal microscope}
-	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccccccc}
-	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
+	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllll}
+	\raisebox{-0.75em}{\begin{tikzpicture}[inner sep=0]
+		\node[above right]{\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}};
+		\draw[ultra thick] ++(0,0.02\TPHorizModule) -- ++(1.55\TPHorizModule,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
+	\end{tikzpicture}}&
 	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
@@
 \begin{textblock}{2}(2,16)
-	\addvspace{4.4em}
-	\hfill bottom surface, \textsc{gdl} 5\%\\
+	\addvspace{2.4em}
+	bottom surface, \textsc{gdl} 5\%\\
+	\begin{tikzpicture}[inner sep=0]
+		\draw[ultra thick] (0,0) -- ++(1.11\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{500}{\micro\metre}};
+	\end{tikzpicture}
+
 	\includegraphics[width=2\TPHorizModule]{plastic_confocal}
@@
 		\draw[Accent2,radius=0.09\TPHorizModule] ++(b.north east) +(-0.435\TPHorizModule,-0.375\TPHorizModule) circle;
+		\draw[ultra thick] (a.south west) ++(0,0.25em) -- ++(0.5\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{2}{\centi\metre}};
 	\end{tikzpicture}

### 2013-09-10 poster_Rome.tex
- title "gels" - mail @ens-lyon - remove "non equilibrium creep" box and "fractures need complete overshoot" in "take away" - add ERC funding - "take home" rather than "take away" - "A good model for solids" -> "An original yielding scenario" - Capitals in "take home" - "tilled" -> "tiled" ? - show lambda on a pattern figure and a fracture figure - fix scale bar of thickness dependence of patterns

@@
 \hypersetup{%
-	pdftitle={Effects of a real world size distribution on the heterogeneous crystallisation of hard sphere colloids},%the title
+	pdftitle={Wrinkling and yielding of protein gels},%the title
 	pdfauthor={Mathieu Leocmach},%your name
@@
 \Title{
-Wrinkling and yielding of protein gel
+Wrinkling and yielding of protein gels
 }
 
-\LHead{Mathieu Leocmach}\qquad\texttt{\color{Accent1}mathieu.leocmach@polytechnique.org}\\
+\LHead{Mathieu Leocmach}\qquad\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}\\
 \LHead{Mathieu Nespoulous, Christophe Perge, Nicolas Taberlet, Thomas Gibaud,
@@
 	\end{tikzpicture}\\
-	Tilled, enhanced relief
+	Tiled, enhanced relief
 \end{textblock}
@@
 	\node[above=0 of a] {\SI{800}{\micro\metre}};
-	\draw[ultra thick](a.north west) ++(0,2.25em) -- ++(0.0797\textwidth,0) node[midway, above, inner sep=0] {\SI{1}{\centi\metre}};
+	\draw[ultra thick](a.north west) ++(0,2em) -- ++(0.2\textwidth,0) node[midway, above, inner sep=0] {\SI{5}{\milli\metre}};
+	\draw[ultra thick, Accent2, <->] (a.south west) ++(0.18\textwidth, 0.05\textwidth) -- ++(0.133\textwidth,0) node[midway,above]{$\lambda$};
 	\end{tikzpicture}}\\
@@
 		\draw[ultra thick] (a.south west) ++(0,0.25em) -- ++(0.5\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{2}{\centi\metre}};
+		\draw[ultra thick, Accent2, <->] (a.south west) ++(0.51\TPHorizModule, \TPHorizModule) -- ++(0.09\TPHorizModule,0) node[midway,above=0.25em,font=\small]{$\lambda$};
 	\end{tikzpicture}
@@
 \begin{textblock}{5}(10,22.5)
-	\Head{Non equilibrium Creep}
-	\Subhead{Fractures need complete overshoot}
-	\begin{itemize}
-		\item Wait only \SI{40}{\minute}. $G^\prime$ at maximum.
-		\item Start creep with $\sigma > \SI{200}{\pascal}$.
-		\item[$\Rightarrow$] Break before $G^\prime$ falls.
-		\item[$\Rightarrow$] No fracture
-	\end{itemize}
+	\Head{Funding}
+	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lb{0.6\textwidth}}
+	\includegraphics[width=0.3\textwidth]{NEW-Logo-ERC-OUTLINE}&
+	The research leading to these results has received funding from the European Research Council under the European Union's Seventh \linebreak Framework Programme (FP7/2007-2013) /\linebreak ERC grant agreement n°~258803.\\
+	\end{tabular*}
 \end{textblock}
@@
 \begin{textblock}{2.875}(12,2.625)%real width of 3
-	\Norulehead{Take away}
+	\Norulehead{Take home}
 	\Subhead{A new kind of pattern}
 	\begin{itemize}
-		\item nested
-		\item self generated
-		\item boundary conditions
+		\item Nested
+		\item Self generated
+		\item Boundary conditions
 	\end{itemize}
-	\Subhead{A good model for solids}
+	\Subhead{An original yielding scenario}
 	\begin{itemize}
@@
 		\item Detachment creates a free surface
-		\item Fractures need complete overshoot
 	\end{itemize}

### 2013-09-10 poster_Rome.tex
global mark size "Casein gel" -> "Acid-induced gels", units, width arrows in creep and fractures gray background to increase readability in speckle's label

@@
 \usepgfplotslibrary{groupplots}
-\pgfplotsset{every axis/.append style={xlabel near ticks,ylabel near ticks}}
+\pgfplotsset{every axis/.append style={xlabel near ticks,ylabel near ticks,mark size={0.2em}}}
 \pgfplotsset{every axis plot post/.append style={very thick}}
@@
 \begin{textblock}{3}(0,2.5)
-	\Head{Casein gel}
+	\Head{Acid-induced gels}
 
@@
 	\begin{axis}[%
-		xmin=0,
+		xmin=0,xmax=8,
 		ymin=0, ymax=7,
@@
 	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
+	\node[base left=0] at (axis cs:8,4.6) {isoelectric};
 	\end{axis}
 	\begin{axis}[%
-		xmin=0,
+		xmin=0,xmax=8,
 		ymin=0,
-		ylabel=\textcolor{Accent2}{$G^\prime$},
+		ylabel=\textcolor{Accent2}{$G^\prime$ (\si{\pascal})},
 		axis y line*=right,
@@
 		ylabel near ticks,
-		mark size={0.01\textwidth},
 		%legend style={legend pos=north west,}
@@
 		\addplot+[no marks, Accent2] file {Y34_strain_rate_log.txt};
-		\draw[<->] (axis cs:1e-1,1) -- (axis cs:1.3e3,5e-4) node[midway,above,rotate=-21] {Andrade};
+		\draw[<->] (axis cs:1e-1,1) -- (axis cs:1.3e3,5e-4) node[midway,above,rotate=-21] {Andrade} node[pos=0.25,below=1em] {$t^{-0.86}$};
 		\node at (axis cs:1.3e3,1e-4) (det){};
@@
 	\end{loglogaxis}
-	\draw[<-, Accent1, ultra thick] (det) -- ++(0,-4em) node[left] {Fractures appear};
-	\draw[<-, Accent1, ultra thick] (touch) -- ++(0,3em) node[left] {Fractures meet};
+	\draw[<-, Accent1, ultra thick] (det) +(0,0.1em) -- ++(0,-4em) node[left] {Fractures appear} -- ++(1em,0);
+	\draw[<-, Accent1, ultra thick] (touch) +(0,-0.1em) -- ++(0,3em) node[left] {Fractures meet} -- ++(1em,0);
 	\end{tikzpicture}\hfill
@@
 		legend style={font=\small\sc}, legend pos=south west,
-		mark size={0.2em},
 		]
@@
 		ylabel near ticks,
-		%mark size={0.01\textwidth},
 		legend cell align=left,
@@
 			]{Y137_speckleBF_tr.png};
-			\draw[Accent2] (axis cs:172,0)--(axis cs:172,2) node[pos=0.66, left, font=\small] {free from the rotor};
-			\draw[Accent1] (axis cs:187,0)--(axis cs:187,2) node[pos=0.33, left, font=\small] {global yield};
+			\draw[Accent2] (axis cs:172,0)--(axis cs:172,2) node[pos=0.66, left, font=\small, fill=gray, inner sep=1pt] {free from the rotor};
+			\draw[Accent1] (axis cs:187,0)--(axis cs:187,2) node[pos=0.33, left, font=\small, fill=gray, inner sep=1pt] {global yield};
 	\end{groupplot}

### 2013-09-10 poster_Rome.tex
forgotten file
logos full ref of Gallot et al.

@@
 
-\LHead{\textsc{Laboratoire de Physique -- UMR 5672, Ecole Normale Supérieure de Lyon}}
+\LHead{\raisebox{-0.3em}{\includegraphics[height=\baselineskip]{CNRSfilaire-Q}\qquad\includegraphics[height=\baselineskip]{logo_UCBL}\qquad}\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}\raisebox{-0.3em}{\qquad\includegraphics[height=\baselineskip]{logo_ens-lyon}}}
 
@@
 \begin{textblock}{3}(0,10)
-	Corresponds with the fall of $G^\prime$.
-
 	\Subhead{Length scales}
@@
 	\begin{itemize}
-	\item phase separation patterns 
+	\item Phase separation patterns 
 	\item Turing patterns
-	\item usual wrinkles
+	\item Usual wrinkles
 	\end{itemize}	 
@@
 	\end{tikzpicture}\\
-	Tiled, enhanced relief
+	Tiled, enhanced elevation
 \end{textblock}
@@
 	\draw[->] (rh) -- +(0,1em) -| (pc.north);
-	\node[below right=1em and 0 of mce.south west, text width=0.18\textwidth] {Gallot et al. RSI (2013).};
+	\node[above right=0 and 0.1\textwidth of sch.south east, text width=0.25\textwidth, align=right] {Gallot et al. \linebreak\textit{Rev. Sci. Instrum.} 84, 045107 (2013).};
 	\node[below right =of sch.north east, text width=0.32\textwidth] {3\% of \SI{30}{\micro\metre} acrylamide beads as contrast agent};

### 2013-10-24 plis_figures.tex
Figure generator for the wrinkling paper. fig:acidgel

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage[utf8]{inputenc}

\usepackage{siunitx}

\usepackage{graphicx}

\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, positioning, intersections, decorations.markings, decorations.shapes, mindmap, shapes.geometric, matrix, patterns}

\tikzset{lab/.style={above right, inner sep=0, font=\sffamily\Large\bfseries}}

\usepackage{pgfplots}

\pgfplotsset{every axis/.append style={xlabel near ticks,ylabel near ticks,mark size={0.2em}}}

\definecolor{Main}{rgb}{1, 0.57, 0}
\definecolor{Accent1}{rgb}{1,0.28,0}
\definecolor{Accent2}{rgb}{1,0.74,0}

\author{Mathieu Leocmach}
\begin{document}

\begin{figure}
	\begin{tikzpicture}
	\begin{axis}[%
		name=pH,
		xmin=0,xmax=8,
		ymin=0, ymax=7,
		xlabel={time (h)}, ylabel={\textcolor{Accent1}{pH}},
		extra tick style={grid=major},%
		extra y ticks={4.6}, extra y tick labels={},%
		extra x ticks={0.55, 0.64, 0.72, 0.80, 1, 1.27, 2.5}, extra x tick labels={},%
		axis y line*=left,
		width=0.925\columnwidth,
		height=0.5\columnwidth,
		]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
	\node[base left=0] at (axis cs:8,4.6) {isoelectric};
	\end{axis}
	\begin{axis}[%
		xmin=0,xmax=8,
		ymin=0,
		ylabel=\textcolor{Accent2}{$G^\prime$ (\si{\pascal})},
		axis y line*=right,
		axis x line=none,
		width=0.925\columnwidth,
		height=0.5\columnwidth,
		]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
	\end{axis}
	\node[lab, yshift=0.5em] at (pH.outer south west) {a};
	\begin{scope}[shift=(pH.outer south west), yshift=-2em]
		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\columnwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\columnwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
		\draw[line width=2pt,Accent1] (0.05\columnwidth,-2.5em) -- (0.95\columnwidth,-2.5em) (0.05\columnwidth,-1pt) -- (0.95\columnwidth,-1pt) node[pos=0, below right, inner sep=1pt, text width=0.4\columnwidth] {acrylamide brush ($10\sim 100$ nm)};
		\fill[gray] (0,0) rectangle (0.05\columnwidth,-2.5em) (\columnwidth,0) rectangle (0.95\columnwidth,-2.5em) node[pos=1, above left] {spacer};
		\draw[<->] (0.75\columnwidth,-2pt) -- (0.75\columnwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
		\draw[<->] (0.05\columnwidth,-4.25em) -- (0.95\columnwidth,-4.25em) node[midway,below] (L) {$L\sim 2.5\,cm$};
		\node[lab] at (0,0 |- L.base) {b};
	\end{scope}
	\begin{scope}[inner sep=0, shift=(pH.outer south west), yshift=-10.5em]
		\node[anchor=west] (a) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2}};
		\node at (0.5\columnwidth,0) (b) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}};
		\node[anchor=east] at (\columnwidth,0) (c) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
		\draw[Accent2] (a.north west) ++(0.126\columnwidth, -0.086\columnwidth) rectangle +(0.1\columnwidth,-0.075\columnwidth) -- (b.south west) (a.north west) ++(0.226\columnwidth, -0.086\columnwidth) -- (b.north west);
		\draw[Main] (b.north west) ++(0.117\columnwidth,-0.117\columnwidth) rectangle +(0.093\columnwidth,-0.068\columnwidth) -- (c.south west) (b.north west) ++(0.211\columnwidth,-0.117\columnwidth) --(c.north west);
		\draw[Accent1] (c.north west) ++(0.094\columnwidth,0) rectangle +(0.117\columnwidth,-0.117\columnwidth);
		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.106\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.158\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.099\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
	\node[lab] at (0,0 |- M.base) {c};
	\node[lab] at (b.west |- M.base) {d};
	\node[lab] at (c.west |- M.base) {e};
	\end{scope}
	\end{tikzpicture}
	\caption{Acid induced gel. (a) Evolution of the pH and of the storage modulus with sticky boundary conditions. (b) Sketch of the cell with no adhesion of the protein on the top and bottom plates. (c-e) Resulting patterns in the same sample ($e=\SI{100}{\micro\metre}$) at different magnification by (c-d) reflexion or (e) transmission optical microscopy.}
	\label{fig:acidgel}
\end{figure}

\end{document}

### 2013-10-25 plis_figures.tex
add fig:dynamics Fix scale bars

@@
 		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\columnwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
-		\draw[line width=2pt,Accent1] (0.05\columnwidth,-2.5em) -- (0.95\columnwidth,-2.5em) (0.05\columnwidth,-1pt) -- (0.95\columnwidth,-1pt) node[pos=0, below right, inner sep=1pt, text width=0.4\columnwidth] {acrylamide brush ($10\sim 100$ nm)};
+		\draw[line width=2pt,Accent1] (0.05\columnwidth,-2.5em) -- (0.95\columnwidth,-2.5em) (0.05\columnwidth,-1pt) -- (0.95\columnwidth,-1pt) node[pos=0, below right, inner sep=1pt, text width=0.4\columnwidth] {acrylamide brush ($10\sim \SI{100}{\nano\metre}$)};
 		\fill[gray] (0,0) rectangle (0.05\columnwidth,-2.5em) (\columnwidth,0) rectangle (0.95\columnwidth,-2.5em) node[pos=1, above left] {spacer};
-		\draw[<->] (0.75\columnwidth,-2pt) -- (0.75\columnwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
-		\draw[<->] (0.05\columnwidth,-4.25em) -- (0.95\columnwidth,-4.25em) node[midway,below] (L) {$L\sim 2.5\,cm$};
+		\draw[<->] (0.75\columnwidth,-2pt) -- (0.75\columnwidth,-2.5em) node[midway,left] {$e\sim \SI{100}{\micro\metre}$};
+		\draw[<->] (0.05\columnwidth,-4.25em) -- (0.95\columnwidth,-4.25em) node[midway,below] (L) {$L\sim\SI{2}{\centi\metre}$};
 		\node[lab] at (0,0 |- L.base) {b};
@@
 
+\begin{figure*}
+	\begin{tikzpicture}
+	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em] (m){
+	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
+	\includegraphics[width=0.13\textwidth]{prise_0100_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0130_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0160_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0190_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0250_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0360_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0799_resized.jpg}\\
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
+	};
+	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.023\textwidth,0);
+	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
+%	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}p{2\TPHorizModule}llllll}
+%	33 min\hfill \begin{tikzpicture}[inner sep=0]
+%		\draw[ultra thick] (0,0) -- ++(0.355\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{1}{\centi\metre}};
+%	\end{tikzpicture}&
+%	 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
+
+	\end{tikzpicture}
+	\caption{Dynamics of pattern formation for $e\approx\SI{100}{\micro\metre}$.
+(top) By light transmission microscopy.
+(bottom) Reconstructed from fluorescent confocal microscopy.
+Scale bars are \SI{1}{\milli\metre}.}
+	\label{fig:dynamics}
+\end{figure*}
+
 \end{document}

### 2013-10-29 plis_figures.tex
Cleanup
fig:thickness

@@
 
+%PNAS fonts
+\usepackage{mathptmx}
+\usepackage{tgheros}
+
 \usepackage{siunitx}
+\usepackage{multirow}
 
@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,shapes,backgrounds, positioning, intersections, decorations.markings, decorations.shapes, mindmap, shapes.geometric, matrix, patterns}
+\usetikzlibrary{calc, positioning, intersections, matrix, patterns}
 
@@
 
-\pgfplotsset{every axis/.append style={xlabel near ticks,ylabel near ticks,mark size={0.2em}}}
+\pgfplotsset{compat=1.3, every axis/.append style={mark size={0.2em}}}
 
@@
 	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
-%	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}p{2\TPHorizModule}llllll}
-%	33 min\hfill \begin{tikzpicture}[inner sep=0]
-%		\draw[ultra thick] (0,0) -- ++(0.355\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{1}{\centi\metre}};
-%	\end{tikzpicture}&
-%	 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
-
 	\end{tikzpicture}
@@
 
+\begin{figure}
+	\begin{tikzpicture}
+	\begin{axis}[%
+		name=g,
+		width=\columnwidth,
+		height=0.4\columnwidth,
+		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
+		xtick={0.1,0.3,...,0.9},
+		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		]
+		\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+		\addplot+[domain=0:3, no marks, Accent1, forget plot] {4.23*x};
+	\end{axis}
+
+	\matrix[matrix of nodes, inner sep=0, matrix anchor=south west, column sep=0.033\columnwidth, row sep=0.5em] at ($(g.outer north west)+(0,0.5em)$) (m) {
+		\SI{50}{\micro\metre} & \SI{100}{\micro\metre}\\
+		\includegraphics[width=0.225\columnwidth]{pattern_50um.jpg} &
+		\includegraphics[width=0.225\columnwidth]{pattern_100um.jpg}\\
+		\SI{250}{\micro\metre} & \SI{450}{\micro\metre}\\
+		\includegraphics[width=0.225\columnwidth]{pattern_250um.jpg} &
+		\includegraphics[width=0.225\columnwidth]{pattern_450um.jpg}\\
+	};
+	\node[inner sep=0, below left] at ($(m-2-1.north west) + (\columnwidth,0)$) (x800) {\includegraphics[width=0.45\columnwidth]{pattern_800um_x1.jpg}};
+	\node[anchor=base] at (x800.north |- m-1-2.base) {\SI{800}{\micro\metre}};
+	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.18\columnwidth, 0.05\columnwidth) -- ++(0.133\columnwidth,0) node[midway,above]{$\lambda$};
+	\draw[ultra thick](x800.south east) ++(0,-0.5em) -- ++(-0.2\columnwidth,0) node[midway, below] {\SI{5}{\milli\metre}};
+	\end{tikzpicture}
+	\caption{Cell thickness dependence of the patterns. Scale bar is common for all 5 transmission microscopy pictures.}
+	\label{fig:thickness}
+\end{figure}
+
 \end{document}

### 2013-11-12 2013_11_12_discussion.tex
Présentation pour discussion du 12 Nov 2013

\documentclass{beamer}
\usetheme{Boadilla}
\usecolortheme{crane}
\usefonttheme[stillsansseriflarge, stillsansserifsmall]{serif}
%\usefonttheme[stillsansseriflarge, stillsansserifsmall]{structurebold}
%\usefonttheme[onlymath]{serif}
\useinnertheme{default}
\setbeamertemplate{navigation symbols}{}

\usepackage{fontspec}
%\usepackage{kpfonts}
\setmainfont[Numbers=OldStyle]{Tex Gyre Pagella}
\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}

\usepackage[frenchb]{babel}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{siunitx}
\sisetup{
  mathsmu = \text{µ},
  textmu  = µ
}
\usepackage{booktabs}
\usepackage{multirow}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}

\usepackage{tikz}
\usetikzlibrary{calc, positioning, intersections, matrix, patterns}
\usepackage{pgfplots}
\usepgfplotslibrary{groupplots}

\pgfplotsset{compat=1.3, every axis/.append style={mark size={0.2em}}}

\usepgfplotslibrary{external}
\tikzset{external/system call={lualatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"}}

\definecolor{Main}{rgb}{1, 0.57, 0}
\definecolor{Accent1}{rgb}{1,0.28,0}
\definecolor{Accent2}{rgb}{1,0.74,0}

%\tikzexternalize

\usepackage{multimedia}

%bibliography
\usepackage{natbib}
%\usepackage{bibentry}
\def\newblock{\hskip .11em plus .33em minus .07em}

\institute{ENS Lyon}
\title{Yaourt pliss\'{e}, yaourt flamb\'{e}}

\author[M. Leocmach]{Mathieu Leocmach}
\date{13 Novembre 2013}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={small}}
\tikzset{external/force remake=false}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}[plain]
	\titlepage
\end{frame}

\begin{frame}{Motifs}
	\begin{tikzpicture}[inner sep=0, very thick]
	\node[anchor=west] (a) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2}};
		\node at (0.5\columnwidth,0) (b) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}};
		\node[anchor=east] at (\columnwidth,0) (c) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
		\draw[Accent2] (a.north west) ++(0.126\columnwidth, -0.086\columnwidth) rectangle +(0.1\columnwidth,-0.075\columnwidth) -- (b.south west) (a.north west) ++(0.226\columnwidth, -0.086\columnwidth) -- (b.north west) rectangle (b.south east);
		\draw[Main] (b.north west) ++(0.117\columnwidth,-0.117\columnwidth) rectangle +(0.093\columnwidth,-0.068\columnwidth) -- (c.south west) (b.north west) ++(0.211\columnwidth,-0.117\columnwidth) --(c.north west) rectangle (c.south east);
		\draw[Accent1] (c.north west) ++(0.094\columnwidth,0) rectangle +(0.117\columnwidth,-0.117\columnwidth);
		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.106\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.158\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.099\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
		\node[above=0 of a] {Reflexion Macroscope};
		\node[above=0 of b] {Reflexion Macroscope};
		\node[above=0 of c, text width=0.3\columnwidth] {Transmission\linebreak Microscope (tiled)};
	\end{tikzpicture}
\end{frame}

\begin{frame}{Acid-induced gels}

	In millipore water
	\begin{itemize}
	\item 4\% Sodium caseinate (milk protein)\hfill
	Isoelectric point $\approx 4.6$
	\item 4\% $\delta$-gluconolactone (\textsc{gdl})\hfill
	Slow hydrolysis into gluconic acid
	\end{itemize}
	\vfill
	\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group size=1 by 2, 
			xlabels at=edge bottom, 
			xticklabels at=edge bottom,
			vertical sep=1em,
			},
		xmin=0, xmax=8,
		xlabel={time (h)}, 
		extra tick style={grid=major},%
		extra y ticks={4.6}, extra y tick labels={},%
		extra x ticks={0.55, 0.64, 0.72, 0.80, 1, 1.27, 2.5}, extra x tick labels={},%
		width=\columnwidth,
		height=0.25\columnwidth,
		]
	\nextgroupplot[ylabel=pH, ymin=0, ymax=7]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
	\node[base left] at (axis cs:8,4.6) {isoelectric};
	\nextgroupplot[ylabel=$G^\prime$ (\si{\pascal}), ymin=0]
	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
	\end{groupplot}
	\end{tikzpicture}
\end{frame}

\begin{frame}{Cellule de microscopie}
\begin{center}
\begin{tikzpicture}
\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\columnwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\columnwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
		\draw[line width=2pt,Accent1] (0.05\columnwidth,-2.5em) -- (0.95\columnwidth,-2.5em) (0.05\columnwidth,-1pt) -- (0.95\columnwidth,-1pt) node[pos=0, below right, inner sep=1pt, text width=0.5\columnwidth] {acrylamide brush ($10\sim \SI{100}{\nano\metre}$)};
		\fill[gray] (0,0) rectangle (0.05\columnwidth,-2.5em) (\columnwidth,0) rectangle (0.95\columnwidth,-2.5em) node[pos=1, above left] {spacer};
		\draw[<->] (0.75\columnwidth,-2pt) -- (0.75\columnwidth,-2.5em) node[midway,left] {$e\sim \SI{100}{\micro\metre}$};
		\draw[<->] (0.05\columnwidth,-4.25em) -- (0.95\columnwidth,-4.25em) node[midway,below] (L) {$L\sim\SI{2}{\centi\metre}$};
\end{tikzpicture}
\end{center}

\begin{itemize}
\item Pas d'accroche sur les plus grandes surfaces grâce aux brosses.
\item Accroche sur les extrémités (spacers solides, pas d'interface libre)
\item Pas de contrainte extérieure
\end{itemize}
\end{frame}

\begin{frame}{Formation}
\begin{columns}
\column{0.5\textwidth}
\movie[externalviewer]{\includegraphics[width=\textwidth]{cas4_GDL4_motif_final.jpg}}{cas4_GDL_4_100um_coat_1_1_comp_4.avi}\\
\begin{tikzpicture}
\draw[ultra thick](0,0)-- (0.757\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
\end{tikzpicture}
\column{0.5\textwidth}
\movie[externalviewer]{\includegraphics[width=\textwidth]{pattern_100um.jpg}}{cas4_GDL_4_100um_coat_macroscope_2.avi}\\
\begin{tikzpicture}
\draw[ultra thick](0,0)-- (0.177\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
\end{tikzpicture}
\end{columns}
\end{frame}

\begin{frame}{Formation}
	\begin{tikzpicture}
	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em, ampersand replacement=\&] (m){
	33 min \& 38 min \& 43 min \\
	\includegraphics[width=0.32\textwidth]{prise_0100_resized.jpg}\&
	\includegraphics[width=0.32\textwidth]{prise_0130_resized.jpg}\&
	\includegraphics[width=0.32\textwidth]{prise_0160_resized.jpg}\\
	48 min \& 1h \& 1h15\\
	\includegraphics[width=0.32\textwidth]{prise_0190_resized.jpg}\&
	\includegraphics[width=0.32\textwidth]{prise_0250_resized.jpg}\&
	\includegraphics[width=0.32\textwidth]{prise_0360_resized.jpg}\\
	};
	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.059\textwidth,0);
	\end{tikzpicture}
\end{frame}

\begin{frame}{Microscopie confocale}
%\begin{columns}
%\column{0.5\textwidth}
	\begin{block}{Procedure}
	\begin{itemize}
		\item Fluorescently label 10\% of the casein
		\item A confocal slice every \SI{5}{\micro\metre}
		\item $(\SI{1.3}{\milli\metre})^2 \times \SI{120}{\micro\metre}$
		\hfill $\Rightarrow$ Protein density field
		\item Detect top and bottom surfaces of the gel phase
	\end{itemize}
	\end{block}
%\column{0.5\textwidth}
\begin{center}
	\begin{tikzpicture}[inner sep=0]
		\node[above right]{\includegraphics[width=0.6\textwidth]{collage_cut}};
		\draw[ultra thick] ++(0,0.5em) -- ++(0.2\textwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
	\end{tikzpicture}
\end{center}
%\end{columns}
\end{frame}

\begin{frame}{Dynamique en confocal}
\begin{columns}
\column{0.5\textwidth}
\movie[externalviewer]{\includegraphics[width=\textwidth]{plots_final.jpg}}{plots.avi}
\column{0.5\textwidth}
\movie[externalviewer]{\includegraphics[width=\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}}{paraview_cas3p3_fluo0p8_GDL4_2_flat2_comp.avi}
\end{columns}
\begin{tikzpicture}
	\begin{groupplot}[%
		group style={group name=g, group size=1 by 2},
		xmin=0, xtick={20,40,...,160},
		xlabel={time (min)}, 
		extra tick style={grid=major},%
		extra x ticks={33, 38, 43, 48, 58, 77, 150}, extra x tick labels={},%
		width=\columnwidth,
		height=0.25\columnwidth,
		]
	\nextgroupplot[ylabel={$V_\text{gel}/V_\text{cell}$}, ymin=0.5, ymax=1]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60.+3}]{cas4_GDL4_fluo0p4_2.volume};

	\nextgroupplot[ylabel={$A/A_\text{cell}-1$ (\%)}, ymin=0, ymax=2]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60.+3}]{cas4_GDL4_fluo0p4_2.area};
	\end{groupplot}
\end{tikzpicture}
\end{frame}

\begin{frame}{Dynamique}
\begin{tikzpicture}
	\matrix[
		matrix of nodes, inner sep=0, 
		column sep=0.015\textwidth, row sep=0.5em,
		ampersand replacement=\&] (m){
	33 min \& 38 min \& 43 min \& 48 min \& 1h \& 1h15 \& 2h30\\
	\includegraphics[width=0.13\textwidth]{prise_0100_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{prise_0130_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{prise_0160_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{prise_0190_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{prise_0250_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{prise_0360_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{prise_0799_resized.jpg}\\
	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}\&
	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
	};
	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.023\textwidth,0);
	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
\end{tikzpicture}
\begin{columns}[T]
\column{0.29\textwidth}
\hrulefill\\
Motif primaire
\begin{itemize}
\item Parfois labyrinthe, parfois plots
\item Du sol au plafond
\end{itemize}
Rides sans substrat ?
\column{0.7\textwidth}
\hrulefill\\
Flambage $\Rightarrow$ harmoniques

\includegraphics[width=\columnwidth]{Roman_nouveau_plis.jpg}

Roman \& Pocheau, EPL (1999).
\end{columns}
\end{frame}

\begin{frame}{Longueurs d'onde en fonction du gap}
	\begin{tikzpicture}
	\matrix[
		matrix of nodes, inner sep=0, matrix anchor=south east, 
		column sep=0.033\columnwidth, row sep=0.5em,
		ampersand replacement=\&
		] (m) {
		\SI{50}{\micro\metre} \& \SI{100}{\micro\metre}\\
		\includegraphics[width=0.225\columnwidth]{pattern_50um.jpg} \&
		\includegraphics[width=0.225\columnwidth]{pattern_100um.jpg}\\
		\SI{250}{\micro\metre} \& \SI{450}{\micro\metre}\\
		\includegraphics[width=0.225\columnwidth]{pattern_250um.jpg} \&
		\includegraphics[width=0.225\columnwidth]{pattern_450um.jpg}\\
	};
	\node[inner sep=0, below=1.5em of m] (x800) {\includegraphics[width=0.45\columnwidth]{pattern_800um_x1.jpg}};
	\node[above = 0 of x800.north] {\SI{800}{\micro\metre}};
	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.18\columnwidth, 0.05\columnwidth) -- ++(0.133\columnwidth,0) node[midway,above]{$\lambda$};
	\draw[ultra thick](x800.north east) ++(0,-0.5em) -- ++(-0.2\columnwidth,0) node[midway, below] {\SI{5}{\milli\metre}};
	\begin{axis}[%
		name=g,
		at={($(m-2-1.west)+(\columnwidth,0)$)},
		anchor=outer north east,
		width=0.5\columnwidth,
		height=0.5\columnwidth,
		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
		xtick={0.1,0.3,...,0.9},
		ymin=0, ymax=4.23, ylabel={$\lambda$ (\si{\milli\metre})},
		]
		\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
		\addplot+[domain=0:3, no marks, Accent1, forget plot] {4.23*x};
	\end{axis}
	\node[anchor=base west] at (x800.south east) {Longueur d'onde du motif primaire};
	\node [above = 0 of g] {$\lambda = 4.23 e$};
	\end{tikzpicture}
\end{frame}

\begin{frame}{Motif primaire: modèle}
Comment plisser (et non flamber) sans substrat ?
\begin{itemize}
\item Notre substrat, c'est l'eau
\hfill {\footnotesize Boudaoud \& Chaïeb, PRE (2001)}
\begin{itemize}
\item $E_s = \frac{3\eta}{\tau}$\qquad avec $\tau$ le temps caractéristique de l'écoulement
\item Epaisseur du substrat $H = \frac{e-h}{2} \approx e/4$. Substrat fin $H<\lambda$.
\end{itemize}
\item Pour un substrat d'un seul côté \hfill{\footnotesize Cerda \& Mahadevan, PRL (2003)} 
\end{itemize}
\begin{columns}
\column{0.6\textwidth}
\[
\lambda = \underbrace{\sqrt{hH}}_{\lessapprox 0.35e} \left(\underbrace{\frac{E_f}{E_s}}_{\approx 87\cdot 10^6}\right)^\frac{1}{6}
\approx 7.4e
\]
\column{0.4\textwidth}
\begin{align*}
G^\prime_f &\approx \SI{500}{\pascal}\Rightarrow
E_f \approx \SI{1300}{\pascal}\\
\tau &\approx \SI{200}{\second}\\
\eta &\approx \SI{1}{\milli\pascal\second}
\end{align*}
\end{columns}
\vfill
\begin{itemize}
\item Amplitude maximum contrainte $A=e-h\approx h\Rightarrow U_B \approx U_S$ \item Parfois rides parfois plots 
\hfill {\footnotesize Vandeparre \textit{et al.}, Soft Matter (2011)}
\end{itemize}
\end{frame}

\begin{frame}{Longueurs d'onde et composition}
\begin{columns}[T]\small

\column{0.45\textwidth}
\begin{block}{Moins de GDL}
\begin{itemize}
		\item Cinétique lente
		\item Gel plus fort
		\item Moins d'overshoot
	\end{itemize}
\end{block}
\includegraphics[width=0.75\textwidth]{cas4_GDL1.jpg}1\%
\begin{itemize}
\item Motif primaire uniquement 
\item Même longueur d'onde
\item Difficile d'obtenir des motifs,
\item Motif transitoires ?
\end{itemize}

\column{0.45\textwidth}
\begin{block}{Plus de GDL}
\begin{tikzpicture}[inner sep=0]
		\node (a) {\includegraphics[width=0.75\textwidth]{plastic_confocal}};
		\draw[ultra thick] (a.south west) -- ++(0.41\textwidth,0) node[pos=0.5, above=0, font=\small] {\SI{500}{\micro\metre}};
\end{tikzpicture}
5\%
\end{block}
\begin{itemize}
\item Encore plus d'harmonique
\item Même longueur d'onde primaire
\item Plasticité
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}{Diagramme de phase}
\begin{tikzpicture}
\begin{axis}[%
	name=phasediag,
	width=0.9\textwidth,
	height=0.6\textwidth,
	xmin=0,xmax=4.5,xlabel={\% caseinate}, xtick={1,2,3,4},
	ymin=0,ymax=5,ylabel={\% GDL},
	legend style={legend pos=north west,},
	xlabel near ticks,
	ylabel near ticks,
	]
\node at (axis cs:1,1) {\includegraphics[width=0.15\textwidth]{cas1_GDL1.jpg}};
\node at (axis cs:2,1) {\includegraphics[width=0.15\textwidth]{cas2_GDL1.jpg}};
\node at (axis cs:2,2) {\includegraphics[width=0.15\textwidth]{cas2_GDL2.jpg}};
\node at (axis cs:3,1) {\includegraphics[width=0.15\textwidth]{cas3_GDL1.jpg}};
\node at (axis cs:3,2) {\includegraphics[width=0.15\textwidth]{cas3_GDL2.jpg}};
\node at (axis cs:3,3) {\includegraphics[width=0.15\textwidth]{cas3_GDL3.jpg}};
\node at (axis cs:4,1) {\includegraphics[width=0.15\textwidth]{cas4_GDL1.jpg}};
\node at (axis cs:4,2) {\includegraphics[width=0.15\textwidth]{cas4_GDL2.jpg}};
\node at (axis cs:4,3) {\includegraphics[width=0.15\textwidth]{cas4_GDL3.jpg}};
\node at (axis cs:4,4) {\includegraphics[width=0.15\textwidth]{pattern_100um.jpg}};
\addplot[no marks, Accent1] coordinates {(0,0) (5,5)} node[pos=0.05,above] {\SI{1}{\hour}};
\addplot[no marks, Main] coordinates {(0,0) (10,5)} node[pos=0.075] {\SI{2}{\hour}};
\addplot[no marks, Accent2] coordinates {(0,0) (8,2)} node[pos=0.2,below] {\SI{8}{\hour}};

\node at (axis cs:1,4) {$\lambda = \sqrt{hH} \left(\frac{E_f}{E_s}\right)^\frac{1}{6}$};

\end{axis}
\draw[->] ($(phasediag.north east)+(1em,0)$) -- ($(phasediag.south east)+(1em,0)$) node[midway,right,text width=0.1\textwidth] {\begin{align*}
\frac{G^{\prime}_\infty}{G^{\prime}_\text{max}}&\nearrow\\
\delta&\searrow
\end{align*}};
\draw[->] ($(phasediag.north west)+(0,1em)$) -- ($(phasediag.north east)+(0,1em)$) node[midway,above] {$G^{\prime}_\text{max}, E_f, h\nearrow$};
\end{tikzpicture}
\end{frame}

\end{document}

### 2014-01-17 creep/creep_yaourt.tex
Creep paper. Figures and captions

\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4}
\usepackage{graphicx}
\usepackage{dcolumn,xcolor,ulem}
\usepackage{amsmath} 
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage[latin1]{inputenc}
%\usepackage{mathrsfs}

\newcommand{\cor}[1]{\textcolor{red}{#1}}
\newcommand{\change}[1]{\textcolor{blue}{#1}}
\newcommand\gp{\dot\gamma}
\newcommand\gl{\gamma_{\rm loc}}
\newcommand\gpm{\dot\gamma_{\rm min}}
\newcommand\taum{\tau_{\rm min}}

\begin{document}

%\widowpenalty=1000
%\clubpenalty=1000
%\preprint{APS/123-QED}

\title{Creep and failure of a model soft solid}

\author{M.~Leocmach}
\email{mathieu.leocmach@ens-lyon.fr}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{C.~Perge}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{N.~Taberlet}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\affiliation{UFR de Physique, Universit\'e Claude Bernard Lyon 1, 43 boulevard du 11 novembre, 69100 Villeurbanne, France}
\author{T.~Divoux}
\affiliation{Centre de Recherche Paul Pascal, CNRS UPR 8642 - 115 avenue Schweitzer, 33600 Pessac, France}
\author{S.~Manneville}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}

\date{\today}

\begin{abstract}
Yoghurts.
\end{abstract}

\pacs{}
\maketitle

%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=7cm,clip]{Fig1.eps}
\caption{(color online) Strain response $\gamma(t)$ in a 4\%~wt. casein gel acidified with 1\%~wt. GDL for an imposed shear stress $\sigma=200$ ($\bullet$), 300 ($\blacktriangledown$), 400 ($\blacksquare$), 550 ($\blacktriangle$), and 1000~Pa ($\blacklozenge$) from right to left. The gap width is 1~mm. The gray dotted line shows $\gamma=1$. Inset: failure time $\tau_f$ vs $\sigma$. The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ and $A=40.6\pm 0.5$~s.Pa$^\beta$.
\label{fig1}}
\end{figure} 
%%%%%%%%%%%%

%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=7cm,clip]{Fig2.eps}
\caption{(color online) Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize each of the three successive creep regimes. $\gpm$ is the minimum shear rate reached at $\taum$ (see text and Suppl. Fig.~2). (a)~Primary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in logarithmic scales. The yellow line is $\gp(t)\sim t^{-0.85}$. Inset: Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 1XXX~\%. Red lines are power laws $G'\sim G''\sim f^{0.15}$. (b)~Secondary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in linear scales. The yellow curve is a sixth order polynomial fit to the data showing that $\taum\simeq 0.6\tau_f$. Inset: $\taum$ vs $\tau_f$. The red line is the best linear fit $\taum=0.598\tau_f$. (c)~Tertiary creep: $\gp(t)/\gpm$ vs $(\tau_f-t)/\tau_f$ in logarithmic scales with a reversed horizontal axis. The yellow line is $\gp(t)\sim (\tau_f-t)^{-1}$. 
\label{fig2}}
\end{figure} 
%%%%%%%%%%%%

%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=18cm,clip]{Fig3.eps}
\caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes. Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images, whose position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$. Experiment performed on a 4\%~wt. casein gel seeded with 3\%~wt. polyamide spheres and acidified with 1\%~wt. GDL under $\sigma=300$~Pa. The gap width is 2~mm. See also Suppl. Movie~1.
\label{fig3}}
\end{figure*} 
%%%%%%%%%%%%

%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=18cm,clip]{Fig4.eps}
\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ (black $\bullet$), $1.7\,10^{-2}$ (blue $\blacktriangledown$), and 0.15 (red $\blacksquare$). Solid lines are linear profiles. (c)~Spatiotemporal diagram of the local velocity $\langle v(r,z,t)\rangle_r$ averaged over the radial direction $r$ and plotted in linear color levels as a function of $z$ and $t/\tau_f$. (d)~Standard deviation $\delta_z v(t)$ of $\langle v(r,z,t)\rangle_r$ taken over the vertical direction $z$ (thick black line) together with corresponding standard deviation $\delta_r v(t)$ computed over the radial direction $r$ on the $z$-average $\langle v(r,z,t)\rangle_z$ (thin red line). (e)~Fracture length $\ell(t)$ vs $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$) and normalized by the height $H$ of the Couette cell. Gray dots show the visualization data for the longest fracture which leads to the failure of the sample at $\tau_f$. Red lines are the best fits $\ell(t)=\alpha+\beta\log(1-t/\tau_f)$ to the visualization data. Same experiment as in Fig.~\ref{fig3}. See also Suppl. Movie~1.
\label{fig4}}
\end{figure*} 
%%%%%%%%%%%%

\begin{acknowledgments}
The authors thank Thomas Gibaud for fruitful discussions and Alan Parker for precious advice and for providing the casein and GDL. This work was funded by the Institut Universitaire de France and by the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement No.~258803. 
\end{acknowledgments}

\bibliography{../../bibseb} 
%\bibliographystyle{jfm}

\clearpage
\newpage
\setcounter{figure}{0}

\section*{\large Supplemental material}

\subsection*{Supplemental movies}

Supplemental Movie~1 shows the creep experiment analyzed in Figs.~\ref{fig3} and \ref{fig4} and performed under $\sigma=300$~Pa on a 4\%~wt. casein gel seeded with 3\%~wt. polyamide spheres and acidified with 1\%~wt. GDL. The gap width of the Couette cell is 2~mm and its height is 60~mm. Images recorded by a standard webcam are shown on the top left. Velocity maps $v(r,z,t)$ inferred from ultrasonic imaging by averaging over 4~s are shown on the top right using linear color levels. The vertical position $z=0$ on the ultrasonic images corresponds to about 20~mm below the top of the Couette cell. The two bottom graphs show the shear rate response $\gp(t)$ (left) and the strain response $\gamma(t)$.

Supplemental Movie~2 shows the failure of a 4\%~wt. casein gel acidified with 1\%~wt. GDL in a plate-plate geometry for an imposed shear stress $\sigma=XXX$~Pa. The plate diameter is 40XXX~mm and the gap width is 1XXX~mm. In this case fractures grow parallel to the vorticitty direction, i.e. along the radial direction. The shear rate response is fully similar to that in the Couette geometry.

\subsection*{Supplemental figures}

 %%%%%%%%%%%%
\begin{figure}[h]
\centering
\includegraphics[width=7cm,clip]{SuppFig1.eps}
\caption{(color online) Gelation process of a 4\%~wt. casein suspension seeded with 3\%~wt. polyamide spheres and acidified with 1\%~wt. GDL. (a)~pH vs time $t$. (b)~Linear viscoelastic moduli $G'$ (top, in blue) and $G''$ (bottom, in red) recorded as a function of time $t$ for fixed frequency $f=1$~Hz and strain amplitude 0.1XXX~\%. Dashed lines indicate that the system starts to build significant elastic properties when the pH decreases below about 5. Lighter colors in (b) correspond to a sample free of polyamide spheres, showing that the addition of acoustic contrast agents does not affect significantly the gelation process and the final viscoelastic properties of the gel.
\label{suppfig1}}
\end{figure}

 %%%%%%%%%%%%
\begin{figure}[h]
\centering
\includegraphics[width=7cm,clip]{SuppFig2.eps}
\caption{(color online) Shear rate response $\gp(t)$ in a 4\%~wt. casein gel acidified with 1\%~wt. GDL for an imposed shear stress $\sigma=200$ ($\bullet$), 300 ($\blacktriangledown$), 400 ($\blacksquare$), 550 ($\blacktriangle$), and 1000~Pa ($\blacklozenge$) from right to left. The gap width is 1~mm. The yellow line shows the power-law behavior $\gp(t)\sim t^{-0.85}$.
\label{suppfig2}}
\end{figure} 
%%%%%%%%%%%%

\end{document}

### 2014-01-17 creep/creep_yaourt.tex
Remarque Mathieu Captions

@@
 Inset: failure time $\tau_f$ vs $\sigma$.
-The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ and $A=40.6\pm 0.5$~s.Pa$^\beta$.
+The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ (Basquin law) with $\beta=5.45\pm 0.05$ and $A=40.6\pm 0.5$~s.Pa$^\beta$.
 \label{fig1}}
@@
 \includegraphics[width=7cm,clip]{Fig2.eps}
-\caption{(color online) Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize each of the three successive creep regimes.
+\caption{(color online)  Three successive creep regimes.
+Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize each of the three successive creep regimes.
 $\gpm$ is the minimum shear rate reached at $\taum$ (see text and Suppl.
@@
 (a)~Primary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in logarithmic scales.
-The yellow line is $\gp(t)\sim t^{-0.85}$.
+The yellow line is $\gp(t)\sim t^{-0.85}$ (Andrade creep).
 Inset: Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 1XXX~\%.
@@
 The red line is the best linear fit $\taum=0.598\tau_f$.
-(c)~Tertiary creep: $\gp(t)/\gpm$ vs $(\tau_f-t)/\tau_f$ in logarithmic scales with a reversed horizontal axis.
+(c)~Tertiary creep: $\gp(t)/\gpm$ vs $(\tau_f-t)/\tau_f$ in logarithmic scales.
 The yellow line is $\gp(t)\sim (\tau_f-t)^{-1}$.
+Note that the time runs leftwards in (c).
 \label{fig2}}
@@
 \includegraphics[width=18cm,clip]{Fig3.eps}
-\caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
+\caption{(color online) Fractures along the vorticity direction.
+Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
 Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images, whose position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$.
@@
 \includegraphics[width=18cm,clip]{Fig4.eps}
-\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ (black $\bullet$), $1.7\,10^{-2}$ (blue $\blacktriangledown$), and 0.15 (red $\blacksquare$).
+\caption{(color online) Local insights into creep and rupture.
+(a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ (black $\bullet$), $1.7\,10^{-2}$ (blue $\blacktriangledown$), and 0.15 (red $\blacksquare$).
 Solid lines are linear profiles.
@@
 (d)~Standard deviation $\delta_z v(t)$ of $\langle v(r,z,t)\rangle_r$ taken over the vertical direction $z$ (thick black line) together with corresponding standard deviation $\delta_r v(t)$ computed over the radial direction $r$ on the $z$-average $\langle v(r,z,t)\rangle_z$ (thin red line).
-(e)~Fracture length $\ell(t)$ vs $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$) and normalized by the height $H$ of the Couette cell.
+(e)~Fracture length $\ell(t)$ normalized by the height $H$ of the Couette cell as a function of $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$).
 Gray dots show the visualization data for the longest fracture which leads to the failure of the sample at $\tau_f$.
@@
 The yellow line shows the power-law behavior $\gp(t)\sim t^{-0.85}$.
+Same experiments as in Figs.~\ref{fig1} and \ref{fig2} of the main text.
 \label{suppfig2}}

### 2014-01-17 creep/creep_yaourt.tex
Premier jet Thibaut

--- 2014-01-17 creep/creep_yaourt.tex
@@
 
+\newcommand{\ti}[1]{\textbf{\color{blue}#1}} % remarques de Thibaut
+\newcommand{\titi}[1]{\sout{\textbf{\color{blue}#1}}} % remarques de Thibaut
+
 \begin{document}
@@
 
-\title{Creep and failure of a model soft solid}
+%\title{Stress-induced yielding of a protein gel: the rupture of a model soft solid}
+
+ \title{Stress-induced fracture of a biopolymer gel}
 
 \author{M.~Leocmach}
-\email{mathieu.leocmach@ens-lyon.fr}
 \affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
@@
 \author{T.~Divoux}
-\affiliation{Centre de Recherche Paul Pascal, CNRS UPR 8642 - 115 avenue Schweitzer, 33600 Pessac, France}
+\email{divoux@crpp-bordeaux.cnrs.fr}
+\affiliation{Centre de Recherche Paul Pascal, CNRS UPR 8641 - 115 avenue Schweitzer, 33600 Pessac, France}
 \author{S.~Manneville}
@@
 \begin{abstract}
-Yoghurts.
+Biopolymer gels are known to behave qualitatively as soft solids: under external stress they display macroscopic fractures followed by an \textit{irreversible} yielding which has not been much investigated.
+As such, biopolymers strongly differ from the far more studied yield stress fluids such as emulsions, colloidal gels and glasses, etc.
+which also display a solid behavior at rest, but can be shear rejuvenated after yielding.
+In this letter, combining optical and ultrasonic imaging to rheology we shed new light on the failure scenario of an acid-induced protein gel.
+Stress-controlled experiments reveal a brittle-like rupture: after a power-law primary creep regime, fully accounted for by the linear elastic properties of the gel, fractures nucleate and slowly grow as the log of time perpendicularly to the shear, up to the suddden rupture.
+The total duration of this failure process surprisingly follows a decreasing power-law with the applied shear stress, strongly reminiscent of Basquin's law of solids.
+These results are discussed in the frame of theoretical approaches including Fiber's Bundles Models.
+
+%Acid-induced protein gels are soft materials that are known to behave qualitatively as soft solids.
+Indeed, under external shear their non-linear behavior feature macroscopic fractures followed by an \textit{irreversible} yielding, and once broken, no shear or rest period allow them to recover their initial elastic properties contrarily to soft glassy materials such as emulsions, colloidal gels and glasses, etc.
+which can be shear rejuvenated.
+
+%The failure time follows a decreasing power-law with the applied shear stress strongly reminiscent of Basquin's law. 
+%Our results are discussed within the frame of Fiber's Bundles Models which account for the fracture growth but fail to reproduce the scaling of the faillure time.
+
+%Under external stress, their non-linear behavior display brittle-like features through the growth of fractures followed by an irreversible yielding.
+In this letter, we report on the stress induced rupture scenario of such protein gels combining optics and ultrasonic velocimetry to rheology.
+We show that protein gels experience a three step rupture .
+The rupture time decreases as an exponentially with the applied shear-stress
 \end{abstract}
@@
 
+%Under a constant external load, a solid deforms and may eventually fail.
+The resulting slow deformation refered as creep has been investigated over a century \cite{Andrade:1910} and differs depending on the nature of the material.
+Brittle materials such as ceramics, cold metals, etc.
+exhibit low plastic deformation prior rupture that is characterized by fast crack propagation perpendicular to the applied stress and a clear-cut fracture, whereas ductile materials such as metals at room temperature, polymers, etc.
+display large plastic deformation and even flow prior any rupture is observed [XXX].
+Interestingly, these two categories defined for hard materials hold true for very soft solids such as viscoelastic fluids and yield stress fluids.
+On the one hand, ductile-like behavior is commonly observed for soft glassy materials such as emulsions, foams microgels, colloidal glass and gels, etc \cite{Gibaud:2010,Schall:2010,Divoux:2011,Siebenburger:2012,Sprakel:2011} which experience extensive plastic deformations and flow without fracture under external stress.
+On the other hand, brittleness has been reported on viscoelastic materials, mostly on transient networks \cite{Ligoure:2013}, and thermoreversible gels \cite{Bonn:1998,Baumberger:2006,Daniels:2007,Brenner:2013}.
+
+Under a constant external load, a solid deforms and may eventually fail.
+The resulting slow deformation refered to as {\it creep} has been investigated over a century on solid materials \cite{Andrade:1910,Nechad:2005,Miguel:2002} but only a handful of papers have dealt with such issue on soft solids \cite{Higgs:1990,vanVliet:1995}, none of which reporting local measurements.
+Soft solids denotes gels made through the reversible association of biopolymers such as proteins or polysaccharides and play key role in food science, from set-style yogourt or cheese, to jelly based deserts, etc.
+\cite{Mezzenga:2005}.
+These materials behave as elastic solids under small deformations, but display remarkable non-linear behavior featuring fractures followed by irreversible rupture \cite{Bonn:1998,Baumberger:2006,Daniels:2007,Brenner:2013}.
+The irreversibility stems directly  from the existence of an external control parameter - the temperature (resp.
+the pH) in the case of thermoreversible gels (resp.
+acid-induced gels) - which, makes the system impossible to shear-rejuvenate after failure, contrarily to other soft glassy materials such as emulsions, colloidal gels and glasses \cite{Cloitre:2000}.
+
+  If a great deal of effort has been made to use proteins and polysaccharides separately or mixed to design soft materials with specific mechanical properties and texture \cite{Dickinson:2006,Gibaud:2012}, their non-linear mechanical behavior under stress has been only partially addressed \cite{vanVliet:1995}.
+Several simple questions remain unanswered: {\it What is the spatially resolved scenario of creep and fracture under applied shear stress}?
+And {\it how far does the analogy with the failure scenario of brittle solid goes}?
 %%%%%%%%%%%%
@@
 \centering
-\includegraphics[width=7cm,clip]{Fig1.eps}
+\includegraphics[width=0.9\columnwidth]{Fig1.eps}
 \caption{(color online) Strain response $\gamma(t)$ in a 4\%~wt.
@@
 %%%%%%%%%%%%
+In the present letter, we report on the stress-induced fracture scenario of acid-induced protein gels by means of rheology coupled to optical and ultrasonic imaging.
+These gels are formed reversibly by acidification of sodium caseinate solution \cite{Lucey:1998}, and at fixed low pH values, display irreversible fractures under large strain \cite{vanVliet:1995} which makes them perfect candidates to quantify brittleness in biopolymer gels and tackle the above-mentioned questions.
+Here we demonstrate thanks to local measurements that under external load, these model soft solids display a brittle-like irreversible yielding which separates into three phases: $(i)$ a primary or Andrade-like creep regime during which the strain field is homogeneous in bulk, $(ii)$ a secondary creep during which fractures nucleate and grow logarithmically from the boundaries toward the center of the experimental cell and $(iii)$ a finite time singularity concomittent to the sudden failure of the gel.
+The failure time displays a non-trivial power-law scaling with the applied shear-stress that contrasts with the exponential scalings usually reported for solid brittle materials \cite{Vanel:2009}.
 
@@
 (b)~Secondary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in linear scales.
-The yellow curve is a sixth order polynomial fit to the data showing that $\taum\simeq 0.6\tau_f$.
+The yellow curve is a sixth order polynomial fit to the data showing that $\taum\simeq 0.6~\tau_f$.
 Inset: $\taum$ vs $\tau_f$.
@@
 
+\textit{Experimental set-up and protocol.-} Gels are prepared by disolving 4\% wt.
+of sodium caseinate powder (Firmenich) in demineralized water under gentle mixing (XX rpm) at T$=35^{\circ}$C.
+To induce the gelation, the solution is acidified with 1\% wt.
+glucono-$\delta$-lactone (GDL) in powder (Firmenich).
+The solution is then stirred for the GDL to dissolve before being poured into the gap of a polished Plexiglas Couette cell (height 50~mm, inner rotating cylinder 23~mm, outer fixed cylinder 25~mm, gap e=2~mm) immersed into a temperature controlled water tank at T=(25.0$\pm$0.1)$^{\circ}$C.
+Rheological data are recorded during the gel formation  via a stress-controled rheometer (ARG2, TA Instruments) through small amplitude oscillatory shear (see Fig.~1 in the Supplemental Material).
+The gelation is over when the increasing linear elastic ($G'$) and viscous ($G''$) modulii reach a plateau with $G'>G''$.
+A constant stress is then applied to the sample.
+Images of the gel are recorded simultaneously to the rheology thanks to a webcam (XXX).
+Displacement field and velocity maps in the (r, z) plane can also be recorded simultaneously to rheology by a custom-made ultrasonic scannner detailed in \cite{Gallot:2013}.
+In this case, the solution is seeded prior to the acidification with accoustic tracers -here 3\% wt.
+polyamide spheres, of diameter 10~$\mu$m (Orgasol 2002 EXD Nat 1, Arkema) that do not modify the final gel properties [see Supplemental Fig.~\ref{suppfig1}~(b)].
+Yielding being irreversible, each creep experiment requires to prepare a new sample in-situ.
+
 %%%%%%%%%%%%
-\begin{figure*}
+\begin{figure*}[t]
 \centering
@@
 %%%%%%%%%%%%
+
+\textit{Macroscopic behaviour.-} Under constant applied shear stress $\sigma$, the global deformation $\gamma$ displays the same robust evolution with time (Fig.~\ref{fig1}): it grows almost logarithmicaly with time up to $\gamma \sim 1$, than accelerates up to the moment the gel breaks.
+ The failure time $\tau_f$ of the gel decreases as a power-law of $\sigma$ [Fig.~\ref{fig1}~(inset)] and allow us to rescal all the experiments on a single mastercurve.[see fig.~\ref{suppfig2} in the Supplemental Materials]
+Figure~\ref{fig2} displays such mastercurve as the temporal evolution of the shear rate $\dot \gamma$ and each subplot (a), (b) and (c) uses a scale that lay the emphasis respectively on the primary, the secondary and the tertiary regime.
+In the primary creep regime [Fig.~\ref{fig2}~(a)], shear rates decrease as the following power-law:  $\dot \gamma(t) \sim~\dot \gamma_{min} (t/\tau_f)^{-\alpha}$, with $\alpha=0.85$ recalling the primary creep observed in solids refered to as the Andrade creep \cite{Miguel:2002}.
+Interestingly, here this scaling is directly linked to the linear viscoelasticity which predicts for such power-law creep the following scaling: $G(\omega) \sim \omega^{1-\alpha}$.
+Inset of figure~\ref{fig2}~(a) confirms that $G'(\omega)$ and $G"\sim \omega ^{0.15}$, strongly suggesting that the primary creep is purely elastic.
+During the secondary creep regime, the shear-rate reaches a minimum value at $t=\tau_{min}= (0.61\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}~(b)-(inset)] as similarly reported for solid composite materials \cite{Nechad:2005}, colloidal gels of attractive particles \cite{Grenard:2013} and Fiber Bundles Models \cite{Kovacs:2008,Jagla:2011}.
+In the tertiary creep regime, the shear rate increases by several orders of magnitude scaling as $(\tau_f-t)^{-1}$ and display a finite time singularity.
+Finally, the failure time of the gel is a decreasing function of the applied shear stress in agreement with what has been reported qualitativeley in similar systems \cite{vanVliet:1995}, and scales as the following power-law: $\tau_f \sim \sigma^{-5.45\pm 0.05}$, a key result on which we come back in the discussion section.
+
+\textit{Local behaviour.-} Local measurements allow us to get deeper insight into the creep regimes [see movie~1 in the Supplemental material].
+During the primary regime, no cracks or failure can be seen from direct visualization [Fig.~\ref{fig3}.~(a)] while ultrasonic imaging reveals that within the spatial resolution of a few microns, the strain field is homogeneous [Fig~\ref{fig4}~(b)].
+If plastic events take place in this primary regime, they occur at smaller scale.
+However, given the agreement between the Andrade exponent and the frequency scaling of the linear elastic modulus discussed above, it is more likely that no or a negligible amount of plastic events take place during this regime, coroborating the optical and ultrasonic measurements.
+In the secondary creep, fractures nucleate from the top and bottom edges of the Couette cell and start growing perpendicularly to the shear.
+In agreement, velocity maps become heterogeneous [Fig.~\ref{fig3}.~(b)-(c)] and display intermittent fluctuations along the cell height corresponding to the independant growth of the vertical cracks [Fig~\ref{fig4}~(d)].
+In the tertiary creep regime, cracks have grown along the vorticity, in a periodic fashion in the orthoradial direction while the velocity map directly correlates with the crack growth [Fig.~\ref{fig3}.~(d)].
+Finally, over secondary and tertiary regimes, direct visualization and ultrasonic velocimetry show that the cracks display a continuous logarithmic growth with no remarkable change between the two regimes [Fig~\ref{fig4}~(e)].
+Note that these three regimes do not depend on the geometry, and are also observed to be quantitatively the same in plate-plate geometry [see movie~2 in the Supplemental Materials].
+
+\textit{Discussion.-} As a central result, acid-induced casein gels display a remarkable brittle-like rupture scenario, different from other soft materials such as transient-network systems \cite{Ligoure:2013}, because truly irreversible.
+The pH is key to the irreversibility and plays the role of an additional control parameter that, when maintained at values lower than the isoelectric point, prevent the system from healing.
+Comparatively, Yield Stress Fluids (YSF) that display a solid-like behavior at rest, either because of a jammed microstructure or attractive interactions \cite{Sciortino:2005,Bonnecaze:2010}, but do not present such external control parameter, display a similar rheological signature during their stress-induced yielding.
+Indeed the S-shaped curved pictured in figure~\ref{fig2}~(a) has been also reported in \cite{Bauer:2006,Gibaud:2010,Divoux:2011,Siebenburger:2012,Grenard:2013}.
+However, their yielding scenario is fundamentally different.
+Local measurements conducted on both type of soft glassy materials have revealed that no fractures take place in the bulk: the microstructure rather reorganizes and flows in a ductile way.
+Our study reveals that acid-induced gels behave elastically up to the rupture point and that fractures develop and grow logarithmically, perpendicularly to the shear direction prior an irreversible failure which makes casein gels model {\it brittle soft solids}.
 
@@
 
+Let us examine in more detail the rupture process and compare it to what is seen on brittle solid materials.
+Here, the primary creep and its power-law exponent ($\dot \gamma \sim t^{-\alpha}$, with $\alpha=0.85$) is fully accounted for by the linear rheology while local measurements reveals in agreement an homogeneous bulk deformation.
+This result strongly contrasts with the primary creep of both solid materials and YSFs \cite{Bauer:2006,Divoux:2011,Siebenburger:2012,Grenard:2013} that usually exhibit a power-law creep with slightly smaller exponents (0.3 $\leq \alpha \leq 0.7$) that are not directly related to linear rheology.
+Besides, the linear behavior of YSF generally differs from a simple power-law scaling, while in the case of solids, the primary creep is irreversible and has been attributed to dislocation interactions in the case or cristaline materials \cite{Miguel:2002} or localized plastic rearragements which nature is still unclear in amorphous materials \cite{Rosti:2010}.
+\ti{ XXX The exponent in the case of casein gel is robust and neither depend on the geometry nor on the GDL concentration within the explored range ...
+mentioned or left aside for the long paper?
+XXX}
+
+The second key feature of the failure process of casein gels is the logarithmic growth of fractures observed in the second and tertiary creep regimes [Fig.~\ref{fig4}~(d)].
+Such evolution  has been also reported on solid disordered materials displaying brittle rupture (such as paper) and interpreted in the framework of Griffith-like or Local Energy Barrier (LEB) models \cite{Vanel:2009}.
+However, these approaches predict an exponential scaling of the rupture time with the applied stress, which gives a less convincing fit of the experimental data than the decreasing power-law we proposed in figure~\ref{fig1}~(inset).
+This result strongly suggests that another mechanism is at stake in the failure process of casein gels.
+\ti{XXX 1) Should we discuss in more details the LEB and Griffith models?
+2) Should we mention that the log growth differs from the algebric growth of capillary fractures observed in thermo-reversible gels \cite{Daniels:2007} ??
+XXX}
+
+The last key feature of this brittle scenario is the power-law dependance of the failure time with the applied shear stress...XXX
+
+\begin{itemize}
+\item FBM are relevant here because the gel microstructure is formed of strands \cite{Kalab:1983,Roefs:1990}.
+FBM that rely on elastic fibers with a local yield strain (or stress) and introduce a non-linear Eyring rheology or a local coupling lead to a different scaling law than the one we find here for the failure time with the applied stress.
+In their law appears a ``yield stress" \cite{Kun:2003,Nechad:2005,Kovacs:2008,Jagla:2011} that is not present here.
+The promising approach combines elastic fibers with a local yield strain and take into account damage accumulation as in \cite{Kun:2007,Halasz:2012}.
+\item Idem Compare to the scalings observed in YSF - in particular microgels display a similar power-law scaling, except that the yield point appears explicitely in the scaling as the critical point in a dynamical phase transition \cite{Divoux:2012}.
+Propose an explenation or at least raise a point.
+\item re-Read \& mention somewhere the following ref. \cite{Chaudhuri:2013} 
+\item Note for us, that some materials display a brittle or a ductile behavior depending on the strain-rate they experience - see the review by Mikko Alava on ``paper" or even foams- travail de J.C Tsai.
+It urges for systematic step shear experiments on casein gels :-).
+\item conclude   
+\end{itemize}
+
+%Acidification is reversible: one can switch reversibly from the gel to the casein solution by tuning the pH respectively below or the casein isoelectric point (pH$\sim$~4.6).
+
 \begin{acknowledgments}
-The authors thank Thomas Gibaud for fruitful discussions and Alan Parker for precious advice and for providing the casein and GDL.
+The authors thank Thomas Gibaud for fruitful discussions, Miloslav Kal\'ab for kindly providing us with unobtainable references and Alan Parker for precious advice and for providing the casein and GDL.
 This work was funded by the Institut Universitaire de France and by the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement No.~258803.
@@
 
-\bibliography{../../bibseb} 
-%\bibliographystyle{jfm}
+\begin{thebibliography}{}
+
+\bibitem{Andrade:1910} E.N. Da C. Andrade Proc. R. Soc. London A {\bf 84}, 1 (1910).
+
+\bibitem{Nechad:2005} H. Nechad et al., Phys. Rev. Lett. {\bf 94}, 045501 (2005); J. Mech. Phys. Solids {\bf 53}, 1099-1127 (2005).
+
+\bibitem{Miguel:2002} M.C. Miguel, A. Vespignani, M. Zaiser and S. Zapperi, Phys. Rev. Lett. {\bf 89}, 165501 (2002); M.C. Miguel, L. Laurson and M.J. Alava, Eur. Phys. J. B {\bf 64}, 443--450 (2008).
+
+\bibitem{Higgs:1990} P.G. Higgs and S.B. Ross-Murphy, Int. J. Biol. Macromol. {\bf 12}, 233--240, (1990).
+
+\bibitem{vanVliet:1995} T.
+van Vliet and P.
+Walstra, Faraday Discuss.
+{\bf 101}, 359--370 (1995); T.
+van Vliet, H.J.M.
+van Dijk, P.
+Zon and P.
+Walstra, Colloid Polym.
+Sci.
+{\bf 269}, 620--627 (1991); L.G.B.
+Bremer et al., Colloids Surf.
+{\bf 51}, 159--170 (1990).
+
+\bibitem{Mezzenga:2005} R. Mezzenga, P. Schurtenberger, A. Burbidge and M. Michel, Nature Materials {4}, 729--740 (2005).
+
+\bibitem{Brenner:2013} T. Brenner et al., J. Non-Newtonian fluid Mech. {\bf 196}, 1--7 (2013).
+
+\bibitem{Bonn:1998} D. Bonn et al., Science {280} 265--267 (1998).
+
+\bibitem{Baumberger:2006} T. Baumberger, C. Caroli and D. Martina, Nature Materials {\bf 5}, 552-555 (2006).
+
+\bibitem{Daniels:2007} K. Daniels et al., Phys. Rev. Lett. {\bf 99}, 124501 (2007).
+
+\bibitem{Cloitre:2000} M.
+Cloitre, R.
+Borrega and L.
+Leibler, Phys.
+Rev.
+Lett.
+{\bf 85} 4819--4822 (2000); V.
+Viasnoff and F.
+Lequeux, Phys.
+Rev.
+Lett.
+{\bf 89} 065701 (2002); Coussot et al., Phys.
+Rev.
+Lett.
+{\bf 88}, 175501 (2002).
+
+\bibitem{Dickinson:2006} E. Dickinson, Soft Matter {\bf 2}, 642--652 (2006).
+
+\bibitem{Gibaud:2012} T. Gibaud et al., Faraday Discuss., {\bf 158}, 267--284 (2012).
+
+\bibitem{Lucey:1998} J.A. Lucey and H. Singh, Food Research Int. {\bf 30}, 529--542 (1998).
+
+%\bibitem{vanVliet:1991} T. van Vliet, H.J.M. van Dijk, P. Zon and P. Walstra, Colloid Polym. Sci. {\bf 269}, 620--627 (1991).
+
+%\bibitem{Bremer:1990} L.G.B. Bremer et al., Colloids Surf. {\bf 51}, 159--170 (1990).
+
+\bibitem{Vanel:2009} L. Vanel et al., J. Phys. D: Appl. Phys. {\bf 42} 214007 (2009).
+
+\bibitem{Gallot:2013} T. Gallot et al., Rev. Sci. Instrum. {\bf 84}, 045107 (2013).
+
+\bibitem{Grenard:2013} V. Grenard, T. Divoux, N. Taberlet and S. Manneville, Soft Matter (2013) - DOI: 10.1039/C3SM52548A
+
+\bibitem{Jagla:2011} E.A. Jagla, Phys. Rev. E {\bf 83} 046119 (2011).
+
+\bibitem{Kovacs:2008} K. Kov\'acs et al., Phys. Rev. E \textbf{77}, 036102 (2008).
+
+\bibitem{Sciortino:2005} F. Sciortino and P. Tartaglia, Advances in Physics {\bf 54}, 471--524 (2005).
+
+\bibitem{Bonnecaze:2010} R.T. Bonnecaze and M. Cloitre, Adv. Polym. Sci. {\bf 236}, 117--161 (2010).
+
+\bibitem{Bauer:2006} T. Bauer, J. Oberdisse and L. Ramos, Phys. Rev. Lett. {\bf 97}, 258303 (2006).
+
+\bibitem{Divoux:2011} T. Divoux, C. Barentin and S. Manneville, Soft Matter {\bf 7}, 8409--8418 (2011).
+
+\bibitem{Siebenburger:2012} Siebenb\"urger, M. Ballauff and T. Voigtmann, Phys. Rev.
+Lett., {\bf 108}, 255701 (2012)
+
+\bibitem{Gibaud:2010} T. Gibaud, D. Frelat and S. Manneville, Soft Matter {\bf 6}, 3482--3488 (2010).
+
+\bibitem{Rosti:2010} J. Rosti, J. Koivisto, L. Laurson, M.J. Alava, Phys. Rev. Lett. {\bf 105}, 100601 (2010).
+
+%\bibitem{Sprakel:2011} J. Sprakel et al., Phys. Rev. Lett. {\bf 106}, 248303 (2011).
+
+\bibitem{Kun:2003} F. Kun et al., Phys. Rev. E {\bf 67} 061802 (2003).
+
+%\bibitem{Ghofraniha:2012} N. Ghofraniha et al., Soft Matter {\bf 8}, 6214 (2012).
+
+\bibitem{Ligoure:2013} C. Ligoure and S. Mora, Rheol. Acta {\bf 52}, 91-114 (2013).
+
+\bibitem{Kun:2007} F. Kun et al., J. Stat. Mech. P02003 (2007); F. Kun et al., Phys. Rev. Lett. {\bf 100}, 094301 (2008).
+
+\bibitem{Roefs:1990} S.P.F.M. Roefs, A.E.A. De Groot-Mostert and T. Van Vliet, Colloids Surf. {\bf 50}, 141--159 (1990).
+
+\bibitem{Kalab:1983} M. Kal\'ab, P. Allan-Wojtas and B.E. Phipps-Todd, Food Microstructure {\bf 2}, 51--66 (1983).
+
+\bibitem{Divoux:2012} T. Divoux, D. Tamarii, C. Barentin, S. Teitel and S. Manneville, Soft Matter {\bf 8}, 4151---164 (2012).
+
+\bibitem{Halasz:2012} Z. Hal\'asz, Z. Danku and F. Kun, Phys. Rev. E {\bf 85}, 016116 (2012).
+
+%\bibitem{Roefs:1986} S.P.F.M.
+Roefs, {\it Structure of acid casein gels.
+A study of gels formed after acidification in the cold.
+PhD Thesis}, Wageningen Agricultural University, The Netherlands (1986).
+
+%\bibitem{Schall:2010} P. Schall and M. van Hecke, Annu. Rev. Fluid Mech. {\bf 42} 67--88 (2010).
+
+\bibitem{Chaudhuri:2013} P. Chaudhuri and J. Horbach, Phys. Rev. E {\bf 88}, 040301(R) (2013).
+
+\end{thebibliography}
 
@@
 
+Figure~\ref{suppfig1} display the gelation process of a 4\%.
+sodium caseinate solution seeded with 3\% wt.
+polyamide spheres.
+The glucono-$\delta$-lactone contains an ester function which, once added to the solution, hydrolyzes spontaneously and leads to a slow decrease of the pH [Fig.~\ref{suppfig1}~(a)] towards the casein isoelectric point (pH$\simeq 4.6$) at which casein particles start to aggregate.
+The gelation starts in fact earlier at pH=5 as already discussed in \cite{Lucey:1998supp}.
+At this point, the elastic and viscous modulii display a brutal increase, then overshoot and converge toward their steady state values, where $G'~>~G''$.
+Note that the maximum of the overshoot is reached at the isoelectric point.
+The decrease of both modulii at lower pH is usually attributed to the overacidification which enhances the repulsive electrostatic interactions between caseinate particles of net positive charge \cite{Dickinson:2002supp}.
+
  %%%%%%%%%%%%
@@
 
+\begin{thebibliography}{}
+
+\bibitem{Lucey:1998supp} J.A. Lucey and H. Singh, Food Research International A {\bf 30}, 529--542 (1998).
+
+\bibitem{Dickinson:2002supp} E. Dickison and L.M. Merino, Food Hydrocolloids {\bf 16}, 321--331 (2002).
+
+\end{thebibliography}
+
 \end{document}

### 2014-01-17 creep/creep_yaourt.tex
Couleur des symboles dans les captions

@@
 casein gel acidified with 1\%~wt.
-GDL for an imposed shear stress $\sigma=200$ ($\bullet$), 300 ($\blacktriangledown$), 400 ($\blacksquare$), 550 ($\blacktriangle$), and 1000~Pa ($\blacklozenge$) from right to left.
+GDL for an imposed shear stress $\sigma=200$ (\textcolor{red!25!black}{$\bullet$}), 300 (\textcolor{red!50!black}{$\blacktriangledown$}), 400 (\textcolor{red!75!black}{$\blacksquare$}), 550 (\textcolor{red}{$\blacktriangle$}), and 1000~Pa (\textcolor{orange!50!red}{$\blacklozenge$}) from right to left.
 The gap width is 1~mm.
@@
 \includegraphics[width=18cm,clip]{Fig4.eps}
-\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ (black $\bullet$), $1.7\,10^{-2}$ (blue $\blacktriangledown$), and 0.15 (red $\blacksquare$).
+\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ ($\bullet$), $1.7\,10^{-2}$ (\textcolor{blue}{$\blacktriangledown$}), and 0.15 (\textcolor{red}{$\blacksquare$}).
 Solid lines are linear profiles.

### 2014-01-24 plis_figures.tex
Start to rework fig:wrinkles with data from cas8

@@
 \usepackage{pgfplots}
+\usepgfplotslibrary{groupplots}
 
@@
 (b) Sketch of the cell with no adhesion of the protein on the top and bottom plates.
-(c-e) Resulting patterns in the same sample ($e=\SI{100}{\micro\metre}$) at different magnification by (c-d) reflexion or (e) transmission optical microscopy.}
+(c-e) Resulting patterns in the same sample ($e=\SI{100}{\micro\metre}$) at different magnification by (c-d) reflexion or (e) transmission optical microscopy (tiled).}
 	\label{fig:acidgel}
@@
 (top) By light transmission microscopy.
-(bottom) Reconstructed from fluorescent confocal microscopy.
+(bottom) Reconstructed from fluorescent confocal microscopy (corresponds to the squared area in Fig.~\ref{fig:acidgel}e).
 Scale bars are \SI{1}{\milli\metre}.}
@@
 
+\begin{figure}
+	\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={group name=g, group size=1 by 2},
+		xmin=0, xmax=200, xtick={20,40,...,180},
+		xlabel={time (min)}, 
+		extra tick style={grid=major},%
+		extra x tick labels={},%
+		width=\columnwidth,
+		height=0.5\columnwidth,
+		]
+	\nextgroupplot[ylabel={Relative volume}, ymin=0.3, ymax=1, extra x ticks={12.55}]
+	\addplot+[no marks,Accent1] table{volume_rel_half_cas8.txt};
+	\node[rotate=90, anchor=base] at (axis cs:12.55,0.75) (syn) {synæresis};
+	\node[right=0 of syn.south west] {swelling};
+
+	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=4, extra x ticks={26.7, 42.8}]
+	\addplot+[no marks,Accent1] table{excess_area_pc_half_cas8.txt};
+	\end{groupplot}
+
+	%\node[inner sep=0, anchor=south east] (g-1-1.outer north west){\includegraphics[width=0.45\columnwidth]{plastic_confocal}};
+	\end{tikzpicture}
+	\caption{Syneresis and swelling (c) Evolution of the volume of the gel phase. (d) Evolution of the midsurface area.}
+	\label{fig:wrinkles}
+\end{figure}
+
 \end{document}

### 2014-01-27 creep/creep_yaourt.tex
V3.0 : Sébastien - encore rechangé le titre (à discuter si besoin)
- pas mal repris l'abstract en lien avec les messages de l'intro et de la conclu

- axé l'intro sur les "biogels" : après mûre réflexion, le démarrage sur les solides durs et le creep historique d'Andrade (certes plus rigoureux) me semblait moins sexy… Pour les refs de la première phrase, je n'ai mis que des revues alors que les refs de la 2ème phrase sont plus précises. Il y a clairement certaines approximations dans le 1er paragraphe mais je ne pense pas y dire trop de "bonneries". On peut débattre de tout cela à l'occasion d

--- 2014-01-17 creep/creep_yaourt.tex
@@
 
-\title{Creep and failure of a model soft solid}
+ \title{Creep and brittle failure of a protein gel under stress}
 
@@
 \affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
-\author{N.~Taberlet}
-\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
-\affiliation{UFR de Physique, Universit\'e Claude Bernard Lyon 1, 43 boulevard du 11 novembre, 69100 Villeurbanne, France}
 \author{T.~Divoux}
-\affiliation{Centre de Recherche Paul Pascal, CNRS UPR 8642 - 115 avenue Schweitzer, 33600 Pessac, France}
+%\email{divoux@crpp-bordeaux.cnrs.fr}
+\affiliation{Centre de Recherche Paul Pascal, CNRS UPR 8641 - 115 avenue Schweitzer, 33600 Pessac, France}
 \author{S.~Manneville}
@@
 \begin{abstract}
-Yoghurts.
+Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids in that they rupture under an external load.
+Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is typical of brittle solids: after a power-law creep regime fully accounted for by linear viscoelasticity and homogeneous deformation, fractures nucleate and grow logarithmically perpendicularly to shear up to sudden rupture.
+The failure time surprisingly follows a decreasing power-law with the applied shear stress strongly reminiscent of the Basquin law of fatigue for solids.
+These results are in excellent agreement with recent fiber-bundle models that include damage accumulation on elastic fibers and exemplify protein gels as model brittle soft solids.
 \end{abstract}
@@
 
+%Under a constant external load, a solid deforms and may eventually fail.
+The resulting slow deformation, referred to as {\it creep}, has been investigated over a century on solid materials \cite{Andrade:1910,Nechad:2005,Miguel:2002} but only a handful of papers have dealt with creep on soft solids, i.e.
+gels made through the reversible association of biopolymers such as proteins or polysaccharides \cite{Higgs:1990,vanVliet:1995}.
+These materials, whose physical properties play a key role not only in food science, from set-style yogurt or cheese to jelly-based deserts \cite{Mezzenga:2005} but also in biomechanics, biochemistry and microbiology \cite{Viovy:2000,Grinnell:2010}, behave as elastic solids under small deformations but display remarkable nonlinear behavior featuring fractures prior to irreversible rupture \cite{Bonn:1998,Baumberger:2006,Daniels:2007,Brenner:2013}.
+Irreversibility directly stems from the existence of an external control parameter, namely temperature (resp.
+pH) in the case of thermoreversible gels (resp.
+acid-induced gels), which makes the system impossible to shear-rejuvenate after failure, contrary to other soft glassy materials such as emulsions, colloidal gels and glasses \cite{Cloitre:2000,Divoux:2012,Siebenburger:2012}.
+While great effort has been devoted to designing protein- and polysaccharide-based soft solids with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}, their nonlinear mechanical behavior under a shear stress has only been partially addressed \cite{vanVliet:1995} and several simple issues remain unexplored such as the spatially resolved scenario for creep and fracture or the physical relevance of the analogy with the failure of hard, brittle solids.
+
+Biogels formed through the self-association of polysaccharide coils, collagen, actin filaments or attractive globular proteins play a major role in biochemistry and microbiology \cite{Viovy:2000}, biological networks and cell mechanics \cite{Stricker:2010} as well as in food science \cite{Mezzenga:2005}.
+These biomaterials all behave as elastic solids under small deformations but display remarkable nonlinear behavior generally featuring stress- or strain-stiffening \cite{Gardel:2004} and fractures prior to irreversible rupture \cite{Bonn:1998,Baumberger:2006,Daniels:2007}.
+Irreversibility stems from the existence of an external control parameter, e.g.
+temperature or pH in the case of thermoreversible or acid-induced gels respectively, which makes such biogels impossible to shear-rejuvenate after failure, contrary to other soft glassy materials such as emulsions, colloidal gels and glasses \cite{Cloitre:2000,Divoux:2012,Siebenburger:2012}.
+So far huge effort has been devoted to designing protein gels with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}.
+However, their mechanical behavior under a shear stress deep into the nonlinear regime has only been partially addressed \cite{Higgs:1990,vanVliet:1995} and several simple issues remain unexplored such as the spatially resolved rupture scenario or the physical relevance of the analogy brittle failure in hard solids.
+
 %%%%%%%%%%%%
@@
 \centering
-\includegraphics[width=7cm,clip]{Fig1.eps}
+\includegraphics[width=7cm]{Fig1.eps}
 \caption{(color online) Strain response $\gamma(t)$ in a 4\%~wt.
 casein gel acidified with 1\%~wt.
-GDL for an imposed shear stress $\sigma=200$ ($\bullet$), 300 ($\blacktriangledown$), 400 ($\blacksquare$), 550 ($\blacktriangle$), and 1000~Pa ($\blacklozenge$) from right to left.
+GDL for an imposed shear stress $\sigma=200$ (\textcolor{red!25!black}{$\bullet$}), 300 (\textcolor{red!50!black}{$\blacktriangledown$}), 400 (\textcolor{red!75!black}{$\blacksquare$}), 550 (\textcolor{red}{$\blacktriangle$}), and 1000~Pa (\textcolor{orange!50!red}{$\blacklozenge$}) from right to left.
 The gap width is 1~mm.
-The gray dotted line shows $\gamma=1$.
+Gray dashes show $\gamma=1$.
 Inset: failure time $\tau_f$ vs $\sigma$.
-The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ (Basquin law) with $\beta=5.45\pm 0.05$ and $A=40.6\pm 0.5$~s.Pa$^\beta$.
+The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ and $A=40.6\pm 0.5$~s.Pa$^\beta$.
 \label{fig1}}
@@
 %%%%%%%%%%%%
+
+In this Letter we report on the stress-induced fracture scenario of protein gels by means of rheology coupled to optical and ultrasonic imaging.
+Gels formed reversibly by slow acidification of a sodium caseinate solution \cite{Lucey:1998} display fractures under large strain at fixed low pH values \cite{vanVliet:1995}, which makes them perfect candidates to quantify brittleness in soft solids and tackle the above-mentioned issues.
+We demonstrate that under external load, these casein gels display brittle-like failure that results from two successive physical processes: a reversible primary (or Andrade-like) creep regime characterized by a homogeneous strain field and the irreversible nucleation and growth of fractures leading to gel failure.
+Our results are in full agreement with the predictions of some recent fiber-bundle models and hint to universal features of brittle failure common to both soft and hard solids.
 
@@
 \includegraphics[width=7cm,clip]{Fig2.eps}
-\caption{(color online)  Three successive creep regimes.
-Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize each of the three successive creep regimes.
+\caption{(color online) Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize each of the three successive regimes.
 $\gpm$ is the minimum shear rate reached at $\taum$ (see text and Suppl.
 Fig.~2).
+The yellow line shows the best fit of $\gp(t)$ using Eq.~(\ref{e.fit}) with $\alpha=0.85$, leading to $\lambda=0.378\pm 0.002$ and $\mu=0.187\pm 0.002$.
 (a)~Primary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in logarithmic scales.
-The yellow line is $\gp(t)\sim t^{-0.85}$ (Andrade creep).
-Inset: Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 1XXX~\%.
+Inset: Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 0.1~\%.
 Red lines are power laws $G'\sim G''\sim f^{0.15}$.
 (b)~Secondary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in linear scales.
-The yellow curve is a sixth order polynomial fit to the data showing that $\taum\simeq 0.6\tau_f$.
+Gray dashes show the minimum of Eq.~(\ref{e.fit}) reached at $\taum=0.556\tau_f$.
 Inset: $\taum$ vs $\tau_f$.
-The red line is the best linear fit $\taum=0.598\tau_f$.
-(c)~Tertiary creep: $\gp(t)/\gpm$ vs $(\tau_f-t)/\tau_f$ in logarithmic scales.
-The yellow line is $\gp(t)\sim (\tau_f-t)^{-1}$.
-Note that the time runs leftwards in (c).
-\label{fig2}}
+The red line is $\taum=0.56\tau_f$.
+(c)~Tertiary creep: $\gp(t)/\gpm$ vs $(\tau_f-t)/\tau_f$ in logarithmic scales with a reversed horizontal axis.\label{fig2}}
 \end{figure} 
 %%%%%%%%%%%%
+
+Gels are prepared by dissolving sodium caseinate powder (Firmenich) at 4\% wt.
+in deionized water under gentle mixing at 35$^{\circ}$C and 500~rpm.
+To induce gelation, 1\% wt.
+glucono-$\delta$-lactone (GDL) in powder (Firmenich) is dissolved in the solution and its hydrolysis progressively lowers the pH over the course of 8~hours [Fig.~\ref{suppfig1}(a) in the Supplemental Material].
+While still liquid, the solution is poured into the gap of a polished Plexiglas Couette cell immersed into a temperature-controlled water tank at 25.0$\pm$0.1$^{\circ}$C \cite{note}.
+Rheological data are recorded during gel formation by a stress-controlled rheometer through small amplitude oscillatory shear at frequency $f=1$~Hz (Fig.~1 in the Supplemental Material) \cite{note}.
+Gelation is complete when the elastic ($G'$) and viscous ($G''$) moduli reach a plateau with $G'\gg G''$.
+A constant stress $\sigma$ is then applied to the sample at time $t=0$ and the subsequent strain response $\gamma(t)$ is monitored.
+Images of the gel are recorded simultaneously to the rheology (Logitech Webcam Pro 9000).
+The local velocity and strain fields can also be imaged in the gradient--vorticity plane $(r,z)$ simultaneously to rheology by a custom-made ultrasonic scanner detailed in \cite{Gallot:2013}.
+In this case, prior to acidification, the sodium caseinate solution is seeded with acoustic tracers here 3\% wt.
+polyamide spheres (Orgasol 2002 ES3 NAT 3, Arkema, diameter 30~$\mu$m, density 1.02) that do not modify the final gel properties [Fig.~\ref{suppfig1}(b) in the Supplemental Material].
+Failure being irreversible, each creep experiment requires to prepare a fresh sample in-situ.
+
+Under a constant applied shear stress $\sigma$, the global strain $\gamma(t)$ displays a robust time dependence (Fig.~\ref{fig1}): $\gamma(t)$ slowly grows with time up to $\gamma \sim 1$ then accelerates until the gel fails at a well-defined time $\tau_f$.
+These three successive steps are better highlighted in Fig.~\ref{fig2} by focusing on the global shear rate $\gp(t)$.
+Figures~\ref{fig3} and \ref{fig4} gather the results from {\it local} measurements and are discussed below together with each of the successive regimes inferred from {\it global} data.
+Note that these failure dynamics do not depend on the geometry as checked in a plate-plate geometry [see Movie~1 in the Supplemental Materials].
+
+%%%%%%%%%%%%
+\begin{figure*}[t]
+\centering
+\includegraphics[width=17.5cm,clip]{Fig3.eps}
+\caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
+Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images, whose position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$.
+The gap width is 2~mm and $r$ is the radial distance to the inner rotating cylinder.
+Experiment performed on a 4\%~wt.
+casein gel seeded with 3\%~wt.
+polyamide spheres and acidified with 1\%~wt.
+GDL under $\sigma=300$~Pa.
+See also Movie~2 in the Supplemental Material.
+\label{fig3}}
+\end{figure*} 
+%%%%%%%%%%%%
+
+As seen in the inset of Fig.~\ref{fig1}, the failure time $\tau_f$ sharply decreases as a power-law of $\sigma$ with an exponent $\beta\simeq 5.5$.
+$\tau_f$ further allows us to rescale all the shear rate data $\gp(t)$ onto the single master curve of Fig.~\ref{fig2}(a) by plotting $\gp/\gp_{\rm min}$ vs $t/\tau_f$, where $\gp_{\rm min}$ is the minimum shear rate reached at a time $\tau_{\rm min}$ [see also Fig.~\ref{suppfig2} in the Supplemental Material for the original $\gp(t)$ data].
+On about four decades for $t\lesssim 0.1\tau_f$, the shear rate decreases as a power-law $\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$.
+This is strongly reminiscent of the {\it primary creep} observed in solids and referred to as Andrade creep \cite{Andrade:1910,Miguel:2002,Nechad:2005}.
+Interestingly, here, this scaling can be fully accounted for by linear viscoelasticity without invoking any irreversible plasticity.
+Indeed casein gels display a power-law rheology $G'(f)\sim G''(f)\sim f^{0.15}$ [inset of Fig.~\ref{fig2}(a)], which corresponds to a compliance $J(t)\equiv\gamma(t)/\sigma\sim t^{0.15}$ in the linear deformation regime \cite{Jaishankar:2013}.
+This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$ and suggests that primary creep is purely reversible without unrecoverable plastic events.
+Homogeneity and linear response are further confirmed, at least up to the available spatial ($\sim 10~\mu$m) and temporal ($\sim 1$~s) resolutions, by direct visualization which shows no sign of crack or fracture [Fig.~\ref{fig3}(a)] and by ultrasonic imaging which reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)].
+
+For $0.1\lesssim t/\tau_f\lesssim 0.9$, $\gp(t)$ departs from power-law behavior and goes through a minimum value at time $\tau_{\rm min}= (0.56\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}(b) and inset] as similarly reported for solid composite materials \cite{Nechad:2005} and fiber-bundles models (FBMs) \cite{Kovacs:2008,Jagla:2011}.
+During this {\it secondary creep} regime, regularly-spaced cracks nucleate from the top and bottom edges of the Couette cell and start growing perpendicular to the applied stress [Fig.~\ref{fig3}(b), see also Movie~2 in the Supplemental Material].
+These macroscopic fractures are invaded with water expelled from the surrounding gel matrix.
+Although fractures have not yet entered the ultrasonic region of interest, velocity maps become heterogeneous along the $z$ direction [Fig.~\ref{fig3}(b, left)] and display intermittent fluctuations [Fig~\ref{fig4}(c,d)].
+The level of these fluctuations, which can also be seen on the global response for the same applied stress [$\blacktriangledown$ in Fig.~\ref{fig2}(b)], is poorly reproducible and stress-dependent.
+We propose that such intermittency results from self-healing microscopic cracks which precede the tip of the fracture.
+Since heterogeneity is much smaller along the radial direction [compare red and black lines in Fig~\ref{fig4}(d)], these ``precursors'' either extend across the full 2-mm gap or induce long-range displacements that span the gap.
+In any case, such noisy features are typical of slow crack growth in solids \cite{Vanel:2009,Kun:2007} and more generally of avalanche-like behavior \cite{Sethna:2001}.
+
+In the {\it tertiary creep} regime, for $t\gtrsim 0.9\tau_f$, the shear rate increases by more than four orders of magnitude and diverges as $(\tau_f-t)^{-1}$ [Fig.~\ref{fig2}(c)].
+This finite-time singularity corresponds to the final growth of the fractures along the vorticity direction $z$ as they accelerate and eventually meet in the middle of the Couette cell at time $\tau_f$ [Fig.~\ref{fig3}(c-d)].
+Ultrasonic velocity maps directly correlate with the crack growth and reveal the complex structure of the displacement field at the tip and around the fracture.
+In particular, Fig.~\ref{fig3}(d, left) suggests that the fracture is initiated at the inner cylinder at $r=0$ and the presence of large positive and negative velocities in the vicinity of the crack tip is indicative of strong compression and recoil of the gel matrix.
+Finally, the fracture length $\ell(t)$ is observed to grow logarithmically with $(\tau_f-t)$ upon approaching $\tau_f$ [Fig~\ref{fig4}(d)].
+In other words one has ${\rm d}\ell/{\rm d}t\sim\gp(t)\sim (\tau_f-t)^{-1}$, which indicates that the global shear rate is essentially governed by fracture-induced displacements.
 
@@
 \centering
-\includegraphics[width=18cm,clip]{Fig3.eps}
-\caption{(color online) Fractures along the vorticity direction.
-Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
-Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images, whose position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$.
-Experiment performed on a 4\%~wt.
-casein gel seeded with 3\%~wt.
-polyamide spheres and acidified with 1\%~wt.
-GDL under $\sigma=300$~Pa.
-The gap width is 2~mm.
-See also Suppl.
-Movie~1.
-\label{fig3}}
-\end{figure*} 
-%%%%%%%%%%%%
-
-%%%%%%%%%%%%
-\begin{figure*}
-\centering
-\includegraphics[width=18cm,clip]{Fig4.eps}
-\caption{(color online) Local insights into creep and rupture.
-(a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ (black $\bullet$), $1.7\,10^{-2}$ (blue $\blacktriangledown$), and 0.15 (red $\blacksquare$).
+\includegraphics[width=17.5cm,clip]{Fig4.eps}
+\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ ($\bullet$), $1.7\,10^{-2}$ (\textcolor{blue}{$\blacktriangledown$}), and 0.15 (\textcolor{red}{$\blacksquare$}).
 Solid lines are linear profiles.
@@
 (d)~Standard deviation $\delta_z v(t)$ of $\langle v(r,z,t)\rangle_r$ taken over the vertical direction $z$ (thick black line) together with corresponding standard deviation $\delta_r v(t)$ computed over the radial direction $r$ on the $z$-average $\langle v(r,z,t)\rangle_z$ (thin red line).
-(e)~Fracture length $\ell(t)$ normalized by the height $H$ of the Couette cell as a function of $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$).
+(e)~Fracture length $\ell(t)$ vs $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$) and normalized by the height $H$ of the Couette cell.
 Gray dots show the visualization data for the longest fracture which leads to the failure of the sample at $\tau_f$.
-Red lines are the best fits $\ell(t)=\alpha+\beta\log(1-t/\tau_f)$ to the visualization data.
-Same experiment as in Fig.~\ref{fig3}.
-See also Suppl.
-Movie~1.
+Red lines are the best fits $\ell(t)=a+b\log(1-t/\tau_f)$ to the visualization data.
 \label{fig4}}
@@
 
+Let us now summarize and discuss the most prominent results of this Letter.
+First we have shown that casein gels display a remarkable failure scenario similar to that of brittle solids and characterized by the same three successive creep regimes.
+Here, the primary creep and its power-law exponent [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] are fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
+Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Jaishankar:2013,Gobeaux:2010} but contrasts with the primary creep of solid materials where smaller exponents are found (0.3 $\lesssim \alpha \lesssim 0.7$).
+Besides primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.
+Protein gels rather appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials.
+
+The logarithmic fracture growth in the tertiary creep regime constitutes our second important result.
+Such an evolution is also commonly reported in disordered solid materials displaying brittle rupture (such as paper) and interpreted in the framework of Griffith-like models based on global or local energy barriers \cite{Vanel:2009}.
+However these approaches all predict exponential scalings for $\tau_f(\sigma)$ while our data are best fitted by the decreasing power-law $\tau_f \sim \sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ [inset of Fig.~\ref{fig1}].
+This last key result suggests that thermally activated crack growth is not relevant to the present soft system.
+Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Suresh:2006}.
+Basquin law has also been recently predicted by FBMs that combine elastic fibers with a local yield strain and take into account damage accumulation \cite{Kun:2007,Halasz:2012}.
+Interestingly assuming damage accumulation to be proportional to $\sigma^\gamma$ directly leads to Basquin law with $\beta=\gamma$ and large values of $\gamma$, typically larger than 5 as found here for $\beta$, lead to macroscopic cracks due to the simultaneous rupture of a large number of fibers \cite{Halasz:2012}.
+More generally FBMs under compressive or elongational load predict three successive creep regimes exactly alike Fig.~\ref{fig2} with similar proportionality between $\tau_{\rm min}$ and $\tau_f$ and finite-time singularity \cite{Nechad:2005,Jagla:2011}.
+This highlights the relevance of FBMs in the context of creep in protein gels, whose microstructure indeed appears to be formed of strands \cite{Kalab:1983,Roefs:1990}, but also urges to study FBMs in shear geometries to check whether they would be able to predict fracture growth as observed here.
+
+Finally, we emphasize that the (reversible) Andrade-like creep and the (irreversible) crack growth are two  physical processes that effectively superimpose to yield the global rheological response.
+Indeed, as seen from the yellow line in Fig.~\ref{fig2}, the master curve $\gp(t)/\gp_{\rm min}$ vs $t/\tau_f$ is perfectly fitted by:
+%the master curve $\tilde{\gp}(t)=\gp(t)/\gp_{\rm min}$ vs $\tilde{t}=t/\tau_f$ is perfectly fitted by:
+\begin{equation}
+\frac{\gp(t)}{\gp_{\rm min}}=\lambda \left(\frac{t}{\tau_f}\right)^{-\alpha}+ \frac{\mu}{1-t/\tau_f}\,,
+%\tilde{\gp}(\tilde{t})=\lambda \tilde{t}^{-\alpha}+ \frac{\mu}{1-\tilde{t}}\,,
+\label{e.fit}
+\end{equation}
+with only two adjustable parameters $\lambda$ and $\mu$ once $\alpha=0.85$ is fixed.
+This allows us to interpret the secondary creep regime as a mere crossover from creep to crack growth.
+To conclude, the present time- and space-resolved study shows that protein gels qualify as model, brittle soft solids.
+The next step consists in a statistical study of the fluctuations associated to crack nucleation and growth as well as a systematic investigation of other systems in order to check for universality in the irreversible creep rupture of soft solids in general and of biogels in particular.
+Such a study is expected to have important implications in understanding the behavior of biomaterials under extreme stress conditions.
+
+%which separates into three successive phases: (i)~a primary (or Andrade-like) creep regime during which the strain field is homogeneous, (ii)~a secondary creep during which fractures nucleate and grow logarithmically, and (iii)~a finite-time singularity concomittant with the sudden failure of the gel.
+The failure time displays a non-trivial power-law scaling with the applied shear-stress that contrasts with the exponential scalings usually reported for solid brittle materials \cite{Vanel:2009}.
+
+% within the spatial resolution of a few microns, the strain field is homogeneous [Fig~\ref{fig4}~(b)].
+If plastic events take place in this primary regime, they occur at smaller scale.
+However, given the agreement between the Andrade exponent and the frequency scaling of the linear elastic modulus discussed above, it is more likely that no or a negligible amount of plastic events take place during this regime, corroborating the optical and ultrasonic measurements.
+
+%Compare to the scalings observed in YSF - in particular microgels display a similar power-law scaling, except that the yield point appears explicitely in the scaling as the critical point in a dynamical phase transition \cite{Divoux:2012}.
+Propose an explanation or at least raise a point.
+\cite{Chaudhuri:2013}
+
+%FBM that rely on elastic fibers with a local yield strain (or stress) and introduce a non-linear Eyring rheology or a local coupling lead to a different scaling law than the one we find here for the failure time with the applied stress.
+In their law appears a ``yield stress" \cite{Kun:2003,Nechad:2005,Kovacs:2008,Jagla:2011} that is not present here.
+
+%different from other soft materials such as transient-network systems \cite{Ligoure:2013}, because truly irreversible.
+The pH is key to the irreversibility and plays the role of an additional control parameter that, when maintained at values lower than the isoelectric point, prevent the system from healing.
+
+%Comparatively, Yield Stress Fluids (YSF) that display a solid-like behavior at rest, either because of a jammed microstructure or attractive interactions \cite{Sciortino:2005,Bonnecaze:2010}, but do not present such external control parameter, display a similar rheological signature during their stress-induced yielding.
+Indeed the S-shaped curved pictured in figure~\ref{fig2}~(a) has been also reported in \cite{Bauer:2006,Gibaud:2010,Divoux:2011,Siebenburger:2012,Grenard:2013}.
+However, their yielding scenario is fundamentally different.
+Local measurements conducted on both type of soft glassy materials have revealed that no fractures take place in the bulk: the microstructure rather reorganizes and flows in a ductile way.
+Our study reveals that acid-induced gels behave elastically up to the rupture point and that fractures develop and grow logarithmically, perpendicularly to the shear direction prior to an irreversible failure which makes casein gels model {\it brittle soft solids}.
+
+%and yield stress fluids (YSFs) \cite{Bauer:2006,Divoux:2011,Siebenburger:2012,Grenard:2013} that usually exhibit a power-law creep with smaller exponents (0.3 $\lesssim \alpha \lesssim 0.7$) which are not directly related to linear rheology.
+Besides, the linear behavior of YSF generally differs from a simple power-law scaling, while in the case of solids
+
 \begin{acknowledgments}
-The authors thank Thomas Gibaud for fruitful discussions and Alan Parker for precious advice and for providing the casein and GDL.
+The authors thank T.~Gibaud, S.~Lindstr\"om and N.~Taberlet for fruitful discussions, M.~Kal\'ab for kindly providing us with unobtainable references, and A.~Parker at Firmenich for precious advice and for providing the casein and GDL.
 This work was funded by the Institut Universitaire de France and by the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement No.~258803.
@@
 
-\bibliography{../../bibseb} 
-%\bibliographystyle{jfm}
+\begin{thebibliography}{}
+
+\bibitem{Viovy:2000} J.L. Viovy, Rev. Mod. Phys. {\bf 72}, 813--872 (2000); Dorfman et al., Chem. Rev. {\bf 113}, 2584--2667 (2013).
+
+\bibitem{Stricker:2010} J. Stricker, T. Falzone and M.L. Gardel, J. Biomech. {\bf 43},  9--14 (2010); F. Grinnell and W.M. Petroll, Annu. Rev. Cell Dev. Biol. {\bf 26}, 335--361 (2010).
+
+\bibitem{Mezzenga:2005} R. Mezzenga, P. Schurtenberger, A. Burbidge and M. Michel, Nature Materials  {\bf 4}, 729--740 (2005).
+
+\bibitem{Gardel:2004} M.L. Gardel et al., Science {\bf 304}, 1301--1305 (2004); C. Storm et al., Nature {\bf 435}, 191--194 (2005)
+
+\bibitem{Bonn:1998} D. Bonn et al., Science {\bf 280}, 265--267 (1998).
+
+\bibitem{Baumberger:2006} T. Baumberger, C. Caroli and D. Martina, Nature Materials {\bf 5}, 552-555 (2006).
+
+\bibitem{Daniels:2007} K. Daniels et al., Phys. Rev. Lett. {\bf 99}, 124501 (2007).
+
+%\bibitem{Brenner:2013} T. Brenner et al., J. Non-Newtonian fluid Mech. {\bf 196}, 1--7 (2013).
+
+\bibitem{Cloitre:2000} M.
+Cloitre, R.
+Borrega and L.
+Leibler, Phys.
+Rev.
+Lett.
+{\bf 85}, 4819--4822 (2000); V.
+Viasnoff and F.
+Lequeux, Phys.
+Rev.
+Lett.
+{\bf 89}, 065701 (2002); Coussot et al., Phys.
+Rev.
+Lett.
+{\bf 88}, 175501 (2002).
+
+\bibitem{Divoux:2012} T. Divoux et al., Soft Matter {\bf 8}, 4151--4164 (2012).
+
+\bibitem{Siebenburger:2012} Siebenb\"urger, M. Ballauff and T. Voigtmann, Phys. Rev.
+Lett., {\bf 108}, 255701 (2012)
+
+\bibitem{Dickinson:2006} E. Dickinson, Soft Matter {\bf 2}, 642--652 (2006).
+
+\bibitem{Gibaud:2012} T. Gibaud et al., Faraday Discuss. {\bf 158}, 267--284 (2012).
+
+\bibitem{Higgs:1990} P.G. Higgs and S.B. Ross-Murphy, Int. J. Biol. Macromol. {\bf 12}, 233--240 (1990).
+
+\bibitem{vanVliet:1995} T.
+van Vliet and P.
+Walstra, Faraday Discuss.
+{\bf 101}, 359--370 (1995); T.
+van Vliet, H.J.M.
+van Dijk, P.
+Zon and P.
+Walstra, Colloid Polym.
+Sci.
+{\bf 269}, 620--627 (1991); L.G.B.
+Bremer et al., Colloids Surf.
+{\bf 51}, 159--170 (1990).
+
+\bibitem{Lucey:1998} J.A. Lucey and H. Singh, Food Research Int. {\bf 30}, 529--542 (1998).
+
+\bibitem{note} Rheological data shown in Figs.~\ref{fig1} and \ref{fig2} and in Supplemental Figs.~1 and 2 were obtained with an MCR 301 rheometer (Anton Paar) in a Couette cell of height 28~mm and gap 1~mm with an inner rotating cylinder of radius 24~mm.
+Optical and ultrasonic imaging coupled to rheometry (Figs.~\ref{fig3} and \ref{fig4}) were performed on an AR G2 rheometer (TA Instruments) in a Couette cell of height 60~mm and gap 2~mm with an inner rotating cylinder of radius 23~mm.
+A detailed, quantitative study of the influence of the gap width on the failure dynamics is left for future work.
+
+\bibitem{Gallot:2013} T. Gallot et al., Rev. Sci. Instrum. {\bf 84}, 045107 (2013).
+
+\bibitem{Andrade:1910} E.N. Da C. Andrade Proc. R. Soc. London A {\bf 84}, 1--12 (1910).
+
+\bibitem{Miguel:2002} M.C. Miguel, A. Vespignani, M. Zaiser and S. Zapperi, Phys. Rev. Lett. {\bf 89}, 165501 (2002); M.C. Miguel, L. Laurson and M.J. Alava, Eur. Phys. J. B {\bf 64}, 443--450 (2008).
+
+\bibitem{Nechad:2005} H. Nechad et al., Phys. Rev. Lett. {\bf 94}, 045501 (2005); J. Mech. Phys. Solids {\bf 53}, 1099-1127 (2005).
+
+\bibitem{Jaishankar:2013} A. Jaishankar and G.H. McKinley, Proc. R. Soc. A {\bf 469}, 20120284 (2013).
+
+\bibitem{Kovacs:2008} K. Kov\'acs et al., Phys. Rev. E \textbf{77}, 036102 (2008).
+
+\bibitem{Jagla:2011} E.A. Jagla, Phys. Rev. E {\bf 83}, 046119 (2011).
+
+\bibitem{Vanel:2009} L. Vanel et al., J. Phys. D: Appl. Phys. {\bf 42}, 214007 (2009).
+
+\bibitem{Kun:2007} F. Kun et al., J. Stat. Mech. P02003 (2007); Phys. Rev. Lett. {\bf 100}, 094301 (2008); J. Stat. Mech. P01021 (2009)
+
+\bibitem{Sethna:2001} J.P. Sethna, K.A. Dahmen and C.R. Myers, Nature {\bf 410}, 242--250 (2001).
+
+\bibitem{Gobeaux:2010} F. Gobeaux et al., Soft Matter {\bf 6}, 3769--3777 (2010).
+
+\bibitem{Rosti:2010} J. Rosti, J. Koivisto, L. Laurson, M.J. Alava, Phys. Rev. Lett. {\bf 105}, 100601 (2010).
+
+\bibitem{Basquin:1910} O.H. Basquin, Proceedings of the American Society for Testing and Materials ASTEA, {\bf 10}, 625--630 (1910).
+
+\bibitem{Suresh:2006} S. Suresh, {\it Fatigue of Materials} (Cambridge University Press, Cambridge, England, 2006)
+
+\bibitem{Halasz:2012} Z. Hal\'asz, Z. Danku and F. Kun, Phys. Rev. E {\bf 85}, 016116 (2012).
+
+\bibitem{Roefs:1990} S.P.F.M. Roefs, A.E.A. De Groot-Mostert and T. Van Vliet, Colloids Surf. {\bf 50}, 141--159 (1990).
+
+\bibitem{Kalab:1983} M. Kal\'ab, P. Allan-Wojtas and B.E. Phipps-Todd, Food Microstruct. {\bf 2}, 51--66 (1983).
+
+%\bibitem{vanVliet:1991} T. van Vliet, H.J.M. van Dijk, P. Zon and P. Walstra, Colloid Polym. Sci. {\bf 269}, 620--627 (1991).
+
+%\bibitem{Bremer:1990} L.G.B. Bremer et al., Colloids Surf. {\bf 51}, 159--170 (1990).
+
+%\bibitem{Grenard:2013} V. Grenard, T. Divoux, N. Taberlet and S. Manneville, Soft Matter (2013) - DOI: 10.1039/C3SM52548A
+
+%\bibitem{Sciortino:2005} F. Sciortino and P. Tartaglia, Advances in Physics {\bf 54}, 471--524 (2005).
+
+%\bibitem{Bonnecaze:2010} R.T. Bonnecaze and M. Cloitre, Adv. Polym. Sci. {\bf 236}, 117--161 (2010).
+
+%\bibitem{Bauer:2006} T. Bauer, J. Oberdisse and L. Ramos, Phys. Rev. Lett. {\bf 97}, 258303 (2006).
+
+%\bibitem{Divoux:2011} T. Divoux, C. Barentin and S. Manneville, Soft Matter {\bf 7}, 8409--8418 (2011).
+
+%\bibitem{Gibaud:2010} T. Gibaud, D. Frelat and S. Manneville, Soft Matter {\bf 6}, 3482--3488 (2010).
+
+%\bibitem{Sprakel:2011} J. Sprakel et al., Phys. Rev. Lett. {\bf 106}, 248303 (2011).
+
+%\bibitem{Kun:2003} F. Kun et al., Phys. Rev. E {\bf 67}, 061802 (2003).
+
+%\bibitem{Ghofraniha:2012} N. Ghofraniha et al., Soft Matter {\bf 8}, 6214 (2012).
+
+%\bibitem{Ligoure:2013} C. Ligoure and S. Mora, Rheol. Acta {\bf 52}, 91-114 (2013).
+
+%\bibitem{Roefs:1986} S.P.F.M.
+Roefs, {\it Structure of acid casein gels.
+A study of gels formed after acidification in the cold.
+PhD Thesis}, Wageningen Agricultural University, The Netherlands (1986).
+
+%\bibitem{Schall:2010} P. Schall and M. van Hecke, Annu. Rev. Fluid Mech. {\bf 42} 67--88 (2010).
+
+%\bibitem{Chaudhuri:2013} P. Chaudhuri and J. Horbach, Phys. Rev. E {\bf 88}, 040301(R) (2013).
+
+\end{thebibliography}
 
@@
 
-Supplemental Movie~1 shows the creep experiment analyzed in Figs.~\ref{fig3} and \ref{fig4} and performed under $\sigma=300$~Pa on a 4\%~wt.
+Supplemental Movie~1 shows the failure of a 4\%~wt.
+casein gel acidified with 1\%~wt.
+GDL in a plate-plate geometry for an imposed shear stress $\sigma=XXX$~Pa.
+The plate diameter is 40XXX~mm and the gap width is 1XXX~mm.
+Fractures grow parallel to the vorticity direction, i.e.
+along the radial direction in this case.
+The global rheological response is fully similar to that measured in the Couette geometry.
+
+Supplemental Movie~2 shows the creep experiment analyzed in Figs.~\ref{fig3} and \ref{fig4} and performed under $\sigma=300$~Pa on a 4\%~wt.
 casein gel seeded with 3\%~wt.
@@
 The gap width of the Couette cell is 2~mm and its height is 60~mm.
-Images recorded by a standard webcam are shown on the top left.
-Velocity maps $v(r,z,t)$ inferred from ultrasonic imaging by averaging over 4~s are shown on the top right using linear color levels.
+Images recorded by a standard webcam (Logitech Webcam Pro 9000) are shown in the top left panel.
+Velocity maps $v(r,z,t)$ inferred from ultrasonic imaging by averaging over 4~s are shown in the top right panel using linear color levels.
 The vertical position $z=0$ on the ultrasonic images corresponds to about 20~mm below the top of the Couette cell.
-The two bottom graphs show the shear rate response $\gp(t)$ (left) and the strain response $\gamma(t)$.
-
-Supplemental Movie~2 shows the failure of a 4\%~wt.
-casein gel acidified with 1\%~wt.
-GDL in a plate-plate geometry for an imposed shear stress $\sigma=XXX$~Pa.
-The plate diameter is 40XXX~mm and the gap width is 1XXX~mm.
-In this case fractures grow parallel to the vorticitty direction, i.e.
-along the radial direction.
-The shear rate response is fully similar to that in the Couette geometry.
+The two bottom graphs show the global shear rate response $\gp(t)$ (left) and the strain response $\gamma(t)$ recorded by the rheometer (AR G2, TA Instruments) simultaneously to optical and ultrasonic imaging.
 
@@
 
+Supplemental Figure~\ref{suppfig1} illustrates the gelation process of a 4\%.
+sodium caseinate solution seeded with 3\% wt.
+polyamide spheres through time-resolved measurements of the elastic and viscous moduli, respectively $G'$ and $G''$, under small amplitude oscillatory shear.
+The glucono-$\delta$-lactone contains an ester function which, once added to the solution at time $t=0$, hydrolyzes spontaneously and leads to a slow decrease of the pH [Fig.~\ref{suppfig1}(a)] towards the casein isoelectric point (${\rm pH}\simeq 4.6$) at which casein particles aggregate.
+Gelation actually starts earlier at ${\rm pH}\simeq 5$ as shown by gray dashed lines and as already discussed in Ref.~\cite{Lucey:1998supp}.
+At this point the elastic and viscous moduli display a sudden increase then overshoot and converge towards their steady-state values where $G'\gg G''$ [Fig.~\ref{suppfig1}(b)].
+Note that the maximum of the overshoot is reached around the isoelectric point.
+The decrease of both moduli at lower pH is usually attributed to the over-acidification which enhances the repulsive electrostatic interactions between caseinate particles of net positive charge \cite{Dickinson:2002supp}.
+
+Supplemental Figure~\ref{suppfig2} shows the shear rate $\gp(t)$ obtained by differentiating the raw strain of Fig.~\ref{fig1} in the main text.
+For $0.1\lesssim t/\tau_f\lesssim 0.9$, i.e.
+in the secondary creep regime, $\gp(t)$ is smoothed using a moving average over $\delta t\simeq 0.01\tau_f$ in order to remove high-frequency noise due to differentiation at very small shear rates.
+The same data normalized by its minimum value $\gp_{\rm min}$ is plotted in Fig.~\ref{fig2}(a) as a function of $t/\tau_f$ leading to a nice collapse onto a single curve whatever the imposed shear stress.
+
  %%%%%%%%%%%%
-\begin{figure}[h]
+\begin{figure}[th]
 \centering
@@
 (a)~pH vs time $t$.
-(b)~Linear viscoelastic moduli $G'$ (top, in blue) and $G''$ (bottom, in red) recorded as a function of time $t$ for fixed frequency $f=1$~Hz and strain amplitude 0.1XXX~\%.
+(b)~Linear viscoelastic moduli $G'$ (top, in blue) and $G''$ (bottom, in red) recorded as a function of time $t$ for fixed frequency $f=1$~Hz and strain amplitude 0.1~\%.
 Dashed lines indicate that the system starts to build significant elastic properties when the pH decreases below about 5.
@@
 casein gel acidified with 1\%~wt.
-GDL for an imposed shear stress $\sigma=200$ ($\bullet$), 300 ($\blacktriangledown$), 400 ($\blacksquare$), 550 ($\blacktriangle$), and 1000~Pa ($\blacklozenge$) from right to left.
+GDL for an imposed shear stress $\sigma=200$ (\textcolor{red!25!black}{$\bullet$}), 300 (\textcolor{red!50!black}{$\blacktriangledown$}), 400 (\textcolor{red!75!black}{$\blacksquare$}), 550 (\textcolor{red}{$\blacktriangle$}), and 1000~Pa (\textcolor{orange!50!red}{$\blacklozenge$}) from right to left.
 The gap width is 1~mm.
 The yellow line shows the power-law behavior $\gp(t)\sim t^{-0.85}$.
-Same experiments as in Figs.~\ref{fig1} and \ref{fig2} of the main text.
 \label{suppfig2}}
@@
 
+\begin{thebibliography}{}
+
+\bibitem{Lucey:1998supp} J.A. Lucey and H. Singh, Food Res. Int. A {\bf 30}, 529--542 (1998).
+
+\bibitem{Dickinson:2002supp} E. Dickinson and L.M. Merino, Food Hydrocolloid {\bf 16}, 321--331 (2002).
+
+\end{thebibliography}
+
 \end{document}

### 2014-01-27 plis_figures.tex
scale z in the computation of the area
Label regimes x axes at bottom only

@@
 	\begin{groupplot}[%
-		group style={group name=g, group size=1 by 2},
+		group style={
+			group name=g, group size=1 by 2,
+			xticklabels at=edge bottom,
+			vertical sep=0,
+			},
 		xmin=0, xmax=200, xtick={20,40,...,180},
-		xlabel={time (min)}, 
 		extra tick style={grid=major},%
@@
 		]
-	\nextgroupplot[ylabel={Relative volume}, ymin=0.3, ymax=1, extra x ticks={12.55}]
+	\nextgroupplot[ylabel={Rel. volume}, ymin=0.3, ymax=1, extra x ticks={12.55}]
 	\addplot+[no marks,Accent1] table{volume_rel_half_cas8.txt};
 	\node[rotate=90, anchor=base] at (axis cs:12.55,0.75) (syn) {synæresis};
-	\node[right=0 of syn.south west] {swelling};
+	\node[above right=0 of syn.south west] {swelling};
 
-	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=4, extra x ticks={26.7, 42.8}]
+	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=1.2, xlabel={time (min)}, extra x ticks={26.7, 42.8}]
 	\addplot+[no marks,Accent1] table{excess_area_pc_half_cas8.txt};
+	\node at (axis cs:13.35,1) {flat};
+	\node[rotate=90, anchor=east] at (axis cs:34.75,1) {wrinkling};
+	\node[anchor=west] at (axis cs:45,0.25) {wrinkling and buckling};
 	\end{groupplot}

### 2014-01-27 plis_figures.tex
Slices in fig:wrinkles

@@
 	\end{groupplot}
+	\matrix[matrix of nodes, matrix anchor=north west, inner sep=0, column sep=0.02\textwidth, row sep=2pt] at (g c1r2.outer south west) (m){
+	Synæresis & Swelling\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t000.png}&
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t023.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t006.png}&
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t049.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t008.png}&
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t051.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t010.png}&
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t053.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t012.png}&
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t078.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t014.png}&
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t088.png}\\
+	};
+	\begin{scope}[->,gray,ultra thick]
+		\draw (m-2-1.east) -- (m-7-1.east) -- (m-2-2.west) -- (m-7-2.west);
+	\end{scope}
 
@@
 	\end{tikzpicture}
-	\caption{Syneresis and swelling (c) Evolution of the volume of the gel phase. (d) Evolution of the midsurface area.}
+	\caption{Synæresis and swelling (c) Evolution of the volume of the gel phase. (d) Evolution of the midsurface area.}
 	\label{fig:wrinkles}

### 2014-01-27 plis_figures.tex
fig:wrinkles numbering subfigures caption

@@
 
-\pgfplotsset{compat=1.3, every axis/.append style={mark size={0.2em}}}
+\pgfplotsset{compat=1.3, every axis/.append style={thick, mark size={0.2em}}}
 
@@
 		extra x tick labels={},%
-		width=\columnwidth,
+		width=1.04\columnwidth,
 		height=0.5\columnwidth,
@@
 	\addplot+[no marks,Accent1] table{volume_rel_half_cas8.txt};
+	\addplot+[only marks,Accent2, mark=+] table{volume_rel_half_cas8_toi.txt};
 	\node[rotate=90, anchor=base] at (axis cs:12.55,0.75) (syn) {synæresis};
 	\node[above right=0 of syn.south west] {swelling};
+	\node[lab,below left=0.2em] at (rel axis cs:1,1) {a};
 
@@
 	\addplot+[no marks,Accent1] table{excess_area_pc_half_cas8.txt};
+	\addplot+[only marks,Accent2, mark=+] table{excess_area_pc_half_cas8_toi.txt};
 	\node at (axis cs:13.35,1) {flat};
@@
 	\node[anchor=west] at (axis cs:45,0.25) {wrinkling and buckling};
+	\node[lab,above left=0.2em] at (rel axis cs:1,0) (b) {b};
 	\end{groupplot}
+
 	\matrix[matrix of nodes, matrix anchor=north west, inner sep=0, column sep=0.02\textwidth, row sep=2pt] at (g c1r2.outer south west) (m){
@@
 	};
-	\begin{scope}[->,gray,ultra thick]
-		\draw (m-2-1.east) -- (m-7-1.east) -- (m-2-2.west) -- (m-7-2.west);
-	\end{scope}
+	\draw[>->,gray,ultra thick] (m-2-1.east) -- (m-7-1.east) -- (m-2-2.west) -- (m-7-2.west);
+	\node[lab,anchor=base east] at (m-1-2.base -| b.east) {c};
 
@@
 	\end{tikzpicture}
-	\caption{Synæresis and swelling (c) Evolution of the volume of the gel phase. (d) Evolution of the midsurface area.}
+	\caption{Synæresis and swelling, cas.
+8\%, GDL 7.2\% (a) Evolution of the volume of the gel phase.
+(b) Evolution of the midsurface area.
+(c) Confocal XZ slices of the cell (field of view is $1\times\SI{0.1}{\milli\metre}$), colors indicate the protein concentration.
+Times of the snapshots are marked as crosses on (a) and (b).
+Note that swelling starts \SI{14}{\minute} before wrinkling.
+The thickness increase is $\approx\SI{0.2}{px}$, thus not directly visible on the pictures.}
 	\label{fig:wrinkles}

### 2014-01-27 plis_figures.tex
fig:acidgel remove yticks because not sure of synchronisation

@@
 		extra y ticks={4.6}, extra y tick labels={},%
-		extra x ticks={0.55, 0.64, 0.72, 0.80, 1, 1.27, 2.5}, extra x tick labels={},%
+		%extra x ticks={0.55, 0.64, 0.72, 0.80, 1, 1.27, 2.5}, extra x tick labels={},%
 		axis y line*=left,

### 2014-01-27 plis_figures.tex
Remarques de Sebastien: Couleur de la courbe de G' Ligne d'isoelectrique sans label Fleches pour indiquer spacer et brush, plus d'épaisseur de brush Volume en % et temps toutes les 30 min

@@
 	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
-	\node[base left=0] at (axis cs:8,4.6) {isoelectric};
+	%\node[base left=0] at (axis cs:8,4.6) {isoelectric};
 	\end{axis}
@@
 		ymin=0,
-		ylabel=\textcolor{Accent2}{$G^\prime$ (\si{\pascal})},
+		ylabel=\textcolor{Main}{$G^\prime$ (\si{\pascal})},
 		axis y line*=right,
@@
 		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
+	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
 	\end{axis}
@@
 	\begin{scope}[shift=(pH.outer south west), yshift=-2em]
-		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\columnwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
+		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\columnwidth,1.5em) node[midway,fill=white,inner sep=1pt] (glabeltop) {glass};
 		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\columnwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
-		\draw[line width=2pt,Accent1] (0.05\columnwidth,-2.5em) -- (0.95\columnwidth,-2.5em) (0.05\columnwidth,-1pt) -- (0.95\columnwidth,-1pt) node[pos=0, below right, inner sep=1pt, text width=0.4\columnwidth] {acrylamide brush ($10\sim \SI{100}{\nano\metre}$)};
-		\fill[gray] (0,0) rectangle (0.05\columnwidth,-2.5em) (\columnwidth,0) rectangle (0.95\columnwidth,-2.5em) node[pos=1, above left] {spacer};
+		%acrylamide
+		\begin{scope}[Accent1]
+			\draw[line width=2pt] (0.05\columnwidth,-2.5em) -- (0.95\columnwidth,-2.5em) (0.05\columnwidth,-1pt) -- (0.95\columnwidth,-1pt);
+			\node[inner sep=1pt, anchor=west] at (0.05\columnwidth,-1.25em) (brushlabel) {acrylamide brush};
+			\draw[->] (brushlabel.base) -- +(0,-0.8em);
+			\draw[->] (brushlabel.north) -- +(0,+0.4em);
+		\end{scope}
+		%spacer
+		\begin{scope}[gray]
+			\fill (0,0) rectangle (0.05\columnwidth,-2.5em) (\columnwidth,0) rectangle (0.95\columnwidth,-2.5em);
+			\node[above left=0] at (\columnwidth,1.5em) (spacerlabel){spacer};
+			\draw[->] (spacerlabel.base) -- (0.975\columnwidth,1pt);
+		\end{scope}
+		%
 		\draw[<->] (0.75\columnwidth,-2pt) -- (0.75\columnwidth,-2.5em) node[midway,left] {$e\sim \SI{100}{\micro\metre}$};
@@
 (a) Evolution of the pH and of the storage modulus with sticky boundary conditions.
+Horizontal line is the isoelectric pH.
 (b) Sketch of the cell with no adhesion of the protein on the top and bottom plates.
@@
 			},
-		xmin=0, xmax=200, xtick={20,40,...,180},
+		xmin=0, xmax=200, xtick={30,60,...,180},
 		extra tick style={grid=major},%
@@
 		]
-	\nextgroupplot[ylabel={Rel. volume}, ymin=0.3, ymax=1, extra x ticks={12.55}]
-	\addplot+[no marks,Accent1] table{volume_rel_half_cas8.txt};
-	\addplot+[only marks,Accent2, mark=+] table{volume_rel_half_cas8_toi.txt};
-	\node[rotate=90, anchor=base] at (axis cs:12.55,0.75) (syn) {synæresis};
+	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, extra x ticks={12.55}]
+	\addplot+[no marks,Accent1] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8.txt};
+	\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
+	\node[rotate=90, anchor=base east] at (axis cs:12.55,100) (syn) {synæresis};
 	\node[above right=0 of syn.south west] {swelling};

### 2014-01-28 plis_figures.tex
Color primary pattern
crop, enhance contrast and zoom pictures in fig:acidgel

@@
 			\node[inner sep=1pt, anchor=west] at (0.05\columnwidth,-1.25em) (brushlabel) {acrylamide brush};
-			\draw[->] (brushlabel.base) -- +(0,-0.8em);
-			\draw[->] (brushlabel.north) -- +(0,+0.4em);
+			\draw[->,thick] (brushlabel.base) -- +(0,-0.8em);
+			\draw[->,thick] (brushlabel.north) -- +(0,+0.4em);
 		\end{scope}
@@
 			\node[above left=0] at (\columnwidth,1.5em) (spacerlabel){spacer};
-			\draw[->] (spacerlabel.base) -- (0.975\columnwidth,1pt);
+			\draw[->,thick] (spacerlabel.base) -- (0.975\columnwidth,1pt);
 		\end{scope}
@@
 	\end{scope}
-	\begin{scope}[inner sep=0, shift=(pH.outer south west), yshift=-10.5em]
-		\node[anchor=west] (a) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2}};
-		\node at (0.5\columnwidth,0) (b) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}};
-		\node[anchor=east] at (\columnwidth,0) (c) {\includegraphics[width=0.3\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
-		\draw[Accent2] (a.north west) ++(0.126\columnwidth, -0.086\columnwidth) rectangle +(0.1\columnwidth,-0.075\columnwidth) -- (b.south west) (a.north west) ++(0.226\columnwidth, -0.086\columnwidth) -- (b.north west);
-		\draw[Main] (b.north west) ++(0.117\columnwidth,-0.117\columnwidth) rectangle +(0.093\columnwidth,-0.068\columnwidth) -- (c.south west) (b.north west) ++(0.211\columnwidth,-0.117\columnwidth) --(c.north west);
-		\draw[Accent1] (c.north west) ++(0.094\columnwidth,0) rectangle +(0.117\columnwidth,-0.117\columnwidth);
-		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.106\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
-		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.158\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
-		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.099\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
+	%pictures
+	\begin{scope}[inner sep=0, shift=(pH.outer south west), yshift=-8em]
+		\node[anchor=north west] (a) {\includegraphics[width=0.28\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2_crop}};
+		\node[anchor=north] at (0.44\columnwidth,0) (b) {\includegraphics[width=0.28\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6_crop}};
+		\node[anchor=north east] at (\columnwidth,0) (c) {\includegraphics[width=0.4\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
+		%zooms
+		\draw[Accent2] (a.north west) ++(0.111\columnwidth, -0.072\columnwidth) rectangle +(0.137\columnwidth,-0.137\columnwidth) -- (b.south west) (a.north west) ++(0.248\columnwidth, -0.072\columnwidth) -- (b.north west);
+		\draw[Main] (b.north west) ++(0.097\columnwidth,-0.143\columnwidth) rectangle +(0.128\columnwidth,-0.081\columnwidth) -- (c.south west) (b.north west) ++(0.225\columnwidth,-0.143\columnwidth) --(c.north west);
+		\draw[Accent1] (c.north west) ++(0.125\columnwidth,0) rectangle +(0.156\columnwidth,-0.156\columnwidth);
+		%scale bars
+		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.178\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
+		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.176\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
+		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
 	\node[lab] at (0,0 |- M.base) {c};
@@
 	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
-	\includegraphics[width=0.13\textwidth]{prise_0100_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0130_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0160_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0190_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0250_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0360_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0799_resized.jpg}\\
+	\includegraphics[width=0.13\textwidth]{prise_0100_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0130_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0160_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0190_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0250_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0360_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0799_color.jpg}\\
 	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
@@
 			},
-		xmin=0, xmax=200, xtick={30,60,...,180},
+		xmin=0, xmax=200, xtick={0,30,...,180},
 		extra tick style={grid=major},%

### 2014-01-29 plis_figures.tex
Disposition of fig:thickness, more space for pictures, no text but symbols

@@
 \usepackage[utf8]{inputenc}
+\usepackage{amssymb}
 
@@
 	\begin{tikzpicture}
+	\node[inner sep=0, anchor=north] (x800) {\includegraphics[height=0.5\columnwidth]{pattern_800um_x1.jpg}};
+
+	\matrix[matrix of nodes, inner sep=0, matrix anchor=south, column sep=0, row sep=0.5em] at ($(x800.north)+(0,0.5em)$) (m) {		
+		\includegraphics[height=0.25\columnwidth, angle=90, trim=512px 0 0 0, clip=true]{pattern_50um.jpg} & 
+		\includegraphics[height=0.25\columnwidth, angle=90, trim=512px 0 0 0, clip=true]{pattern_100um.jpg} &
+		\includegraphics[height=0.25\columnwidth,angle=90]{pattern_250um.jpg} & 
+		\includegraphics[height=0.25\columnwidth,angle=90]{pattern_450um.jpg} \\
+	};
+
+	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.376\columnwidth, 0.09\columnwidth) -- ++(0.278\columnwidth,0) node[midway,above]{$\lambda$};
+	\draw[ultra thick](m-1-1.north west) ++(0,0.5em) -- ++(0.167\columnwidth,0) node[midway, above] {\SI{2}{\milli\metre}};
+	%marks
+	\begin{scope}[anchor=north west,Accent1,font=\Large]
+		\node[font=\huge] at (m-1-1.north west) {$\bullet$};
+		\node at (m-1-2.north west) {$\blacktriangle$};
+		\node at (m-1-3.north west) {$\blacksquare$};
+		\node at (m-1-4.north west) {$\blacklozenge$};
+		\node at (x800.north west) {$\bigstar$}; 
+	\end{scope}
+
 	\begin{axis}[%
 		name=g,
-		width=\columnwidth,
-		height=0.4\columnwidth,
+		anchor=outer north west,
+		at={(m.north west)},
+		width=0.375\columnwidth,
+		height=0.2\columnwidth,
 		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
-		xtick={0.1,0.3,...,0.9},
+		xtick={0,0.4,...,0.9},
 		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		label shift={-0.5em},
+		scale only axis,
 		]
-		\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+		\addplot+[only marks,Accent1, mark options={fill=Accent1}, mark=*] coordinates {(0.05,0.261)};
+		 \addplot+[only marks,Accent1, mark options={fill=Accent1}, mark=triangle*] coordinates {(0.1,0.458)}; 
+		 \addplot+[only marks,Accent1, mark options={fill=Accent1}, mark=square*] coordinates {(0.250,0.916)};
+		 \addplot+[only marks,Accent1, mark options={fill=Accent1}, mark=diamond*] coordinates {(0.450,2.075)};
+		 \addplot+[only marks,Accent1, mark options={fill=Accent1}, mark=text, text mark={$\bigstar$}] coordinates {(0.800,3.333)};
 		\addplot+[domain=0:3, no marks, Accent1, forget plot] {4.23*x};
 	\end{axis}
-
-	\matrix[matrix of nodes, inner sep=0, matrix anchor=south west, column sep=0.033\columnwidth, row sep=0.5em] at ($(g.outer north west)+(0,0.5em)$) (m) {
-		\SI{50}{\micro\metre} & \SI{100}{\micro\metre}\\
-		\includegraphics[width=0.225\columnwidth]{pattern_50um.jpg} &
-		\includegraphics[width=0.225\columnwidth]{pattern_100um.jpg}\\
-		\SI{250}{\micro\metre} & \SI{450}{\micro\metre}\\
-		\includegraphics[width=0.225\columnwidth]{pattern_250um.jpg} &
-		\includegraphics[width=0.225\columnwidth]{pattern_450um.jpg}\\
-	};
-	\node[inner sep=0, below left] at ($(m-2-1.north west) + (\columnwidth,0)$) (x800) {\includegraphics[width=0.45\columnwidth]{pattern_800um_x1.jpg}};
-	\node[anchor=base] at (x800.north |- m-1-2.base) {\SI{800}{\micro\metre}};
-	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.18\columnwidth, 0.05\columnwidth) -- ++(0.133\columnwidth,0) node[midway,above]{$\lambda$};
-	\draw[ultra thick](x800.south east) ++(0,-0.5em) -- ++(-0.2\columnwidth,0) node[midway, below] {\SI{5}{\milli\metre}};
 	\end{tikzpicture}

### 2014-01-29 plis_figures.tex
Reorganise fig:thickness in line + graph + x800. Markers to indicate respective thicknesses

--- 2014-01-28 plis_figures.tex
@@
 
+\usepackage{amssymb}
 \usepackage{siunitx}
@@
 	\begin{tikzpicture}
+	\matrix[matrix of nodes, inner sep=0, column sep=0.01\columnwidth, row sep=0.01\columnwidth] (m) {
+		\includegraphics[width=0.245\columnwidth]{pattern_50um.jpg}&
+		\includegraphics[width=0.245\columnwidth]{pattern_100um.jpg}&
+		\includegraphics[width=0.245\columnwidth]{pattern_250um.jpg}&
+		\includegraphics[width=0.245\columnwidth]{pattern_450um.jpg}\\
+	};
+	\node[inner sep=0, below left=0.01\columnwidth and 0 of m.south east] (x800) {\includegraphics[width=0.49\columnwidth]{pattern_800um_x1.jpg}};
+	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.196\columnwidth, 0.05\columnwidth) -- ++(0.145\columnwidth,0) node[midway,above]{$\lambda$};
+	\draw[ultra thick,Main](x800.north east) -- ++(-0.217\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};
+
 	\begin{axis}[%
 		name=g,
-		width=\columnwidth,
-		height=0.4\columnwidth,
+		anchor=left of north west,
+		at={($(m.south west)+(0,-0.01\columnwidth)$)},
+		scale only axis,
+		width=0.4\columnwidth,
+		height=0.175\columnwidth,
 		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
-		xtick={0.1,0.3,...,0.9},
+		xtick={0,0.4,...,0.9},
 		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		label shift=-0.5em,
+		cycle multi list={%
+			Accent1,no marks\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=*\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=triangle*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=triangle*\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=square*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=square*\\%
+		}
 		]
-		\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
-		\addplot+[domain=0:3, no marks, Accent1, forget plot] {4.23*x};
+		\addplot+[domain=0:3] {4.23*x};
+		\addplot coordinates {(0.05,0.261)};\label{pgfplots:50}
+		\addplot coordinates {(0.1,0.458)};\label{pgfplots:100}
+		\addplot coordinates {(0.250,0.916)};\label{pgfplots:250}
+		\addplot coordinates {(0.450,2.075)};\label{pgfplots:450}
+		\addplot coordinates {(0.800,3.333)};\label{pgfplots:800}
+		%\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
 	\end{axis}
-
-	\matrix[matrix of nodes, inner sep=0, matrix anchor=south west, column sep=0.033\columnwidth, row sep=0.5em] at ($(g.outer north west)+(0,0.5em)$) (m) {
-		\SI{50}{\micro\metre} & \SI{100}{\micro\metre}\\
-		\includegraphics[width=0.225\columnwidth]{pattern_50um.jpg} &
-		\includegraphics[width=0.225\columnwidth]{pattern_100um.jpg}\\
-		\SI{250}{\micro\metre} & \SI{450}{\micro\metre}\\
-		\includegraphics[width=0.225\columnwidth]{pattern_250um.jpg} &
-		\includegraphics[width=0.225\columnwidth]{pattern_450um.jpg}\\
-	};
-	\node[inner sep=0, below left] at ($(m-2-1.north west) + (\columnwidth,0)$) (x800) {\includegraphics[width=0.45\columnwidth]{pattern_800um_x1.jpg}};
-	\node[anchor=base] at (x800.north |- m-1-2.base) {\SI{800}{\micro\metre}};
-	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.18\columnwidth, 0.05\columnwidth) -- ++(0.133\columnwidth,0) node[midway,above]{$\lambda$};
-	\draw[ultra thick](x800.south east) ++(0,-0.5em) -- ++(-0.2\columnwidth,0) node[midway, below] {\SI{5}{\milli\metre}};
+	\node[below right=0 of m-1-1.north west] {\ref{pgfplots:50}};
+	\node[below right=0 of m-1-2.north west] {\ref{pgfplots:100}};
+	\node[below right=0 of m-1-3.north west] {\ref{pgfplots:250}};
+	\node[below right=0 of m-1-4.north west] {\ref{pgfplots:450}};
+	\node[below right=0 of x800.north west] {\ref{pgfplots:800}};
 	\end{tikzpicture}
-	\caption{Cell thickness dependence of the patterns. Scale bar is common for all 5 transmission microscopy pictures.}
+	\caption{Cell thickness dependence of the patterns. Each transmission microscopy picture corresponds to a point on the graph. Scale bar is common for all pictures.}
 	\label{fig:thickness}

### 2014-01-31 creep/creep_yaourt.tex
Intro: Single equation, fundamental issues remain unexplored no slippage at the Plexiglas walls Monkman-Grant relation no self healing Re-work conclusion SI: sigma and diameter of plane-plane

@@
 Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is typical of brittle solids: after a power-law creep regime fully accounted for by linear viscoelasticity and homogeneous deformation, fractures nucleate and grow logarithmically perpendicularly to shear up to sudden rupture.
+A single equation accounting for those two successive processes nicely captures the full rheological response.
 The failure time surprisingly follows a decreasing power-law with the applied shear stress strongly reminiscent of the Basquin law of fatigue for solids.
@@
 So far huge effort has been devoted to designing protein gels with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}.
-However, their mechanical behavior under a shear stress deep into the nonlinear regime has only been partially addressed \cite{Higgs:1990,vanVliet:1995} and several simple issues remain unexplored such as the spatially resolved rupture scenario or the physical relevance of the analogy brittle failure in hard solids.
+However, their mechanical behavior under a shear stress deep into the nonlinear regime has only been partially addressed \cite{vanVliet:1995} and several fundamental issues remain unexplored such as the spatially resolved rupture scenario or the physical relevance of the analogy brittle failure in hard solids.
 
@@
 Gelation is complete when the elastic ($G'$) and viscous ($G''$) moduli reach a plateau with $G'\gg G''$.
-A constant stress $\sigma$ is then applied to the sample at time $t=0$ and the subsequent strain response $\gamma(t)$ is monitored.
+A constant stress $\sigma$ is then applied to the sample from time $t=0$ and the subsequent strain response $\gamma(t)$ is monitored.
 Images of the gel are recorded simultaneously to the rheology (Logitech Webcam Pro 9000).
@@
 Interestingly, here, this scaling can be fully accounted for by linear viscoelasticity without invoking any irreversible plasticity.
-Indeed casein gels display a power-law rheology $G'(f)\sim G''(f)\sim f^{0.15}$ [inset of Fig.~\ref{fig2}(a)], which corresponds to a compliance $J(t)\equiv\gamma(t)/\sigma\sim t^{0.15}$ in the linear deformation regime \cite{Jaishankar:2013}.
+Indeed casein gels display a power-law rheology $G'(f)\sim G''(f)\sim f^{0.15}$ [inset of Fig.~\ref{fig2}(a)], which corresponds to a compliance $J(t)\equiv\gamma(t)/\sigma\sim t^{0.15}$ in the linear deformation regime \cite{Tschoegl:1989}.
 This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$ and suggests that primary creep is purely reversible without unrecoverable plastic events.
-Homogeneity and linear response are further confirmed, at least up to the available spatial ($\sim 10~\mu$m) and temporal ($\sim 1$~s) resolutions, by direct visualization which shows no sign of crack or fracture [Fig.~\ref{fig3}(a)] and by ultrasonic imaging which reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)].
+Homogeneity and linear response are further confirmed, at least up to the available spatial ($\sim 10~\mu$m) and temporal ($\sim 1$~s) resolutions, by direct visualization which shows no sign of crack or fracture [Fig.~\ref{fig3}(a)] and by ultrasonic imaging which reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)] and no slippage at the Plexiglas walls.
 
 For $0.1\lesssim t/\tau_f\lesssim 0.9$, $\gp(t)$ departs from power-law behavior and goes through a minimum value at time $\tau_{\rm min}= (0.56\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}(b) and inset] as similarly reported for solid composite materials \cite{Nechad:2005} and fiber-bundles models (FBMs) \cite{Kovacs:2008,Jagla:2011}.
+This linearity between $\tau_{\rm min}$ and $\tau_f$ is also known as the Monkman-Grant relation in the context of metals \cite{Sundararajan:1989} and allows one to ``predict'' the failure time from the intermediate-time response.
 During this {\it secondary creep} regime, regularly-spaced cracks nucleate from the top and bottom edges of the Couette cell and start growing perpendicular to the applied stress [Fig.~\ref{fig3}(b), see also Movie~2 in the Supplemental Material].
@@
 Although fractures have not yet entered the ultrasonic region of interest, velocity maps become heterogeneous along the $z$ direction [Fig.~\ref{fig3}(b, left)] and display intermittent fluctuations [Fig~\ref{fig4}(c,d)].
-The level of these fluctuations, which can also be seen on the global response for the same applied stress [$\blacktriangledown$ in Fig.~\ref{fig2}(b)], is poorly reproducible and stress-dependent.
-We propose that such intermittency results from self-healing microscopic cracks which precede the tip of the fracture.
+The level of these fluctuations, which can also be seen on the global response for the same applied stress [\textcolor{red!50!black}{$\blacktriangledown$} in Fig.~\ref{fig2}(b)], is poorly reproducible and stress-dependent.
+We propose that such intermittency arises from microscopic cracks ahead of the fracture tip.
 Since heterogeneity is much smaller along the radial direction [compare red and black lines in Fig~\ref{fig4}(d)], these ``precursors'' either extend across the full 2-mm gap or induce long-range displacements that span the gap.
@@
 
-Let us now summarize and discuss the most prominent results of this Letter.
-First we have shown that casein gels display a remarkable failure scenario similar to that of brittle solids and characterized by the same three successive creep regimes.
-Here, the primary creep and its power-law exponent [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] are fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
-Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Jaishankar:2013,Gobeaux:2010} but contrasts with the primary creep of solid materials where smaller exponents are found (0.3 $\lesssim \alpha \lesssim 0.7$).
-Besides primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.
-Protein gels rather appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials.
-
-The logarithmic fracture growth in the tertiary creep regime constitutes our second important result.
-Such an evolution is also commonly reported in disordered solid materials displaying brittle rupture (such as paper) and interpreted in the framework of Griffith-like models based on global or local energy barriers \cite{Vanel:2009}.
-However these approaches all predict exponential scalings for $\tau_f(\sigma)$ while our data are best fitted by the decreasing power-law $\tau_f \sim \sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ [inset of Fig.~\ref{fig1}].
-This last key result suggests that thermally activated crack growth is not relevant to the present soft system.
-Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Suresh:2006}.
-Basquin law has also been recently predicted by FBMs that combine elastic fibers with a local yield strain and take into account damage accumulation \cite{Kun:2007,Halasz:2012}.
-Interestingly assuming damage accumulation to be proportional to $\sigma^\gamma$ directly leads to Basquin law with $\beta=\gamma$ and large values of $\gamma$, typically larger than 5 as found here for $\beta$, lead to macroscopic cracks due to the simultaneous rupture of a large number of fibers \cite{Halasz:2012}.
-More generally FBMs under compressive or elongational load predict three successive creep regimes exactly alike Fig.~\ref{fig2} with similar proportionality between $\tau_{\rm min}$ and $\tau_f$ and finite-time singularity \cite{Nechad:2005,Jagla:2011}.
-This highlights the relevance of FBMs in the context of creep in protein gels, whose microstructure indeed appears to be formed of strands \cite{Kalab:1983,Roefs:1990}, but also urges to study FBMs in shear geometries to check whether they would be able to predict fracture growth as observed here.
-
 Finally, we emphasize that the (reversible) Andrade-like creep and the (irreversible) crack growth are two  physical processes that effectively superimpose to yield the global rheological response.
@@
 with only two adjustable parameters $\lambda$ and $\mu$ once $\alpha=0.85$ is fixed.
-This allows us to interpret the secondary creep regime as a mere crossover from creep to crack growth.
+The remarkable collapse of the whole data set to such a simple equation allows us to interpret the secondary creep regime as a mere crossover from creep to crack growth.
+
+Let us now summarize and discuss the most prominent results of this Letter.
+First we have shown that casein gels display a remarkable failure scenario similar to that of brittle solids and characterized by the same three successive creep regimes.
+Here, the primary creep and its power-law exponent [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] are fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
+Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Gobeaux:2010,Jaishankar:2013} but contrasts with the primary creep of solid materials where smaller exponents are found (0.3 $\lesssim \alpha \lesssim 0.7$).
+Besides primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.
+Protein gels rather appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials.
+
+The logarithmic fracture growth in the tertiary creep regime constitutes our second important result.
+Such an evolution is also commonly reported in disordered solid materials displaying brittle rupture and interpreted in the framework of Griffith-like models based on global or local energy barriers \cite{Vanel:2009}.
+However these approaches all predict exponential scalings for $\tau_f(\sigma)$ while our data are best fitted by the decreasing power-law $\tau_f \sim \sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ [inset of Fig.~\ref{fig1}].
+This last key result suggests that thermally activated crack growth is not relevant to the present soft system.
+Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Suresh:2006}.
+Basquin law has also been recently predicted by FBMs that combine elastic fibers with a local yield strain and take into account damage accumulation \cite{Kun:2007,Halasz:2012}.
+Interestingly assuming damage accumulation to be proportional to $\sigma^\gamma$ directly leads to Basquin law with $\beta=\gamma$ and large values of $\gamma$, typically larger than 5 as found here for $\beta$, lead to macroscopic cracks due to the simultaneous rupture of a large number of fibers \cite{Halasz:2012}.
+More generally FBMs under compressive or elongational load predict three successive creep regimes exactly alike Fig.~\ref{fig2} with similar proportionality between $\tau_{\rm min}$ and $\tau_f$ and finite-time singularity \cite{Nechad:2005,Jagla:2011}.
+This highlights the relevance of FBMs in the context of creep in protein gels, whose microstructure indeed appears to be formed of strands \cite{Kalab:1983,Roefs:1990}, but also urges to study FBMs in shear geometries to check whether they would be able to predict fracture growth as observed here.
+
 To conclude, the present time- and space-resolved study shows that protein gels qualify as model, brittle soft solids.
-The next step consists in a statistical study of the fluctuations associated to crack nucleation and growth as well as a systematic investigation of other systems in order to check for universality in the irreversible creep rupture of soft solids in general and of biogels in particular.
+Prompted by the remarkable simplicity of Eq.~(\ref{e.fit}) that encompasses Andrade creep, the Monkman-Grant relation and finite-time singularity within a single equation, future modeling will undoubtedly focus on the microscopic ingredients that may give rise to this expression.
+The next experimental step consists in a statistical study of the fluctuations associated to crack nucleation and growth as well as a systematic investigation of other systems in order to check for universality in the irreversible creep rupture of soft solids in general and of biogels such as actin, alginate or agar gels in particular.
 Such a study is expected to have important implications in understanding the behavior of biomaterials under extreme stress conditions.
@@
 
-\bibitem{Mezzenga:2005} R. Mezzenga, P. Schurtenberger, A. Burbidge and M. Michel, Nature Materials  {\bf 4}, 729--740 (2005).
+\bibitem{Mezzenga:2005} R. Mezzenga, P. Schurtenberger, A. Burbidge and M. Michel, Nature Mater.  {\bf 4}, 729--740 (2005).
 
@@
 
-\bibitem{Baumberger:2006} T. Baumberger, C. Caroli and D. Martina, Nature Materials {\bf 5}, 552-555 (2006).
+\bibitem{Baumberger:2006} T. Baumberger, C. Caroli and D. Martina, Nature Mater. {\bf 5}, 552-555 (2006).
 
@@
 
-\bibitem{Higgs:1990} P.G. Higgs and S.B. Ross-Murphy, Int. J. Biol. Macromol. {\bf 12}, 233--240 (1990).
+%\bibitem{1990} P.G. Higgs and S.B. Ross-Murphy, Int. J. Biol. Macromol. {\bf 12}, 233--240 (1990).
 
@@
 
-\bibitem{Jaishankar:2013} A. Jaishankar and G.H. McKinley, Proc. R. Soc. A {\bf 469}, 20120284 (2013).
+\bibitem{Tschoegl:1989}  N.W. Tschoegl, {\it The Phenomenological Theory of Linear Viscoelastic Behavior} (Springer, Berlin, Germany, 1989).
 
@@
 \bibitem{Jagla:2011} E.A. Jagla, Phys. Rev. E {\bf 83}, 046119 (2011).
+
+\bibitem{Sundararajan:1989} G. Sundararajan, Mater. Sci. Eng. A {\bf 112}, 205-214 (1989).
 
@@
 \bibitem{Gobeaux:2010} F. Gobeaux et al., Soft Matter {\bf 6}, 3769--3777 (2010).
+
+\bibitem{Jaishankar:2013} A. Jaishankar and G.H. McKinley, Proc. R. Soc. A {\bf 469}, 20120284 (2013).
 
@@
 casein gel acidified with 1\%~wt.
-GDL in a plate-plate geometry for an imposed shear stress $\sigma=XXX$~Pa.
-The plate diameter is 40XXX~mm and the gap width is 1XXX~mm.
+GDL in a plate-plate geometry for an imposed shear stress $\sigma=120$~Pa.
+The plate diameter is 50~mm and the gap width is 1~mm.
 Fractures grow parallel to the vorticity direction, i.e.

### 2014-01-31 creep/creep_yaourt.tex
Balance bibliography Andrade not in intro rheology -> creep xp master curve in caption of fig2 nor on system composition global shear rate and fracture growth related, no causality No discussion of smaller exponents and rework the sentences around SI: GDL4% in Plane-plane

@@
 \begin{abstract}
-Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids in that they rupture under an external load.
+Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids and to rupture under an external load.
 Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is typical of brittle solids: after a power-law creep regime fully accounted for by linear viscoelasticity and homogeneous deformation, fractures nucleate and grow logarithmically perpendicularly to shear up to sudden rupture.
 A single equation accounting for those two successive processes nicely captures the full rheological response.
-The failure time surprisingly follows a decreasing power-law with the applied shear stress strongly reminiscent of the Basquin law of fatigue for solids.
+The failure time follows a decreasing power-law with the applied shear stress strongly reminiscent of the Basquin law of fatigue for solids.
 These results are in excellent agreement with recent fiber-bundle models that include damage accumulation on elastic fibers and exemplify protein gels as model brittle soft solids.
@@
 Biogels formed through the self-association of polysaccharide coils, collagen, actin filaments or attractive globular proteins play a major role in biochemistry and microbiology \cite{Viovy:2000}, biological networks and cell mechanics \cite{Stricker:2010} as well as in food science \cite{Mezzenga:2005}.
-These biomaterials all behave as elastic solids under small deformations but display remarkable nonlinear behavior generally featuring stress- or strain-stiffening \cite{Gardel:2004} and fractures prior to irreversible rupture \cite{Bonn:1998,Baumberger:2006,Daniels:2007}.
+These biomaterials all behave as elastic solids under small deformations but display remarkable nonlinear behavior generally featuring stress- or strain-stiffening \cite{Gardel:2004} and fractures prior to irreversible rupture \cite{Bonn:1998,Baumberger:2006}.
 Irreversibility stems from the existence of an external control parameter, e.g.
@@
 So far huge effort has been devoted to designing protein gels with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}.
-However, their mechanical behavior under a shear stress deep into the nonlinear regime has only been partially addressed \cite{vanVliet:1995} and several fundamental issues remain unexplored such as the spatially resolved rupture scenario or the physical relevance of the analogy brittle failure in hard solids.
+However, their mechanical behavior deep into the nonlinear regime has only been partially addressed \cite{vanVliet:1995,Pouzot:2006} and several fundamental issues remain unexplored such as the spatially resolved rupture scenario or the physical relevance of the analogy with brittle failure in hard solids.
 
@@
 
-In this Letter we report on the stress-induced fracture scenario of protein gels by means of rheology coupled to optical and ultrasonic imaging.
-Gels formed reversibly by slow acidification of a sodium caseinate solution \cite{Lucey:1998} display fractures under large strain at fixed low pH values \cite{vanVliet:1995}, which makes them perfect candidates to quantify brittleness in soft solids and tackle the above-mentioned issues.
-We demonstrate that under external load, these casein gels display brittle-like failure that results from two successive physical processes: a reversible primary (or Andrade-like) creep regime characterized by a homogeneous strain field and the irreversible nucleation and growth of fractures leading to gel failure.
+In this Letter we report on stress-induced fracture in protein gels by means of creep experiments coupled to optical and ultrasonic imaging.
+Gels formed by slow acidification of a sodium caseinate solution \cite{Lucey:1998} display fractures under large strain at fixed low pH values \cite{vanVliet:1995}, which makes them perfect candidates to quantify brittleness in soft solids and tackle the above-mentioned issues.
+We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: a reversible primary creep regime characterized by a homogeneous strain field and the irreversible nucleation and fracture growth leading to gel failure.
 Our results are in full agreement with the predictions of some recent fiber-bundle models and hint to universal features of brittle failure common to both soft and hard solids.
@@
 \includegraphics[width=7cm,clip]{Fig2.eps}
-\caption{(color online) Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize each of the three successive regimes.
+\caption{(color online) Normalized shear rate responses $\gp(t)/\gpm$ corresponding to the data of Fig.~\ref{fig1} and plotted so as to emphasize the three successive regimes.
 $\gpm$ is the minimum shear rate reached at $\taum$ (see text and Suppl.
 Fig.~2).
-The yellow line shows the best fit of $\gp(t)$ using Eq.~(\ref{e.fit}) with $\alpha=0.85$, leading to $\lambda=0.378\pm 0.002$ and $\mu=0.187\pm 0.002$.
+The yellow line shows the master curve inferred from fitting $\gp(t)$ by Eq.~(\ref{e.fit}) with $\alpha=0.85$, leading to $\lambda=0.378\pm 0.002$ and $\mu=0.187\pm 0.002$.
 (a)~Primary creep: $\gp(t)/\gpm$ vs $t/\tau_f$ in logarithmic scales.
-Inset: Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 0.1~\%.
+Inset: Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 0.1\%.
 Red lines are power laws $G'\sim G''\sim f^{0.15}$.
@@
 Figures~\ref{fig3} and \ref{fig4} gather the results from {\it local} measurements and are discussed below together with each of the successive regimes inferred from {\it global} data.
-Note that these failure dynamics do not depend on the geometry as checked in a plate-plate geometry [see Movie~1 in the Supplemental Materials].
+Note that these failure dynamics do not qualitatively depend on geometry nor on system composition (see Movie~1 in the Supplemental Materials).
 
@@
 \caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
-Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images, whose position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$.
+Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images.
+Their position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$.
 The gap width is 2~mm and $r$ is the radial distance to the inner rotating cylinder.
@@
 
-For $0.1\lesssim t/\tau_f\lesssim 0.9$, $\gp(t)$ departs from power-law behavior and goes through a minimum value at time $\tau_{\rm min}= (0.56\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}(b) and inset] as similarly reported for solid composite materials \cite{Nechad:2005} and fiber-bundles models (FBMs) \cite{Kovacs:2008,Jagla:2011}.
-This linearity between $\tau_{\rm min}$ and $\tau_f$ is also known as the Monkman-Grant relation in the context of metals \cite{Sundararajan:1989} and allows one to ``predict'' the failure time from the intermediate-time response.
+For $0.1\lesssim t/\tau_f\lesssim 0.9$, $\gp(t)$ departs from power-law behavior and goes through a minimum value at time $\tau_{\rm min}= (0.56\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}(b) and inset] as similarly reported for metals \cite{Sundararajan:1989}, solid composite materials \cite{Nechad:2005} and fiber-bundles models (FBMs) \cite{Kovacs:2008,Jagla:2011}.
+This linearity between $\tau_{\rm min}$ and $\tau_f$, also known as the Monkman-Grant relation \cite{Monkman:1956}, allows one to ``predict'' the failure time from the intermediate-time response.
 During this {\it secondary creep} regime, regularly-spaced cracks nucleate from the top and bottom edges of the Couette cell and start growing perpendicular to the applied stress [Fig.~\ref{fig3}(b), see also Movie~2 in the Supplemental Material].
@@
 Since heterogeneity is much smaller along the radial direction [compare red and black lines in Fig~\ref{fig4}(d)], these ``precursors'' either extend across the full 2-mm gap or induce long-range displacements that span the gap.
-In any case, such noisy features are typical of slow crack growth in solids \cite{Vanel:2009,Kun:2007} and more generally of avalanche-like behavior \cite{Sethna:2001}.
+In any case, such noisy features are typical of slow crack growth in solids \cite{Vanel:2009,Kun:2007} and more generally of avalanche behavior \cite{Sethna:2001}.
 
@@
 Finally, the fracture length $\ell(t)$ is observed to grow logarithmically with $(\tau_f-t)$ upon approaching $\tau_f$ [Fig~\ref{fig4}(d)].
-In other words one has ${\rm d}\ell/{\rm d}t\sim\gp(t)\sim (\tau_f-t)^{-1}$, which indicates that the global shear rate is essentially governed by fracture-induced displacements.
+In other words one has ${\rm d}\ell/{\rm d}t\sim\gp(t)\sim (\tau_f-t)^{-1}$, which indicates that the global shear rate is linked to fracture-induced displacements.
 
@@
 Red lines are the best fits $\ell(t)=a+b\log(1-t/\tau_f)$ to the visualization data.
+Same experiment as in Fig.~\ref{fig3} and Supplemental Movie~2.
 \label{fig4}}
@@
 Here, the primary creep and its power-law exponent [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] are fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
-Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Gobeaux:2010,Jaishankar:2013} but contrasts with the primary creep of solid materials where smaller exponents are found (0.3 $\lesssim \alpha \lesssim 0.7$).
-Besides primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.
-Protein gels rather appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials.
+Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Gobeaux:2010,Jaishankar:2013} as well as hard-sphere-like colloidal glasses \cite{Siebenburger:2012}.
+Protein gels thus appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials.
+By contrast primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.
 
@@
 However these approaches all predict exponential scalings for $\tau_f(\sigma)$ while our data are best fitted by the decreasing power-law $\tau_f \sim \sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ [inset of Fig.~\ref{fig1}].
-This last key result suggests that thermally activated crack growth is not relevant to the present soft system.
-Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Suresh:2006}.
+This last key result suggests that thermally activated crack growth is not relevant to the present soft solid.
+Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Kohout:2000}.
 Basquin law has also been recently predicted by FBMs that combine elastic fibers with a local yield strain and take into account damage accumulation \cite{Kun:2007,Halasz:2012}.
@@
 To conclude, the present time- and space-resolved study shows that protein gels qualify as model, brittle soft solids.
-Prompted by the remarkable simplicity of Eq.~(\ref{e.fit}) that encompasses Andrade creep, the Monkman-Grant relation and finite-time singularity within a single equation, future modeling will undoubtedly focus on the microscopic ingredients that may give rise to this expression.
+Prompted by the remarkable simplicity of Eq.~(\ref{e.fit}) that encompasses Andrade creep, the Monkman-Grant relation and finite-time singularity within a single equation, future modeling will undoubtedly focus on the microscopic ingredients needed to predict quantitatively the Monkman-Grant prefactor and the Basquin exponent $\beta$.
 The next experimental step consists in a statistical study of the fluctuations associated to crack nucleation and growth as well as a systematic investigation of other systems in order to check for universality in the irreversible creep rupture of soft solids in general and of biogels such as actin, alginate or agar gels in particular.
@@
 
-\bibitem{Gardel:2004} M.L. Gardel et al., Science {\bf 304}, 1301--1305 (2004); C. Storm et al., Nature {\bf 435}, 191--194 (2005)
+\bibitem{Gardel:2004} M.L. Gardel et al., Science {\bf 304}, 1301--1305 (2004); C. Storm et al., Nature {\bf 435}, 191--194 (2005).
 
@@
 
-\bibitem{Daniels:2007} K. Daniels et al., Phys. Rev. Lett. {\bf 99}, 124501 (2007).
+%\bibitem{Daniels:2007} K. Daniels et al., Phys. Rev. Lett. {\bf 99}, 124501 (2007).
 
@@
 \bibitem{Siebenburger:2012} Siebenb\"urger, M. Ballauff and T. Voigtmann, Phys. Rev.
-Lett., {\bf 108}, 255701 (2012)
+Lett., {\bf 108}, 255701 (2012).
 
@@
 
-\bibitem{vanVliet:1995} T.
-van Vliet and P.
-Walstra, Faraday Discuss.
-{\bf 101}, 359--370 (1995); T.
-van Vliet, H.J.M.
-van Dijk, P.
-Zon and P.
-Walstra, Colloid Polym.
-Sci.
-{\bf 269}, 620--627 (1991); L.G.B.
-Bremer et al., Colloids Surf.
-{\bf 51}, 159--170 (1990).
+\bibitem{vanVliet:1995} T. van Vliet and P. Walstra, Faraday Discuss. {\bf 101}, 359--370 (1995); L.G.B. Bremer et al., Colloids Surf. {\bf 51}, 159--170 (1990).
+
+\bibitem{Pouzot:2006} M. Pouzot, J. Coll. Interf. Sci. {\bf 293} 376--383 (2006).
 
@@
 
-\bibitem{Andrade:1910} E.N. Da C. Andrade Proc. R. Soc. London A {\bf 84}, 1--12 (1910).
+\bibitem{Andrade:1910} E.N. Da C. Andrade, Proc. R. Soc. London A {\bf 84}, 1--12 (1910).
 
@@
 
+\bibitem{Sundararajan:1989} G. Sundararajan, Mater. Sci. Eng. A {\bf 112}, 205-214 (1989).
+
 \bibitem{Kovacs:2008} K. Kov\'acs et al., Phys. Rev. E \textbf{77}, 036102 (2008).
@@
 
-\bibitem{Sundararajan:1989} G. Sundararajan, Mater. Sci. Eng. A {\bf 112}, 205-214 (1989).
+\bibitem{Monkman:1956} F. C. Monkman and N. J. Grant, Proc. Am. Soc. Test.
+Mater. {\bf 56}, 593-- (1956).
 
@@
 
-\bibitem{Kun:2007} F. Kun et al., J. Stat. Mech. P02003 (2007); Phys. Rev. Lett. {\bf 100}, 094301 (2008); J. Stat. Mech. P01021 (2009)
+\bibitem{Kun:2007} F. Kun et al., J. Stat. Mech. P02003 (2007); Phys. Rev. Lett. {\bf 100}, 094301 (2008); J. Stat. Mech. P01021 (2009).
 
@@
 
-\bibitem{Suresh:2006} S. Suresh, {\it Fatigue of Materials} (Cambridge University Press, Cambridge, England, 2006)
+\bibitem{Kohout:2000} J. Kohout, Fatigue Fract. Eng. Mater. Struct. {\bf 23}, 969--977 (2000).
+
+%\bibitem{Suresh:2006} S. Suresh, {\it Fatigue of Materials} (Cambridge University Press, Cambridge, England, 2006).
 
@@
 
+\bibitem{Kalab:1983} M. Kal\'ab, P. Allan-Wojtas and B.E. Phipps-Todd, Food Microstruct. {\bf 2}, 51--66 (1983).
+
 \bibitem{Roefs:1990} S.P.F.M. Roefs, A.E.A. De Groot-Mostert and T. Van Vliet, Colloids Surf. {\bf 50}, 141--159 (1990).
-
-\bibitem{Kalab:1983} M. Kal\'ab, P. Allan-Wojtas and B.E. Phipps-Todd, Food Microstruct. {\bf 2}, 51--66 (1983).
 
@@
 Supplemental Movie~1 shows the failure of a 4\%~wt.
-casein gel acidified with 1\%~wt.
+casein gel acidified with 4\%~wt.
 GDL in a plate-plate geometry for an imposed shear stress $\sigma=120$~Pa.
@@
 
-Supplemental Figure~\ref{suppfig1} illustrates the gelation process of a 4\%.
+Supplemental Figure~\ref{suppfig1} illustrates the gelation process of a 4\% wt.
 sodium caseinate solution seeded with 3\% wt.
 polyamide spheres through time-resolved measurements of the elastic and viscous moduli, respectively $G'$ and $G''$, under small amplitude oscillatory shear.
-The glucono-$\delta$-lactone contains an ester function which, once added to the solution at time $t=0$, hydrolyzes spontaneously and leads to a slow decrease of the pH [Fig.~\ref{suppfig1}(a)] towards the casein isoelectric point (${\rm pH}\simeq 4.6$) at which casein particles aggregate.
+GDL contains an ester function which, once added to the solution at time $t=0$, hydrolyzes spontaneously and leads to a slow decrease of the pH [Fig.~\ref{suppfig1}(a)] towards the casein isoelectric point (${\rm pH}\simeq 4.6$) at which casein particles aggregate.
 Gelation actually starts earlier at ${\rm pH}\simeq 5$ as shown by gray dashed lines and as already discussed in Ref.~\cite{Lucey:1998supp}.
@@
 Note that the maximum of the overshoot is reached around the isoelectric point.
-The decrease of both moduli at lower pH is usually attributed to the over-acidification which enhances the repulsive electrostatic interactions between caseinate particles of net positive charge \cite{Dickinson:2002supp}.
+The decrease of both moduli at lower pH is usually attributed to the over-acidification which enhances the repulsive electrostatic interactions between casein particles of net positive charge \cite{Dickinson:2002supp}.
 
@@
 (a)~pH vs time $t$.
-(b)~Linear viscoelastic moduli $G'$ (top, in blue) and $G''$ (bottom, in red) recorded as a function of time $t$ for fixed frequency $f=1$~Hz and strain amplitude 0.1~\%.
+(b)~Linear viscoelastic moduli $G'$ (top, in blue) and $G''$ (bottom, in red) recorded as a function of time $t$ for fixed frequency $f=1$~Hz and strain amplitude 0.1\%.
 Dashed lines indicate that the system starts to build significant elastic properties when the pH decreases below about 5.

### 2014-04-07 plis_figures.tex
zoom with nodes

@@
 		%zooms
-		\draw[Accent2] (a.north west) ++(0.111\columnwidth, -0.072\columnwidth) rectangle +(0.137\columnwidth,-0.137\columnwidth) -- (b.south west) (a.north west) ++(0.248\columnwidth, -0.072\columnwidth) -- (b.north west);
-		\draw[Main] (b.north west) ++(0.097\columnwidth,-0.143\columnwidth) rectangle +(0.128\columnwidth,-0.081\columnwidth) -- (c.south west) (b.north west) ++(0.225\columnwidth,-0.143\columnwidth) --(c.north west);
-		\draw[Accent1] (c.north west) ++(0.125\columnwidth,0) rectangle +(0.156\columnwidth,-0.156\columnwidth);
+		\node[minimum width = 0.137\columnwidth, minimum height=0.137\columnwidth, anchor=north west, draw=Accent2] at ($(a.north west) +(0.111\columnwidth,-0.072\columnwidth)$) (bz){};
+		\draw[Accent2] (bz.north east) -- (b.north west) (bz.south east) -- (b.south west);
+		\node[minimum width = 0.128\columnwidth, minimum height=0.081\columnwidth, anchor=north west, draw=Main] at ($(b.north west) +(0.097\columnwidth,-0.143\columnwidth)$) (cz){};
+		\draw[Main] (cz.north east) -- (c.north west) (cz.south east) -- (c.south west);
+		\node[minimum width = 0.156\columnwidth, minimum height=0.156\columnwidth, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\columnwidth,0)$) (dz) {};
 		%scale bars

### 2014-04-22 creep/creep_yaourt.tex
- primary creep accounted for -> exponent of the primary creep - reversible -> kinecically reversible - creep/recovery => viscous dissipation and negligible amount of plastic events - viré les pre-cracks - Aknowledge Gareth

@@
 Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids and to rupture under an external load.
-Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is typical of brittle solids: after a power-law creep regime fully accounted for by linear viscoelasticity and homogeneous deformation, fractures nucleate and grow logarithmically perpendicularly to shear up to sudden rupture.
+Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is typical of brittle solids: \textcolor{blue}{after an homogeneous-deformation power-law creep regime whoes exponent is fully accounted for by linear viscoelasticity}, fractures nucleate and grow logarithmically perpendicularly to shear up to sudden rupture.
 A single equation accounting for those two successive processes nicely captures the full rheological response.
@@
 Gels formed by slow acidification of a sodium caseinate solution \cite{Lucey:1998} display fractures under large strain at fixed low pH values \cite{vanVliet:1995}, which makes them perfect candidates to quantify brittleness in soft solids and tackle the above-mentioned issues.
-We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: a reversible primary creep regime characterized by a homogeneous strain field and the irreversible nucleation and fracture growth leading to gel failure.
+We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: a \textcolor{blue}{kinetically} reversible primary creep regime characterized by a homogeneous strain field and the irreversible nucleation and fracture growth leading to gel failure.
 Our results are in full agreement with the predictions of some recent fiber-bundle models and hint to universal features of brittle failure common to both soft and hard solids.
@@
 \caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
-Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images.
+Ultrasonic images are \textcolor{blue}{azimuthal} velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images.
 Their position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$.
@@
 This is strongly reminiscent of the {\it primary creep} observed in solids and referred to as Andrade creep \cite{Andrade:1910,Miguel:2002,Nechad:2005}.
-Interestingly, here, this scaling can be fully accounted for by linear viscoelasticity without invoking any irreversible plasticity.
+Interestingly, here, \textcolor{blue}{the exponent} can be fully accounted for by linear viscoelasticity without invoking any irreversible plasticity.
 Indeed casein gels display a power-law rheology $G'(f)\sim G''(f)\sim f^{0.15}$ [inset of Fig.~\ref{fig2}(a)], which corresponds to a compliance $J(t)\equiv\gamma(t)/\sigma\sim t^{0.15}$ in the linear deformation regime \cite{Tschoegl:1989}.
-This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$ and suggests that primary creep is purely reversible without unrecoverable plastic events.
+This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$ and suggests that primary creep is \textcolor{blue}{\textit{kinetically} reversible, i.e.
+with a negligible amount of unrecoverable plastic events}.
 Homogeneity and linear response are further confirmed, at least up to the available spatial ($\sim 10~\mu$m) and temporal ($\sim 1$~s) resolutions, by direct visualization which shows no sign of crack or fracture [Fig.~\ref{fig3}(a)] and by ultrasonic imaging which reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)] and no slippage at the Plexiglas walls.
+\textcolor{blue}{To go one step further we have performed creep/recovery experiments limited to the primary creep regime (see Fig.~\ref{suppfig3}) that show almost complete recovery.
+The remaining deformation can be attributed mainly to viscous dissipation in the solvent since when the strain is reset to zero, i.e.
+the moving tool is moved back to its initial position, the same loading induces the same response.
+Undistinguishable linear reology measured before each cycle (see Fig.~\ref{suppfig4}) further rules out significant plastic events, although microscopic observation may found a negligible amount of damaged fibers.}
 
@@
 The level of these fluctuations, which can also be seen on the global response for the same applied stress [\textcolor{red!50!black}{$\blacktriangledown$} in Fig.~\ref{fig2}(b)], is poorly reproducible and stress-dependent.
-We propose that such intermittency arises from microscopic cracks ahead of the fracture tip.
-Since heterogeneity is much smaller along the radial direction [compare red and black lines in Fig~\ref{fig4}(d)], these ``precursors'' either extend across the full 2-mm gap or induce long-range displacements that span the gap.
-In any case, such noisy features are typical of slow crack growth in solids \cite{Vanel:2009,Kun:2007} and more generally of avalanche behavior \cite{Sethna:2001}.
+\textcolor{blue}{Such intermittency may arises from the growth of cracks outside the region of interest, inducing elastic deformation propagating in the sample.
+The much smaller heterogeneity along the radial direction [compare red and black lines in Fig~\ref{fig4}(d)] is coherent with induced long-range displacements that span the gap.} In any case, such noisy features are typical of slow crack growth in solids \cite{Vanel:2009,Kun:2007} and more generally of avalanche behavior \cite{Sethna:2001}.
 
@@
 
-Finally, we emphasize that the (reversible) Andrade-like creep and the (irreversible) crack growth are two  physical processes that effectively superimpose to yield the global rheological response.
+Finally, we emphasize that the (\textcolor{blue}{kinetically} reversible) Andrade-like creep and the (irreversible) crack growth are two  physical processes that effectively superimpose to yield the global rheological response.
 Indeed, as seen from the yellow line in Fig.~\ref{fig2}, the master curve $\gp(t)/\gp_{\rm min}$ vs $t/\tau_f$ is perfectly fitted by:
@@
 First we have shown that casein gels display a remarkable failure scenario similar to that of brittle solids and characterized by the same three successive creep regimes.
-Here, the primary creep and its power-law exponent [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] are fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
+Here, the power-law exponent of the primary creep [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] is fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
 Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Gobeaux:2010,Jaishankar:2013} as well as hard-sphere-like colloidal glasses \cite{Siebenburger:2012}.
-Protein gels thus appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials.
+Protein gels thus appear to remain purely elastic and \textcolor{blue}{kinetically} reversible throughout primary creep, a hallmark of brittle materials.
 By contrast primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.
@@
 \begin{acknowledgments}
-The authors thank T.~Gibaud, S.~Lindstr\"om and N.~Taberlet for fruitful discussions, M.~Kal\'ab for kindly providing us with unobtainable references, and A.~Parker at Firmenich for precious advice and for providing the casein and GDL.
+The authors thank T.~Gibaud, S.~Lindstr\"om\textcolor{blue}{, G.H.~McKinley} and N.~Taberlet for fruitful discussions, M.~Kal\'ab for kindly providing us with unobtainable references, and A.~Parker at Firmenich for precious advice and for providing the casein and GDL.
 This work was funded by the Institut Universitaire de France and by the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement No.~258803.

### 2014-04-23 creep/creep_yaourt.tex
Much more care in the use of "brittle" (disapear from the title) yield stress fluids in intro Move the discussion on reversibility to the discussion. Reduce out of ROI fracture influence Aknowledge Mikko Alava Tabuteau:2009, Skrzeszewska:2010 SupplFig3 et 4

--- 2014-01-31 creep/creep_yaourt.tex
@@
 
- \title{Creep and brittle failure of a protein gel under stress}
+ \title{Creep and fracture of a protein gel under stress}
 
@@
 Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids and to rupture under an external load.
-Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is typical of brittle solids: after a power-law creep regime fully accounted for by linear viscoelasticity and homogeneous deformation, fractures nucleate and grow logarithmically perpendicularly to shear up to sudden rupture.
+Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is reminiscent of brittle solids: after a primary creep regime characterized by a macroscopically homogeneous deformation and a power-law behavior which exponent is fully accounted for by linear viscoelasticity, fractures nucleate and grow logarithmically perpendicularly to shear, up to the sudden rupture of the gel.
 A single equation accounting for those two successive processes nicely captures the full rheological response.
-The failure time follows a decreasing power-law with the applied shear stress strongly reminiscent of the Basquin law of fatigue for solids.
-These results are in excellent agreement with recent fiber-bundle models that include damage accumulation on elastic fibers and exemplify protein gels as model brittle soft solids.
+The failure time follows a decreasing power-law with the applied shear stress, similar to the Basquin law of fatigue for solids.
+These results are in excellent agreement with recent fiber-bundle models that include damage accumulation on elastic fibers and exemplify protein gels as model, brittle-like soft solids.
 \end{abstract}
@@
 
-%Under a constant external load, a solid deforms and may eventually fail.
-The resulting slow deformation, referred to as {\it creep}, has been investigated over a century on solid materials \cite{Andrade:1910,Nechad:2005,Miguel:2002} but only a handful of papers have dealt with creep on soft solids, i.e.
-gels made through the reversible association of biopolymers such as proteins or polysaccharides \cite{Higgs:1990,vanVliet:1995}.
-These materials, whose physical properties play a key role not only in food science, from set-style yogurt or cheese to jelly-based deserts \cite{Mezzenga:2005} but also in biomechanics, biochemistry and microbiology \cite{Viovy:2000,Grinnell:2010}, behave as elastic solids under small deformations but display remarkable nonlinear behavior featuring fractures prior to irreversible rupture \cite{Bonn:1998,Baumberger:2006,Daniels:2007,Brenner:2013}.
-Irreversibility directly stems from the existence of an external control parameter, namely temperature (resp.
-pH) in the case of thermoreversible gels (resp.
-acid-induced gels), which makes the system impossible to shear-rejuvenate after failure, contrary to other soft glassy materials such as emulsions, colloidal gels and glasses \cite{Cloitre:2000,Divoux:2012,Siebenburger:2012}.
-While great effort has been devoted to designing protein- and polysaccharide-based soft solids with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}, their nonlinear mechanical behavior under a shear stress has only been partially addressed \cite{vanVliet:1995} and several simple issues remain unexplored such as the spatially resolved scenario for creep and fracture or the physical relevance of the analogy with the failure of hard, brittle solids.
-
 Biogels formed through the self-association of polysaccharide coils, collagen, actin filaments or attractive globular proteins play a major role in biochemistry and microbiology \cite{Viovy:2000}, biological networks and cell mechanics \cite{Stricker:2010} as well as in food science \cite{Mezzenga:2005}.
@@
 Irreversibility stems from the existence of an external control parameter, e.g.
-temperature or pH in the case of thermoreversible or acid-induced gels respectively, which makes such biogels impossible to shear-rejuvenate after failure, contrary to other soft glassy materials such as emulsions, colloidal gels and glasses \cite{Cloitre:2000,Divoux:2012,Siebenburger:2012}.
-So far huge effort has been devoted to designing protein gels with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}.
+temperature or pH in the case of thermoreversible or acid-induced gels respectively.
+This makes such biogels fundamentally different from other soft glassy materials such as emulsions, colloidal gels and glasses that can be rejuvenated by shear \cite{Cloitre:2000,Caton:2008,Divoux:2011,Siebenburger:2012} or transient networks where fractures spontaneously heal \cite{Tabuteau:2009,Skrzeszewska:2010}.
+So far, huge effort has been devoted to the design of protein gels with specific properties and textures at rest \cite{Dickinson:2006,Gibaud:2012}.
 However, their mechanical behavior deep into the nonlinear regime has only been partially addressed \cite{vanVliet:1995,Pouzot:2006} and several fundamental issues remain unexplored such as the spatially resolved rupture scenario or the physical relevance of the analogy with brittle failure in hard solids.
@@
 casein gel acidified with 1\%~wt.
-GDL for an imposed shear stress $\sigma=200$ (\textcolor{red!25!black}{$\bullet$}), 300 (\textcolor{red!50!black}{$\blacktriangledown$}), 400 (\textcolor{red!75!black}{$\blacksquare$}), 550 (\textcolor{red}{$\blacktriangle$}), and 1000~Pa (\textcolor{orange!50!red}{$\blacklozenge$}) from right to left.
+GDL for an imposed shear stress $\sigma=200$ (\textcolor{red!25!black}{$\bullet$}), 300 (\textcolor{red!50!black}{$\blacktriangledown$}), 400 (\textcolor{red!75!black}{$\blacksquare$}), 550 (\textcolor{red}{$\blacktriangle$}) and 1000~Pa (\textcolor{orange!50!red}{$\blacklozenge$}) from right to left.
 The gap width is 1~mm.
@@
 In this Letter we report on stress-induced fracture in protein gels by means of creep experiments coupled to optical and ultrasonic imaging.
-Gels formed by slow acidification of a sodium caseinate solution \cite{Lucey:1998} display fractures under large strain at fixed low pH values \cite{vanVliet:1995}, which makes them perfect candidates to quantify brittleness in soft solids and tackle the above-mentioned issues.
-We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: a reversible primary creep regime characterized by a homogeneous strain field and the irreversible nucleation and fracture growth leading to gel failure.
-Our results are in full agreement with the predictions of some recent fiber-bundle models and hint to universal features of brittle failure common to both soft and hard solids.
+Gels formed by slow acidification of a sodium caseinate solution display fractures under large strain at fixed low pH values \cite{vanVliet:1995,Lucey:1998}, which makes them perfect candidates to quantify the rupture of soft solids and tackle the above-mentioned issues.
+We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: (i) a primary creep regime characterized by a macroscopically homogeneous strain field where dissipation is dominated by viscous flow through the gel matrix and (ii) the irreversible nucleation and growth of fractures leading to gel failure.
+Our results are in full agreement with the predictions of some recent fiber-bundle models and hint to universal features of failure common to both soft and hard solids.
 
@@
 \includegraphics[width=17.5cm,clip]{Fig3.eps}
-\caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$], and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
-Ultrasonic images are velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images.
+\caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$] and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
+Ultrasonic images are azimuthal velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images.
 Their position relative to the picture of the cell reflects the actual arrangement of the ultrasonic probe along the vertical direction $z$.
@@
 $\tau_f$ further allows us to rescale all the shear rate data $\gp(t)$ onto the single master curve of Fig.~\ref{fig2}(a) by plotting $\gp/\gp_{\rm min}$ vs $t/\tau_f$, where $\gp_{\rm min}$ is the minimum shear rate reached at a time $\tau_{\rm min}$ [see also Fig.~\ref{suppfig2} in the Supplemental Material for the original $\gp(t)$ data].
-On about four decades for $t\lesssim 0.1\tau_f$, the shear rate decreases as a power-law $\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$.
-This is strongly reminiscent of the {\it primary creep} observed in solids and referred to as Andrade creep \cite{Andrade:1910,Miguel:2002,Nechad:2005}.
-Interestingly, here, this scaling can be fully accounted for by linear viscoelasticity without invoking any irreversible plasticity.
+On about four decades for $t\lesssim 0.1\tau_f$, the shear rate decreases as a power-law $\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.04$.
+This is strongly reminiscent of the {\it primary creep} observed in solids and referred to as Andrade creep \cite{Andrade:1910,Miguel:2002,Nechad:2005,Rosti:2010}.
+Interestingly, here, the exponent can be inferred from linear viscoelasticity.
 Indeed casein gels display a power-law rheology $G'(f)\sim G''(f)\sim f^{0.15}$ [inset of Fig.~\ref{fig2}(a)], which corresponds to a compliance $J(t)\equiv\gamma(t)/\sigma\sim t^{0.15}$ in the linear deformation regime \cite{Tschoegl:1989}.
-This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$ and suggests that primary creep is purely reversible without unrecoverable plastic events.
-Homogeneity and linear response are further confirmed, at least up to the available spatial ($\sim 10~\mu$m) and temporal ($\sim 1$~s) resolutions, by direct visualization which shows no sign of crack or fracture [Fig.~\ref{fig3}(a)] and by ultrasonic imaging which reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)] and no slippage at the Plexiglas walls.
+This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$.
+Moreover ultrasonic imaging reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)] and no slippage at the Plexiglas walls.
+Together with direct visualization [Fig.~\ref{fig3}(a)], these local measurements demonstrate that during primary creep there is no macroscopic strain localization or fracture, although we cannot rule out microscopic rearrangements below the available spatial ($\sim10~\mu$m) and temporal ($\sim1$~s) resolutions.
 
-For $0.1\lesssim t/\tau_f\lesssim 0.9$, $\gp(t)$ departs from power-law behavior and goes through a minimum value at time $\tau_{\rm min}= (0.56\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}(b) and inset] as similarly reported for metals \cite{Sundararajan:1989}, solid composite materials \cite{Nechad:2005} and fiber-bundles models (FBMs) \cite{Kovacs:2008,Jagla:2011}.
+For $0.1\lesssim t/\tau_f\lesssim 0.9$, $\gp(t)$ departs from power-law behavior and goes through a minimum value at time $\tau_{\rm min}= (0.56\pm0.04)\tau_f$ independently of the applied stress [Fig.~\ref{fig2}(b) and inset] as similarly reported for metals \cite{Sundararajan:1989}, solid composite materials \cite{Nechad:2005} and fiber-bundle models (FBMs) \cite{Kovacs:2008,Jagla:2011}.
 This linearity between $\tau_{\rm min}$ and $\tau_f$, also known as the Monkman-Grant relation \cite{Monkman:1956}, allows one to ``predict'' the failure time from the intermediate-time response.
@@
 The level of these fluctuations, which can also be seen on the global response for the same applied stress [\textcolor{red!50!black}{$\blacktriangledown$} in Fig.~\ref{fig2}(b)], is poorly reproducible and stress-dependent.
-We propose that such intermittency arises from microscopic cracks ahead of the fracture tip.
-Since heterogeneity is much smaller along the radial direction [compare red and black lines in Fig~\ref{fig4}(d)], these ``precursors'' either extend across the full 2-mm gap or induce long-range displacements that span the gap.
-In any case, such noisy features are typical of slow crack growth in solids \cite{Vanel:2009,Kun:2007} and more generally of avalanche behavior \cite{Sethna:2001}.
+Such intermittency may arise from crack growth outside the region of interest which induces long-range elastic deformation propagating across the sample mainly along the vertical direction [see the lower level of fluctuations along the radial direction in Fig~\ref{fig4}(d)].
 
@@
 \includegraphics[width=17.5cm,clip]{Fig4.eps}
-\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ ($\bullet$), $1.7\,10^{-2}$ (\textcolor{blue}{$\blacktriangledown$}), and 0.15 (\textcolor{red}{$\blacksquare$}).
+\caption{(color online) (a)~Local velocity $\langle v(r,z,t)\rangle_z$ and (b)~local strain field $\langle\gl(r,z,t)\rangle_z$ averaged over the vertical direction $z$ at various times during primary creep: $t/\tau_f=1.9\,10^{-3}$ ($\bullet$), $1.7\,10^{-2}$ (\textcolor{blue}{$\blacktriangledown$}) and 0.15 (\textcolor{red}{$\blacksquare$}).
 Solid lines are linear profiles.
@@
 
-Finally, we emphasize that the (reversible) Andrade-like creep and the (irreversible) crack growth are two  physical processes that effectively superimpose to yield the global rheological response.
+Finally, we emphasize that the Andrade-like creep and the crack growth are two  physical processes that effectively superimpose to yield the global rheological response.
 Indeed, as seen from the yellow line in Fig.~\ref{fig2}, the master curve $\gp(t)/\gp_{\rm min}$ vs $t/\tau_f$ is perfectly fitted by:
@@
 First we have shown that casein gels display a remarkable failure scenario similar to that of brittle solids and characterized by the same three successive creep regimes.
-Here, the primary creep and its power-law exponent [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.02$] are fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
+Here, the power-law exponent of the primary creep [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.04$] is fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
 Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Gobeaux:2010,Jaishankar:2013} as well as hard-sphere-like colloidal glasses \cite{Siebenburger:2012}.
-Protein gels thus appear to remain purely elastic and reversible throughout primary creep, a hallmark of brittle materials.
-By contrast primary creep in solids is irreversible and has been attributed to dislocation interactions in the case of crystalline materials \cite{Miguel:2002} or to localized plastic rearrangements in amorphous materials \cite{Rosti:2010}.
+To better understand this link, additional creep and recovery tests were performed within the primary regime (Figs.~3 and 4 in the Supplemental Material).
+Successive loading/unloading sequences on a fresh sample show that (i)~the strain is not fully recovered by a few percents, (ii)~the viscoelastic properties of the material are unaltered and (iii)~strain responses are indistinguishable from one test to another.
+This strongly suggests that the small irreversibility observed in the primary regime mainly stems from viscous dissipation (which accounts for the nonzero viscous modulus $G''$) due to the solvent flow within the gel fibrous matrix and involves only a minute amount of local damage of the gel network.
+Protein gels thus appear to experience a {\it kinetically} reversible primary creep, i.e.
+dominated by viscous dissipation rather than by plasticity.
 
@@
 Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Kohout:2000}.
-Basquin law has also been recently predicted by FBMs that combine elastic fibers with a local yield strain and take into account damage accumulation \cite{Kun:2007,Halasz:2012}.
+Basquin law has also been recently predicted for creep experiments by FBMs that combine elastic fibers with a local yield strain and take into account damage accumulation \cite{Kun:2007,Halasz:2012}.
 Interestingly assuming damage accumulation to be proportional to $\sigma^\gamma$ directly leads to Basquin law with $\beta=\gamma$ and large values of $\gamma$, typically larger than 5 as found here for $\beta$, lead to macroscopic cracks due to the simultaneous rupture of a large number of fibers \cite{Halasz:2012}.
-More generally FBMs under compressive or elongational load predict three successive creep regimes exactly alike Fig.~\ref{fig2} with similar proportionality between $\tau_{\rm min}$ and $\tau_f$ and finite-time singularity \cite{Nechad:2005,Jagla:2011}.
+More generally FBMs under elongational load predict three successive creep regimes exactly alike Fig.~\ref{fig2} with similar proportionality between $\tau_{\rm min}$ and $\tau_f$ and finite-time singularity \cite{Nechad:2005,Jagla:2011}.
 This highlights the relevance of FBMs in the context of creep in protein gels, whose microstructure indeed appears to be formed of strands \cite{Kalab:1983,Roefs:1990}, but also urges to study FBMs in shear geometries to check whether they would be able to predict fracture growth as observed here.
 
-To conclude, the present time- and space-resolved study shows that protein gels qualify as model, brittle soft solids.
+To conclude, the present time- and space-resolved study exemplifies protein gels as model, brittle-like soft solids.
 Prompted by the remarkable simplicity of Eq.~(\ref{e.fit}) that encompasses Andrade creep, the Monkman-Grant relation and finite-time singularity within a single equation, future modeling will undoubtedly focus on the microscopic ingredients needed to predict quantitatively the Monkman-Grant prefactor and the Basquin exponent $\beta$.
@@
 
-%which separates into three successive phases: (i)~a primary (or Andrade-like) creep regime during which the strain field is homogeneous, (ii)~a secondary creep during which fractures nucleate and grow logarithmically, and (iii)~a finite-time singularity concomittant with the sudden failure of the gel.
-The failure time displays a non-trivial power-law scaling with the applied shear-stress that contrasts with the exponential scalings usually reported for solid brittle materials \cite{Vanel:2009}.
-
-% within the spatial resolution of a few microns, the strain field is homogeneous [Fig~\ref{fig4}~(b)].
-If plastic events take place in this primary regime, they occur at smaller scale.
-However, given the agreement between the Andrade exponent and the frequency scaling of the linear elastic modulus discussed above, it is more likely that no or a negligible amount of plastic events take place during this regime, corroborating the optical and ultrasonic measurements.
-
-%Compare to the scalings observed in YSF - in particular microgels display a similar power-law scaling, except that the yield point appears explicitely in the scaling as the critical point in a dynamical phase transition \cite{Divoux:2012}.
-Propose an explanation or at least raise a point.
-\cite{Chaudhuri:2013}
-
-%FBM that rely on elastic fibers with a local yield strain (or stress) and introduce a non-linear Eyring rheology or a local coupling lead to a different scaling law than the one we find here for the failure time with the applied stress.
-In their law appears a ``yield stress" \cite{Kun:2003,Nechad:2005,Kovacs:2008,Jagla:2011} that is not present here.
-
-%different from other soft materials such as transient-network systems \cite{Ligoure:2013}, because truly irreversible.
-The pH is key to the irreversibility and plays the role of an additional control parameter that, when maintained at values lower than the isoelectric point, prevent the system from healing.
-
-%Comparatively, Yield Stress Fluids (YSF) that display a solid-like behavior at rest, either because of a jammed microstructure or attractive interactions \cite{Sciortino:2005,Bonnecaze:2010}, but do not present such external control parameter, display a similar rheological signature during their stress-induced yielding.
-Indeed the S-shaped curved pictured in figure~\ref{fig2}~(a) has been also reported in \cite{Bauer:2006,Gibaud:2010,Divoux:2011,Siebenburger:2012,Grenard:2013}.
-However, their yielding scenario is fundamentally different.
-Local measurements conducted on both type of soft glassy materials have revealed that no fractures take place in the bulk: the microstructure rather reorganizes and flows in a ductile way.
-Our study reveals that acid-induced gels behave elastically up to the rupture point and that fractures develop and grow logarithmically, perpendicularly to the shear direction prior to an irreversible failure which makes casein gels model {\it brittle soft solids}.
-
-%and yield stress fluids (YSFs) \cite{Bauer:2006,Divoux:2011,Siebenburger:2012,Grenard:2013} that usually exhibit a power-law creep with smaller exponents (0.3 $\lesssim \alpha \lesssim 0.7$) which are not directly related to linear rheology.
-Besides, the linear behavior of YSF generally differs from a simple power-law scaling, while in the case of solids
-
 \begin{acknowledgments}
-The authors thank T.~Gibaud, S.~Lindstr\"om and N.~Taberlet for fruitful discussions, M.~Kal\'ab for kindly providing us with unobtainable references, and A.~Parker at Firmenich for precious advice and for providing the casein and GDL.
+The authors thank M.
+Alava, T.~Gibaud, S.~Lindstr\"om, G.~H.~McKinley and N.~Taberlet for fruitful discussions, M.~Kal\'ab for kindly providing us with unobtainable references, and A.~Parker at Firmenich for precious advice and for providing the casein and GDL.
 This work was funded by the Institut Universitaire de France and by the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement No.~258803.
@@
 
-\bibitem{Divoux:2012} T. Divoux et al., Soft Matter {\bf 8}, 4151--4164 (2012).
+\bibitem{Caton:2008} F. Caton and C. Baravian, Rheol. Acta {\bf 47}, 601--607 (2008).
 
-\bibitem{Siebenburger:2012} Siebenb\"urger, M. Ballauff and T. Voigtmann, Phys. Rev.
+\bibitem{Divoux:2011} T. Divoux et al., Soft Matter {\bf 7}, 8409--8418 (2011).
+
+\bibitem{Siebenburger:2012} M. Siebenb\"urger, M. Ballauff and T. Voigtmann, Phys. Rev.
 Lett., {\bf 108}, 255701 (2012).
+
+\bibitem{Tabuteau:2009} H. Tabuteau et al., Phys. Rev. Lett., {\bf 102}, 155501 (2009); C. Ligoure and S. Mora, Rheol. Acta {\bf 52}, 91--114 (2013).
+
+\bibitem{Skrzeszewska:2010} P. J. Skrzeszewska et al., Macromolecules, {\bf 43}, 3542--3548 (2010).
 
@@
 
+\bibitem{Rosti:2010} J. Rosti, J. Koivisto, L. Laurson, M.J. Alava, Phys. Rev. Lett. {\bf 105}, 100601 (2010).
+
 \bibitem{Tschoegl:1989}  N.W. Tschoegl, {\it The Phenomenological Theory of Linear Viscoelastic Behavior} (Springer, Berlin, Germany, 1989).
@@
 \bibitem{Monkman:1956} F. C. Monkman and N. J. Grant, Proc. Am. Soc. Test.
-Mater. {\bf 56}, 593-- (1956).
+Mater. {\bf 56}, 593--620 (1956).
 
-\bibitem{Vanel:2009} L. Vanel et al., J. Phys. D: Appl. Phys. {\bf 42}, 214007 (2009).
-
-\bibitem{Kun:2007} F. Kun et al., J. Stat. Mech. P02003 (2007); Phys. Rev. Lett. {\bf 100}, 094301 (2008); J. Stat. Mech. P01021 (2009).
-
-\bibitem{Sethna:2001} J.P. Sethna, K.A. Dahmen and C.R. Myers, Nature {\bf 410}, 242--250 (2001).
+%\bibitem{Sethna:2001} J.P. Sethna, K.A. Dahmen and C.R. Myers, Nature {\bf 410}, 242--250 (2001).
 
@@
 
-\bibitem{Rosti:2010} J. Rosti, J. Koivisto, L. Laurson, M.J. Alava, Phys. Rev. Lett. {\bf 105}, 100601 (2010).
+\bibitem{Vanel:2009} L. Vanel et al., J. Phys. D: Appl. Phys. {\bf 42}, 214007 (2009).
+
+\bibitem{Kun:2007} F. Kun et al., J. Stat. Mech. P02003 (2007); Phys. Rev. Lett. {\bf 100}, 094301 (2008); J. Stat. Mech. P01021 (2009).
 
@@
 polyamide spheres through time-resolved measurements of the elastic and viscous moduli, respectively $G'$ and $G''$, under small amplitude oscillatory shear.
-GDL contains an ester function which, once added to the solution at time $t=0$, hydrolyzes spontaneously and leads to a slow decrease of the pH [Fig.~\ref{suppfig1}(a)] towards the casein isoelectric point (${\rm pH}\simeq 4.6$) at which casein particles aggregate.
+GDL contains an ester function which, once added to the solution at time $t=0$, hydrolyzes spontaneously and leads to a slow decrease of the pH [Suppl.
+Fig.~\ref{suppfig1}(a)] towards the casein isoelectric point (${\rm pH}\simeq 4.6$) at which casein particles aggregate.
 Gelation actually starts earlier at ${\rm pH}\simeq 5$ as shown by gray dashed lines and as already discussed in Ref.~\cite{Lucey:1998supp}.
-At this point the elastic and viscous moduli display a sudden increase then overshoot and converge towards their steady-state values where $G'\gg G''$ [Fig.~\ref{suppfig1}(b)].
+At this point the elastic and viscous moduli display a sudden increase then overshoot and converge towards their steady-state values where $G'\gg G''$ [Suppl.
+Fig.~\ref{suppfig1}(b)].
 Note that the maximum of the overshoot is reached around the isoelectric point.
@@
 
-Supplemental Figure~\ref{suppfig2} shows the shear rate $\gp(t)$ obtained by differentiating the raw strain of Fig.~\ref{fig1} in the main text.
-For $0.1\lesssim t/\tau_f\lesssim 0.9$, i.e.
-in the secondary creep regime, $\gp(t)$ is smoothed using a moving average over $\delta t\simeq 0.01\tau_f$ in order to remove high-frequency noise due to differentiation at very small shear rates.
-The same data normalized by its minimum value $\gp_{\rm min}$ is plotted in Fig.~\ref{fig2}(a) as a function of $t/\tau_f$ leading to a nice collapse onto a single curve whatever the imposed shear stress.
-
  %%%%%%%%%%%%
-\begin{figure}[th]
+\begin{figure}[t]
 \centering
@@
 %%%%%%%%%%%%
+
+Supplemental Figure~\ref{suppfig2} shows the shear rate $\gp(t)$ obtained by differentiating the raw strain of Fig.~\ref{fig1} in the main text.
+For $0.1\lesssim t/\tau_f\lesssim 0.9$, i.e.
+in the secondary creep regime, $\gp(t)$ is smoothed using a moving average over $\delta t\simeq 0.01\tau_f$ in order to remove high-frequency noise due to differentiation at very small shear rates.
+The same data normalized by its minimum value $\gp_{\rm min}$ is plotted in Fig.~\ref{fig2}(a) as a function of $t/\tau_f$ leading to a nice collapse onto a single curve whatever the imposed shear stress.
+
+\clearpage
+\newpage
+
+ %%%%%%%%%%%%
+\begin{figure}
+\centering
+\includegraphics[width=8cm,clip]{SuppFig3.eps}
+\caption{(color online) Three creep and recovery tests preformed successively in the primary creep regime on the same 4\%~wt.
+casein gel acidified with 1\%~wt.
+GDL.
+Prior to each test, the viscoelastic moduli are measured over 450~s under small amplitude oscillations around the reference position $\gamma=0$ (see Suppl.
+Fig.~\ref{suppfig4}).
+A constant stress $\sigma=100$~Pa is then applied for 600~s and finally the relaxation at $\sigma=0$ is followed over 8,000~s.
+(a)~Full strain response $\gamma(t)$ vs time.
+(b)~Superimposed strain responses during the three successive creep tests at $\sigma=100$~Pa and plotted in logarithmic scales as a function of the time since stress application.
+The yellow line is $\gamma(t)= 0.048+0.083 t^{0.15}$.
+(c)~Superimposed strain responses during the three recovery phases and plotted in logarithmic scales as a function of the time since stress is removed.
+\label{suppfig3}}
+\end{figure} 
+%%%%%%%%%%%%
+
+ %%%%%%%%%%%%
+\begin{figure}
+\centering
+\includegraphics[width=6.5cm,clip]{SuppFig4.eps}
+\caption{(color online) Linear viscoelastic moduli $G'$ (top) and $G''$ (bottom) as a function of frequency $f$ for a strain amplitude of 0.1\% recorded before each creep and recovery test shown in Suppl.
+Fig.~\ref{suppfig3}.
+The symbols \textcolor{red!25!black}{$\bullet$}, \textcolor{red!50!black}{$\blacktriangledown$} and \textcolor{red!75!black}{$\blacksquare$} respectively correspond to the first, second and third tests.
+Black lines are power laws $G'\sim G''\sim f^{0.15}$.
+\label{suppfig4}}
+\end{figure} 
+%%%%%%%%%%%%
+
+Successive creep and recovery experiments performed on the same sample within the primary creep regime are presented in Supplemental Figures~\ref{suppfig3} and \ref{suppfig4}.
+When a stress of 100~Pa is applied at time $t=0$, an ``instantaneous'' elastic strain of about 10\% is recorded, followed by the slow Andrade creep already investigated in the main text and characterized by an exponent of 0.15 [Suppl.
+Fig.~\ref{suppfig3}(b)] that coincides with that of the frequency-dependence of both viscoelastic moduli $G'$ and $G''$ [Suppl.
+Fig.~\ref{suppfig4}].
+When stress is removed at $t=600$~s, the strain drops by about 6\% and then slowly relaxes over several hours.
+After 8,000~s we observe that the initial strain is almost fully recovered: only a few percents are not (here 2\% out of a total maximum strain of 27\%).
+
+Such an irrecoverable strain can be attributed either to viscous flow within the porous structure of the gel or to plastic events i.e.
+to local rupture of the gel network, and in the most general case to a combination of both.
+Interestingly, when the strain is reset to zero between two successive creep and recovery tests [i.e.
+the moving tool is moved back to its initial position, see the corresponding trace in Suppl.
+Fig.~\ref{suppfig3}(a)], the strain responses are indistinguishable [Suppl.
+Fig.~\ref{suppfig3}(b) and (c)].
+If significant plastic deformation had occurred during the first loading--unloading cycle, one would have expected the second and third responses to the same cycle to differ significantly, e.g.
+through larger maximum strains or larger unrecoverable strains.
+
+Finally Supplemental Figure 4 shows that both viscoelastic moduli $G'$ and $G''$ do not change significantly from one creep and recovery test to another and that their scaling with the frequency remains the same.
+This indicates that the gel microstructure is mostly undamaged by creep within the primary regime.
 

### 2014-05-09 plis_figures.tex
Zoom on wrinkling regime in semilog scale, exponential fits.

@@
 \usepackage{amssymb}
-\usepackage{siunitx}
+\usepackage[separate-uncertainty = true,multi-part-units=single]{siunitx}
 \usepackage{multirow}
+\usepackage{textcomp}
 
@@
 		extra x tick labels={},%
-		width=1.04\columnwidth,
+		width=1.03\columnwidth,
 		height=0.5\columnwidth,
@@
 
-	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=1.2, xlabel={time (min)}, extra x ticks={26.7, 42.8}]
+	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=1.2, xlabel={time (min)}, extra x ticks={26.2, 42.8}]
 	\addplot+[no marks,Accent1] table{excess_area_pc_half_cas8.txt};
@@
 	\node[lab,above left=0.2em] at (rel axis cs:1,0) (b) {b};
+	\coordinate (wr1) at (axis cs:26.2,0);
+	\coordinate (wr2) at (axis cs:42.8,0);
 	\end{groupplot}
 
-	\matrix[matrix of nodes, matrix anchor=north west, inner sep=0, column sep=0.02\textwidth, row sep=2pt] at (g c1r2.outer south west) (m){
-	Synæresis & Swelling\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t000.png}&
+	\begin{semilogyaxis}[
+		name={zoom},
+		at={(g c1r2.below south west)},
+		anchor=above north west,
+		width=0.5\columnwidth,
+		height=0.6\columnwidth,
+		xmin=26.2, xmax=42.8, xtick={25,30,...,180},
+		ylabel={Excess area (\textperthousand)}, ymin=0.06, ymax=5,
+		xlabel={time (min)},
+		ytickten={-1,0}, yticklabels={0.1,1},
+		]
+		\addplot+[only marks,Accent1, mark=o] table[y expr={\thisrowno{1}*10}]{excess_area_pc_half_cas8.txt};
+		\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*10}]{excess_area_pc_half_cas8_toi.txt};
+		\addplot+[black, no marks,domain=25:32.5] {10*exp(x/2.71-14.6)};
+		\addplot+[black, no marks,domain=25:45] {10*exp(x/5.60-8.81)};
+		\node[lab,above left=0.2em] at (rel axis cs:1,0) (c) {c};
+	\end{semilogyaxis}
+	\draw[help lines] (wr1)--(zoom.north west);
+	\draw[help lines] (wr2)--(zoom.north east);
+
+	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, column sep=0.02\textwidth, row sep=2pt] at (g c1r2.outer south east) (m){
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t000.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t006.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t008.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t010.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t012.png}\\
+	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t014.png}\\
 	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t023.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t006.png}&
 	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t049.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t008.png}&
 	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t051.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t010.png}&
 	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t053.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t012.png}&
 	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t078.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t014.png}&
 	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t088.png}\\
 	};
-	\draw[>->,gray,ultra thick] (m-2-1.east) -- (m-7-1.east) -- (m-2-2.west) -- (m-7-2.west);
-	\node[lab,anchor=base east] at (m-1-2.base -| b.east) {c};
+	%\draw[>->,gray,ultra thick] (m-2-1.east) -- (m-7-1.east) -- (m-2-2.west) -- (m-7-2.west);
+	\node[lab,anchor=base east] at (m-1-1.north -| b.east) {d};
 
@@
 8\%, GDL 7.2\% (a) Evolution of the volume of the gel phase.
-(b) Evolution of the midsurface area.
-(c) Confocal XZ slices of the cell (field of view is $1\times\SI{0.1}{\milli\metre}$), colors indicate the protein concentration.
-Times of the snapshots are marked as crosses on (a) and (b).
+(b-c) Evolution of the midsurface area, straight lines are exponential fits with time constants \SI{170 \pm 7}{\second} and \SI{337 \pm 3}{\second}.
+(d) Successive confocal XZ slices of the cell (field of view is $1\times\SI{0.1}{\milli\metre}$), colors indicate the protein concentration.
+Times of the snapshots are marked as crosses on (a-c).
 Note that swelling starts \SI{14}{\minute} before wrinkling.

### 2014-05-13 plis_figures.tex
add mean position on top of figure 3

@@
 		group style={
-			group name=g, group size=1 by 2,
+			group name=g, group size=1 by 3,
 			xticklabels at=edge bottom,
@@
 		]
+	\nextgroupplot[ylabel={Position (\%)}, ymin=0, ymax=100]
+	\addplot+[no marks,Accent1] table[y expr={\thisrowno{1}*100}]{mean_alt_rel_half_cas8.txt};
+	\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{mean_alt_rel_half_cas8_toi.txt};
+	\node[lab,below left=0.2em] at (rel axis cs:1,1) {a};
+
 	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, extra x ticks={12.55}]
@@
 	\node[above right=0 of syn.south west] {swelling};
-	\node[lab,below left=0.2em] at (rel axis cs:1,1) {a};
+	\node[lab,below left=0.2em] at (rel axis cs:1,1) {b};
 
@@
 	\node[anchor=west] at (axis cs:45,0.25) {wrinkling and buckling};
-	\node[lab,above left=0.2em] at (rel axis cs:1,0) (b) {b};
+	\node[lab,above left=0.2em] at (rel axis cs:1,0) (c) {c};
 	\coordinate (wr1) at (axis cs:26.2,0);
@@
 		name={zoom},
-		at={(g c1r2.below south west)},
+		at={(g c1r3.below south west)},
 		anchor=above north west,
@@
 		\addplot+[black, no marks,domain=25:45] {10*exp(x/5.60-8.81)};
-		\node[lab,above left=0.2em] at (rel axis cs:1,0) (c) {c};
+		\node[lab,above left=0.2em] at (rel axis cs:1,0) (d) {d};
 	\end{semilogyaxis}
@@
 
-	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, column sep=0.02\textwidth, row sep=2pt] at (g c1r2.outer south east) (m){
+	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, column sep=0.02\textwidth, row sep=2pt] at (g c1r3.outer south east) (m){
 	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t000.png}\\
@@
 	%\draw[>->,gray,ultra thick] (m-2-1.east) -- (m-7-1.east) -- (m-2-2.west) -- (m-7-2.west);
-	\node[lab,anchor=base east] at (m-1-1.north -| b.east) {d};
+	\node[lab,anchor=base east] at (m-1-1.north -| c.east) {e};
 
@@
 	\caption{Synæresis and swelling, cas.
-8\%, GDL 7.2\% (a) Evolution of the volume of the gel phase.
-(b-c) Evolution of the midsurface area, straight lines are exponential fits with time constants \SI{170 \pm 7}{\second} and \SI{337 \pm 3}{\second}.
-(d) Successive confocal XZ slices of the cell (field of view is $1\times\SI{0.1}{\milli\metre}$), colors indicate the protein concentration.
-Times of the snapshots are marked as crosses on (a-c).
+8\%, GDL 7.2\% (a) Average relative altitude, (b) volume (c-d) midsurface area of the gel phase.
+Straight lines are exponential fits with time constants \SI{170 \pm 7}{\second} and \SI{337 \pm 3}{\second}.
+(e) Successive confocal XZ slices (field of view is $1\times\SI{0.1}{\milli\metre}$), colors indicate the protein concentration.
+Times of the snapshots are marked as crosses on (a-d).
 Note that swelling starts \SI{14}{\minute} before wrinkling.

### 2014-05-14 2014_05_15_discussion.tex
Modèle des cloques

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[francais]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{kpfonts}

\usepackage{siunitx}

\author{Mathieu Leocmach}
\begin{document}

\begin{figure}
\includegraphics[width=\textwidth]{plots_0117.png}
\caption{Un exemple de cloque en train de croitre (premier plan). Au fond : une cloque qui a touché une paroi. La barre fait \SI{1}{\milli\metre}.}
\end{figure}

On considère une cloque de diamètre $\lambda$ dont le contour touche la paroi et dont le sommet s'est élevé d'une hauteur $A(t)$ à l'instant $t$. On considère que $A \ll \lambda$, donc l'aire reste $\lambda^2$.

La courbure $\kappa$ est de l'ordre de $A/\lambda^2$ et l'énergie de courbure est
\begin{equation}
U_B \sim B  \kappa^2  \lambda^2 \sim B  \left(\frac{A}{\lambda}\right)^2
\end{equation}
avec $B$ le module de flexion $B \sim E h^3 /12 / (1-nu^2) \sim G' h^3/10$

La déflexion du film crée une différence de pression $P$ entre le dessous et le dessus. Le travail associé est $W \sim P  A  \lambda^2$

Si on égale $U_B$ et $W$ (je dois faire une hypothèse quand je fais ça, mais je n'arrive pas à mettre le doigt dessus), on trouve $P \sim BA/\lambda^4$

Une autre équation vient de la loi de Darcy qui donne le débit volumique $v$ à travers un poreux 
\begin{equation}
\vec{v} = - \frac{l_p^2}{\eta}  \vec{\nabla} P
\label{eq:Darcy}
\end{equation}
avec $l_p$ la taille des pores, $\eta$ la viscosité du solvant et $\vec{\nabla} P \sim -P/h$.

Le volume sous le film est $V \sim A \lambda^2$ et sa dérivée temporelle est
\begin{equation}
\frac{dV}{dt} \sim \lambda^2  \frac{dA}{dt}
\label{eq:var_vol}
\end{equation}
qui s'exprime aussi en fonction du débit à travers la surface
\begin{equation}
\frac{dV}{dt} \sim \lambda^2  v
\label{eq:debit}
\end{equation}

En utilisant les équations (\ref{eq:Darcy}), (\ref{eq:var_vol}) et (\ref{eq:debit}), on obtient
\begin{equation}
\frac{dA}{dt} \sim v \sim \frac{l_p^2  P}{eta * h}
\end{equation}

Et en réutilisant l'expression de la pression ci-dessus, on trouve
\begin{equation}
\frac{dA}{dt} \sim \frac{A}{\tau}
\end{equation}
donc une croissance exponentielle de temps caractéristique
\begin{equation}
\tau = \frac{\eta h  \lambda^4} {B l_p^2}
\end{equation}

Application numérique :
\begin{align}
\lambda \sim & \SI{1}{\milli\metre}\\
h \sim & \SI{50}{\micro\metre}\\
G^\prime \sim & \SI{500}{\pascal}\\
B \sim & 50 \times (50\cdot 10^{-6})^3 \sim \SI{6e-12}{\joule}\\
\tau \sim & \frac{10^{-3} \times 50\cdot 10^{-6} \times (10^{-3})^4}{6 \cdot 10^{-12} \times (4 \cdot 10^{-6})^{-2}} \sim \SI{500}{\second}
\end{align}

\end{document}

### 2014-05-15 2014_05_15_discussion.tex
Analyse plus correcte suivant Vella PRL 2009

@@
 
-On considère une cloque de diamètre $\lambda$ dont le contour touche la paroi et dont le sommet s'est élevé d'une hauteur $A(t)$ à l'instant $t$.
+On considère une cloque de diamètre $\lambda$ dont le contour touche la paroi inférieure et dont le sommet s'est élevé d'une hauteur $A(t)$ à l'instant $t$.
 On considère que $A \ll \lambda$, donc l'aire reste $\lambda^2$.
+
+En généralisant Vella PRL 2009, on considère $\sigma$ une contrainte s'exerçant uniformément sur la cloque suivant $z$. Minimiser l'énargie revient à minimiser la fonctionnelle
+\begin{equation}
+\mathcal{F} = \int_0^{\lambda/2}\left\lbrace\frac{1}{2} B \theta_s^2 + \sigma z - \Sigma \left[1-\frac{\Delta \lambda}{\lambda} - \cos\theta\right] \right\rbrace ds
+\end{equation}
+avec $s$ l'abcisse curviligne, $\theta(s)$ l'angle avec l'horizontale, $B$ le module de flexion $B \sim E h^3 /12 / (1-\nu^2) \sim G' h^3/10$.
+
+Une longueur caractéristique apparaît :
+\begin{equation}
+\lambda^* = \left(\frac{B}{\sigma}\right)^\frac{1}{3}
+\end{equation}
+au dessus de laquelle $\sigma$ est capable d'écraser la cloque.
+
+\paragraph*{Si $\sigma = \Delta\rho g h$}, c'est une longueur élastogravitaire
+
+Dans notre cas $\Delta\rho\simeq 10\%w = \SI{100}{\kilo\gram\per\metre^3}$, d'où
+\begin{equation}
+\lambda^* \sim \left(\frac{\num{6e-12}}{100\times 10 \num{50e-6}}\right)^\frac{1}{3} \sim \SI{5}{\centi\metre}
+\end{equation}
+Donc la gravité n'influe pas, ce qui est confirmé par un échantillon vertical plissant.
+
+\paragraph*{Si $\sigma$ est une pression de Darcy $P$} telle que 
+\begin{equation}
+\vec{v} = - \frac{l_p^2}{\eta}  \vec{\nabla} P
+\label{eq:Darcy}
+\end{equation}
+avec $v$ le flux volumique à travers la surface, c'est-à-dire la vitesse de déplacement du film, $l_p$ la taille des pores, $\eta$ la viscosité du solvant et $\vec{\nabla} P \sim -P/h$. Alors,
+\begin{equation}
+\lambda^* \sim \left(\frac{B l_p^2}{\eta v h}\right)^\frac{1}{3}
+\end{equation}
+
+On a mesuré $v\simeq \SI{0.2}{\micro\metre\per\second}$, ce qui donne 
+\begin{equation}
+\lambda^* \sim \left(\frac{\num{6e-12} (\num{4e-6})^2}{10^{-3} \times \num{0.2e-6} \times \num{50e-6}}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}
+\end{equation}
+
+Il manque une équation qui nous permettrait d'obtenir $v$ ou $\lambda$ à partir des données du problème.
+
+\pagebreak
 
@@
 \end{equation}
-avec $B$ le module de flexion $B \sim E h^3 /12 / (1-nu^2) \sim G' h^3/10$
+avec $B$ le module de flexion $B \sim E h^3 /12 / (1-\nu^2) \sim G' h^3/10$
 

### 2014-05-21 2014_05_15_discussion.tex
- la raison pour laquelle on a lambda proportionnel à l'épaisseur, avec une justification du coefficient.  - le cas où la position initiale du gel est au milieu du gap et non contre une parois. Du coup on a deux écoulements possibles, l'un à travers le poreux, l'autre à travers les films d'eau. De façon pas intuitive du tout, c'est bien le poreux qui domine.

@@
 
+\usepackage{pgfplots}
+
 \author{Mathieu Leocmach}
 \begin{document}
+
+\section{Gel en contact avec le fond}
 
@@
 
-En généralisant Vella PRL 2009, on considère $\sigma$ une contrainte s'exerçant uniformément sur la cloque suivant $z$. Minimiser l'énargie revient à minimiser la fonctionnelle
+En généralisant Vella PRL 2009, on considère $\sigma$ une contrainte s'exerçant uniformément sur la cloque suivant $z$. Minimiser l'énergie revient à minimiser la fonctionnelle
 \begin{equation}
@@
 \lambda^* = \left(\frac{B}{\sigma}\right)^\frac{1}{3}
+\label{eq:lstar}
 \end{equation}
@@
 \begin{equation}
-\vec{v} = - \frac{l_p^2}{\eta}  \vec{\nabla} P
+\vec{v} = - \frac{\ell_p^2}{\eta}  \vec{\nabla} P
 \label{eq:Darcy}
 \end{equation}
-avec $v$ le flux volumique à travers la surface, c'est-à-dire la vitesse de déplacement du film, $l_p$ la taille des pores, $\eta$ la viscosité du solvant et $\vec{\nabla} P \sim -P/h$. Alors,
+avec $v$ le flux volumique à travers la surface, c'est-à-dire la vitesse de déplacement du film, $\ell_p$ la taille des pores, $\eta$ la viscosité du solvant et $\vec{\nabla} P \sim -P/h$. Alors,
 \begin{equation}
-\lambda^* \sim \left(\frac{B l_p^2}{\eta v h}\right)^\frac{1}{3}
+\lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3}
 \end{equation}
 
-On a mesuré $v\simeq \SI{0.2}{\micro\metre\per\second}$, ce qui donne 
+On a mesuré $v\simeq \SI{0.2}{\micro\metre\per\second}$, et $\ell_p=\SI{4}{\micro\metre}$, ce qui donne 
 \begin{equation}
@@
 
-Il manque une équation qui nous permettrait d'obtenir $v$ ou $\lambda$ à partir des données du problème.
+D'après Kolinski PRL 2009, aux faibles déformations la longueur d'onde effective est $\lambda \sim \epsilon^{1/7} \lambda^*$ avec $\epsilon$ le gain relatif de longueur du film.
+L'amplitude, elle, s'écrit $A \sim \epsilon^{4/7} \lambda^*$.
+On remarque que la base de la cloque n'évolue plus après que son sommet ait touché le plafond, c'est à dire que la longueur d'onde est sélectionnée quand $A=e-h\equiv H$.
+
+Ca nous donne $\lambda \sim \epsilon^{-3/4} H$. On retrouve ainsi la proportionalité entre la longueur d'onde et l'épaisseur de la cellule.
+\begin{equation}
+\epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-4/3} \simeq \left(15\times 2\right)^{-4/3} \simeq 1\%
+\end{equation}
+Ce qui correspond tout à fait aux mesures expérimentales de la racine carrée de l'excès d'aire au moment où les cloques se forment.
+
+On avait remarqué que lorsque des cloques se forment, la longueur d'onde est environ moitié plus petite que lorsque des plis se forment de façon homogène.
+Si les plis se forment bien à $\lambda^*$ (voir plus bas), les cloques ont une taille de $\epsilon^{1/7} \lambda^*$, or avec la valeur ci-dessus, on trouve $\epsilon^{1/7}\simeq 0.5$.
 
@@
 
-La courbure $\kappa$ est de l'ordre de $A/\lambda^2$ et l'énergie de courbure est
+\section{Gel au milieu de la cellule}
+
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth, height=0.25\textwidth, scale only axis,
+	domain=0:360, no markers, ymin=-4, ymax=4,xmin=0,xmax=360,
+	axis y line=none, xtick=\empty,
+	]
+\addplot+[orange]{sin(x)+0.5};
+\addplot+[orange]{sin(x)-0.5};
+\addplot+[dashed, black]{0.5};
+\addplot+[dashed, black]{-0.5};
+\draw[<->] (axis cs:180,0.5) -- (axis cs:180,4) node[midway, left] {$H$}; 
+\draw[<->] (axis cs:30,-0.5) -- (axis cs:30,-4) node[midway, left] {$H$};
+\node[below] at (axis cs:90,0.5) (pb1) {$P_1$};
+\node[above] at (axis cs:90,1.5) (ph2) {$P_2$};
+\node[above] at (axis cs:270,-0.5) (ph1) {$P_1$};
+\node[below] at (axis cs:270,-1.5) (pb2) {$P_2$};
+\draw[thick, ->] (ph2) -- (pb1);
+\draw[thick, ->] (pb2) -- (pb1);
+\end{axis}
+\end{tikzpicture}
+\end{figure}
+
+On considère maintenant une couche de gel d'épaisseur $h$ centrée à mi-hauteur de la cellule.
+La distance moyenne entre une surface du film et la paroi la plus proche est notée $H$.
+On considère une perturbation sinusoïdale.
+
+Par symmétrie, la différence de pression à travers le poreux est la même qu'entre un minimum et un maximum d'altitude du film.
+Sur cette demi-longueur d'onde, on suppose un écoulement de Poiseuille dans la couche d'eau d'épaisseur $H$.
+
 \begin{equation}
-U_B \sim B  \kappa^2  \lambda^2 \sim B  \left(\frac{A}{\lambda}\right)^2
+\Delta P \sim 12\eta\frac{\lambda}{H^2} u
 \end{equation}
-avec $B$ le module de flexion $B \sim E h^3 /12 / (1-\nu^2) \sim G' h^3/10$
-
-La déflexion du film crée une différence de pression $P$ entre le dessous et le dessus. Le travail associé est $W \sim P  A  \lambda^2$
-
-Si on égale $U_B$ et $W$ (je dois faire une hypothèse quand je fais ça, mais je n'arrive pas à mettre le doigt dessus), on trouve $P \sim BA/\lambda^4$
-
-Une autre équation vient de la loi de Darcy qui donne le débit volumique $v$ à travers un poreux 
+où $u$ est la valeur moyenne de la vitesse suivant $x$. Dans l'hypothèse de lubrification, la vitesse suivant $z$ est $v\sim \frac{H}{\lambda} u$, donc
 \begin{equation}
-\vec{v} = - \frac{l_p^2}{\eta}  \vec{\nabla} P
-\label{eq:Darcy}
-\end{equation}
-avec $l_p$ la taille des pores, $\eta$ la viscosité du solvant et $\vec{\nabla} P \sim -P/h$.
-
-Le volume sous le film est $V \sim A \lambda^2$ et sa dérivée temporelle est
-\begin{equation}
-\frac{dV}{dt} \sim \lambda^2  \frac{dA}{dt}
-\label{eq:var_vol}
-\end{equation}
-qui s'exprime aussi en fonction du débit à travers la surface
-\begin{equation}
-\frac{dV}{dt} \sim \lambda^2  v
-\label{eq:debit}
+\Delta P \sim 12\eta\frac{\lambda^2}{H^3} v
 \end{equation}
 
-En utilisant les équations (\ref{eq:Darcy}), (\ref{eq:var_vol}) et (\ref{eq:debit}), on obtient
+En injectant ce résultat dans l'équation (\ref{eq:lstar}), on obtient la longueur au delà de laquelle l'écoulement de Poiseuille interdit la déformation :
 \begin{equation}
-\frac{dA}{dt} \sim v \sim \frac{l_p^2  P}{eta * h}
+\lambda^*_\text{Poiseuille} \sim \left(\frac{B H^3}{12\eta v}\right)^\frac{1}{5}
 \end{equation}
 
-Et en réutilisant l'expression de la pression ci-dessus, on trouve
+En prenant $H=\SI{30}{\micro\metre}$, on obtient
 \begin{equation}
-\frac{dA}{dt} \sim \frac{A}{\tau}
+\lambda^*_\text{Poiseuille} \sim \left(\frac{\num{6e-12} (\num{30e-6})^3}{12 \times
+10^{-3} \times \num{0.2e-6}}\right)^\frac{1}{5} \simeq \SI{2}{\milli\metre}
 \end{equation}
-donc une croissance exponentielle de temps caractéristique
+Ce qui est comparable à la longueur d'onde de Darcy pour notre expérience typique (caséine 4\%, GDL 4\% dans l'eau pure à \SI{21}{\celsius}).
+
+Pour une longueur d'onde donnée, les vitesses de croissances respectives sont
+\begin{align}
+v_\text{Darcy} &\sim \frac{B \ell_p^2}{\eta h \lambda^3}\\
+v_\text{Poiseuille} &\sim \frac{B H^3}{12\eta \lambda^5}
+\end{align}
+D'où leur rapport
 \begin{equation}
-\tau = \frac{\eta h  \lambda^4} {B l_p^2}
+\frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3}
 \end{equation}
 
-Application numérique :
-\begin{align}
-\lambda \sim & \SI{1}{\milli\metre}\\
-h \sim & \SI{50}{\micro\metre}\\
-G^\prime \sim & \SI{500}{\pascal}\\
-B \sim & 50 \times (50\cdot 10^{-6})^3 \sim \SI{6e-12}{\joule}\\
-\tau \sim & \frac{10^{-3} \times 50\cdot 10^{-6} \times (10^{-3})^4}{6 \cdot 10^{-12} \times (4 \cdot 10^{-6})^{-2}} \sim \SI{500}{\second}
-\end{align}
+On note que ce rapport est indépendant des propriétés viscoélastiques du système ($B$ et $\eta$) et est donc pûrement géométrique.
+La croissance des modes de Darcy est plus rapide quand les pores sont gros, la longueur d'onde importante et le système fin.
+
+Pour les deux concentrations où on connaît la taille des pores (4\% de caséine à \SI{21}{\celsius}, on suppose $h=\SI{40}{\micro\metre}$ et $H=\SI{30}{\micro\metre}$ dans les deux cas) on a les rapports de vitesses :
+\begin{itemize} 
+\item 4\% de GDL, dans l'eau pure : 37
+\item 8\% de GDL, dans 60\% de glycérol : 75
+\end{itemize}
+Le mode de Darcy semble donc être cinétiquement favorisé dans (la pluspart de) nos expériences.
 

### 2014-05-23 poster_Okinawa.tex
Poster pour Okinawa : change le titre, les auteurs et les logos

--- 2013-09-10 poster_Rome.tex
@@
 \hypersetup{%
-	pdftitle={Wrinkling and yielding of protein gels},%the title
+	pdftitle={Wrinkling and nested buckling in a confined protein gel},%the title
 	pdfauthor={Mathieu Leocmach},%your name
@@
 \Title{
-Wrinkling and yielding of protein gels
+Wrinkling of a confined porous layer
 }
 
-\LHead{Mathieu Leocmach}\qquad\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}\\
-\LHead{Mathieu Nespoulous, Christophe Perge, Nicolas Taberlet, Thomas Gibaud,
-Thibaut Divoux, Sébastien Manneville}
+\smallskip
+\LHead{Mathieu Leocmach, Mathieu Nespoulous, Thomas Gibaud, Sébastien Manneville
+}\hfill\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}\\
 
-\LHead{\raisebox{-0.3em}{\includegraphics[height=\baselineskip]{CNRSfilaire-Q}\qquad\includegraphics[height=\baselineskip]{logo_UCBL}\qquad}\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}\raisebox{-0.3em}{\qquad\includegraphics[height=\baselineskip]{logo_ens-lyon}}}
+\LHead{\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}\hfill\raisebox{-1.5\baselineskip}{\includegraphics[height=2\baselineskip,clip=true, trim=0 10mm 0 0]{NEW-Logo-ERC-OUTLINE}\qquad\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\qquad\includegraphics[height=2\baselineskip]{logo_ens-lyon}}}
 

### 2014-05-23 poster_Okinawa.tex
Logo UMS (PLATIM), ERC agreement en petit en bas

@@
 
-\LHead{\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}\hfill\raisebox{-1.5\baselineskip}{\includegraphics[height=2\baselineskip,clip=true, trim=0 10mm 0 0]{NEW-Logo-ERC-OUTLINE}\qquad\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\qquad\includegraphics[height=2\baselineskip]{logo_ens-lyon}}}
+\LHead{\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}\hfill\raisebox{-1.5\baselineskip}{
+	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
+	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
+	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
+	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
+}}
 
@@
 
-\begin{textblock}{4}(0,16)
-	\Head{Composition}
-	\Subhead{Less gdl}
-	\begin{itemize}
-		\item slower kinetics
-		\item stronger gel
-		\item weaker overshoot
-		\item same fundamental \linebreak wavelength
-		\item less harmonics
-	\end{itemize}
-	\Subhead{More gdl}
-	\begin{itemize}
-		\item plasticity
-	\end{itemize}
-\end{textblock}
-\begin{textblock}{2}(2,16)
-	\addvspace{2.4em}
-	bottom surface, \textsc{gdl} 5\%\\
-	\begin{tikzpicture}[inner sep=0]
-		\draw[ultra thick] (0,0) -- ++(1.11\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{500}{\micro\metre}};
-	\end{tikzpicture}
-
-	\includegraphics[width=2\TPHorizModule]{plastic_confocal}
-\end{textblock}
-
-\begin{textblock}{4}(0,20)
-	\Head{Thickness}
-	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccb{0.45\textwidth}}
-	\SI{50}{\micro\metre} & \SI{100}{\micro\metre}&
-	\multirow{4}{*}{\begin{tikzpicture}
-	\node[inner sep=0] (a) {\includegraphics[width=0.45\textwidth]{pattern_800um_x1.jpg}};
-	\node[above=0 of a] {\SI{800}{\micro\metre}};
-	\draw[ultra thick](a.north west) ++(0,2em) -- ++(0.2\textwidth,0) node[midway, above, inner sep=0] {\SI{5}{\milli\metre}};
-	\draw[ultra thick, Accent2, <->] (a.south west) ++(0.18\textwidth, 0.05\textwidth) -- ++(0.133\textwidth,0) node[midway,above]{$\lambda$};
-	\end{tikzpicture}}\\
-
-	\includegraphics[width=0.225\textwidth]{pattern_50um.jpg} & \includegraphics[width=0.225\textwidth]{pattern_100um.jpg}&
-	\\
-
-	\SI{250}{\micro\metre} & \SI{450}{\micro\metre} &\\
-
-	\includegraphics[width=0.225\textwidth]{pattern_250um.jpg} & \includegraphics[width=0.225\textwidth]{pattern_450um.jpg}& \\
-	\end{tabular*}
-
-	\vspace{1em}
-
-	\begin{tikzpicture}
-	\begin{axis}[%
-		width=0.95\textwidth,
-		height=0.4\textwidth,
-		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
-		xtick={0.1,0.3,...,0.9},
-		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
-		xlabel near ticks,
-		ylabel near ticks,
-		%legend style={legend pos=north west,}
-		]
-		\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
-		\addplot+[domain=0:3, no marks, Accent1, forget plot] {4.23*x};
-	\end{axis}
-	\end{tikzpicture}
-\end{textblock}
-
-\begin{textblock}{4}(5,16)
-	\Head{Creep and fractures}
-	Couette cell, no brush, the gel sticks, no pattern.\\
-	Wait \SI{8}{\hour} for equilibration, then apply constant stress.
-
-	\begin{tikzpicture}
-	\begin{loglogaxis}[%
-		name=a,
-		xlabel={$t\,(s)$},
-		ylabel={Shear rate $(s^{-1})$},
-		]
-		\addplot+[no marks, Accent2] file {Y34_strain_rate_log.txt};
-		\draw[<->] (axis cs:1e-1,1) -- (axis cs:1.3e3,5e-4) node[midway,above,rotate=-21] {Andrade} node[pos=0.25,below=1em] {$t^{-0.86}$};
-		\node at (axis cs:1.3e3,1e-4) (det){};
-		\node at (axis cs:2.3e3,1e4) (touch){};
-	\end{loglogaxis}
-	\draw[<-, Accent1, ultra thick] (det) +(0,0.1em) -- ++(0,-4em) node[left] {Fractures appear} -- ++(1em,0);
-	\draw[<-, Accent1, ultra thick] (touch) +(0,-0.1em) -- ++(0,3em) node[left] {Fractures meet} -- ++(1em,0);
-	\end{tikzpicture}\hfill
-	\begin{tikzpicture}[inner sep=0]
-		\node[above right] at (0,1em) {\includegraphics[width=1.5\TPHorizModule, trim=0 6mm 2cm 1cm,clip=true]{Y210_SSC300_avant_div.jpg}};
-		\node[below right] (b){\includegraphics[width=1.5\TPHorizModule, trim=0 6mm 2cm 1cm,clip=true]{Y210_SSC300_premiere_frac.jpg}};
-		\node[below left=0 of b.north east]{\includegraphics[width=0.75\TPHorizModule]{detail_premiere_frac.jpg}};
-		\draw[Accent2,radius=0.07\TPHorizModule] ++(b.north east) +(-1.14\TPHorizModule,-1.05\TPHorizModule) circle;
-		\draw[Accent2,radius=0.09\TPHorizModule] ++(b.north east) +(-0.435\TPHorizModule,-0.375\TPHorizModule) circle;
-		\draw[ultra thick] (a.south west) ++(0,0.25em) -- ++(0.5\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{2}{\centi\metre}};
-		\draw[ultra thick, Accent2, <->] (a.south west) ++(0.51\TPHorizModule, \TPHorizModule) -- ++(0.09\TPHorizModule,0) node[midway,above=0.25em,font=\small]{$\lambda$};
-	\end{tikzpicture}
-
-	\begin{tabular}{b{0.55\textwidth}b{0.45\textwidth}}
-	\begin{tikzpicture}
-	\begin{loglogaxis}[%
-		ylabel={Breaking time (s)},ylabel near ticks,
-		xlabel={Stress $\sigma$ (Pa)},
-		xmin=10,
-		legend style={font=\small\sc}, legend pos=south west,
-		]
-		\addplot+[only marks, Accent1, mark options={fill=Accent1}] coordinates {(40, 67225) (50,29860) (100,398) (200,4) (300,0.06)};
-		\addplot+[only marks, Accent2, mark options={fill=Accent2}] table[x index=0, y index=1]{cas4_GDL1_t_cassure_sigma.txt};
-		\legend{gdl 4\%, gdl 1\%};
-		\addplot+[Accent1, no marks, forget plot, domain=40:300] gnuplot {9.2e+15*x^{-6.8}};
-		\addplot+[Accent2, no marks, forget plot, domain=150:1000] gnuplot {3.93e+17*x^{-5.44}};
-	\end{loglogaxis}
-	\end{tikzpicture}&
-	Irreversible fractures\linebreak
-	\Subhead{Fracture\linebreak wavelength}
-	\begin{itemize}
-		\item stress-independent
-		\item composition-indep.
-		\item $\propto$ gap
-	\end{itemize}\\
-	\end{tabular}
-
-	\begin{tikzpicture}
-	\begin{axis}[%
-		width=0.95\textwidth,
-		height=0.4\textwidth,
-		xmin=0,xmax=3.5,xlabel={gap (\si{\milli\metre})},
-		xtick={0, 0.5,...,3},
-		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
-		xlabel near ticks,
-		ylabel near ticks,
-		legend cell align=left,
-		legend style={font=\small}, legend pos=north west,
-		]
-		\addplot+[%
-			only marks, Accent1, mark options={fill=Accent1},
-			error bars/.cd,
-			y dir=both,y explicit,
-			x dir=none,x fixed=0.05,
-			]
-			coordinates {
-				(0.5,0.6) +-(0,0.12)
-				(1, 1.33) +-(0,0.14)
-				(1.5,1.5) +-(0,0.14)
-				(1.5, 1.25) +- (0,0.77)
-				(3,3.75) +- (0,0.4)
-				(3, 4.2) +-(0,0.4)};
-		\addplot+[%
-			only marks, Main, mark options={fill=Main},
-			error bars/.cd,
-			y dir=both,y explicit,
-			x dir=none,x fixed=0.05,
-			]
-			coordinates
-			{(2.5,3.2) +- (0,0.84)};
-		\addplot+[%
-			only marks, Accent2, mark options={fill=Accent2}, mark=triangle,
-			error bars/.cd,
-			y dir=both,y explicit,
-			x dir=none,x fixed=0.05,
-			]
-			coordinates
-			{(2,2.29) +- (0,0.13)};
-		\legend{$R_\text{out}=\SI{25}{\milli\metre}$, $R_\text{out}=\SI{37.5}{\milli\metre}$, $R_\text{out}=\SI{40}{\milli\metre}$};
-		\addplot+[domain=0:3.5, no marks, Accent1, forget plot] {1.25*x};
-	\end{axis}
-	\end{tikzpicture}
-\end{textblock}
-
-\begin{textblock}{5}(10,16)
-	\Head{Ultrasound speckles}
-	\Subhead{Probe deformation inside the gap}
-
-	\begin{tikzpicture}[ultra thick, font=\small]
-	\node[above right, inner sep=0] (sch) {\includegraphics[width=0.4\textwidth]{usv_3D_image}};
-	\draw[<-] (0.3\textwidth,0.11\textwidth) -- +(1em,0) node[right] {water in};
-	\draw[->] (0.3\textwidth,0.07\textwidth) -- +(1em,0) node[right] {water out};
-	\node[right] at (0.2\textwidth, 0.3\textwidth) (rh) {rheometer};
-	\node[above right] at (0.2\textwidth, 0) (arraylabel) {array of transducers};
-	\draw[->,dotted]	 (arraylabel.west) -- (0.1\textwidth, 0.1\textwidth);
-	\draw[->] (0.06\textwidth, 0.15\textwidth) -- (0,0.15\textwidth) node[draw, left, text width=0.18\textwidth] (mce) {multi-channel electronics};
-	\node[above=of mce, draw] (pc) {PC};
-	\draw[->] (mce)-- (pc);
-	\draw[->] (rh) -- +(0,1em) -| (pc.north);
-	\node[above right=0 and 0.1\textwidth of sch.south east, text width=0.25\textwidth, align=right] {Gallot et al. \linebreak\textit{Rev. Sci. Instrum.} 84, 045107 (2013).};
-	\node[below right =of sch.north east, text width=0.32\textwidth] {3\% of \SI{30}{\micro\metre} acrylamide beads as contrast agent};
-	\end{tikzpicture}
-
-	\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=grp,
-			group size=2 by 1, 
-			horizontal sep=\TPHorizModule,
-		},
-		width=2\TPHorizModule,
-		xmin=0,xmax=200,
-		xlabel={t},
-		axis on top,
-		xtick={28, 100, 172},
-		xticklabels={49h30, 50h, 50h30},
-		]
-		\nextgroupplot[ymin=0, ymax=32,	ylabel={$z (\si{\milli\metre})$}, title={$r=\SI{200}{\micro\metre}$}]
-		\addplot graphics[
-			xmin=0,xmax=200,
-			ymin=0,ymax=32,
-			]{Y137_speckleBF_tz.png};
-		\addplot[no marks,Accent2, domain=0:32]  ({187-0.272*(x-17.6)^2}, x);
-		\draw[Accent1] (axis cs:187,0)--(axis cs:187,32);
-
-		\nextgroupplot[ymin=0, ymax=2,	ylabel={$r (\si{\milli\metre})$}, title={$z=\SI{8}{\milli\metre}$}]
-		\addplot graphics[
-			xmin=0,xmax=200,
-			ymin=0,ymax=2,
-			]{Y137_speckleBF_tr.png};
-			\draw[Accent2] (axis cs:172,0)--(axis cs:172,2) node[pos=0.66, left, font=\small, fill=gray, inner sep=1pt] {free from the rotor};
-			\draw[Accent1] (axis cs:187,0)--(axis cs:187,2) node[pos=0.33, left, font=\small, fill=gray, inner sep=1pt] {global yield};
-	\end{groupplot}
-	\node[left=of grp c2r1.south west] {rotor};
-	\node[left=of grp c2r1.north west] {stator};
-	\end{tikzpicture}
-
-	Fractures appear when the gel detaches from the rotor. Creates a free surface that allows synæresis, relaxation of the gelation-induced \linebreak stress and possibly wrinkling.
-\end{textblock}
-
-\begin{textblock}{5}(10,22.5)
-	\Head{Funding}
-	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lb{0.6\textwidth}}
-	\includegraphics[width=0.3\textwidth]{NEW-Logo-ERC-OUTLINE}&
-	The research leading to these results has received funding from the European Research Council under the European Union's Seventh \linebreak Framework Programme (FP7/2007-2013) /\linebreak ERC grant agreement n°~258803.\\
-	\end{tabular*}
+\begin{textblock}{15}(0,25)
+\small{\texttt{The research leading to these results has received funding from the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement n°~258803.}}
 \end{textblock}

### 2014-05-23 poster_Okinawa.tex
Split G' and pH in groupplot import zooms from plis_figure New take home message

@@
 % places on the page.
-\usepackage[absolute,overlay,%showboxes
+\usepackage[absolute,overlay,showboxes
 ]{textpos}
@@
 \def\newblock{\hskip .11em plus .33em minus .07em}
+
+\newlength{\mylength}
 
@@
 
-	In millipore water
-	\begin{itemize}
-	\item[4\%] Sodium caseinate (milk protein)\\
-	Isoelectric point $\approx 4.6$
-	\item[4\%] $\delta$-gluconolactone (\textsc{gdl})\\
-	Slow hydrolysis into gluconic acid
-	\end{itemize}
+	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (4\%w \textsc{gdl})
 
 	\begin{tikzpicture}
-	\begin{axis}[%
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=1 by 2,
+			xticklabels at=edge bottom,
+			vertical sep=0,
+			},
 		xmin=0,xmax=8,
-		ymin=0, ymax=7,
-		xlabel={time (h)}, ylabel={\textcolor{Accent1}{pH}},
+		scale only axis,
+		width=\textwidth-4em,
+		height=0.3\textwidth,
 		extra tick style={grid=major},%
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.5em}
+		]
+	\nextgroupplot[
+		ymin=0, ymax=7, ylabel={\textcolor{Accent1}{pH}},
 		extra y ticks={4.6}, extra y tick labels={},%
-		extra x ticks={0.55, 0.64, 0.72, 0.80, 1, 1.27, 2.5}, extra x tick labels={},%
-		axis y line*=left,
-		width=0.7\textwidth,
 		]
@@
 	\node[base left=0] at (axis cs:8,4.6) {isoelectric};
-	\end{axis}
-	\begin{axis}[%
-		xmin=0,xmax=8,
-		ymin=0,
-		ylabel=\textcolor{Accent2}{$G^\prime$ (\si{\pascal})},
-		axis y line*=right,
-		axis x line=none,
-		width=0.7\textwidth,
+
+	\nextgroupplot[
+		xlabel={time (h)},
+		ymin=0, ylabel=\textcolor{Accent2}{$G^\prime$ (\si{\pascal})}
 		]
 	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
-	\end{axis}
+	\end{groupplot}
 	\end{tikzpicture}
+
+For confocal microscopy we label 10\% of the casein.
 \end{textblock}
@@
 
+	%\Subhead{Side view}\\
 	\tikzsetnextfilename{cell_brushes}
@@
 		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\textwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
-		\draw[line width=2pt,Accent1] (0.05\textwidth,-2.5em) -- (0.95\textwidth,-2.5em) (0.05\textwidth,-1pt) -- (0.95\textwidth,-1pt) node[below,pos=0.30] {acrylamide brush ($10\sim 100$ nm)};
+		\draw[line width=2pt,Accent1] (0.05\textwidth,-2.5em) -- (0.95\textwidth,-2.5em) (0.05\textwidth,-1pt) -- (0.95\textwidth,-1pt) node[below,pos=0.30, text width=0.4\textwidth] {acrylamide brush ($10\sim 100$ nm)\linebreak $\Rightarrow$ No adhesion};
 		\fill[gray] (0,0) rectangle (0.05\textwidth,-2.5em) (\textwidth,0) rectangle (0.95\textwidth,-2.5em) node[pos=1, above left] {spacer};
 		\draw[<->] (0.75\textwidth,-2pt) -- (0.75\textwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
-		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] {$L\sim 2.5\,cm$};
+		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] (L){$L\sim 2.5\,cm$};
+		\node[anchor=north west, inner sep=0] at ($(L.north) - (0.5\textwidth,0)$) {\Subhead{Side view}};
 	\end{tikzpicture}
 
-	\begin{itemize}
-	\item No adhesion on top and bottom (polymer brushes).
-	\item Adhesion at the edge (solid spacers, no free interface)
-	\end{itemize}
+	\Subhead{Top view}\\
+	\begin{tikzpicture}[inner sep=0, very thick]
+	\setlength{\mylength}{\columnwidth}
+	\node[anchor=north west] (a) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2_crop}};
+		\node[anchor=north] at (0.44\mylength,0) (b) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6_crop}};
+		\node[anchor=north east] at (\mylength,0) (c) {\includegraphics[width=0.4\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
+		%zooms
+		\node[minimum width = 0.137\mylength, minimum height=0.137\mylength, anchor=north west, draw=Accent2] at ($(a.north west) +(0.111\mylength,-0.072\mylength)$) (bz){};
+		\draw[Accent2] (bz.north east) -- (b.north west) (bz.south east) -- (b.south west);
+		\node[minimum width = 0.128\mylength, minimum height=0.081\mylength, anchor=north west, draw=Main] at ($(b.north west) +(0.097\mylength,-0.143\mylength)$) (cz){};
+		\draw[Main] (cz.north east) -- (c.north west) (cz.south east) -- (c.south west);
+		\node[minimum width = 0.156\mylength, minimum height=0.156\mylength, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\mylength,0)$) (dz) {};
+		%scale bars
+		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.178\mylength,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
+		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.176\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
+		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
 
-	\begin{tikzpicture}[inner sep=0, very thick]
-	\node[anchor=west] (a) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2}};
-	\node at (3.5\TPHorizModule,0) (b) {\includegraphics{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6}};
-	\node[anchor=east] at (7\TPHorizModule,0) (c) {\includegraphics[height=65mm]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
-	\draw[Accent2] (a.north west) ++(36.4mm, -25mm) rectangle +(28.9mm,-21.7mm) -- (b.south west) (a.north west) ++(65.3mm, -25mm) -- (b.north west);
-	\draw[Main] (b.north west) ++(34mm,-34mm) rectangle +(27mm,-19.7mm) -- (c.south west) (b.north west) ++(61mm,-34mm) --(c.north west);
-	\draw[Accent1] (c.south west) ++(29mm,29mm) rectangle +(36mm,36mm);
-	\draw[ultra thick] (a.south west) ++(0,-0.25em) -- ++(30.588mm,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\centi\metre}};
-	\draw[ultra thick] (b.south west) ++(0,-0.25em) -- ++(45.774mm,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
-	\draw[ultra thick] (c.south west) ++(0,-0.25em) -- ++(30.415mm,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
+		%final volumic view
+		%\node[anchor=north east] at (\columnwidth,0) (c) {\includegraphics[width=0.4\mylength]{recol_volume_plis}};
 	\end{tikzpicture}
@@
 	\Norulehead{Take home}
-	\Subhead{A new kind of pattern}
+	\Subhead{Poroelasticity dominates}
 	\begin{itemize}
-		\item Nested
-		\item Self generated
-		\item Boundary conditions
+		\item gravity
+		\item lubrication
 	\end{itemize}
-	\Subhead{An original yielding scenario}
+	\Subhead{Two modes}
 	\begin{itemize}
-		\item Andrade creep over 6 decades in time
-		\item Irreversible fractures
+		\item blistering
+		\item wrinkling
 	\end{itemize}
-	\Subhead{A link with fractures}
+	\Subhead{Patterns}
 	\begin{itemize}
-		\item Wrinkles can induce plasticity
-		\item Wavelengths have the same dependence
-		\item Detachment creates a free surface
+		\item self-generated
+		\item nested
+		\item kinetic wavelength
 	\end{itemize}
+
 \end{textblock}%Conclusion

### 2014-05-23 poster_Okinawa.tex
clean up and titles

@@
 \begin{textblock}{15}(0,8)
-	\Head{Pattern kinetics by transmission macroscope}
-	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}p{2\TPHorizModule}llllll}
-	33 min\hfill \begin{tikzpicture}[inner sep=0]
-		\draw[ultra thick] (0,0) -- ++(0.355\TPHorizModule,0) node[pos=0.5, above=0, font=\small] {\SI{1}{\centi\metre}};
-	\end{tikzpicture}&
-	 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
-	\includegraphics[width=2\TPHorizModule]{prise_0100_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{prise_0130_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{prise_0160_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{prise_0190_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{prise_0250_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{prise_0360_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{prise_0799_resized.jpg}\\
-	\end{tabular*}
-\end{textblock}
+	\Head{Initial contact with a substrate: Poroelastic blistering}
 
-\begin{textblock}{3}(0,10)
-	\Subhead{Length scales}
-	\begin{itemize}
-	\item A fundamental that does not evolve
-	\item Nested structures as ``harmonics''
-	\end{itemize}
-	\Subhead{Qualitatively different from}
-	\begin{itemize}
-	\item Phase separation patterns 
-	\item Turing patterns
-	\item Usual wrinkles
-	\end{itemize}	 
-\end{textblock}
-
-\begin{textblock}{11}(4,11)
-	\Head{Confocal Microscopy}
-\end{textblock}
-
-\begin{textblock}{3}(4,11)
-	\addvspace{2em}
-	\Subhead{Procedure}
-	\begin{itemize}
-		\item Fluorescently label 20\% of the casein
-		\item A confocal slice every \SI{5}{\micro\metre}
-		\item $(\SI{1.3}{\milli\metre})^2 \times \SI{120}{\micro\metre}$
-		\item[$\Rightarrow$] Protein density field
-		\item Detect top and bottom surfaces of the gel phase
-	\end{itemize}
-\end{textblock}
-
-\begin{textblock}{2}(8,11)
-	\addvspace{2em}
-	\Subhead{Final state}
-
-	\begin{tikzpicture}[inner sep=0]
-		\node[above right]{\includegraphics[width=\textwidth]{collage_cut}};
-		\draw[ultra thick] ++(0,0.5em) -- ++(0.33\textwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
-	\end{tikzpicture}\\
-	Tiled, enhanced elevation
-\end{textblock}
-
-\begin{textblock}{4}(11,11)
-	\addvspace{2em}
-	\Subhead{Synæresis and swelling}
-
-	\begin{tikzpicture}
-	\begin{axis}[%
-		xmin=0, xtick={20,40,...,160},
-		ymin=0.5, ymax=1, 
-		xlabel={time (min)}, ylabel={$V_\text{gel}/V_\text{cell}$},
-		extra tick style={grid=major},%
-		extra x ticks={33, 38, 43, 48, 58, 77, 150}, extra x tick labels={},%
-		%width=2.65\TPHorizModule,
-		width=3.65\TPHorizModule,
-		%width=0.8\textwidth, 
-		height=1.5\TPVertModule,
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60.+3}]{cas4_GDL4_fluo0p4_2.volume};
-	\end{axis}
-	\end{tikzpicture}
 \end{textblock}
@@
 \begin{textblock}{15}(0,14)
-	\Head{Wrinkling kinetics by confocal microscope}
-	\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}lllllll}
-	\raisebox{-0.75em}{\begin{tikzpicture}[inner sep=0]
-		\node[above right]{\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}};
-		\draw[ultra thick] ++(0,0.02\TPHorizModule) -- ++(1.55\TPHorizModule,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
-	\end{tikzpicture}}&
-	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
-	\includegraphics[width=2\TPHorizModule]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
-\end{tabular*}
+	\Head{No initial contact with a substrate: Poiseuille vs Darcy on a flying carpet}
+	\begin{tikzpicture}
+	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em] (m){
+	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
+	\includegraphics[width=0.13\textwidth]{prise_0100_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0130_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0160_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0190_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0250_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0360_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0799_color.jpg}\\
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
+	};
+	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.023\textwidth,0);
+	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
+	\end{tikzpicture}
 \end{textblock}
@@
 	\end{itemize}
-	\Subhead{Patterns}
+	\Subhead{Nested patterns}
 	\begin{itemize}
-		\item self-generated
-		\item nested
+		\item generated by swelling
 		\item kinetic wavelength

### 2014-05-26 2014_05_15_discussion.tex
XZ slices of blisters. Good volumic image of forming blister Volume and area of blisters Blister sketch , remarks and equations

@@
 \begin{equation}
-\lambda^* \sim \left(\frac{\num{6e-12}}{100\times 10 \num{50e-6}}\right)^\frac{1}{3} \sim \SI{5}{\centi\metre}
+\lambda^* \sim \left(\frac{\num{6e-12}}{100\times 10 \times \num{50e-6}}\right)^\frac{1}{3} \sim \SI{0.4}{\milli\metre}
 \end{equation}
-Donc la gravité n'influe pas, ce qui est confirmé par un échantillon vertical plissant.
+Donc la gravité pourrait influer, ce qui est infirmé par un échantillon vertical plissant.
 
@@
 
-Ca nous donne $\lambda \sim \epsilon^{-3/4} H$. On retrouve ainsi la proportionalité entre la longueur d'onde et l'épaisseur de la cellule.
+Ca nous donne $\lambda \sim \epsilon^{-3/7} H$. On retrouve ainsi la proportionalité entre la longueur d'onde et l'épaisseur de la cellule.
 \begin{equation}
-\epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-4/3} \simeq \left(15\times 2\right)^{-4/3} \simeq 1\%
+\epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-4/3} \simeq \left(15\times 2\right)^{-7/3} \simeq \num{3e-4}
 \end{equation}
-Ce qui correspond tout à fait aux mesures expérimentales de la racine carrée de l'excès d'aire au moment où les cloques se forment.
+Ce qui correspond tout à fait aux mesures expérimentales de l'excès d'aire au moment où les cloques se forment.
 
 On avait remarqué que lorsque des cloques se forment, la longueur d'onde est environ moitié plus petite que lorsque des plis se forment de façon homogène.
-Si les plis se forment bien à $\lambda^*$ (voir plus bas), les cloques ont une taille de $\epsilon^{1/7} \lambda^*$, or avec la valeur ci-dessus, on trouve $\epsilon^{1/7}\simeq 0.5$.
+Si les plis se forment bien à $\lambda^*$ (voir plus bas), les cloques ont une taille de $\epsilon^{1/7} \lambda^*$, or avec la valeur ci-dessus, on trouve $\epsilon^{1/7}\simeq 0.3$.
 

### 2014-05-26 poster_Okinawa.tex
XZ slices of blisters. Good volumic image of forming blister Volume and area of blisters Blister sketch , remarks and equations

@@
 		extra tick style={grid=major},%
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.5em}
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
 		]
@@
 \begin{textblock}{15}(0,8)
-	\Head{Initial contact with a substrate: Poroelastic blistering}
+	\Head{Initial flat contact with a substrate: Poroelastic blistering}
 
+\end{textblock}
+
+\begin{textblock}{3}(0,8.5)
+	\begin{tikzpicture}
+	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png}\\
+	};
+	\begin{scope}[above left, white, fill=gray]
+		\node at (m-1-1.south east) {\SI{15}{\minute}};
+		\node at (m-2-1.south east) {\SI{22}{\minute}};
+		\node at (m-3-1.south east) {\SI{1}{\hour}~26};
+		\node at (m-4-1.south east) {\SI{1}{\hour}~30};
+		\node at (m-5-1.south east) {\SI{1}{\hour}~31};
+		\node at (m-6-1.south east) {\SI{1}{\hour}~35};
+		\node at (m-7-1.south east) {\SI{2}{\hour}~25};
+		\node at (m-8-1.south east) {\SI{2}{\hour}~40};
+	\end{scope}
+	\draw[<->, ultra thick] ($(m-8-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
+	\end{tikzpicture}
+
+	\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=1 by 2,
+			xticklabels at=edge bottom,
+			vertical sep=0,
+			},
+		xmin=0, xmax=170, xtick={0,30,...,150},
+		extra tick style={grid=major},%
+		extra x tick labels={},%
+		scale only axis,
+		width=\textwidth-4em,
+		height=0.3\textwidth,
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
+		]
+
+	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em}]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_cloques.txt};
+	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
+
+	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=2, ytick={0,0.5,1,1.5}, xlabel={time (min)}, every axis y label/.append style={xshift=-1em}]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt};
+	%\addplot+[only marks,Accent2, mark=+] table{excess_area_pc_half_cas8_toi.txt};
+	\end{groupplot}
+	\end{tikzpicture}
+
+\end{textblock}
+
+\begin{textblock}{3}(4,8.5)
+\includegraphics[width=\textwidth]{volume_cloque_t116.jpg}
+
+\begin{tikzpicture}[ultra thick]
+\begin{axis}[
+	name=a,
+	width=\textwidth, height=0.25\textwidth, scale only axis,
+	domain=-180:180, no markers, ymin=0, ymax=4,xmin=-180,xmax=180,
+	axis lines=none, xtick=\empty,
+	]
+	\addplot+[Accent1]{cos(x)+1};
+	\addplot+[Accent1]{cos(x)+1.5};
+	\draw[->, ultra thick] (axis cs:90,1.25) -- +(45:0.05\textwidth) node[pos=1, above] {$\vec{v}$};
+	\node[above] at (axis cs:-90,1.5) {$P_2$};
+	\node[below right] at (axis cs:-90,1.25) {$P_1$};
+\end{axis}
+\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle (\textwidth,-1em);
+\draw[<->] ($(a.south west)+(0,-1.1em)$) -- +(\textwidth,0) node[midway, below] {$\lambda$};
+\end{tikzpicture}
+
+\begin{itemize}
+\item Synæresis then swelling
+\item First blister after $\approx\SI{1}{\hour}$ of swelling
+\item Independent blisters
+\item $\lambda \ll L \Rightarrow$ not buckling
+\item Once in contact with ceiling, $\lambda$ fixed
+\item Cascade buckling\linebreak \hfill\textit{Roman \& Pocheau 1999}
+\end{itemize}
+\end{textblock}
+
+\begin{textblock}{3}(8,8.5)
+Bending energy $B$ vs vertical stress $\sigma$:
+\[ \lambda^* = \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]
+\textit{Kolinski et al 2009, Vella et al 2009}
+
+Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p$)
+\[\sigma = \Delta P \sim \eta h v/\ell_p^2   \]
 \end{textblock}

### 2014-05-27 poster_Okinawa.tex
Introduces epsilon

@@
 	\node[below right] at (axis cs:-90,1.25) {$P_1$};
+	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
 \end{axis}
@@
 \item Independent blisters
-\item $\lambda \ll L \Rightarrow$ not buckling
+\item $h,H,e \ll \lambda \ll L$
 \item Once in contact with ceiling, $\lambda$ fixed
 \item Cascade buckling\linebreak \hfill\textit{Roman \& Pocheau 1999}
+\item Independent of gravity
 \end{itemize}
@@
 \begin{textblock}{3}(8,8.5)
+\Subhead{Ruck in a rug analogy}\\
+\textit{Kolinski et al 2009, Vella et al 2009}\\
 Bending energy $B$ vs vertical stress $\sigma$:
 \[ \lambda^* = \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]
-\textit{Kolinski et al 2009, Vella et al 2009}
 
+Actual shape depends on relative excess length $\epsilon$: $\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
+
+\Subhead{Poroelasticity}\\
 Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p$)

### 2014-05-30 poster_Okinawa.tex
Coupes, volume, aire des plis Altitude du fond d'une cloque (n'avance plus après contact avec le plafond)

@@
 	};
-	\begin{scope}[above left, white, fill=gray]
+	\begin{scope}[above left, white, fill=gray, inner ysep=0]
 		\node at (m-1-1.south east) {\SI{15}{\minute}};
@@
 \begin{textblock}{3}(4,8.5)
+
+\begin{tikzpicture}
+\begin{axis}[
+	scale only axis,
+	width=\textwidth-4em,
+	height=0.3\textwidth,
+	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
+	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
+	cycle list={
+		{Accent2}, 
+		{black!15!Accent2},
+		{black!30!Accent2},
+		{black!45!Accent2},
+		{black!60!Accent2},
+		{black!75!Accent2},
+		{black}
+	},
+	no marks,
+	]
+	\foreach \x in {2,3,..., 8}
+		\addplot table[y index=\x]{alts_bottom_cloque.txt};
+
+	\draw[ultra thick, Accent1] (axis cs:580,0) -- (axis cs:580,50);
+\end{axis}
+\end{tikzpicture}
 \includegraphics[width=\textwidth]{volume_cloque_t116.jpg}
 
+%\Subhead{Phenomenology}
+\begin{itemize}
+\item Synæresis then swelling
+\item First blister after $\approx\SI{1}{\hour}$ of swelling
+\item Independent blisters
+\item $h,H,e \ll \lambda \ll L$
+\item Once in contact with ceiling, $\lambda$ fixed
+\item Cascade buckling\linebreak \hfill\textit{Roman \& Pocheau 1999}
+\item Independent of gravity
+\end{itemize}
+\end{textblock}
+
+\begin{textblock}{3}(8,8.5)
 \begin{tikzpicture}[ultra thick]
@@
 
-\begin{itemize}
-\item Synæresis then swelling
-\item First blister after $\approx\SI{1}{\hour}$ of swelling
-\item Independent blisters
-\item $h,H,e \ll \lambda \ll L$
-\item Once in contact with ceiling, $\lambda$ fixed
-\item Cascade buckling\linebreak \hfill\textit{Roman \& Pocheau 1999}
-\item Independent of gravity
-\end{itemize}
-\end{textblock}
-
-\begin{textblock}{3}(8,8.5)
 \Subhead{Ruck in a rug analogy}\\
@@
 \Subhead{Poroelasticity}\\
-Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p$)
+Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p \approx \SI{4}{\micro\metre}$)
 \[\sigma = \Delta P \sim \eta h v/\ell_p^2   \]
@@
 	\end{tikzpicture}
+\end{textblock}
+
+\begin{textblock}{3}(0,17)
+	\begin{tikzpicture}
+	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t000.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t016.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t032.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t038.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t040.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t043.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t055.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t070.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t123.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t332.png}\\
+	};
+	\begin{scope}[above left, white, fill=gray, inner ysep=0]
+		\node at (m-1-1.south east) {\SI{15}{\minute}};
+		\node at (m-2-1.south east) {\SI{23}{\minute}};
+		\node at (m-3-1.south east) {\SI{32}{\minute}};
+		\node at (m-4-1.south east) {\SI{35}{\minute}};
+		\node at (m-5-1.south east) {\SI{36}{\minute}};
+		\node at (m-6-1.south east) {\SI{38}{\minute}};
+		\node at (m-7-1.south east) {\SI{44}{\minute}};
+		\node at (m-8-1.south east) {\SI{53}{\minute}};
+		\node at (m-9-1.south east) {\SI{1}{\hour}~21};
+		\node at (m-10-1.south east) {\SI{3}{\hour}~15};
+	\end{scope}
+	\draw[<->, ultra thick] ($(m-10-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
+	\end{tikzpicture}
+
+	\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=1 by 2,
+			xticklabels at=edge bottom,
+			vertical sep=0,
+			},
+		xmin=0, xmax=180, xtick={0,30,...,150},
+		extra tick style={grid=major},%
+		extra x tick labels={},%
+		scale only axis,
+		width=\textwidth-4em,
+		height=0.3\textwidth,
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
+		]
+
+	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em}]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_plis.txt};
+	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
+
+	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=6, ytick={0,2,4}, 
+	xlabel={time (min)}, every axis y label/.append style={xshift=-1em}]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_plis.txt};
+	%\addplot+[only marks,Accent2, mark=+] table{excess_area_pc_half_cas8_toi.txt};
+	\end{groupplot}
+	\end{tikzpicture}
+\end{textblock}
+
+\begin{textblock}{7}(4,17)
+\begin{tikzpicture}
+\begin{axis}[
+	name=a,
+	width=\textwidth, height=0.25\textwidth, scale only axis,
+	domain=0:360, no markers, ymin=-4, ymax=4,xmin=0,xmax=360,
+	axis lines=none, xtick=\empty,
+	]
+	\fill[gray!20] (axis cs:85,4) rectangle (axis cs:95,-4) (axis cs:265,4) rectangle (axis cs:275,-4);
+	\addplot+[Accent1]{sin(x)+0.5};
+	\addplot+[Accent1]{sin(x)-0.5};
+	\addplot+[dashed, black]{0.5};
+	\addplot+[dashed, black]{-0.5};
+	\draw[<->, help lines] (axis cs:180,0.5) -- (axis cs:180,4) node[midway, left] {$H$}; 
+	\draw[<->, help lines] (axis cs:30,-0.5) -- (axis cs:30,-4) node[midway, left] {$H$};
+	\node[below] at (axis cs:90,-1.5) (pb1) {$P_1$};
+	\node[above] at (axis cs:90,1.5) (ph2) {$P_2$};
+	\node[above] at (axis cs:270,1.5) (ph1) {$P_1$};
+	\node[below] at (axis cs:270,-1.5) (pb2) {$P_2$};
+	\draw[thick, ->] (ph2) -- (axis cs:90,0.5) node[midway, left] {Darcy} node[midway, right] {$v$};
+	\draw[thick, ->] (pb2) -- (pb1) node[midway, above] {Poiseuille} node[midway, below] {$u \sim \frac{\lambda}{H} v$};
+\end{axis}
+\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle +(\textwidth,-1em) (a.north west) rectangle +(\textwidth,1em);
+\end{tikzpicture}
+
+\begin{itemize}
+\item The same $\Delta P$ can be relaxed by a Poiseuille flow
+\[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
+\item For the same wavelength
+\[\frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30\]
+\end{itemize}
+
 \end{textblock}

### 2014-06-02 poster_Okinawa.tex
Single blister excess area

@@
 
-	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=2, ytick={0,0.5,1,1.5}, xlabel={time (min)}, every axis y label/.append style={xshift=-1em}]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt};
-	%\addplot+[only marks,Accent2, mark=+] table{excess_area_pc_half_cas8_toi.txt};
+	\nextgroupplot[
+		ylabel={Excess area (\%)}, ymin=0, ymax=4, ytick={0,1,2,3},
+		xlabel={time (min)}, 
+		every axis y label/.append style={xshift=-1em},
+%		legend style={legend pos=north west} 
+		]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, below left] {all};
+	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{3}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, left] {single blister};
+	%\legend{all, single blister};
 	\end{groupplot}

### 2014-06-02 poster_Okinawa.tex
Quickly fill in the gaps : thickness, viscosity, etc.

@@
 	pdfauthor={Mathieu Leocmach},%your name
+	hidelinks=true,
 }
@@
 
-For confocal microscopy we label 10\% of the casein.
+\Subhead{Confocal microscopy}\\ 10\% labelled casein
 \end{textblock}
@@
 
-\begin{textblock}{15}(0,8)
+\begin{textblock}{11}(0,8)
 	\Head{Initial flat contact with a substrate: Poroelastic blistering}
@@
 		extra tick style={grid=major},%
-		extra x tick labels={},%
+		extra x ticks = {21,68}, extra x tick labels={},%
 		scale only axis,
@@
 
-	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em}]
+	\nextgroupplot[
+		ylabel={Volume (\%)}, ymin=20, ymax=100, 
+		ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em},
+		]
 	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_cloques.txt};
 	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
+	\node[anchor=south west] at (rel axis cs:0,0) {\rotatebox{90}{synæresis}};
+	\node at (axis cs:68,80) {swelling};
 
@@
 		every axis y label/.append style={xshift=-1em},
-%		legend style={legend pos=north west} 
 		]
@@
 	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{3}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, left] {single blister};
+	\draw[->] (axis cs: 21, 3) -- (axis cs: 68, 3) node[midway, below] {\SI{1}{\hour}};
 	%\legend{all, single blister};
@@
 	\end{tikzpicture}
-
-\end{textblock}
-
-\begin{textblock}{3}(4,8.5)
 
@@
 	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
+	ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em},
 	cycle list={
@@
 \end{tikzpicture}
+
+\end{textblock}
+
+\begin{textblock}{3}(4,8.5)
 \includegraphics[width=\textwidth]{volume_cloque_t116.jpg}
 
-%\Subhead{Phenomenology}
+\Subhead{Phenomenology}
 \begin{itemize}
@@
 \end{itemize}
-\end{textblock}
 
-\begin{textblock}{3}(8,8.5)
 \begin{tikzpicture}[ultra thick]
@@
 
-Actual shape depends on relative excess length $\epsilon$: $\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
-
-\Subhead{Poroelasticity}\\
-Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p \approx \SI{4}{\micro\metre}$)
-\[\sigma = \Delta P \sim \eta h v/\ell_p^2   \]
+%Actual shape depends on relative excess length $\epsilon$: $\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
 \end{textblock}
 
-\begin{textblock}{15}(0,14)
+\begin{textblock}{3}(8,8.5)
+\Subhead{Poroelasticity}\\
+Usual view (\textit{Biot 1964}): pore pressure\linebreak increases transiently the bending energy.
+\begin{itemize}
+\item Solvent must flow through the gel
+\begin{itemize}
+	\item Volumic flux $v\simeq \SI{0.2}{\micro\metre\per\second}$
+	\item Pore size $\ell_p \approx \SI{4}{\micro\metre}$
+\end{itemize}
+\item $\sigma$ is a Darcy pressure
+\[\sigma = \Delta P \sim \eta h v/\ell_p^2   \]
+\item Poroelastic length
+\[\lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}\]
+\end{itemize}
+\Subhead{Wet paper ?}
+
+\smallskip
+\includegraphics[width=\textwidth]{large1}
+\end{textblock}
+
+\begin{textblock}{3}(12,8)
+	\Head{Cell thickness}
+\begin{tikzpicture}
+	\matrix[matrix of nodes, inner sep=0, column sep=0.02\columnwidth, row sep=0.02\columnwidth] (m) {
+		\includegraphics[width=0.49\columnwidth]{pattern_50um.jpg}&
+		\includegraphics[width=0.49\columnwidth]{pattern_100um.jpg}\\
+		\includegraphics[width=0.49\columnwidth]{pattern_250um.jpg}&
+		\includegraphics[width=0.49\columnwidth]{pattern_450um.jpg}\\
+	};
+	\node[inner sep=0, below=0.02\columnwidth of m.south] (x800) {\includegraphics[width=0.98\columnwidth]{pattern_800um_x1.jpg}};
+	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.25\columnwidth, 0.02\columnwidth) -- ++(45:0.29\columnwidth) node[midway,above]{$\lambda$};
+	\draw[line width=0.2em,Main](x800.north east) -- ++(-0.434\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};
+
+	\begin{axis}[%
+		name=g,
+		anchor=left of north west,
+		at={($(x800.south west)+(-0.01\columnwidth,-0.01\columnwidth)$)},
+		scale only axis,
+		width=\textwidth-3.25em,
+		height=0.3\textwidth,
+		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
+		xtick={0,0.4,...,0.9},
+		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		%label shift=-0.5em,
+		cycle multi list={%
+			Accent1,no marks\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=*\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=triangle*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=triangle*\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=square*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=square*\\%
+		},
+		mark size={0.3em},
+		]
+		\addplot+[domain=0:3] {4.23*x};
+		\addplot coordinates {(0.05,0.261)};\label{pgfplots:50}
+		\addplot coordinates {(0.1,0.458)};\label{pgfplots:100}
+		\addplot coordinates {(0.250,0.916)};\label{pgfplots:250}
+		\addplot coordinates {(0.450,2.075)};\label{pgfplots:450}
+		\addplot coordinates {(0.800,3.333)};\label{pgfplots:800}
+		%\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+	\end{axis}
+	\begin{scope}[mark size={0.3em}]
+	\node[below right=0 of m-1-1.north west] {\ref{pgfplots:50}};
+	\node[below right=0 of m-1-2.north west] {\ref{pgfplots:100}};
+	\node[below right=0 of m-2-1.north west] {\ref{pgfplots:250}};
+	\node[below right=0 of m-2-2.north west] {\ref{pgfplots:450}};
+	\node[below right=0 of x800.north west] {\ref{pgfplots:800}};
+	\end{scope}
+	\end{tikzpicture}
+	Shape of a ruck depends on relative excess length $\epsilon$: $\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$.
+
+	At contact: $A=H\equiv e-h\approx e/2$
+
+	\[ \lambda \sim \epsilon^{-3/7} H \]
+	%\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq \sqrt{1.6\%} \]
+
+\end{textblock}
+
+\begin{textblock}{15}(0,16)
 	\Head{No initial contact with a substrate: Poiseuille vs Darcy on a flying carpet}
@@
 
-\begin{textblock}{3}(0,17)
+\begin{textblock}{3}(0,19)
 	\begin{tikzpicture}
@@
 
-\begin{textblock}{7}(4,17)
+\begin{textblock}{7}(4,19)
 \begin{tikzpicture}
@@
 	\node[below] at (axis cs:270,-1.5) (pb2) {$P_2$};
-	\draw[thick, ->] (ph2) -- (axis cs:90,0.5) node[midway, left] {Darcy} node[midway, right] {$v$};
-	\draw[thick, ->] (pb2) -- (pb1) node[midway, above] {Poiseuille} node[midway, below] {$u \sim \frac{\lambda}{H} v$};
+	\draw[line width=0.1em, ->] (ph2) -- (axis cs:90,0.5) node[midway, left] {Darcy} node[midway, right] {$v$};
+	\draw[line width=0.1em, ->] (pb2) -- (pb1) node[midway, above] {Poiseuille} node[midway, below] {$u \sim \frac{\lambda}{H} v$};
 \end{axis}
@@
 \end{tikzpicture}
+\end{textblock}
 
+\begin{textblock}{3}(4,21.5)
+\Subhead{Phenomenology}
 \begin{itemize}
-\item The same $\Delta P$ can be relaxed by a Poiseuille flow
+\item Global destabilisation
+\item Labyrinth pattern
+\item First wrinkle after $\approx\SI{13}{\minute}$ of swelling
+\item Next generations = blisters
+\end{itemize}
+\end{textblock}
+
+\begin{textblock}{3}(8,21.5)
+\begin{itemize}
+\item The same $\Delta P$ can be relaxed by\linebreak a Poiseuille flow
 \[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
@@
 \end{itemize}
+Flow through the pores dominate
+
+\end{textblock}
+
+\begin{textblock}{3}(12,19)
+\Head{Viscosity}
+\begin{tikzpicture}
+	\begin{loglogaxis}[%
+		scale only axis,
+		width=\textwidth-3.25em,
+		height=0.4\textwidth,
+		xmin=0.6,xmax=11,xlabel={viscosity (\si{\milli\pascal\second})},
+		xtick={0.7, 1, 2, 5, 10}, xticklabels={$0.7$, $1$, $2$, $5$, $10$},
+		ytick={1,2,3,4}, yticklabels={$1$,$2$,$3$,$4$},
+		ymin=1, ymax=5, ylabel={$\lambda$ (\si{\milli\metre})},
+		mark size={0.3em},
+		]
+		\addplot+[only marks,Accent1, mark options={fill=Accent1}] table{eta_lambda_eau_pure.txt};
+		\addplot+[only marks,Accent2, mark options={fill=Accent2}] table{eta_lambda_eau_glycerol.txt};
+		%\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+	\end{loglogaxis}
+\end{tikzpicture}
 

### 2014-06-02 poster_Okinawa.tex
Reorganisation complete du poster. Cadres groupés par question scientifique. Plus de liste phénoménologique

--- 2014-06-02 poster_Okinawa.tex
@@
 \usepackage[T1]{fontenc}
+\usepackage{kpfonts}
 \usepackage{fontspec}
@@
 \begin{textblock}{15}(0,8)
-	\Head{Initial flat contact with a substrate: Poroelastic blistering}
-
+	\Head{Dynamics: wrinkling patterns and nested buckling}
+	%\Head{No initial contact with a substrate: Poiseuille vs Darcy on a flying carpet}
+	\begin{tikzpicture}
+	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em] (m){
+	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
+	\includegraphics[width=0.13\textwidth]{prise_0100_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0130_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0160_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0190_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0250_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0360_color.jpg}&
+	\includegraphics[width=0.13\textwidth]{prise_0799_color.jpg}\\
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
+	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
+	};
+	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.023\textwidth,0);
+	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
+	\end{tikzpicture}
 \end{textblock}
 
-\begin{textblock}{3}(0,8.5)
+\begin{textblock}{7}(0,11)
+	\Head{Synæresis and swelling cause buckling}
+\end{textblock}
+
+\begin{textblock}{3}(0,11.5)
+\Subhead{Initial contact with a substrate}\\
 	\begin{tikzpicture}
@@
 
-	\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		xmin=0, xmax=170, xtick={0,30,...,150},
-		extra tick style={grid=major},%
-		extra x tick labels={},%
-		scale only axis,
-		width=\textwidth-4em,
-		height=0.3\textwidth,
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
-		]
+Cascade buckling \textit{Roman \& Pocheau 1999}
 
-	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em}]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_cloques.txt};
-	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
-
-	\nextgroupplot[
-		ylabel={Excess area (\%)}, ymin=0, ymax=4, ytick={0,1,2,3},
-		xlabel={time (min)}, 
-		every axis y label/.append style={xshift=-1em},
-%		legend style={legend pos=north west} 
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, below left] {all};
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{3}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, left] {single blister};
-	%\legend{all, single blister};
-	\end{groupplot}
-	\end{tikzpicture}
-
-\end{textblock}
-
-\begin{textblock}{3}(4,8.5)
-
-\begin{tikzpicture}
-\begin{axis}[
-	scale only axis,
-	width=\textwidth-4em,
-	height=0.3\textwidth,
-	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
-	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
-	cycle list={
-		{Accent2}, 
-		{black!15!Accent2},
-		{black!30!Accent2},
-		{black!45!Accent2},
-		{black!60!Accent2},
-		{black!75!Accent2},
-		{black}
-	},
-	no marks,
-	]
-	\foreach \x in {2,3,..., 8}
-		\addplot table[y index=\x]{alts_bottom_cloque.txt};
-
-	\draw[ultra thick, Accent1] (axis cs:580,0) -- (axis cs:580,50);
-\end{axis}
-\end{tikzpicture}
-\includegraphics[width=\textwidth]{volume_cloque_t116.jpg}
-
-%\Subhead{Phenomenology}
-\begin{itemize}
-\item Synæresis then swelling
-\item First blister after $\approx\SI{1}{\hour}$ of swelling
-\item Independent blisters
-\item $h,H,e \ll \lambda \ll L$
-\item Once in contact with ceiling, $\lambda$ fixed
-\item Cascade buckling\linebreak \hfill\textit{Roman \& Pocheau 1999}
-\item Independent of gravity
-\end{itemize}
-\end{textblock}
-
-\begin{textblock}{3}(8,8.5)
-\begin{tikzpicture}[ultra thick]
-\begin{axis}[
-	name=a,
-	width=\textwidth, height=0.25\textwidth, scale only axis,
-	domain=-180:180, no markers, ymin=0, ymax=4,xmin=-180,xmax=180,
-	axis lines=none, xtick=\empty,
-	]
-	\addplot+[Accent1]{cos(x)+1};
-	\addplot+[Accent1]{cos(x)+1.5};
-	\draw[->, ultra thick] (axis cs:90,1.25) -- +(45:0.05\textwidth) node[pos=1, above] {$\vec{v}$};
-	\node[above] at (axis cs:-90,1.5) {$P_2$};
-	\node[below right] at (axis cs:-90,1.25) {$P_1$};
-	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
-\end{axis}
-\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle (\textwidth,-1em);
-\draw[<->] ($(a.south west)+(0,-1.1em)$) -- +(\textwidth,0) node[midway, below] {$\lambda$};
-\end{tikzpicture}
-
-\Subhead{Ruck in a rug analogy}\\
-\textit{Kolinski et al 2009, Vella et al 2009}\\
-Bending energy $B$ vs vertical stress $\sigma$:
-\[ \lambda^* = \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]
-
-Actual shape depends on relative excess length $\epsilon$: $\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
-
-\Subhead{Poroelasticity}\\
-Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p \approx \SI{4}{\micro\metre}$)
-\[\sigma = \Delta P \sim \eta h v/\ell_p^2   \]
-\end{textblock}
-
-\begin{textblock}{15}(0,14)
-	\Head{No initial contact with a substrate: Poiseuille vs Darcy on a flying carpet}
-	\begin{tikzpicture}
-	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em] (m){
-	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
-	\includegraphics[width=0.13\textwidth]{prise_0100_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0130_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0160_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0190_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0250_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0360_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0799_color.jpg}\\
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
-	};
-	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.023\textwidth,0);
-	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
-	\end{tikzpicture}
-\end{textblock}
-
-\begin{textblock}{3}(0,17)
+\Subhead{No initial contact}\\
 \begin{tikzpicture}
@@
 	\draw[<->, ultra thick] ($(m-10-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
+	\end{tikzpicture}
+
+	Globally unstable wavelength
+\end{textblock}
+
+\begin{textblock}{3}(4,11.5)
+	\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=1 by 2,
+			xticklabels at=edge bottom,
+			vertical sep=0,
+			},
+		xmin=0, xmax=170, xtick={0,30,...,150},
+		extra tick style={grid=major},%
+		extra x ticks = {21,68}, extra x tick labels={},%
+		scale only axis,
+		width=\textwidth-4em,
+		height=0.3\textwidth,
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
+		]
+
+	\nextgroupplot[
+		ylabel={Volume (\%)}, ymin=20, ymax=100, 
+		ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em},
+		]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_cloques.txt};
+	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
+	\node[anchor=south west] at (rel axis cs:0,0) {\rotatebox{90}{synæresis}};
+	\node at (axis cs:68,80) {swelling};
+
+	\nextgroupplot[
+		ylabel={Excess area (\%)}, ymin=0, ymax=4, ytick={0,1,2,3},
+		xlabel={time (min)}, 
+		every axis y label/.append style={xshift=-1em},
+		]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, below left] {all};
+	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{3}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, left] {single blister};
+	\draw[->] (axis cs: 21, 3) -- (axis cs: 68, 3) node[midway, below] {\SI{1}{\hour}};
+	%\legend{all, single blister};
+	\end{groupplot}
 	\end{tikzpicture}
@@
 
-\begin{textblock}{7}(4,17)
+\begin{textblock}{3}(12,12)
+\textit{Kolinski et al 2009, Vella et al 2009}\\
+Bending energy $B$ vs vertical stress $\sigma$:
+\[ \lambda^* = \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]
+\end{textblock}
+
+\begin{textblock}{7}(8,11)
+	\Head{Wavelength: resisting stress needed}
+	\Subhead{Initial contact with a substrate : Ruck in a rug analogy}
+
+\begin{tikzpicture}[ultra thick]
+\begin{axis}[
+	name=a,
+	width=3\TPHorizModule, height=0.75\TPHorizModule, scale only axis,
+	domain=-180:180, no markers, ymin=0, ymax=4,xmin=-180,xmax=180,
+	axis lines=none, xtick=\empty,
+	]
+	\addplot+[Accent1]{cos(x)+1};
+	\addplot+[Accent1]{cos(x)+1.5};
+	\draw[->, ultra thick] (axis cs:90,1.25) -- +(45:0.05\textwidth) node[pos=1, above] {$\vec{v}$};
+	\node[above] at (axis cs:-90,1.5) {$P_2$};
+	\node[below right] at (axis cs:-90,1.25) {$P_1$};
+	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
+\end{axis}
+\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle ($(a.south east)+(0,-1em)$);
+\draw[<->] ($(a.south west)+(0,-1.1em)$) -- ($(a.south east)+(0,-1.1em)$) node[midway, below] {$\lambda$};
+\end{tikzpicture}
+
+Actual shape depends on relative excess length $\epsilon$: $\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
+
+\Subhead{Poroelasticity instead of gravity} \hfill Vertical cells show the same wavelength
+\begin{itemize}
+\item Solvent must flow through the gel(volumic flux $v\simeq \SI{0.2}{\micro\metre\per\second}$, pore size $\ell_p \approx \SI{4}{\micro\metre}$)
+\item[$\Rightarrow$] $\sigma$ is a Darcy pressure
+$\sigma = \Delta P \sim \eta h v/\ell_p^2   $
+\hfill\fcolorbox{Main}{white}{$\Rightarrow$ Poroelastic length
+$\displaystyle \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}$}
+\item Different from usual poroelasticity (\textit{Biot 1964}) where pore pressure increases transiently the bending energy. Here Darcy flow resists large wavelength.
+\end{itemize}
+
+\Subhead{No initial contact with a substrate: Poiseuille vs Darcy on a flying carpet}
+
+\smallskip
 \begin{tikzpicture}
@@
 \item The same $\Delta P$ can be relaxed by a Poiseuille flow
-\[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
+$\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v$
 \item For the same wavelength
-\[\frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30\]
+$\frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$ \hfill$\Rightarrow$ Flow through the pores dominates
 \end{itemize}
+\end{textblock}
+
+\begin{textblock}{7}(0,18)
+	\Head{Cell thickness}
+\begin{tikzpicture}[mark size={0.3em}]
+	\matrix[matrix of nodes, inner sep=0, column sep=0.01\columnwidth, row sep=0.01\columnwidth] (m) {
+		\includegraphics[width=0.245\columnwidth]{pattern_50um.jpg}&
+		\includegraphics[width=0.245\columnwidth]{pattern_100um.jpg}&
+		\includegraphics[width=0.245\columnwidth]{pattern_250um.jpg}&
+		\includegraphics[width=0.245\columnwidth]{pattern_450um.jpg}\\
+	};
+	\node[inner sep=0, below left=0.01\columnwidth and 0 of m.south east] (x800) {\includegraphics[width=0.49\columnwidth]{pattern_800um_x1.jpg}};
+	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.196\columnwidth, 0.05\columnwidth) -- ++(0.145\columnwidth,0) node[midway,above]{$\lambda$};
+	\draw[line width=0.2em,Main](x800.north east) -- ++(-0.217\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};
+
+	\begin{axis}[%
+		name=g,
+		anchor=left of north west,
+		at={($(m.south west)+(0,-0.01\columnwidth)$)},
+		scale only axis,
+		width=3\TPHorizModule-4em,
+		height=0.25\columnwidth-2em,
+		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
+		xtick={0,0.4,...,0.9},
+		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em},
+		cycle multi list={%
+			Accent1,no marks\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=*\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=triangle*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=triangle*\\%
+			Accent1,only marks,every mark/.append style={fill=Accent1},mark=square*\\%
+			Accent1,only marks,every mark/.append style={fill=white},mark=square*\\%
+		},
+		mark size={0.3em}
+		]
+		\addplot+[domain=0:3] {4.23*x};
+		\addplot coordinates {(0.05,0.261)};\label{pgfplots:50}
+		\addplot coordinates {(0.1,0.458)};\label{pgfplots:100}
+		\addplot coordinates {(0.250,0.916)};\label{pgfplots:250}
+		\addplot coordinates {(0.450,2.075)};\label{pgfplots:450}
+		\addplot coordinates {(0.800,3.333)};\label{pgfplots:800}
+		%\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
+	\end{axis}
+	\node[below right=0 of m-1-1.north west] {\ref{pgfplots:50}};
+	\node[below right=0 of m-1-2.north west] {\ref{pgfplots:100}};
+	\node[below right=0 of m-1-3.north west] {\ref{pgfplots:250}};
+	\node[below right=0 of m-1-4.north west] {\ref{pgfplots:450}};
+	\node[below right=0 of x800.north west] {\ref{pgfplots:800}};
+\end{tikzpicture}
+
+\Subhead{Wavelength stop evolving after a blister touches the ceiling}
+\end{textblock}
+
+\begin{textblock}{3}(0,22.25)
+	\begin{tikzpicture}
+\begin{axis}[
+	scale only axis,
+	width=\textwidth-4em,
+	height=0.4\textwidth,
+	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
+	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
+	ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em},
+	cycle list={
+		{Accent2}, 
+		{black!15!Accent2},
+		{black!30!Accent2},
+		{black!45!Accent2},
+		{black!60!Accent2},
+		{black!75!Accent2},
+		{black}
+	},
+	no marks,
+	]
+	\foreach \x in {2,3,..., 8}
+		\addplot table[y index=\x]{alts_bottom_cloque.txt};
+
+	\draw[ultra thick, Accent1] (axis cs:580,0) -- (axis cs:580,50);
+\end{axis}
+\end{tikzpicture}
+\end{textblock}
+
+\begin{textblock}{3}(4,22.25)
+At contact: $A=H\equiv e-h\approx e/2$
+	\[ \lambda \sim \epsilon^{-3/7} H \]
+	\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq 0.6\% \]
 
@@
 
-\begin{textblock}{15}(0,25)
+\begin{textblock}{7}(8,21)
+\Head{Surprising viscosity influence}
+\begin{tikzpicture}
+	\begin{loglogaxis}[%
+		scale only axis,
+		width=3\TPHorizModule-4em,
+		height=2\TPHorizModule,
+		xmin=0.6,xmax=11,xlabel={viscosity (\si{\milli\pascal\second})},
+		xtick={0.7, 1, 2, 5, 10}, xticklabels={$0.7$, $1$, $2$, $5$, $10$},
+		ytick={1,2,3,4}, yticklabels={$1$,$2$,$3$,$4$},
+		ymin=1, ymax=5, ylabel={$\lambda$ (\si{\milli\metre})},
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em},
+		mark size={0.3em},
+		legend style={legend pos=south east},
+		]
+		\addplot+[no marks, domain=0.7:10, black, forget plot] {1.7*x^0.4} node[midway, below] {$\eta^{0.4}$};
+		\addplot+[only marks,Accent1, mark options={fill=Accent1}] table{eta_lambda_eau_pure.txt};
+		\addplot+[only marks,Accent2, mark options={fill=Accent2}] table{eta_lambda_eau_glycerol.txt};
+		\legend{water, water-glycerol};
+	\end{loglogaxis}
+\end{tikzpicture}
+\end{textblock}
+
+\begin{textblock}{3}(12,21.5)
+\begin{itemize}
+\item $\lambda$ increases with $\eta$
+\item Can be explained only by a hidden dependence of $\eta$ on $B$, $\ell_p$, \ldots 
+\item[\Rightarrow] Viscosity influences gel formation.
+\end{itemize}
+If $\ell_p \sim \eta^{-\alpha}$ and $B\sim \mathrm{k_B}T \left(\frac{h}{\ell}\right)^3$, then
+\[ \lambda^* \sim \eta^\frac{\alpha-1}{3} \]
+\[\alpha\simeq 2.2\]
+\end{textblock}
+
+\begin{textblock}{15}(0,24.75)
 \small{\texttt{The research leading to these results has received funding from the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement n°~258803.}}
@@
 	\Norulehead{Take home}
-	\Subhead{Poroelasticity dominates}
+	\Subhead{Original patterns}
 	\begin{itemize}
-		\item gravity
-		\item lubrication
+		\item nested (optical elements ?)
+		\item generated by swelling
+		\item kinetic wavelength selection
 	\end{itemize}
-	\Subhead{Two modes}
+	\Subhead{A new wrinkling mechanism}
 	\begin{itemize}
-		\item blistering
-		\item wrinkling
+		\item porous flow resist bending
+		\item Darcy can beat Poiseuille
+		\item may extend to other porous sheets 
+		\begin{itemize}
+			\item wet paper
+			\item biological membranes
 		\end{itemize}
-	\Subhead{Nested patterns}
+	\end{itemize}
+	\Subhead{Viscosity influences}
 	\begin{itemize}
-		\item generated by swelling
-		\item kinetic wavelength
+		\item porous dissipation, thus wrinkling
+		\item gel formation, thus wrinkling
 	\end{itemize}

### 2014-06-03 poster_Okinawa.tex
Corrections Thomas & Mathieu

@@
 \def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{#1}}}
+
+\renewcommand{\descriptionlabel}[1]{\hspace{\labelsep}\textcolor{Accent2}{\textsc{#1}}}
 
@@
 
-\Subhead{Confocal microscopy}\\ 10\% labelled casein
+\Subhead{Microscopy}
+\begin{description}
+\item[macroscope] transmission \& reflection
+\item[confocal] 10\% labelled casein
+\end{description} 
 \end{textblock}
@@
 		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] (L){$L\sim 2.5\,cm$};
-		\node[anchor=north west, inner sep=0] at ($(L.north) - (0.5\textwidth,0)$) {\Subhead{Side view}};
+		\node[anchor=north west, inner sep=0] at ($(L.north) - (0.5\textwidth,0)$) {\Subhead{Side view of the cell}};
 	\end{tikzpicture}
 
-	\Subhead{Top view}\\
+	\Subhead{Top view of the final pattern}\\
 	\begin{tikzpicture}[inner sep=0, very thick]
@@
 	};
-	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.023\textwidth,0);
-	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
+	\draw[line width=0.2em] ++(m-2-1.north west) -- ++(0.023\textwidth,0) node[pos=0, above right, inner xsep=0] {\SI{1}{\milli\metre}};
+	\draw[line width=0.2em] ++(m-3-1.south west) -- ++(0.1\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
+	\node[anchor=north east, rotate=90] at (m-2-1.north east) (macro) {macroscope};
+	\node[anchor=base, rotate=90] at (m-3-1.east-| macro.base) {confocal};
 	\end{tikzpicture}
@@
 	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png} & \SI{15}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png} & \SI{22}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png} & \SI{1}{\hour}~26\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png} & \SI{1}{\hour}~30\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png} & \SI{1}{\hour}~31\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png} & \SI{1}{\hour}~35\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png} & \SI{2}{\hour}~25\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png} & \SI{2}{\hour}~40\\
 	};
-	\begin{scope}[above left, white, fill=gray, inner ysep=0]
-		\node at (m-1-1.south east) {\SI{15}{\minute}};
-		\node at (m-2-1.south east) {\SI{22}{\minute}};
-		\node at (m-3-1.south east) {\SI{1}{\hour}~26};
-		\node at (m-4-1.south east) {\SI{1}{\hour}~30};
-		\node at (m-5-1.south east) {\SI{1}{\hour}~31};
-		\node at (m-6-1.south east) {\SI{1}{\hour}~35};
-		\node at (m-7-1.south east) {\SI{2}{\hour}~25};
-		\node at (m-8-1.south east) {\SI{2}{\hour}~40};
-	\end{scope}
 	\draw[<->, ultra thick] ($(m-8-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
@@
 	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t000.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t016.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t032.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t038.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t040.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t043.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t055.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t070.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t123.png}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t332.png}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t000.png} & \SI{15}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t016.png} & \SI{23}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t032.png} & \SI{32}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t038.png} & \SI{35}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t040.png} & \SI{36}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t043.png} & \SI{38}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t055.png} & \SI{44}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t070.png} & \SI{53}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t123.png} & \SI{1}{\hour}~21\\
+	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t332.png} & \SI{3}{\hour}~15\\
 	};
-	\begin{scope}[above left, white, fill=gray, inner ysep=0]
-		\node at (m-1-1.south east) {\SI{15}{\minute}};
-		\node at (m-2-1.south east) {\SI{23}{\minute}};
-		\node at (m-3-1.south east) {\SI{32}{\minute}};
-		\node at (m-4-1.south east) {\SI{35}{\minute}};
-		\node at (m-5-1.south east) {\SI{36}{\minute}};
-		\node at (m-6-1.south east) {\SI{38}{\minute}};
-		\node at (m-7-1.south east) {\SI{44}{\minute}};
-		\node at (m-8-1.south east) {\SI{53}{\minute}};
-		\node at (m-9-1.south east) {\SI{1}{\hour}~21};
-		\node at (m-10-1.south east) {\SI{3}{\hour}~15};
-	\end{scope}
 	\draw[<->, ultra thick] ($(m-10-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
@@
 	\Norulehead{Take home}
+	Spontaneous wrinkling of over-acidified casein gels in non-adhesive confinement
+
 	\Subhead{Original patterns}
@@
 		\item nested (optical elements ?)
-		\item generated by swelling
 		\item kinetic wavelength selection
@@
 		\item Darcy can beat Poiseuille
-		\item may extend to other porous sheets 
-		\begin{itemize}
-			\item wet paper
-			\item biological membranes
-		\end{itemize}
+		\item may extend to other porous sheets\linebreak  (wet paper, biological membranes\ldots)
 	\end{itemize}
@@
 	\begin{itemize}
-		\item porous dissipation, thus wrinkling
-		\item gel formation, thus wrinkling
+		\item porous dissipation
+		\item gel formation\hfill
+		\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ thus wrinkling}
 	\end{itemize}

### 2014-06-03 seminaire_Tanakalab.tex
Introduction for the Seminar in Tanaka lab

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[francais]{babel}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
\usefonttheme{structuresmallcapsserif}
\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
\usecolortheme{whale}
%\useoutertheme{infolines}

\usecolortheme[named=Accent1]{structure}

\setbeamercolor{alerted text}{fg=Accent2}
%\setbeamercolor{palette primary}{fg=white}
%\setbeamercolor{palette secondary}{bg=Accent1}
%\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}

\setbeamerfont{page number in head/foot}{size=\large}
\setbeamercolor{page number in head/foot}{fg=Main}
% page/total
%\setbeamertemplate{footline}[frame number]
% pas de total
\setbeamertemplate{footline}{%
    	\hfill%
	\usebeamercolor[fg]{page number in head/foot}%
	\usebeamerfont{page number in head/foot}%
	\insertframenumber\kern1em\vskip2pt%
}

\setbeamersize{text margin left=1em}
\setbeamersize{text margin right=1em}

%font
\usepackage[T1]{fontenc}
\usepackage[oldstylenums]{kpfonts}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[E.N.S. Lyon]{Ecole Normale Supérieure de Lyon}
\title[Yoghurt under stress]{Yoghurt under stress}
\author[M. Leocmach]{Mathieu Leocmach}
\date{1st June 2014}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{compat=1.3, every axis/.append style={footnotesize}}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
		{\footnotesize Leocmach \& Tanaka \textit{Nature Communications} 3, 974 (2012).}
	\end{frame}
}

\begin{frame}[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{Application of biogels}
\begin{tabu}{cX[c]c}
\includegraphics[height=0.3\textheight]{cell_mech} &
\includegraphics[height=0.3\textheight]{electrophoresis} &
\includegraphics[height=0.3\textheight]{cosmetics} \\
Cell mechanics & Electrophoresis & Cosmetics\\
\end{tabu}
\begin{tabu}{X[c]X[c]}
\includegraphics[height=0.3\textheight]{bacterial_culture} &
\includegraphics[height=0.3\textheight]{food} \\
Bacterial culture & Food\\
\end{tabu}

\begin{footnotesize}
\begin{block}{Sources}
\begin{tabu}{lXl}
Wikimedia Commons & www.madaboutscience.com & www.keautystore.com\\
\end{tabu}
\end{block}
\end{footnotesize}
\end{frame}

\begin{frame}{Fundamental issues in biogels}
\begin{itemize}
\item Linear elasticity in not well understood \textit{\scriptsize Gardel et al. Science 2004}
\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
\end{itemize}
\begin{block}{Our questions}
\begin{itemize}
\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
\item What is the role of the solvent in mechanical instabilities?
\end{itemize}

\end{block}
\end{frame}

\end{document}

### 2014-06-03 seminaire_Tanakalab.tex
Collaborators (Creep)

@@
 
-\institute[E.N.S. Lyon]{Ecole Normale Supérieure de Lyon}
+\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Supérieure de Lyon}
 \title[Yoghurt under stress]{Yoghurt under stress}
@@
 		\tableofcontents[currentsection, hideothersubsections]
-		{\footnotesize Leocmach \& Tanaka \textit{Nature Communications} 3, 974 (2012).}
 	\end{frame}
@@
 \end{itemize}
-
 \end{block}
@@
 
+\section{Creep and yielding}
+
+\begin{frame}{Merciless yoghurt breakers}
+
+\begin{tabu}{X[c]X[c]c}
+\includegraphics[height=0.3\textheight]{Chris}&
+\includegraphics[height=0.3\textheight]{Thibaut}&
+\includegraphics[height=0.3\textheight]{Seb}\\
+Christophe Perge & Thibaut Divoux & Sebastien Manneville\\
+PhD student & CNRS researcher & Professor\\
+E.N.S Lyon &  CRPP Bordeaux & E.N.S Lyon\\
+\end{tabu}
+
+\end{frame}
+
+\section{Wrinkling}
+
 \end{document}

### 2014-06-03 seminaire_Tanakalab.tex
Creep experiment description

@@
 
+\section{Our system: Yoghurt}
+
+\begin{frame}{Acid-set protein gel}
+\begin{block}{Recipe}
+\begin{itemize}
+\item Water (\SI{30}{\celsius})
+\item Sodium caseinate (milk protein) 4\%
+\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
+\item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow acidification
+\end{itemize}
+\end{block}
+
+\end{frame}
+
 \section{Creep and yielding}
@@
 \begin{frame}{Merciless yoghurt breakers}
-
 \begin{tabu}{X[c]X[c]c}
@@
 PhD student & CNRS researcher & Professor\\
-E.N.S Lyon &  CRPP Bordeaux & E.N.S Lyon\\
+E.N.S. Lyon &  CRPP Bordeaux & E.N.S. Lyon\\
 \end{tabu}
+\end{frame}
 
+\begin{frame}{The yoghurt creep experiment}
+\begin{tikzpicture}
+\node[anchor=south west] (webcam){\includegraphics[width=0.2\textwidth]{webcam}};
+\node[anchor=south] at (0.4\textwidth,0) (spoon) {\includegraphics[width=0.3\textwidth]{spoon}};
+\node[anchor=south east] at (\textwidth,0) (ultrasound) {\includegraphics[width=0.4\textwidth]{ultrasound}};
+
+\node[anchor=south west] at ($(ultrasound.north west)+(0,2em)$) (rheometer) {\includegraphics[height=0.4\textheight]{rheometer}};
+
+\begin{scope}[draw]
+	\node[below=0of spoon.south] (couette) {A Taylor-Couette cell where the yogurt is made \emph{in situ}};
+	\draw[->] (couette) -- ($(spoon.south)+(0,2em)$);
+	\node[above=0of ultrasound] {ultrasonic imaging};
+	\node[above=0of webcam, text width=0.2\textwidth, align=center] {optical imaging};
+	\draw[->] (spoon.north east) +(235:2em) -- (rheometer);
+	\node[right=0of rheometer, text width=0.2\textwidth, align=center] {A rheometer imposes a constant stress $\sigma$ and records the strain $\gamma(t)$};
+\end{scope}
+\end{tikzpicture}
 \end{frame}

### 2014-06-05 seminaire_Tanakalab.tex
All needed data to make creep figures.\n Yoghurt recipe

@@
 
+\begin{columns}
+\column{0.5\textwidth}
+\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=1 by 2,
+			xticklabels at=edge bottom,
+			vertical sep=0,
+			},
+		xmode=log,
+		xmin=1e2,xmax=3e5,
+		scale only axis,
+		width=\textwidth-4em,
+		height=0.3\textwidth,
+		extra tick style={grid=major},%
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+		]
+	\nextgroupplot[
+		ymin=0, ymax=7, ylabel={pH},
+		%extra y ticks={4.6}, extra y tick labels={},%
+		]
+	\addplot+[no marks,Accent2] table{Y190_cas4_gdl1.pH};
+	\draw[help lines] (axis cs:1e2,4.6) -- (axis cs:1.3979e4,4.6) node[pos=0, below right, font=\scriptsize, text width=0.4\textwidth] {isoelectric point $pH\approx 4.6$}  -- (axis cs:1.3979e4,0);
+
+	\nextgroupplot[
+		xlabel={time (s)},
+		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
+		extra x ticks={1e5}, extra x tick labels={},
+		]
+	\addplot+[no marks,Accent2] table{cas4_GDL1_Y22.prise};
+	\addplot+[no marks,Main] table[y index=2]{cas4_GDL1_Y22.prise};
+	\begin{scope}[font=\scriptsize]
+		\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
+		\draw[<-] (axis cs:1.3979e4,654) -- +(-1em,0) node[left] {gelation};
+	\end{scope}
+	\end{groupplot}
+\end{tikzpicture}
+
+\column{0.5\textwidth}
+\includegraphics[width=\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
+
+\begin{scriptsize}
+\textit{Kaláb (1983),\linebreak Roefs \& van Vliet (1990),\linebreak Lucey \& Singh (1998)}
+\end{scriptsize}
+
+\end{columns}
 \end{frame}

### 2014-06-05 seminaire_Tanakalab.tex
Prises + Frequency sweeps

@@
 \pgfplotsset{compat=1.3, every axis/.append style={footnotesize}}
+
+\pgfplotscreateplotcyclelist{earthy}{%
+red!40!black,
+red!60!black,
+red!80!black,
+red,
+red!80!yellow,
+red!60!yellow,
+red!40!yellow,
+}
 
@@
 \end{scriptsize}
+\end{columns}
+\end{frame}
 
-\end{columns}
+\begin{frame}{Over-acidification: Faster but weaker}
+\begin{tikzpicture}
+\begin{axis}[
+	height=0.8\textheight,
+	width=\textwidth,
+	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
+	cycle list name=earthy,
+	no marks,
+	xmin=0, xmax=20,ymin=0,
+	]
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
+\end{axis}
+\end{tikzpicture}
+
+Caseins regain some solubility at low pH.
+\end{frame}
+
+\begin{frame}{Linear rheology: Power law (visco)elastic solid}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	height=0.8\textheight,
+	width=\textwidth,
+	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
+	cycle list name=earthy,
+	domain={6e-2:70},
+	]
+	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
+	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
+	\addplot{590*x^0.14};
+	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
+	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
+	\addplot{350*x^0.12};
+	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
+	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
+	\addplot{246*x^0.1};
+	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
+	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
+	\addplot{163*x^0.067};
+	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
+	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
+	\addplot{109*x^0.044};
+	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
+	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
+	\addplot{76*x^0.035};
+
+\end{loglogaxis}
+\end{tikzpicture}
 \end{frame}

### 2014-06-06 seminaire_Tanakalab.tex
Yoghurt creep experiment

--- 2014-06-05 seminaire_Tanakalab.tex
@@
 \usepackage{pgfplots}
+\pgfplotsset{compat=1.9}
 \usepgfplotslibrary{groupplots}
 \usepgfplotslibrary{external}
+\tikzexternalize
+\tikzsetexternalprefix{fig_presentation/}
+%\tikzset{external/optimize=false}
+%\tikzset{external/force remake}
 
@@
 \tikzset{every mark/.append style={scale=0.8}}
-\pgfplotsset{compat=1.3, every axis/.append style={footnotesize}}
+\pgfplotsset{every axis/.append style={footnotesize}}
+
+\pgfplotscreateplotcyclelist{earthy}{%
+red!40!black,
+red!60!black,
+red!80!black,
+red,
+red!80!yellow,
+red!60!yellow,
+red!40!yellow,
+}
 
@@
 			},
-		xmode=log,
+		%xmode=log,
 		xmin=1e2,xmax=3e5,
@@
 \end{scriptsize}
+\end{columns}
+\end{frame}
 
-\end{columns}
+\begin{frame}{Linear rheology: Power law (visco)elastic solid}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	height=0.8\textheight,
+	width=\textwidth,
+	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{Loss} moduli (\si{\pascal})},
+	domain={6e-2:70},
+	]
+	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
+	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
+	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime,G^{\prime\prime}\sim f^{0.15}$};
+%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
+%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
+%	\addplot{590*x^0.14};
+%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
+%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
+%	\addplot{350*x^0.12};
+%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
+%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
+%	\addplot{246*x^0.1};
+%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
+%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
+%	\addplot{163*x^0.067};
+%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
+%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
+%	\addplot{109*x^0.044};
+%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
+%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
+%	\addplot{76*x^0.035};
+
+\end{loglogaxis}
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}{Over-acidification: Faster but weaker}
+\begin{tikzpicture}
+\begin{axis}[
+	height=0.8\textheight,
+	width=\textwidth,
+	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
+	cycle list name=earthy,
+	no marks,
+	xmin=0, xmax=20,ymin=0,
+	]
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
+\end{axis}
+\end{tikzpicture}
+
+Caseins regain some solubility at low pH.
 \end{frame}
@@
 
+\tikzset{external/force remake}
+\begin{frame}{The yoghurt creep experiment}
+\begin{tikzpicture}
+\node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{/home/mathieu/Documents/tex/svn/Yaourt/creep_experiment.jpg}};
+\begin{scope}[shift={(a.north west)}, shift={(7.2\baselineskip,-1.1\baselineskip)}, ultra thick]
+	\draw[Accent1] ellipse[x radius=2.9\baselineskip, y radius=0.7\baselineskip] ++(2.9\baselineskip,0)  -- ++(0,-6.6\baselineskip)  ++(-2.9\baselineskip,0) ellipse[x radius=2.9\baselineskip, y radius=0.7\baselineskip] ++(-2.9\baselineskip,0)  -- ++(0,6.6\baselineskip);
+	\shadedraw[Accent2,ball color=Accent2] ellipse[x radius=2\baselineskip, y radius=0.4\baselineskip] ++(2\baselineskip,0)  -- ++(0,-6.6\baselineskip)  ++(-2\baselineskip,0) ellipse[x radius=2\baselineskip, y radius=0.4\baselineskip] ++(-2\baselineskip,0)  -- ++(0,6.6\baselineskip);
+	\fill[Accent2, left color=black,right color=black, middle color=Accent2, ] (-0.3\baselineskip,0) rectangle +(0.6\baselineskip, \baselineskip);
+	\begin{scope}[Main]
+		\node[draw, minimum height=3.5\baselineskip, minimum width=1.2\baselineskip] at (2.4\baselineskip,-3.35\baselineskip) (window){};
+		\node[draw, minimum height=7.5\baselineskip, minimum width=4\baselineskip] at (7.5\baselineskip,-4\baselineskip) (USV){};
+		\draw(window.north east) -- (USV.north west) (window.south east) -- (USV.south west);
+		\node[right, text width=4\baselineskip] at (USV.east) {ultrasound region of interest};
+	\end{scope}
+\end{scope}
+\draw[->] (window.north west) ++(-1em, 0.2em) -- +(1em,0) node[midway,above,inner sep=1pt] {$r$};
+\draw[->] (window.north west) ++(-1em, 0.2em) -- +(0,-1em)  node[midway,left] {$z$};
+\node[text width=4\baselineskip] at (-3\baselineskip,-2.5\baselineskip) {shear rate response};
+\node at (5\baselineskip,-2.5\baselineskip) {strain response};
+\end{tikzpicture}
+\end{frame}
+
 \section{Wrinkling}

### 2014-06-06 seminaire_Tanakalab.tex
creep rheology

@@
 %\tikzset{external/force remake}
+
+%link or play movies
+\usepackage{multimedia}
 
@@
 \begin{frame}{Linear rheology: Power law (visco)elastic solid}
+\[\text{stress}\rightarrow \sigma = G \gamma\leftarrow\text{strain}\]
+\[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
 \begin{tikzpicture}
@@
 	width=\textwidth,
-	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{Loss} moduli (\si{\pascal})},
+	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{loss} moduli (\si{\pascal})},
 	domain={6e-2:70},
@@
 
+%\tikzset{external/force remake}
 \begin{frame}{The yoghurt creep experiment}
@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{The yoghurt creep experiment}
-\begin{tikzpicture}
-\node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{/home/mathieu/Documents/tex/svn/Yaourt/creep_experiment.jpg}};
+\movie[externalviewer]{\begin{tikzpicture}
+\node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{creep_experiment.jpg}};
 \begin{scope}[shift={(a.north west)}, shift={(7.2\baselineskip,-1.1\baselineskip)}, ultra thick]
@@
 \node at (5\baselineskip,-2.5\baselineskip) {strain response};
+\end{tikzpicture}%
+}{/home/mathieu/Dropbox/Yaourt/films/Yaourt141_enhanced.avi}
+\end{frame}
+
+\tikzset{external/force remake}
+\begin{frame}{Creep rheology: Three regimes}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+   name=g,
+	width=\textwidth,
+	height=0.8\textheight,
+	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
+	ymin=0.2,ymax=2,
+	ytick={0.2, 0.4,0.8,1.6}, yticklabels={0.2, 0.4,0.8,1.6},
+	ylabel={strain $\gamma$},
+	extra tick style={grid=major},%
+	extra y ticks={1}, extra y tick labels={1.0},
+	cycle list name=earthy,
+	axis on top,
+	]
+	\fill[Accent1!20]
+		(axis cs:10,0.4) ellipse[rotate=-30, x radius=0.2\textwidth, y radius=0.12\textwidth]
+		(axis cs:2e2,1.6) ellipse[x radius=0.3\textwidth, y radius=0.05\textwidth]
+		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
+		;
+	\node[above right] at (axis cs:1e1,0.2) {primary};
+	\node[below left] at (axis cs:4e5,0.7) {secondary};
+	\node[left] at (axis cs:5,1.6) {tertiary};
+	\addplot table{Y27_200Pa_gamma_decimated.txt} node (s200){};
+	\addplot table{Y38_300Pa_gamma_decimated.txt} node (s300){};
+	\addplot table{Y25_400Pa_gamma_decimated.txt}  node (s400){};
+	\addplot table{Y32_550Pa_gamma_decimated.txt}  node (s550){};
+	\addplot table{Y39_1000Pa_gamma_decimated.txt} node (s1000){};
+\end{loglogaxis}
+\begin{scope}[anchor=base, every node/.style={yshift=0.2em}]
+	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
+	\node[red!60!black] at (s300 |- g.outer north) {300};
+	\node[red!80!black] at (s400 |- g.outer north) {400};
+	\node[red] at (s550 |- g.outer north) {550};
+	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
+\end{scope}
 \end{tikzpicture}
+
+Failure at $\gamma\approx 1$ for a well defined time $\tau_f$
 \end{frame}

### 2014-06-15 seminaire_Tanakalab.tex
gelation linear time Villi

@@
 \date{1st June 2014}
+\titlegraphic{
+	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
+	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
+	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
+	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
+	}
 
@@
 \end{itemize}
+\includegraphics[height=3\baselineskip]{Villi_sq}
 \begin{block}{Our questions}
@@
 
-\section{Our system: Yoghurt}
+%\section{Our system: Yoghurt}
 
-\begin{frame}{Acid-set protein gel}
+\tikzset{external/force remake}
+\begin{frame}{Acid-set protein gel (Yoghurt)}
 \begin{block}{Recipe}
@@
 		%xmode=log,
-		xmin=1e2,xmax=3e5,
+		%xmin=1e2,xmax=3e5,
+		xmin=0, xmax=20,
 		scale only axis,
@@
 		ymin=0, ymax=7, ylabel={pH},
-		%extra y ticks={4.6}, extra y tick labels={},%
+		extra y ticks={4.6}, extra y tick labels={},%
 		]
-	\addplot+[no marks,Accent2] table{Y190_cas4_gdl1.pH};
-	\draw[help lines] (axis cs:1e2,4.6) -- (axis cs:1.3979e4,4.6) node[pos=0, below right, font=\scriptsize, text width=0.4\textwidth] {isoelectric point $pH\approx 4.6$}  -- (axis cs:1.3979e4,0);
+	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y190_cas4_gdl1.pH};
+	\draw[help lines] (axis cs:0,4.6) -- (axis cs:3.88,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:3.88,0);
 
 	\nextgroupplot[
-		xlabel={time (s)},
+		xlabel={time (\si{\hour})},
 		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
-		extra x ticks={1e5}, extra x tick labels={},
+		extra x ticks={17}, extra x tick labels={},
 		]
-	\addplot+[no marks,Accent2] table{cas4_GDL1_Y22.prise};
-	\addplot+[no marks,Main] table[y index=2]{cas4_GDL1_Y22.prise};
+	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y22.prise};
+	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL1_Y22.prise};
 	\begin{scope}[font=\scriptsize]
-		\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
-		\draw[<-] (axis cs:1.3979e4,654) -- +(-1em,0) node[left] {gelation};
+		%\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
+		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
+		\node[anchor=base,rotate=80] at (axis cs:2.5,400) {gelation};
 	\end{scope}
 	\end{groupplot}
+	\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
 \end{tikzpicture}
@@
 \end{frame}
+\tikzset{external/force remake=false}
 
@@
 \end{tikzpicture}
-\end{frame}
-
-\begin{frame}{Over-acidification: Faster but weaker}
-\begin{tikzpicture}
-\begin{axis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
-	cycle list name=earthy,
-	no marks,
-	xmin=0, xmax=20,ymin=0,
-	]
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
-\end{axis}
-\end{tikzpicture}
-
-Caseins regain some solubility at low pH.
 \end{frame}
@@
 
-\tikzset{external/force remake}
 \begin{frame}{Creep rheology: Three regimes}
@@
 
+%\tikzset{external/force remake}
+\begin{frame}{Over-acidification: Faster but weaker}
+\begin{tikzpicture}
+\begin{axis}[
+	height=0.8\textheight,
+	width=\textwidth,
+	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
+	cycle list name=earthy,
+	no marks,
+	xmin=0, xmax=20,ymin=0,
+	]
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
+\end{axis}
+\end{tikzpicture}
+
+Caseins regain some solubility at low pH.
+\end{frame}
+
 \end{document}

### 2014-06-15 seminaire_Tanakalab.tex
use the enhanced version of the creep_experiment video, update Couette sketch

@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{Acid-set protein gel (Yoghurt)}
@@
 
-\section{Creep and yielding}
+\section*{Creep and yielding}
 
@@
 
-%\tikzset{external/force remake}
+\tikzset{external/force remake}
 \begin{frame}{The yoghurt creep experiment}
@@
 \node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{creep_experiment.jpg}};
-\begin{scope}[shift={(a.north west)}, shift={(7.2\baselineskip,-1.1\baselineskip)}, ultra thick]
-	\draw[Accent1] ellipse[x radius=2.9\baselineskip, y radius=0.7\baselineskip] ++(2.9\baselineskip,0)  -- ++(0,-6.6\baselineskip)  ++(-2.9\baselineskip,0) ellipse[x radius=2.9\baselineskip, y radius=0.7\baselineskip] ++(-2.9\baselineskip,0)  -- ++(0,6.6\baselineskip);
-	\shadedraw[Accent2,ball color=Accent2] ellipse[x radius=2\baselineskip, y radius=0.4\baselineskip] ++(2\baselineskip,0)  -- ++(0,-6.6\baselineskip)  ++(-2\baselineskip,0) ellipse[x radius=2\baselineskip, y radius=0.4\baselineskip] ++(-2\baselineskip,0)  -- ++(0,6.6\baselineskip);
+\begin{scope}[shift={(a.north west)}, shift={(5.8\baselineskip,-1.1\baselineskip)}, ultra thick]
+	\draw[Accent1, x radius=3.1\baselineskip, y radius=0.3\baselineskip] ellipse ++(3.1\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-3.1\baselineskip,0) ellipse[] ++(-3.1\baselineskip,0)  -- ++(0,7.4\baselineskip);
+	\shadedraw[Accent2,ball color=Accent2, x radius=2\baselineskip, y radius=0.1\baselineskip] ellipse ++(2\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-2\baselineskip,0) ellipse[] ++(-2\baselineskip,0)  -- ++(0,7.4\baselineskip);
 	\fill[Accent2, left color=black,right color=black, middle color=Accent2, ] (-0.3\baselineskip,0) rectangle +(0.6\baselineskip, \baselineskip);
 	\begin{scope}[Main]
-		\node[draw, minimum height=3.5\baselineskip, minimum width=1.2\baselineskip] at (2.4\baselineskip,-3.35\baselineskip) (window){};
-		\node[draw, minimum height=7.5\baselineskip, minimum width=4\baselineskip] at (7.5\baselineskip,-4\baselineskip) (USV){};
+		\node[draw, minimum height=4.2\baselineskip, minimum width=1.5\baselineskip] at (2.5\baselineskip,-4\baselineskip) (window){};
+		\node[draw, minimum height=7.2\baselineskip, minimum width=4\baselineskip] at (8.9\baselineskip,-4.2\baselineskip) (USV){};
 		\draw(window.north east) -- (USV.north west) (window.south east) -- (USV.south west);
@@
 \draw[->] (window.north west) ++(-1em, 0.2em) -- +(0,-1em)  node[midway,left] {$z$};
+\draw[<->] ++(-2\baselineskip,-\baselineskip) -- +(-1.1\baselineskip,0) node[midway,above] {\SI{2}{\milli\metre}};
 \node[text width=4\baselineskip] at (-3\baselineskip,-2.5\baselineskip) {shear rate response};

### 2014-06-15 seminaire_Tanakalab.tex
Bonn_fracture and Villi centred
USV velocity map

@@
 \begin{itemize}
-\item Linear elasticity in not well understood \textit{\scriptsize Gardel et al. Science 2004}
+\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
 \item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
@@
 \end{itemize}
-\includegraphics[height=3\baselineskip]{Villi_sq}
+\begin{tabu}{X[c]X[c]}
+\includegraphics[height=6\baselineskip]{Bonn_fracture} &
+\includegraphics[height=6\baselineskip]{Villi_sq}
+\end{tabu}
 \begin{block}{Our questions}
@@
 \begin{frame}{Acid-set protein gel (Yoghurt)}
-\begin{block}{Recipe}
+\begin{block}{Recipe: take your time}
 \begin{itemize}
@@
 \raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
-\item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow acidification
+\item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow homogeneous acidification
 \end{itemize}
@@
 
-\begin{frame}{Merciless yoghurt breakers}
+\begin{frame}{Creep and yielding}
+\structure{Merciless yoghurt breakers}
+
+\bigskip
 \begin{tabu}{X[c]X[c]c}
@@
 \node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{creep_experiment.jpg}};
+\node[below right=\baselineskip of a.north east, inner sep=0] (vmap) {\includegraphics[width=4\baselineskip]{jet_cmap}};
+
+\begin{scope}[font=\footnotesize]
+	\node[above=0 of vmap] {velocity map};
+	\node[below=0 of vmap.south west] {-50};
+	\node[below=0 of vmap.south] (v0) {0};
+	\node[below=0 of vmap.south east] {50};
+	\node[below=0 of v0, inner sep=0]{\si{\micro\metre\per\second}};
+\end{scope}
+
 \begin{scope}[shift={(a.north west)}, shift={(5.8\baselineskip,-1.1\baselineskip)}, ultra thick]
@@
 
-\section{Wrinkling}
+\section*{Wrinkling}
 

### 2014-06-15 seminaire_Tanakalab.tex
Basquin

@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{The yoghurt creep experiment}
@@
 
+\tikzset{external/force remake}
+\begin{frame}{Failure time}
+\begin{columns}
+\column{0.55\textwidth}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xlabel={$\sigma$ (\si{\pascal})},
+	ylabel={$\tau_f$ (\si{\second})},
+	xmin=100, ymin=10, ymax=1e6,
+	xtick={100, 200,..., 1000}, xticklabels={100, 200,,, 500,,,,, 1000},
+	]
+	\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
+	\addplot[no marks, black, domain={100:1000}] {4.2e17*x^(-5.45)} node [midway, above right] {$\tau_f\sim \sigma^{-5.5}$};
+\end{loglogaxis}
+\end{tikzpicture}
+
+\begin{block}{Basquin law}
+\begin{itemize}
+\item fatigue
+\item heterogeneous solids
+\item no yield stress
+\end{itemize}
+\end{block}
+
+\column{0.45\textwidth}
+\begin{tikzpicture}
+\begin{groupplot}[
+	group style={
+		group name=g, group size=1 by 2,
+		},
+	width=\textwidth,
+	height=0.7\textwidth,
+	ymode=log,
+	ylabel={$\tau_f$ (\si{\second})},
+	ymin=10, ymax=1e6,
+	]
+	\nextgroupplot[
+		xlabel={$\sigma$ (\si{\pascal})},
+		xtick={200,600,1000},
+		]
+		\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
+		\addplot[no marks, black, domain={100:1000}] {6.5e5*exp(-x/84)} node [pos=0.25, above right] {$\tau_f\nsim e^{-\frac{\sigma}{\sigma_0}}$};
+
+	\nextgroupplot[
+		xlabel={$1/\sigma$ (\si{\per\pascal})},
+		]
+		\addplot[only marks, Accent1] table[x expr=1/\thisrowno{0}, y index=3]{creep_cas4_gdl1_MCR.txt};
+		\addplot[no marks, black, domain={1e-3:7e-3}] {20.3*exp(1720*x)} node [pos=0.25, below right] {$\tau_f\nsim e^\frac{\sigma_0}{\sigma}$};
+\end{groupplot}
+\end{tikzpicture}
+
+Activated processes unlikely (Eyring, Griffith, Taylor, \ldots)
+\end{columns}
+\end{frame}
+\tikzset{external/force remake=false}
+
 \section*{Wrinkling}

### 2014-06-16 seminaire_Tanakalab.tex
Primary creep, Andrade_1910

@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{Failure time}
@@
 \begin{itemize}
-\item fatigue
+\item fatigue (oscillatory stress)
 \item heterogeneous solids
@@
 
+\tikzset{external/force remake}
+\begin{frame}{Primary creep}
+\begin{columns}
+\column{0.55\textwidth}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
+	ylabel={Strain rate $\dot{\gamma}$ (\si{\per\second})},
+	cycle list name=earthy,
+	no marks,
+	]
+	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[pos=0.75, below left] {$\dot{\gamma}(t)\sim t^{-0.85}$};
+\end{loglogaxis}
+\end{tikzpicture}
+
+\begin{block}{Primary creep $\Rightarrow$ ``Andrade'' creep}
+Power-law creep known in \alert{solids} for over a century
+
+Classically explained by disinclination dynamics in crystals
+\end{block}
+
+\column{0.45\textwidth}
+\includegraphics[width=\textwidth]{Andrade_1910}
+
+\bigskip
+\includegraphics[height=7\baselineskip]{Andrade_1910_fig1}
+\begin{tikzpicture}
+\node[inner sep=0] (a){\includegraphics[height=7\baselineskip]{Andrade_1910_fig2}};
+\node[below right=0 and 1em of a.north west, font=\footnotesize,fill=white] {lead \& copper};
+\node[above left=1em of a.south east, Accent2] {$\gamma-\gamma_0 \sim t^{1/3}$};
+\end{tikzpicture}
+\end{columns}
+\end{frame}
+\tikzset{external/force remake=false}
+
 \section*{Wrinkling}

### 2014-06-16 seminaire_Tanakalab.tex
Primary creep is kinetically reversible

@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{Primary creep}
@@
 
+\tikzset{external/force remake}
+\begin{frame}{Primary creep}
+\begin{block}{Local insight from \movie[externalviewer]{ultrasounds}{Yaourt110_primary.avi}}
+\begin{itemize}
+	\item homogeneous \movie[externalviewer]{strain field}{Yaourt110_profils.avi} (no wall slip, no shear band)
+	\item if present, plastic events are below our resolution (a few \si{\micro\metre})
+\end{itemize}
+\end{block}
+
+\begin{block}{Partial creep and recovery}
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=8\baselineskip,
+	xlabel={time (\si{\second})},
+	ylabel={strain $\gamma$},
+	xmin=-1e3, xmax=2.8e4,
+	]
+	\addplot[black, no marks] table {Y261_oscil_creep_recovery_100Pa_decimated.txt};
+	\begin{scope}[<-,Main, inner sep=0]
+	\draw (axis cs:0,0) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow \SI{100}{\pascal}$};
+	\draw (axis cs:600,0.2727007) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow 0$};
+	\draw (axis cs:8600,0.0159) -- +(0,1em) node[anchor=base east]{$\gamma\leftarrow 0$};
+	\end{scope}
+\end{axis}
+\end{tikzpicture}
+\end{block}
+
+$\Rightarrow$ Primary creep is ``kinetically'' reversible (no detectable structure damage, but viscous dissipation)
+
+\end{frame}
+\tikzset{external/force remake=false}
+
 \section*{Wrinkling}

### 2014-06-16 seminaire_Tanakalab.tex
Dimensionless creep response, relationship with linear rheology

@@
 	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime,G^{\prime\prime}\sim f^{0.15}$};
+	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
 %	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
@@
 	xmin=-1e3, xmax=2.8e4,
+	axis background/.style={fill=white},
 	]
@@
 $\Rightarrow$ Primary creep is ``kinetically'' reversible (no detectable structure damage, but viscous dissipation)
+\end{frame}
+\tikzset{external/force remake=false}
 
+\tikzset{external/force remake}
+\begin{frame}{Dimensionless creep response}
+\begin{columns}
+\column{0.55\textwidth}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
+	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+	cycle list name=earthy,
+	no marks,
+	]
+	\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.2*x^(-0.85)} node[pos=0.75, below left] {$\frac{\dot{\gamma}}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
+\end{loglogaxis}
+\end{tikzpicture}
+
+Linear viscoelasticity accounts for the Andrade exponent
+
+\column{0.45\textwidth}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	height=0.8\textwidth,
+	width=\textwidth,
+	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{$G^\prime$}, \textcolor{Accent2}{$G^{\prime\prime}$}(\si{\pascal})},
+	domain={6e-2:70},
+	ymin=10,
+	]
+	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
+	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
+	\addplot[Main, no marks]{100*x^0.15}  node[midway, below=1em] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
+
+\end{loglogaxis}
+\end{tikzpicture}
+\textcolor{Main}{
+\begin{align*}
+\Rightarrow & J(t) \equiv& \frac{\gamma(t)}{\sigma}\sim& t^{0.15}\\
+\Rightarrow & \dot{\gamma}(t)\sim& t^{0.15-1} =& t^{-0.85}
+\end{align*}
+}
+\end{columns}
 \end{frame}

### 2014-06-16 seminaire_Tanakalab.tex
Tertiary creep, fracture images and length

@@
 	]
-	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node (s200){};
-	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){};
-	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
-	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
-	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node[left] (s200){200};
+	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){300};
+	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){400};
+	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){550};
+	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){1000};
 	\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[pos=0.75, below left] {$\dot{\gamma}(t)\sim t^{-0.85}$};
@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{Dimensionless creep response}
@@
 	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.2*x^(-0.85)} node[pos=0.75, below left] {$\frac{\dot{\gamma}}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
+	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.2*x^(-0.85)} node[pos=0.75, below left] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
 \end{loglogaxis}
@@
 
+\tikzset{external/force remake}
+\begin{frame}{Tertiary creep}
+\begin{columns}
+\column{0.55\textwidth}
+Time to failure
+
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=1e-5, xmax=2, xlabel={$(\tau_f-t)/\tau_f$},
+	x dir=reverse,
+	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+	cycle list name=earthy,
+	no marks,
+	]
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.05/x} node[pos=0.6, below right, inner sep=0] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{\tau_f}{\tau_f-t}\right)$};
+\end{loglogaxis}
+\end{tikzpicture}
+
+Close to failure: 
+\begin{itemize}
+\item finite time singularity
+\item logarithmic divergence
+\end{itemize}
+$\Rightarrow$ dominated by fracture growth
+
+\column{0.45\textwidth}
+\begin{block}{Fracture growth}
+\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-15-00.jpg}\hfill
+\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-20-00.jpg}\hfill
+\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-21-15.jpg}
+
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=1e-3, xmax=1, xlabel={$(\tau_f-t)/\tau_f$},
+	xmode=log, x dir=reverse,
+	ylabel={$\ell/H$},
+	ymin=0, ymax=0.4,
+	axis background/.style={fill=white},
+	]
+	\addplot[only marks, Accent1, error bars/.cd, y dir=both, y explicit] table[y error index=2]{Y110_fracturelength.txt};
+	\addplot[Accent2, ultra thick, no marks, domain={1e-3:1}] {8.9e-3 - 0.06 * ln(x)};
+\end{axis}
+\end{tikzpicture}
+\end{block}
+
+\end{columns}
+\end{frame}
+\tikzset{external/force remake=false}
+
 \section*{Wrinkling}

### 2014-06-17 seminaire_Tanakalab.tex
Accepted in PRL gap at the right place

@@
 \end{tabu}
+
+\bigskip
+Just accepted in PRL, \textit{ArXiv 1401.8234}
+
 \end{frame}
@@
 	\end{scope}
+
+\draw[<->,thin] ++(-2\baselineskip,-\baselineskip) -- +(-1.1\baselineskip,0) node[midway,above] {\SI{2}{\milli\metre}};
 \end{scope}
@@
 \draw[->] (window.north west) ++(-1em, 0.2em) -- +(0,-1em)  node[midway,left] {$z$};
-\draw[<->] ++(-2\baselineskip,-\baselineskip) -- +(-1.1\baselineskip,0) node[midway,above] {\SI{2}{\milli\metre}};
+
 \node[text width=4\baselineskip] at (-3\baselineskip,-2.5\baselineskip) {shear rate response};
@@
 	ylabel={$\ell/H$},
-	ymin=0, ymax=0.4,
+	ymin=0, ymax=0.5,
 	axis background/.style={fill=white},
@@
 	\addplot[only marks, Accent1, error bars/.cd, y dir=both, y explicit] table[y error index=2]{Y110_fracturelength.txt};
-	\addplot[Accent2, ultra thick, no marks, domain={1e-3:1}] {8.9e-3 - 0.06 * ln(x)};
+	\addplot[Accent2, ultra thick, no marks, domain={1e-3:1}] {- 0.06 * ln(x)} node [pos=0.25,left=1em] {$\frac{\ell}{H}\sim \ln\frac{\tau_f-t}{\tau_f}$};
 \end{axis}

### 2014-06-17 seminaire_Tanakalab.tex
Secondary creep, Monkman-Grant, summary symbols in earthy list, fits in yellow

@@
 \pgfplotscreateplotcyclelist{earthy}{%
-red!40!black,
-red!60!black,
-red!80!black,
-red,
-red!80!yellow,
+{red!40!black,mark=o},
+{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
+{red!80!black,mark=square},
+{red,mark=triangle},
+{red!80!yellow, mark=diamond},
 red!60!yellow,
@@
 	cycle list name=earthy,
-	no marks,
+	clip mode=individual,
 	]
@@
 	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.2*x^(-0.85)} node[pos=0.75, below left] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
+	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.378*x^(-0.85)} node[Accent2,pos=0.75, below left] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
 \end{loglogaxis}
@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{Tertiary creep}
@@
 	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.05/x} node[pos=0.6, below right, inner sep=0] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{\tau_f}{\tau_f-t}\right)$};
+	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.187/x} node[Accent2, pos=0.6, below right, inner sep=0] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{\tau_f}{\tau_f-t}\right)$};
 \end{loglogaxis}
@@
 	ymin=0, ymax=0.5,
+	clip mode=individual,
 	axis background/.style={fill=white},
@@
 \end{block}
-
 \end{columns}
@@
 \tikzset{external/force remake=false}
+
+%\tikzset{external/force remake}
+\begin{frame}{Secondary creep: a crossover}
+\begin{columns}
+\column{0.55\textwidth}
+Creep responses in linear scales
+
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=0, xmax=1, xlabel={$t/\tau_f$},
+	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+	ymin=0.5, ymax=4,
+	cycle list name=earthy,
+	restrict y to domain=0.5:10,
+	clip mode=individual,
+	]
+	\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot[yellow, ultra thick, domain={0.01:0.99}] {0.378*x^(-0.85) + 0.187/(1-x)};
+	\draw[<-] (axis cs:0.56,1) -- +(0,1em) node[above] {$\tau_\text{min}$};
+\end{axis}
+\end{tikzpicture}
+\begin{block}{Master curve}
+\[\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}} = \underbrace{\lambda \left(\frac{t}{\tau_f}\right)^{-0.85}}_\text{Andrade} + \underbrace{\frac{\mu}{1 - t/\tau_f}}_\text{fractures}\]
+\end{block}
+
+\column{0.45\textwidth}
+\structure{Monkman-Grant relation}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xlabel={$\tau_f$ (\si{\second})},
+	ylabel={$\tau_\text{min}$ (\si{\second})},
+	xmin=10, xmax=1e6, ymin=10, ymax=1e6,
+	]
+	\addplot[only marks, Accent1] table[x index=3, y index=4]{creep_cas4_gdl1_MCR.txt};
+	\addplot[no marks, black, domain={10:1e6}] {0.56*x)};
+\end{loglogaxis}
+\end{tikzpicture}
+\[\tau_\text{min} \approx 0.6 \tau_f\]
+``Early time'' response allows to predict failure time
+\end{columns}
+\end{frame}
+\tikzset{external/force remake=false}
+
+\begin{frame}{Creep and yielding summary}
+\begin{itemize}
+\item failure involves a single time scale $\tau_f \sim \sigma^{-5.5}$ Basquin law
+\item reversible, homogeneous creep $\rightarrow$ irreversible fracture growth
+\item a single expression captures all the global rheological response
+\end{itemize}
+\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=3 by 1,
+			horizontal sep=3em,
+			},
+		width=0.35\textwidth,
+		cycle list name=earthy,
+		clip mode=individual,
+		]
+	\nextgroupplot[
+		xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
+		ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+		xmode=log,ymode=log,
+		ytick={1e-4, 1e-2, 1},
+		]
+		\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*x^(-0.85) + 0.187/(1-x)};
+
+	\nextgroupplot[
+		xmin=0, xmax=1, xlabel={$t/\tau_f$},
+		ymin=0.5, ymax=4,
+		restrict y to domain=0.5:10,
+		]
+		\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		\addplot[yellow, ultra thick, domain={0.01:0.99}] {0.378*x^(-0.85) + 0.187/(1-x)};
+
+	\nextgroupplot[
+		xmin=1e-5, xmax=2, xlabel={$(\tau_f-t)/\tau_f$},
+		x dir=reverse,
+		xmode=log,ymode=log,
+		ytick={1e-4, 1e-2, 1},
+	]
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*(1-x)^(-0.85) + 0.187/x};
+	\end{groupplot}
+\end{tikzpicture}
+\end{frame}
 

### 2014-06-17 seminaire_Tanakalab.tex
Theory and Models

@@
 \begin{frame}{Creep and yielding}
+Just accepted in PRL, \textit{ArXiv 1401.8234}
+
+\bigskip
 \structure{Merciless yoghurt breakers}
@@
 \bigskip
-Just accepted in PRL, \textit{ArXiv 1401.8234}
+\begin{description}[Special thanks]
+\item[Special thanks] to Alan Parker (Firmenich) for prompting this study and providing casein \& GDL
+\end{description}
 
@@
 
+\begin{frame}{What about theories and models?}
+\begin{columns}
+\column{0.35\textwidth}
+\begin{block}{Discrete dislocation dynamics (dense)}
+\textit{\footnotesize Miguel et al., PRL 2002}
+
+\hspace{1em}\includegraphics[width=\textwidth-2em]{Miguel_2002_img}
+
+\medskip
+\hspace{1em}\includegraphics[width=\textwidth-2em]{Miguel_2002_graph}
+
+predicts Andrade creep \alert{but} no fracture
+\end{block}
+\column{0.65\textwidth}
+\begin{columns}
+\column{0.5\textwidth+1em}
+\structure{Fibre bundle model~\#1}
+
+\textit{\footnotesize Jagla et al., PRE 2011}
+\begin{itemize}
+\item elastic fibres
+\item local yield strain
+\item local coupling
+\end{itemize}
+
+\hspace{1em}\includegraphics[width=\textwidth-1em]{Jagla_2011_gamma}
+\[\tau_f\sim\left(\sigma-\alert{\sigma_c}\right)^{-1.25}\]
+
+\column{0.5\textwidth-1em}
+\includegraphics[width=\textwidth]{Jagla_2011_3regimes}
+\end{columns}
+predicts Andrade creep and failure \alert{but} critical stress
+\end{columns}
+\end{frame}
+
+\begin{frame}{What about theories and models?}
+\structure{Fibre bundle model~\#2}
+
+elastic fibres + local yield strain + damage accumulation
+\begin{block}{\textit{\footnotesize Kun et al. J. Stat Mech 2006, PRL 2008}}
+\begin{columns}
+\column{0.7\textwidth}
+\begin{tikzpicture}[inner sep=0]
+\node(a){\includegraphics[height=5\baselineskip]{Kun_2006_basquin.png}};
+\node[fill=white,yshift=-0.5em] at (a) {$\tau_f \sim \sigma^{-\beta}$};
+\end{tikzpicture}\,
+\includegraphics[height=5\baselineskip]{Kun_2008_Basquin}\,
+\includegraphics[height=5\baselineskip]{asphalt}
+\column{0.3\textwidth}
+$\beta$ exponent of damage accumulation
+\[\dot{\gamma}\sim \left(\tau_f-t\right)^{-1/(1+\beta)}\]
+\end{columns}
+\end{block}
+
+\begin{columns}
+\column{0.65\textwidth}
+\begin{block}{\textit{\footnotesize Halász et al., PRE 2012}}
+\begin{tikzpicture}[inner sep=0, ultra thick, font=\footnotesize]
+\node (g) {\includegraphics[height=7\baselineskip]{Halasz_2012_Basquin}};
+\node[draw=green!50!black, below right=0 and \baselineskip of g.north east] (c) {\includegraphics[height=4\baselineskip]{Halasz_2012_crack}};
+\node[draw=red, above right=0 and 6\baselineskip of g.south east] (d) {\includegraphics[height=4\baselineskip]{Halasz_2012_disordered}};
+\draw[->,green!50!black] (c.west) -- ++(-2\baselineskip,0) node[above] {$\beta=5$} node[below right=0 of c.north east, text width=5\baselineskip]{single crack propagation};
+\draw[->, red] (d.south west) -- +(-5\baselineskip,0) --($(g.south east) +(0, \baselineskip)$) node[below=1em] {$\beta=1$} node[above left=0 of d.south west] {diffusive damage};
+\end{tikzpicture}
+\end{block}
+\column{0.35\textwidth}
+Predicts Basquin law \& fractures \alert{but} no Andrade creep \& too slow divergence
+\end{columns}
+\end{frame}
+
 \begin{frame}{Creep and yielding summary}
@@
 		xmode=log,ymode=log,
-		ytick={1e-4, 1e-2, 1},
+		xtick={1e-4, 1e-2, 1},
 		]
@@
 		xmode=log,ymode=log,
-		ytick={1e-4, 1e-2, 1},
+		xtick={1e-4, 1e-2, 1},
 	]
@@
 \end{tikzpicture}
+\begin{itemize}
+\item a model soft solid well captured by fibre-bundle models
+\end{itemize}
 \end{frame}

### 2014-06-17 creep_beamer.tex
split creep_beamer.tex from seminaire_Tanakalab.tex

--- 2014-06-17 seminaire_Tanakalab.tex
@@
-\documentclass[xcolor=table]{beamer}
-\usepackage[utf8]{inputenc}
-\usepackage[francais]{babel}
-%\usetheme{Boadilla}
-%\usecolortheme{rose}
-%\usecolortheme{crane}
-\usefonttheme{structuresmallcapsserif}
-\setbeamertemplate{navigation symbols}{}
-
-\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
-\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
-\definecolor{Accent2}{rgb}{0.54,0.1,0.4}
-
-\usecolortheme{rose}
-%\useinnertheme[shadow]{circles}
-\usecolortheme{whale}
-%\useoutertheme{infolines}
-
-\usecolortheme[named=Accent1]{structure}
-
-\setbeamercolor{alerted text}{fg=Accent2}
-%\setbeamercolor{palette primary}{fg=white}
-%\setbeamercolor{palette secondary}{bg=Accent1}
-%\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}
-
-\setbeamerfont{page number in head/foot}{size=\large}
-\setbeamercolor{page number in head/foot}{fg=Main}
-% page/total
-%\setbeamertemplate{footline}[frame number]
-% pas de total
-\setbeamertemplate{footline}{%
-    	\hfill%
-	\usebeamercolor[fg]{page number in head/foot}%
-	\usebeamerfont{page number in head/foot}%
-	\insertframenumber\kern1em\vskip2pt%
-}
-
-\setbeamersize{text margin left=1em}
-\setbeamersize{text margin right=1em}
-
-%font
-\usepackage[T1]{fontenc}
-\usepackage[oldstylenums]{kpfonts}
-
-%proper math and math symbols
-%\usepackage{amsmath}
-\usepackage{amssymb}
-
-\usepackage{datenumber,fp}
-
-\usepackage{siunitx}
-
-\usepackage{tabu}
-\usepackage{multirow}
-\usepackage{booktabs}
-
-% Allow the usage of graphics (.jpg, .png, etc.) in the document
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
-%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
-
-\usepackage{pgfplots}
-\pgfplotsset{compat=1.9}
-\usepgfplotslibrary{groupplots}
-\usepgfplotslibrary{external}
-\tikzexternalize
-\tikzsetexternalprefix{fig_presentation/}
-%\tikzset{external/optimize=false}
-%\tikzset{external/force remake}
-
-%link or play movies
-\usepackage{multimedia}
-
-%beamer related package
-
-\usepackage{todonotes}
-\presetkeys{todonotes}{inline}{}
-
-%bibliography
-\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
-%minimal citations
-\AtEveryCitekey{%
-	\clearfield{title}
-	\clearfield{pages}
-	\clearfield{volume}
-	\clearfield{number}
-	\clearfield{month}}
-\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{%
-  \printtext{\bibstring{in}\intitlepunct}}}
-%\bibliography{biblio}
-
-\newcolumntype{P}[1]{>{\raggedright}p{#1}}
-
-\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Supérieure de Lyon}
-\title[Yoghurt under stress]{Yoghurt under stress}
-\author[M. Leocmach]{Mathieu Leocmach}
-\date{1st June 2014}
-\titlegraphic{
-	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
-	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
-	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
-	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
-	}
-
-\begin{document}
-\tikzset{every mark/.append style={scale=0.8}}
-\pgfplotsset{every axis/.append style={footnotesize}}
-
-\pgfplotscreateplotcyclelist{earthy}{%
-{red!40!black,mark=o},
-{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
-{red!80!black,mark=square},
-{red,mark=triangle},
-{red!80!yellow, mark=diamond},
-red!60!yellow,
-red!40!yellow,
-}
-
-\AtBeginSection[]{
-	\addtocounter{framenumber}{-1}
-	\begin{frame}[plain]
-		\tableofcontents[currentsection, hideothersubsections]
-	\end{frame}
-}
-
-\begin{frame}[plain]
-	\titlepage
-\end{frame}
-
-\setcounter{framenumber}{0}
-
-\begin{frame}{Application of biogels}
-\begin{tabu}{cX[c]c}
-\includegraphics[height=0.3\textheight]{cell_mech} &
-\includegraphics[height=0.3\textheight]{electrophoresis} &
-\includegraphics[height=0.3\textheight]{cosmetics} \\
-Cell mechanics & Electrophoresis & Cosmetics\\
-\end{tabu}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=0.3\textheight]{bacterial_culture} &
-\includegraphics[height=0.3\textheight]{food} \\
-Bacterial culture & Food\\
-\end{tabu}
-
-\begin{footnotesize}
-\begin{block}{Sources}
-\begin{tabu}{lXl}
-Wikimedia Commons & www.madaboutscience.com & www.keautystore.com\\
-\end{tabu}
-\end{block}
-\end{footnotesize}
-\end{frame}
-
-\begin{frame}{Fundamental issues in biogels}
-\begin{itemize}
-\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
-\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
-\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
-\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
-\end{itemize}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=6\baselineskip]{Bonn_fracture} &
-\includegraphics[height=6\baselineskip]{Villi_sq}
-\end{tabu}
-\begin{block}{Our questions}
-\begin{itemize}
-\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
-\item What is the role of the solvent in mechanical instabilities?
-\end{itemize}
-\end{block}
-\end{frame}
-
-%\section{Our system: Yoghurt}
-
-%\tikzset{external/force remake}
-\begin{frame}{Acid-set protein gel (Yoghurt)}
-\begin{block}{Recipe: take your time}
-\begin{itemize}
-\item Water (\SI{30}{\celsius})
-\item Sodium caseinate (milk protein) 4\%
-\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
-\item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow homogeneous acidification
-\end{itemize}
-\end{block}
-
-\begin{columns}
-\column{0.5\textwidth}
-\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		%xmode=log,
-		%xmin=1e2,xmax=3e5,
-		xmin=0, xmax=20,
-		scale only axis,
-		width=\textwidth-4em,
-		height=0.3\textwidth,
-		extra tick style={grid=major},%
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
-		]
-	\nextgroupplot[
-		ymin=0, ymax=7, ylabel={pH},
-		extra y ticks={4.6}, extra y tick labels={},%
-		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y190_cas4_gdl1.pH};
-	\draw[help lines] (axis cs:0,4.6) -- (axis cs:3.88,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:3.88,0);
-
-	\nextgroupplot[
-		xlabel={time (\si{\hour})},
-		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
-		extra x ticks={17}, extra x tick labels={},
-		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y22.prise};
-	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL1_Y22.prise};
-	\begin{scope}[font=\scriptsize]
-		%\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
-		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
-		\node[anchor=base,rotate=80] at (axis cs:2.5,400) {gelation};
-	\end{scope}
-	\end{groupplot}
-	\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
-\end{tikzpicture}
-
-\column{0.5\textwidth}
-\includegraphics[width=\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
-
-\begin{scriptsize}
-\textit{Kaláb (1983),\linebreak Roefs \& van Vliet (1990),\linebreak Lucey \& Singh (1998)}
-\end{scriptsize}
-\end{columns}
-\end{frame}
-\tikzset{external/force remake=false}
-
-\begin{frame}{Linear rheology: Power law (visco)elastic solid}
-\[\text{stress}\rightarrow \sigma = G \gamma\leftarrow\text{strain}\]
-\[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{loss} moduli (\si{\pascal})},
-	domain={6e-2:70},
-	]
-	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot{590*x^0.14};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot{350*x^0.12};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot{246*x^0.1};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot{163*x^0.067};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot{109*x^0.044};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot{76*x^0.035};
-
-\end{loglogaxis}
-\end{tikzpicture}
-\end{frame}
-
-\section*{Creep and yielding}
-
 \begin{frame}{Creep and yielding}
@@
 \end{frame}
-
-\section*{Wrinkling}
-
-%\tikzset{external/force remake}
-\begin{frame}{Over-acidification: Faster but weaker}
-\begin{tikzpicture}
-\begin{axis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
-	cycle list name=earthy,
-	no marks,
-	xmin=0, xmax=20,ymin=0,
-	]
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
-\end{axis}
-\end{tikzpicture}
-
-Caseins regain some solubility at low pH.
-\end{frame}
-
-\end{document}

### 2014-06-17 seminaire_Tanakalab.tex
split creep_beamer.tex from seminaire_Tanakalab.tex

--- 2014-06-17 seminaire_Tanakalab.tex
@@
 
-\begin{frame}{Creep and yielding}
-\structure{Merciless yoghurt breakers}
-
-\bigskip
-\begin{tabu}{X[c]X[c]c}
-\includegraphics[height=0.3\textheight]{Chris}&
-\includegraphics[height=0.3\textheight]{Thibaut}&
-\includegraphics[height=0.3\textheight]{Seb}\\
-Christophe Perge & Thibaut Divoux & Sebastien Manneville\\
-PhD student & CNRS researcher & Professor\\
-E.N.S. Lyon &  CRPP Bordeaux & E.N.S. Lyon\\
-\end{tabu}
-
-\bigskip
-Just accepted in PRL, \textit{ArXiv 1401.8234}
-
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{The yoghurt creep experiment}
-\begin{tikzpicture}
-\node[anchor=south west] (webcam){\includegraphics[width=0.2\textwidth]{webcam}};
-\node[anchor=south] at (0.4\textwidth,0) (spoon) {\includegraphics[width=0.3\textwidth]{spoon}};
-\node[anchor=south east] at (\textwidth,0) (ultrasound) {\includegraphics[width=0.4\textwidth]{ultrasound}};
-
-\node[anchor=south west] at ($(ultrasound.north west)+(0,2em)$) (rheometer) {\includegraphics[height=0.4\textheight]{rheometer}};
-
-\begin{scope}[draw]
-	\node[below=0of spoon.south] (couette) {A Taylor-Couette cell where the yogurt is made \emph{in situ}};
-	\draw[->] (couette) -- ($(spoon.south)+(0,2em)$);
-	\node[above=0of ultrasound] {ultrasonic imaging};
-	\node[above=0of webcam, text width=0.2\textwidth, align=center] {optical imaging};
-	\draw[->] (spoon.north east) +(235:2em) -- (rheometer);
-	\node[right=0of rheometer, text width=0.2\textwidth, align=center] {A rheometer imposes a constant stress $\sigma$ and records the strain $\gamma(t)$};
-\end{scope}
-\end{tikzpicture}
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{The yoghurt creep experiment}
-\movie[externalviewer]{\begin{tikzpicture}
-\node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{creep_experiment.jpg}};
-\node[below right=\baselineskip of a.north east, inner sep=0] (vmap) {\includegraphics[width=4\baselineskip]{jet_cmap}};
-
-\begin{scope}[font=\footnotesize]
-	\node[above=0 of vmap] {velocity map};
-	\node[below=0 of vmap.south west] {-50};
-	\node[below=0 of vmap.south] (v0) {0};
-	\node[below=0 of vmap.south east] {50};
-	\node[below=0 of v0, inner sep=0]{\si{\micro\metre\per\second}};
-\end{scope}
-
-\begin{scope}[shift={(a.north west)}, shift={(5.8\baselineskip,-1.1\baselineskip)}, ultra thick]
-	\draw[Accent1, x radius=3.1\baselineskip, y radius=0.3\baselineskip] ellipse ++(3.1\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-3.1\baselineskip,0) ellipse[] ++(-3.1\baselineskip,0)  -- ++(0,7.4\baselineskip);
-	\shadedraw[Accent2,ball color=Accent2, x radius=2\baselineskip, y radius=0.1\baselineskip] ellipse ++(2\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-2\baselineskip,0) ellipse[] ++(-2\baselineskip,0)  -- ++(0,7.4\baselineskip);
-	\fill[Accent2, left color=black,right color=black, middle color=Accent2, ] (-0.3\baselineskip,0) rectangle +(0.6\baselineskip, \baselineskip);
-	\begin{scope}[Main]
-		\node[draw, minimum height=4.2\baselineskip, minimum width=1.5\baselineskip] at (2.5\baselineskip,-4\baselineskip) (window){};
-		\node[draw, minimum height=7.2\baselineskip, minimum width=4\baselineskip] at (8.9\baselineskip,-4.2\baselineskip) (USV){};
-		\draw(window.north east) -- (USV.north west) (window.south east) -- (USV.south west);
-		\node[right, text width=4\baselineskip] at (USV.east) {ultrasound region of interest};
-	\end{scope}
-
-\draw[<->,thin] ++(-2\baselineskip,-\baselineskip) -- +(-1.1\baselineskip,0) node[midway,above] {\SI{2}{\milli\metre}};
-\end{scope}
-\draw[->] (window.north west) ++(-1em, 0.2em) -- +(1em,0) node[midway,above,inner sep=1pt] {$r$};
-\draw[->] (window.north west) ++(-1em, 0.2em) -- +(0,-1em)  node[midway,left] {$z$};
-
-\node[text width=4\baselineskip] at (-3\baselineskip,-2.5\baselineskip) {shear rate response};
-\node at (5\baselineskip,-2.5\baselineskip) {strain response};
-\end{tikzpicture}%
-}{/home/mathieu/Dropbox/Yaourt/films/Yaourt141_enhanced.avi}
-\end{frame}
-\tikzset{external/force remake=false}
-
-\begin{frame}{Creep rheology: Three regimes}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-   name=g,
-	width=\textwidth,
-	height=0.8\textheight,
-	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
-	ymin=0.2,ymax=2,
-	ytick={0.2, 0.4,0.8,1.6}, yticklabels={0.2, 0.4,0.8,1.6},
-	ylabel={strain $\gamma$},
-	extra tick style={grid=major},%
-	extra y ticks={1}, extra y tick labels={1.0},
-	cycle list name=earthy,
-	axis on top,
-	]
-	\fill[Accent1!20]
-		(axis cs:10,0.4) ellipse[rotate=-30, x radius=0.2\textwidth, y radius=0.12\textwidth]
-		(axis cs:2e2,1.6) ellipse[x radius=0.3\textwidth, y radius=0.05\textwidth]
-		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
-		;
-	\node[above right] at (axis cs:1e1,0.2) {primary};
-	\node[below left] at (axis cs:4e5,0.7) {secondary};
-	\node[left] at (axis cs:5,1.6) {tertiary};
-	\addplot table{Y27_200Pa_gamma_decimated.txt} node (s200){};
-	\addplot table{Y38_300Pa_gamma_decimated.txt} node (s300){};
-	\addplot table{Y25_400Pa_gamma_decimated.txt}  node (s400){};
-	\addplot table{Y32_550Pa_gamma_decimated.txt}  node (s550){};
-	\addplot table{Y39_1000Pa_gamma_decimated.txt} node (s1000){};
-\end{loglogaxis}
-\begin{scope}[anchor=base, every node/.style={yshift=0.2em}]
-	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
-	\node[red!60!black] at (s300 |- g.outer north) {300};
-	\node[red!80!black] at (s400 |- g.outer north) {400};
-	\node[red] at (s550 |- g.outer north) {550};
-	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
-\end{scope}
-\end{tikzpicture}
-
-Failure at $\gamma\approx 1$ for a well defined time $\tau_f$
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{Failure time}
-\begin{columns}
-\column{0.55\textwidth}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xlabel={$\sigma$ (\si{\pascal})},
-	ylabel={$\tau_f$ (\si{\second})},
-	xmin=100, ymin=10, ymax=1e6,
-	xtick={100, 200,..., 1000}, xticklabels={100, 200,,, 500,,,,, 1000},
-	]
-	\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
-	\addplot[no marks, black, domain={100:1000}] {4.2e17*x^(-5.45)} node [midway, above right] {$\tau_f\sim \sigma^{-5.5}$};
-\end{loglogaxis}
-\end{tikzpicture}
-
-\begin{block}{Basquin law}
-\begin{itemize}
-\item fatigue (oscillatory stress)
-\item heterogeneous solids
-\item no yield stress
-\end{itemize}
-\end{block}
-
-\column{0.45\textwidth}
-\begin{tikzpicture}
-\begin{groupplot}[
-	group style={
-		group name=g, group size=1 by 2,
-		},
-	width=\textwidth,
-	height=0.7\textwidth,
-	ymode=log,
-	ylabel={$\tau_f$ (\si{\second})},
-	ymin=10, ymax=1e6,
-	]
-	\nextgroupplot[
-		xlabel={$\sigma$ (\si{\pascal})},
-		xtick={200,600,1000},
-		]
-		\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
-		\addplot[no marks, black, domain={100:1000}] {6.5e5*exp(-x/84)} node [pos=0.25, above right] {$\tau_f\nsim e^{-\frac{\sigma}{\sigma_0}}$};
-
-	\nextgroupplot[
-		xlabel={$1/\sigma$ (\si{\per\pascal})},
-		]
-		\addplot[only marks, Accent1] table[x expr=1/\thisrowno{0}, y index=3]{creep_cas4_gdl1_MCR.txt};
-		\addplot[no marks, black, domain={1e-3:7e-3}] {20.3*exp(1720*x)} node [pos=0.25, below right] {$\tau_f\nsim e^\frac{\sigma_0}{\sigma}$};
-\end{groupplot}
-\end{tikzpicture}
-
-Activated processes unlikely (Eyring, Griffith, Taylor, \ldots)
-\end{columns}
-\end{frame}
-\tikzset{external/force remake=false}
-
-%\tikzset{external/force remake}
-\begin{frame}{Primary creep}
-\begin{columns}
-\column{0.55\textwidth}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
-	ylabel={Strain rate $\dot{\gamma}$ (\si{\per\second})},
-	cycle list name=earthy,
-	no marks,
-	]
-	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node[left] (s200){200};
-	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){300};
-	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){400};
-	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){550};
-	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){1000};
-	\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[pos=0.75, below left] {$\dot{\gamma}(t)\sim t^{-0.85}$};
-\end{loglogaxis}
-\end{tikzpicture}
-
-\begin{block}{Primary creep $\Rightarrow$ ``Andrade'' creep}
-Power-law creep known in \alert{solids} for over a century
-
-Classically explained by disinclination dynamics in crystals
-\end{block}
-
-\column{0.45\textwidth}
-\includegraphics[width=\textwidth]{Andrade_1910}
-
-\bigskip
-\includegraphics[height=7\baselineskip]{Andrade_1910_fig1}
-\begin{tikzpicture}
-\node[inner sep=0] (a){\includegraphics[height=7\baselineskip]{Andrade_1910_fig2}};
-\node[below right=0 and 1em of a.north west, font=\footnotesize,fill=white] {lead \& copper};
-\node[above left=1em of a.south east, Accent2] {$\gamma-\gamma_0 \sim t^{1/3}$};
-\end{tikzpicture}
-\end{columns}
-\end{frame}
-\tikzset{external/force remake=false}
-
-%\tikzset{external/force remake}
-\begin{frame}{Primary creep}
-\begin{block}{Local insight from \movie[externalviewer]{ultrasounds}{Yaourt110_primary.avi}}
-\begin{itemize}
-	\item homogeneous \movie[externalviewer]{strain field}{Yaourt110_profils.avi} (no wall slip, no shear band)
-	\item if present, plastic events are below our resolution (a few \si{\micro\metre})
-\end{itemize}
-\end{block}
-
-\begin{block}{Partial creep and recovery}
-\begin{tikzpicture}
-\begin{axis}[
-	width=\textwidth,
-	height=8\baselineskip,
-	xlabel={time (\si{\second})},
-	ylabel={strain $\gamma$},
-	xmin=-1e3, xmax=2.8e4,
-	axis background/.style={fill=white},
-	]
-	\addplot[black, no marks] table {Y261_oscil_creep_recovery_100Pa_decimated.txt};
-	\begin{scope}[<-,Main, inner sep=0]
-	\draw (axis cs:0,0) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow \SI{100}{\pascal}$};
-	\draw (axis cs:600,0.2727007) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow 0$};
-	\draw (axis cs:8600,0.0159) -- +(0,1em) node[anchor=base east]{$\gamma\leftarrow 0$};
-	\end{scope}
-\end{axis}
-\end{tikzpicture}
-\end{block}
-
-$\Rightarrow$ Primary creep is ``kinetically'' reversible (no detectable structure damage, but viscous dissipation)
-\end{frame}
-\tikzset{external/force remake=false}
-
-%\tikzset{external/force remake}
-\begin{frame}{Dimensionless creep response}
-\begin{columns}
-\column{0.55\textwidth}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
-	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
-	cycle list name=earthy,
-	no marks,
-	]
-	\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
-	\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
-	\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
-	\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
-	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.2*x^(-0.85)} node[pos=0.75, below left] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
-\end{loglogaxis}
-\end{tikzpicture}
-
-Linear viscoelasticity accounts for the Andrade exponent
-
-\column{0.45\textwidth}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	height=0.8\textwidth,
-	width=\textwidth,
-	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{$G^\prime$}, \textcolor{Accent2}{$G^{\prime\prime}$}(\si{\pascal})},
-	domain={6e-2:70},
-	ymin=10,
-	]
-	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Main, no marks]{100*x^0.15}  node[midway, below=1em] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
-
-\end{loglogaxis}
-\end{tikzpicture}
-\textcolor{Main}{
-\begin{align*}
-\Rightarrow & J(t) \equiv& \frac{\gamma(t)}{\sigma}\sim& t^{0.15}\\
-\Rightarrow & \dot{\gamma}(t)\sim& t^{0.15-1} =& t^{-0.85}
-\end{align*}
-}
-\end{columns}
-\end{frame}
-\tikzset{external/force remake=false}
-
-\tikzset{external/force remake}
-\begin{frame}{Tertiary creep}
-\begin{columns}
-\column{0.55\textwidth}
-Time to failure
-
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=1e-5, xmax=2, xlabel={$(\tau_f-t)/\tau_f$},
-	x dir=reverse,
-	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
-	cycle list name=earthy,
-	no marks,
-	]
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[Accent2, ultra thick, domain={1e-4:0.1}] {0.05/x} node[pos=0.6, below right, inner sep=0] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{\tau_f}{\tau_f-t}\right)$};
-\end{loglogaxis}
-\end{tikzpicture}
-
-Close to failure: 
-\begin{itemize}
-\item finite time singularity
-\item logarithmic divergence
-\end{itemize}
-$\Rightarrow$ dominated by fracture growth
-
-\column{0.45\textwidth}
-\begin{block}{Fracture growth}
-\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-15-00.jpg}\hfill
-\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-20-00.jpg}\hfill
-\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-21-15.jpg}
-
-\begin{tikzpicture}
-\begin{axis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=1e-3, xmax=1, xlabel={$(\tau_f-t)/\tau_f$},
-	xmode=log, x dir=reverse,
-	ylabel={$\ell/H$},
-	ymin=0, ymax=0.5,
-	axis background/.style={fill=white},
-	]
-	\addplot[only marks, Accent1, error bars/.cd, y dir=both, y explicit] table[y error index=2]{Y110_fracturelength.txt};
-	\addplot[Accent2, ultra thick, no marks, domain={1e-3:1}] {- 0.06 * ln(x)} node [pos=0.25,left=1em] {$\frac{\ell}{H}\sim \ln\frac{\tau_f-t}{\tau_f}$};
-\end{axis}
-\end{tikzpicture}
-\end{block}
-
-\end{columns}
-\end{frame}
-\tikzset{external/force remake=false}
+\include{creep_beamer}
 

### 2014-06-17 seminaire_Tanakalab.tex
People

@@
 	}
+
+\includeonly{}
 
@@
 
+\begin{frame}{Wrinkling of a confined porous layer}
+
+\bigskip
+\structure{Fluorescent yoghurt makers}
+
+\bigskip
+\begin{tabu}{X[c]X[c]c}
+\includegraphics[height=0.3\textheight,clip=true,trim=3cm 7cm  2cm 5cm]{Mathieu}&
+\includegraphics[height=0.3\textheight]{Thomas}&
+\includegraphics[height=0.3\textheight]{Seb}\\
+Mathieu Nespoulous & Thomas Gibaud & Sebastien Manneville\\
+Associate Professor & CNRS researcher & Professor\\
+Aix-Marseille Université & E.N.S. Lyon & E.N.S. Lyon\\
+\end{tabu}
+\end{frame}
+
 %\tikzset{external/force remake}

### 2014-06-17 seminaire_Tanakalab.tex
Cell, patterns, macroscope dynamics

@@
 	}
+
+\newlength{\mylength}
 
@@
 \end{tabu}
+
+\bigskip
+\structure{Imaging plateform} BioSciences Gerland - Lyon Sud (UMS~3444)
 \end{frame}
@@
 
+\begin{frame}{Sealed cell, spontaneous pattern}
+
+	\structure{Side view of the cell}\\
+	\tikzsetnextfilename{cell_brushes}
+	\tikzset{external/force remake}
+	\begin{tikzpicture}[font=\footnotesize]
+		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\textwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
+		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\textwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
+		\draw[line width=2pt,Accent1] (0.05\textwidth,-2.5em) -- (0.95\textwidth,-2.5em) (0.05\textwidth,-1pt) -- (0.95\textwidth,-1pt) node[below,pos=0.30, text width=0.5\textwidth] {acrylamide brush ($10\sim 100$ nm)\linebreak $\Rightarrow$ No adhesion};
+		\fill[gray] (0,0) rectangle (0.05\textwidth,-2.5em) (\textwidth,0) rectangle (0.95\textwidth,-2.5em) node[pos=1, above left] {spacer};
+		\draw[<->] (0.75\textwidth,-2pt) -- (0.75\textwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
+		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] (L){$L\sim 2.5\,cm$};
+		%\node[anchor=north west, inner sep=0] at ($(L.north) - (0.5\textwidth,0)$) {\structure{Side view of the cell}};
+	\end{tikzpicture}
+
+	\bigskip
+	\structure{Top view of the final pattern}\\
+	\begin{tikzpicture}[inner sep=0, very thick]
+	\setlength{\mylength}{\columnwidth}
+	\node[anchor=north west] (a) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2_crop}};
+		\node[anchor=north] at (0.44\mylength,0) (b) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6_crop}};
+		\node[anchor=north east] at (\mylength,0) (c) {\includegraphics[width=0.4\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
+		%zooms
+		\node[minimum width = 0.137\mylength, minimum height=0.137\mylength, anchor=north west, draw=Accent2] at ($(a.north west) +(0.111\mylength,-0.072\mylength)$) (bz){};
+		\draw[Accent2] (bz.north east) -- (b.north west) (bz.south east) -- (b.south west);
+		\node[minimum width = 0.128\mylength, minimum height=0.081\mylength, anchor=north west, draw=Main] at ($(b.north west) +(0.097\mylength,-0.143\mylength)$) (cz){};
+		\draw[Main] (cz.north east) -- (c.north west) (cz.south east) -- (c.south west);
+		\node[minimum width = 0.156\mylength, minimum height=0.156\mylength, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\mylength,0)$) (dz) {};
+		%scale bars
+		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.178\mylength,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
+		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.176\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
+		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
+	\end{tikzpicture}
+\end{frame}
+
+\begin{frame}{Dynamics (transmission macroscope)}
+\movie[externalviewer]{\begin{tikzpicture}
+\node[inner sep=0] (a) {\includegraphics[width=\textwidth]{prise_0799_resized.jpg}};
+\draw[line width=0.2em] ++(a.north west) -- ++(0.177\textwidth,0) node[pos=0, above right, inner xsep=0] {\SI{1}{\milli\metre}};
+\end{tikzpicture}
+}{cas4_GDL_4_100um_coat_macroscope_2.avi}
+\end{frame}
+
+\begin{frame}{Dynamics (transmission macroscope)}
+\begin{tikzpicture}
+	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em, ampersand replacement=\&] (m){
+	33 min \& 38 min \& 43 min \\
+	\includegraphics[width=0.32\textwidth]{prise_0100_color.jpg}\&
+	\includegraphics[width=0.32\textwidth]{prise_0130_color.jpg}\&
+	\includegraphics[width=0.32\textwidth]{prise_0160_color.jpg}\\
+	48 min \& 1h \& 1h15 \\
+	\includegraphics[width=0.32\textwidth]{prise_0190_color.jpg}\&
+	\includegraphics[width=0.32\textwidth]{prise_0250_color.jpg}\&
+	\includegraphics[width=0.32\textwidth]{prise_0360_color.jpg}\\
+	};
+	\draw[line width=0.2em] ++(m-4-1.south west) -- ++(0.056\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
+\end{tikzpicture}
+
+\begin{itemize}
+\item nesting patterns
+\end{itemize}
+\end{frame}
+
 \end{document}

### 2014-06-17 seminaire_Tanakalab.tex
confocal, cascade buckling

@@
 \item nesting patterns
+\item length scale becomes smaller
 \end{itemize}
@@
 
+\begin{frame}{What is going on in the thickness?}
+\begin{columns}
+\column{0.4\textwidth}
+\begin{itemize}
+\item fluorescent casein
+\item confocal microscope
+\item[$\Rightarrow$] 3D concentration field
+\end{itemize}
+
+\begin{tikzpicture}
+\draw[ultra thick] (0,) -- +(0.786\textwidth,0) node[midway, above] {\SI{1}{\milli\metre}};
+\end{tikzpicture}
+\setlength{\tabcolsep}{1pt}
+\begin{tabular}{p{\columnwidth}l}
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}& \SI{15}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png} & \SI{22}{\minute}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png} & \SI{1}{\hour}~26\onslide<2->{\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png} & \SI{1}{\hour}~30\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png} & \SI{1}{\hour}~31\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png} & \SI{1}{\hour}~35\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png} & \SI{2}{\hour}~25\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png} & \SI{2}{\hour}~40\\}
+\end{tabular}
+
+\begin{itemize}
+\item synæresis and swelling
+\item<2-> cause buckling
+\end{itemize}
+\column{0.1\textwidth}
+
+\column{0.5\textwidth}
+\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=1 by 3,
+			xticklabels at=edge bottom,
+			vertical sep=0,
+			},
+		xmin=0, xmax=170, xtick={0,30,...,150},
+		extra tick style={grid=major},%
+		extra x ticks = {21,68}, extra x tick labels={},%
+		scale only axis,
+		width=\textwidth-4em,
+		height=5\baselineskip,
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-1em}
+		]
+	\nextgroupplot[
+		ylabel={$G^\prime$ (\si{\pascal})},
+		ymin=0,
+		]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60}]{cas4_GDL4_Y271.prise};
+
+	\nextgroupplot[
+		ylabel={Volume (\%)}, ymin=20, ymax=100, 
+		ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em},
+		]
+	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_cloques.txt};
+	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
+	\node[anchor=south west] at (rel axis cs:0,0) {\rotatebox{90}{synæresis}};
+	\node at (axis cs:68,80) {swelling};
+
+	\nextgroupplot[
+		ylabel={Excess area (\%)}, ymin=0, ymax=4, ytick={0,1,2,3},
+		xlabel={time (min)}, 
+		every axis y label/.append style={xshift=-1em},
+		]
+	\only<2->{\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, below left] {all};
+	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{3}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, left] {single blister};
+	\draw[->] (axis cs: 21, 3) -- (axis cs: 68, 3) node[midway, below] {\SI{1}{\hour}};}
+	%\legend{all, single blister};
+	\end{groupplot}
+	\end{tikzpicture}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Dynamics (confocal microscope)}
+\movie[externalviewer]{\begin{tikzpicture}
+	\node[inner sep=0] (a) {\includegraphics[width=\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}};
+	\draw[line width=0.2em] ++(a.south west) -- ++(0.75\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
+
+	\node[above=1em of a, inner sep=0] (c) {\includegraphics[width=0.4\textwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
+
+	\node[minimum width = 0.156\textwidth, minimum height=0.156\textwidth, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\textwidth,0)$) (dz) {};
+	\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\textwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
+\end{tikzpicture}
+}{volume_plis.avi}
+\end{frame}
+
+\begin{frame}{Dynamics (confocal microscope)}
+\begin{tikzpicture}
+	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em, ampersand replacement=\&] (m){
+	33 min \& 38 min \& 43 min \\
+	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}\&
+	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}\&
+	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}\\
+	48 min \& 1h \& 1h15 \\
+	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}\&
+	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}\&
+	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}\\
+	};
+	\draw[line width=0.2em] ++(m-4-1.south west) -- ++(0.24\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
+\end{tikzpicture}
+
+\begin{columns}
+\column{0.32\textwidth}
+\begin{itemize}
+\item confinement
+\item cascade buckling
+\item wavelength division
+\end{itemize}
+\column{0.68\textwidth}
+\includegraphics[width=\columnwidth]{Roman_1999_confined.jpg}
+
+\textit{\footnotesize Roman \& Pocheau, EPL (1999)}
+\end{columns}
+
+\begin{center}
+\alert{Where does the initial wavelength come from?}
+\end{center}
+\end{frame}
+
 \end{document}

### 2014-06-18 seminaire_Tanakalab.tex
Wrinkling mechanism: carpet, Darcy, Biot, Poiseuille

@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
+\usetikzlibrary{arrows,arrows.meta,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
 %\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
@@
 	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60}]{cas4_GDL4_Y271.prise};
+	\node[below left] at (rel axis cs:1,1) {rheometer (stick)};
 
@@
 
+\begin{frame}<1-2>[label=carpet]{Wavelength: resisting stress needed}
+\structure{Simple case} Initial flat contact with the substrate: carpet
+\begin{columns}
+\column{0.5\textwidth}
+\temporal<2>{
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png}\\
+	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png}\\
+
+	\includegraphics[width=\textwidth]{volume_cloque_t116}\\
+	\begin{tikzpicture}
+	\draw[ultra thick] (0,) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
+	\end{tikzpicture}
+
+	\structure{In our case}
+	\begin{itemize}
+	\item Vertical cells show the same wavelength
+	\item $\sigma \neq \rho g h$
+	\end{itemize}
+}{
+
+\bigskip
+\begin{itemize}
+\item solvent must flow through gel
+\begin{description}[$\ell_p$]
+\item[$v$] gel velocity = -volumic flux $\simeq \SI{0.2}{\micro\metre\per\second}$
+\item[$\ell_p$] pore size $ \approx \SI{4}{\micro\metre}$
+\end{description}
+\item $\sigma$ is a Darcy pressure
+\[\sigma = \Delta P \sim \eta h v/\ell_p^2\]
+\item poroelastic length
+\[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}\]
+\end{itemize}
+}{
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=0.6\textwidth,
+	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
+	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
+	cycle list name=earthy,
+	no marks,
+	]
+	\draw[ultra thick, Accent2] (axis cs:580,0) -- (axis cs:580,50);
+	\foreach \x in {2,3,..., 8}
+		\addplot table[y index=\x]{alts_bottom_cloque.txt};
+\end{axis}
+\end{tikzpicture}
+\begin{itemize}
+\item $\lambda$ frozen at ceiling touch 
+\item $A=H\equiv e-h$
+\item $\lambda \sim \epsilon^{-3/7} H \sim e$
+	%\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq 0.6\% \]
+\end{itemize}
+}
+
+\column{0.5\textwidth}
+\begin{block}{Ruck in a rug}
+\begin{tikzpicture}[ultra thick]
+\begin{axis}[
+	name=a,
+	width=\textwidth, height=3\baselineskip, scale only axis,
+	domain=-180:180, no markers, ymin=0, ymax=3,xmin=-180,xmax=180,
+	axis lines=none, xtick=\empty,
+	axis background/.style={fill=white},
+	]
+	\addplot+[Accent1]{cos(x)+1};
+	\addplot+[Accent1]{cos(x)+1.5};
+	\draw[->, Accent2, ultra thick] (axis cs:90,1.25) -- (axis cs:90,2.25) node[pos=1, right] {$v$};
+	\draw[Main, ultra thick, {Circle}->] (axis cs:-90,1.5) -- (axis cs:-90,0.25) node[right] {$\sigma$};
+	%\node[above] at (axis cs:-90,1.5) {$P_2$};
+	%\node[below right] at (axis cs:-90,1.25) {$P_1$};
+	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
+	\draw[<->] (axis cs:0,2) -- (axis cs:0,2.5) node[midway, right] {$h$};
+\end{axis}
+\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle ($(a.south east)+(0,-1em)$);
+\draw[<->] ($(a.south west)+(0,-1.1em)$) -- ($(a.south east)+(0,-1.1em)$) node[midway, below] {$\lambda$};
+\end{tikzpicture}
+
+\textit{\footnotesize Kolinski et al., Vella et al. PRL 2009}\\
+\begin{description}[$B$]
+\item[$\epsilon$] excess area \hfill$\Rightarrow$ buckling 
+\item[$B$] bending modulus \hfill$\Rightarrow \lambda\nearrow$
+\item[$\sigma$] vertical stress \hfill$\Rightarrow \lambda\searrow$
+\end{description}
+\[ \lambda^* \equiv \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]
+$\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
+\end{block}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Poroelasticity}
+\begin{columns}\column{0.15\textwidth}
+\includegraphics[width=\textwidth]{Henry_Darcy}
+\column{0.85\textwidth}
+\begin{block}{Henry Darcy (1803-1858)}
+\begin{description}[1803]
+\item[1846] granted lifelong free water
+\item[1856] \textit{Les fontaines publiques de la ville de Dijon}
+\end{description}
+\end{block}
+\end{columns}
+
+\begin{columns}
+\column{0.15\textwidth}
+\includegraphics[width=\textwidth]{Maurice_Anthony_Biot}
+\column{0.58\textwidth}
+\begin{block}{Maurice Anthony Biot (1905-1985)}
+\begin{description}[1803]
+%\item[1932] Ph.D. under von Kármán (Cal Tech)
+\item[1932-1942] earthquake engineering
+\item[1935-1962] theory of poroelasticity
+\end{description}
+\end{block}
+\column{0.27\textwidth}
+\includegraphics[width=\textwidth]{Biot_1964_Buckling_Porous}
+\end{columns}
+
+\medskip
+\begin{footnotesize}
+\textit{Biot, J. Appl. Mech. 1964}: Buckling of a porous slab embedded in an impervious infinite viscous or viscoelastic medium
+\end{footnotesize}
+\begin{itemize}
+\item $B$ transiently higher due to pore pressure
+\item $\sigma$ comes only from the surrounding medium $\rightarrow 0$
+\item effect of flow through the porous slab not considered
+\end{itemize}
+\end{frame}
+
+\begin{frame}{No initial contact with a substrate}
+\structure{Poiseuille vs Darcy on a flying carpet}
+\begin{tikzpicture}
+\begin{axis}[
+	name=a,
+	width=\textwidth, height=0.25\textwidth, scale only axis,
+	domain=0:360, no markers, ymin=-4, ymax=4,xmin=0,xmax=360,
+	axis lines=none, xtick=\empty,
+	]
+	\fill[gray!20] (axis cs:85,4) rectangle (axis cs:95,-4) (axis cs:265,4) rectangle (axis cs:275,-4);
+	\addplot+[Accent1]{sin(x)+0.5};
+	\addplot+[Accent1]{sin(x)-0.5};
+	\addplot+[dashed, black]{0.5};
+	\addplot+[dashed, black]{-0.5};
+	\draw[<->, help lines] (axis cs:180,0.5) -- (axis cs:180,4) node[midway, left] {$H$}; 
+	\draw[<->, help lines] (axis cs:30,-0.5) -- (axis cs:30,-4) node[midway, left] {$H$};
+	\node[below] at (axis cs:90,-1.5) (pb1) {$P_1$};
+	\node[above] at (axis cs:90,1.5) (ph2) {$P_2$};
+	\node[above] at (axis cs:270,1.5) (ph1) {$P_1$};
+	\node[below] at (axis cs:270,-1.5) (pb2) {$P_2$};
+	\draw[line width=0.1em, ->] (ph2) -- (axis cs:90,0.5) node[midway, left] {Darcy} node[midway, right] {$v$};
+	\draw[line width=0.1em, ->] (pb2) -- (pb1) node[midway, above] {Poiseuille} node[midway, below] {$u \sim \frac{\lambda}{H} v$};
+\end{axis}
+\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle +(\textwidth,-1em) (a.north west) rectangle +(\textwidth,1em);
+\end{tikzpicture}
+
+\begin{columns}
+\column{0.83\textwidth}
+\begin{itemize}
+\item The same $\Delta P$ can be relaxed by a Poiseuille flow
+\[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
+\item For the same wavelength
+\hfill$\displaystyle \frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$
+\item Flow through the pores dominates
+\end{itemize}
+\column{0.17\textwidth}
+\includegraphics[width=\textwidth]{Jean-Leonard-Marie_Poiseuille.jpg}\\
+\hfill\structure{1797-1869}
+\end{columns}
+\end{frame}
+
+\againframe<3>{carpet}
+
 \end{document}

### 2014-06-18 seminaire_Tanakalab.tex
Cell thickness and debuging of a ill placed force remake

@@
 	\tikzsetnextfilename{cell_brushes}
-	\tikzset{external/force remake}
 	\begin{tikzpicture}[font=\footnotesize]
@@
 \begin{block}{Ruck in a rug}
+\tikzsetnextfilename{ruck}%
 \begin{tikzpicture}[ultra thick]
@@
 \againframe<3>{carpet}
+\tikzset{external/force remake}
+\begin{frame}{Cell thickness}
+\begin{tikzpicture}[mark size={0.3em}]
+	\matrix[matrix of nodes, inner sep=0, column sep=0.01\columnwidth, row sep=0.01\columnwidth, ampersand replacement=\&] (m) {
+		\includegraphics[width=0.245\columnwidth]{pattern_50um.jpg}\&
+		\includegraphics[width=0.245\columnwidth]{pattern_100um.jpg}\&
+		\includegraphics[width=0.245\columnwidth]{pattern_250um.jpg}\&
+		\includegraphics[width=0.245\columnwidth]{pattern_450um.jpg}\\
+	};
+	\node[inner sep=0, below left=0.01\columnwidth and 0 of m.south east] (x800) {\includegraphics[width=0.49\columnwidth]{pattern_800um_x1.jpg}};
+	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.196\columnwidth, 0.05\columnwidth) -- ++(0.145\columnwidth,0) node[midway,above]{$\lambda$};
+	\draw[line width=0.2em,Main](x800.north east) -- ++(-0.217\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};
+
+	\begin{axis}[%
+		name=g,
+		anchor=left of north west,
+		at={($(m.south west)+(0,-0.01\columnwidth)$)},
+		scale only axis,
+		width=0.5\textwidth-3em,
+		height=0.25\columnwidth-2em,
+		xmin=0,xmax=1,xlabel={$e$ (\si{\milli\metre})},
+		xtick={0,0.4,...,0.9},
+		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
+		]
+		\addplot[no marks, domain=0:3] {4.23*x};
+		\begin{scope}[inner sep=0, minimum size=0.8em]
+			\node[circle, fill=Accent1] at (axis cs:0.05,0.261){};
+			\node[circle, fill=white, draw=Accent1] at (axis cs:0.1,0.458) {};
+			\node[regular polygon, regular polygon sides=3, fill=Accent1] at (axis cs:0.250,0.916){};
+			\node[regular polygon, regular polygon sides=3, fill=white, draw=Accent1] at (axis cs:0.450,2.075){};
+			\node[fill=Accent1] at (axis cs:0.800,3.333){};
+		\end{scope}
+	\end{axis}
+	\begin{scope}[inner sep=0, minimum size=0.8em, transform canvas={xshift=0.5em, yshift=-0.5em}]
+		\node[circle, fill=Accent1] at (m-1-1.north west) {};
+		\node[circle, fill=white, draw=Accent1] at (m-1-2.north west){};
+		\node[regular polygon, regular polygon sides=3, fill=Accent1] at (m-1-3.north west){};
+		\node[regular polygon, regular polygon sides=3, fill=white, draw=Accent1] at (m-1-4.north west) {};
+		\node[fill=Accent1] at (x800.north west) {};
+	\end{scope}
+\end{tikzpicture}
+\end{frame}
+\tikzset{external/force remake=false}
 

### 2014-06-18 seminaire_Tanakalab.tex
Viscosity

@@
 \againframe<3>{carpet}
-\tikzset{external/force remake}
+
+%\tikzset{external/force remake}
 \begin{frame}{Cell thickness}
@@
 \end{frame}
-\tikzset{external/force remake=false}
+%\tikzset{external/force remake=false}
+
+\tikzset{external/force remake}
+\begin{frame}{Surprising viscosity influence}
+\begin{columns}
+\column{0.5\textwidth}
+\[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3}\]
+\begin{tikzpicture}
+	\begin{loglogaxis}[%
+		width=\textwidth,
+		height=0.8\textwidth,
+		xmin=0.6,xmax=11,xlabel={viscosity (\si{\milli\pascal\second})},
+		xtick={0.7, 1, 2, 5, 10}, xticklabels={$0.7$, $1$, $2$, $5$, $10$},
+		ytick={1,2,3,4}, yticklabels={$1$,$2$,$3$,$4$},
+		ymin=1, ymax=5, ylabel={$\lambda$ (\si{\milli\metre})},
+		mark size={0.3em},
+		legend style={legend pos=south east},
+		]
+		\addplot+[no marks, domain=0.7:10, black, forget plot] {1.7*x^0.4} node[midway, below] {$\eta^{0.4}$};
+		\addplot+[only marks,Accent1, mark=o] table{eta_lambda_eau_pure.txt};
+		\addplot+[only marks,Accent2, mark=square] table{eta_lambda_eau_glycerol.txt};
+		\legend{water, water-glycerol};
+	\end{loglogaxis}
+\end{tikzpicture}
+
+\column{0.5\textwidth}
+\begin{itemize}
+\item $\lambda$ increases with $\eta$
+\item Can be explained only by a hidden dependence of the viscosity on $B$, $\ell_p$, \ldots 
+\item[$\Rightarrow$] Viscosity influences gel formation.
+\end{itemize}
+%If $\ell_p \sim \eta^{-\alpha}$ and $B\sim \mathrm{k_B}T \left(\frac{h}{\ell}\right)^3$, then
+%\[ \lambda^* \sim \eta^\frac{\alpha-1}{3} \]
+%\[\alpha\simeq 2.2\]
+\end{columns}
+\end{frame}
+
+\begin{frame}{Viscosity of the solvent during gel formation}
+\begin{tabu}{X[c]X[c]}
+water $\eta=\SI{1}{\milli\pascal\second}$ &
+glycerol 60\% $\eta=\SI{10}{\milli\pascal\second}$\\
+\includegraphics[width=0.45\textwidth]{XYslice_gly0}&
+\includegraphics[width=0.45\textwidth]{XYslice_gly60}
+\end{tabu}
+\tikz\draw[line width=0.2em] (0,0) -- ++(0.045\textwidth,0) node[midway, below] {\SI{10}{\micro\metre}};
+\end{frame}
 

### 2014-06-18 plis_beamer.tex
split the wrinkling part of the seminar

--- 2014-06-18 seminaire_Tanakalab.tex
@@
-\documentclass[xcolor=table]{beamer}
-\usepackage[utf8]{inputenc}
-\usepackage[francais]{babel}
-%\usetheme{Boadilla}
-%\usecolortheme{rose}
-%\usecolortheme{crane}
-\usefonttheme{structuresmallcapsserif}
-\setbeamertemplate{navigation symbols}{}
-
-\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
-\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
-\definecolor{Accent2}{rgb}{0.54,0.1,0.4}
-
-\usecolortheme{rose}
-%\useinnertheme[shadow]{circles}
-\usecolortheme{whale}
-%\useoutertheme{infolines}
-
-\usecolortheme[named=Accent1]{structure}
-
-\setbeamercolor{alerted text}{fg=Accent2}
-%\setbeamercolor{palette primary}{fg=white}
-%\setbeamercolor{palette secondary}{bg=Accent1}
-%\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}
-
-\setbeamerfont{page number in head/foot}{size=\large}
-\setbeamercolor{page number in head/foot}{fg=Main}
-% page/total
-%\setbeamertemplate{footline}[frame number]
-% pas de total
-\setbeamertemplate{footline}{%
-    	\hfill%
-	\usebeamercolor[fg]{page number in head/foot}%
-	\usebeamerfont{page number in head/foot}%
-	\insertframenumber\kern1em\vskip2pt%
-}
-
-\setbeamersize{text margin left=1em}
-\setbeamersize{text margin right=1em}
-
-%font
-\usepackage[T1]{fontenc}
-\usepackage[oldstylenums]{kpfonts}
-
-%proper math and math symbols
-%\usepackage{amsmath}
-\usepackage{amssymb}
-
-\usepackage{datenumber,fp}
-
-\usepackage{siunitx}
-
-\usepackage{tabu}
-\usepackage{multirow}
-\usepackage{booktabs}
-
-% Allow the usage of graphics (.jpg, .png, etc.) in the document
-\usepackage{graphicx}
-\usepackage{tikz}
-\usetikzlibrary{arrows,arrows.meta,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
-%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
-
-\usepackage{pgfplots}
-\pgfplotsset{compat=1.9}
-\usepgfplotslibrary{groupplots}
-\usepgfplotslibrary{external}
-\tikzexternalize
-\tikzsetexternalprefix{fig_presentation/}
-%\tikzset{external/optimize=false}
-%\tikzset{external/force remake}
-
-%link or play movies
-\usepackage{multimedia}
-
-%beamer related package
-
-\usepackage{todonotes}
-\presetkeys{todonotes}{inline}{}
-
-%bibliography
-\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
-%minimal citations
-\AtEveryCitekey{%
-	\clearfield{title}
-	\clearfield{pages}
-	\clearfield{volume}
-	\clearfield{number}
-	\clearfield{month}}
-\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{%
-  \printtext{\bibstring{in}\intitlepunct}}}
-%\bibliography{biblio}
-
-\newcolumntype{P}[1]{>{\raggedright}p{#1}}
-
-\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Supérieure de Lyon}
-\title[Yoghurt under stress]{Yoghurt under stress}
-\author[M. Leocmach]{Mathieu Leocmach}
-\date{1st June 2014}
-\titlegraphic{
-	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
-	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
-	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
-	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
-	}
-
-\newlength{\mylength}
-
-\includeonly{}
-
-\begin{document}
-\tikzset{every mark/.append style={scale=0.8}}
-\pgfplotsset{every axis/.append style={footnotesize}}
-
-\pgfplotscreateplotcyclelist{earthy}{%
-{red!40!black,mark=o},
-{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
-{red!80!black,mark=square},
-{red,mark=triangle},
-{red!80!yellow, mark=diamond},
-red!60!yellow,
-red!40!yellow,
-}
-
-\AtBeginSection[]{
-	\addtocounter{framenumber}{-1}
-	\begin{frame}[plain]
-		\tableofcontents[currentsection, hideothersubsections]
-	\end{frame}
-}
-
-\begin{frame}[plain]
-	\titlepage
-\end{frame}
-
-\setcounter{framenumber}{0}
-
-\begin{frame}{Application of biogels}
-\begin{tabu}{cX[c]c}
-\includegraphics[height=0.3\textheight]{cell_mech} &
-\includegraphics[height=0.3\textheight]{electrophoresis} &
-\includegraphics[height=0.3\textheight]{cosmetics} \\
-Cell mechanics & Electrophoresis & Cosmetics\\
-\end{tabu}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=0.3\textheight]{bacterial_culture} &
-\includegraphics[height=0.3\textheight]{food} \\
-Bacterial culture & Food\\
-\end{tabu}
-
-\begin{footnotesize}
-\begin{block}{Sources}
-\begin{tabu}{lXl}
-Wikimedia Commons & www.madaboutscience.com & www.keautystore.com\\
-\end{tabu}
-\end{block}
-\end{footnotesize}
-\end{frame}
-
-\begin{frame}{Fundamental issues in biogels}
-\begin{itemize}
-\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
-\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
-\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
-\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
-\end{itemize}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=6\baselineskip]{Bonn_fracture} &
-\includegraphics[height=6\baselineskip]{Villi_sq}
-\end{tabu}
-\begin{block}{Our questions}
-\begin{itemize}
-\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
-\item What is the role of the solvent in mechanical instabilities?
-\end{itemize}
-\end{block}
-\end{frame}
-
-%\section{Our system: Yoghurt}
-
-%\tikzset{external/force remake}
-\begin{frame}{Acid-set protein gel (Yoghurt)}
-\begin{block}{Recipe: take your time}
-\begin{itemize}
-\item Water (\SI{30}{\celsius})
-\item Sodium caseinate (milk protein) 4\%
-\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
-\item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow homogeneous acidification
-\end{itemize}
-\end{block}
-
-\begin{columns}
-\column{0.5\textwidth}
-\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		%xmode=log,
-		%xmin=1e2,xmax=3e5,
-		xmin=0, xmax=20,
-		scale only axis,
-		width=\textwidth-4em,
-		height=0.3\textwidth,
-		extra tick style={grid=major},%
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
-		]
-	\nextgroupplot[
-		ymin=0, ymax=7, ylabel={pH},
-		extra y ticks={4.6}, extra y tick labels={},%
-		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y190_cas4_gdl1.pH};
-	\draw[help lines] (axis cs:0,4.6) -- (axis cs:3.88,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:3.88,0);
-
-	\nextgroupplot[
-		xlabel={time (\si{\hour})},
-		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
-		extra x ticks={17}, extra x tick labels={},
-		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y22.prise};
-	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL1_Y22.prise};
-	\begin{scope}[font=\scriptsize]
-		%\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
-		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
-		\node[anchor=base,rotate=80] at (axis cs:2.5,400) {gelation};
-	\end{scope}
-	\end{groupplot}
-	\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
-\end{tikzpicture}
-
-\column{0.5\textwidth}
-\includegraphics[width=\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
-
-\begin{scriptsize}
-\textit{Kaláb (1983),\linebreak Roefs \& van Vliet (1990),\linebreak Lucey \& Singh (1998)}
-\end{scriptsize}
-\end{columns}
-\end{frame}
-\tikzset{external/force remake=false}
-
-\begin{frame}{Linear rheology: Power law (visco)elastic solid}
-\[\text{stress}\rightarrow \sigma = G \gamma\leftarrow\text{strain}\]
-\[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{loss} moduli (\si{\pascal})},
-	domain={6e-2:70},
-	]
-	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot{590*x^0.14};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot{350*x^0.12};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot{246*x^0.1};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot{163*x^0.067};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot{109*x^0.044};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot{76*x^0.035};
-
-\end{loglogaxis}
-\end{tikzpicture}
-\end{frame}
-
-\section*{Creep and yielding}
-
-\include{creep_beamer}
-
-\section*{Wrinkling}
-
 \begin{frame}{Wrinkling of a confined porous layer}
@@
 \begin{tabu}{X[c]X[c]}
-water $\eta=\SI{1}{\milli\pascal\second}$ &
-glycerol 60\% $\eta=\SI{10}{\milli\pascal\second}$\\
+water, $\eta=\SI{1}{\milli\pascal\second}$ &
+glycerol 60\%, $\eta=\SI{10}{\milli\pascal\second}$\\
 \includegraphics[width=0.45\textwidth]{XYslice_gly0}&
@@
 
-\end{document}
+\begin{frame}{Conclusion}
+Spontaneous wrinkling of over-acidified casein gels in non-adhesive confinement
+
+	\structure{Original patterns}
+	\begin{itemize}
+		\item nested (optical elements ?)
+		\item kinetic wavelength selection
+	\end{itemize}
+	\structure{A new wrinkling mechanism}
+	\begin{itemize}
+		\item porous flow resist bending
+		\item Darcy can beat Poiseuille
+		\item may extend to other porous sheets (wet paper, biological membranes\ldots)
+	\end{itemize}
+	\structure{Viscosity influences}
+	\begin{itemize}
+		\item porous dissipation
+		\item gel formation (theory?)
+		\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ thus wrinkling}
+	\end{itemize}
+	\structure{To do:} Measure everything to plot $\lambda_\text{theory}$ vs $\lambda_\text{experiment}$
+\end{frame}

### 2014-06-18 seminaire_Tanakalab.tex
Conclusion
split the wrinkling part of the seminar

--- 2014-06-18 seminaire_Tanakalab.tex
@@
 
-\includeonly{}
+%\includeonly{}
 
@@
 
-\begin{frame}{Wrinkling of a confined porous layer}
-
-\bigskip
-\structure{Fluorescent yoghurt makers}
-
-\bigskip
-\begin{tabu}{X[c]X[c]c}
-\includegraphics[height=0.3\textheight,clip=true,trim=3cm 7cm  2cm 5cm]{Mathieu}&
-\includegraphics[height=0.3\textheight]{Thomas}&
-\includegraphics[height=0.3\textheight]{Seb}\\
-Mathieu Nespoulous & Thomas Gibaud & Sebastien Manneville\\
-Associate Professor & CNRS researcher & Professor\\
-Aix-Marseille Université & E.N.S. Lyon & E.N.S. Lyon\\
-\end{tabu}
-
-\bigskip
-\structure{Imaging plateform} BioSciences Gerland - Lyon Sud (UMS~3444)
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{Over-acidification: Faster but weaker}
-\begin{tikzpicture}
-\begin{axis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
-	cycle list name=earthy,
-	no marks,
-	xmin=0, xmax=20,ymin=0,
-	]
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
-\end{axis}
-\end{tikzpicture}
-
-Caseins regain some solubility at low pH.
-\end{frame}
-
-\begin{frame}{Sealed cell, spontaneous pattern}
-
-	\structure{Side view of the cell}\\
-	\tikzsetnextfilename{cell_brushes}
-	\tikzset{external/force remake}
-	\begin{tikzpicture}[font=\footnotesize]
-		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\textwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
-		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\textwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
-		\draw[line width=2pt,Accent1] (0.05\textwidth,-2.5em) -- (0.95\textwidth,-2.5em) (0.05\textwidth,-1pt) -- (0.95\textwidth,-1pt) node[below,pos=0.30, text width=0.5\textwidth] {acrylamide brush ($10\sim 100$ nm)\linebreak $\Rightarrow$ No adhesion};
-		\fill[gray] (0,0) rectangle (0.05\textwidth,-2.5em) (\textwidth,0) rectangle (0.95\textwidth,-2.5em) node[pos=1, above left] {spacer};
-		\draw[<->] (0.75\textwidth,-2pt) -- (0.75\textwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
-		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] (L){$L\sim 2.5\,cm$};
-		%\node[anchor=north west, inner sep=0] at ($(L.north) - (0.5\textwidth,0)$) {\structure{Side view of the cell}};
-	\end{tikzpicture}
-
-	\bigskip
-	\structure{Top view of the final pattern}\\
-	\begin{tikzpicture}[inner sep=0, very thick]
-	\setlength{\mylength}{\columnwidth}
-	\node[anchor=north west] (a) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2_crop}};
-		\node[anchor=north] at (0.44\mylength,0) (b) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6_crop}};
-		\node[anchor=north east] at (\mylength,0) (c) {\includegraphics[width=0.4\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
-		%zooms
-		\node[minimum width = 0.137\mylength, minimum height=0.137\mylength, anchor=north west, draw=Accent2] at ($(a.north west) +(0.111\mylength,-0.072\mylength)$) (bz){};
-		\draw[Accent2] (bz.north east) -- (b.north west) (bz.south east) -- (b.south west);
-		\node[minimum width = 0.128\mylength, minimum height=0.081\mylength, anchor=north west, draw=Main] at ($(b.north west) +(0.097\mylength,-0.143\mylength)$) (cz){};
-		\draw[Main] (cz.north east) -- (c.north west) (cz.south east) -- (c.south west);
-		\node[minimum width = 0.156\mylength, minimum height=0.156\mylength, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\mylength,0)$) (dz) {};
-		%scale bars
-		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.178\mylength,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
-		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.176\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
-		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
-	\end{tikzpicture}
-\end{frame}
-
-\begin{frame}{Dynamics (transmission macroscope)}
-\movie[externalviewer]{\begin{tikzpicture}
-\node[inner sep=0] (a) {\includegraphics[width=\textwidth]{prise_0799_resized.jpg}};
-\draw[line width=0.2em] ++(a.north west) -- ++(0.177\textwidth,0) node[pos=0, above right, inner xsep=0] {\SI{1}{\milli\metre}};
-\end{tikzpicture}
-}{cas4_GDL_4_100um_coat_macroscope_2.avi}
-\end{frame}
-
-\begin{frame}{Dynamics (transmission macroscope)}
-\begin{tikzpicture}
-	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em, ampersand replacement=\&] (m){
-	33 min \& 38 min \& 43 min \\
-	\includegraphics[width=0.32\textwidth]{prise_0100_color.jpg}\&
-	\includegraphics[width=0.32\textwidth]{prise_0130_color.jpg}\&
-	\includegraphics[width=0.32\textwidth]{prise_0160_color.jpg}\\
-	48 min \& 1h \& 1h15 \\
-	\includegraphics[width=0.32\textwidth]{prise_0190_color.jpg}\&
-	\includegraphics[width=0.32\textwidth]{prise_0250_color.jpg}\&
-	\includegraphics[width=0.32\textwidth]{prise_0360_color.jpg}\\
-	};
-	\draw[line width=0.2em] ++(m-4-1.south west) -- ++(0.056\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
-\end{tikzpicture}
-
-\begin{itemize}
-\item nesting patterns
-\item length scale becomes smaller
-\end{itemize}
-\end{frame}
-
-\begin{frame}{What is going on in the thickness?}
-\begin{columns}
-\column{0.4\textwidth}
-\begin{itemize}
-\item fluorescent casein
-\item confocal microscope
-\item[$\Rightarrow$] 3D concentration field
-\end{itemize}
-
-\begin{tikzpicture}
-\draw[ultra thick] (0,) -- +(0.786\textwidth,0) node[midway, above] {\SI{1}{\milli\metre}};
-\end{tikzpicture}
-\setlength{\tabcolsep}{1pt}
-\begin{tabular}{p{\columnwidth}l}
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}& \SI{15}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png} & \SI{22}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png} & \SI{1}{\hour}~26\onslide<2->{\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png} & \SI{1}{\hour}~30\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png} & \SI{1}{\hour}~31\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png} & \SI{1}{\hour}~35\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png} & \SI{2}{\hour}~25\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png} & \SI{2}{\hour}~40\\}
-\end{tabular}
-
-\begin{itemize}
-\item synæresis and swelling
-\item<2-> cause buckling
-\end{itemize}
-\column{0.1\textwidth}
-
-\column{0.5\textwidth}
-\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 3,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		xmin=0, xmax=170, xtick={0,30,...,150},
-		extra tick style={grid=major},%
-		extra x ticks = {21,68}, extra x tick labels={},%
-		scale only axis,
-		width=\textwidth-4em,
-		height=5\baselineskip,
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-1em}
-		]
-	\nextgroupplot[
-		ylabel={$G^\prime$ (\si{\pascal})},
-		ymin=0,
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/60}]{cas4_GDL4_Y271.prise};
-	\node[below left] at (rel axis cs:1,1) {rheometer (stick)};
-
-	\nextgroupplot[
-		ylabel={Volume (\%)}, ymin=20, ymax=100, 
-		ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em},
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_cloques.txt};
-	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
-	\node[anchor=south west] at (rel axis cs:0,0) {\rotatebox{90}{synæresis}};
-	\node at (axis cs:68,80) {swelling};
-
-	\nextgroupplot[
-		ylabel={Excess area (\%)}, ymin=0, ymax=4, ytick={0,1,2,3},
-		xlabel={time (min)}, 
-		every axis y label/.append style={xshift=-1em},
-		]
-	\only<2->{\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, below left] {all};
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{3}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, left] {single blister};
-	\draw[->] (axis cs: 21, 3) -- (axis cs: 68, 3) node[midway, below] {\SI{1}{\hour}};}
-	%\legend{all, single blister};
-	\end{groupplot}
-	\end{tikzpicture}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Dynamics (confocal microscope)}
-\movie[externalviewer]{\begin{tikzpicture}
-	\node[inner sep=0] (a) {\includegraphics[width=\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}};
-	\draw[line width=0.2em] ++(a.south west) -- ++(0.75\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
-
-	\node[above=1em of a, inner sep=0] (c) {\includegraphics[width=0.4\textwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
-
-	\node[minimum width = 0.156\textwidth, minimum height=0.156\textwidth, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\textwidth,0)$) (dz) {};
-	\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\textwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
-\end{tikzpicture}
-}{volume_plis.avi}
-\end{frame}
-
-\begin{frame}{Dynamics (confocal microscope)}
-\begin{tikzpicture}
-	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em, ampersand replacement=\&] (m){
-	33 min \& 38 min \& 43 min \\
-	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}\&
-	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}\&
-	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}\\
-	48 min \& 1h \& 1h15 \\
-	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}\&
-	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}\&
-	\includegraphics[width=0.32\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}\\
-	};
-	\draw[line width=0.2em] ++(m-4-1.south west) -- ++(0.24\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
-\end{tikzpicture}
-
-\begin{columns}
-\column{0.32\textwidth}
-\begin{itemize}
-\item confinement
-\item cascade buckling
-\item wavelength division
-\end{itemize}
-\column{0.68\textwidth}
-\includegraphics[width=\columnwidth]{Roman_1999_confined.jpg}
-
-\textit{\footnotesize Roman \& Pocheau, EPL (1999)}
-\end{columns}
-
-\begin{center}
-\alert{Where does the initial wavelength come from?}
-\end{center}
-\end{frame}
-
-\begin{frame}<1-2>[label=carpet]{Wavelength: resisting stress needed}
-\structure{Simple case} Initial flat contact with the substrate: carpet
-\begin{columns}
-\column{0.5\textwidth}
-\temporal<2>{
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png}\\
-
-	\includegraphics[width=\textwidth]{volume_cloque_t116}\\
-	\begin{tikzpicture}
-	\draw[ultra thick] (0,) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
-	\end{tikzpicture}
-
-	\structure{In our case}
-	\begin{itemize}
-	\item Vertical cells show the same wavelength
-	\item $\sigma \neq \rho g h$
-	\end{itemize}
-}{
-
-\bigskip
-\begin{itemize}
-\item solvent must flow through gel
-\begin{description}[$\ell_p$]
-\item[$v$] gel velocity = -volumic flux $\simeq \SI{0.2}{\micro\metre\per\second}$
-\item[$\ell_p$] pore size $ \approx \SI{4}{\micro\metre}$
-\end{description}
-\item $\sigma$ is a Darcy pressure
-\[\sigma = \Delta P \sim \eta h v/\ell_p^2\]
-\item poroelastic length
-\[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}\]
-\end{itemize}
-}{
-\begin{tikzpicture}
-\begin{axis}[
-	width=\textwidth,
-	height=0.6\textwidth,
-	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
-	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
-	cycle list name=earthy,
-	no marks,
-	]
-	\draw[ultra thick, Accent2] (axis cs:580,0) -- (axis cs:580,50);
-	\foreach \x in {2,3,..., 8}
-		\addplot table[y index=\x]{alts_bottom_cloque.txt};
-\end{axis}
-\end{tikzpicture}
-\begin{itemize}
-\item $\lambda$ frozen at ceiling touch 
-\item $A=H\equiv e-h$
-\item $\lambda \sim \epsilon^{-3/7} H \sim e$
-	%\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq 0.6\% \]
-\end{itemize}
-}
-
-\column{0.5\textwidth}
-\begin{block}{Ruck in a rug}
-\begin{tikzpicture}[ultra thick]
-\begin{axis}[
-	name=a,
-	width=\textwidth, height=3\baselineskip, scale only axis,
-	domain=-180:180, no markers, ymin=0, ymax=3,xmin=-180,xmax=180,
-	axis lines=none, xtick=\empty,
-	axis background/.style={fill=white},
-	]
-	\addplot+[Accent1]{cos(x)+1};
-	\addplot+[Accent1]{cos(x)+1.5};
-	\draw[->, Accent2, ultra thick] (axis cs:90,1.25) -- (axis cs:90,2.25) node[pos=1, right] {$v$};
-	\draw[Main, ultra thick, {Circle}->] (axis cs:-90,1.5) -- (axis cs:-90,0.25) node[right] {$\sigma$};
-	%\node[above] at (axis cs:-90,1.5) {$P_2$};
-	%\node[below right] at (axis cs:-90,1.25) {$P_1$};
-	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
-	\draw[<->] (axis cs:0,2) -- (axis cs:0,2.5) node[midway, right] {$h$};
-\end{axis}
-\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle ($(a.south east)+(0,-1em)$);
-\draw[<->] ($(a.south west)+(0,-1.1em)$) -- ($(a.south east)+(0,-1.1em)$) node[midway, below] {$\lambda$};
-\end{tikzpicture}
-
-\textit{\footnotesize Kolinski et al., Vella et al. PRL 2009}\\
-\begin{description}[$B$]
-\item[$\epsilon$] excess area \hfill$\Rightarrow$ buckling 
-\item[$B$] bending modulus \hfill$\Rightarrow \lambda\nearrow$
-\item[$\sigma$] vertical stress \hfill$\Rightarrow \lambda\searrow$
-\end{description}
-\[ \lambda^* \equiv \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]
-$\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
-\end{block}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Poroelasticity}
-\begin{columns}\column{0.15\textwidth}
-\includegraphics[width=\textwidth]{Henry_Darcy}
-\column{0.85\textwidth}
-\begin{block}{Henry Darcy (1803-1858)}
-\begin{description}[1803]
-\item[1846] granted lifelong free water
-\item[1856] \textit{Les fontaines publiques de la ville de Dijon}
-\end{description}
-\end{block}
-\end{columns}
-
-\begin{columns}
-\column{0.15\textwidth}
-\includegraphics[width=\textwidth]{Maurice_Anthony_Biot}
-\column{0.58\textwidth}
-\begin{block}{Maurice Anthony Biot (1905-1985)}
-\begin{description}[1803]
-%\item[1932] Ph.D. under von Kármán (Cal Tech)
-\item[1932-1942] earthquake engineering
-\item[1935-1962] theory of poroelasticity
-\end{description}
-\end{block}
-\column{0.27\textwidth}
-\includegraphics[width=\textwidth]{Biot_1964_Buckling_Porous}
-\end{columns}
-
-\medskip
-\begin{footnotesize}
-\textit{Biot, J. Appl. Mech. 1964}: Buckling of a porous slab embedded in an impervious infinite viscous or viscoelastic medium
-\end{footnotesize}
-\begin{itemize}
-\item $B$ transiently higher due to pore pressure
-\item $\sigma$ comes only from the surrounding medium $\rightarrow 0$
-\item effect of flow through the porous slab not considered
-\end{itemize}
-\end{frame}
-
-\begin{frame}{No initial contact with a substrate}
-\structure{Poiseuille vs Darcy on a flying carpet}
-\begin{tikzpicture}
-\begin{axis}[
-	name=a,
-	width=\textwidth, height=0.25\textwidth, scale only axis,
-	domain=0:360, no markers, ymin=-4, ymax=4,xmin=0,xmax=360,
-	axis lines=none, xtick=\empty,
-	]
-	\fill[gray!20] (axis cs:85,4) rectangle (axis cs:95,-4) (axis cs:265,4) rectangle (axis cs:275,-4);
-	\addplot+[Accent1]{sin(x)+0.5};
-	\addplot+[Accent1]{sin(x)-0.5};
-	\addplot+[dashed, black]{0.5};
-	\addplot+[dashed, black]{-0.5};
-	\draw[<->, help lines] (axis cs:180,0.5) -- (axis cs:180,4) node[midway, left] {$H$}; 
-	\draw[<->, help lines] (axis cs:30,-0.5) -- (axis cs:30,-4) node[midway, left] {$H$};
-	\node[below] at (axis cs:90,-1.5) (pb1) {$P_1$};
-	\node[above] at (axis cs:90,1.5) (ph2) {$P_2$};
-	\node[above] at (axis cs:270,1.5) (ph1) {$P_1$};
-	\node[below] at (axis cs:270,-1.5) (pb2) {$P_2$};
-	\draw[line width=0.1em, ->] (ph2) -- (axis cs:90,0.5) node[midway, left] {Darcy} node[midway, right] {$v$};
-	\draw[line width=0.1em, ->] (pb2) -- (pb1) node[midway, above] {Poiseuille} node[midway, below] {$u \sim \frac{\lambda}{H} v$};
-\end{axis}
-\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle +(\textwidth,-1em) (a.north west) rectangle +(\textwidth,1em);
-\end{tikzpicture}
-
-\begin{columns}
-\column{0.83\textwidth}
-\begin{itemize}
-\item The same $\Delta P$ can be relaxed by a Poiseuille flow
-\[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
-\item For the same wavelength
-\hfill$\displaystyle \frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$
-\item Flow through the pores dominates
-\end{itemize}
-\column{0.17\textwidth}
-\includegraphics[width=\textwidth]{Jean-Leonard-Marie_Poiseuille.jpg}\\
-\hfill\structure{1797-1869}
-\end{columns}
-\end{frame}
-
-\againframe<3>{carpet}
+\include{plis_beamer}
 

### 2014-06-19 creep_beamer.tex
video relative link unscaled primary creep stress labels -> vertical

@@
 \end{tikzpicture}%
-}{/home/mathieu/Dropbox/Yaourt/films/Yaourt141_enhanced.avi}
+}{Yaourt141_enhanced.avi}
 \end{frame}
@@
 \begin{loglogaxis}[
+	name=g,
 	width=\textwidth,
@@
 	]
-	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node[left] (s200){200};
-	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){300};
-	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){400};
-	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){550};
-	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){1000};
+	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
 	\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[pos=0.75, below left] {$\dot{\gamma}(t)\sim t^{-0.85}$};
 \end{loglogaxis}
+\begin{scope}[anchor=base east, every node/.style={ rotate=90}]
+	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
+	\node[red!60!black] at (s300 |- g.outer north) {300};
+	\node[red!80!black] at (s400 |- g.outer north) {400};
+	\node[red] at (s550 |- g.outer north) {550};
+	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
+\end{scope}
 \end{tikzpicture}

### 2014-06-28 creep_beamer.tex
ellipse of tertiary regime

@@
 		(axis cs:10,0.4) ellipse[rotate=-30, x radius=0.2\textwidth, y radius=0.12\textwidth]
-		(axis cs:2e2,1.6) ellipse[x radius=0.3\textwidth, y radius=0.05\textwidth]
+		(axis cs:2e2,1.6) ellipse[x radius=0.35\textwidth, y radius=0.05\textwidth]
 		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
@@
 \begin{block}{\textit{\footnotesize Halász et al., PRE 2012}}
+\tikzsetnextfilename{Halasz}
 \begin{tikzpicture}[inner sep=0, ultra thick, font=\footnotesize]

### 2014-06-28 plis_beamer.tex
lighter version of the XYslices
Add videos for carpet and flying carpet case Re-work conclusion to fit in the "mechanics of biogels" theme

@@
 \end{tikzpicture}
-}{volume_plis.avi}
+}{volume_plis_comp.avi}
 \end{frame}
@@
 
-	\includegraphics[width=\textwidth]{volume_cloque_t116}\\
+	\movie[externalviewer]{\includegraphics[width=\textwidth]{volume_cloque_t116}{volume_cloques_comp.avi}\\
 	\begin{tikzpicture}
@@
 glycerol 60\%, $\eta=\SI{10}{\milli\pascal\second}$\\
-\includegraphics[width=0.45\textwidth]{XYslice_gly0}&
-\includegraphics[width=0.45\textwidth]{XYslice_gly60}
+\includegraphics[width=0.45\textwidth]{XYslice_gly0.jpg}&
+\includegraphics[width=0.45\textwidth]{XYslice_gly60.jpg}
 \end{tabu}
 \tikz\draw[line width=0.2em] (0,0) -- ++(0.045\textwidth,0) node[midway, below] {\SI{10}{\micro\metre}};
+
+\structure{To do:} Measure everything to plot $\lambda_\text{theory}$ vs $\lambda_\text{experiment}$
 \end{frame}
 
-\begin{frame}{Conclusion}
-Spontaneous wrinkling of over-acidified casein gels in non-adhesive confinement
+\begin{frame}{Solvent is important in biogel mechanics}
 
-	\structure{Original patterns}
+\structure{Gel formation}
 \begin{itemize}
-		\item nested (optical elements ?)
-		\item kinetic wavelength selection
+	\item non adhesive boundary conditions $\Rightarrow$ synæresis
+	\item gel properties depend on solvent viscosity
 \end{itemize}
-	\structure{A new wrinkling mechanism}
+
+\bigskip
+\structure{Non monotonic $p$H dependence}
+\begin{itemize}
+	\item spontaneous re-swelling $\Rightarrow$ biaxial compression
+	\item lateral confinement $\Rightarrow$ cascade buckling, nested structures.
+\end{itemize}
+
+\bigskip
+\structure{Gel poroelasticity}
 \begin{itemize}
@@
 	\item Darcy can beat Poiseuille
-		\item may extend to other porous sheets (wet paper, biological membranes\ldots)
 \end{itemize}
-	\structure{Viscosity influences}
-	\begin{itemize}
-		\item porous dissipation
-		\item gel formation (theory?)
-		\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ thus wrinkling}
-	\end{itemize}
-	\structure{To do:} Measure everything to plot $\lambda_\text{theory}$ vs $\lambda_\text{experiment}$
+
 \end{frame}

### 2014-07-08 creep/creep_yaourt.tex
Resubmission 2

@@
 Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids and to rupture under an external load.
-Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is reminiscent of brittle solids: after a primary creep regime characterized by a macroscopically homogeneous deformation and a power-law behavior which exponent is fully accounted for by linear viscoelasticity, fractures nucleate and grow logarithmically perpendicularly to shear, up to the sudden rupture of the gel.
+Combining optical and ultrasonic imaging to shear rheology we show that the failure scenario of a protein gel is reminiscent of brittle solids: after a primary creep regime characterized by a power-law behavior which exponent is fully accounted for by linear viscoelasticity, fractures nucleate and grow logarithmically perpendicularly to shear, up to the sudden rupture of the gel.
 A single equation accounting for those two successive processes nicely captures the full rheological response.
-The failure time follows a decreasing power-law with the applied shear stress, similar to the Basquin law of fatigue for solids.
+The failure time follows a decreasing power law with the applied shear stress, similar to the Basquin law of fatigue for solids.
 These results are in excellent agreement with recent fiber-bundle models that include damage accumulation on elastic fibers and exemplify protein gels as model, brittle-like soft solids.
@@
 
-\pacs{}
+\pacs{82.35.Pq, 47.57.Qk, 83.60.-a, 83.80.Kn}
 \maketitle
@@
 Inset: failure time $\tau_f$ vs $\sigma$.
-The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ and $A=40.6\pm 0.5$~s.Pa$^\beta$.
+The red line is the best power-law fit $\tau_f=A\sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ and $A=(4.2\pm 0.1)\,10^{17}$~s.Pa$^\beta$.
 \label{fig1}}
@@
 Gels formed by slow acidification of a sodium caseinate solution display fractures under large strain at fixed low pH values \cite{vanVliet:1995,Lucey:1998}, which makes them perfect candidates to quantify the rupture of soft solids and tackle the above-mentioned issues.
-We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: (i) a primary creep regime characterized by a macroscopically homogeneous strain field where dissipation is dominated by viscous flow through the gel matrix and (ii) the irreversible nucleation and growth of fractures leading to gel failure.
+We demonstrate that under an external load, these casein gels display brittle-like failure that results from two successive physical processes: (i) a primary creep regime where dissipation is dominated by viscous flow through the gel matrix without any detectable macroscopic strain localization and (ii) the irreversible nucleation and growth of fractures leading to gel failure.
 Our results are in full agreement with the predictions of some recent fiber-bundle models and hint to universal features of failure common to both soft and hard solids.
@@
 glucono-$\delta$-lactone (GDL) in powder (Firmenich) is dissolved in the solution and its hydrolysis progressively lowers the pH over the course of 8~hours [Fig.~\ref{suppfig1}(a) in the Supplemental Material].
-While still liquid, the solution is poured into the gap of a polished Plexiglas Couette cell immersed into a temperature-controlled water tank at 25.0$\pm$0.1$^{\circ}$C \cite{note}.
-Rheological data are recorded during gel formation by a stress-controlled rheometer through small amplitude oscillatory shear at frequency $f=1$~Hz (Fig.~1 in the Supplemental Material) \cite{note}.
+While still liquid, the solution is poured into the gap of a polished Plexiglas concentric cylinder (or Taylor-Couette, TC) shear cell immersed into a temperature-controlled water tank at 25.0$\pm$0.1$^{\circ}$C \cite{note}.
+Rheological data are recorded during gel formation by a stress-controlled rheometer through small amplitude oscillatory shear at frequency $f=1$~Hz (Fig.~1 in the Supplemental Material).
 Gelation is complete when the elastic ($G'$) and viscous ($G''$) moduli reach a plateau with $G'\gg G''$.
@@
 polyamide spheres (Orgasol 2002 ES3 NAT 3, Arkema, diameter 30~$\mu$m, density 1.02) that do not modify the final gel properties [Fig.~\ref{suppfig1}(b) in the Supplemental Material].
+The radial position $r$ is measured from the inner rotating cylinder and the vertical axis $z$ points downwards with the origin about 15~mm from the top of the TC cell.
 Failure being irreversible, each creep experiment requires to prepare a fresh sample in-situ.
@@
 Figures~\ref{fig3} and \ref{fig4} gather the results from {\it local} measurements and are discussed below together with each of the successive regimes inferred from {\it global} data.
-Note that these failure dynamics do not qualitatively depend on geometry nor on system composition (see Movie~1 in the Supplemental Materials).
+Note that similar failure dynamics are observed for other casein and GDL concentrations, for different gap widths of the TC cell and in a different geometry namely a plate-plate cell (see Movie~1 in the Supplemental Material).
 
@@
 \includegraphics[width=17.5cm,clip]{Fig3.eps}
-\caption{(color online) Ultrasonic images (left) and direct visualization of the Couette cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$] and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
+\caption{(color online) Ultrasonic images (left) and direct visualization of the TC cell (right) recorded simultaneously at various times in the primary [(a)~$t/\tau_f=0.02$], secondary [(b)~$t/\tau_f=0.57\simeq\taum/\tau_f$] and tertiary  [(c) $t/\tau_f=0.93$ and (d) 0.99] creep regimes.
 Ultrasonic images are azimuthal velocity maps $v(r,z,t)$ computed by averaging over 4~s and coded using the linear color levels shown below the images.
@@
 
-As seen in the inset of Fig.~\ref{fig1}, the failure time $\tau_f$ sharply decreases as a power-law of $\sigma$ with an exponent $\beta\simeq 5.5$.
+As seen in the inset of Fig.~\ref{fig1}, the failure time $\tau_f$ sharply decreases as a power law of $\sigma$ with an exponent $\beta\simeq 5.5$.
 $\tau_f$ further allows us to rescale all the shear rate data $\gp(t)$ onto the single master curve of Fig.~\ref{fig2}(a) by plotting $\gp/\gp_{\rm min}$ vs $t/\tau_f$, where $\gp_{\rm min}$ is the minimum shear rate reached at a time $\tau_{\rm min}$ [see also Fig.~\ref{suppfig2} in the Supplemental Material for the original $\gp(t)$ data].
-On about four decades for $t\lesssim 0.1\tau_f$, the shear rate decreases as a power-law $\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.04$.
+On about four decades for $t\lesssim 0.1\tau_f$, the shear rate decreases as a power law $\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.04$.
 This is strongly reminiscent of the {\it primary creep} observed in solids and referred to as Andrade creep \cite{Andrade:1910,Miguel:2002,Nechad:2005,Rosti:2010}.
@@
 This nicely corresponds to the observed $\gp(t)\sim t^{-0.85}$.
-Moreover ultrasonic imaging reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with insignificant $z$-dependence [Fig.~\ref{fig3}(a, left)] and no slippage at the Plexiglas walls.
-Together with direct visualization [Fig.~\ref{fig3}(a)], these local measurements demonstrate that during primary creep there is no macroscopic strain localization or fracture, although we cannot rule out microscopic rearrangements below the available spatial ($\sim10~\mu$m) and temporal ($\sim1$~s) resolutions.
+Moreover ultrasonic imaging reveals velocity and strain fields averaged over the vertical direction $z$ that linearly decrease with the radial position $r$ within the gap [Fig.~\ref{fig4}(a-b)] with no slippage at the Plexiglas walls [see arrows in Fig.~\ref{fig4}(a)].
+Together with direct visualization [Fig.~\ref{fig3}(a)], these local measurements demonstrate that during primary creep there is no macroscopic strain localization or fracture, although we cannot rule out rearrangements below the available spatial ($\sim10~\mu$m) and/or temporal ($\sim1$~s) resolutions due to limited signal-to-noise ratio at very low shear rates [Fig.~\ref{fig3}(a, left)].
 
@@
 This linearity between $\tau_{\rm min}$ and $\tau_f$, also known as the Monkman-Grant relation \cite{Monkman:1956}, allows one to ``predict'' the failure time from the intermediate-time response.
-During this {\it secondary creep} regime, regularly-spaced cracks nucleate from the top and bottom edges of the Couette cell and start growing perpendicular to the applied stress [Fig.~\ref{fig3}(b), see also Movie~2 in the Supplemental Material].
+During this {\it secondary creep} regime, regularly-spaced cracks nucleate from the top and bottom edges of the cell and start growing perpendicular to the applied stress [Fig.~\ref{fig3}(b), see also Movie~2 in the Supplemental Material].
 These macroscopic fractures are invaded with water expelled from the surrounding gel matrix.
@@
 In the {\it tertiary creep} regime, for $t\gtrsim 0.9\tau_f$, the shear rate increases by more than four orders of magnitude and diverges as $(\tau_f-t)^{-1}$ [Fig.~\ref{fig2}(c)].
-This finite-time singularity corresponds to the final growth of the fractures along the vorticity direction $z$ as they accelerate and eventually meet in the middle of the Couette cell at time $\tau_f$ [Fig.~\ref{fig3}(c-d)].
+This finite-time singularity corresponds to the final growth of the fractures along the vorticity direction $z$ as they accelerate and eventually meet in the middle of the cell at time $\tau_f$ [Fig.~\ref{fig3}(c-d)].
 Ultrasonic velocity maps directly correlate with the crack growth and reveal the complex structure of the displacement field at the tip and around the fracture.
@@
 Solid lines are linear profiles.
+The arrows in (a) indicate the velocity of the inner cylinder inferred from the current shear rate.
 (c)~Spatiotemporal diagram of the local velocity $\langle v(r,z,t)\rangle_r$ averaged over the radial direction $r$ and plotted in linear color levels as a function of $z$ and $t/\tau_f$.
 (d)~Standard deviation $\delta_z v(t)$ of $\langle v(r,z,t)\rangle_r$ taken over the vertical direction $z$ (thick black line) together with corresponding standard deviation $\delta_r v(t)$ computed over the radial direction $r$ on the $z$-average $\langle v(r,z,t)\rangle_z$ (thin red line).
-(e)~Fracture length $\ell(t)$ vs $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$) and normalized by the height $H$ of the Couette cell.
+(e)~Fracture length $\ell(t)$ vs $(\tau_f-t)/\tau_f$ as inferred from direct visualization ($\bullet$, average over 6 different fractures, error bars show the standard deviation) and from ultrasonic imaging ($\circ$) and normalized by the height $H$ of the TC cell.
 Gray dots show the visualization data for the longest fracture which leads to the failure of the sample at $\tau_f$.
@@
 First we have shown that casein gels display a remarkable failure scenario similar to that of brittle solids and characterized by the same three successive creep regimes.
-Here, the power-law exponent of the primary creep [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.04$] is fully accounted for by linear viscoelasticity and local measurements consistently display a homogeneous bulk deformation.
+Here, the power-law exponent of the primary creep [$\gp(t)\sim t^{-\alpha}$ with $\alpha=0.85\pm 0.04$] is fully accounted for by linear viscoelasticity.
 Such a link between creep and viscoelasticity is shared by other biopolymer gels with power-law rheology \cite{Gobeaux:2010,Jaishankar:2013} as well as hard-sphere-like colloidal glasses \cite{Siebenburger:2012}.
@@
 Such an evolution is also commonly reported in disordered solid materials displaying brittle rupture and interpreted in the framework of Griffith-like models based on global or local energy barriers \cite{Vanel:2009}.
-However these approaches all predict exponential scalings for $\tau_f(\sigma)$ while our data are best fitted by the decreasing power-law $\tau_f \sim \sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ [inset of Fig.~\ref{fig1}].
-This last key result suggests that thermally activated crack growth is not relevant to the present soft solid.
+However these approaches all predict exponential scalings for $\tau_f(\sigma)$ while our data are best fitted by the decreasing power law $\tau_f \sim \sigma^{-\beta}$ with $\beta=5.45\pm 0.05$ [inset of Fig.~\ref{fig1}].
+This last key result suggests that thermally activated crack growth does not play any prominent role.
 Rather the power-law scaling is strikingly reminiscent of the Basquin law of fatigue found for a variety of heterogeneous or cellular materials under cyclic deformation \cite{Kun:2007,Basquin:1910,Kohout:2000}.
@@
 To conclude, the present time- and space-resolved study exemplifies protein gels as model, brittle-like soft solids.
-Prompted by the remarkable simplicity of Eq.~(\ref{e.fit}) that encompasses Andrade creep, the Monkman-Grant relation and finite-time singularity within a single equation, future modeling will undoubtedly focus on the microscopic ingredients needed to predict quantitatively the Monkman-Grant prefactor and the Basquin exponent $\beta$.
+Prompted by the remarkable simplicity of Eq.~(\ref{e.fit}) that encompasses Andrade-like creep, the Monkman-Grant relation and finite-time singularity within a single equation, future modeling will undoubtedly focus on the microscopic ingredients needed to predict quantitatively the Monkman-Grant prefactor and the Basquin exponent $\beta$.
 The next experimental step consists in a statistical study of the fluctuations associated to crack nucleation and growth as well as a systematic investigation of other systems in order to check for universality in the irreversible creep rupture of soft solids in general and of biogels such as actin, alginate or agar gels in particular.
@@
 
-\bibitem{note} Rheological data shown in Figs.~\ref{fig1} and \ref{fig2} and in Supplemental Figs.~1 and 2 were obtained with an MCR 301 rheometer (Anton Paar) in a Couette cell of height 28~mm and gap 1~mm with an inner rotating cylinder of radius 24~mm.
-Optical and ultrasonic imaging coupled to rheometry (Figs.~\ref{fig3} and \ref{fig4}) were performed on an AR G2 rheometer (TA Instruments) in a Couette cell of height 60~mm and gap 2~mm with an inner rotating cylinder of radius 23~mm.
-A detailed, quantitative study of the influence of the gap width on the failure dynamics is left for future work.
+\bibitem{note} Rheological data shown in Figs.~\ref{fig1} and \ref{fig2} and in Supplemental Figs.~1 and 2 were obtained with an MCR 301 rheometer (Anton Paar) in a TC cell of height 28~mm and gap 1~mm with an inner rotating cylinder of radius 24~mm.
+Optical and ultrasonic imaging coupled to rheometry (Figs.~\ref{fig3} and \ref{fig4}) were performed on an AR G2 rheometer (TA Instruments) in a TC cell of height 60~mm and gap 2~mm with an inner rotating cylinder of radius 23~mm.
+We checked that the commanded shear stress $\sigma$ is reached within 20~ms.
+A detailed, quantitative study of the influence of the gap width and of the temperature on the failure dynamics is left for future work.
 
@@
 along the radial direction in this case.
-The global rheological response is fully similar to that measured in the Couette geometry.
+The global rheological response is fully similar to that measured in the Taylor-Couette (TC) geometry.
 
@@
 GDL.
-The gap width of the Couette cell is 2~mm and its height is 60~mm.
+The gap width of the TC cell is 2~mm and its height is 60~mm.
 Images recorded by a standard webcam (Logitech Webcam Pro 9000) are shown in the top left panel.
 Velocity maps $v(r,z,t)$ inferred from ultrasonic imaging by averaging over 4~s are shown in the top right panel using linear color levels.
-The vertical position $z=0$ on the ultrasonic images corresponds to about 20~mm below the top of the Couette cell.
+The vertical position $z=0$ on the ultrasonic images corresponds to about 15~mm below the top of the TC cell.
 The two bottom graphs show the global shear rate response $\gp(t)$ (left) and the strain response $\gamma(t)$ recorded by the rheometer (AR G2, TA Instruments) simultaneously to optical and ultrasonic imaging.
@@
 in the secondary creep regime, $\gp(t)$ is smoothed using a moving average over $\delta t\simeq 0.01\tau_f$ in order to remove high-frequency noise due to differentiation at very small shear rates.
-The same data normalized by its minimum value $\gp_{\rm min}$ is plotted in Fig.~\ref{fig2}(a) as a function of $t/\tau_f$ leading to a nice collapse onto a single curve whatever the imposed shear stress.
-
-\clearpage
-\newpage
+For the largest applied shear stress $\sigma=1000$~Pa at times $t<0.1$~s, oscillations are observed in $\gamma(t)$ and consequently in $\gp(t)$ (negative values of the shear rate do not show in Suppl.
+Fig.~\ref{suppfig2} due to logarithmic scales).
+These short-time oscillations, which are well documented in the rheology literature \cite{Benmouffok-Benbelkacem:2010}, arise from the coupling between the sample viscoelasticity and the instrument inertia.
+Here they impose an experimental limit to the extension of the power-law scaling $\gp(t)\sim t^{-0.85}$ at short times.
+The shear rate data normalized by its minimum value $\gp_{\rm min}$ is plotted in Fig.~\ref{fig2}(a) as a function of $t/\tau_f$ once all oscillations have died out.
+This leads to a nice collapse onto a single curve whatever the imposed shear stress.
 
@@
 Successive creep and recovery experiments performed on the same sample within the primary creep regime are presented in Supplemental Figures~\ref{suppfig3} and \ref{suppfig4}.
-When a stress of 100~Pa is applied at time $t=0$, an ``instantaneous'' elastic strain of about 10\% is recorded, followed by the slow Andrade creep already investigated in the main text and characterized by an exponent of 0.15 [Suppl.
+When a stress of 100~Pa is applied at time $t=0$, an ``instantaneous'' elastic strain of about 10\% is recorded, followed by the slow Andrade-like creep already investigated in the main text and characterized by an exponent of 0.15 [Suppl.
 Fig.~\ref{suppfig3}(b)] that coincides with that of the frequency-dependence of both viscoelastic moduli $G'$ and $G''$ [Suppl.
@@
 
+\bibitem{Benmouffok-Benbelkacem:2010} G. Benmouffok-Benbelkacem et al., Rheol Acta {\bf 49}, 305--314 (2010).
+
 \end{thebibliography}

### 2014-09-11 permeability_measurement.tex
Transcribe permeability measurement method from Roefs PhD thesis (1986)

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{kpfonts}
\usepackage{siunitx}
\usepackage{tikz}
\usetikzlibrary{calc}
\author{Roefs, Sebastianus PFM}
\title{Permeability measurements}
\date{1986}
\begin{document}
\maketitle

\subsection*{General}
The resistance of a liquid flow through a porous medium depends on the spatial distribution of the solid phase. In case of a laminar flow through a homogeneous fixed matrix generally Darcy's law is obeyed. For a flow in one direction the liquid flux, $v$, can be read as:
\begin{equation}
v = -B/\eta \nabla P
\label{eq:Darcy}
\end{equation}
\begin{description}
\item[$v$] liquid flux (i.e. volume flow rate/cross-sectional area)(\si{\metre\per\second})
\item[$B$] permeability coefficient (\si{\metre^2})
\item[$\eta$] viscosity of the flowing liquid (\si{\pascal\second})
\item[$\nabla P$] pressure gradient. In our case $\nabla P$ has only one component in the $x$-direction (i.e. parallel to the measuring tubes). So $\nabla P = dP/dx$ (\si{\pascal\per\metre})
\end{description}

The permeability coefficient B depends on the geometry, scale and spatial distribution of the percolated matrix. For flow through a porous medium a Reynolds number $Re$ can be defined (Scheidegger, 1960):
\begin{equation}
Re = v\rho\delta/\eta
\label{eq:Reynolds}
\end{equation}
\begin{description}
\item[$v$] liquid flux (\si{\metre\per\second})
\item[$\rho$] density of the flowing liquid (\si{\kilo\gram\per\cubic\metre})
\item[$\eta$] viscosity of the flowing liquid (\si{\pascal\second})
\item[$\delta$] diameter associated with the porous medium, i.e some length corresponding to an average grain or pore diameter (\si{\metre})
\end{description}

In case of laminar flow it is a prerequisite for applying Darcy's law that $Re$ should not exceed a certain critical value, which varies from 0.1 to 75 depending on the porous medium (Scheidegger, 1960).

\subsection*{The tube method}
\begin{figure}
\center
\begin{tikzpicture}
\node[minimum height=0.25\textwidth, minimum width=0.5\textwidth, fill=blue!20, anchor=north] (whey) at (0,0) {};
\begin{scope}[every node/.style={minimum width=0.05\textwidth}]
%where the tubes will be
\node[minimum height=0.4\textwidth, anchor=west] (tubel) at ($(whey.north)+(-0.12\textwidth,0)$) {};
\node[minimum height=0.4\textwidth, anchor=east] (tuber) at ($(whey.north)+(0.12\textwidth,0)$) {};
%Jurin in tubel
\node[minimum height=2em, anchor=west, fill=blue!20] (wheyinfty) at (tubel.west)  {};
%lack of liquid at t2
\node[minimum height=1em, anchor=north east, fill=white] (wheyt2) at (tuber.east) {};
%lack of liquid at t1
\node[minimum height=\baselineskip, anchor=north, fill=blue!10] (wheyt1) at (wheyt2.south) {};
\end{scope}
%gel
\fill[gray] (wheyt1.south west) +(0,-0.5em) rectangle (tuber.south east);
%draw tubes
\draw[very thick] (tubel.south west) -- (tubel.north west) (tubel.south east) -- (tubel.north east)
(tuber.south west) -- (tuber.north west) (tuber.south east) -- (tuber.north east);

%labels
\draw[<-] (wheyinfty.north west) -- +(-1em,0) node[left]{$h(\infty)$};
\draw[<-] (wheyt2.south east) -- +(1em,0) node[right]{$h(t_2)$};
\draw[<-] (wheyt1.south east) -- +(1em,0) node[right]{$h(t_1)$};
\draw  (wheyt1.south east) ++(0,-0.5em) -- +(1em,0) coordinate[pos=0.9] (Ht) (tuber.south east) -- +(1em,0) coordinate[pos=0.9] (Hb);
\draw[<->] (Ht) -- (Hb) node[midway, right] (H) {$H$};
\draw[<-] (H -| tuber) -- +(-0.05\textwidth,0) node[left]{gel};
\node[text=white,anchor=north west] at (whey.north west) {whey};
\node at (tubel.north) {$1$};
\node at (tuber.north) {$2$};
\end{tikzpicture}
\caption{Schematic representation of permeability measurement.}
\label{fig:tubexp}
\end{figure}
By measuring the liquid flux, $v$, of usually whey through open glass tubes filled with casein gels, the permeability coefficient, $B$, could be calculated from the known pressure gradient, $\nabla P$, according to equation \ref{eq:Darcy}. The apparatus used was developed by Van Dijk (1982) for measurements on rennet gels. Glass tubes open at both ends with an inner diameter of \SI{3.7}{\milli\metre} and a length of \SI{25}{\centi\metre} were placed in a precooled vat. The tubes placed in a holder rested on a plexiglass base. Acidified skimmilk or sodium caseinate was added to the vat to such an extent that the tubes were filled over a length of about \SI{10}{\centi\metre}. The gelation vat was heated according to the procedure given in section 2.3 (see original thesis). After the required gelation time at the appropriate temperature the tubes were with- drawn from the gelation vat, cleaned at the outside and placed in a rack, which was placed in the measuring vat made of clear plexiglass. The level of the percolating liquid in the measuring vat (usually whey or a sodium phosphate solution) was adjusted in such a way that the initial pressure gradient was about \SIrange{4.5e3}{6.0e3}{\pascal\per\metre}. The whey level in the tubes was read at regular time intervals with the help of a travelling microscope. The length of the intervals depended on the permeability of the gel matrix. From the five readings usually made four permeability coefficients could be calculated. The order of magnitude of the flow velocity, $v$, varied from \SIrange{3e-6}{8e-8}{\metre\per\second}. Assuming a value of \SI{1e-7}{\metre} for the parameter $\delta$ (i.e. a low estimate of the approximate diameter of casein particles) in eq.~\ref{eq:Reynolds} while $\rho\simeq \SI{1e3}{\kilo\gram\per\cubic\metre}$ and $\eta \simeq \SI{1e-3}{\pascal\second}$, the Reynolds number, $Re$, for the flow through casein gels varied from \SIrange{8e-9}{3e-7}{}. Taking for $\delta$ a pore diameter of \SIrange{1e-6}{1e-5}{\metre} instead of a particle diameter of \SI{1e-7}{\metre}, $Re$ will be 10 to 100 times larger, but still sufficiently small. There was therefore no risk for turbulent flow in our experiments. Variation of the pressure gradient from \SIrange{4.0}{11.0e3}{\pascal\per\metre} did not influence the calculated permeability coefficients, so that the application of Darcy's law for flow through acid casein gels seems allowable.

The calculation of $B$ from the experimental readings is somewhat complicated, because the pressure gradient is changing with time as the whey level in the glass tubes rises (see fig.~\ref{fig:tubexp}). On the other hand because of the rather homogeneous character of the gel and the one dimensional direction of flow (x-direction) the pressure gradient at a certain moment is constant and negative over the whole gel and is given by:
\begin{equation}
\frac{dP}{dx}(t) = -\rho g\frac{h(\infty)-h(t)}{H}
\end{equation}
\begin{description}
\item[$h(\infty)$] hight of the whey level in the reference tube (\si{\metre})
\item[$h(t)$] height of the whey level in the gel tube (\si{\metre})
\item[$H$] length of the casein gel (\si{\metre})
\item[$g$] gravitational acceleration (\si{\metre\per\square\second})
\end{description}

Because of the large volume outside the tubes $h(\infty)$ is assumed to be constant. The liquid flux, $v$, also changes with time:
\begin{equation}
v(t) = \frac{dh}{dt}(t)
\end{equation}

In this case the equation of Darcy (eq.~\eqref{eq:Darcy}) can be rewritten as:
\begin{equation}
 v(t) = \frac{dh}{dt}(t) = \frac{B}{\eta}\rho g\frac{h(\infty)-h(t)}{H}
\end{equation}

Integration, solving the integration constant and rewriting leads to an expression for $B$, depending on $h(t)$ at $t=t_1$ and $t=t_2$ (Van Dijk, 1982).
\begin{equation}
B = -\frac{\eta H \ln\left(\frac{h(\infty)-h(t_2)}{h(\infty)-h(t_1)}\right)}{\rho g (t_2-t_1)}
\end{equation}

Within the experimental accuracy the permeability coefficient, $B$, of acid casein gels did not change with time during the experiment.

\end{document}

### 2014-09-11 permeability_measurement.tex
Missing x axis
Bibliography

@@
 \node at (tuber.north) {$2$};
+\draw[->] (whey.south west) ++(45:0.075\textwidth) -- +(0,0.1\textwidth) node[above] {$x$};
 \end{tikzpicture}
@@
 Acidified skimmilk or sodium caseinate was added to the vat to such an extent that the tubes were filled over a length of about \SI{10}{\centi\metre}.
-The gelation vat was heated according to the procedure given in section 2.3 (see original thesis).
-After the required gelation time at the appropriate temperature the tubes were with- drawn from the gelation vat, cleaned at the outside and placed in a rack, which was placed in the measuring vat made of clear plexiglass.
+The gelation vat was heated according to the procedure given in section~2.3\footnote{see original thesis}.
+After the required gelation time at the appropriate temperature the tubes were withdrawn from the gelation vat, cleaned at the outside and placed in a rack, which was placed in the measuring vat made of clear plexiglass.
 The level of the percolating liquid in the measuring vat (usually whey or a sodium phosphate solution) was adjusted in such a way that the initial pressure gradient was about \SIrange{4.5e3}{6.0e3}{\pascal\per\metre}.
@@
 
+\section*{Bibliography}
+Dijk, H.J.M, van (1982). Doctoral thesis Agricultural University Wageningen, the Netherlands.\\
+Scheidegger, A.E. (1960). "The physics of flow through porous media". Chapter VI and VII, p. 112-193. Oxford University Press London.\\
+
 \end{document}

### 2014-09-18 creep_beamer.tex
Adapt to wall slip workshop. Fix broken tex. tikzpictures re-generated for each overlay.

@@
 \begin{scope}[shift={(a.north west)}, shift={(5.8\baselineskip,-1.1\baselineskip)}, ultra thick]
-	\draw[Accent1, x radius=3.1\baselineskip, y radius=0.3\baselineskip] ellipse ++(3.1\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-3.1\baselineskip,0) ellipse[] ++(-3.1\baselineskip,0)  -- ++(0,7.4\baselineskip);
-	\shadedraw[Accent2,ball color=Accent2, x radius=2\baselineskip, y radius=0.1\baselineskip] ellipse ++(2\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-2\baselineskip,0) ellipse[] ++(-2\baselineskip,0)  -- ++(0,7.4\baselineskip);
+	\draw[Accent1, x radius=3.1\baselineskip, y radius=0.3\baselineskip] ellipse[] ++(3.1\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-3.1\baselineskip,0) ellipse[] ++(-3.1\baselineskip,0)  -- ++(0,7.4\baselineskip);
+	\shadedraw[Accent2,ball color=Accent2, x radius=2\baselineskip, y radius=0.1\baselineskip] ellipse[] ++(2\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-2\baselineskip,0) ellipse[] ++(-2\baselineskip,0)  -- ++(0,7.4\baselineskip);
 	\fill[Accent2, left color=black,right color=black, middle color=Accent2, ] (-0.3\baselineskip,0) rectangle +(0.6\baselineskip, \baselineskip);
@@
 \end{frame}
-\tikzset{external/force remake=false}
+%\tikzset{external/force remake=false}
 

### 2014-09-18 plis_beamer.tex
Adapt to wall slip workshop. Fix broken tex. tikzpictures re-generated for each overlay.

@@
 Associate Professor & CNRS researcher & Professor\\
-Aix-Marseille Université & E.N.S. Lyon & E.N.S. Lyon\\
+Aix-Marseille Universit\'{e} & E.N.S. Lyon & E.N.S. Lyon\\
 \end{tabu}
@@
 \begin{frame}{Over-acidification: Faster but weaker}
+\tikzsetnextfilename{gelation_GDL}
 \begin{tikzpicture}
@@
 	\structure{Top view of the final pattern}\\
+	\tikzsetnextfilename{patterns_zooms}
 	\begin{tikzpicture}[inner sep=0, very thick]
@@
 \begin{frame}{Dynamics (transmission macroscope)}
-\movie[externalviewer]{\begin{tikzpicture}
+\movie[externalviewer]{\tikzsetnextfilename{dynamics_still}\begin{tikzpicture}
 \node[inner sep=0] (a) {\includegraphics[width=\textwidth]{prise_0799_resized.jpg}};
@@
 \begin{frame}{Dynamics (transmission macroscope)}
+\tikzsetnextfilename{snapshots_macroscope}
 \begin{tikzpicture}
@@
 
+\tikzsetnextfilename{scalebar_fluo}
 \begin{tikzpicture}
@@
 \column{0.5\textwidth}
+\tikzsetnextfilename{volume_area}
 \begin{tikzpicture}
@@
 \begin{frame}{Dynamics (confocal microscope)}
+\tikzsetnextfilename{dynamics_confocal_still}
 \movie[externalviewer]{\begin{tikzpicture}
@@
 \begin{frame}{Dynamics (confocal microscope)}
+\tikzsetnextfilename{snapshots_confocal}
 \begin{tikzpicture}
@@
 
-	\movie[externalviewer]{\includegraphics[width=\textwidth]{volume_cloque_t116}{volume_cloques_comp.avi}\\
+	\movie[externalviewer]{\includegraphics[width=\textwidth]{volume_cloque_t116}}{volume_cloques_comp.avi}\\
+	\only<1>{\tikzsetnextfilename{cloques_still}
 	\begin{tikzpicture}
 	\draw[ultra thick] (0,) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
-	\end{tikzpicture}
+	\end{tikzpicture}}
 
@@
 }{
+\tikzsetnextfilename{alts_bottom_cloque}
 \begin{tikzpicture}
@@
 	\draw[->, Accent2, ultra thick] (axis cs:90,1.25) -- (axis cs:90,2.25) node[pos=1, right] {$v$};
-	\draw[Main, ultra thick, {Circle}->] (axis cs:-90,1.5) -- (axis cs:-90,0.25) node[right] {$\sigma$};
+	\draw[Main, ultra thick,->] (axis cs:-90,1.5) -- (axis cs:-90,0.25) node[right] {$\sigma$};
 	%\node[above] at (axis cs:-90,1.5) {$P_2$};
@@
 \structure{Poiseuille vs Darcy on a flying carpet}
+\tikzsetnextfilename{flying_carpet}%
 \begin{tikzpicture}
@@
 \begin{frame}{Cell thickness}
+\tikzsetnextfilename{thickness}
 \begin{tikzpicture}[mark size={0.3em}]
@@
 
-\tikzset{external/force remake}
+%\tikzset{external/force remake}
 \begin{frame}{Surprising viscosity influence}
@@
 \[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3}\]
+\tikzsetnextfilename{viscosity_dependence}
 \begin{tikzpicture}

### 2014-09-18 wokshop_wallslip.tex
Adapt to wall slip workshop. Fix broken tex. tikzpictures re-generated for each overlay.

--- 2014-06-18 seminaire_Tanakalab.tex
@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,arrows.meta,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
+\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
 %\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
@@
 \usepgfplotslibrary{external}
+\makeatletter
+\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
+\tikzset{
+  beamer externalizing/.style={%
+    execute at end picture={%
+      \tikzifexternalizing{%
+        \ifbeamer@anotherslide
+        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
+        \fi
+      }{}%
+    }%
+  },
+  external/optimize=false
+}
+\let\orig@tikzsetnextfilename=\tikzsetnextfilename
+\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
+\makeatother
+
+\tikzset{every picture/.style={beamer externalizing}}
 \tikzexternalize
@@
 
-\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Supérieure de Lyon}
+\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
 \title[Yoghurt under stress]{Yoghurt under stress}
@@
 
-%\includeonly{}
+%\includeonly{creep_beamer}
 
@@
 
-\begin{frame}{Application of biogels}
+\begin{frame}{Application of biogels \textup{\normalsize(physical gels
+$E_{bond}\gg k_B T$)}}
 \begin{tabu}{cX[c]c}
@@
 \column{0.5\textwidth}
+\tikzsetnextfilename{gelation}
 \begin{tikzpicture}
@@
 \[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
+\tikzsetnextfilename{power_law_rheology}
 \begin{tikzpicture}
@@
 	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};	
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot{590*x^0.14};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot{350*x^0.12};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot{246*x^0.1};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot{163*x^0.067};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot{109*x^0.044};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot{76*x^0.035};
-
 \end{loglogaxis}
@@
 \end{frame}
-
-\section*{Creep and yielding}
-
-\include{creep_beamer}
 

### 2014-09-18 plis_beamer.tex
Adapt intro

@@
 \structure{Imaging plateform} BioSciences Gerland - Lyon Sud (UMS~3444)
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{Over-acidification: Faster but weaker}
-\tikzsetnextfilename{gelation_GDL}
-\begin{tikzpicture}
-\begin{axis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
-	cycle list name=earthy,
-	no marks,
-	xmin=0, xmax=20,ymin=0,
-	]
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
-\end{axis}
-\end{tikzpicture}
-
-Caseins regain some solubility at low pH.
 \end{frame}

### 2014-09-18 seminaire_Tanakalab.tex
Definition of biogels
Adapt intro

@@
 
+%\tikzset{external/force remake}
+\begin{frame}{Over-acidification: Faster but weaker}
+\tikzsetnextfilename{gelation_GDL}
+\begin{tikzpicture}
+\begin{axis}[
+	height=0.8\textheight,
+	width=\textwidth,
+	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
+	cycle list name=earthy,
+	no marks,
+	xmin=0, xmax=20,ymin=0,
+	]
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
+\end{axis}
+\end{tikzpicture}
+
+Caseins regain some solubility at low pH.
+\end{frame}
+
 \include{plis_beamer}

### 2014-09-18 wokshop_wallslip.tex
Adapt intro

@@
 \institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
-\title[Yoghurt under stress]{Yoghurt under stress}
+\title[Yoghurt wrinkling]{Wrinkling of a confined porous layer}
 \author[M. Leocmach]{Mathieu Leocmach}
-\date{1st June 2014}
+\date{24\textsuperscript{th} September 2014}
 \titlegraphic{
@@
 \begin{itemize}
-\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
-\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
-\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
+\item Gel stability and solvent expulsion (syn\ae{}resis) \textit{\scriptsize Bartlett et al. PRE 2012}
 \item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
@@
 \begin{tabu}{X[c]X[c]}
-\includegraphics[height=6\baselineskip]{Bonn_fracture} &
-\includegraphics[height=6\baselineskip]{Villi_sq}
+\includegraphics[height=8\baselineskip]{faisselle} &
+\includegraphics[height=8\baselineskip]{Villi_sq}
 \end{tabu}
@@
 \begin{itemize}
-\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
+\item What is the role of boundary conditions?
 \item What is the role of the solvent in mechanical instabilities?
@@
 		%xmin=1e2,xmax=3e5,
-		xmin=0, xmax=20,
+		xmin=0, xmax=4,
 		scale only axis,
@@
 		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y190_cas4_gdl1.pH};
-	\draw[help lines] (axis cs:0,4.6) -- (axis cs:3.88,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:3.88,0);
+	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y189_cas4_gdl4.pH};
+	\draw[help lines] (axis cs:0,4.6) -- (axis cs:0.26,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:0.26,0);
 
@@
 		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y22.prise};
-	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL1_Y22.prise};
+	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise};
+	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL4_Y271.prise};
 	\begin{scope}[font=\scriptsize]
@@
 		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
-		\node[anchor=base,rotate=80] at (axis cs:2.5,400) {gelation};
+		\node[anchor=base,rotate=80] at (axis cs:0.2,300) {gelation};
 	\end{scope}
 	\end{groupplot}
-	\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
+	%\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
 \end{tikzpicture}
@@
 \end{frame}
-\tikzset{external/force remake=false}
+%\tikzset{external/force remake=false}
 
-\begin{frame}{Linear rheology: Power law (visco)elastic solid}
-\[\text{stress}\rightarrow \sigma = G \gamma\leftarrow\text{strain}\]
-\[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
-\tikzsetnextfilename{power_law_rheology}
+%\tikzset{external/force remake}
+\begin{frame}{Over-acidification: Faster but weaker}
+\tikzsetnextfilename{gelation_GDL}
 \begin{tikzpicture}
-\begin{loglogaxis}[
+\begin{axis}[
 	height=0.8\textheight,
 	width=\textwidth,
-	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{Storage} and \textcolor{Accent2}{loss} moduli (\si{\pascal})},
-	domain={6e-2:70},
+	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
+	cycle list name=earthy,
+	no marks,
+	xmin=0, xmax=20,ymin=0,
 	]
-	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};	
-\end{loglogaxis}
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
+	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
+\end{axis}
 \end{tikzpicture}
+
+Caseins regain some solubility at low pH.
 \end{frame}

### 2014-09-22 plis_beamer.tex
Remove old speculations and partial scalings. Better present Poiseuille. Use permeability instead of pore size. What can we change experimentally.

@@
 \item[$v$] gel velocity = -volumic flux $\simeq \SI{0.2}{\micro\metre\per\second}$
-\item[$\ell_p$] pore size $ \approx \SI{4}{\micro\metre}$
+\item[$\alpha$] permeability $ \approx \SI{6e-14}{\square\metre}$
 \end{description}
 \item $\sigma$ is a Darcy pressure
-\[\sigma = \Delta P \sim \eta h v/\ell_p^2\]
+\[\sigma = \Delta P \sim \eta h v/\alpha\]
 \item poroelastic length
-\[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}\]
+\[ \lambda^* \sim \left(\frac{B \alpha}{\eta v h}\right)^\frac{1}{3} \simeq \SI{0.5}{\milli\metre}\]
 \end{itemize}
@@
 \item $\lambda$ frozen at ceiling touch 
-\item $A=H\equiv e-h$
-\item $\lambda \sim \epsilon^{-3/7} H \sim e$
+\item Amplitude $A=e-h$
+\item $\epsilon = \left(\frac{e-h}{\lambda^*}\right)^\frac{7}{4}$
+\item $\lambda \sim \left(e-h\right)^\frac{1}{4} \left(\lambda^*\right)^\frac{3}{4}$
 	%\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq 0.6\% \]
@@
 \begin{columns}
-\column{0.83\textwidth}
-\begin{itemize}
-\item The same $\Delta P$ can be relaxed by a Poiseuille flow
-\[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
-\item For the same wavelength
-\hfill$\displaystyle \frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$
-\item Flow through the pores dominates
-\end{itemize}
 \column{0.17\textwidth}
@@
 \hfill\structure{1797-1869}
+
+\column{0.80\textwidth}
+The same $\Delta P$ can be relaxed by a Poiseuille flow
+\begin{align*}
+\Delta P &\sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\\
+\lambda^* &= \left(\frac{B H^2}{12\eta\lambda}\right)^{1/3}
+\end{align*}
+\textit{\footnotesize Huang \& Suo, J. Appl. Phys. (2002)}
+%\item For the same wavelength
+%\hfill$\displaystyle \frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$
+%\item Flow through the pores dominates
+
 \end{columns}
@@
 
-%\tikzset{external/force remake}
-\begin{frame}{Cell thickness}
-\tikzsetnextfilename{thickness}
-\begin{tikzpicture}[mark size={0.3em}]
-	\matrix[matrix of nodes, inner sep=0, column sep=0.01\columnwidth, row sep=0.01\columnwidth, ampersand replacement=\&] (m) {
-		\includegraphics[width=0.245\columnwidth]{pattern_50um.jpg}\&
-		\includegraphics[width=0.245\columnwidth]{pattern_100um.jpg}\&
-		\includegraphics[width=0.245\columnwidth]{pattern_250um.jpg}\&
-		\includegraphics[width=0.245\columnwidth]{pattern_450um.jpg}\\
-	};
-	\node[inner sep=0, below left=0.01\columnwidth and 0 of m.south east] (x800) {\includegraphics[width=0.49\columnwidth]{pattern_800um_x1.jpg}};
-	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.196\columnwidth, 0.05\columnwidth) -- ++(0.145\columnwidth,0) node[midway,above]{$\lambda$};
-	\draw[line width=0.2em,Main](x800.north east) -- ++(-0.217\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};
-
-	\begin{axis}[%
-		name=g,
-		anchor=left of north west,
-		at={($(m.south west)+(0,-0.01\columnwidth)$)},
-		scale only axis,
-		width=0.5\textwidth-3em,
-		height=0.25\columnwidth-2em,
-		xmin=0,xmax=1,xlabel={$e$ (\si{\milli\metre})},
-		xtick={0,0.4,...,0.9},
-		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
-		]
-		\addplot[no marks, domain=0:3] {4.23*x};
-		\begin{scope}[inner sep=0, minimum size=0.8em]
-			\node[circle, fill=Accent1] at (axis cs:0.05,0.261){};
-			\node[circle, fill=white, draw=Accent1] at (axis cs:0.1,0.458) {};
-			\node[regular polygon, regular polygon sides=3, fill=Accent1] at (axis cs:0.250,0.916){};
-			\node[regular polygon, regular polygon sides=3, fill=white, draw=Accent1] at (axis cs:0.450,2.075){};
-			\node[fill=Accent1] at (axis cs:0.800,3.333){};
-		\end{scope}
-	\end{axis}
-	\begin{scope}[inner sep=0, minimum size=0.8em, transform canvas={xshift=0.5em, yshift=-0.5em}]
-		\node[circle, fill=Accent1] at (m-1-1.north west) {};
-		\node[circle, fill=white, draw=Accent1] at (m-1-2.north west){};
-		\node[regular polygon, regular polygon sides=3, fill=Accent1] at (m-1-3.north west){};
-		\node[regular polygon, regular polygon sides=3, fill=white, draw=Accent1] at (m-1-4.north west) {};
-		\node[fill=Accent1] at (x800.north west) {};
-	\end{scope}
-\end{tikzpicture}
-\end{frame}
-%\tikzset{external/force remake=false}
-
-%\tikzset{external/force remake}
-\begin{frame}{Surprising viscosity influence}
+\begin{frame}{Poiseuille vs Darcy}
 \begin{columns}
 \column{0.5\textwidth}
-\[ \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3}\]
-\tikzsetnextfilename{viscosity_dependence}
-\begin{tikzpicture}
-	\begin{loglogaxis}[%
-		width=\textwidth,
-		height=0.8\textwidth,
-		xmin=0.6,xmax=11,xlabel={viscosity (\si{\milli\pascal\second})},
-		xtick={0.7, 1, 2, 5, 10}, xticklabels={$0.7$, $1$, $2$, $5$, $10$},
-		ytick={1,2,3,4}, yticklabels={$1$,$2$,$3$,$4$},
-		ymin=1, ymax=5, ylabel={$\lambda$ (\si{\milli\metre})},
-		mark size={0.3em},
-		legend style={legend pos=south east},
-		]
-		\addplot+[no marks, domain=0.7:10, black, forget plot] {1.7*x^0.4} node[midway, below] {$\eta^{0.4}$};
-		\addplot+[only marks,Accent1, mark=o] table{eta_lambda_eau_pure.txt};
-		\addplot+[only marks,Accent2, mark=square] table{eta_lambda_eau_glycerol.txt};
-		\legend{water, water-glycerol};
-	\end{loglogaxis}
-\end{tikzpicture}
-
+\begin{block}{Observable wavelengths}
+Same magnitude (typical sample)
+\begin{align*}
+\lambda_D &\sim \Upsilon^\frac{1}{4} h^\frac{1}{2} \alpha^\frac{1}{4}\simeq \SI{0.23}{\milli\metre}\\
+\lambda_P &\sim \Upsilon^\frac{1}{6} h^\frac{1}{2} H^\frac{1}{2}\simeq \SI{0.25}{\milli\metre}
+\end{align*}
+\end{block}
+with $\Upsilon = \frac{E(e-h)}{\eta v}\simeq 10^{6-8}$
 \column{0.5\textwidth}
+Limiting case:
 \begin{itemize}
-\item $\lambda$ increases with $\eta$
-\item Can be explained only by a hidden dependence of the viscosity on $B$, $\ell_p$, \ldots 
-\item[$\Rightarrow$] Viscosity influences gel formation.
+\item Same growth speed
+\item Same wavelength
+\item[$\Rightarrow$] $H^* = \Upsilon^\frac{1}{6} \alpha^\frac{1}{2}\simeq \SI{6}{\micro\metre}$
 \end{itemize}
-%If $\ell_p \sim \eta^{-\alpha}$ and $B\sim \mathrm{k_B}T \left(\frac{h}{\ell}\right)^3$, then
-%\[ \lambda^* \sim \eta^\frac{\alpha-1}{3} \]
-%\[\alpha\simeq 2.2\]
+Darcy can be faster than Poiseuille when $H<H^*$.
 \end{columns}
+\begin{block}{What can we change experimentally?}
+$H$ and $v$ are measured a posteriori but cannot be controlled.
+\begin{description}[leftmargin=8em]
+\item[Geometry] $e$, $h$
+\item[Composition] $E$, $\alpha$
+\item[Solvent] $\eta$ \alert{but also $E$ and $\alpha$}
+\end{description}
+\end{block}
 \end{frame}
@@
 
-\structure{To do:} Measure everything to plot $\lambda_\text{theory}$ vs $\lambda_\text{experiment}$
+May be explained by ``viscoelastic phase separation'' arguments
+\end{frame}
+
+\begin{frame}{Darcy never beats Poiseuille}
+
 \end{frame}

### 2014-10-08 poster_Palavas.tex
poster pour Palavas les flots

--- 2014-06-03 poster_Okinawa.tex
@@
 \usepackage{kpfonts}
+%\usepackage{roboto}
 \usepackage{fontspec}
@@
 \setmainfont{Tex Gyre Pagella}
-\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}
+%\setsansfont[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}
+\setsansfont{Choplin-Medium-DEMO}
+%\setsansfont{RobotoSlab-Bold}
+%\setsansfont[StylisticSet=4]{MEgalopolisExtra}
 %\renewcommand*\sfdefault{ugq}
@@
 
+\usepackage{tabu}
 \usepackage{multirow}
@@
 \usepackage{pgfplots}
+\pgfplotsset{compat=1.9}
+\usepackage{pgfplotstable}
 %\usepgfplotslibrary{units}
@@
 \usepgfplotslibrary{external}
-%\tikzexternalize
-%\tikzsetexternalprefix{fig_Rome/}
+\tikzexternalize
+\tikzsetexternalprefix{fig_poster_palavas/}
 \tikzset{external/system call={lualatex \tikzexternalcheckshellescape -halt-on-error -interaction=batchmode -jobname "\image" "\texsource"}}
@@
 \def\Head#1{\Norulehead{\dotfill #1}}
-\def\LHead#1{\noindent{\LARGE #1}\smallskip}
+\def\LHead#1{{\LARGE #1}\smallskip}
 \def\Subhead#1{\noindent{\large\color{Accent1}\textsc{#1}}}
-\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{#1}}}
+\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{\uppercase{#1}}}}
 
@@
 \begin{document}
-%\tikzset{every mark/.append style={scale=0.8}}
-%\pgfplotsset{every axis/.append style={small}}
+\pgfplotscreateplotcyclelist{earthy}{%
+red!40!black,
+red!60!black,
+red!80!black,
+red,
+red!80!yellow,
+red!60!yellow,
+red!40!yellow,
+}
 
@@
 % but it looks awful.
-\begin{textblock}{15}(0,0)
-\centering
+%\begin{textblock}{15}(0,0)
+\begin{center}
 \Title{
-Wrinkling of a confined porous layer
+Role of the microstructure in the failure of a protein gel under stress
 }
+\end{center}
 
 \smallskip
-\LHead{Mathieu Leocmach, Mathieu Nespoulous, Thomas Gibaud, Sébastien Manneville
-}\hfill\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}\\
+\LHead{Mathieu Leocmach, Christophe Perge, Thomas Gibaud, Sébastien Manneville
+}\hfill\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}
 
-\LHead{\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}\hfill\raisebox{-1.5\baselineskip}{
-	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
-	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
-	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
+\LHead{\textsc{Laboratoire de Physique, Ecole Normale Supérieure de Lyon}}
+
+\LHead{Thibaut Divoux\qquad
+\textsc{Centre de Recherche Paul Pascal}
+\hfill\raisebox{0pt}[0pt]{
+	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\;
+	\includegraphics[height=2\baselineskip]{logo_ums_grand}\;
+	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\;
+	\includegraphics[height=2\baselineskip]{CRPP}\;
 	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
-}}
+}}\\
+
+%\end{textblock}
+
+\begin{textblock}{11}(0,3)
+	\Head{Over-acidified protein gels: beyond isoelectric pH}
+
+	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (\textsc{gdl})
+
+	\tikzsetnextfilename{prise_cas4}
+	\input{fig_poster_palavas/prise_cas4.pgf}
 
@@
 
-\begin{textblock}{3}(0,2.5)
-	\Head{Acid-induced gels}
-
-	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (4\%w \textsc{gdl})
-
-	\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		xmin=0,xmax=8,
-		scale only axis,
-		width=\textwidth-4em,
-		height=0.3\textwidth,
-		extra tick style={grid=major},%
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
-		]
-	\nextgroupplot[
-		ymin=0, ymax=7, ylabel={\textcolor{Accent1}{pH}},
-		extra y ticks={4.6}, extra y tick labels={},%
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
-	\node[base left=0] at (axis cs:8,4.6) {isoelectric};
-
-	\nextgroupplot[
-		xlabel={time (h)},
-		ymin=0, ylabel=\textcolor{Accent2}{$G^\prime$ (\si{\pascal})}
-		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
-	\end{groupplot}
-	\end{tikzpicture}
-
-\Subhead{Microscopy}
-\begin{description}
-\item[macroscope] transmission \& reflection
-\item[confocal] 10\% labelled casein
-\end{description} 
-\end{textblock}
-
-\begin{textblock}{7}(4,2.5)
-	\Head{Sealed cell, spontaneous pattern}
-
-	%\Subhead{Side view}\\
-	\tikzsetnextfilename{cell_brushes}
-	\tikzset{external/force remake}
-	\begin{tikzpicture}
-		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\textwidth,1.5em) node[midway,fill=white,inner sep=1pt] {glass};
-		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\textwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
-		\draw[line width=2pt,Accent1] (0.05\textwidth,-2.5em) -- (0.95\textwidth,-2.5em) (0.05\textwidth,-1pt) -- (0.95\textwidth,-1pt) node[below,pos=0.30, text width=0.4\textwidth] {acrylamide brush ($10\sim 100$ nm)\linebreak $\Rightarrow$ No adhesion};
-		\fill[gray] (0,0) rectangle (0.05\textwidth,-2.5em) (\textwidth,0) rectangle (0.95\textwidth,-2.5em) node[pos=1, above left] {spacer};
-		\draw[<->] (0.75\textwidth,-2pt) -- (0.75\textwidth,-2.5em) node[midway,left] {$e\sim 100\,\mu m$};
-		\draw[<->] (0.05\textwidth,2.25em) -- (0.95\textwidth,2.25em) node[midway,above] (L){$L\sim 2.5\,cm$};
-		\node[anchor=north west, inner sep=0] at ($(L.north) - (0.5\textwidth,0)$) {\Subhead{Side view of the cell}};
-	\end{tikzpicture}
-
-	\Subhead{Top view of the final pattern}\\
-	\begin{tikzpicture}[inner sep=0, very thick]
-	\setlength{\mylength}{\columnwidth}
-	\node[anchor=north west] (a) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2_crop}};
-		\node[anchor=north] at (0.44\mylength,0) (b) {\includegraphics[width=0.28\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6_crop}};
-		\node[anchor=north east] at (\mylength,0) (c) {\includegraphics[width=0.4\mylength]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
-		%zooms
-		\node[minimum width = 0.137\mylength, minimum height=0.137\mylength, anchor=north west, draw=Accent2] at ($(a.north west) +(0.111\mylength,-0.072\mylength)$) (bz){};
-		\draw[Accent2] (bz.north east) -- (b.north west) (bz.south east) -- (b.south west);
-		\node[minimum width = 0.128\mylength, minimum height=0.081\mylength, anchor=north west, draw=Main] at ($(b.north west) +(0.097\mylength,-0.143\mylength)$) (cz){};
-		\draw[Main] (cz.north east) -- (c.north west) (cz.south east) -- (c.south west);
-		\node[minimum width = 0.156\mylength, minimum height=0.156\mylength, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\mylength,0)$) (dz) {};
-		%scale bars
-		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.178\mylength,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
-		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.176\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
-		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\mylength,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
-
-		%final volumic view
-		%\node[anchor=north east] at (\columnwidth,0) (c) {\includegraphics[width=0.4\mylength]{recol_volume_plis}};
-	\end{tikzpicture}
+\begin{textblock}{11}(0,8.5)
+	\Head{Linear rheology, from power-law to glassy}
+	\tikzsetnextfilename{sweep_cas4}
+	\input{fig_poster_palavas/sweep_cas4.pgf}
 
@@
 
-\begin{textblock}{15}(0,8)
-	\Head{Dynamics: wrinkling patterns and nested buckling}
-	%\Head{No initial contact with a substrate: Poiseuille vs Darcy on a flying carpet}
-	\begin{tikzpicture}
-	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em] (m){
-	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
-	\includegraphics[width=0.13\textwidth]{prise_0100_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0130_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0160_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0190_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0250_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0360_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0799_color.jpg}\\
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
-	};
-	\draw[line width=0.2em] ++(m-2-1.north west) -- ++(0.023\textwidth,0) node[pos=0, above right, inner xsep=0] {\SI{1}{\milli\metre}};
-	\draw[line width=0.2em] ++(m-3-1.south west) -- ++(0.1\textwidth,0) node[pos=0, below right, inner xsep=0] {\SI{1}{\milli\metre}};
-	\node[anchor=north east, rotate=90] at (m-2-1.north east) (macro) {macroscope};
-	\node[anchor=base, rotate=90] at (m-3-1.east-| macro.base) {confocal};
-	\end{tikzpicture}
-\end{textblock}
+\begin{textblock}{3}(12,8.5)
+	\Head{Microscopy}
+	Final state without over-acidification
 
-\begin{textblock}{7}(0,11)
-	\Head{Synæresis and swelling cause buckling}
-\end{textblock}
-
-\begin{textblock}{3}(0,11.5)
-\Subhead{Initial contact with a substrate}\\
-	\begin{tikzpicture}
-	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png} & \SI{15}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png} & \SI{22}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png} & \SI{1}{\hour}~26\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png} & \SI{1}{\hour}~30\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png} & \SI{1}{\hour}~31\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png} & \SI{1}{\hour}~35\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png} & \SI{2}{\hour}~25\\
-	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png} & \SI{2}{\hour}~40\\
-	};
-	\draw[<->, ultra thick] ($(m-8-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
-	\end{tikzpicture}
-
-Cascade buckling \textit{Roman \& Pocheau 1999}
-
-\Subhead{No initial contact}\\
-\begin{tikzpicture}
-	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t000.png} & \SI{15}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t016.png} & \SI{23}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t032.png} & \SI{32}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t038.png} & \SI{35}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t040.png} & \SI{36}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t043.png} & \SI{38}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t055.png} & \SI{44}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t070.png} & \SI{53}{\minute}\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t123.png} & \SI{1}{\hour}~21\\
-	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t332.png} & \SI{3}{\hour}~15\\
-	};
-	\draw[<->, ultra thick] ($(m-10-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
-	\end{tikzpicture}
-
-	Globally unstable wavelength
-\end{textblock}
-
-\begin{textblock}{3}(4,11.5)
-	\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		xmin=0, xmax=170, xtick={0,30,...,150},
-		extra tick style={grid=major},%
-		extra x ticks = {21,68}, extra x tick labels={},%
-		scale only axis,
-		width=\textwidth-4em,
-		height=0.3\textwidth,
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
-		]
-
-	\nextgroupplot[
-		ylabel={Volume (\%)}, ymin=20, ymax=100, 
-		ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em},
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_cloques.txt};
-	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
-	\node[anchor=south west] at (rel axis cs:0,0) {\rotatebox{90}{synæresis}};
-	\node at (axis cs:68,80) {swelling};
-
-	\nextgroupplot[
-		ylabel={Excess area (\%)}, ymin=0, ymax=4, ytick={0,1,2,3},
-		xlabel={time (min)}, 
-		every axis y label/.append style={xshift=-1em},
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, below left] {all};
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{3}*100}]{relative_volume_excess_area_cloques.txt} node[pos=1, left] {single blister};
-	\draw[->] (axis cs: 21, 3) -- (axis cs: 68, 3) node[midway, below] {\SI{1}{\hour}};
-	%\legend{all, single blister};
-	\end{groupplot}
-	\end{tikzpicture}
-
-	\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		xmin=0, xmax=180, xtick={0,30,...,150},
-		extra tick style={grid=major},%
-		extra x tick labels={},%
-		scale only axis,
-		width=\textwidth-4em,
-		height=0.3\textwidth,
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
-		]
-
-	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em}]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_plis.txt};
-	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
-
-	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=6, ytick={0,2,4}, 
-	xlabel={time (min)}, every axis y label/.append style={xshift=-1em}]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_plis.txt};
-	%\addplot+[only marks,Accent2, mark=+] table{excess_area_pc_half_cas8_toi.txt};
-	\end{groupplot}
-	\end{tikzpicture}
-\end{textblock}
-
-\begin{textblock}{3}(12,12)
-\textit{Kolinski et al 2009, Vella et al 2009}\\
-Bending energy $B$ vs vertical stress $\sigma$:
-\[ \lambda^* = \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]
-\end{textblock}
-
-\begin{textblock}{7}(8,11)
-	\Head{Wavelength: resisting stress needed}
-	\Subhead{Initial contact with a substrate : Ruck in a rug analogy}
-
-\begin{tikzpicture}[ultra thick]
-\begin{axis}[
-	name=a,
-	width=3\TPHorizModule, height=0.75\TPHorizModule, scale only axis,
-	domain=-180:180, no markers, ymin=0, ymax=4,xmin=-180,xmax=180,
-	axis lines=none, xtick=\empty,
-	]
-	\addplot+[Accent1]{cos(x)+1};
-	\addplot+[Accent1]{cos(x)+1.5};
-	\draw[->, ultra thick] (axis cs:90,1.25) -- +(45:0.05\textwidth) node[pos=1, above] {$\vec{v}$};
-	\node[above] at (axis cs:-90,1.5) {$P_2$};
-	\node[below right] at (axis cs:-90,1.25) {$P_1$};
-	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
-\end{axis}
-\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle ($(a.south east)+(0,-1em)$);
-\draw[<->] ($(a.south west)+(0,-1.1em)$) -- ($(a.south east)+(0,-1.1em)$) node[midway, below] {$\lambda$};
-\end{tikzpicture}
-
-Actual shape depends on relative excess length $\epsilon$: $\lambda \sim \epsilon^{1/7} \lambda^*$ and $A \sim \epsilon^{4/7} \lambda^*$
-
-\Subhead{Poroelasticity instead of gravity} \hfill Vertical cells show the same wavelength
-\begin{itemize}
-\item Solvent must flow through the gel(volumic flux $v\simeq \SI{0.2}{\micro\metre\per\second}$, pore size $\ell_p \approx \SI{4}{\micro\metre}$)
-\item[$\Rightarrow$] $\sigma$ is a Darcy pressure
-$\sigma = \Delta P \sim \eta h v/\ell_p^2   $
-\hfill\fcolorbox{Main}{white}{$\Rightarrow$ Poroelastic length
-$\displaystyle \lambda^* \sim \left(\frac{B \ell_p^2}{\eta v h}\right)^\frac{1}{3} \simeq \SI{2}{\milli\metre}$}
-\item Different from usual poroelasticity (\textit{Biot 1964}) where pore pressure increases transiently the bending energy. Here Darcy flow resists large wavelength.
-\end{itemize}
-
-\Subhead{No initial contact with a substrate: Poiseuille vs Darcy on a flying carpet}
-
-\smallskip
-\begin{tikzpicture}
-\begin{axis}[
-	name=a,
-	width=\textwidth, height=0.25\textwidth, scale only axis,
-	domain=0:360, no markers, ymin=-4, ymax=4,xmin=0,xmax=360,
-	axis lines=none, xtick=\empty,
-	]
-	\fill[gray!20] (axis cs:85,4) rectangle (axis cs:95,-4) (axis cs:265,4) rectangle (axis cs:275,-4);
-	\addplot+[Accent1]{sin(x)+0.5};
-	\addplot+[Accent1]{sin(x)-0.5};
-	\addplot+[dashed, black]{0.5};
-	\addplot+[dashed, black]{-0.5};
-	\draw[<->, help lines] (axis cs:180,0.5) -- (axis cs:180,4) node[midway, left] {$H$}; 
-	\draw[<->, help lines] (axis cs:30,-0.5) -- (axis cs:30,-4) node[midway, left] {$H$};
-	\node[below] at (axis cs:90,-1.5) (pb1) {$P_1$};
-	\node[above] at (axis cs:90,1.5) (ph2) {$P_2$};
-	\node[above] at (axis cs:270,1.5) (ph1) {$P_1$};
-	\node[below] at (axis cs:270,-1.5) (pb2) {$P_2$};
-	\draw[line width=0.1em, ->] (ph2) -- (axis cs:90,0.5) node[midway, left] {Darcy} node[midway, right] {$v$};
-	\draw[line width=0.1em, ->] (pb2) -- (pb1) node[midway, above] {Poiseuille} node[midway, below] {$u \sim \frac{\lambda}{H} v$};
-\end{axis}
-\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle +(\textwidth,-1em) (a.north west) rectangle +(\textwidth,1em);
-\end{tikzpicture}
-
-\begin{itemize}
-\item The same $\Delta P$ can be relaxed by a Poiseuille flow
-$\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v$
-\item For the same wavelength
-$\frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30$ \hfill$\Rightarrow$ Flow through the pores dominates
-\end{itemize}
-\end{textblock}
-
-\begin{textblock}{7}(0,18)
-	\Head{Cell thickness}
-\begin{tikzpicture}[mark size={0.3em}]
-	\matrix[matrix of nodes, inner sep=0, column sep=0.01\columnwidth, row sep=0.01\columnwidth] (m) {
-		\includegraphics[width=0.245\columnwidth]{pattern_50um.jpg}&
-		\includegraphics[width=0.245\columnwidth]{pattern_100um.jpg}&
-		\includegraphics[width=0.245\columnwidth]{pattern_250um.jpg}&
-		\includegraphics[width=0.245\columnwidth]{pattern_450um.jpg}\\
-	};
-	\node[inner sep=0, below left=0.01\columnwidth and 0 of m.south east] (x800) {\includegraphics[width=0.49\columnwidth]{pattern_800um_x1.jpg}};
-	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.196\columnwidth, 0.05\columnwidth) -- ++(0.145\columnwidth,0) node[midway,above]{$\lambda$};
-	\draw[line width=0.2em,Main](x800.north east) -- ++(-0.217\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};
-
-	\begin{axis}[%
-		name=g,
-		anchor=left of north west,
-		at={($(m.south west)+(0,-0.01\columnwidth)$)},
-		scale only axis,
-		width=3\TPHorizModule-4em,
-		height=0.25\columnwidth-2em,
-		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
-		xtick={0,0.4,...,0.9},
-		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em},
-		cycle multi list={%
-			Accent1,no marks\\%
-			Accent1,only marks,every mark/.append style={fill=Accent1},mark=*\\%
-			Accent1,only marks,every mark/.append style={fill=white},mark=*\\%
-			Accent1,only marks,every mark/.append style={fill=Accent1},mark=triangle*\\%
-			Accent1,only marks,every mark/.append style={fill=white},mark=triangle*\\%
-			Accent1,only marks,every mark/.append style={fill=Accent1},mark=square*\\%
-			Accent1,only marks,every mark/.append style={fill=white},mark=square*\\%
-		},
-		mark size={0.3em}
-		]
-		\addplot+[domain=0:3] {4.23*x};
-		\addplot coordinates {(0.05,0.261)};\label{pgfplots:50}
-		\addplot coordinates {(0.1,0.458)};\label{pgfplots:100}
-		\addplot coordinates {(0.250,0.916)};\label{pgfplots:250}
-		\addplot coordinates {(0.450,2.075)};\label{pgfplots:450}
-		\addplot coordinates {(0.800,3.333)};\label{pgfplots:800}
-		%\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
-	\end{axis}
-	\node[below right=0 of m-1-1.north west] {\ref{pgfplots:50}};
-	\node[below right=0 of m-1-2.north west] {\ref{pgfplots:100}};
-	\node[below right=0 of m-1-3.north west] {\ref{pgfplots:250}};
-	\node[below right=0 of m-1-4.north west] {\ref{pgfplots:450}};
-	\node[below right=0 of x800.north west] {\ref{pgfplots:800}};
-\end{tikzpicture}
-
-\Subhead{Wavelength stop evolving after a blister touches the ceiling}
-\end{textblock}
-
-\begin{textblock}{3}(0,22.25)
-	\begin{tikzpicture}
-\begin{axis}[
-	scale only axis,
-	width=\textwidth-4em,
-	height=0.4\textwidth,
-	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
-	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
-	ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em},
-	cycle list={
-		{Accent2}, 
-		{black!15!Accent2},
-		{black!30!Accent2},
-		{black!45!Accent2},
-		{black!60!Accent2},
-		{black!75!Accent2},
-		{black}
-	},
-	no marks,
-	]
-	\foreach \x in {2,3,..., 8}
-		\addplot table[y index=\x]{alts_bottom_cloque.txt};
-
-	\draw[ultra thick, Accent1] (axis cs:580,0) -- (axis cs:580,50);
-\end{axis}
-\end{tikzpicture}
-\end{textblock}
-
-\begin{textblock}{3}(4,22.25)
-At contact: $A=H\equiv e-h\approx e/2$
-	\[ \lambda \sim \epsilon^{-3/7} H \]
-	\[ \epsilon \sim \left(\frac{\lambda}{e}\frac{e}{H}\right)^{-7/3} \simeq 0.6\% \]
+	\tikzsetnextfilename{structure_finale}
+	\input{fig_poster_palavas/structure_finale.pgf}
 
@@
 
-\begin{textblock}{7}(8,21)
-\Head{Surprising viscosity influence}
-\begin{tikzpicture}
-	\begin{loglogaxis}[%
-		scale only axis,
-		width=3\TPHorizModule-4em,
-		height=2\TPHorizModule,
-		xmin=0.6,xmax=11,xlabel={viscosity (\si{\milli\pascal\second})},
-		xtick={0.7, 1, 2, 5, 10}, xticklabels={$0.7$, $1$, $2$, $5$, $10$},
-		ytick={1,2,3,4}, yticklabels={$1$,$2$,$3$,$4$},
-		ymin=1, ymax=5, ylabel={$\lambda$ (\si{\milli\metre})},
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em},
-		mark size={0.3em},
-		legend style={legend pos=south east},
-		]
-		\addplot+[no marks, domain=0.7:10, black, forget plot] {1.7*x^0.4} node[midway, below] {$\eta^{0.4}$};
-		\addplot+[only marks,Accent1, mark options={fill=Accent1}] table{eta_lambda_eau_pure.txt};
-		\addplot+[only marks,Accent2, mark options={fill=Accent2}] table{eta_lambda_eau_glycerol.txt};
-		\legend{water, water-glycerol};
-	\end{loglogaxis}
-\end{tikzpicture}
+\begin{textblock}{11}(0,11.5)
+	\Head{Evolution of the microscructure: preserved network, swelling aggregates}
+	\tikzsetnextfilename{structure_confocal}
+	\input{fig_poster_palavas/structure_confocal.pgf}
+
 \end{textblock}
 
-\begin{textblock}{3}(12,21.5)
-\begin{itemize}
-\item $\lambda$ increases with $\eta$
-\item Can be explained only by a hidden dependence of $\eta$ on $B$, $\ell_p$, \ldots 
-\item[\Rightarrow] Viscosity influences gel formation.
-\end{itemize}
-If $\ell_p \sim \eta^{-\alpha}$ and $B\sim \mathrm{k_B}T \left(\frac{h}{\ell}\right)^3$, then
-\[ \lambda^* \sim \eta^\frac{\alpha-1}{3} \]
-\[\alpha\simeq 2.2\]
+\begin{textblock}{11}(0,15.5)
+	\Head{Three creep regimes: much more chaotic when over-acidified}
+	\tikzsetnextfilename{creep_cas4_gdl4}
+	\input{fig_poster_palavas/creep_cas4_gdl4.pgf}
+
+\end{textblock}
+
+\begin{textblock}{3}(12,15.5)
+	\Head{Monkman-Grant}
+	Predict failure from the minimum
+
+	\tikzsetnextfilename{MonkmanGrant}
+	\input{fig_poster_palavas/MonkmanGrant.pgf}
+
+\end{textblock}
+
+\begin{textblock}{11}(0,18.5)
+	\Head{Basquin law rescaling}
+	\begin{tabu}{XX}
+	\Subhead{Composition} rescaled by elastic modulus&
+	\Subhead{Geometry} rescaled as a fracture nucleation rate
+	\end{tabu}
+
+	\tikzsetnextfilename{basquin}
+	\input{fig_poster_palavas/basquin.pgf}
+
+\end{textblock}
+
+\begin{textblock}{3}(12,20)
+	\Head{Wavelength}
+	\tikzsetnextfilename{lambda}
+	\input{fig_poster_palavas/lambda.pgf}
+
 \end{textblock}
@@
 \TPMargin*{ 0.125\TPHorizModule }
-\begin{textblock}{2.875}(12,2.625)%real width of 3
+\begin{textblock}{2.875}(12,3.125)%real width of 3
 	\Norulehead{Take home}

### 2014-10-13 poster_Palavas.tex
Mise en page avec minipage plutôt que TextPos, garde la grille

@@
 % display. (\documentclass[a0,draft]{a0poster})
+
+\usepackage[margin=40mm]{geometry}
 
@@
 
+\usepackage{framed}
+\colorlet{shadecolor}{lightgray!50!white}
+
 \definecolor{Main}{rgb}{1, 0.57, 0}
@@
 \def\Subhead#1{\noindent{\large\color{Accent1}\textsc{#1}}}
-\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{\uppercase{#1}}}}
+\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{\fontspec[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}{#1}}}}
 
@@
 % but it looks awful.
-%\begin{textblock}{15}(0,0)
+
 \begin{center}
@@
 
-\smallskip
+\vfill
 \LHead{Mathieu Leocmach, Christophe Perge, Thomas Gibaud, Sébastien Manneville
@@
 
-%\end{textblock}
-
-\begin{textblock}{11}(0,3)
+\vfill
+\begin{minipage}[b][21.5\TPVertModule]{11\TPHorizModule}
 	\Head{Over-acidified protein gels: beyond isoelectric pH}
 
-	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (\textsc{gdl})
+	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (\textsc{gdl})\\ \textit{Bremer et al., Colloids Surf.
+(1990); van Vliet \& Walstra, Faraday Discuss.
+(1995); Lucey \& Singh, Food Res.
+Int.
+(1997); Leocmach et al.
+PRL (2014)}
 
+	\smallskip
 	\tikzsetnextfilename{prise_cas4}
 	\input{fig_poster_palavas/prise_cas4.pgf}
+	\begin{tabu}{X[5]XX[5]}
+	Caseins regain solubility below isoelectric $\Rightarrow$ overshoot in $G^\prime$&&
+	Single time scale $(t_m-t_i)$ governs gelation and softening.\\
+	How does it translate in term of microstructure?&&Diverge when $pH_\infty\rightarrow pI$
+	\end{tabu}
 
-\end{textblock}
-
-\begin{textblock}{11}(0,8.5)
-	\Head{Linear rheology, from power-law to glassy}
+	\vfill
+	\Head{Linear rheology, from power-law to soft glassy}
 	\tikzsetnextfilename{sweep_cas4}
 	\input{fig_poster_palavas/sweep_cas4.pgf}
+	\begin{tabu}{X[4]X[2]X[5]}
+	Near isoeletric, power law rheology typical of gels.&&
+	Minimum in loss modulus appear with over-acidification.\\
+	Flatter with over-acidification.&%&
+	\multicolumn{2}{l}{\textcolor{Accent1}{$\blacktriangleright$} Over-acidified behaviour typical of soft glassy materials}\\
+	&\textit{Sollich, PRE (1998)}&\textit{Mason, Curr. Opin. Colloid Interface Sci. (1999)}\\
+	\end{tabu}
 
-\end{textblock}
+	\vfill
+	\Head{Evolution of the microscructure: preserved network, swelling aggregates}
+	\tikzsetnextfilename{structure_confocal}
+	\input{fig_poster_palavas/structure_confocal.pgf}
+	\begin{tabu}{X[5]XX[5]}
+	Maximum $G^\prime$ corresponds to maximum contrast between phases $\chi$&&
+	\textcolor{Accent1}{$\blacktriangleright$} Strongest gel when attraction is strongest\\
+	Characteristic size $\xi$ increases during over-acidification&&
+	\textcolor{Accent1}{$\blacktriangleright$} Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy\\
+	\end{tabu}
 
-\begin{textblock}{3}(12,8.5)
+	\vfill
+	\Head{Three creep regimes: much more erratic when over-acidified}
+	\tikzsetnextfilename{creep_cas4_gdl4}
+	\input{fig_poster_palavas/creep_cas4_gdl4.pgf}
+	\begin{tabu}{X[4]X[3]X[3]}
+	$\tau_f$: divergence of the strain $\gamma$, failure time&
+	\multicolumn{2}{l}{Robust Andrade creep, compatible with power-law rheology \hfill\textit{Leocmach et. al, PRL (2014)}}\\
+	$\tau_\text{min}$: minimum of shear rate $\dot{\gamma}$, fracture nucleation&
+	Robust finite time divergence&
+	\textcolor{Accent1}{$\blacktriangleright$}Possible to reform bonds? Glassy?
+	\end{tabu}
+
+	\vfill
+	\Head{Predict fracture nucleation time: Basquin law rescaling}
+	\begin{tabu}{XX}
+	\hspace{\TPHorizModule}\Subhead{Composition} rescaled by elastic modulus&
+	\hspace{\TPHorizModule}\Subhead{Geometry} rescaled as a fracture nucleation rate
+	\end{tabu}
+
+	\tikzsetnextfilename{basquin}
+	\input{fig_poster_palavas/basquin.pgf}
+	\begin{tabu}{X[5]XX[5]}
+	Robust Basquin law despite glassy behaviour&&\textcolor{Accent1}{$\blacktriangleright$} Failure determined by network breaking \& fracture nucleation
+	\end{tabu}
+
+\end{minipage}
+\hfill
+\begin{minipage}[b][21.5\TPVertModule]{3\TPHorizModule}
+\begin{shaded*}
+	\Norulehead{Take home}
+	\Subhead{Over acidification}
+	\begin{itemize}
+		\item protein aggregates swell
+		\item power-law rheology, decreasing exponent
+		\item dissipation becomes glassy
+	\end{itemize}
+	\Subhead{Creep response}
+	\begin{itemize}
+		\item Andrade creep
+		\item finite time singularity
+		\item erratic transition
+	\end{itemize}
+	\Subhead{Fracture time prediction}
+	\begin{itemize}
+		\item composition influence limited to $G^\prime$
+		\item geometry factors as a nucleation rate
+	\end{itemize}
+\end{shaded*}
+
+	\vfill
 	\Head{Microscopy}
-	Final state without over-acidification
+	Final state without over-acidification\\
 
@@
 
-\end{textblock}
-
-\begin{textblock}{11}(0,11.5)
-	\Head{Evolution of the microscructure: preserved network, swelling aggregates}
-	\tikzsetnextfilename{structure_confocal}
-	\input{fig_poster_palavas/structure_confocal.pgf}
-
-\end{textblock}
-
-\begin{textblock}{11}(0,15.5)
-	\Head{Three creep regimes: much more chaotic when over-acidified}
-	\tikzsetnextfilename{creep_cas4_gdl4}
-	\input{fig_poster_palavas/creep_cas4_gdl4.pgf}
-
-\end{textblock}
-
-\begin{textblock}{3}(12,15.5)
+	\vfill
 	\Head{Monkman-Grant}
@@
 	\input{fig_poster_palavas/MonkmanGrant.pgf}
+	Height-dependent near isoelectric\\
+	Upper bound for over-acidified
 
-\end{textblock}
-
-\begin{textblock}{11}(0,18.5)
-	\Head{Basquin law rescaling}
-	\begin{tabu}{XX}
-	\Subhead{Composition} rescaled by elastic modulus&
-	\Subhead{Geometry} rescaled as a fracture nucleation rate
-	\end{tabu}
-
-	\tikzsetnextfilename{basquin}
-	\input{fig_poster_palavas/basquin.pgf}
-
-\end{textblock}
-
-\begin{textblock}{3}(12,20)
+	\vfill
 	\Head{Wavelength}
@@
 	\input{fig_poster_palavas/lambda.pgf}
+	Compatible with nucleation rate
 
-\end{textblock}
+\end{minipage}
 
-\begin{textblock}{15}(0,24.75)
+\vfill
 \small{\texttt{The research leading to these results has received funding from the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement n°~258803.}}
-\end{textblock}
-
-\textblockcolour{lightgray!50!white}
-\TPMargin*{ 0.125\TPHorizModule }
-\begin{textblock}{2.875}(12,3.125)%real width of 3
-	\Norulehead{Take home}
-	Spontaneous wrinkling of over-acidified casein gels in non-adhesive confinement
-
-	\Subhead{Original patterns}
-	\begin{itemize}
-		\item nested (optical elements ?)
-		\item kinetic wavelength selection
-	\end{itemize}
-	\Subhead{A new wrinkling mechanism}
-	\begin{itemize}
-		\item porous flow resist bending
-		\item Darcy can beat Poiseuille
-		\item may extend to other porous sheets\linebreak  (wet paper, biological membranes\ldots)
-	\end{itemize}
-	\Subhead{Viscosity influences}
-	\begin{itemize}
-		\item porous dissipation
-		\item gel formation\hfill
-		\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ thus wrinkling}
-	\end{itemize}
-
-\end{textblock}%Conclusion
 

### 2014-10-13 soundbite_palavas.tex
soundbite Palavas

--- 2014-09-18 wokshop_wallslip.tex
@@
 \pgfplotsset{compat=1.9}
+\usepackage{pgfplotstable}
 \usepgfplotslibrary{groupplots}
 \usepgfplotslibrary{external}
-\makeatletter
-\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
-\tikzset{
-  beamer externalizing/.style={%
-    execute at end picture={%
-      \tikzifexternalizing{%
-        \ifbeamer@anotherslide
-        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
-        \fi
-      }{}%
-    }%
-  },
-  external/optimize=false
-}
-\let\orig@tikzsetnextfilename=\tikzsetnextfilename
-\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
-\makeatother
-
-\tikzset{every picture/.style={beamer externalizing}}
 \tikzexternalize
-\tikzsetexternalprefix{fig_presentation/}
+\tikzsetexternalprefix{fig_soundbite_palavas/}
 %\tikzset{external/optimize=false}
@@
 \institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
-\title[Yoghurt wrinkling]{Wrinkling of a confined porous layer}
+\title{Role of the microstructure in the failure of a protein gel under stress}
 \author[M. Leocmach]{Mathieu Leocmach}
-\date{24\textsuperscript{th} September 2014}
+\date{}
 \titlegraphic{
+	\begin{tabu}{X[c]X[c]X[c]X[c]}
+		\includegraphics[height=3\baselineskip]{Chris}&
+		\includegraphics[height=3\baselineskip]{Thomas}&
+		\includegraphics[height=3\baselineskip]{Thibaut}&
+		\includegraphics[height=3\baselineskip]{Seb}\\
+		Christophe Perge & Thomas\linebreak Gibaud & Thibaut\linebreak Divoux & Sebastien Manneville\\
+	\end{tabu}
+
+	\vfill
 	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
@@
 	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
+	\includegraphics[height=2\baselineskip]{CRPP}\quad
 	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
@@
 \pgfplotscreateplotcyclelist{earthy}{%
-{red!40!black,mark=o},
-{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
-{red!80!black,mark=square},
-{red,mark=triangle},
-{red!80!yellow, mark=diamond},
+red!40!black,
+red!60!black,
+red!80!black,
+red,
+red!80!yellow,
 red!60!yellow,
@@
 
-\begin{frame}{Application of biogels \textup{\normalsize(physical gels
-$E_{bond}\gg k_B T$)}}
-\begin{tabu}{cX[c]c}
-\includegraphics[height=0.3\textheight]{cell_mech} &
-\includegraphics[height=0.3\textheight]{electrophoresis} &
-\includegraphics[height=0.3\textheight]{cosmetics} \\
-Cell mechanics & Electrophoresis & Cosmetics\\
-\end{tabu}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=0.3\textheight]{bacterial_culture} &
-\includegraphics[height=0.3\textheight]{food} \\
-Bacterial culture & Food\\
-\end{tabu}
+\begin{frame}{Over-acidified protein gels: beyond isoelectric pH}
 
-\begin{footnotesize}
-\begin{block}{Sources}
-\begin{tabu}{lXl}
-Wikimedia Commons & www.madaboutscience.com & www.keautystore.com\\
-\end{tabu}
-\end{block}
-\end{footnotesize}
+	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (\textsc{gdl})\\ %\textit{Bremer et al., Colloids Surf.
+(1990); van Vliet \& Walstra, Faraday Discuss.
+(1995); Lucey \& Singh, Food Res.
+Int.
+(1997); Leocmach et al.
+PRL (2014)}
+
+	\smallskip
+	\tikzsetnextfilename{prise_cas4}
+	\input{fig_soundbite_palavas/prise_cas4.pgf}
+	\begin{itemize}
+	\item Caseins regain charges, solubility below isoelectric $\Rightarrow$ $G^\prime\searrow$
+	\item The sample is still solid
+	\end{itemize}
 \end{frame}
 
-\begin{frame}{Fundamental issues in biogels}
+\begin{frame}{Linear rheology, from power-law to soft glassy}
+	\tikzsetnextfilename{sweep_cas4}
+	\input{fig_soundbite_palavas/sweep_cas4.pgf}
 	\begin{itemize}
-\item Gel stability and solvent expulsion (syn\ae{}resis) \textit{\scriptsize Bartlett et al. PRE 2012}
-\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
+	\item Near isoelectric, power law rheology $G^\prime, G^{\prime\prime}$ typical of gels 
+	\item Over-acidification lowers exponent and creates a minimum in $G^{\prime\prime}$
+	\item Transition to soft-glassy behaviour \\\textit{\footnotesize Sollich, PRE (1998); Mason, Curr. Opin. Colloid Interface Sci. (1999)}
 	\end{itemize}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=8\baselineskip]{faisselle} &
-\includegraphics[height=8\baselineskip]{Villi_sq}
-\end{tabu}
-\begin{block}{Our questions}
-\begin{itemize}
-\item What is the role of boundary conditions?
-\item What is the role of the solvent in mechanical instabilities?
-\end{itemize}
-\end{block}
 \end{frame}
 
-%\section{Our system: Yoghurt}
-
-%\tikzset{external/force remake}
-\begin{frame}{Acid-set protein gel (Yoghurt)}
-\begin{block}{Recipe: take your time}
-\begin{itemize}
-\item Water (\SI{30}{\celsius})
-\item Sodium caseinate (milk protein) 4\%
-\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
-\item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow homogeneous acidification
-\end{itemize}
-\end{block}
-
-\begin{columns}
-\column{0.5\textwidth}
-\tikzsetnextfilename{gelation}
-\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		%xmode=log,
-		%xmin=1e2,xmax=3e5,
-		xmin=0, xmax=4,
-		scale only axis,
-		width=\textwidth-4em,
-		height=0.3\textwidth,
-		extra tick style={grid=major},%
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
-		]
-	\nextgroupplot[
-		ymin=0, ymax=7, ylabel={pH},
-		extra y ticks={4.6}, extra y tick labels={},%
-		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{Y189_cas4_gdl4.pH};
-	\draw[help lines] (axis cs:0,4.6) -- (axis cs:0.26,4.6) node[pos=1, above right, font=\scriptsize] {isoelectric point $pH\approx 4.6$}  -- (axis cs:0.26,0);
-
-	\nextgroupplot[
-		xlabel={time (\si{\hour})},
-		ymin=0, ylabel={$\textcolor{Accent2}{G^\prime}, \textcolor{Main}{G^{\prime\prime}}$ (\si{\pascal})},
-		extra x ticks={17}, extra x tick labels={},
-		]
-	\addplot+[no marks,Accent2] table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise};
-	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600}, y index=2]{cas4_GDL4_Y271.prise};
-	\begin{scope}[font=\scriptsize]
-		%\node[above right] at (rel axis cs:0,0) {casein ``micelles''};
-		%\draw[<-] (axis cs:3.88,654) -- +(1em,0) node[right] {gelation};
-		\node[anchor=base,rotate=80] at (axis cs:0.2,300) {gelation};
-	\end{scope}
-	\end{groupplot}
-	%\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
-\end{tikzpicture}
-
-\column{0.5\textwidth}
-\includegraphics[width=\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
-
-\begin{scriptsize}
-\textit{Kaláb (1983),\linebreak Roefs \& van Vliet (1990),\linebreak Lucey \& Singh (1998)}
-\end{scriptsize}
-\end{columns}
-\end{frame}
-%\tikzset{external/force remake=false}
-
-%\tikzset{external/force remake}
-\begin{frame}{Over-acidification: Faster but weaker}
-\tikzsetnextfilename{gelation_GDL}
-\begin{tikzpicture}
-\begin{axis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
-	cycle list name=earthy,
-	no marks,
-	xmin=0, xmax=20,ymin=0,
-	]
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
-\end{axis}
-\end{tikzpicture}
-
-Caseins regain some solubility at low pH.
+\begin{frame}{Questions}
+\begin{description}
+\item[What] is the change in microstructure?
+\item[How] does it affects the failure mechanism?
+\end{description}
 \end{frame}
 
-\section*{Wrinkling}
+\begin{frame}{Microscructure: retain network, swell aggregates}
+	\tikzsetnextfilename{structure_confocal}
+	\input{fig_soundbite_palavas/structure_confocal.pgf}
+	\begin{itemize}
+	\item Maximum $G^\prime$ corresponds to maximum contrast between phases
+	\item Characteristic size $\xi$ increases during over-acidification
+	\item Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy
+	\end{itemize}
+\end{frame}
 
-\include{plis_beamer}
+\begin{frame}{Three creep regimes + softness}
+\vfill
+	\tikzsetnextfilename{creep_cas4_gdl4}
+	\input{fig_soundbite_palavas/creep_cas4_gdl4.pgf}
+%	\begin{description}
+%		\item[$\tau_f$] divergence of the strain
+%		\item[$\tau_\text{min}$] minimum of shear rate $\dot{\gamma}$
+%	\end{description}
+%	\begin{itemize}
+%		\item Erratic position of $\tau_\text{min}$
+%		\item Possible to reform bonds? Glassy?
+%	\end{itemize}
+\begin{itemize}
+%\item Failure scenario almost conserved: ``fibre bundle'' network
+%\item Other dissipation due to softness: overcompressed gel
+\item Prediction of failure from when fractures nucleate more difficult
+\end{itemize}
+
+\end{frame}
+
+\begin{frame}[plain]
+	\titlepage
+\end{frame}
+
+\appendix
+\newcounter{finalframe}
+\setcounter{finalframe}{\value{framenumber}}
+
+\begin{frame}{Basquin law rescaling}
+	\tikzsetnextfilename{basquin}
+	\input{fig_soundbite_palavas/basquin.pgf}
+\end{frame}
+
+\begin{frame}{Monkman-Grant: Predict failure from nucleation}
+	\tikzsetnextfilename{MonkmanGrant}
+	\input{fig_soundbite_palavas/MonkmanGrant.pgf}
+\end{frame}
+
+\begin{frame}{Wavelength}
+	\tikzsetnextfilename{lambda}
+	\input{fig_soundbite_palavas/lambda.pgf}
+	\textcolor{Accent1}{$\blacktriangleright$} $\lambda\propto e$ compatible with nucleation rate
+
+\end{frame}
 

### 2014-10-18 creep_beamer.tex
seminaire Gulliver. Manque Darcy_vs_Poiseuille.tikz

@@
 \begin{frame}{The yoghurt creep experiment}
+\tikzsetnextfilename{creep_spoon}
 \begin{tikzpicture}
@@
 \begin{frame}{The yoghurt creep experiment}
-\movie[externalviewer]{\begin{tikzpicture}
+\movie[externalviewer]{\tikzsetnextfilename{creep_xp}\begin{tikzpicture}
 \node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{creep_experiment.jpg}};
@@
 \begin{frame}{Creep rheology: Three regimes}
+\tikzsetnextfilename{three_regimes}
 \begin{tikzpicture}
@@
 \column{0.55\textwidth}
+\tikzsetnextfilename{basquin}
 \begin{tikzpicture}
@@
 \column{0.45\textwidth}
+\tikzsetnextfilename{basqun_otherfits}
 \begin{tikzpicture}
@@
 \end{frame}
-\tikzset{external/force remake=false}
 
@@
 \column{0.55\textwidth}
+\tikzsetnextfilename{primary_creep}
 \begin{tikzpicture}
@@
 \bigskip
-\includegraphics[height=7\baselineskip]{Andrade_1910_fig1}
+\includegraphics[height=7\baselineskip]{Andrade_1910_fig1}%
+\tikzsetnextfilename{Andrade_historical}%
 \begin{tikzpicture}
@@
 \begin{block}{Partial creep and recovery}
+\tikzsetnextfilename{recovery}
 \begin{tikzpicture}
@@
 	height=8\baselineskip,
-	xlabel={time (\si{\second})},
+	xlabel={time (\si{\hour})},
 	ylabel={strain $\gamma$},
-	xmin=-1e3, xmax=2.8e4,
+	xmin=-0.15, xmax=7.5, xtick={0, 1, ..., 7},
 	axis background/.style={fill=white},
 	]
-	\addplot[black, no marks] table {Y261_oscil_creep_recovery_100Pa_decimated.txt};
+	\addplot[black, no marks] table[x expr={\thisrowno{0}/3600}] {Y261_oscil_creep_recovery_100Pa_decimated.txt};
 	\begin{scope}[<-,Main, inner sep=0]
 	\draw (axis cs:0,0) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow \SI{100}{\pascal}$};
-	\draw (axis cs:600,0.2727007) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow 0$};
-	\draw (axis cs:8600,0.0159) -- +(0,1em) node[anchor=base east]{$\gamma\leftarrow 0$};
+	\draw (axis cs:1/6,0.2727007) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow 0$};
+	\draw (axis cs:2.388,0.0159) -- +(0,1em) node[anchor=base east]{$\gamma\leftarrow 0$};
 	\end{scope}
@@
 \column{0.55\textwidth}
+\tikzsetnextfilename{primary_rescaled}
 \begin{tikzpicture}
@@
 \column{0.45\textwidth}
+\tikzsetnextfilename{linear_rheology}
 \begin{tikzpicture}
@@
 
+\tikzsetnextfilename{tertiary_rescaled}
 \begin{tikzpicture}
@@
 
+\tikzsetnextfilename{fracture_length}
 \begin{tikzpicture}
@@
 
+\tikzsetnextfilename{secondary_rescaled}
 \begin{tikzpicture}
@@
 \structure{Monkman-Grant relation}
+\tikzsetnextfilename{Monkman-Grant}
 \begin{tikzpicture}
@@
 \column{0.7\textwidth}
+\tikzsetnextfilename{Kun_2006_basquin}
 \begin{tikzpicture}[inner sep=0]
@@
 \end{itemize}
+\tikzsetnextfilename{creep_summary}
 \begin{tikzpicture}

### 2014-10-23 2014_10_23_memo.tex
Theory giving prefactors and the mixt model between Darcy and Poiseuille

\documentclass[12pt,a4paper]{revtex4}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{kpfonts}

\usepackage{siunitx}

\begin{document}
\author{Mathieu Leocmach}
\title{Wrinkling à la Biot}
\maketitle

Mechanical equilibrium of a plate of thickness $h$, young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_0$ along $x$ and a lateral load $q$ along $z$:
\begin{equation}
\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial w^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q,
\end{equation}
where $w(x,t)$ is the deflection of the plate along $z$. It can be generalised as
\begin{equation}
B\frac{h^3}{12}\frac{\partial^4 w}{\partial w^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = -B_s \frac{\partial w}{\partial x},
\label{eq:mecheq}
\end{equation}
where $B=\frac{E}{1-\nu^2}$ and $B_s$ are the respective moduli (in \si{\pascal}) of the plate and the (effective) substrate. In general $B$ and $B_s$ are operators.

Let us suppose a sinusoidal deflection with a time dependent amplitude $w(x,t) = A(t) \sin kx$. Equation~\ref{eq:mecheq} becomes
\begin{equation}
B\frac{h^3}{12}k^4 - \sigma_0 h k^2 + B_s k =0,
\end{equation}
and the compression load can be expressed as
\begin{equation}
\sigma_0 = B \frac{h^2}{12} k^2 + \frac{B_s}{hk}.
\label{eq:sigma0}
\end{equation}

The dominant wavelength (wavevector) is the one minimizing the compression load.

\section{Elastic film in an infinite elastic medium}

$B$ and $B_s$ are constants, not $k$-dependant operators. We minimise $\sigma_0$:
\begin{equation}
0 = \frac{\partial \sigma_0}{\partial (kh)} = \frac{B}{6} hk -  \frac{B_s}{(hk)^2},
\end{equation}
which yields $hk = \left(\frac{6B_s}{B}\right)^{1/3}$ and thus the dominant wavelength is
\begin{equation}
\lambda_d = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
\label{eq:lambdaElEl}
\end{equation}

\section{Elastic film in an infinite viscous medium}

\citet{Biot1957} gives the modulus of the infinite viscous medium of viscosity $\eta$ as the operator
\begin{equation}
B_s = 4\eta\frac{\partial}{\partial t} = 4\frac{\eta}{\tau},
\end{equation}
if we look for an time dependent amplitude $A(t)= A_0 e^{t/\tau}$.

Still neither $B$ nor $B_s$ are $k$-dependant operators, so we can use directly Eq.~(\ref{eq:lambdaElEl})
\begin{equation}
\lambda_d = \pi h \left(\frac{B\tau}{3\eta}\right)^{1/3}.
\end{equation}

Since Eq.~(\ref{eq:mecheq}) is linear, the case of an elastic film on an semi-infinite viscous medium is dealt with by simply taking half of $B_s$.

\section{Elastic film on a viscous layer}
The layer thickness $H$ is supposed much smaller than $\lambda_d$, we are thus in the lubrication approximation.
\begin{equation}
\left \{
\begin{array}{c @{=} c}
	\frac{\partial p}{\partial z} & 0\\
	\frac{\partial^2 u}{\partial z^2} & \frac{1}{\eta}\frac{\partial p}{\partial x}
\end{array}
\right.,
\label{eq:lubrication}
\end{equation}
with $p$ the pressure in the fluid and $u$ the fluid velocity along $x$.

Since we work at small deflections, we can neglect the displacement of the plate along $x$ and set the boundary conditions to $u(z=0) = u(z=H) = 0$, and a double integration along $z$ of Eqs.~(\ref{eq:lubrication}) yields
\begin{equation}
u = \frac{1}{2\eta}\frac{\partial p}{\partial x}z(z-H).
\end{equation}

The flux along $x$ is thus of the Poiseuille form
\begin{equation}
Q = \int_0^H u dz = -\frac{H^3}{12\eta}\frac{\partial p}{\partial x}.
\label{eq:PoiseuilleFlux}
\end{equation}

Mass conservation writes
\begin{equation}
\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} = 0.
\label{eq:conservation}
\end{equation}

Keeping only linear order terms
\begin{equation}
12\eta\frac{\partial w}{\partial t} = 3H^2\frac{\partial w}{\partial x}\frac{\partial p}{\partial x} + H^3\frac{\partial^2 p}{\partial x^2} \approx H^3\frac{\partial^2 p}{\partial x^2}.
\label{eq:conservationLin}
\end{equation}

In operator language
\begin{equation}
\frac{12\eta}{\tau} \frac{w}{H^3k^2} = -q
\end{equation}
which allows to identify $B_s = \frac{12\eta}{\tau H^3k^3}$. Since $B_s$ depends on $k$, we come back to Eq.~(\ref{eq:sigma0}) that we differentiate:
\begin{equation}
0 = \frac{\partial\sigma_0}{\partial k}
 = \frac{Bh^2 k}{6} - \frac{48}{hH^3}\frac{\eta}{\tau}\frac{1}{k^5},
\end{equation}
which yields 
\begin{equation}
k_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
\end{equation}
with $\Upsilon = \frac{B\tau}{\eta}$ the viscoelastic factor. The dominant wavelength for Poiseuille model is thus
\begin{equation}
\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
\end{equation}

The scaling is then identical to the (much more involved) derivation by \citet{Huang2002}.

\section{Elastic film between two viscous layers}
We use once again the linearity of Eq.~(\ref{eq:mecheq}) to find the modulus of the effective substrate as 
\begin{equation}
B_s = \frac{12\eta}{\tau k^3} \left(\frac{1}{H_1^3}+\frac{1}{H_2^3}\right) = \frac{12\eta}{\tau H^3k^3}.
\end{equation}

We can use the results of the previous section using an effective $H$ so that
\begin{equation}
\frac{1}{H^3} = \frac{1}{H_1^3}+\frac{1}{H_2^3}.
\end{equation}

Note that $H$ goes to zero if any of the $H_1,H_2$ goes to zero.

\section{Porous elastic film with no possible flow}
The lateral load derives from the Darcy law through the film of permeability $\alpha$:
\begin{equation}
-q = p_2-p_1 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t} = \frac{\eta}{\tau}\frac{h}{\alpha k} \frac{\partial w}{\partial x},
\end{equation}
which allows to identify $B_s = \frac{\eta}{\tau}\frac{h}{\alpha k}$. Since $B_s$ depends on $k$, we come back to Eq.~(\ref{eq:sigma0}) that we differentiate:
\begin{equation}
0 = \frac{\partial\sigma_0}{\partial k}
 = \frac{Bh^2 k}{6} - \frac{2}{\alpha}\frac{\eta}{\tau}\frac{1}{k^3},
\end{equation}
which yields 
\begin{equation}
k_D^4 = \frac{12}{h^2\alpha\Upsilon}.
\end{equation}

The dominant wavelength for Darcy model is thus
\begin{equation}
\lambda_D = 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}.
\end{equation}

\section{Porous elastic film between two viscous layers}
We write mass conservation over the whole height of the cell
\begin{equation}
\frac{\partial Q_1}{\partial x} + \frac{\partial Q_2}{\partial x} = 0.
\end{equation}

Since we only consider sinusoidal variations $Q_1 + Q_2 = 0$. Using the expression of the Poiseuille flux Eq.~(\ref{eq:PoiseuilleFlux}), we obtain
\begin{equation}
H_1^3 \frac{\partial p_1}{\partial x} + H_2^3 \frac{\partial p_2}{\partial x} = 0,
\end{equation}
and thus
\begin{equation}
H_1^3 p_1 + H^2 p_2 = 0.
\label{eq:pressures}
\end{equation}

We now write the mass conservation in the lower viscous layer
\begin{equation}
\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} + v = 0,
\label{eq:conservationDarcy}
\end{equation}
which is the same as Eq.~(\ref{eq:conservation}) with an added leak $v$ due to the porosity. $v$ can be expressed by the Darcy law
\begin{equation}
v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}.
\label{eq:Darcy}
\end{equation}

Combining Eqs.~(\ref{eq:pressures}, \ref{eq:conservationDarcy}, \ref{eq:Darcy}) we obtain
\begin{align}
\frac{\partial w}{\partial t} &= \frac{H_1^3}{12\eta} \frac{\partial^2 p_1}{\partial x^2} - \frac{\alpha}{h\eta} \left(\frac{H_1}{H}\right)^3 p_1\\
\frac{\eta}{\tau}\left(\frac{H}{H_1}\right)^3 w &= -\left(\frac{H^3 k^2}{12} + \frac{\alpha}{h}\right)p_1.
\end{align}

The lateral load is
\begin{equation}
q = p_1-p_2 = \left(\frac{H_1}{H}\right)^3 p_1 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}},
\end{equation}
which allows to identify 
\begin{equation}
B_s = \frac{\eta}{\tau}\frac{1}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}}.
\end{equation}

Note that we recover a pure Poiseuille substrate when $\alpha \rightarrow 0$ and a pure Darcy substrate when $H \rightarrow 0$. Analytic minimisation of $\sigma_0$ knowing $B_s$ is possible but heavy. Instead we rewrite Eq.~(\ref{eq:sigma0}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
\begin{align}
\frac{12\sigma_0}{k_D^2 h^2 B} &= \left(\frac{k}{k_D}\right)^2 + \frac{1}{2\frac{k_D^2}{k_P^6}k^4  + \left(\frac{k}{k_D}\right)^2}\\
&= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}
\end{align}
and we minimise this expression numerically.

\bibliography{../../bib/Yaourt}

\end{document}

### 2014-11-03 model_mixte.tex
Présentation théorie ILM

\begin{frame}{Mechanical equilibrium of a plate}
\begin{columns}
\column{0.55\textwidth}
\begin{equation*}
\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q
\end{equation*}
\column{0.45\textwidth}
\begin{description}
\item[$\sigma_0$] compression load
\item[$q$] lateral load
\item[$w$] deflection
\end{description}
\end{columns}

\begin{columns}
\column{0.55\textwidth}
\begin{equation*}
B\frac{h^3}{12}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = -B_s \frac{\partial w}{\partial x}
\label{eq:mecheq}
\end{equation*}
\column{0.45\textwidth}
\begin{description}
\item[$B$] plate modulus
\item[$B_s$] substrate modulus
\end{description}
\end{columns}

\bigskip
\begin{columns}
\column{0.7\textwidth}
Sinusoidal deflection with a time exponential amplitude $w(x,t) = A_0 e^{t/\tau} \sin kx$
\column{0.3\textwidth}
\begin{equation*}
\sigma_0 = B \frac{h^2}{12} k^2 + \frac{B_s}{hk}
\label{eq:sigma0}
\end{equation*}
\end{columns}

\bigskip
The dominant wavelength (wavevector $k$) is the one minimizing the compression load.

\end{frame}

\begin{frame}{Elastic film in an infinite elastic medium}

$B$ and $B_s$ are constants, not $k$-dependant operators. We minimise $\sigma_0$:
\begin{equation*}
0 = \frac{\partial \sigma_0}{\partial (kh)} = \frac{B}{6} hk -  \frac{B_s}{(hk)^2}
\end{equation*}
which yields $hk = \left(\frac{6B_s}{B}\right)^{1/3}$ and thus the dominant wavelength is
\begin{equation*}
\lambda_d = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
\label{eq:lambdaElEl}
\end{equation*}

\end{frame}

\begin{frame}{Elastic film in an infinite viscous medium}

Biot (1957) gives the modulus of the infinite viscous medium of viscosity $\eta$ as the operator
\begin{equation*}
B_s = 4\eta\frac{\partial}{\partial t} = 4\frac{\eta}{\tau},
\end{equation*}
if we look for an time dependent amplitude $A(t)= A_0 e^{t/\tau}$.

Still neither $B$ nor $B_s$ are $k$-dependant operators
\begin{equation}
\lambda_d = \pi h \left(\frac{B\tau}{3\eta}\right)^{1/3}.
\end{equation}

Since the mechanical equilibrium equation is linear, the case of an elastic film on an semi-infinite viscous medium is dealt with by simply taking half of $B_s$.

\end{frame}

\begin{frame}{Elastic film on a viscous layer (Poiseuille)}
\begin{columns}
\column{0.65\textwidth}
\begin{itemize}
\item Lubrication approximation
\item Negligible $x$-displacement of the plate
\end{itemize}
\column{0.35\textwidth}
\begin{equation}
Q = -\frac{H^3}{12\eta}\frac{\partial p}{\partial x}
\label{eq:PoiseuilleFlux}
\end{equation}
\end{columns}

\begin{block}{Mass conservation}
\begin{equation*}
\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} = 0 \textcolor{Accent1}{\qquad\Rightarrow\qquad} 
12\eta\frac{\partial w}{\partial t} \approx H^3\frac{\partial^2 p}{\partial x^2}
\textcolor{Accent1}{\qquad\Rightarrow\qquad}
B_s = \frac{12\eta}{\tau H^3k^3}
\label{eq:conservationLin}
\end{equation*}
\end{block}

\begin{align*}
0 = \frac{\partial\sigma_0}{\partial k}
 = \frac{Bh^2 k}{6} - \frac{48}{hH^3}\frac{\eta}{\tau}\frac{1}{k^5}
 \textcolor{Accent1}{\qquad\Rightarrow\qquad} &
k_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon}\\
&\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
\end{align*}

\end{frame}

\begin{frame}{Elastic film between two viscous layers}
We use once again the linearity of Eq.~(\ref{eq:mecheq}) to find the modulus of the effective substrate as 
\begin{equation}
B_s = \frac{12\eta}{\tau k^3} \left(\frac{1}{H_1^3}+\frac{1}{H_2^3}\right) = \frac{12\eta}{\tau H^3k^3}.
\end{equation}

We can use the results of the previous section using an effective $H$ so that
\begin{equation}
\frac{1}{H^3} = \frac{1}{H_1^3}+\frac{1}{H_2^3}.
\end{equation}

Note that $H$ goes to zero if any of the $H_1,H_2$ goes to zero.

\end{frame}

\begin{frame}{Porous elastic film with no possible flow (Darcy)}
\begin{block}{Darcy law$\Rightarrow$ lateral load}
\begin{equation}
-q = p_2-p_1 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t} = \frac{\eta}{\tau}\frac{h}{\alpha k} \frac{\partial w}{\partial x} 
\textcolor{Accent1}{\qquad\Rightarrow\qquad} 
B_s = \frac{\eta}{\tau}\frac{h}{\alpha k}
\end{equation}
\end{block}

\begin{align*}
0 = \frac{\partial\sigma_0}{\partial k}
 = \frac{Bh^2 k}{6} - \frac{2}{\alpha}\frac{\eta}{\tau}\frac{1}{k^3} \textcolor{Accent1}{\qquad\Rightarrow\qquad} &
k_D^4 = \frac{12}{h^2\alpha\Upsilon}\\
& \lambda_D = 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}
\end{align*}

\end{frame}

\begin{frame}{Porous elastic film between two viscous layers}
\begin{block}{Mass conservation over the whole height of the cell}
\begin{equation*}
\frac{\partial Q_1}{\partial x} + \frac{\partial Q_2}{\partial x} = 0  \textcolor{Accent1}{\qquad\Rightarrow\qquad} 
\begin{array}{cc}
Q_1 + Q_2 = 0\\
H_1^3 p_1 + H_2^3 p_2 = 0
\end{array}
\end{equation*}
\end{block}

\begin{block}{Mass conservation in the lower viscous layer + Darcy}
\begin{equation*}
\frac{\partial Q_1}{\partial x} + \frac{\partial w}{\partial t} + v = 0
\qquad\text{with}
\qquad v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}
\label{eq:conservationDarcy}
\end{equation*}
\end{block}

\begin{equation*}
\text{\structure{lateral load}}\qquad q = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}}
\textcolor{Accent1}{\quad\Rightarrow\quad}
B_s = \frac{\eta}{\tau}\frac{1}{\frac{H^3 k^3}{12} + \frac{\alpha k}{h}}
\end{equation*}

\begin{description}
\item[$\alpha \rightarrow 0$] Poiseuille
\item[$H \rightarrow 0$] Darcy
\end{description}

\end{frame}
\begin{frame}{Poiseuille and Darcy}
Analytic minimisation of $\sigma_0$ knowing $B_s$ is possible but heavy.

Instead we express $\sigma_0$ in terms of previously calculated $\lambda_P$ and $\lambda_D$
\begin{equation*}
\frac{12\sigma_0}{k_D^2 h^2 B} 
= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}
\end{equation*}
and we minimise this expression numerically.

\end{frame}

### 2014-11-11 plis_figures.tex
Include in tex and generate figures.

--- 2014-04-07 plis_figures.tex
@@
-\documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage[utf8]{inputenc}
-
-%PNAS fonts
-\usepackage{mathptmx}
-\usepackage{tgheros}
-
-\usepackage{amssymb}
+\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4-1}
 \usepackage{siunitx}
-\usepackage{multirow}
-
-\usepackage{graphicx}
-
 \usepackage{tikz}
-\usetikzlibrary{calc, positioning, intersections, matrix, patterns}
-
-\tikzset{lab/.style={above right, inner sep=0, font=\sffamily\Large\bfseries}}
+\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap,fit}
+%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
 
 \usepackage{pgfplots}
+\pgfplotsset{compat=1.9}
 \usepgfplotslibrary{groupplots}
-
-\pgfplotsset{compat=1.3, every axis/.append style={thick, mark size={0.2em}}}
+\usepgfplotslibrary{external}
+\tikzsetexternalprefix{fig_plis/}
+\tikzexternalize
+\tikzset{external/force remake}
 
@@
 
+\begin{document}
 \author{Mathieu Leocmach}
-\begin{document}
 
 \begin{figure}
-	\begin{tikzpicture}
-	\begin{axis}[%
-		name=pH,
-		xmin=0,xmax=8,
-		ymin=0, ymax=7,
-		xlabel={time (h)}, ylabel={\textcolor{Accent1}{pH}},
-		extra tick style={grid=major},%
-		extra y ticks={4.6}, extra y tick labels={},%
-		%extra x ticks={0.55, 0.64, 0.72, 0.80, 1, 1.27, 2.5}, extra x tick labels={},%
-		axis y line*=left,
-		width=0.925\columnwidth,
-		height=0.5\columnwidth,
-		]
-	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}/3600.+0.05}]{Y189_28800s.pH};
-	%\node[base left=0] at (axis cs:8,4.6) {isoelectric};
-	\end{axis}
-	\begin{axis}[%
-		xmin=0,xmax=8,
-		ymin=0,
-		ylabel=\textcolor{Main}{$G^\prime$ (\si{\pascal})},
-		axis y line*=right,
-		axis x line=none,
-		width=0.925\columnwidth,
-		height=0.5\columnwidth,
-		]
-	\addplot+[no marks,Main] table[x expr={\thisrowno{0}/3600.+0.05}]{Y235_28800s.prise};
-	\end{axis}
-	\node[lab, yshift=0.5em] at (pH.outer south west) {a};
-	\begin{scope}[shift=(pH.outer south west), yshift=-2em]
-		\fill[pattern=north east lines,pattern color=Accent2] (0,0) rectangle (\columnwidth,1.5em) node[midway,fill=white,inner sep=1pt] (glabeltop) {glass};
-		\fill[pattern=north east lines,pattern color=Accent2] (0,-2.5em) rectangle (\columnwidth,-4em) node[midway,fill=white,inner sep=1pt] {glass};
-		%acrylamide
-		\begin{scope}[Accent1]
-			\draw[line width=2pt] (0.05\columnwidth,-2.5em) -- (0.95\columnwidth,-2.5em) (0.05\columnwidth,-1pt) -- (0.95\columnwidth,-1pt);
-			\node[inner sep=1pt, anchor=west] at (0.05\columnwidth,-1.25em) (brushlabel) {acrylamide brush};
-			\draw[->,thick] (brushlabel.base) -- +(0,-0.8em);
-			\draw[->,thick] (brushlabel.north) -- +(0,+0.4em);
-		\end{scope}
-		%spacer
-		\begin{scope}[gray]
-			\fill (0,0) rectangle (0.05\columnwidth,-2.5em) (\columnwidth,0) rectangle (0.95\columnwidth,-2.5em);
-			\node[above left=0] at (\columnwidth,1.5em) (spacerlabel){spacer};
-			\draw[->,thick] (spacerlabel.base) -- (0.975\columnwidth,1pt);
-		\end{scope}
-		%
-		\draw[<->] (0.75\columnwidth,-2pt) -- (0.75\columnwidth,-2.5em) node[midway,left] {$e\sim \SI{100}{\micro\metre}$};
-		\draw[<->] (0.05\columnwidth,-4.25em) -- (0.95\columnwidth,-4.25em) node[midway,below] (L) {$L\sim\SI{2}{\centi\metre}$};
-		\node[lab] at (0,0 |- L.base) {b};
-	\end{scope}
-	%pictures
-	\begin{scope}[inner sep=0, shift=(pH.outer south west), yshift=-8em]
-		\node[anchor=north west] (a) {\includegraphics[width=0.28\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom2_crop}};
-		\node[anchor=north] at (0.44\columnwidth,0) (b) {\includegraphics[width=0.28\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_zoom6_crop}};
-		\node[anchor=north east] at (\columnwidth,0) (c) {\includegraphics[width=0.4\columnwidth]{cas3p2_fluo0p8_GDL4_50um_coating_2_transmission}};
-		%zooms
-		\node[minimum width = 0.137\columnwidth, minimum height=0.137\columnwidth, anchor=north west, draw=Accent2] at ($(a.north west) +(0.111\columnwidth,-0.072\columnwidth)$) (bz){};
-		\draw[Accent2] (bz.north east) -- (b.north west) (bz.south east) -- (b.south west);
-		\node[minimum width = 0.128\columnwidth, minimum height=0.081\columnwidth, anchor=north west, draw=Main] at ($(b.north west) +(0.097\columnwidth,-0.143\columnwidth)$) (cz){};
-		\draw[Main] (cz.north east) -- (c.north west) (cz.south east) -- (c.south west);
-		\node[minimum width = 0.156\columnwidth, minimum height=0.156\columnwidth, anchor=north west, draw=Accent1] at ($(c.north west) + (0.125\columnwidth,0)$) (dz) {};
-		%scale bars
-		\draw[ultra thick] (a.south east) ++(0,-0.25em) -- ++(-0.178\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] (M) {\SI{1}{\centi\metre}};
-		\draw[ultra thick] (b.south east) ++(0,-0.25em) -- ++(-0.176\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{5}{\milli\metre}};
-		\draw[ultra thick] (c.south east) ++(0,-0.25em) -- ++(-0.132\columnwidth,0) node[pos=0.5, below=0.25em, font=\small] {\SI{1}{\milli\metre}};
-	\node[lab] at (0,0 |- M.base) {c};
-	\node[lab] at (b.west |- M.base) {d};
-	\node[lab] at (c.west |- M.base) {e};
-	\end{scope}
-	\end{tikzpicture}
-	\caption{Acid induced gel.
-(a) Evolution of the pH and of the storage modulus with sticky boundary conditions.
-Horizontal line is the isoelectric pH.
-(b) Sketch of the cell with no adhesion of the protein on the top and bottom plates.
-(c-e) Resulting patterns in the same sample ($e=\SI{100}{\micro\metre}$) at different magnification by (c-d) reflexion or (e) transmission optical microscopy (tiled).}
-	\label{fig:acidgel}
+	\tikzsetnextfilename{dynamics}
+	\input{fig_plis/dynamics.pgf}
+	\caption{Dynamics of pattern formation. Left: light transmission microscopy. Right: Reconstructed from fluorescent confocal microscopy. Scale bars are \SI{1}{\milli\metre}.}
+	\label{fig:dynamics}
+\end{figure}
+
+\begin{figure}
+	\tikzsetnextfilename{acidification}
+	\input{fig_plis/acidification.pgf}
+	\caption{Acid-induced protein gels properties behave non-monotonously with pH.
+(a) pH decrease in 4\%w sodium caseinate solution acidified by 4\%w GDL.
+Horizontal line indicates isoelectric pH of casen (b) Evolution of storage modulus measured in a rheometer (full adhesion to rotor and stator).
+(c) Evolution of pore size measured by confocal microscopy in a half coated microscopy cell (no adhesion to cell ceiling, syn\ae{}resis and swelling allowed).}
+	\label{fig:acidification}
+\end{figure}
+
+\begin{figure}
+	\tikzsetnextfilename{sideview}
+	\input{fig_plis/sideview.pgf}
+	\caption{Modelisation  and measurements.
+(a) Schematic side view of the cell.
+The constant thickness gel sheet (orange) is surrounded by water.
+Yellow lines show the path of transmitted light through the gel phase.
+A region of interest is highlighted in gray.
+(b) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
+Scale bar is \SI{100}{\micro\metre} (real size ratio).
+Evolution of (c) volume of the gel phase relative to cell volume and (d) excess area measured by confocal microscopy.}
+	\label{fig:sideview}
 \end{figure}
@@
 \begin{figure*}
-	\begin{tikzpicture}
-	\matrix[matrix of nodes, inner sep=0, column sep=0.015\textwidth, row sep=0.5em] (m){
-	33 min & 38 min & 43 min & 48 min & 1h & 1h15 & 2h30\\
-	\includegraphics[width=0.13\textwidth]{prise_0100_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0130_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0160_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0190_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0250_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0360_color.jpg}&
-	\includegraphics[width=0.13\textwidth]{prise_0799_color.jpg}\\
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t047_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t056_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t065_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t074_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t092_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t125_crop_resized.jpg}&
-	\includegraphics[width=0.13\textwidth]{cas3p2_fluo0p8_GDL4_2_t260_crop_resized.jpg}\\
-	};
-	\draw[ultra thick] ++(m-2-1.south west) -- ++(0.023\textwidth,0);
-	\draw[ultra thick] ++(m-3-1.south west) -- ++(0.1\textwidth,0);
-	\end{tikzpicture}
-	\caption{Dynamics of pattern formation for $e\approx\SI{100}{\micro\metre}$.
-(top) By light transmission microscopy.
-(bottom) Reconstructed from fluorescent confocal microscopy (corresponds to the squared area in Fig.~\ref{fig:acidgel}e).
-Scale bars are \SI{1}{\milli\metre}.}
-	\label{fig:dynamics}
+	\tikzsetnextfilename{Darcy_vs_Poiseuille}
+	\input{fig_plis/Darcy_vs_Poiseuille.pgf}
+	\caption{Comparing model predictions with measured wavelengths.
+Continuous line is the perfect match ($\lambda_{th}=\lambda_{xp}$), dashed line is the best linear fit through the origin (prefactor is 0.81 in a, 0.75 in b), dotted line is the best affine fit ($\lambda_{xp}=0.56\lambda_{th}+\SI{0.28}{\milli\metre}$).}
+	\label{fig:DarcyPoiseuille}
 \end{figure*}
 
-\begin{figure}
-	\begin{tikzpicture}
-	\matrix[matrix of nodes, inner sep=0, column sep=0.01\columnwidth, row sep=0.01\columnwidth] (m) {
-		\includegraphics[width=0.245\columnwidth]{pattern_50um.jpg}&
-		\includegraphics[width=0.245\columnwidth]{pattern_100um.jpg}&
-		\includegraphics[width=0.245\columnwidth]{pattern_250um.jpg}&
-		\includegraphics[width=0.245\columnwidth]{pattern_450um.jpg}\\
-	};
-	\node[inner sep=0, below left=0.01\columnwidth and 0 of m.south east] (x800) {\includegraphics[width=0.49\columnwidth]{pattern_800um_x1.jpg}};
-	\draw[ultra thick, Accent2, <->] (x800.south west) ++(0.196\columnwidth, 0.05\columnwidth) -- ++(0.145\columnwidth,0) node[midway,above]{$\lambda$};
-	\draw[ultra thick,Main](x800.north east) -- ++(-0.217\columnwidth,0) node[midway,below] {\SI{5}{\milli\metre}};
-
-	\begin{axis}[%
-		name=g,
-		anchor=left of north west,
-		at={($(m.south west)+(0,-0.01\columnwidth)$)},
-		scale only axis,
-		width=0.4\columnwidth,
-		height=0.175\columnwidth,
-		xmin=0,xmax=1,xlabel={gap (\si{\milli\metre})},
-		xtick={0,0.4,...,0.9},
-		ymin=0,ylabel={$\lambda$ (\si{\milli\metre})},
-		label shift=-0.5em,
-		cycle multi list={%
-			Accent1,no marks\\%
-			Accent1,only marks,every mark/.append style={fill=Accent1},mark=*\\%
-			Accent1,only marks,every mark/.append style={fill=white},mark=*\\%
-			Accent1,only marks,every mark/.append style={fill=Accent1},mark=triangle*\\%
-			Accent1,only marks,every mark/.append style={fill=white},mark=triangle*\\%
-			Accent1,only marks,every mark/.append style={fill=Accent1},mark=square*\\%
-			Accent1,only marks,every mark/.append style={fill=white},mark=square*\\%
-		}
-		]
-		\addplot+[domain=0:3] {4.23*x};
-		\addplot coordinates {(0.05,0.261)};\label{pgfplots:50}
-		\addplot coordinates {(0.1,0.458)};\label{pgfplots:100}
-		\addplot coordinates {(0.250,0.916)};\label{pgfplots:250}
-		\addplot coordinates {(0.450,2.075)};\label{pgfplots:450}
-		\addplot coordinates {(0.800,3.333)};\label{pgfplots:800}
-		%\addplot+[only marks,Accent1, mark options={fill=Accent1}] coordinates {(0.05,0.261) (0.1,0.458) (0.250,0.916) (0.450,2.075) (0.800,3.333)};
-	\end{axis}
-	\node[below right=0 of m-1-1.north west] {\ref{pgfplots:50}};
-	\node[below right=0 of m-1-2.north west] {\ref{pgfplots:100}};
-	\node[below right=0 of m-1-3.north west] {\ref{pgfplots:250}};
-	\node[below right=0 of m-1-4.north west] {\ref{pgfplots:450}};
-	\node[below right=0 of x800.north west] {\ref{pgfplots:800}};
-	\end{tikzpicture}
-	\caption{Cell thickness dependence of the patterns. Each transmission microscopy picture corresponds to a point on the graph. Scale bar is common for all pictures.}
-	\label{fig:thickness}
-\end{figure}
-
-\begin{figure}
-	\begin{tikzpicture}
-	\begin{groupplot}[%
-		group style={
-			group name=g, group size=1 by 2,
-			xticklabels at=edge bottom,
-			vertical sep=0,
-			},
-		xmin=0, xmax=200, xtick={0,30,...,180},
-		extra tick style={grid=major},%
-		extra x tick labels={},%
-		width=1.04\columnwidth,
-		height=0.5\columnwidth,
-		]
-	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, extra x ticks={12.55}]
-	\addplot+[no marks,Accent1] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8.txt};
-	\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};
-	\node[rotate=90, anchor=base east] at (axis cs:12.55,100) (syn) {synæresis};
-	\node[above right=0 of syn.south west] {swelling};
-	\node[lab,below left=0.2em] at (rel axis cs:1,1) {a};
-
-	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=1.2, xlabel={time (min)}, extra x ticks={26.7, 42.8}]
-	\addplot+[no marks,Accent1] table{excess_area_pc_half_cas8.txt};
-	\addplot+[only marks,Accent2, mark=+] table{excess_area_pc_half_cas8_toi.txt};
-	\node at (axis cs:13.35,1) {flat};
-	\node[rotate=90, anchor=east] at (axis cs:34.75,1) {wrinkling};
-	\node[anchor=west] at (axis cs:45,0.25) {wrinkling and buckling};
-	\node[lab,above left=0.2em] at (rel axis cs:1,0) (b) {b};
-	\end{groupplot}
-
-	\matrix[matrix of nodes, matrix anchor=north west, inner sep=0, column sep=0.02\textwidth, row sep=2pt] at (g c1r2.outer south west) (m){
-	Synæresis & Swelling\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t000.png}&
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t023.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t006.png}&
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t049.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t008.png}&
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t051.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t010.png}&
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t053.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t012.png}&
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t078.png}\\
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t014.png}&
-	\includegraphics[width=0.48\columnwidth]{coupe_av10_cas8_t088.png}\\
-	};
-	\draw[>->,gray,ultra thick] (m-2-1.east) -- (m-7-1.east) -- (m-2-2.west) -- (m-7-2.west);
-	\node[lab,anchor=base east] at (m-1-2.base -| b.east) {c};
-
-	%\node[inner sep=0, anchor=south east] (g-1-1.outer north west){\includegraphics[width=0.45\columnwidth]{plastic_confocal}};
-	\end{tikzpicture}
-	\caption{Synæresis and swelling, cas.
-8\%, GDL 7.2\% (a) Evolution of the volume of the gel phase.
-(b) Evolution of the midsurface area.
-(c) Confocal XZ slices of the cell (field of view is $1\times\SI{0.1}{\milli\metre}$), colors indicate the protein concentration.
-Times of the snapshots are marked as crosses on (a) and (b).
-Note that swelling starts \SI{14}{\minute} before wrinkling.
-The thickness increase is $\approx\SI{0.2}{px}$, thus not directly visible on the pictures.}
-	\label{fig:wrinkles}
-\end{figure}
-
 \end{document}

### 2014-11-11 fig_plis/article_plis.tex
Canevas pour l'article : figures, auteurs, affiliations et remerciements

--- 2014-11-11 plis_figures.tex
@@
 \documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4-1}
+\usepackage[utf8]{inputenc}
+
+\usepackage{graphicx}
+\usepackage{xcolor}
+\usepackage{hyperref}
 \usepackage{siunitx}
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap,fit}
-%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
-
-\usepackage{pgfplots}
-\pgfplotsset{compat=1.9}
-\usepgfplotslibrary{groupplots}
-\usepgfplotslibrary{external}
-\tikzsetexternalprefix{fig_plis/}
-\tikzexternalize
-\tikzset{external/force remake}
 
@@
 \begin{document}
+\title{Wrinkling of a porous layer}
 \author{Mathieu Leocmach}
+\email{mathieu.leocmach@ens-lyon.fr}
+\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
+\author{Mathieu Nespoulous}
+\affiliation{Aix-Marseille Université, CNRS, MADIREL UMR 7246, Marseille, France}
+\author{Sébastien Manneville}
+\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
+\author{Thomas Gibaud}
+\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
+
+\begin{abstract}
+
+\end{abstract}
+
+\maketitle
+
+\begin{acknowledgments}
+The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay and L.
+Mahadevan for theoretical insights at various sage of the research.
+Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
+Special thanks to Madame Biot who made M.A.
+Biot work accessible through poronet website.
+ML thank the Region Rhône Alpes and the Programme d'avenir Lyon- Saint Etienne (PALSE NoGELPo) for postoctoral grant.
+TG and MN acknowledge funding from ANR XXXX.
+ML and SM acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No.
+258803.
+\end{acknowledgments}
+
+\bibliographystyle{naturemag3}
 
 \begin{figure}
-	\tikzsetnextfilename{dynamics}
-	\input{fig_plis/dynamics.pgf}
-	\caption{Dynamics of pattern formation. Left: light transmission microscopy. Right: Reconstructed from fluorescent confocal microscopy. Scale bars are \SI{1}{\milli\metre}.}
+	\includegraphics{dynamics}
+	\caption{Dynamics of pattern formation.
+Left: light transmission microscopy.
+Part of the initial pattern is highlighted in yellow to show that it is not evolving.
+Right: 3D reconstruction from fluorescent confocal microscopy.
+Scale bars are \SI{1}{\milli\metre}.}
 	\label{fig:dynamics}
@@
 \begin{figure}
-	\tikzsetnextfilename{acidification}
-	\input{fig_plis/acidification.pgf}
+	\includegraphics{acidification}
 	\caption{Acid-induced protein gels properties behave non-monotonously with pH.
@@
 \begin{figure}
-	\tikzsetnextfilename{sideview}
-	\input{fig_plis/sideview.pgf}
+	\includegraphics{sideview}
 	\caption{Modelisation  and measurements.
@@
 \begin{figure*}
-	\tikzsetnextfilename{Darcy_vs_Poiseuille}
-	\input{fig_plis/Darcy_vs_Poiseuille.pgf}
+	\includegraphics{Darcy_vs_Poiseuille}
 	\caption{Comparing model predictions with measured wavelengths.

### 2014-11-12 fig_plis/article_plis.tex
Method summary Author contributions & information Alan Parker & Catherine Barentin URL of poronet

@@
 \usepackage{siunitx}
+%\usepackage{multirow}
 
@@
 \begin{document}
-\title{Wrinkling of a porous layer}
+\title{Wrinkling of a confined porous layer}
 \author{Mathieu Leocmach}
-\email{mathieu.leocmach@ens-lyon.fr}
 \affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
@@
 \author{Thomas Gibaud}
+\email{thomas.gibaud@ens-lyon.fr}
 \affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
@@
 
-\begin{acknowledgments}
-The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay and L.
-Mahadevan for theoretical insights at various sage of the research.
-Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
-Special thanks to Madame Biot who made M.A.
-Biot work accessible through poronet website.
-ML thank the Region Rhône Alpes and the Programme d'avenir Lyon- Saint Etienne (PALSE NoGELPo) for postoctoral grant.
-TG and MN acknowledge funding from ANR XXXX.
-ML and SM acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No.
-258803.
-\end{acknowledgments}
+\section*{Methods summary}
+To prevent adsorption of caseins to surfaces, microscope slides and cover-slips were cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush was polymerized from methacrylate groups.
+Sodium caseinate (Firmenich) was labelled with Dylight~550 NHS ESTER (Thermo Scientific) and then purified by centrifugation.
+Unlabelled sodium caseinate powder was dissolved in deionised water (or water+glycerol mixture).
+To induce gelation glucono-$\delta$-lactone (GDL) in powder (Firmenich) was added to the solution.
+\SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
+3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
+We used either 10x (air) lens to observe wrinkling or 63x (oil) lens to observe the microstructure of the gel.
+Larger scope pictures were obtained either by stitching fluorescent microscope data or by transmitted/reflected light macroscope (Nikkon SMZ745T/Leica DMS1000).
+Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
 
 \bibliographystyle{naturemag3}
+
+\paragraph*{Acknowledgements}
+Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
+The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay and L. Mahadevan for theoretical insights at various sage of the research and Alan Parker at Firmenich for providing the casein and GDL. 
+Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
+Catherine Barentin provided the (late) glass Couette cell of SM XXXX.
+Special thanks to Madame Biot who made M.A. Biot work accessible through \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 
+ML thank the Region Rhône Alpes and the Programme d'avenir Lyon - Saint Etienne (PALSE NoGELPo) for postoctoral grant. 
+TG and MN acknowledge funding from ANR XXXX. 
+ML and SM acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No. 258803.
+
+\paragraph*{Author Contributions}
+ML and MN conducted preliminary experiments.
+ML, SM and TG designed the experiments and interpreted the results.
+ML and TG performed quantitative experiments.
+ML performed data analysis and modelling.
+ML, SM and TG wrote the manuscript.
+
+\paragraph*{Author Information} 
+The authors declare that they have no competing financial interests. 
+Correspondence and requests for materials should be addressed to TG (\href{mailto:thomas.gibaud@ens-lyon.fr}{thomas.gibaud@ens-lyon.fr}).
 
@@
 Left: light transmission microscopy.
-Part of the initial pattern is highlighted in yellow to show that it is not evolving.
+The same area is highlighted in yellow on all panels to show that the initial pattern is not evolving.
 Right: 3D reconstruction from fluorescent confocal microscopy.
@@
 	\caption{Comparing model predictions with measured wavelengths.
-Continuous line is the perfect match ($\lambda_{th}=\lambda_{xp}$), dashed line is the best linear fit through the origin (prefactor is 0.81 in a, 0.75 in b), dotted line is the best affine fit ($\lambda_{xp}=0.56\lambda_{th}+\SI{0.28}{\milli\metre}$).}
+Dashed line is the best linear fit through the origin (prefactor is 0.81 in a, 0.65 in b), continuous line is the best affine fit ($\lambda_{xp}=0.56\lambda_{th}+\SI{0.28}{\milli\metre}$).}
 	\label{fig:DarcyPoiseuille}

### 2014-11-13 fig_plis/article_plis.tex
tab:data
permeability in nm^2
fig:mesh

@@
 
+\clearpage
+\newpage
+\setcounter{figure}{0}
+
+\section*{Supplemental material}
+
+\begin{table*}
+\begin{tabular}{SSSScSrrScSSSScSSSSSS}
+%\toprule
+\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v$} && {$\Upsilon$} & {$H^*/H$} & {$\lambda_{xp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
+{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.97 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
+4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 0.45 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
+4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 0.24 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
+4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 0.23 & 2.46 & 1.69 & 2.56 & 2.62 \\ 
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 0.14 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
+8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 0.05 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 2.66 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
+4 & 4 & 35 & 82.3 && 2.7 & 1858 & 13000 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 4.03 & 0.61 & 1.27 & 0.56 & 1.27 \\ 
+4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 0.34 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
+\end{tabular}
+\caption{Characteristics of the samples discussed in the main text}
+\label{tab:data}
+\end{table*}
+
+\begin{figure}
+	\includegraphics{mesh}
+	\caption{Gel microstructure.
+(a) Detail of a confocal micrograph of a 4\%w casein, 4\%w GDL in water, when only the slide is coated (not the cover slip).
+Time corresponds to the minimum in mesh size.
+Scale bar is \SI{10}{\micro\metre}.
+(b) Corresponding radial averaged Fourier transform.
+Black line is the best fit by Eq.XXX with $\xi=\SI{4.9}{\micro\metre}$ and $d=2.4$.
+Inset shows the center of the non radial average spectrum (arbitrary logarithmic color scale).
+Scale bar is \SI{10}{\per\micro\metre}.
+(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{1.5}{\micro\metre}$ and $d=2.5$.}
+	\label{fig:mesh}
+\end{figure}
+
 \end{document}

### 2014-11-13 plis_figures.tex
fig:mesh

@@
 
+\begin{figure}
+	\tikzsetnextfilename{mesh}
+	\input{fig_plis/mesh.pgf}
+	\caption{Gel microstructure.
+(a) Detail of a confocal micrograph of a 4\%w casein, 4\%w GDL in water, when only the slide is coated (not the cover slip).
+Time corresponds to the minimum in mesh size.
+Scale bar is \SI{10}{\micro\metre}.
+(b) Corresponding radial averaged Fourier transform.
+Black line is the best fit by Eq.XXX with $\xi=\SI{4.9}{\micro\metre}$ and $d=2.4$.
+Inset shows the center of the non radial average spectrum (arbitrary logarithmic color scale).
+Scale bar is \SI{10}{\per\micro\metre}.
+(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{1.5}{\micro\metre}$ and $d=2.5$.}
+	\label{fig:mesh}
+\end{figure}
+
 \end{document}

### 2014-11-14 fig_plis/article_plis.tex
permeability : Figure : switch tube 1 and 2 in the sketch, use gray instead of yellow, add a second graph for glycerol 50%, use figure* Supplementary method

@@
 
+\subsection*{Permeability measurements}
+\begin{figure*}
+	\includegraphics{permeability}
+	\caption{Permeability measurements.
+(a) Schematic representation of the experiment.
+(b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
+Black line is the best exponential fit $Ae^{-t/\tau}$.
+(b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
+(c) Idem in 50\%w glycerol, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
+	\label{fig:permeability}
+\end{figure*}
+
+To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}.
+First, the gel of interest is formed at the lower extremity of a thin glass tube.
+The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition).
+Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4.
+The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
+Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
+A webcam (Logitech Webcam Pro 9000) records the rise of the liquid in each tube.
+Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.
+
+Darcy law through the gel of permeability $\alpha$ and height $H$ writes
+\begin{equation}
+\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
+\end{equation}
+where $\Delta P$ is the the hydrostatic pressure relative to $z_\infty$. We can thus write
+\begin{equation}
+\frac{d(z_\infty-z)}{dt} = -\frac{z_\infty-z}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
+\end{equation}
+
+By fitting $z_\infty-z(t)$ by an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
+Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosities and low permeability.
+Obtaining $z_\infty$ without tube 2 would be time prohibitive.
+
 \end{document}

### 2014-11-14 fig_plis/article_plis.tex
fig:growth

@@
 
+\subsection*{Characteristic time measurement}
+\begin{figure}
+	\includegraphics{growth}
+	\caption{Growth of the instability.
+(a) Evolution of the gel velocity $v$, integrated over the whole surface.
+Inset: Evolution of the amplitude (peak to peak measurement) in semi logarithmic scale.
+Black line is the best exponential fit with $\tau=\SI{90}{\second}$.
+(b) Evolution of the bottom profile of the gel from the onset of instability (yellow) to the contact to both walls (brown).
+Note that the lateral size of the blister evolve in this regime.
+Thick black curve is the profile at much longer time.
+Note that blister foot (highlighted by the orange vertical line) has not moved since contact.}
+	\label{fig:growth}
+\end{figure}
+
+In most of our samples, it was possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
+However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both highest and lowest point of the surface.
+Even when valid, such local measurement is noisy and we prefer the measure of the vertical displacement of the gel $v$ that can be integrated over the whole field of view.
+
+We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls.
+We thus used consistently this approximation of $\tau$ for all our samples.
+This is coherent with the fact that the blister size $\approx\lambda/2$ does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is decided when $A(t) = A_\text{max}$.
+
 \end{document}

### 2014-11-14 plis_figures.tex
permeability : Figure : switch tube 1 and 2 in the sketch, use gray instead of yellow, add a second graph for glycerol 50%, use figure* Supplementary method
fig:growth

@@
 
+\begin{figure*}
+	\tikzsetnextfilename{permeability}
+	\input{fig_plis/permeability.pgf}
+	\caption{Permeability measurements.
+(a) Schematic representation of the experiment.
+(b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
+Black line is the best exponential fit $Ae^{-t/\tau}$.
+(b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
+(c) Idem in 50\%w glycerol, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
+	\label{fig:permeability}
+\end{figure*}
+
+\begin{figure}
+	\tikzsetnextfilename{growth}
+	\input{fig_plis/growth.pgf}
+	\caption{Growth of the instability.
+(a) Evolution of the gel velocity $v$, integrated over the whole surface.
+Inset: Evolution of the amplitude (peak to peak measurement) in semi logarithmic scale.
+Black line is the best exponential fit with $\tau=\SI{90}{\second}$.
+(b) Evolution of the bottom profile of the gel from the onset of instability (yellow) to the contact to both walls (brown).
+Note that the lateral size of the blister evolve in this regime.
+Thick black curve is the profile at much longer time.
+Note that blister foot (highlighted by the orange vertical line) has not moved since contact.}
+	\label{fig:growth}
+\end{figure}
+
 \end{document}

### 2014-11-15 fig_plis/article_plis.tex
confocal images segmentation

@@
 3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
-We used either 10x (air) lens to observe wrinkling or 63x (oil) lens to observe the microstructure of the gel.
+We used either 10x (air) lens to measure the geometry of wrinkling (see SM) or 63x (oil) lens to observe the microstructure of the gel.
 Larger scope pictures were obtained either by stitching fluorescent microscope data or by transmitted/reflected light macroscope (Nikkon SMZ745T/Leica DMS1000).
@@
 Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
-The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay and L. Mahadevan for theoretical insights at various sage of the research and Alan Parker at Firmenich for providing the casein and GDL. 
+The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay and L. Mahadevan for theoretical insights at various sage of the research. 
 Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
-Catherine Barentin provided the (late) glass Couette cell of SM XXXX.
+Alan Parker at Firmenich provided the casein and GDL; Catherine Barentin provided the (late) glass Couette cell of SM XXXX.
 Special thanks to Madame Biot who made M.A. Biot work accessible through \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 
@@
 
+\subsection*{Confocal images segmentation}
+
+At 10x magnification, we acquired sacks of (depending on cell thickness) 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
+Each confocal slice was $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
+At this low resolution details of the gel microstructure are undistinguishable ($\xi<$px for all compositions) and the gel appear as a continuous medium.
+
+We first denoise each plane by a small Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
+Then, for each $(x,y)$, we focus on the z-dependant intensity profile $I(z)$.
+We define the position of the top surface of the gel $z_\text{top}$ as where $I(z)$ first crosses half of its maximum value.
+Since the criterion is local, we do not need to correct for background intensity variation.
+In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.
+
+From these data, we obtain for each stack the volume and thickness of the gel phase, area of the midsurface and peak-to-peak amplitude.
+Velocity is obtained by time differentiation.
+Once the instability has saturated on floor and ceiling of the cell, peaks in the histograms of $z_\text{top}$, and $z_\text{bottom}$ yield the measure of the positions of the walls, thus to $e$ and retrospectively to $H_1$ and $H_2$.
+
 \subsection*{Characteristic time measurement}

### 2014-11-15 fig_plis/article_plis.tex
Thank O P-L Microstructure measurement

@@
 Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
-The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay and L. Mahadevan for theoretical insights at various sage of the research. 
+The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various sage of the research. 
 Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
@@
 
+\subsection*{Microstructure measurement}
+
 \begin{figure}
@@
 (b) Corresponding radial averaged Fourier transform.
-Black line is the best fit by Eq.XXX with $\xi=\SI{4.9}{\micro\metre}$ and $d=2.4$.
+Black line is the best fit by Eq.XXX with $\xi=\SI{4.9}{\micro\metre}$ and $\gamma=2.4$.
 Inset shows the center of the non radial average spectrum (arbitrary logarithmic color scale).
 Scale bar is \SI{10}{\per\micro\metre}.
-(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{1.5}{\micro\metre}$ and $d=2.5$.}
+(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{1.5}{\micro\metre}$ and $\gamma=2.5$.}
 	\label{fig:mesh}
 \end{figure}
+
+To follow the evolution of the microstructure in a gel undergoing syn\ae{}resis and swelling, we had to keep it in the thin focal plane of the 63x objective lens of the confocal microscope.
+To do so, we coated only the microscope slide with acrylamide brush to allow syn\ae{}resis on this side.
+The cover-slip was cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
+Neither buckling nor wrinkling was detected in such configuration.
+
+We acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (see Fig.~\ref{fig:mesh}a and c).
+After Hann windowing we performed Fourier transform (see insets of Fig.~\ref{fig:mesh}b and d) and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:mesh}b and d) that we fit by the following fractal form:
+\begin{equation}
+I(q) = \frac{\chi}{1+(\xi q)^\gamma}
+\end{equation}
+were the susceptibility $\chi$ is related to the contrast of fluorescence between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and the exponent $\gamma$ is loosely related to the fractal dimension.
 
@@
 We thus used consistently this approximation of $\tau$ for all our samples.
-This is coherent with the fact that the blister size $\approx\lambda/2$ does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is decided when $A(t) = A_\text{max}$.
+This is coherent with the fact that the blister size $\propto\lambda$ does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is decided when $A(t) = A_\text{max}$.
 

### 2014-11-25 fig_plis/article_plis.tex
Measured permeability of gly35
conrrection ech 4 après avoir trouvé le champ du confocal sur l'image macroscope
ajoute I(q,t) à fig:dynamics
Plis secondaires moyennées, avec barre d'erreur Utilise H/H* et non l'inverse

@@
 	\caption{Dynamics of pattern formation.
-Left: light transmission microscopy.
+(a) Light transmission microscopy.
 The same area is highlighted in yellow on all panels to show that the initial pattern is not evolving.
-Right: 3D reconstruction from fluorescent confocal microscopy.
+(b) 3D reconstruction from fluorescent confocal microscopy.
+(c) Time dependent Fourier spectrum of (a).
+First peak saturating around \SI{17}{\per\milli\metre} corresponds to the mean distance between two lines of the primary pattern, i.e.
+$\lambda/2$.
 Scale bars are \SI{1}{\milli\metre}.}
@@
 \cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v$} && {$\Upsilon$} & {$H^*/H$} & {$\lambda_{xp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
+{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_{xp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
 {\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
 \cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.97 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
-4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 0.45 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
-4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 0.24 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
-4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 0.23 & 2.46 & 1.69 & 2.56 & 2.62 \\ 
+%ech2
+4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 1.03 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
+& & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.18 & 0.45 & 1.00 & 0.39 & 1.00\\
+4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 2.23 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
+4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 4.25 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
+4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 4.36 & 1.62 & 1.69 & 2.56 & 2.62 \\
+& & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.17 & 0.81 & 1.37 & 0.6 & 1.37 \\
 \cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 0.14 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
-8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 0.05 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
+8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 7.23 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
+& & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.81 & 0.50 & 0.77 & 0.55 & 0.82\\
+8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 18.51 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
+& & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.61 & 0.75 & 0.82 & 0.59 & 0.83 \\
 \cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 2.66 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
-4 & 4 & 35 & 82.3 && 2.7 & 1858 & 13000 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 4.03 & 0.61 & 1.27 & 0.56 & 1.27 \\ 
-4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 0.34 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
+4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.38 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
+4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.24 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
+4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 2.94 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
 \end{tabular}
-\caption{Characteristics of the samples discussed in the main text}
+\caption{Characteristics of the samples discussed in the main text. Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.}
 \label{tab:data}

### 2014-11-25 fig_plis/article_plis.tex
Prefacteur de H* Inclu les données des cloques secondaires Différentie les points à H<H* et H>H*

@@
 	\caption{Comparing model predictions with measured wavelengths.
-Dashed line is the best linear fit through the origin (prefactor is 0.81 in a, 0.65 in b), continuous line is the best affine fit ($\lambda_{xp}=0.56\lambda_{th}+\SI{0.28}{\milli\metre}$).}
+Symbols are empty when $H<H^*$ (Darcy mode) and full when $H>H^*$ (Poiseuille mode).
+Orange dots come from primary pattern, yellow squares from secondary blisters.
+Dashed lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode, (b) in Poiseuille mode (c) all points.
+Prefactors are 0.63, 0.69 and 0.67 respectively.
+Continuous line is the best affine fit ($\lambda_{xp}=0.52\lambda_{th}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
 	\label{fig:DarcyPoiseuille}
@@
 %ech2
-4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 1.03 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
-& & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.18 & 0.45 & 1.00 & 0.39 & 1.00\\
-4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 2.23 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
-4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 4.25 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
-4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 4.36 & 1.62 & 1.69 & 2.56 & 2.62 \\
-& & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.17 & 0.81 & 1.37 & 0.6 & 1.37 \\
+4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
+& & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
+4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
+4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
+4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
+& & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
 \cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 7.23 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
-& & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.81 & 0.50 & 0.77 & 0.55 & 0.82\\
-8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 18.51 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
-& & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.61 & 0.75 & 0.82 & 0.59 & 0.83 \\
+8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
+& & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
+8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
+& & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
 \cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.38 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
-4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.24 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
-4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 2.94 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
+4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
+4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
+4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
 \end{tabular}

### 2014-12-04 fig_plis/article_plis.tex
fig:dynamics c BW
Dérivation du modèle mixte

@@
 We used either 10x (air) lens to measure the geometry of wrinkling (see SM) or 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures were obtained either by stitching fluorescent microscope data or by transmitted/reflected light macroscope (Nikkon SMZ745T/Leica DMS1000).
+Larger scope pictures, obtained either by stitching fluorescent microscope data or by transmitted/reflected light macroscope (Nikkon SMZ745T/Leica DMS1000), were used to measure the wavelength of the primary pattern.
+Secondary pattern wavelength is set to half of the primary wavelength, according to Fig.~\ref{fig:dynamics}c.
 Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
@@
 (b) 3D reconstruction from fluorescent confocal microscopy.
-(c) Time dependent Fourier spectrum of (a).
-First peak saturating around \SI{17}{\per\milli\metre} corresponds to the mean distance between two lines of the primary pattern, i.e.
-$\lambda/2$.
+(c) Fourier spectrum of binary altitudes of (a) corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at \SI{2}{\hour}.
 Scale bars are \SI{1}{\milli\metre}.}
@@
 	\includegraphics{sideview}
-	\caption{Modelisation  and measurements.
+	\caption{Model and measurements.
 (a) Schematic side view of the cell.
@@
 	\caption{Comparing model predictions with measured wavelengths.
-Symbols are empty when $H<H^*$ (Darcy mode) and full when $H>H^*$ (Poiseuille mode).
-Orange dots come from primary pattern, yellow squares from secondary blisters.
-Dashed lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode, (b) in Poiseuille mode (c) all points.
+Dots come from primary pattern, squares from secondary blisters.
+Dashed lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
 Prefactors are 0.63, 0.69 and 0.67 respectively.
@@
 
+\subsection*{Darcy+Poiseuille model}
+
+Mechanical equilibrium of a plate of thickness $h$, young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_0$ along $x$ and a lateral load $q$ along $z$ writes:
+\begin{equation}
+\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q,
+\label{eq:mecheq}
+\end{equation}
+where $w(x,t)$ is the deflection of the plate along $z$.
+
+We suppose an exponentially growing sinusoidal deflection $w(x,t) =  A_0 e^{t/\tau} \sin kx$ of characteristic time $\tau$ and wavevector $k$. In operator language, Eq.~\ref{eq:mecheq} writes:
+\begin{equation}
+\left(\frac{1}{12} B h^3 k^4 - \sigma_0 h k^2\right)w = q,
+\label{eq:mecheqop}
+\end{equation}
+where $B=E/(1-\nu^2)$ is the modulus of the plate.
+
+\subsubsection*{Elastic film in an infinite elastic medium}
+
+The deformation of the substrate of modulus $B_s$ gives the lateral load as $q = -k B_s w$, and thus the compression load writes
+\begin{equation}
+\sigma_0 = B \frac{h^2}{12} k^2 + \frac{B_s}{hk}.
+\label{eq:sigma0}
+\end{equation}
+
+Minimizing $\sigma_0$ function of $k$ yield the well known~\cite{Biot1957,Cerda2003} result $hk = \left(6B_s/B\right)^{1/3}$ and thus the dominant wavelength is
+\begin{equation}
+\lambda_d = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
+\label{eq:lambdaElEl}
+\end{equation}
+
+\subsubsection*{Elastic film on a viscous layer}
+The layer thickness $H$ is supposed much smaller than $\lambda_d$, we are thus in the lubrication approximation.
+Since we work at small deflections, we can neglect the displacement of the plate along $x$.
+The flux along $x$ is thus of the Poiseuille form
+\begin{equation}
+Q = -\frac{H^3}{12\eta}\frac{\partial p}{\partial x}.
+\label{eq:PoiseuilleFlux}
+\end{equation}
+where $p(x)$ is the pressure in the layer of viscosity $\eta$.
+
+Mass conservation writes
+\begin{equation}
+\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} = 0.
+\label{eq:conservation}
+\end{equation}
+Keeping only linear order terms, in operator language
+\begin{equation}
+\frac{12\eta}{\tau} \frac{w}{H^3k^2} = -q.
+\label{eq:PoiseuilleLoad}
+\end{equation}
+
+Combining Eqs.~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
+\begin{equation}
+\sigma_0 = \frac{1}{12}B h^2 k^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h k^4}
+\label{eq:sigma0P}
+\end{equation}
+and can be minimised by $k_P$ such that
+\begin{equation}
+k_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
+\label{eq:kP}
+\end{equation}
+with $\Upsilon = B\tau/\eta$ the viscoelastic factor. The dominant wavelength for Poiseuille model is thus
+\begin{equation}
+\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
+\end{equation}
+
+The scaling is then identical to the (much more involved) derivation by \citet{Huang2002}.
+When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of Eq.~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness
+\begin{equation}
+\frac{1}{H^3} = \frac{1}{H_1^3}+\frac{1}{H_2^3}.
+\end{equation}
+
+\subsubsection*{Porous elastic film with no possible Poiseuille flow}
+The lateral load derives from the Darcy law through the film of permeability $\alpha$:
+\begin{equation}
+-q = p_2-p_1 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
+\label{eq:DarcyLoad}
+\end{equation}
+
+Combining Eqs.~(\ref{eq:mecheqop}) and (\ref{eq:DarcyLoad}), the compression load writes
+\begin{equation}
+\sigma_0 = \frac{1}{12}B h^2 k^2 + \frac{\eta}{\tau}\frac{1}{\alpha k^2}
+\label{eq:sigma0D}
+\end{equation}
+and can be minimised by $k_D$ such that
+\begin{equation}
+k_D^4 = \frac{12}{h^2\alpha\Upsilon}.
+\label{eq:kD}
+\end{equation}
+
+The dominant wavelength for Darcy model is thus
+\begin{equation}
+\lambda_D = 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}.
+\end{equation}
+
+\subsubsection*{Porous elastic film between two viscous layers}
+The mass conservation over the whole height of the cell yield $Q_1 + Q_2 = 0$ and thus
+\begin{equation}
+H_1^3 p_1 + H_2^3 p_2 = 0.
+\label{eq:pressures}
+\end{equation}
+
+The mass conservation in the lower viscous layer is the same as Eq.~(\ref{eq:conservation}) with an added leak $v$ due to the porosity
+\begin{equation}
+\frac{\partial Q_1}{\partial x} + \frac{\partial w}{\partial t} + v = 0.
+\label{eq:conservationDarcy}
+\end{equation}
+
+$v$ can be expressed by the Darcy law
+\begin{equation}
+v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}.
+\label{eq:Darcy}
+\end{equation}
+
+Combining Eqs.~(\ref{eq:pressures}, \ref{eq:conservationDarcy}, \ref{eq:Darcy}) we obtain the lateral load
+\begin{equation}
+q = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}},
+\label{eq:MixtLoad}
+\end{equation}
+
+Note that we recover Eq~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and Eq~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
+We can recast this intuition in term of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $k_P$ and $k_D$:
+\begin{equation}
+H^* = 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}.
+\end{equation}
+
+Analytic minimisation of $\sigma_0$ knowing $q$ is possible but heavy. Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
+\begin{align}
+\frac{12\sigma_0}{k_D^2 h^2 B} &= \left(\frac{k}{k_D}\right)^2 + \frac{1}{2\frac{k_D^2}{k_P^6}k^4  + \left(\frac{k}{k_D}\right)^2}\\
+&= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}
+\end{align}
+and we minimise this expression numerically.
+
 \end{document}

### 2014-12-08 fig_plis/article_plis.tex
Describe figures 1, 2, 3, start model

@@
 \maketitle
+
+In the present work, we study the swelling-induced wrinkling of a porous gel sheet sandwiched between two viscous layers, with the whole system confined between two glass plates.
+As shown in Fig.~\ref{fig:dynamics}, the lateral confinement leads to saturation of the amplitude on the floor and ceiling followed by a cascade of secondary, ternary, etc.
+buckling.
+Such cascade buckling have been observed by~\cite{Roman1999} in a one dimensional macroscopic system, a sheet of steel forced to buckle.
+The 2D nature of our experiment creates nested patterns that could prove useful to design optical elements.
+In addition, one has to note that the primary wavelength is much smaller than the size of the system, implying that we are not in a simple buckling scenario but in a wrinkling instability.
+Our main result is to identify this selection mechanism.
+
+In the following, we will describe the physical chemistry that allows us to form such patterns.
+Then, we will model the situation and identify two possible selection mechanisms: either Darcy flow or Poiseuille flow.
+Finally we will compare our experimental data with the two models.
+Combining both Darcy and Poiseuille mechanisms will allow us to describe our data.
+
+The slow acidification of a protein (sodium caseinate) solution up to the isoelectric pH induces gelation.
+Since caseins adsorb strongly to interfaces, the resulting gel sticks strongly to container walls, allowing rheological measurements.
+As shown in Fig.~\ref{fig:acidification}, further decrease of the pH beyond the isoeletric point weakens the gel.
+This non monotonous behaviour of the elastic modulus translates in a non monotonous mesh size in a microscopy cell (slit geometry) coated so that the casein could not adsorb on the top wall (Fig.~\ref{fig:acidification}c), i.e.
+in syn\ae{}resis (solvent expulsion) followed by swelling.
+
+Figure~\ref{fig:sideview}b shows confocal measurements in the case where both top and bottom walls have been coated.
+Starting from an homogeneous solution (black margins correspond to the glass walls), the gel forms and simultaneously expels solvent on both sides.
+Since we have two clear interfaces between a protein rich phase and a protein poor (water) phase, we model the resulting situation as a porous elastic sheet of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ et $H_2$ (see Fig.~\ref{fig:sideview}a).
+Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in second panel of Fig.~\ref{fig:sideview}b.
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the sheet of gel (Fig.~\ref{fig:sideview}d).
+We can thus confirm that the instability is due to swelling.
+
+Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses while swelling.
+Also, the only situation were be observed wrinkling is when both top and bottom walls are coated.
+We confirmed that at any given time the thickness $h$ of the sheet of gel is homogeneous in $(x,y)$.
+The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is not due to thickness inhomogeneities but to altitude gradients.
+The length of the optical path (along $z$) through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}a.
+
+\section*{Scalings}
+
+The characteristic length $\lambda^*$ of a wrinkling pattern is set by the competition between the flexural modulus of the sheet $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a lateral load $q$~\cite{Vella2009,Kolinski2009}:
+\begin{equation}
+\lambda^* \sim h \left(\frac{B}{q}\right)^{1/3}.
+\label{eq:lstar}
+\end{equation}
+In small deflection actual wavelength $\lambda$ is related to the amplitude $A$ as $\lambda \sim A^{1/4} \lambda^{*3/4}$.
+
+We identify three possible candidates for the lateral load $q$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
+We confirmed that the former was playing no role by successfully repeating our experiment in a cell held vertical, i.e.
+with the weight acting longitudinally rather than laterally.
+
+In a situation where the sheet of gel would be sitting (not sticking) to the floor of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
+Darcy law relates the gradient of pressure (identified to $q/h$) to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{q}{h} = \frac{dA}{dt}$.
+Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
+\begin{equation}
+\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4}
+\end{equation}
+where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the sheet and the solvent at a characteristic time $\tau$.
+This expression stresses the kinetic nature of the wavelength selection.
+However we note that the wavelength of the primary pattern does not evolve with time (see Fig.~\ref{fig:dynamics} and Sup.
+Fig.~\ref{fig:growth}) once the gel is touching both walls, we can deduce the relevant $\tau$ from the growth of the primary pattern (see Supplementary Method for the precise definition).
+
+Conversely, in a situation where the sheet of gel would be impervious ($\alpha\rightarrow 0$), destabilisation over a wavelength $\lambda$ would create a lubrication ($H\ll\lambda$) flow in the viscous layers.
+The longitudinal flux $Q$ is related to the longitudinal pressure gradient $q/\lambda$ by a Poiseuille flow expression $Q = \frac{H^3}{12\eta} \frac{q}{\lambda}$.
+Mass conservation relates
 

### 2014-12-09 fig_plis/article_plis.tex
Poiseuille Model + Experimental tests

@@
 
+\section*{Pattern formation}
+
 The slow acidification of a protein (sodium caseinate) solution up to the isoelectric pH induces gelation.
@@
 This expression stresses the kinetic nature of the wavelength selection.
-However we note that the wavelength of the primary pattern does not evolve with time (see Fig.~\ref{fig:dynamics} and Sup.
-Fig.~\ref{fig:growth}) once the gel is touching both walls, we can deduce the relevant $\tau$ from the growth of the primary pattern (see Supplementary Method for the precise definition).
+However we note that the wavelength of (a generation of) the pattern does not evolve with time (see Fig.~\ref{fig:dynamics} and Sup.
+Fig.~\ref{fig:growth}) once the gel is touching both walls, we can deduce the relevant $\tau$ from the growth of the  pattern (see Supplementary Method for the precise definition).
 
-Conversely, in a situation where the sheet of gel would be impervious ($\alpha\rightarrow 0$), destabilisation over a wavelength $\lambda$ would create a lubrication ($H\ll\lambda$) flow in the viscous layers.
-The longitudinal flux $Q$ is related to the longitudinal pressure gradient $q/\lambda$ by a Poiseuille flow expression $Q = \frac{H^3}{12\eta} \frac{q}{\lambda}$.
-Mass conservation relates
+Conversely, in a the simple case where the sheet of gel is in the middle of the cell ($H_1=H_2$) and impervious ($\alpha\rightarrow 0$), destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
+By symmetry the lateral load across the gel $q$ is also the pressure difference over the wavelength.
+Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim q \frac{H^3}{\eta\lambda^2}$.
+Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
+\begin{equation}
+\lambda_P \sim (hH)^{1/2} \Upsilon^{1/6}
+\end{equation}
+where we recognise the scaling of the wrinkles of an elastic sheet on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural moduli of the film and substrate.
+Indeed, in lubrication or low Reynolds number one can consider a viscous film as elastic with effective moduli $\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
+
+A more complete derivation of the models can be found in the Supplementary Method and yield the following prefactors:
+\begin{align}
+\lambda_D &= 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}\\
+\lambda_P &= \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
+\end{align}
+where $1/H^3 = 1/H_1^3 + 1/H_2^3$. The limiting case between the two models is for $H = H^* \equiv 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}$.
+
+\section*{Experimental tests}
+
+The above models yield wavelengths of the same order in most of the cases we studied.
+Thus only systematic measurements of experimental $\lambda$ and all the parameters involved in either model allows to discriminate between them (see Supplementary Method for details on the measurements).
+As summarized in Supplementary Table~\ref{tab:data}, we changed the thickness of the cell, impacting $h$ and the maximal amplitude.
+We changed the composition and thus the stiffness of the gel and its permeability.
+We also changed systematically the viscosity of the solvent by adding glycerol, however this also impacted the gel stiffness, permeability and $\tau$.
+Finally, the only parameter we had no control on was the initial altitude of the gel before wrinklink ($H$) that to our knowledge varied randomly.
+However, we were able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
+Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
+Here the size of the blister is fixed as half of the size of the primary pattern as show on Fig.~\ref{fig:dynamics}c and $\tau$ is the variable parameter.
+
+Our results are displayed in Fig.~\ref{fig:DarcyPoiseuille}.
+Neither model alone is able to reproduce all the data.
+However considering only the points where $H<H^*$ (respectively $H>H^*$), the Darcy (resp.
+Poiseuille) model fits the data well albeit with a prefactor.
+Note that it would be easy to oversight the failure of the Poiseuille model at small wavelength by adding an unexplained offset.
+Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.
 

### 2014-12-09 fig_plis/article_plis.tex
reformulations et répétitions

@@
 In addition, one has to note that the primary wavelength is much smaller than the size of the system, implying that we are not in a simple buckling scenario but in a wrinkling instability.
-Our main result is to identify this selection mechanism.
+Our main goal is to identify the selection mechanism of the wavelength.
 
@@
 The slow acidification of a protein (sodium caseinate) solution up to the isoelectric pH induces gelation.
-Since caseins adsorb strongly to interfaces, the resulting gel sticks strongly to container walls, allowing rheological measurements.
+Since caseins adsorb strongly to interfaces, the resulting gel sticks to container walls, allowing rheological measurements.
 As shown in Fig.~\ref{fig:acidification}, further decrease of the pH beyond the isoeletric point weakens the gel.
@@
 
-Conversely, in a the simple case where the sheet of gel is in the middle of the cell ($H_1=H_2$) and impervious ($\alpha\rightarrow 0$), destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
+The opposite limiting case is when the sheet of gel is in the middle of the cell ($H_1=H_2$) and impervious ($\alpha\rightarrow 0$), destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
 By symmetry the lateral load across the gel $q$ is also the pressure difference over the wavelength.
@@
 where we recognise the scaling of the wrinkles of an elastic sheet on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural moduli of the film and substrate.
-Indeed, in lubrication or low Reynolds number one can consider a viscous film as elastic with effective moduli $\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
+Indeed, in lubrication or low Reynolds number one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
 

### 2014-12-09 fig_plis/article_plis.tex
Move figures around to have them in front of the corresponding text. Bibliography

@@
-\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4-1}
+\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,
+amsmath,amssymb,prl]{revtex4-1}
 \usepackage[utf8]{inputenc}
@@
 
+\section*{Introduction}
+
+\clearpage
+\newpage
+
+\begin{figure}
+	\includegraphics{dynamics}
+	\caption{Dynamics of pattern formation.
+(a) Light transmission microscopy.
+The same area is highlighted in yellow on all panels to show that the initial pattern is not evolving.
+(b) 3D reconstruction from fluorescent confocal microscopy.
+(c) Fourier spectrum of binary altitudes of (a) corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at \SI{2}{\hour}.
+Scale bars are \SI{1}{\milli\metre}.}
+	\label{fig:dynamics}
+\end{figure}
+
 In the present work, we study the swelling-induced wrinkling of a porous gel sheet sandwiched between two viscous layers, with the whole system confined between two glass plates.
@@
 
-The slow acidification of a protein (sodium caseinate) solution up to the isoelectric pH induces gelation.
-Since caseins adsorb strongly to interfaces, the resulting gel sticks to container walls, allowing rheological measurements.
-As shown in Fig.~\ref{fig:acidification}, further decrease of the pH beyond the isoeletric point weakens the gel.
+\begin{figure}[b]
+	\includegraphics{acidification}
+	\caption{Acid-induced protein gels properties behave non-monotonously with pH.
+(a) pH decrease in 4\%w sodium caseinate solution acidified by 4\%w GDL.
+Horizontal line indicates isoelectric pH of casen (b) Evolution of storage modulus measured in a rheometer (full adhesion to rotor and stator).
+(c) Evolution of pore size measured by confocal microscopy in a half coated microscopy cell (no adhesion to cell ceiling, syn\ae{}resis and swelling allowed).}
+	\label{fig:acidification}
+\end{figure}
+
+The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation.
+Since caseins adsorb strongly to interfaces, the resulting gel sticks to container walls, allowing rheological measurements (see Fig.~\ref{fig:acidification}b).
+Further decrease of the pH beyond the isoeletric point weakens the gel.
 This non monotonous behaviour of the elastic modulus translates in a non monotonous mesh size in a microscopy cell (slit geometry) coated so that the casein could not adsorb on the top wall (Fig.~\ref{fig:acidification}c), i.e.
@@
 
+\begin{figure}
+	\includegraphics{sideview}
+	\caption{Model and measurements.
+(a) Schematic side view of the cell.
+The constant thickness gel sheet (orange) is surrounded by water.
+Yellow lines show the path of transmitted light through the gel phase.
+A region of interest is highlighted in gray.
+(b) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
+Scale bar is \SI{100}{\micro\metre} (real size ratio).
+Evolution of (c) volume of the gel phase relative to cell volume and (d) excess area measured by confocal microscopy.}
+	\label{fig:sideview}
+\end{figure}
+
 \section*{Scalings}
@@
 with the weight acting longitudinally rather than laterally.
+
+\begin{figure*}
+	\includegraphics{Darcy_vs_Poiseuille}
+	\caption{Comparing model predictions with measured wavelengths.
+Dots come from primary pattern, squares from secondary blisters.
+Dashed lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
+Prefactors are 0.63, 0.69 and 0.67 respectively.
+Continuous line is the best affine fit ($\lambda_{xp}=0.52\lambda_{th}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
+	\label{fig:DarcyPoiseuille}
+\end{figure*}
 
@@
 \bibliographystyle{naturemag3}
+\bibliography{Yaourt}
 
@@
 Correspondence and requests for materials should be addressed to TG (\href{mailto:thomas.gibaud@ens-lyon.fr}{thomas.gibaud@ens-lyon.fr}).
-
-\begin{figure}
-	\includegraphics{dynamics}
-	\caption{Dynamics of pattern formation.
-(a) Light transmission microscopy.
-The same area is highlighted in yellow on all panels to show that the initial pattern is not evolving.
-(b) 3D reconstruction from fluorescent confocal microscopy.
-(c) Fourier spectrum of binary altitudes of (a) corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at \SI{2}{\hour}.
-Scale bars are \SI{1}{\milli\metre}.}
-	\label{fig:dynamics}
-\end{figure}
-
-\begin{figure}
-	\includegraphics{acidification}
-	\caption{Acid-induced protein gels properties behave non-monotonously with pH.
-(a) pH decrease in 4\%w sodium caseinate solution acidified by 4\%w GDL.
-Horizontal line indicates isoelectric pH of casen (b) Evolution of storage modulus measured in a rheometer (full adhesion to rotor and stator).
-(c) Evolution of pore size measured by confocal microscopy in a half coated microscopy cell (no adhesion to cell ceiling, syn\ae{}resis and swelling allowed).}
-	\label{fig:acidification}
-\end{figure}
-
-\begin{figure}
-	\includegraphics{sideview}
-	\caption{Model and measurements.
-(a) Schematic side view of the cell.
-The constant thickness gel sheet (orange) is surrounded by water.
-Yellow lines show the path of transmitted light through the gel phase.
-A region of interest is highlighted in gray.
-(b) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
-Scale bar is \SI{100}{\micro\metre} (real size ratio).
-Evolution of (c) volume of the gel phase relative to cell volume and (d) excess area measured by confocal microscopy.}
-	\label{fig:sideview}
-\end{figure}
-
-\begin{figure*}
-	\includegraphics{Darcy_vs_Poiseuille}
-	\caption{Comparing model predictions with measured wavelengths.
-Dots come from primary pattern, squares from secondary blisters.
-Dashed lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
-Prefactors are 0.63, 0.69 and 0.67 respectively.
-Continuous line is the best affine fit ($\lambda_{xp}=0.52\lambda_{th}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
-	\label{fig:DarcyPoiseuille}
-\end{figure*}
 

### 2014-12-16 2014_12_16_memo_fractal.tex
memo fractal dimension

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[francais]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{kpfonts}
\author{Mathieu Leocmach}
\title{Memo fractal dimension}
\date{16th December 2014}
\begin{document}

\maketitle

Thanks to Pierre Borgna, things are clearer.

A fractal in a 2D space has the pair correlation function
\begin{equation}
g(r) -1 \sim r^{d-2} e^{-\frac{r}{\xi}}
\end{equation}

To get the structure factor, we Fourier transform in 2D
\begin{align*}
S(q) -1 &\sim \iint e^{-\imath \vec{r}\cdot\vec{q}} \left(g(r)-1\right) \mathrm{d}^2\vec{r}\\
&\sim \int e^\frac{r}{\xi} r^{d-1} \mathrm{d}r
	\int_0^\pi e^{-\imath rq \cos \theta} \sin\theta \mathrm{d}\theta\\
&\sim \int e^\frac{r}{\xi} r^{d-1} 2\frac{\sin qr}{qr}\mathrm{d}r
\qquad\text{by variable change}\qquad X=cos\theta\\
\end{align*}

Rather than dig in the involved exact calculus, we just look for a power law in the regime where $q\xi \gg 1$, e.g. $r\gg \xi$ and thus $e^\frac{r}{\xi} \approx 1$.
\begin{align*}
S(\lambda q) -1 &\sim \int r^{d-1} 2\frac{\sin \lambda qr}{\lambda qr}\mathrm{d}r\\
&\sim \int \left(\frac{u}{\lambda}\right)^{d-1} 2\frac{\sin qu}{qu}\frac{\mathrm{d}u}{\lambda} \qquad\text{by variable change}\qquad u=\lambda r\\
&\sim \frac{1}{\lambda^d} \left(S(q) -1\right)
\end{align*}

The slope of $S_{2D}(q)$ is thus $-d_{2D}$. We can use the phenomenological law
\begin{equation}
S(q) = \frac{\chi}{1+(\xi q)^{d_{2D}}}
\end{equation}

Then we suppose that the fractal is isotropic, thus codimension is invariant $2-d_{2D} = 3-d_{3D}$ leading to $d_{3D} = d_{2D}+1$.

\end{document}

### 2014-12-17 2014_12_16_memo_fractal.tex
sign
With verified empirical expression

@@
 \title{Memo fractal dimension}
-\date{16th December 2014}
+\date{17th December 2014}
 \begin{document}
@@
 S(q) -1 &\sim \iint e^{-\imath \vec{r}\cdot\vec{q}} \left(g(r)-1\right) \mathrm{d}^2\vec{r}\\
-&\sim \int e^\frac{r}{\xi} r^{d-1} \mathrm{d}r
+&\sim \int e^{-\frac{r}{\xi}} r^{d-1} \mathrm{d}r
 	\int_0^\pi e^{-\imath rq \cos \theta} \sin\theta \mathrm{d}\theta\\
-&\sim \int e^\frac{r}{\xi} r^{d-1} 2\frac{\sin qr}{qr}\mathrm{d}r
-\qquad\text{by variable change}\qquad X=cos\theta\\
+&\sim \int e^{-\frac{r}{\xi}} r^{d-1} 2\frac{\sin qr}{qr}\mathrm{d}r
+\qquad\text{by variable change}\qquad X=cos\theta
+\addtocounter{equation}{1}\tag{\theequation}\label{eq:exact}\\
 \end{align*}
 
-Rather than dig in the involved exact calculus, we just look for a power law in the regime where $q\xi \gg 1$, e.g. $r\gg \xi$ and thus $e^\frac{r}{\xi} \approx 1$.
+Rather than dig in the involved exact calculus, we just look for a power law in the regime where $q\xi \gg 1$, thus $e^{-\frac{r}{\xi}} \approx 1$ in the integral.
 \begin{align*}
@@
 
-The slope of $S_{2D}(q)$ is thus $-d_{2D}$. We can use the phenomenological law
+The slope of $S_{2D}(q)$ is thus $-d_{2D}$. Then we suppose that the fractal is isotropic, thus codimension is invariant $2-d_{2D} = 3-d_{3D}$ leading to $d_{3D} = d_{2D}+1$.
+
+\begin{figure}
+\includegraphics{empirical_fractal}
+\caption{Testing empirical expression. Lines are numerical integration of (Eq.~\ref{eq:exact}) normalized by $2\Gamma(d)$ for $d=1,1.5,2$. Circles are the corresponding empirical expression.}
+\label{fig:empirical}
+\end{figure}
+
+As show in Fig.\ref{fig:empirical} Numerical integration of (Eq.~\ref{eq:exact}) for various $d$ is well fitted by the empirical law:
 \begin{equation}
-S(q) = \frac{\chi}{1+(\xi q)^{d_{2D}}}
+S(q) = \frac{2\chi\Gamma(d)}{\left(1+\left((d+1)\xi q\right)^2\right)^{d/2}}
 \end{equation}
-
-Then we suppose that the fractal is isotropic, thus codimension is invariant $2-d_{2D} = 3-d_{3D}$ leading to $d_{3D} = d_{2D}+1$.
+where $\chi$ is independent of $d$ and represents the relative contrast between the phases.
 

### 2014-12-17 2014_12_16_memo_fractal.tex
Generation of the graph by empirical_fractal.py Inverse circles and lines, less circles.

@@
 \includegraphics{empirical_fractal}
-\caption{Testing empirical expression. Lines are numerical integration of (Eq.~\ref{eq:exact}) normalized by $2\Gamma(d)$ for $d=1,1.5,2$. Circles are the corresponding empirical expression.}
+\caption{Testing empirical expression. Circles are numerical integration of (Eq.~\ref{eq:exact}) normalized by $2\Gamma(d)$ for $d=1,1.5,2$. Lines are the corresponding empirical expression.}
 \label{fig:empirical}

### 2014-12-17 empirical_fractal.py

@@ -1,0 +1,24 @@
+from matplotlib.pyplot import *
+from scipy.optimize import leastsq
+from scipy.integrate import quad
+from scipy.special import gamma
+
+
+f = lambda r,q,xi,d: 2 * r**(d-1)*np.exp(-r/xi)*np.sinc(q*r)
+
+F = np.vectorize(lambda q,xi,d: quad(f, 0, np.inf, args=(q,xi,d))[0])
+
+qs = np.logspace(-2,2,25)
+
+figure('Empirical', figsize=(4.875, 3.25))
+
+for d,c in zip([1,1.5,2], 'rgb'): 
+    plot(qs, 1/(1+((d+1)*qs)**2)**(0.5*d), color=c)
+    plot(qs, F(qs, 1, d)/(2*gamma(d)), 'o', color=c)
+
+gca().set_position([0.16, 0.16, 0.76, 0.76])
+
+xscale('log'); yscale('log')
+xlabel(r'$\xi q$')
+ylabel(r'$S(q)/(2\chi\Gamma(d))$')
+savefig('empirical_fractal.pdf')

### 2014-12-17 fig_plis/article_plis.tex
fig:mesh update caption

@@
 (b) Corresponding radial averaged Fourier transform.
-Black line is the best fit by Eq.XXX with $\xi=\SI{4.9}{\micro\metre}$ and $\gamma=2.4$.
+Black line is the best fit by Eq.XXX with $\xi=\SI{2.0}{\micro\metre}$ and $d_{3D}=2.27$.
 Inset shows the center of the non radial average spectrum (arbitrary logarithmic color scale).
 Scale bar is \SI{10}{\per\micro\metre}.
-(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{1.5}{\micro\metre}$ and $\gamma=2.5$.}
+(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{0.3}{\micro\metre}$ and $d_{3D}=3$.}
 	\label{fig:mesh}

### 2014-12-17 plis_figures.tex
fig:mesh update caption

@@
 (b) Corresponding radial averaged Fourier transform.
-Black line is the best fit by Eq.XXX with $\xi=\SI{4.9}{\micro\metre}$ and $d=2.4$.
+Black line is the best fit by Eq.XXX with $\xi=\SI{2.0}{\micro\metre}$ and $d_{3D}=2.27$.
 Inset shows the center of the non radial average spectrum (arbitrary logarithmic color scale).
 Scale bar is \SI{10}{\per\micro\metre}.
-(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{1.5}{\micro\metre}$ and $d=2.5$.}
+(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{0.3}{\micro\metre}$ and $d_{3D}=3$.}
 	\label{fig:mesh}

### 2014-12-19 fig_plis/article_plis.tex
Disambuguation of q. Lateral load is now \sigma_\perp. Coherently, \sigma_0 is now \sigma_\parallel

@@
 
-The characteristic length $\lambda^*$ of a wrinkling pattern is set by the competition between the flexural modulus of the sheet $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a lateral load $q$~\cite{Vella2009,Kolinski2009}:
+The characteristic length $\lambda^*$ of a wrinkling pattern is set by the competition between the flexural modulus of the sheet $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a lateral load $\sigma_\perp$~\cite{Vella2009,Kolinski2009}:
 \begin{equation}
-\lambda^* \sim h \left(\frac{B}{q}\right)^{1/3}.
+\lambda^* \sim h \left(\frac{B}{\sigma_\perp}\right)^{1/3}.
 \label{eq:lstar}
@@
 
-We identify three possible candidates for the lateral load $q$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
+We identify three possible candidates for the lateral load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
 We confirmed that the former was playing no role by successfully repeating our experiment in a cell held vertical, i.e.
@@
 In a situation where the sheet of gel would be sitting (not sticking) to the floor of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
-Darcy law relates the gradient of pressure (identified to $q/h$) to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{q}{h} = \frac{dA}{dt}$.
+Darcy law relates the gradient of pressure (identified to $\sigma_\perp/h$) to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
@@
 The opposite limiting case is when the sheet of gel is in the middle of the cell ($H_1=H_2$) and impervious ($\alpha\rightarrow 0$), destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
-By symmetry the lateral load across the gel $q$ is also the pressure difference over the wavelength.
-Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim q \frac{H^3}{\eta\lambda^2}$.
+By symmetry the lateral load across the gel $\sigma_\perp$ is also the pressure difference over the wavelength.
+Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_\perp \frac{H^3}{\eta\lambda^2}$.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
@@
 
-Mechanical equilibrium of a plate of thickness $h$, young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_0$ along $x$ and a lateral load $q$ along $z$ writes:
+Mechanical equilibrium of a plate of thickness $h$, young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a lateral load $\sigma_\perp$ along $z$ writes:
 \begin{equation}
-\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q,
+\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_\parallel h \frac{\partial^2 w}{\partial x^2} = \sigma_\perp,
 \label{eq:mecheq}
@@
 \begin{equation}
-\left(\frac{1}{12} B h^3 k^4 - \sigma_0 h k^2\right)w = q,
+\left(\frac{1}{12} B h^3 k^4 - \sigma_\parallel h k^2\right)w = \sigma_\perp,
 \label{eq:mecheqop}
@@
 
-The deformation of the substrate of modulus $B_s$ gives the lateral load as $q = -k B_s w$, and thus the compression load writes
+The deformation of the substrate of modulus $B_s$ gives the lateral load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
 \begin{equation}
-\sigma_0 = B \frac{h^2}{12} k^2 + \frac{B_s}{hk}.
+\sigma_\parallel = B \frac{h^2}{12} k^2 + \frac{B_s}{hk}.
 \label{eq:sigma0}
@@
 
-Minimizing $\sigma_0$ function of $k$ yield the well known~\cite{Biot1957,Cerda2003} result $hk = \left(6B_s/B\right)^{1/3}$ and thus the dominant wavelength is
+Minimizing $\sigma_\parallel$ function of $k$ yield the well known~\cite{Biot1957,Cerda2003} result $hk = \left(6B_s/B\right)^{1/3}$ and thus the dominant wavelength is
 \begin{equation}
@@
 \begin{equation}
-\frac{12\eta}{\tau} \frac{w}{H^3k^2} = -q.
+\frac{12\eta}{\tau} \frac{w}{H^3k^2} = -\sigma_\perp.
 \label{eq:PoiseuilleLoad}
@@
 \begin{equation}
-\sigma_0 = \frac{1}{12}B h^2 k^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h k^4}
+\sigma_\parallel = \frac{1}{12}B h^2 k^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h k^4}
 \label{eq:sigma0P}
@@
 \begin{equation}
--q = p_2-p_1 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
+-\sigma_\perp = p_2-p_1 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
 \label{eq:DarcyLoad}
@@
 \begin{equation}
-\sigma_0 = \frac{1}{12}B h^2 k^2 + \frac{\eta}{\tau}\frac{1}{\alpha k^2}
+\sigma_\parallel = \frac{1}{12}B h^2 k^2 + \frac{\eta}{\tau}\frac{1}{\alpha k^2}
 \label{eq:sigma0D}
@@
 \begin{equation}
-q = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}},
+\sigma_\perp = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}},
 \label{eq:MixtLoad}
@@
 
-Analytic minimisation of $\sigma_0$ knowing $q$ is possible but heavy. Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
+Analytic minimisation of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but heavy. Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
 \begin{align}
-\frac{12\sigma_0}{k_D^2 h^2 B} &= \left(\frac{k}{k_D}\right)^2 + \frac{1}{2\frac{k_D^2}{k_P^6}k^4  + \left(\frac{k}{k_D}\right)^2}\\
+\frac{12\sigma_\parallel}{k_D^2 h^2 B} &= \left(\frac{k}{k_D}\right)^2 + \frac{1}{2\frac{k_D^2}{k_P^6}k^4  + \left(\frac{k}{k_D}\right)^2}\\
 &= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}

### 2015-01-09 fig_plis/article_plis.tex
Ajoute une intro

--- 2014-12-09 fig_plis/article_plis.tex
@@
 \begin{document}
-\title{Wrinkling of a confined porous layer}
+\title{Hierarchical wrinkling in a permeable and confined biogel}
 \author{Mathieu Leocmach}
@@
 \section*{Introduction}
+There are many ways and reasons for a film to wrinkle.
+An elastic film buckles due to excess area with respect to its boundaries.
+Wrinkling is a buckling hindered by a substrate, the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength, as discovered originally in aeronautics [Gough1940, Bijlaard1946].
+On the engineering side, the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear [Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013] or non-linear wrinkling [Efimenko2005, Guvendiren2010, Kim2011a, Brau2011].
+Such patterns can be triggered by temperature dilation, swelling (eventually pH or light induced) or the removal of pre-strain.
+On the biology side, wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles, [bisset1987, genzer2005], difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts [savin2001, ciarletta2014, shyer2013], localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern [trejo2013, asally2012].
+On the physics side, analogy with wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching [Cerda2003], gravity [Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013] (linked to the pioneering work of Smoluchowski [Smoluchowski1927]), boundaries [Vandeparre2011a, Li2013] or adhesion [Vella2009a].
+Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity [Biot1957] or poroelasticity [Biot1964], leading to time-dependent kinetic wavelength selection.
+Stability analysis of a film on a thin viscous substrate was studied more recently [Huang2002] and remains theoretical since on a free interface gravity dominates over viscous substrate.
 
-\clearpage
-\newpage
+Yet, in a biological context, membranes are immersed in a nearly buoyant medium (lymph, blood, mucus), a practical situation that could lead to wrinkling on a thin viscous substrate.
+Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling [Ma2004,Longley2013].
+Such interplay is also at stake in daily life problems such as wrinkles encountered while putting up wall paper [pacione1991].
+
+Here, we have developed permeable model bio-gels that produce hierarchical wrinkling as a result of the interplay between its gelation dynamics and the confinement conditions.
+In slip geometry, we slowly acidify a dispersion of casein, one of the milk proteins.
+As a result, not only did the casein form a gel as expected [leocmach2014] --this is a well know effect used in cheese or yogurt making, but, the gel started to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
+Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the casein network spontaneously shrink and swell at the micron level and wrinkles at the millimeter scale.
+We developed a model than pin down the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength, Poiseuille [poiseuille] flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy [darcy] flows of the solvent through the gel layer for short wavelengths.
 

### 2015-01-09 fig_plis/article_plis.tex
Fusionne les modifications q->sigma_perp effacées à la révision précédente Syntaxe des citations Notes et remarques de la réunion intro du jour: - membrane -> thin tissues (epithelimum, endothelium)

--- 2014-12-19 fig_plis/article_plis.tex
@@
 \section*{Introduction}
+There are many ways and reasons for a film to wrinkle.
+An elastic film buckles due to excess area with respect to its boundaries.
+Wrinkling is a buckling hindered by a substrate, the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength, as discovered originally in aeronautics~\cite{Gough1940, Bijlaard1946}.
+On the engineering side, the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or non-linear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
+Such patterns can be triggered by temperature dilation, swelling (eventually pH or light induced) or the removal of pre-strain.
+On the biology side, wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{bisset1987, genzer2005}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{savin2001, ciarletta2014, shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{trejo2013, asally2012}.
+On the physics side, analogy with wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013} (linked to the pioneering work of Smoluchowski~\cite{Smoluchowski1927}), boundaries~\cite{Vandeparre2011a, Li2013} or adhesion~\cite{Vella2009a}.
+Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
+Stability analysis of a film on a thin viscous substrate was studied more recently~\cite{Huang2002} and remains theoretical since on a free interface gravity dominates over viscous substrate.
 
-\clearpage
-\newpage
+Yet, in a biological context, thin tissues (epithelimum, endothelium) are immersed in a nearly buoyant medium (lymph, blood, mucus), a practical situation that could lead to wrinkling on a thin viscous substrate.
+Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling [Ma2004,Longley2013].
+Such interplay is also at stake in daily life problems such as wrinkles encountered while putting up wall paper [pacione1991].
+
+Here, we have introduced permeable model bio-gels that produce hierarchical wrinkling as a result of the interplay between its gelation dynamics and the confinement conditions.
+In slit geometry, we slowly acidify a dispersion of casein, one of the milk proteins.
+As a result, not only does the casein form a gel as expected~\cite{leocmach2014} --this is a well know effect used in cheese or yogurt making, but, the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
+Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the casein network spontaneously shrink and swell at the micron level and wrinkles at the millimeter scale.
+We developed a model than pin down the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength, Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel layer for short wavelengths.
 
@@
 
+% § à déplacer après \section*{Pattern formation}, éventuellement en fusionnant des éléments avec le paragraphe précédent
 In the present work, we study the swelling-induced wrinkling of a porous gel sheet sandwiched between two viscous layers, with the whole system confined between two glass plates.
@@
 
+% § à supprimer, ou à fusionner avec l'annonce de plan
 In the following, we will describe the physical chemistry that allows us to form such patterns.
@@
 \end{figure}
+
+%discuter la figure 1c
 
@@
 
+% Pour Mathieu : ajouter quelque part des applications numériques pour les trois longueurs d'ondes gravité, Darcy, Poiseuille
 We identify three possible candidates for the lateral load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
@@
 Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.
+
+%Conclusion à écrire
 

### 2015-01-09 fig_plis/article_plis.tex
Fix citations

@@
 Such patterns can be triggered by temperature dilation, swelling (eventually pH or light induced) or the removal of pre-strain.
-On the biology side, wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{bisset1987, genzer2005}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{savin2001, ciarletta2014, shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{trejo2013, asally2012}.
+On the biology side, wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
 On the physics side, analogy with wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013} (linked to the pioneering work of Smoluchowski~\cite{Smoluchowski1927}), boundaries~\cite{Vandeparre2011a, Li2013} or adhesion~\cite{Vella2009a}.
@@
 Yet, in a biological context, thin tissues (epithelimum, endothelium) are immersed in a nearly buoyant medium (lymph, blood, mucus), a practical situation that could lead to wrinkling on a thin viscous substrate.
-Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling [Ma2004,Longley2013].
-Such interplay is also at stake in daily life problems such as wrinkles encountered while putting up wall paper [pacione1991].
+Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
+Such interplay is also at stake in daily life problems such as wrinkles encountered while putting up wall paper~\cite{Pacione1991}.
 
@@
 In slit geometry, we slowly acidify a dispersion of casein, one of the milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{leocmach2014} --this is a well know effect used in cheese or yogurt making, but, the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
+As a result, not only does the casein form a gel as expected~\cite{Leocmach2014} --this is a well know effect used in cheese or yogurt making, but, the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
 Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the casein network spontaneously shrink and swell at the micron level and wrinkles at the millimeter scale.

### 2015-01-09 seminaire_Montpellier.tex
Séminaire à Montpellier, parcourt colloides

--- 2014-09-18 seminaire_Tanakalab.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[francais]{babel}
+\usepackage[british]{babel}
+\usepackage[super]{nth}
 %\usetheme{Boadilla}
@@
 \usepackage{tikz}
-\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, shapes.geometric, matrix,patterns,mindmap}
+\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap}
 %\usetikzlibrary{positioning, patterns,topaths,chains,matrix}
@@
 \author[M. Leocmach]{Mathieu Leocmach}
-\date{1st June 2014}
+\date{20 October 2014}
 \titlegraphic{
@@
 
+\setcounter{framenumber}{0}
+
+\include{parcourt}
+
 \begin{frame}[plain]
@@
 \end{frame}
-
-\setcounter{framenumber}{0}
 
@@
 \column{0.5\textwidth}
+\tikzsetnextfilename{gelation_gdl1}
 \begin{tikzpicture}
@@
 	\end{groupplot}
-	\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
+	%\draw[<-] (g c1r2.south west) ++(0.5em,0) -- +(0,-1em) node[below,font=\footnotesize] {casein ``micelles''};
 \end{tikzpicture}
@@
 \[\textcolor{Accent1}{\text{storage}\rightarrow} G^\prime + \imath G^{\prime\prime} \textcolor{Accent2}{\leftarrow\text{loss}}\]
+\tikzsetnextfilename{powerlaw}
 \begin{tikzpicture}

### 2015-01-20 plis_figures.tex
Schéma en (d) Réorganise les notations pour plus de clarté

@@
 	\caption{Modelisation  and measurements.
-(a) Schematic side view of the cell.
+(a) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
+Scale bar is \SI{100}{\micro\metre} (real size ratio).
+Evolution of (b) volume of the gel phase relative to cell volume and (c) excess area measured by confocal microscopy.
+Crosses correspond to times of (a) (d) Schematic side view of the cell.
 The constant thickness gel sheet (orange) is surrounded by water.
 Yellow lines show the path of transmitted light through the gel phase.
-A region of interest is highlighted in gray.
-(b) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
-Scale bar is \SI{100}{\micro\metre} (real size ratio).
-Evolution of (c) volume of the gel phase relative to cell volume and (d) excess area measured by confocal microscopy.}
+A region of interest is highlighted in gray.}
 	\label{fig:sideview}

### 2015-01-20 fig_plis/article_plis.tex
Modifs Seb

@@
 %\usepackage{multirow}
+\usepackage{bm}
+\usepackage{dcolumn,ulem}
+\usepackage{amsfonts}
 
@@
 
+\newcommand{\seb}[1]{\textbf{\color{blue}#1}} % remarques Seb
+\newcommand{\sseb}[1]{\sout{\textbf{\color{blue}#1}}} % remarques Seb
+
 \begin{document}
-\title{Hierarchical wrinkling in a permeable and confined biogel}
+\title{Hierarchical wrinkling in a confined permeable biogel}
 \author{Mathieu Leocmach}
@@
 \begin{abstract}
-
+\seb{\it Ajouter l'abstract.}
 \end{abstract}
@@
 \section*{Introduction}
+
+\seb{\it Terminologie generale a uniformiser dans tout l'article: sheet $\rightarrow$ layer, floor and ceiling $\rightarrow$ bottom and top walls, impervious $\rightarrow$ impermeable, non-monotonous $\rightarrow$ nonmonotonic, lateral load $\rightarrow$ transverse load, longitudinal/compression load $\rightarrow$ compressional load, laterally $\rightarrow$ transversely.
+A discuter puis a changer partout systematiquement ?
+De facon generale, il reste encore beaucoup de choses entre parentheses: soit c'est utile et on enleve les parentheses, soit c'est inutile et on supprime ?}
+
 There are many ways and reasons for a film to wrinkle.
 An elastic film buckles due to excess area with respect to its boundaries.
-Wrinkling is a buckling hindered by a substrate, the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength, as discovered originally in aeronautics~\cite{Gough1940, Bijlaard1946}.
-On the engineering side, the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or non-linear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
+Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength \sseb{as discovered originally in aeronautics} \cite{Gough1940, Bijlaard1946}.
+On the engineering side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
 Such patterns can be triggered by temperature dilation, swelling (eventually pH or light induced) or the removal of pre-strain.
-On the biology side, wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
-On the physics side, analogy with wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013} (linked to the pioneering work of Smoluchowski~\cite{Smoluchowski1927}), boundaries~\cite{Vandeparre2011a, Li2013} or adhesion~\cite{Vella2009a}.
+On the biology side wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
+On the physics side analogues of wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013} \sseb{(linked to the pioneering work of Smoluchowski~\cite{Smoluchowski1927})} \seb{\it utile?}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Stability analysis of a film on a thin viscous substrate was studied more recently~\cite{Huang2002} and remains theoretical since on a free interface gravity dominates over viscous substrate.
+Stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002} \seb{and to date this confined situation remains without experimental equivalent.} \sseb{as for a free interface gravity dominates}
+\sseb{ over viscous substrate} \seb{Pas tr\`es clair, \`a reformuler~?}
 
-Yet, in a biological context, thin tissues (epithelimum, endothelium) are immersed in a nearly buoyant medium (lymph, blood, mucus), a practical situation that could lead to wrinkling on a thin viscous substrate.
+Yet, in a biological context, thin tissues (epithelimum, endothelium) are \seb{naturally confined and }immersed in a nearly buoyant medium (lymph, blood, mucus), a practical situation that could lead to wrinkling on a thin viscous substrate.
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
-Such interplay is also at stake in daily life problems such as wrinkles encountered while putting up wall paper~\cite{Pacione1991}.
+\sseb{Such interplay is also at stake in daily life problems}
+\sseb{such as wrinkles encountered while putting up}
+\sseb{ wall paper~\cite{Pacione1991}.} \seb{Therefore benchmark experiments that explore the possibility of wrinkling in confined soft materials immersed in a buoyancy-matched viscous medium are in line to get fundamental insights into such a biologically relevant situation.}
 
-Here, we have introduced permeable model bio-gels that produce hierarchical wrinkling as a result of the interplay between its gelation dynamics and the confinement conditions.
+Here, we introduce model permeable biogels that produce hierarchical wrinkling as a result of their interplay between their gelation dynamics and the confinement conditions.
 In slit geometry, we slowly acidify a dispersion of casein, one of the milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Leocmach2014} --this is a well know effect used in cheese or yogurt making, but, the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
-Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the casein network spontaneously shrink and swell at the micron level and wrinkles at the millimeter scale.
-We developed a model than pin down the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength, Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel layer for short wavelengths.
+As a result, not only does the casein form a gel as expected~\cite{Leocmach2014} \seb{\it citer plutot une ref historique} --this is a well-known effect used in cheese or yogurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
+Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimeter scale.
+\seb{We systematically investigate such} swelling-induced wrinkling of a porous gel sheet sandwiched between two viscous layers, the whole system being confined between two glass plates.
+As shown in Fig.~\ref{fig:dynamics} \seb{and in Supplemental Movies}, lateral confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
+Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle \seb{between two solid plates}.
+The two-dimensional nature of our experiment creates nested patterns that could prove useful to design original optical elements.
+\seb{\it Mettre plutot la phrase precedente en discussion/conclusion ?} In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario \seb{cannot explain the observations in Fig.~\ref{fig:dynamics}} and that only a wrinkling instability \seb{can be invoked}.
 
-\begin{figure}
+\begin{figure}[ht!]
 	\includegraphics{dynamics}
-	\caption{Dynamics of pattern formation.
+	\caption{Dynamics of pattern formation \seb{in a confined sheet of casein gel}.
 (a) Light transmission microscopy.
@@
 (b) 3D reconstruction from fluorescent confocal microscopy.
-(c) Fourier spectrum of binary altitudes of (a) corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at \SI{2}{\hour}.
-Scale bars are \SI{1}{\milli\metre}.}
+(c) Fourier spectra of binary altitudes of (a) corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
+Scale bars are \SI{1}{\milli\metre}.
+\seb{\it Indiquer la composition du gel~?}}
 	\label{fig:dynamics}
@@
 
-% § à déplacer après \section*{Pattern formation}, éventuellement en fusionnant des éléments avec le paragraphe précédent
-In the present work, we study the swelling-induced wrinkling of a porous gel sheet sandwiched between two viscous layers, with the whole system confined between two glass plates.
-As shown in Fig.~\ref{fig:dynamics}, the lateral confinement leads to saturation of the amplitude on the floor and ceiling followed by a cascade of secondary, ternary, etc.
-buckling.
-Such cascade buckling have been observed by~\cite{Roman1999} in a one dimensional macroscopic system, a sheet of steel forced to buckle.
-The 2D nature of our experiment creates nested patterns that could prove useful to design optical elements.
-In addition, one has to note that the primary wavelength is much smaller than the size of the system, implying that we are not in a simple buckling scenario but in a wrinkling instability.
-Our main goal is to identify the selection mechanism of the wavelength.
-
-% § à supprimer, ou à fusionner avec l'annonce de plan
-In the following, we will describe the physical chemistry that allows us to form such patterns.
-Then, we will model the situation and identify two possible selection mechanisms: either Darcy flow or Poiseuille flow.
-Finally we will compare our experimental data with the two models.
-Combining both Darcy and Poiseuille mechanisms will allow us to describe our data.
+The main goal of the present study is thus to identify the selection mechanism of the wavelength  \seb{and to discuss its implication for practical situations (?)}.
+In the following we first describe the physical chemistry that allows us to form such patterns.
+Then we  model the situation and identify two possible selection mechanisms, namely porous Darcy flow or viscous Poiseuille flow.
+Comparing our experimental data to these two models we finally show that \sseb{a model based on} a combination of both Darcy and Poiseuille mechanisms accurately describes our data.
+This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
+%Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel layer for short wavelengths.
 
@@
 	\includegraphics{acidification}
-	\caption{Acid-induced protein gels properties behave non-monotonously with pH.
-(a) pH decrease in 4\%w sodium caseinate solution acidified by 4\%w GDL.
-Horizontal line indicates isoelectric pH of casen (b) Evolution of storage modulus measured in a rheometer (full adhesion to rotor and stator).
-(c) Evolution of pore size measured by confocal microscopy in a half coated microscopy cell (no adhesion to cell ceiling, syn\ae{}resis and swelling allowed).}
+	\caption{Acid-induced protein gel properties behave nonmonotonically with pH.
+(a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL.
+Horizontal line \seb{The horizontal line?
+\it je laisse comme ca dans tous les captions.} indicates isoelectric pH of casein.
+(b) Evolution of the storage modulus $G'$ measured in a rheometer (full adhesion to rotor and stator).
+(c) Evolution of pore size measured by confocal microscopy in a half-coated microscopy cell (no adhesion to top wall, syn\ae{}resis and swelling allowed).}
 	\label{fig:acidification}
@@
 
-The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation.
-Since caseins adsorb strongly to interfaces, the resulting gel sticks to container walls, allowing rheological measurements (see Fig.~\ref{fig:acidification}b).
-Further decrease of the pH beyond the isoeletric point weakens the gel.
-This non monotonous behaviour of the elastic modulus translates in a non monotonous mesh size in a microscopy cell (slit geometry) coated so that the casein could not adsorb on the top wall (Fig.~\ref{fig:acidification}c), i.e.
-in syn\ae{}resis (solvent expulsion) followed by swelling.
+The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation (Fig.~\ref{fig:acidification}a).
+Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
+Further decrease of the pH beyond the isoeletric point weakens the gel \seb{due to \dots} \seb{\it Il manque une explication ici ainsi que des references generales sur la prise du yaourt}.
+This nonmonotonic behaviour of the elastic modulus translates into a nonmonotonic mesh size in a microscopy cell (slit geometry) coated so that the casein cannot adsorb on the top wall (Fig.~\ref{fig:acidification}c), i.e.
+into syn\ae{}resis (solvent expulsion) followed by swelling.
 
-Figure~\ref{fig:sideview}b shows confocal measurements in the case where both top and bottom walls have been coated.
-Starting from an homogeneous solution (black margins correspond to the glass walls), the gel forms and simultaneously expels solvent on both sides.
-Since we have two clear interfaces between a protein rich phase and a protein poor (water) phase, we model the resulting situation as a porous elastic sheet of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ et $H_2$ (see Fig.~\ref{fig:sideview}a).
-Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in second panel of Fig.~\ref{fig:sideview}b.
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the sheet of gel (Fig.~\ref{fig:sideview}d).
-We can thus confirm that the instability is due to swelling.
+Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated.
+Starting from a homogeneous solution (black margins correspond to the glass walls), the gel forms and simultaneously expells solvent on both sides.
+Since two interfaces \seb{clearly separate} a protein-rich phase and a protein-poor (water) phase, we model the resulting situation as a porous elastic sheet of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ \seb{in a cell of thickness $e=H_1+h+H_2$} (Fig.~\ref{fig:sideview}d) \seb{\it $e$ n'etait defini nulle part + attention sur la Fig.~3c il vaudrait mieux choisir de montrer $H_1$ et $H_2$ a un endroit ou le gel est horizontal, sinon le schema n'est valable qu'aux petites deformations + du coup ce sera it mieux de mettre le schema en 3b ?}.
+Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in the second panel of Fig.~\ref{fig:sideview}a.
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}b) and excess surface area of the gel layer (Fig.~\ref{fig:sideview}c).
+This confirms that the instability is due to swelling.
 
-Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses while swelling.
-Also, the only situation were be observed wrinkling is when both top and bottom walls are coated.
-We confirmed that at any given time the thickness $h$ of the sheet of gel is homogeneous in $(x,y)$.
-The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is not due to thickness inhomogeneities but to altitude gradients.
-The length of the optical path (along $z$) through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}a.
+Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses to develop while swelling: the only situation were wrinkling was observed is when both top and bottom walls are coated.
+We also confirmed that at any given \seb{point in time and space} the thickness $h$ of the sheet of gel is homogeneous.
+The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is \seb{therefore} not due to thickness inhomogeneities but to altitude gradients.
+\seb{Indeed} the length of the optical path (along \seb{the vertical direction} $z$) through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}d.
 
@@
 	\includegraphics{sideview}
-	\caption{Model and measurements.
-(a) Schematic side view of the cell.
-The constant thickness gel sheet (orange) is surrounded by water.
+	\caption{Measurements and modeling.
+(a) Confocal ($x$,$z$) cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
+Scale bar is \SI{100}{\micro\metre} (real size ratio).
+Evolution of (b) volume of the gel phase relative to cell volume and (c) excess area measured by confocal microscopy.
+Crosses correspond to times in (a).
+(d) Schematic side view of the cell.
+The constant thickness gel layer (orange) is surrounded by water.
 Yellow lines show the path of transmitted light through the gel phase.
-A region of interest is highlighted in gray.
-(b) Confocal XZ cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
-Scale bar is \SI{100}{\micro\metre} (real size ratio).
-Evolution of (c) volume of the gel phase relative to cell volume and (d) excess area measured by confocal microscopy.}
+A region of interest is highlighted in gray.}
 	\label{fig:sideview}
@@
 
-\section*{Scalings}
+\section*{Models and scaling laws}
+
+\seb{\it Ajouter une transition: en l'etat le passage au modele est trop brutal.}
 
@@
 \end{equation}
-In small deflection actual wavelength $\lambda$ is related to the amplitude $A$ as $\lambda \sim A^{1/4} \lambda^{*3/4}$.
+For small deflections the actual wavelength $\lambda$ is related to $\lambda^*$ and to the \seb{wrinkle} amplitude $A$ as $\lambda \sim A^{1/4} \lambda^{*3/4}$.
 
@@
 We identify three possible candidates for the lateral load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
-We confirmed that the former was playing no role by successfully repeating our experiment in a cell held vertical, i.e.
+We confirmed that the former plays no significant role by successfully repeating our experiment in a cell held vertically, i.e.
 with the weight acting longitudinally rather than laterally.
@@
 Prefactors are 0.63, 0.69 and 0.67 respectively.
-Continuous line is the best affine fit ($\lambda_{xp}=0.52\lambda_{th}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
+Continuous line is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all the points in (b).
+\seb{\it Mettre $\lambda_P$ et $\lambda_D$ comme dans le texte + $\lambda_{\rm mixed}$ ?
+Mettre tous les fits de la meme couleur (en gris ou en noir).
+Ajouter les textes ``porous model'', ``viscous model'' et ``mixed model'' (ou equivalent) sur les graphes.}}
 	\label{fig:DarcyPoiseuille}
@@
 
-In a situation where the sheet of gel would be sitting (not sticking) to the floor of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
-Darcy law relates the gradient of pressure (identified to $\sigma_\perp/h$) to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
+In a first limit where the sheet of gel would be sitting (\seb{without} sticking) to the floor of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
+Darcy law relates the pressure gradient (\seb{estimated as} $\sigma_\perp/h$) to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
@@
 where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the sheet and the solvent at a characteristic time $\tau$.
-This expression stresses the kinetic nature of the wavelength selection.
-However we note that the wavelength of (a generation of) the pattern does not evolve with time (see Fig.~\ref{fig:dynamics} and Sup.
+This expression emphasizes the kinetic nature of the wavelength selection \seb{through Darcy flow}.
+However we note that the wavelength of (a generation of) the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup.
 Fig.~\ref{fig:growth}) once the gel is touching both walls, we can deduce the relevant $\tau$ from the growth of the  pattern (see Supplementary Method for the precise definition).
+\seb{\it La phrase precedente n'est pas assez claire et semble contredire le texte en supplementary : a reformuler.}
 
-The opposite limiting case is when the sheet of gel is in the middle of the cell ($H_1=H_2$) and impervious ($\alpha\rightarrow 0$), destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
+The opposite limiting case is when the gel layer lies in the middle of the cell ($H_1=H_2$) and impermeable ($\alpha\rightarrow 0$).
+There destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
 By symmetry the lateral load across the gel $\sigma_\perp$ is also the pressure difference over the wavelength.
@@
 \end{equation}
-where we recognise the scaling of the wrinkles of an elastic sheet on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural moduli of the film and substrate.
-Indeed, in lubrication or low Reynolds number one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
+where we recognize the scaling for the wrinkles of an elastic sheet on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate.
+Indeed within the lubrication approximation or at low Reynolds numbers one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
 
-A more complete derivation of the models can be found in the Supplementary Method and yield the following prefactors:
+A more complete derivation of the models can be found in the Supplementary Method and yields the following prefactors:
 \begin{align}
 \lambda_D &= 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}\\
-\lambda_P &= \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
+\lambda_P &= \pi (hH)^{1/2}\left(\frac{2}{9}\Upsilon\right)^{1/6}
 \end{align}
-where $1/H^3 = 1/H_1^3 + 1/H_2^3$. The limiting case between the two models is for $H = H^* \equiv 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}$.
+where $1/H^3 = 1/H_1^3 + 1/H_2^3$. The crossover between the two models where $\lambda_D=\lambda_P$ is found for $H = H^* \equiv 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}$.
+
+\seb{\it Resumer ce qui precede en une phrase disant dans quel domaine de longueur d'onde on s'attend a trouver quel scaling pour nos manips.}
 
@@
 
-The above models yield wavelengths of the same order in most of the cases we studied.
-Thus only systematic measurements of experimental $\lambda$ and all the parameters involved in either model allows to discriminate between them (see Supplementary Method for details on the measurements).
-As summarized in Supplementary Table~\ref{tab:data}, we changed the thickness of the cell, impacting $h$ and the maximal amplitude.
-We changed the composition and thus the stiffness of the gel and its permeability.
-We also changed systematically the viscosity of the solvent by adding glycerol, however this also impacted the gel stiffness, permeability and $\tau$.
-Finally, the only parameter we had no control on was the initial altitude of the gel before wrinklink ($H$) that to our knowledge varied randomly.
-However, we were able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
+The two above models yield wavelengths that are of the same order in most of the cases under study.
+Thus only systematic measurements of the experimental $\lambda$ \seb{together with estimates for} all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
+As summarized in Supplementary Table~\ref{tab:data} we systematically varied (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
+Finally, the only parameter that we had no control upon was the initial altitude $H$ of the gel before wrinkling that varied randomly.
+However we were able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
-Here the size of the blister is fixed as half of the size of the primary pattern as show on Fig.~\ref{fig:dynamics}c and $\tau$ is the variable parameter.
+Here \seb{the wavelength associated to the secondary} blister is taken as half of that of the primary pattern as shown in Fig.~\ref{fig:dynamics}c and $\tau$ is the variable parameter.
+\seb{\it Pas sur que la fin de cette phrase soit bien claire pour moi.}
 
 Our results are displayed in Fig.~\ref{fig:DarcyPoiseuille}.
-Neither model alone is able to reproduce all the data.
+\seb{Although they provide the correct order of magnitude for $\lambda$,} neither model alone is able to \seb{quantitatively} reproduce the full data set.
 However considering only the points where $H<H^*$ (respectively $H>H^*$), the Darcy (resp.
 Poiseuille) model fits the data well albeit with a prefactor.
-Note that it would be easy to oversight the failure of the Poiseuille model at small wavelength by adding an unexplained offset.
+Note that it would be easy to oversight \seb{\it ``oversight'' $\neq$ verbe: plutot ``oversee'' ?} the failure of the Poiseuille model at small wavelengths by adding an unexplained offset \seb{(dashed line in Fig.~\ref{fig:DarcyPoiseuille}b)}.
 Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.
 
-%Conclusion à écrire
+\seb{\it Ajouter une conclusion.}
 
 \section*{Methods summary}
-To prevent adsorption of caseins to surfaces, microscope slides and cover-slips were cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush was polymerized from methacrylate groups.
+To prevent adsorption of caseins to surfaces, microscope slides and cover slips were cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush was polymerized from methacrylate groups.
 Sodium caseinate (Firmenich) was labelled with Dylight~550 NHS ESTER (Thermo Scientific) and then purified by centrifugation.
-Unlabelled sodium caseinate powder was dissolved in deionised water (or water+glycerol mixture).
+\seb{\it vraiment ``purified''?} Unlabelled sodium caseinate powder was dissolved in deionised water (or water--glycerol mixture).
 To induce gelation glucono-$\delta$-lactone (GDL) in powder (Firmenich) was added to the solution.
-\SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
+\SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution, \seb{stirred} and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
 3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
-We used either 10x (air) lens to measure the geometry of wrinkling (see SM) or 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscope data or by transmitted/reflected light macroscope (Nikkon SMZ745T/Leica DMS1000), were used to measure the wavelength of the primary pattern.
+We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplemental Material) or a 63x (oil) lens to observe the microstructure of the gel.
+Larger scope pictures, obtained either by stitching fluorescent \seb{confocal microscopy images} or by transmitted/reflected light macroscope (Nikkon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern.
 Secondary pattern wavelength is set to half of the primary wavelength, according to Fig.~\ref{fig:dynamics}c.
 Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
+
+\seb{\it Il y a pas mal de petits soucis dans la biblio : manque des page numbers ou refs a completer. Je joins un scan de mes notes pour ajuster le .bib.}
 
@@
 Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
-The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various sage of the research. 
+The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
 Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
-Alan Parker at Firmenich provided the casein and GDL; Catherine Barentin provided the (late) glass Couette cell of SM XXXX.
-Special thanks to Madame Biot who made M.A. Biot work accessible through \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 
-ML thank the Region Rhône Alpes and the Programme d'avenir Lyon - Saint Etienne (PALSE NoGELPo) for postoctoral grant. 
+Alan Parker at Firmenich provided the casein and GDL; \sseb{Catherine Barentin provided the (late)}
+\sseb{glass Couette cell of SM.}
+Special thanks to Madame Biot who made M.A. Biot work accessible through the \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 
+ML thanks the Region Rhône Alpes and the Programme d'Avenir Lyon - Saint Etienne (PALSE NoGELPo) for postoctoral grant. 
 TG and MN acknowledge funding from ANR XXXX. 
@@
 	\caption{Gel microstructure.
-(a) Detail of a confocal micrograph of a 4\%w casein, 4\%w GDL in water, when only the slide is coated (not the cover slip).
-Time corresponds to the minimum in mesh size.
+(a) Detail of a confocal micrograph of a 4\%w casein in water acidified with 4\%w GDL, when only the microscope slide is coated (not the cover slip).
+Images were taken at times corresponding to the minimum in mesh size.
 Scale bar is \SI{10}{\micro\metre}.
-(b) Corresponding radial averaged Fourier transform.
-Black line is the best fit by Eq.XXX with $\xi=\SI{2.0}{\micro\metre}$ and $d_{3D}=2.27$.
-Inset shows the center of the non radial average spectrum (arbitrary logarithmic color scale).
+(b) Corresponding radially averaged Fourier transform.
+\seb{\it Fourier transform or Fourier spectrum or power spectrum?
+Clarify?} Black line is the best fit by Eq.~(\ref{eq:fractalS}) with $\xi=\SI{2.0}{\micro\metre}$ and $d_{3D}=2.27$.
+Inset shows the center of the \sseb{non radial average} spectrum \seb{using a} logarithmic color scale.
 Scale bar is \SI{10}{\per\micro\metre}.
-(c-d) Idem in a 50\%w glycerol solvent, with $\xi=\SI{0.3}{\micro\metre}$ and $d_{3D}=3$.}
+(c-d) Same as (a-b) in a 50\%w \seb{glycerol--water mixture leading to} $\xi=\SI{0.3}{\micro\metre}$ and $d_{3D}=3$.}
 	\label{fig:mesh}
@@
 
-To follow the evolution of the microstructure in a gel undergoing syn\ae{}resis and swelling, we had to keep it in the thin focal plane of the 63x objective lens of the confocal microscope.
+To follow the evolution of the microstructure in a gel undergoing syn\ae{}resis and swelling, it was necessary to keep it into the thin focal plane of the 63x objective lens of the confocal microscope.
 To do so, we coated only the microscope slide with acrylamide brush to allow syn\ae{}resis on this side.
-The cover-slip was cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
+The cover slip was cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
 Neither buckling nor wrinkling was detected in such configuration.
 
-We acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (see Fig.~\ref{fig:mesh}a and c).
-After Hann windowing we performed Fourier transform (see insets of Fig.~\ref{fig:mesh}b and d) and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:mesh}b and d) that we fit by the following fractal form:
+We acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:mesh}a and c).
+After Hann windowing we performed Fourier transform (see insets of Fig.~\ref{fig:mesh}b and d) and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:mesh}b and d) that was fitted by the following fractal form \seb{\it Clarify $I(q)$ vs $S(q)$}:
 \begin{equation}
 I(q) = \frac{\chi}{1+(\xi q)^\gamma}
+\label{eq:fractalS}
 \end{equation}
-were the susceptibility $\chi$ is related to the contrast of fluorescence between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and the exponent $\gamma$ is loosely related to the fractal dimension.
+where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and the exponent $\gamma$ is loosely related to the fractal dimension $d_{3D}$.
+
+\seb{\it Actualiser la fin de cette partie avec nouvelle Eq. (6) et texte correspondant.}
 
@@
 (b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
-(c) Idem in 50\%w glycerol, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
+(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.
+\seb{\it Changer $z(\infty)$ en $z_\infty$ dans le schema.}}
 	\label{fig:permeability}
@@
 Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
-A webcam (Logitech Webcam Pro 9000) records the rise of the liquid in each tube.
+A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
 Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.
 
-Darcy law through the gel of permeability $\alpha$ and height $H$ writes
+Darcy law through the gel of permeability $\alpha$ and height $H$ reads \seb{\it Attention au signe ?} 
 \begin{equation}
-\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
+\frac{dz}{dt} = \seb{-} \frac{\alpha}{\eta}\frac{\Delta P}{H}
 \end{equation}
-where $\Delta P$ is the the hydrostatic pressure relative to $z_\infty$. We can thus write
+where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure \seb{drop through the gel, with $\rho$ the buffer density. {\it Avez-vous mesure $\rho$ independamment?}} We can thus write
 \begin{equation}
@@
 
-By fitting $z_\infty-z(t)$ by an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
-Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosities and low permeability.
+By fitting $z_\infty-z(t)$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
+Note that even with $H$ of the order of a few millimeters high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
 Obtaining $z_\infty$ without tube 2 would be time prohibitive.
 
-\subsection*{Confocal images segmentation}
+\subsection*{Confocal images processing}
 
-At 10x magnification, we acquired sacks of (depending on cell thickness) 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
+At 10x magnification and depending on cell thickness, we acquired stacks 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
 Each confocal slice was $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
-At this low resolution details of the gel microstructure are undistinguishable ($\xi<$px for all compositions) and the gel appear as a continuous medium.
+At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.
 
-We first denoise each plane by a small Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
-Then, for each $(x,y)$, we focus on the z-dependant intensity profile $I(z)$.
-We define the position of the top surface of the gel $z_\text{top}$ as where $I(z)$ first crosses half of its maximum value.
-Since the criterion is local, we do not need to correct for background intensity variation.
+We first denoise each plane by a \seb{short-range} Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
+Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
+We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value \seb{over the given vertical slice}.
+Since this criterion is local, we do not need to correct for background intensity variation.
 In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.
@@
 Velocity is obtained by time differentiation.
-Once the instability has saturated on floor and ceiling of the cell, peaks in the histograms of $z_\text{top}$, and $z_\text{bottom}$ yield the measure of the positions of the walls, thus to $e$ and retrospectively to $H_1$ and $H_2$.
+Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus \seb{of the cell thickness} $e$ and retrospectively of $H_1$ and $H_2$.
 
@@
 	\caption{Growth of the instability.
-(a) Evolution of the gel velocity $v$, integrated over the whole surface.
-Inset: Evolution of the amplitude (peak to peak measurement) in semi logarithmic scale.
+(a) Evolution of the gel velocity $v$ \seb{averaged} over the whole surface.
+Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
 Black line is the best exponential fit with $\tau=\SI{90}{\second}$.
-(b) Evolution of the bottom profile of the gel from the onset of instability (yellow) to the contact to both walls (brown).
-Note that the lateral size of the blister evolve in this regime.
-Thick black curve is the profile at much longer time.
-Note that blister foot (highlighted by the orange vertical line) has not moved since contact.}
+(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
+Note that the lateral size of the blister evolves in this regime.
+Thick black curve is the profile at much longer times.
+Note that the blister foot (highlighted by the orange vertical line) has not moved since contact.
+\seb{\it Utiliser des symboles ($t$, $r$,$ A(t)$) plutot que des mots ?
++ j'ai plusieurs petites questions par rapport a cette figure.}}
 	\label{fig:growth}
@@
 
-In most of our samples, it was possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
-However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both highest and lowest point of the surface.
-Even when valid, such local measurement is noisy and we prefer the measure of the vertical displacement of the gel $v$ that can be integrated over the whole field of view.
+In most of our samples it was possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
+However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
+Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical \seb{velocity ?} $v$ of the gel that can be integrated over the whole field of view.
 
-We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls.
-We thus used consistently this approximation of $\tau$ for all our samples.
-This is coherent with the fact that the blister size $\propto\lambda$ does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is decided when $A(t) = A_\text{max}$.
+We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls \seb{and $v_\text{max}$ is the maximum velocity of the gel}.
+We used this approximation of $\tau$ for all our samples.
+This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
 
-\subsection*{Darcy+Poiseuille model}
+\subsection*{Mixed Darcy--Poiseuille model}
 
-Mechanical equilibrium of a plate of thickness $h$, young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a lateral load $\sigma_\perp$ along $z$ writes:
+\subsubsection*{General framework for wrinkling}
+
+The mechanical equilibrium of a plate of thickness $h$, Young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a lateral load $\sigma_\perp$ along $z$ writes
 \begin{equation}
@@
 where $w(x,t)$ is the deflection of the plate along $z$.
+\seb{\it Cette formulation --stationnaire-- des equations ne semble pas coherente avec le fait de chercher une solution dependant du temps.
+Clarifier ou donner des references...}
 
-We suppose an exponentially growing sinusoidal deflection $w(x,t) =  A_0 e^{t/\tau} \sin kx$ of characteristic time $\tau$ and wavevector $k$. In operator language, Eq.~\ref{eq:mecheq} writes:
+We assume an exponentially growing sinusoidal deflection $w(x,t) =  A_0 e^{t/\tau} \sin kx$ of characteristic time $\tau$ and wavevector $k$.
+\seb{\it En lien avec la remarque precedente, il est delicat de donner un sens a $\tau$ ici en l'absence de l'eq (14).} \sseb{In operator language,} Eq.~(\ref{eq:mecheq}) then reads
 \begin{equation}
@@
 
-The deformation of the substrate of modulus $B_s$ gives the lateral load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
+\seb{\it Vraiment utile de detailler ce cas qui ne correspond pas aux manips ?}
+
+\seb{In the case of an elastic film lying over an infinite elastic medium,} the deformation of the substrate of modulus $B_s$ gives the lateral load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
 \begin{equation}
@@
 
-Minimizing $\sigma_\parallel$ function of $k$ yield the well known~\cite{Biot1957,Cerda2003} result $hk = \left(6B_s/B\right)^{1/3}$ and thus the dominant wavelength is
+Minimizing $\sigma_\parallel$ as a function of $k$ yields the well-known result $hk = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is \seb{\it Il serait bien de rappeler pourquoi on minimise ici $\sigma_\parallel$ : cela ne vient-il pas plus de considerations energetiques que de la selection du mode le plus instable dans une relation de dispersion ?
+Les deux visions sont-elles equivalentes ?
+Je n'ai pas les idees bien claires la-dessus --c'est sans doute relie au pb deja evoque plus haut-- et le lecteur moyen risque de se poser la question de comment on justifie proprement cette minimisation\dots}
 \begin{equation}
@@
 
+\seb{\it Anecdotique : pourquoi $\lambda_d$ plutot que $\lambda$ ?}
+
 \subsubsection*{Elastic film on a viscous layer}
-The layer thickness $H$ is supposed much smaller than $\lambda_d$, we are thus in the lubrication approximation.
-Since we work at small deflections, we can neglect the displacement of the plate along $x$.
+\seb{Let us now turn to the case of an elastic, impermeable film on a viscous layer which is more relevant for the present experimental situation.} Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation.
+\seb{This corresponds to the case where the gel layer is close to one of the cell walls and this specifically applies to secondary blisters.} Further considering only small deflections, we may neglect the displacement of the plate along $x$.
 The flux along $x$ is thus of the Poiseuille form
@@
 \end{equation}
-Keeping only linear order terms, in operator language
+Keeping only linear order terms, one gets \sseb{in operator language}
 \begin{equation}
@@
 \end{equation}
-and can be minimised by $k_P$ such that
+and is minimized at $k_P$ such that
 \begin{equation}
@@
 \end{equation}
-with $\Upsilon = B\tau/\eta$ the viscoelastic factor. The dominant wavelength for Poiseuille model is thus
+with $\Upsilon = B\tau/\eta$ the viscoelastic factor. The dominant wavelength for the Poiseuille model is thus
 \begin{equation}
@@
 \subsubsection*{Porous elastic film with no possible Poiseuille flow}
-The lateral load derives from the Darcy law through the film of permeability $\alpha$:
+\seb{Another limiting case is that of a porous elastic film sandwiched between two viscous layers that are supposed to be large enough so that the previous Poiseuille mechanism can be neglected.} The lateral load derives from the Darcy law through the film of permeability $\alpha$:
 \begin{equation}
--\sigma_\perp = p_2-p_1 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
+\sigma_\perp = p_1-p_2 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
 \label{eq:DarcyLoad}
@@
 \end{equation}
-and can be minimised by $k_D$ such that
+and is minimized at $k_D$ such that
 \begin{equation}
@@
 \subsubsection*{Porous elastic film between two viscous layers}
-The mass conservation over the whole height of the cell yield $Q_1 + Q_2 = 0$ and thus
+\seb{Finally the most realistic scenario mixes both Poiseuille and Darcy mechanisms, i.e.
+considers a porous elastic film between two viscous layers of arbitrary thickness.
+In this case,} the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
 \begin{equation}
@@
 
-Note that we recover Eq~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and Eq~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
-We can recast this intuition in term of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $k_P$ and $k_D$:
+Note that we recover Eq~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and Eq.~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
+We can recast this intuition in terms of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $k_P$ and $k_D$:
 \begin{equation}
@@
 
-Analytic minimisation of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but heavy. Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
+Analytic minimization of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but \seb{cumbersome}.
+Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
 \begin{align}
@@
 \end{align}
-and we minimise this expression numerically.
+and we minimize this expression numerically.
 

### 2015-01-20 plis_figures.tex
fig:patterns

@@
 
+\begin{figure*}
+	\tikzsetnextfilename{patterns}
+	\input{fig_plis/patterns.pgf}
+	\caption{Patterns corresponding to the samples of Table~\ref{tab:data}.
+(a-d) increasing cell thickness.
+(e-f) larger solid content.
+(g-i) increasing glycerol content.
+All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
+Scale is common to all panels (scale bar \SI{1}{\milli\metre}).
+Length of arrows correspond to the measured primary wavelength.}
+	\label{fig:patterns}
+\end{figure*}
+
 \end{document}

### 2015-01-22 fig_plis/article_plis.tex
Terminologie generale uniformisée dans tout l'article: sheet -> film, floor and ceiling -> bottom and top walls, lateral load -> transverse load, laterally -> transversely
Lateral size of the blister -> longitudinal

@@
 Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimeter scale.
-\seb{We systematically investigate such} swelling-induced wrinkling of a porous gel sheet sandwiched between two viscous layers, the whole system being confined between two glass plates.
-As shown in Fig.~\ref{fig:dynamics} \seb{and in Supplemental Movies}, lateral confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
+\seb{We systematically investigate such} swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
+As shown in Fig.~\ref{fig:dynamics} \seb{and in Supplemental Movies}, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
 Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle \seb{between two solid plates}.
@@
 	\includegraphics{dynamics}
-	\caption{Dynamics of pattern formation \seb{in a confined sheet of casein gel}.
+	\caption{Dynamics of pattern formation \seb{in a confined film of casein gel}.
 (a) Light transmission microscopy.
@@
 This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
-%Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel layer for short wavelengths.
+%Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel film for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel film for short wavelengths.
 
@@
 Starting from a homogeneous solution (black margins correspond to the glass walls), the gel forms and simultaneously expells solvent on both sides.
-Since two interfaces \seb{clearly separate} a protein-rich phase and a protein-poor (water) phase, we model the resulting situation as a porous elastic sheet of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ \seb{in a cell of thickness $e=H_1+h+H_2$} (Fig.~\ref{fig:sideview}d) \seb{\it $e$ n'etait defini nulle part + attention sur la Fig.~3c il vaudrait mieux choisir de montrer $H_1$ et $H_2$ a un endroit ou le gel est horizontal, sinon le schema n'est valable qu'aux petites deformations + du coup ce sera it mieux de mettre le schema en 3b ?}.
+Since two interfaces \seb{clearly separate} a protein-rich phase and a protein-poor (water) phase, we model the resulting situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ \seb{in a cell of thickness $e=H_1+h+H_2$} (Fig.~\ref{fig:sideview}d) \seb{\it $e$ n'etait defini nulle part + attention sur la Fig.~3c il vaudrait mieux choisir de montrer $H_1$ et $H_2$ a un endroit ou le gel est horizontal, sinon le schema n'est valable qu'aux petites deformations + du coup ce sera it mieux de mettre le schema en 3b ?}.
 Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in the second panel of Fig.~\ref{fig:sideview}a.
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}b) and excess surface area of the gel layer (Fig.~\ref{fig:sideview}c).
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}b) and excess surface area of the gel film (Fig.~\ref{fig:sideview}c).
 This confirms that the instability is due to swelling.
@@
 Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses to develop while swelling: the only situation were wrinkling was observed is when both top and bottom walls are coated.
-We also confirmed that at any given \seb{point in time and space} the thickness $h$ of the sheet of gel is homogeneous.
+We also confirmed that at any given \seb{point in time and space} the thickness $h$ of the film of gel is homogeneous.
 The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is \seb{therefore} not due to thickness inhomogeneities but to altitude gradients.
@@
 (d) Schematic side view of the cell.
-The constant thickness gel layer (orange) is surrounded by water.
+The constant thickness gel film (orange) is surrounded by water.
 Yellow lines show the path of transmitted light through the gel phase.
@@
 
-The characteristic length $\lambda^*$ of a wrinkling pattern is set by the competition between the flexural modulus of the sheet $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a lateral load $\sigma_\perp$~\cite{Vella2009,Kolinski2009}:
+The characteristic length $\lambda^*$ of a wrinkling pattern is set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$~\cite{Vella2009,Kolinski2009}:
 \begin{equation}
@@
 % Pour Mathieu : ajouter quelque part des applications numériques pour les trois longueurs d'ondes gravité, Darcy, Poiseuille
-We identify three possible candidates for the lateral load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
+We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
 We confirmed that the former plays no significant role by successfully repeating our experiment in a cell held vertically, i.e.
-with the weight acting longitudinally rather than laterally.
+with the weight acting longitudinally rather than transversely.
 
@@
 
-In a first limit where the sheet of gel would be sitting (\seb{without} sticking) to the floor of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
+In a first limit where the film of gel would be sitting (\seb{without} sticking) to the bottom wall of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
 Darcy law relates the pressure gradient (\seb{estimated as} $\sigma_\perp/h$) to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
@@
 \end{equation}
-where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the sheet and the solvent at a characteristic time $\tau$.
+where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
 This expression emphasizes the kinetic nature of the wavelength selection \seb{through Darcy flow}.
@@
 
-The opposite limiting case is when the gel layer lies in the middle of the cell ($H_1=H_2$) and impermeable ($\alpha\rightarrow 0$).
+The opposite limiting case is when the gel film lies in the middle of the cell ($H_1=H_2$) and impermeable ($\alpha\rightarrow 0$).
 There destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
-By symmetry the lateral load across the gel $\sigma_\perp$ is also the pressure difference over the wavelength.
+By symmetry the transverse load across the gel $\sigma_\perp$ is also the pressure difference over the wavelength.
 Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_\perp \frac{H^3}{\eta\lambda^2}$.
@@
 \end{equation}
-where we recognize the scaling for the wrinkles of an elastic sheet on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate.
+where we recognize the scaling for the wrinkles of an elastic film on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate.
 Indeed within the lubrication approximation or at low Reynolds numbers one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
@@
 (b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
-Note that the lateral size of the blister evolves in this regime.
+Note that the longitudinal size of the blister evolves in this regime.
 Thick black curve is the profile at much longer times.
@@
 
-The mechanical equilibrium of a plate of thickness $h$, Young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a lateral load $\sigma_\perp$ along $z$ writes
+The mechanical equilibrium of a plate of thickness $h$, Young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a transverse load $\sigma_\perp$ along $z$ writes
 \begin{equation}
@@
 
-\seb{In the case of an elastic film lying over an infinite elastic medium,} the deformation of the substrate of modulus $B_s$ gives the lateral load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
+\seb{In the case of an elastic film lying over an infinite elastic medium,} the deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
 \begin{equation}
@@
 \seb{Let us now turn to the case of an elastic, impermeable film on a viscous layer which is more relevant for the present experimental situation.} Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation.
-\seb{This corresponds to the case where the gel layer is close to one of the cell walls and this specifically applies to secondary blisters.} Further considering only small deflections, we may neglect the displacement of the plate along $x$.
+\seb{This corresponds to the case where the gel film is close to one of the cell walls and this specifically applies to secondary blisters.} Further considering only small deflections, we may neglect the displacement of the plate along $x$.
 The flux along $x$ is thus of the Poiseuille form
@@
 \subsubsection*{Porous elastic film with no possible Poiseuille flow}
-\seb{Another limiting case is that of a porous elastic film sandwiched between two viscous layers that are supposed to be large enough so that the previous Poiseuille mechanism can be neglected.} The lateral load derives from the Darcy law through the film of permeability $\alpha$:
+\seb{Another limiting case is that of a porous elastic film sandwiched between two viscous layers that are supposed to be large enough so that the previous Poiseuille mechanism can be neglected.} The transverse load derives from the Darcy law through the film of permeability $\alpha$:
 \begin{equation}
@@
 
-Combining Eqs.~(\ref{eq:pressures}, \ref{eq:conservationDarcy}, \ref{eq:Darcy}) we obtain the lateral load
+Combining Eqs.~(\ref{eq:pressures}, \ref{eq:conservationDarcy}, \ref{eq:Darcy}) we obtain the transverse load
 \begin{equation}

### 2015-01-22 fig_plis/article_plis.tex
chasse aux parenthèses

@@
 
-\seb{\it Terminologie generale a uniformiser dans tout l'article: sheet $\rightarrow$ layer, floor and ceiling $\rightarrow$ bottom and top walls, impervious $\rightarrow$ impermeable, non-monotonous $\rightarrow$ nonmonotonic, lateral load $\rightarrow$ transverse load, longitudinal/compression load $\rightarrow$ compressional load, laterally $\rightarrow$ transversely.
-A discuter puis a changer partout systematiquement ?
-De facon generale, il reste encore beaucoup de choses entre parentheses: soit c'est utile et on enleve les parentheses, soit c'est inutile et on supprime ?}
-
 There are many ways and reasons for a film to wrinkle.
@@
 On the engineering side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
-Such patterns can be triggered by temperature dilation, swelling (eventually pH or light induced) or the removal of pre-strain.
+Such patterns can be triggered by temperature dilation, swelling or the removal of pre-strain.
 On the biology side wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
-On the physics side analogues of wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013} \sseb{(linked to the pioneering work of Smoluchowski~\cite{Smoluchowski1927})} \seb{\it utile?}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+On the physics side analogues of wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
@@
 
-Yet, in a biological context, thin tissues (epithelimum, endothelium) are \seb{naturally confined and }immersed in a nearly buoyant medium (lymph, blood, mucus), a practical situation that could lead to wrinkling on a thin viscous substrate.
+Yet, in a biological context, thin tissues, e.g.
+epithelimum or endothelium, are \seb{naturally confined and }immersed in a nearly buoyant medium, e.g.
+lymph, blood or mucus, a practical situation that could lead to wrinkling on a thin viscous substrate.
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
@@
 \it je laisse comme ca dans tous les captions.} indicates isoelectric pH of casein.
-(b) Evolution of the storage modulus $G'$ measured in a rheometer (full adhesion to rotor and stator).
-(c) Evolution of pore size measured by confocal microscopy in a half-coated microscopy cell (no adhesion to top wall, syn\ae{}resis and swelling allowed).}
+(b) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator.
+(c) Evolution of pore size measured by confocal microscopy in a half-coated microscopy cell with no adhesion to top wall, syn\ae{}resis and swelling allowed.}
 	\label{fig:acidification}
@@
 Further decrease of the pH beyond the isoeletric point weakens the gel \seb{due to \dots} \seb{\it Il manque une explication ici ainsi que des references generales sur la prise du yaourt}.
-This nonmonotonic behaviour of the elastic modulus translates into a nonmonotonic mesh size in a microscopy cell (slit geometry) coated so that the casein cannot adsorb on the top wall (Fig.~\ref{fig:acidification}c), i.e.
+This nonmonotonic behaviour of the elastic modulus translates into a nonmonotonic mesh size in a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall (Fig.~\ref{fig:acidification}c), i.e.
 into syn\ae{}resis (solvent expulsion) followed by swelling.
@@
 Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated.
-Starting from a homogeneous solution (black margins correspond to the glass walls), the gel forms and simultaneously expells solvent on both sides.
-Since two interfaces \seb{clearly separate} a protein-rich phase and a protein-poor (water) phase, we model the resulting situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ \seb{in a cell of thickness $e=H_1+h+H_2$} (Fig.~\ref{fig:sideview}d) \seb{\it $e$ n'etait defini nulle part + attention sur la Fig.~3c il vaudrait mieux choisir de montrer $H_1$ et $H_2$ a un endroit ou le gel est horizontal, sinon le schema n'est valable qu'aux petites deformations + du coup ce sera it mieux de mettre le schema en 3b ?}.
+Starting from a homogeneous solution (first panel Fig.~\ref{fig:sideview}a where black margins correspond to the glass walls), the gel forms and simultaneously expels solvent on both sides.
+Since two interfaces \seb{clearly separate} a protein-rich phase and a protein-poor phase, we model the resulting situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ \seb{in a cell of thickness $e=H_1+h+H_2$} (Fig.~\ref{fig:sideview}d) \seb{\it $e$ n'etait defini nulle part + attention sur la Fig.~3c il vaudrait mieux choisir de montrer $H_1$ et $H_2$ a un endroit ou le gel est horizontal, sinon le schema n'est valable qu'aux petites deformations + du coup ce sera it mieux de mettre le schema en 3b ?}.
 Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in the second panel of Fig.~\ref{fig:sideview}a.
@@
 The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is \seb{therefore} not due to thickness inhomogeneities but to altitude gradients.
-\seb{Indeed} the length of the optical path (along \seb{the vertical direction} $z$) through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}d.
+\seb{Indeed} the length of the optical path along \seb{the vertical direction} $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}d.
 
@@
 	\caption{Measurements and modeling.
-(a) Confocal ($x$,$z$) cuts showing syn\ae{}resis, swelling and (cascade) wrinkling.
+(a) Confocal ($x$,$z$) cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
 Scale bar is \SI{100}{\micro\metre} (real size ratio).
@@
 
-In a first limit where the film of gel would be sitting (\seb{without} sticking) to the bottom wall of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
-Darcy law relates the pressure gradient (\seb{estimated as} $\sigma_\perp/h$) to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
+In a first limit where the film of gel would be sitting without sticking to the bottom wall of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
+Darcy law relates the pressure gradient, estimated as $\sigma_\perp/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
@@
 This expression emphasizes the kinetic nature of the wavelength selection \seb{through Darcy flow}.
-However we note that the wavelength of (a generation of) the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup.
+However we note that the wavelength of a generation of the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup.
 Fig.~\ref{fig:growth}) once the gel is touching both walls, we can deduce the relevant $\tau$ from the growth of the  pattern (see Supplementary Method for the precise definition).
@@
 Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
-Alan Parker at Firmenich provided the casein and GDL; \sseb{Catherine Barentin provided the (late)}
-\sseb{glass Couette cell of SM.}
+Alan Parker at Firmenich provided the casein and GDL; 
 Special thanks to Madame Biot who made M.A. Biot work accessible through the \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 
@@
 	\caption{Gel microstructure.
-(a) Detail of a confocal micrograph of a 4\%w casein in water acidified with 4\%w GDL, when only the microscope slide is coated (not the cover slip).
+(a) Detail of a confocal micrograph of a 4\%w casein in water acidified with 4\%w GDL, when only the microscope slide is coated, not the cover slip.
 Images were taken at times corresponding to the minimum in mesh size.
@@
 Thick black curve is the profile at much longer times.
-Note that the blister foot (highlighted by the orange vertical line) has not moved since contact.
+Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.
 \seb{\it Utiliser des symboles ($t$, $r$,$ A(t)$) plutot que des mots ?
@@
 
-The scaling is then identical to the (much more involved) derivation by \citet{Huang2002}.
 When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of Eq.~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness 

### 2015-01-22 fig_plis/article_plis.tex
Incorpore les modifs de Sébastien sur le texte principal: - Références historiques sur le Yaourt -Meilleure explication de ce qui se passe microscopiquement - Défini $e$ - Réissère les citations Darcy et Poiseuille - Remet Catherine, sans le Couette

@@
 An elastic film buckles due to excess area with respect to its boundaries.
-Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength \sseb{as discovered originally in aeronautics} \cite{Gough1940, Bijlaard1946}.
+Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}.
 On the engineering side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
@@
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002} \seb{and to date this confined situation remains without experimental equivalent.} \sseb{as for a free interface gravity dominates}
-\sseb{ over viscous substrate} \seb{Pas tr\`es clair, \`a reformuler~?}
+Stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002}.
+Since in practice gravity sets the wavelength in such situation~\cite{Smoluchowski1927}, the wrinkling on a confined viscous substrate has not been studied experimentally.
 
 Yet, in a biological context, thin tissues, e.g.
-epithelimum or endothelium, are \seb{naturally confined and }immersed in a nearly buoyant medium, e.g.
-lymph, blood or mucus, a practical situation that could lead to wrinkling on a thin viscous substrate.
+epithelimum or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g.
+lymph, blood or mucus, a practical situation were a thin viscous substrate would set the wrinkling wavelength.
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
-\sseb{Such interplay is also at stake in daily life problems}
-\sseb{such as wrinkles encountered while putting up}
-\sseb{ wall paper~\cite{Pacione1991}.} \seb{Therefore benchmark experiments that explore the possibility of wrinkling in confined soft materials immersed in a buoyancy-matched viscous medium are in line to get fundamental insights into such a biologically relevant situation.}
+Therefore benchmark experiments that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line to get fundamental insights into such a biologically relevant situation.
 
@@
 In slit geometry, we slowly acidify a dispersion of casein, one of the milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Leocmach2014} \seb{\it citer plutot une ref historique} --this is a well-known effect used in cheese or yogurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
+As a result, not only does the casein form a gel as expected~\cite{Bremer1989} --this is a well-known effect used in cheese or yogurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
 Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimeter scale.
-\seb{We systematically investigate such} swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
-As shown in Fig.~\ref{fig:dynamics} \seb{and in Supplemental Movies}, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
-Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle \seb{between two solid plates}.
-The two-dimensional nature of our experiment creates nested patterns that could prove useful to design original optical elements.
-\seb{\it Mettre plutot la phrase precedente en discussion/conclusion ?} In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario \seb{cannot explain the observations in Fig.~\ref{fig:dynamics}} and that only a wrinkling instability \seb{can be invoked}.
+We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
+As shown in Fig.~\ref{fig:dynamics} and in Supplemental Movies, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
+Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates.
+In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Fig.~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.
 
@@
 	\includegraphics{dynamics}
-	\caption{Dynamics of pattern formation \seb{in a confined film of casein gel}.
+	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w).
 (a) Light transmission microscopy.
@@
 Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-Scale bars are \SI{1}{\milli\metre}.
-\seb{\it Indiquer la composition du gel~?}}
+Scale bars are \SI{1}{\milli\metre}.}
 	\label{fig:dynamics}
@@
 Then we  model the situation and identify two possible selection mechanisms, namely porous Darcy flow or viscous Poiseuille flow.
-Comparing our experimental data to these two models we finally show that \sseb{a model based on} a combination of both Darcy and Poiseuille mechanisms accurately describes our data.
+Comparing our experimental data to these two models we finally show that a combination of both Darcy and Poiseuille mechanisms accurately describes our data.
 This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
@@
 (a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL.
-Horizontal line \seb{The horizontal line?
-\it je laisse comme ca dans tous les captions.} indicates isoelectric pH of casein.
+The horizontal line indicates isoelectric pH of casein.
 (b) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator.
@@
 
-The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation (Fig.~\ref{fig:acidification}a).
+The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation~\cite{Bremer1989} (Fig.~\ref{fig:acidification}a).
 Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
-Further decrease of the pH beyond the isoeletric point weakens the gel \seb{due to \dots} \seb{\it Il manque une explication ici ainsi que des references generales sur la prise du yaourt}.
-This nonmonotonic behaviour of the elastic modulus translates into a nonmonotonic mesh size in a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall (Fig.~\ref{fig:acidification}c), i.e.
-into syn\ae{}resis (solvent expulsion) followed by swelling.
+Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplemental Fig.~\ref{fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
+In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplemental Fig.~\ref{fig:nonmonotonic}).
 
@@
 Starting from a homogeneous solution (first panel Fig.~\ref{fig:sideview}a where black margins correspond to the glass walls), the gel forms and simultaneously expels solvent on both sides.
-Since two interfaces \seb{clearly separate} a protein-rich phase and a protein-poor phase, we model the resulting situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ \seb{in a cell of thickness $e=H_1+h+H_2$} (Fig.~\ref{fig:sideview}d) \seb{\it $e$ n'etait defini nulle part + attention sur la Fig.~3c il vaudrait mieux choisir de montrer $H_1$ et $H_2$ a un endroit ou le gel est horizontal, sinon le schema n'est valable qu'aux petites deformations + du coup ce sera it mieux de mettre le schema en 3b ?}.
+Since two interfaces clearly separate a protein-rich phase and a protein-poor phase, we model the resulting situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
+Wavelength being much larger than thickness, we can assume small strain in the film and lubrication in the layers.
 Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in the second panel of Fig.~\ref{fig:sideview}a.
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}b) and excess surface area of the gel film (Fig.~\ref{fig:sideview}c).
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d).
 This confirms that the instability is due to swelling.
@@
 Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses to develop while swelling: the only situation were wrinkling was observed is when both top and bottom walls are coated.
-We also confirmed that at any given \seb{point in time and space} the thickness $h$ of the film of gel is homogeneous.
-The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is \seb{therefore} not due to thickness inhomogeneities but to altitude gradients.
-\seb{Indeed} the length of the optical path along \seb{the vertical direction} $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}d.
+We also confirmed that at any given point in time and space the thickness $h$ of the film of gel is homogeneous.
+The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is therefore not due to thickness inhomogeneities but to altitude gradients.
+Indeed the length of the optical path along the vertical direction $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}b.
 
@@
 \end{equation}
-For small deflections the actual wavelength $\lambda$ is related to $\lambda^*$ and to the \seb{wrinkle} amplitude $A$ as $\lambda \sim A^{1/4} \lambda^{*3/4}$.
+For small deflections the actual wavelength $\lambda$ is related to $\lambda^*$ and to the wrinkle amplitude $A$ as $\lambda \sim A^{1/4} \lambda^{*3/4}$.
 
@@
 Prefactors are 0.63, 0.69 and 0.67 respectively.
-Continuous line is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all the points in (b).
-\seb{\it Mettre $\lambda_P$ et $\lambda_D$ comme dans le texte + $\lambda_{\rm mixed}$ ?
-Mettre tous les fits de la meme couleur (en gris ou en noir).
-Ajouter les textes ``porous model'', ``viscous model'' et ``mixed model'' (ou equivalent) sur les graphes.}}
+Continuous line is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
 	\label{fig:DarcyPoiseuille}
@@
 In a first limit where the film of gel would be sitting without sticking to the bottom wall of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
-Darcy law relates the pressure gradient, estimated as $\sigma_\perp/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
+Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_\perp/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
@@
 where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
-This expression emphasizes the kinetic nature of the wavelength selection \seb{through Darcy flow}.
+This expression emphasizes the kinetic nature of the wavelength selection hrough Darcy flow.
 However we note that the wavelength of a generation of the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup.
-Fig.~\ref{fig:growth}) once the gel is touching both walls, we can deduce the relevant $\tau$ from the growth of the  pattern (see Supplementary Method for the precise definition).
-\seb{\it La phrase precedente n'est pas assez claire et semble contredire le texte en supplementary : a reformuler.}
+Fig.~\ref{fig:growth}) once the gel is touching both walls.
+The relevant $\tau$ corresponds to the times it takes for the amplitude to saturate (see Supplementary Method for the precise definition).
 
@@
 By symmetry the transverse load across the gel $\sigma_\perp$ is also the pressure difference over the wavelength.
-Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_\perp \frac{H^3}{\eta\lambda^2}$.
+Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_\perp \frac{H^3}{\eta\lambda^2}$~\cite{Poiseuille1842}.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
@@
 The two above models yield wavelengths that are of the same order in most of the cases under study.
-Thus only systematic measurements of the experimental $\lambda$ \seb{together with estimates for} all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
+Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
 As summarized in Supplementary Table~\ref{tab:data} we systematically varied (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
@@
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
-Here \seb{the wavelength associated to the secondary} blister is taken as half of that of the primary pattern as shown in Fig.~\ref{fig:dynamics}c and $\tau$ is the variable parameter.
-\seb{\it Pas sur que la fin de cette phrase soit bien claire pour moi.}
+Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Fig.~\ref{fig:dynamics}c.
 
 Our results are displayed in Fig.~\ref{fig:DarcyPoiseuille}.
-\seb{Although they provide the correct order of magnitude for $\lambda$,} neither model alone is able to \seb{quantitatively} reproduce the full data set.
+Although they provide the correct order of magnitude for $\lambda$, neither model alone is able to quantitatively reproduce the full data set.
 However considering only the points where $H<H^*$ (respectively $H>H^*$), the Darcy (resp.
 Poiseuille) model fits the data well albeit with a prefactor.
-Note that it would be easy to oversight \seb{\it ``oversight'' $\neq$ verbe: plutot ``oversee'' ?} the failure of the Poiseuille model at small wavelengths by adding an unexplained offset \seb{(dashed line in Fig.~\ref{fig:DarcyPoiseuille}b)}.
+Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset \seb{(dashed line in Fig.~\ref{fig:DarcyPoiseuille}b)}.
 Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.
@@
 \seb{\it Ajouter une conclusion.}
+%optical elements ?
+%wallpaper ?
+%role of gravity
 
@@
 To prevent adsorption of caseins to surfaces, microscope slides and cover slips were cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush was polymerized from methacrylate groups.
-Sodium caseinate (Firmenich) was labelled with Dylight~550 NHS ESTER (Thermo Scientific) and then purified by centrifugation.
-\seb{\it vraiment ``purified''?} Unlabelled sodium caseinate powder was dissolved in deionised water (or water--glycerol mixture).
+Sodium caseinate (Firmenich) was labelled with Dylight~550 NHS ESTER (Thermo Scientific).
+Excess dye was removed by centrifugation.
+Unlabelled sodium caseinate powder was dissolved in deionised water (or water--glycerol mixture).
 To induce gelation glucono-$\delta$-lactone (GDL) in powder (Firmenich) was added to the solution.
-\SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution, \seb{stirred} and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
+\SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution, stirred and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
 3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
 We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplemental Material) or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent \seb{confocal microscopy images} or by transmitted/reflected light macroscope (Nikkon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern.
+Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern (Supplementary Figure~\ref{fig:patterns}).
 Secondary pattern wavelength is set to half of the primary wavelength, according to Fig.~\ref{fig:dynamics}c.
@@
 The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
-Precious experimental ideas came from Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
+Precious experimental ideas came from Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
 Alan Parker at Firmenich provided the casein and GDL; 

### 2015-01-22 fig_plis/article_plis.tex
Fix bibtext Allow at most 12 authors before putting "et al."

@@
 Stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002}.
-Since in practice gravity sets the wavelength in such situation~\cite{Smoluchowski1927}, the wrinkling on a confined viscous substrate has not been studied experimentally.
+Since in practice gravity sets the wavelength in such situation~\cite{Smoluchowski1910}, the wrinkling on a confined viscous substrate has not been studied experimentally.
 
@@
 Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
+Permeability is measured using a protocol found in \cite{VanDijk1986} (see Materials and Supplementary Fig.~\ref{fig:permeability}).
 
@@
 The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
-Precious experimental ideas came from Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
+Precious experimental ideas came from Thibaut Divoux, Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
 Alan Parker at Firmenich provided the casein and GDL; 

### 2015-01-22 fig_plis/article_plis.tex
Use data from ech23 and ech25 Little pore size change in half coat. Thick lines

@@
 \begin{figure}[ht!]
-	\includegraphics{dynamics}
+	\includegraphics{dynamics}%
 	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w).
@@
 Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-Scale bars are \SI{1}{\milli\metre}.}
-	\label{fig:dynamics}
+Scale bars are \SI{1}{\milli\metre}.}%
+	\label{fig:dynamics}%
 \end{figure}
@@
 \begin{figure}[b]
-	\includegraphics{acidification}
+	\includegraphics{acidification}%
 	\caption{Acid-induced protein gel properties behave nonmonotonically with pH.
@@
 (b) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator.
-(c) Evolution of pore size measured by confocal microscopy in a half-coated microscopy cell with no adhesion to top wall, syn\ae{}resis and swelling allowed.}
-	\label{fig:acidification}
+(c) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
+allowed syn\ae{}resis and swelling (orange dashed line).}%
+	\label{fig:acidification}%
 \end{figure}
@@
 Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplemental Fig.~\ref{fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
-In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplemental Fig.~\ref{fig:nonmonotonic}).
+In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplemental Fig.~\ref{fig:nonmonotonic}).
 

### 2015-01-22 fig_plis/article_plis.tex
transition au début de Models and scaling laws et résumé à la fin

@@
 
-\seb{\it Ajouter une transition: en l'etat le passage au modele est trop brutal.}
+Longitudinally constrained swelling of the gel film is the origin and driving force of the buckling instability.
+The gel area becomes in excess compared to the available area, thus the gel must fold.
+However, this does not explains how a wavelength is selected.
 
@@
 We confirmed that the former plays no significant role by successfully repeating our experiment in a cell held vertically, i.e.
+with the weight acting longitudinally rather than transversely.
+%Since the gel is almost buoyant with the surrounding medium, we expect that the weight plays no significant role, with an elasto-gravitational length at least centimetric.
+We confirmed this insight by successfully repeating our experiment in a cell held vertically, i.e.
 with the weight acting longitudinally rather than transversely.
@@
 
-\seb{\it Resumer ce qui precede en une phrase disant dans quel domaine de longueur d'onde on s'attend a trouver quel scaling pour nos manips.}
+To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$ the main dissipation mechanism would be the flow through the porous film, whereas for $H\gg H^*$ the flow in the viscous layers would dominate at set the wavelength.
 
@@
 Poiseuille) model fits the data well albeit with a prefactor.
-Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset \seb{(dashed line in Fig.~\ref{fig:DarcyPoiseuille}b)}.
+Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Fig.~\ref{fig:DarcyPoiseuille}b).
 Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.

### 2015-01-23 fig_plis/article_plis.tex
Split fig:dynamics, panel c in Supplementary. Supplementary method that describe fig:patterns and fig:generations

@@
 (a) Light transmission microscopy.
-The same area is highlighted in yellow on all panels to show that the initial pattern is not evolving.
+Successive generations of patterns are highlighted in colour to stress the absence of growth after formation.
 (b) 3D reconstruction from fluorescent confocal microscopy.
-(c) Fourier spectra of binary altitudes of (a) corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
 Scale bars are \SI{1}{\milli\metre}.}%
@@
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
-Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Fig.~\ref{fig:dynamics}c.
+Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Fig.~\ref{fig:generations}.
 
@@
 We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplemental Material) or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern (Supplementary Figure~\ref{fig:patterns}).
-Secondary pattern wavelength is set to half of the primary wavelength, according to Fig.~\ref{fig:dynamics}c.
+Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}).
+Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
 Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
 Permeability is measured using a protocol found in \cite{VanDijk1986} (see Materials and Supplementary Fig.~\ref{fig:permeability}).
-
-\seb{\it Il y a pas mal de petits soucis dans la biblio : manque des page numbers ou refs a completer. Je joins un scan de mes notes pour ajuster le .bib.}
 
@@
 
+\subsection*{Wavelength measurement}
+
+\begin{figure*}
+	\includegraphics{patterns}
+	\caption{Patterns corresponding to the samples of Table~\ref{tab:data}.
+(a-d) increasing cell thickness.
+(e-f) larger solid content.
+(g-i) increasing glycerol content.
+All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
+Scale is common to all panels (scale bar \SI{1}{\milli\metre}).
+Length of arrows correspond to the measured primary wavelength.}
+	\label{fig:patterns}
+\end{figure*}
+
+\begin{figure}
+	\includegraphics{generations}%
+	\caption{Fourier spectra of binary altitudes of Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
+Scale bars is \SI{1}{\milli\metre}.}%
+	\label{fig:generations}%
+\end{figure}
+
+To measure wrinkle wavelength, we used large field pictures of the final state.
+These picture were obtained either by stitching fluorescent microscopy images obtained at 10x magnification (Fig.~\ref{fig:patterns} except c and e), or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x (Fig.~\ref{fig:patterns}c and e).
+The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
+White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
+It is thus possible to reconstruct a binary picture of the pattern (inset of Fig.~\ref{fig:generations}).
+Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final picture to pure first generation.
+
+Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
+When only the first generation pattern is present a peak clearly indicates the dominant mode.
+When second generation has appeared, the dominant wavelength is halved and the peak much weaker.
+This allows us to use secondary pattern is our analysis by setting their wavelength at half of the primary wavelength.
+Further generations do not seem to appear simultaneously and the peak is weak and broad.
+This could be a signature of a transition to chaos of a non-linear oscillator~\cite{Brau2011}.
+
 \end{document}

### 2015-01-24 fig_plis/article_plis.tex
Cite LeGoff2014 Supplemental -> Supplementary

@@
 We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
-As shown in Fig.~\ref{fig:dynamics} and in Supplemental Movies, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
+As shown in Fig.~\ref{fig:dynamics} and in Supplementary Movies 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
 Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates.
@@
 Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
-Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplemental Fig.~\ref{fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
-In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplemental Fig.~\ref{fig:nonmonotonic}).
+Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplementary Fig.~\ref{fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
+In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{fig:nonmonotonic}).
 
@@
 where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
-This expression emphasizes the kinetic nature of the wavelength selection hrough Darcy flow.
+This expression emphasizes the kinetic nature of the wavelength selection through Darcy flow.
 However we note that the wavelength of a generation of the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup.
 Fig.~\ref{fig:growth}) once the gel is touching both walls.
+This arrest may be due to bending rigidity~\cite{LeGoff2014}.
 The relevant $\tau$ corresponds to the times it takes for the amplitude to saturate (see Supplementary Method for the precise definition).
@@
 3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
-We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplemental Material) or a 63x (oil) lens to observe the microstructure of the gel.
+We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel.
 Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}).
@@
 
-\section*{Supplemental material}
+\section*{Supplementary material}
 
@@
 	\includegraphics{generations}%
-	\caption{Fourier spectra of binary altitudes of Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
+	\caption{Fourier spectra of binary altitudes of Main Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
 Inset: reconstructed binary altitudes at time \SI{2}{\hour}.

### 2015-01-24 fig_plis/article_plis.tex
sign in permeability

@@
 (b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
-(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.
-\seb{\it Changer $z(\infty)$ en $z_\infty$ dans le schema.}}
+(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
 	\label{fig:permeability}
@@
 
-Darcy law through the gel of permeability $\alpha$ and height $H$ reads \seb{\it Attention au signe ?} 
+Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
 \begin{equation}
-\frac{dz}{dt} = \seb{-} \frac{\alpha}{\eta}\frac{\Delta P}{H}
+\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
 \end{equation}
-where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure \seb{drop through the gel, with $\rho$ the buffer density. {\it Avez-vous mesure $\rho$ independamment?}} We can thus write
+where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel.
+We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer.
+We can thus write
 \begin{equation}
-\frac{d(z_\infty-z)}{dt} = -\frac{z_\infty-z}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
+\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
 \end{equation}
 
-By fitting $z_\infty-z(t)$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
+By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
 Note that even with $H$ of the order of a few millimeters high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.

### 2015-01-24 2014_10_23_memo.tex
Correct typo
Growth rate maximisation

@@
 \usepackage{siunitx}
+\usepackage{hyperref}
 
@@
 \begin{equation}
-\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial w^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q,
+\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = q,
 \end{equation}
@@
 \begin{equation}
-B\frac{h^3}{12}\frac{\partial^4 w}{\partial w^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = -B_s \frac{\partial w}{\partial x},
+B\frac{h^3}{12}\frac{\partial^4 w}{\partial x^4} + \sigma_0 h \frac{\partial^2 w}{\partial x^2} = -B_s \frac{\partial w}{\partial x},
 \label{eq:mecheq}
@@
 k_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
+\label{eq:Poiseuillek}
 \end{equation}
@@
 \begin{equation}
-H_1^3 p_1 + H^2 p_2 = 0.
+H_1^3 p_1 + H_2^3 p_2 = 0.
 \label{eq:pressures}
@@
 \begin{equation}
-\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} + v = 0,
+\frac{\partial Q_1}{\partial x} + \frac{\partial w}{\partial t} + v = 0,
 \label{eq:conservationDarcy}
@@
 \begin{equation}
-B_s = \frac{\eta}{\tau}\frac{1}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}}.
+B_s = \frac{\eta}{\tau}\frac{1}{\frac{H^3 k^3}{12} + \frac{\alpha k}{h}}.
 \end{equation}
@@
 
+\section{Growth rate maximisation}
+
+Rather than finding the mode that minimises the stress, it is possible to find the mode that has the maximum growth rate. Disclaimer: both methods always give the same result.
+
+In the case of an elastic film on a viscous layer the mechanical equilibrium writes:
+\begin{equation}
+\left(\frac{B h^3}{12}k^4 - \sigma_0 h k^2\right) w = -\frac{12\eta}{H^3 k^2}\frac{\partial w}{\partial t},
+\end{equation}
+which is of the form $\tau \frac{\partial w}{\partial t}=w$, with
+\begin{equation}
+\tau = \frac{12^2\eta}{H^3 h (12 \sigma_0 k^4 - B h^2 k^6)}.
+\end{equation}
+
+To minimise $\tau$ we must have
+\begin{equation}
+0 = \frac{\partial}{\partial k}\left(12 \sigma_0 k^4 - B h^2 k^6\right) = 6k^3\left( 8 \sigma_0 - B h^2 k^2\right).
+\end{equation}
+Thus the minimum $\tau$ is
+\begin{equation}
+\tau_\text{min} = \frac{2\times12^2\eta}{H^3 h^3 k^6 B},
+\end{equation}
+from which we express the wave number
+\begin{equation}
+\frac{2\time 12}{k^6} = H^3 h^3 \frac{B\tau_min}{\eta} = H^3 h^3  \Upsilon,
+\end{equation}
+which is the same as Eq.~(\ref{eq:Poiseuillek}).
+
 \bibliography{../../bib/Yaourt}

### 2015-01-24 fig_plis/article_plis.tex
Modèles: pourquoi minimiser la contrainte de compression; inversion des cas limites

@@
 
-We first denoise each plane by a \seb{short-range} Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
+We first denoise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
 Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
-We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value \seb{over the given vertical slice}.
+We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line.
 Since this criterion is local, we do not need to correct for background intensity variation.
@@
 Velocity is obtained by time differentiation.
-Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus \seb{of the cell thickness} $e$ and retrospectively of $H_1$ and $H_2$.
+Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
 
@@
 	\caption{Growth of the instability.
-(a) Evolution of the gel velocity $v$ \seb{averaged} over the whole surface.
+(a) Evolution of the gel velocity $v$ averaged over the whole surface.
 Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
@@
 Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.
-\seb{\it Utiliser des symboles ($t$, $r$,$ A(t)$) plutot que des mots ?
-+ j'ai plusieurs petites questions par rapport a cette figure.}}
+\seb{\it j'ai plusieurs petites questions par rapport a cette figure.}}
 	\label{fig:growth}
@@
 However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
-Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical \seb{velocity ?} $v$ of the gel that can be integrated over the whole field of view.
+Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
 
-We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls \seb{and $v_\text{max}$ is the maximum velocity of the gel}.
+We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
 We used this approximation of $\tau$ for all our samples.
@@
 \end{equation}
-where $w(x,t)$ is the deflection of the plate along $z$.
-\seb{\it Cette formulation --stationnaire-- des equations ne semble pas coherente avec le fait de chercher une solution dependant du temps.
-Clarifier ou donner des references...}
+where $w(x,t)$ is the deflection of the plate along $z$~\cite{Biot1957}. Since we deal with slow rates of deformation, inertia is neglected.
 
-We assume an exponentially growing sinusoidal deflection $w(x,t) =  A_0 e^{t/\tau} \sin kx$ of characteristic time $\tau$ and wavevector $k$.
-\seb{\it En lien avec la remarque precedente, il est delicat de donner un sens a $\tau$ ici en l'absence de l'eq (14).} \sseb{In operator language,} Eq.~(\ref{eq:mecheq}) then reads
+We assume that the deflection can be decomposed in exponentially growing sinusoidal modes and we consider one of them $w(x,t) =  A_0 e^{t/\tau} \sin kx$ of characteristic time $\tau$ and wavevector $k$.
+Eq.~(\ref{eq:mecheq}) then reads
 \begin{equation}
@@
 
+Following \cite{Biot1957}, we consider the case where the compression load $\sigma_\parallel$ is increased quasistatically.
+The observed mode is the one that arises at the lowest possible $\sigma_\parallel$, i.e.
+the mode minimising $\sigma_\parallel$.
+
 \subsubsection*{Elastic film in an infinite elastic medium}
 
-\seb{\it Vraiment utile de detailler ce cas qui ne correspond pas aux manips ?}
-
-\seb{In the case of an elastic film lying over an infinite elastic medium,} the deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
+In this framework, we recall here the well known case of an elastic film lying over an infinite elastic medium.
+The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
 \begin{equation}
@@
 
-Minimizing $\sigma_\parallel$ as a function of $k$ yields the well-known result $hk = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is \seb{\it Il serait bien de rappeler pourquoi on minimise ici $\sigma_\parallel$ : cela ne vient-il pas plus de considerations energetiques que de la selection du mode le plus instable dans une relation de dispersion ?
-Les deux visions sont-elles equivalentes ?
-Je n'ai pas les idees bien claires la-dessus --c'est sans doute relie au pb deja evoque plus haut-- et le lecteur moyen risque de se poser la question de comment on justifie proprement cette minimisation\dots}
+Minimizing $\sigma_\parallel$ as a function of $k$ yields the well-known result $hk = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
 \begin{equation}
-\lambda_d = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
+\lambda = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
 \label{eq:lambdaElEl}
@@
 
-\seb{\it Anecdotique : pourquoi $\lambda_d$ plutot que $\lambda$ ?}
-
 \subsubsection*{Elastic film on a viscous layer}
-\seb{Let us now turn to the case of an elastic, impermeable film on a viscous layer which is more relevant for the present experimental situation.} Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation.
-\seb{This corresponds to the case where the gel film is close to one of the cell walls and this specifically applies to secondary blisters.} Further considering only small deflections, we may neglect the displacement of the plate along $x$.
+Let us now turn to the case of an elastic, impermeable film on a single viscous layer.
+As \cite{Huang2002} we neglect gravity, although it is difficult to do so in practice at a free interface.
+Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation.
+Further considering only small deflections, we may neglect the displacement of the plate along $x$.
 The flux along $x$ is thus of the Poiseuille form
@@
 \end{equation}
-Keeping only linear order terms, one gets \sseb{in operator language}
+Keeping only linear order terms, one gets
 \begin{equation}
@@
 
+We note that the mode minimising $\sigma_\parallel$ is also maximising the growth rate
+\begin{equation}
+\frac{1}{\tau} = \frac{H^3k^2}{12^2\eta}\left(12\sigma_\parallel - Bh^3k^4\right).
+\end{equation}
+
 When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of Eq.~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness 
@@
 \end{equation}
+This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface the gravity can indeed be neglected.
 
 \subsubsection*{Porous elastic film with no possible Poiseuille flow}
-\seb{Another limiting case is that of a porous elastic film sandwiched between two viscous layers that are supposed to be large enough so that the previous Poiseuille mechanism can be neglected.} The transverse load derives from the Darcy law through the film of permeability $\alpha$:
+Another limiting case is that of a porous elastic film immersed in a viscous medium where longitudinal resistance to flow is infinite.
+It is the case in our experiments when the film is at contact with one of the walls.
+The only way to deform the film is to flow some liquid through the film of permeability $\alpha$.
+The transverse load derives from the Darcy law:
 \begin{equation}
@@
 \end{equation}
+Here also the dominant mode maximises the growth rate.
 
 \subsubsection*{Porous elastic film between two viscous layers}
-\seb{Finally the most realistic scenario mixes both Poiseuille and Darcy mechanisms, i.e.
+Finally the most realistic scenario mixes both Poiseuille and Darcy mechanisms, i.e.
 considers a porous elastic film between two viscous layers of arbitrary thickness.
-In this case,} the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
+In this case, the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
 \begin{equation}
@@
 
-Analytic minimization of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but \seb{cumbersome}.
-Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
+Analytic minimization of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome. Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
 \begin{align}

### 2015-01-25 plis_figures.tex
Split fig:dynamics, panel c in Supplementary. Supplementary method that describe fig:patterns and fig:generations
Insert nonmonotonic Do not use cycle list name=earthy Add time arrow on I(q)

@@
 
+\begin{figure}
+	\tikzsetnextfilename{generations}
+	\input{fig_plis/generations.pgf}%
+	\caption{Fourier spectra of binary altitudes of Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
+Scale bars is \SI{1}{\milli\metre}.}%
+	\label{fig:generations}%
+\end{figure}
+
+\begin{figure*}
+	\tikzsetnextfilename{nonmonotonic}
+	\input{fig_plis/nonmonotonic.pgf}%
+	\caption{Microstructure evolution.
+(a) Details of confocal images in fully adhesive conditions (left) and with no adhesion on top (right).
+Scale bar is \SI{10}{\micro\metre}.
+Arrows indicates the pore size.
+(b) Corresponding Fourier spectra in second situation and same times.
+Lines are fits by Eq.~\ref{eq:fractalS}.
+(c-d) Evolution of susceptibility and fractal dimension in both situations (black line and orange dashed line respectively).
+Symbols indicates the times of the images in (a).}%
+	\label{fig:nonmonotonic}%
+\end{figure*}
+
 \end{document}

### 2015-01-25 fig_plis/article_plis.tex
tab:data in main text remplace fig:mesh par fig:nonmonotonic

@@
 
+\begin{table*}
+\begin{tabular}{SSSScSrrScSSSScSSSSSS}
+%\toprule
+\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_{xp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
+{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+%ech2
+4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
+& & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
+4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
+4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
+4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
+& & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
+& & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
+8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
+& & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
+\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
+4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
+4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
+\end{tabular}
+\caption{Characteristics of the samples discussed in the main text. Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.}
+\label{tab:data}
+\end{table*}
+
 The two above models yield wavelengths that are of the same order in most of the cases under study.
 Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
-As summarized in Supplementary Table~\ref{tab:data} we systematically varied (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
+As summarized in Table~\ref{tab:data} we systematically varied (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
 Finally, the only parameter that we had no control upon was the initial altitude $H$ of the gel before wrinkling that varied randomly.
@@
 
-\begin{table*}
-\begin{tabular}{SSSScSrrScSSSScSSSSSS}
-%\toprule
-\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_{xp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
-{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-%ech2
-4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
-& & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
-4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
-4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
-4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
-& & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
-& & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
-8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
-& & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
-4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
-4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
-\end{tabular}
-\caption{Characteristics of the samples discussed in the main text. Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.}
-\label{tab:data}
-\end{table*}
-
 \subsection*{Microstructure measurement}
 
-\begin{figure}
-	\includegraphics{mesh}
-	\caption{Gel microstructure.
-(a) Detail of a confocal micrograph of a 4\%w casein in water acidified with 4\%w GDL, when only the microscope slide is coated, not the cover slip.
-Images were taken at times corresponding to the minimum in mesh size.
+\begin{figure*}
+	\includegraphics{nonmonotonic}%
+	\caption{Microstructure evolution.
+(a) Details of confocal images in fully adhesive conditions (left) and with no adhesion on top (right).
 Scale bar is \SI{10}{\micro\metre}.
-(b) Corresponding radially averaged Fourier transform.
-\seb{\it Fourier transform or Fourier spectrum or power spectrum?
-Clarify?} Black line is the best fit by Eq.~(\ref{eq:fractalS}) with $\xi=\SI{2.0}{\micro\metre}$ and $d_{3D}=2.27$.
-Inset shows the center of the \sseb{non radial average} spectrum \seb{using a} logarithmic color scale.
-Scale bar is \SI{10}{\per\micro\metre}.
-(c-d) Same as (a-b) in a 50\%w \seb{glycerol--water mixture leading to} $\xi=\SI{0.3}{\micro\metre}$ and $d_{3D}=3$.}
-	\label{fig:mesh}
-\end{figure}
+Arrows indicates the pore size.
+(b) Corresponding Fourier spectra in second situation and same times.
+Lines are fits by Eq.~\ref{eq:fractalS}.
+(c-d) Evolution of susceptibility and fractal dimension in both situations (black line and orange dashed line respectively).
+Symbols indicates the times of the images in (a).}%
+	\label{fig:nonmonotonic}%
+\end{figure*}
 
-To follow the evolution of the microstructure in a gel undergoing syn\ae{}resis and swelling, it was necessary to keep it into the thin focal plane of the 63x objective lens of the confocal microscope.
-To do so, we coated only the microscope slide with acrylamide brush to allow syn\ae{}resis on this side.
+To follow the evolution of the microstructure of the gel, we acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:nonmonotonic}a).
+After Hann windowing we performed Fourier transform and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:nonmonotonic}b) that was fitted by the following fractal form:
+\begin{equation}
+I(q) = \frac{2\chi\Gamma(d)}{\left(1+\left((d+1)\xi q\right)^2\right)^{d/2}}
+\label{eq:fractalS}
+\end{equation}
+where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension.
+$\Gamma$ is the gamma function.
+
+Figure~\ref{fig:acidification}c of the main text and Supplementary Fig.~\ref{fig:nonmonotonic}c and d show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
+The later situation is a proxy to the situation where we observe wrinkling, since it was necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
+To do so, we coated only the microscope slide with acrylamide brush.
 The cover slip was cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
@@
 
-We acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:mesh}a and c).
-After Hann windowing we performed Fourier transform (see insets of Fig.~\ref{fig:mesh}b and d) and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:mesh}b and d) that was fitted by the following fractal form \seb{\it Clarify $I(q)$ vs $S(q)$}:
-\begin{equation}
-I(q) = \frac{\chi}{1+(\xi q)^\gamma}
-\label{eq:fractalS}
-\end{equation}
-where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and the exponent $\gamma$ is loosely related to the fractal dimension $d_{3D}$.
+In situation (i), the gel keeps its topology throughout the gelation and over acidification.
+Left of Fig.~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again.
+Syn\ae{}resis and swelling are thus local processes.
+By contrast in situation (ii), the gel as a whole compresses while forming.
+Since our focal plane is a fixed depth, we cannot follow the topology of the same subset of the network, however the pore size is clearly smaller than in situation (ii).
+The isoelectric gel, at minimum gel volume, is probably isotropic or only marginally anisotropic in $z$.
+This explains that subsequent swelling is not occurring only along $z$ but also in plane, causing buckling.
 
-\seb{\it Actualiser la fin de cette partie avec nouvelle Eq. (6) et texte correspondant.}
+$\xi$ was also systematically measured in the same way at the end of each wrinkling experiment and reported in table~\ref{tab:data}.
+%Final fractal dimension was always close to 2 within error bars.
+Comparing $\chi$ across different setting of the microscope is not relevant.
 
@@
 	\includegraphics{patterns}
-	\caption{Patterns corresponding to the samples of Table~\ref{tab:data}.
+	\caption{Patterns corresponding to the samples of Table~\ref{tab:data} of the main text.
 (a-d) increasing cell thickness.

### 2015-01-25 fig_plis/article_plis.tex
Réorganise les Supplementary pour coller à l'ordre du texte principal

@@
 Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in the second panel of Fig.~\ref{fig:sideview}a.
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d).
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods).
 This confirms that the instability is due to swelling.
@@
 
-\begin{figure*}
+\begin{figure}
 	\includegraphics{nonmonotonic}%
 	\caption{Microstructure evolution.
-(a) Details of confocal images in fully adhesive conditions (left) and with no adhesion on top (right).
+(a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row).
 Scale bar is \SI{10}{\micro\metre}.
@@
 	\label{fig:nonmonotonic}%
-\end{figure*}
+\end{figure}
 
@@
 
-\subsection*{Permeability measurements}
-\begin{figure*}
-	\includegraphics{permeability}
-	\caption{Permeability measurements.
-(a) Schematic representation of the experiment.
-(b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
-Black line is the best exponential fit $Ae^{-t/\tau}$.
-(b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
-(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
-	\label{fig:permeability}
-\end{figure*}
-
-To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}.
-First, the gel of interest is formed at the lower extremity of a thin glass tube.
-The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition).
-Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4.
-The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
-Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
-A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
-Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.
-
-Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
-\begin{equation}
-\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
-\end{equation}
-where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel.
-We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer.
-We can thus write
-\begin{equation}
-\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
-\end{equation}
-
-By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
-Note that even with $H$ of the order of a few millimeters high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
-Obtaining $z_\infty$ without tube 2 would be time prohibitive.
-
 \subsection*{Confocal images processing}
@@
 Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
-
-\subsection*{Characteristic time measurement}
-\begin{figure}
-	\includegraphics{growth}
-	\caption{Growth of the instability.
-(a) Evolution of the gel velocity $v$ averaged over the whole surface.
-Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
-Black line is the best exponential fit with $\tau=\SI{90}{\second}$.
-(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
-Note that the longitudinal size of the blister evolves in this regime.
-Thick black curve is the profile at much longer times.
-Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.
-\seb{\it j'ai plusieurs petites questions par rapport a cette figure.}}
-	\label{fig:growth}
-\end{figure}
-
-In most of our samples it was possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
-However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
-Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
-
-We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
-We used this approximation of $\tau$ for all our samples.
-This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
 
@@
 
+\begin{figure}
+	\includegraphics{generations}%
+	\caption{Fourier spectra of binary altitudes of Main Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
+Scale bars is \SI{1}{\milli\metre}.}%
+	\label{fig:generations}%
+\end{figure}
+
 \begin{figure*}
@@
 \end{figure*}
-
-\begin{figure}
-	\includegraphics{generations}%
-	\caption{Fourier spectra of binary altitudes of Main Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-Scale bars is \SI{1}{\milli\metre}.}%
-	\label{fig:generations}%
-\end{figure}
 
@@
 
+\subsection*{Characteristic time measurement}
+\begin{figure}
+	\includegraphics{growth}
+	\caption{Growth of the instability.
+(a) Evolution of the gel velocity $v$ averaged over the whole surface.
+Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
+Black line is the best exponential fit with $\tau=\SI{90}{\second}$.
+(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
+Note that the longitudinal size of the blister evolves in this regime.
+Thick black curve is the profile at much longer times.
+Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.
+\seb{\it j'ai plusieurs petites questions par rapport a cette figure.}}
+	\label{fig:growth}
+\end{figure}
+
+In most of our samples it was possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
+However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
+Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
+
+We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
+We used this approximation of $\tau$ for all our samples.
+This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
+
+\subsection*{Permeability measurements}
+\begin{figure*}
+	\includegraphics{permeability}
+	\caption{Permeability measurements.
+(a) Schematic representation of the experiment.
+(b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
+Black line is the best exponential fit $Ae^{-t/\tau}$.
+(b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
+(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
+	\label{fig:permeability}
+\end{figure*}
+
+To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}.
+First, the gel of interest is formed at the lower extremity of a thin glass tube.
+The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition).
+Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4.
+The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
+Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
+A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
+Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.
+
+Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
+\begin{equation}
+\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
+\end{equation}
+where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel.
+We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer.
+We can thus write
+\begin{equation}
+\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
+\end{equation}
+
+By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
+Note that even with $H$ of the order of a few millimeters high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
+Obtaining $z_\infty$ without tube 2 would be time prohibitive.
+
 \end{document}

### 2015-01-26 fig_plis/article_plis.tex
Typos

@@
 In slit geometry, we slowly acidify a dispersion of casein, one of the milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Bremer1989} --this is a well-known effect used in cheese or yogurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit wall is turned off.
-Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimeter scale.
+As a result, not only does the casein form a gel as expected~\cite{Bremer1989} --this is a well-known effect used in cheese or yogurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
+Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
 We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
@@
 \end{figure}
-
-%discuter la figure 1c
 

### 2015-01-26 fig_plis/article_plis.tex
Placement de figures
Indique à quoi correspond chaque pic de fig:growth

@@
 
-\begin{figure}[b]
+\begin{figure}
 	\includegraphics{acidification}%
@@
 
-\begin{figure}
+\begin{figure}[b]
 	\includegraphics{nonmonotonic}%
@@
 
-$\xi$ was also systematically measured in the same way at the end of each wrinkling experiment and reported in table~\ref{tab:data}.
+$\xi$ was also systematically measured in the same way at the end of each wrinkling experiment and reported in Table~\ref{tab:data} of the main text.
 %Final fractal dimension was always close to 2 within error bars.
@@
 Thick black curve is the profile at much longer times.
-Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.
-\seb{\it j'ai plusieurs petites questions par rapport a cette figure.}}
+Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
 	\label{fig:growth}

### 2015-01-26 fig_plis/article_plis.tex
Dans les modèles le nombre d'onde k-> q

@@
 
-We assume that the deflection can be decomposed in exponentially growing sinusoidal modes and we consider one of them $w(x,t) =  A_0 e^{t/\tau} \sin kx$ of characteristic time $\tau$ and wavevector $k$.
+We assume that the deflection can be decomposed in exponentially growing sinusoidal modes and we consider one of them $w(x,t) =  A_0 e^{t/\tau} \sin qx$ of characteristic time $\tau$ and wavevector $q$.
 Eq.~(\ref{eq:mecheq}) then reads
 \begin{equation}
-\left(\frac{1}{12} B h^3 k^4 - \sigma_\parallel h k^2\right)w = \sigma_\perp,
+\left(\frac{1}{12} B h^3 q^4 - \sigma_\parallel h q^2\right)w = \sigma_\perp,
 \label{eq:mecheqop}
@@
 In this framework, we recall here the well known case of an elastic film lying over an infinite elastic medium.
-The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -k B_s w$, and thus the compression load writes
+The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -q B_s w$, and thus the compression load writes
 \begin{equation}
-\sigma_\parallel = B \frac{h^2}{12} k^2 + \frac{B_s}{hk}.
+\sigma_\parallel = B \frac{h^2}{12} q^2 + \frac{B_s}{hq}.
 \label{eq:sigma0}
@@
 
-Minimizing $\sigma_\parallel$ as a function of $k$ yields the well-known result $hk = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
+Minimizing $\sigma_\parallel$ as a function of $q$ yields the well-known result $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
 \begin{equation}
@@
 \begin{equation}
-\frac{12\eta}{\tau} \frac{w}{H^3k^2} = -\sigma_\perp.
+\frac{12\eta}{\tau} \frac{w}{H^3q^2} = -\sigma_\perp.
 \label{eq:PoiseuilleLoad}
@@
 \begin{equation}
-\sigma_\parallel = \frac{1}{12}B h^2 k^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h k^4}
+\sigma_\parallel = \frac{1}{12}B h^2 q^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h q^4}
 \label{eq:sigma0P}
 \end{equation}
-and is minimized at $k_P$ such that
+and is minimized at $q_P$ such that
 \begin{equation}
-k_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
-\label{eq:kP}
+q_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
+\label{eq:qP}
 \end{equation}
@@
 \begin{equation}
-\frac{1}{\tau} = \frac{H^3k^2}{12^2\eta}\left(12\sigma_\parallel - Bh^3k^4\right).
+\frac{1}{\tau} = \frac{H^3q^2}{12^2\eta}\left(12\sigma_\parallel - Bh^3q^4\right).
 \end{equation}
@@
 \begin{equation}
-\sigma_\parallel = \frac{1}{12}B h^2 k^2 + \frac{\eta}{\tau}\frac{1}{\alpha k^2}
+\sigma_\parallel = \frac{1}{12}B h^2 q^2 + \frac{\eta}{\tau}\frac{1}{\alpha q^2}
 \label{eq:sigma0D}
 \end{equation}
-and is minimized at $k_D$ such that
+and is minimized at $q_D$ such that
 \begin{equation}
-k_D^4 = \frac{12}{h^2\alpha\Upsilon}.
-\label{eq:kD}
+q_D^4 = \frac{12}{h^2\alpha\Upsilon}.
+\label{eq:qD}
 \end{equation}
@@
 \begin{equation}
-\sigma_\perp = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 k^2}{12} + \frac{\alpha}{h}},
+\sigma_\perp = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 q^2}{12} + \frac{\alpha}{h}},
 \label{eq:MixtLoad}
@@
 Note that we recover Eq~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and Eq.~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
-We can recast this intuition in terms of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $k_P$ and $k_D$:
+We can recast this intuition in terms of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $q_P$ and $q_D$:
 \begin{equation}
@@
 \begin{align}
-\frac{12\sigma_\parallel}{k_D^2 h^2 B} &= \left(\frac{k}{k_D}\right)^2 + \frac{1}{2\frac{k_D^2}{k_P^6}k^4  + \left(\frac{k}{k_D}\right)^2}\\
+\frac{12\sigma_\parallel}{q_D^2 h^2 B} &= \left(\frac{q}{q_D}\right)^2 + \frac{1}{2\frac{q_D^2}{q_P^6}q^4  + \left(\frac{q}{q_D}\right)^2}\\
 &= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}

### 2015-01-26 fig_plis/article_plis.tex
Check captions: "The", permutation of panels tab:data fix notations

@@
 	\caption{Measurements and modeling.
-(a) Confocal ($x$,$z$) cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
-Scale bar is \SI{100}{\micro\metre} (real size ratio).
-Evolution of (b) volume of the gel phase relative to cell volume and (c) excess area measured by confocal microscopy.
-Crosses correspond to times in (a).
-(d) Schematic side view of the cell.
+(a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
+The scale bar is \SI{100}{\micro\metre} (real size ratio).
+(b) Schematic side view of the cell.
 The constant thickness gel film (orange) is surrounded by water.
 Yellow lines show the path of transmitted light through the gel phase.
-A region of interest is highlighted in gray.}
+A region of interest is highlighted in gray.
+(c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
+Crosses correspond to times in (a).}
 	\label{fig:sideview}
@@
 Dots come from primary pattern, squares from secondary blisters.
-Dashed lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
+Lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
 Prefactors are 0.63, 0.69 and 0.67 respectively.
-Continuous line is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
+The dashed line is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
 	\label{fig:DarcyPoiseuille}
@@
 \cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_{xp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
+{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
 {\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
@@
 \end{tabular}
-\caption{Characteristics of the samples discussed in the main text. Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.}
+\caption{Fully characterised samples at the basis of Fig.~\ref{fig:DarcyPoiseuille}.
+Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.
+`cas', `GDL' and `gly' indicate the weight fraction of sodium caseinate, GDL and glycerol respectively.
+$\xi$ is the pore size at the end of each experiment.}
 \label{tab:data}
@@
 (a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row).
-Scale bar is \SI{10}{\micro\metre}.
+The scale bar is \SI{10}{\micro\metre}.
 Arrows indicates the pore size.
@@
 Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-Scale bars is \SI{1}{\milli\metre}.}%
+The scale bar is \SI{1}{\milli\metre}.}%
 	\label{fig:generations}%
@@
 All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
-Scale is common to all panels (scale bar \SI{1}{\milli\metre}).
+The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
 Length of arrows correspond to the measured primary wavelength.}
@@
 Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
-Black line is the best exponential fit with $\tau=\SI{90}{\second}$.
+The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
 (b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
 Note that the longitudinal size of the blister evolves in this regime.
-Thick black curve is the profile at much longer times.
+The thick black curve is the profile at much longer times and is at lower altitude due to swelling.
 Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
@@
 (b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
-Black line is the best exponential fit $Ae^{-t/\tau}$.
+The black line is the best exponential fit $Ae^{-t/\tau}$.
 (b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.

### 2015-01-26 fig_plis/article_plis.tex
Justifie rapidement la forme fractale

@@
 Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
-Permeability is measured using a protocol found in \cite{VanDijk1986} (see Materials and Supplementary Fig.~\ref{fig:permeability}).
+Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}).
 
@@
 Arrows indicates the pore size.
-(b) Corresponding Fourier spectra in second situation and same times.
+(b) Corresponding Fourier spectra when no adhesion on top and same times.
 Lines are fits by Eq.~\ref{eq:fractalS}.
-(c-d) Evolution of susceptibility and fractal dimension in both situations (black line and orange dashed line respectively).
+(c-d) Evolution of susceptibility and fractal dimension in full adhesive (black line) and no adhesion on top (orange dashed line) situations.
 Symbols indicates the times of the images in (a).}%
@@
 $\Gamma$ is the gamma function.
+Eq.~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
 

### 2015-01-26 fig_plis/article_plis.tex
Ajoute  un colonne # à tab:data pour identifier les échantillons Ordre de grandeur des 3 contraintes transverses possibles.

@@
 \begin{table*}
-\begin{tabular}{SSSScSrrScSSSScSSSSSS}
+\begin{tabular}{cSSSScSrrScSSSScSSSSSS}
 %\toprule
-\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
-{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
+&\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
+\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
+&{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
 %ech2
-4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
-& & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
-4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
-4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
-4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
-& & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
-& & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
-8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
-& & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
-\cline{1-4} \cline{6-9} \cline{11-14} \cline{16-21}\\[-2ex]
-4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
-4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
-4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
+a & 4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
+& & & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
+b & 4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
+c & 4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
+d & 4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
+& & & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
+e & 8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
+& & & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
+f & 8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
+& & & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
+g & 4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
+h & 4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
+i & 4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
 \end{tabular}
@@
 
-The two above models yield wavelengths that are of the same order in most of the cases under study.
+For a typical gel sheet (Table~\ref{tab:data}b) the buoyancy after syn\ae{}resis is estimated as $\Delta\rho=\SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$.
+
+Darcy and Poiseuille models yield wavelengths that are of the same order in most of the cases under study.
 Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
@@
 Eq.~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
+%En 3D on peut citer \cite{Martin1987,Chen1986}
 
@@
 \end{equation}
-This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface the gravity can indeed be neglected.
+This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface gravity can indeed be neglected.
 

### 2015-01-26 fig_plis/article_plis.tex
Amplitude dépendante du temps dans un deuxième temps

@@
 epithelimum or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g.
-lymph, blood or mucus, a practical situation were a thin viscous substrate would set the wrinkling wavelength.
+lymph, blood or mucus, a practical situation where a thin viscous substrate would set the wrinkling wavelength.
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
@@
 
-We assume that the deflection can be decomposed in exponentially growing sinusoidal modes and we consider one of them $w(x,t) =  A_0 e^{t/\tau} \sin qx$ of characteristic time $\tau$ and wavevector $q$.
-Eq.~(\ref{eq:mecheq}) then reads
+We decomposed the deflection in sinusoidal modes $w(x,t) =  A \sin qx$ of wavevector $q$. Eq.~(\ref{eq:mecheq}) then reads
 \begin{equation}
@@
 \end{equation}
-Keeping only linear order terms, one gets
+
+Assuming an exponentially growing amplitude $A(t) = A_0 \exp(t/\tau)$ and keeping only linear order terms, one gets
 \begin{equation}

### 2015-02-12 fig_plis/article_plis.tex
Modifs de Thomas: - ajoute abstract et conclusion - fait passer le détail de la littérature des plis de l'intro à "Models and scaling laws" - pattern formation -> system design. Explication en profondeur de la non-monotonicité de la solubilité de la caséine et ses conséquences. - dynamics of the wrinkling process: meilleure description de fig:sideview - utilise directement lambda et non lambda* - refait passer tab:data en Supplementary - Remise en ordre des Method summary - Titration - Sample characteristics pour introduire tab:data - Supplementary Movies

@@
 \begin{abstract}
-\seb{\it Ajouter l'abstract.}
+The slow acidification of milk proteins such as casein is well known to result in the formation of gels – this is indeed one of the key steps to make cheese or yogurt.
+Here, using the casein proteins as  building blocks and the acidifier  Glucono delta-lactone, we design model thin bio-gel layers that wrinkles in cascade.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gels shrink and swell creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel.
+We have develop a hydrodynamic model that predict correctly the wrinkling wavelength.
+Our results should be universal for proteins as it is based on pH induced charge stabilasion/destablisation and control of boundary conditons.
+As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical puposes.
 \end{abstract}
@@
 
-There are many ways and reasons for a film to wrinkle.
+%There are many ways and reasons for a film to wrinkle.
 An elastic film buckles due to excess area with respect to its boundaries.
@@
 
-Yet, in a biological context, thin tissues, e.g.
+There are many ways and reasons for a film to wrinkle.
+An elastic film buckles due to excess area with respect to its boundaries and wrinkles when buckling is hindered by a substrate.
+The selection of the preferred wrinkling wavelength is peculiar to the situation and is related to the mismatch of elastic properties between the surface and the substrate~\cite{Gough1940, Bijlaard1946} XXX.
+On the biology side wrinkling-controlled morphogenesis is ubiquitous.
+Ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}.
+Difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}.
+Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
+On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear behaviour~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
+Such patterns can be triggered by temperature dilation refXXX, swelling refXXX or the removal of pre-strain refXXX.
+
+%ph swelling: hu_1998_Polymer gels with engineered environmentally and kim_2010_Dynamic display of biomolecular patterns
+%compressive stress: kim_2011_Hierarchical folding of elastic membranes under
+
+Yet,  benchmark experiments that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line, not only from a physics point of view, but also to get fundamental insights into biologically relevant situation.
+Indeed, stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002} and remains theoretical because, on a free interface, gravity dominates over viscous substrate.
+In a biological context, thin tissues, e.g.
 epithelimum or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g.
@@
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
-Therefore benchmark experiments that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line to get fundamental insights into such a biologically relevant situation.
 
@@
 
-The main goal of the present study is thus to identify the selection mechanism of the wavelength \seb{and to discuss its implication for practical situations (?)}.
+The main goal of the present study is thus to identify the selection mechanism of the wavelength.
 In the following we first describe the physical chemistry that allows us to form such patterns.
@@
 
-\section*{Pattern formation}
+\section*{System design}
 
@@
 
-The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation~\cite{Bremer1989} (Fig.~\ref{fig:acidification}a).
+To form wrinkled patterns such as the one oberved in figure 1XXX, ones needs to take advantage of the pH sensitivity of casein to control the route to gelation, to fine tune the gel properties and the boundary conditions.
+
+the stability of casein dispersion strongly depend on the pH of the solution.
+Supplementary movie 3XXX shows the titration of a casein dispertion at 1\%w by 1 molar HCl.
+At pH=7, the casein is well dispersed then it floculate between pH= XXX to XXX and finaly at pH=XXX the solution become clear again.
+However as one increases the casein concentration up to 4\%w, at low pH the casein redispersion is not total, Supplementary movie 4XXX.
+The state diagram of casein as a function of the pH  is related to charge stabilization.
+At neutral pH casein are charged negatively and repel each other.
+As the pH decreases to the isoelectric point of the caseine (pH=4.5), caseine molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability].
+Beyond the isoelectric point casein acquire a positive net charge and therefore partialy re-disolve.
+In figure 2cXXX, we quantify this effect and measure with UV absorption spectroscopie the amount of free casein within the 4\%w dispersion as a function of pH (see Supplementary Methods).
+Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction is 1.
+Upon decreasing the pH of the solution to the isolectric point the free casein fraction drop to almost 0, almost all the casein molecule are aggregated.
+Further decrease of the pH of the solution lead to a partial redisolution of the casein agregate.
+At pH=2 the fraction of free casein is XXX.
+
+%On the Stability of Casein Micelles, Pieter Walstra J. of dairy science 1990
+
+To form a space spaning gel, floculation need to be prevented.
+We use of Glucono delta-lactone (GDL) a molecule  that slowly hydrlolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, fig.2aXXX.
+This GDL controled acidification permits the formation of a bulk gel phase.
+We monitor the GDL induced gelation with a rheometer in cone plate geometry.
+As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elatic and viscous modulus of the casein dispersion as a function of time.
+the mechanical properties of the gel are closely related to the fraction of free casein.
+First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finaly the elastic modulus decreases down to 100 Pa.
+
+To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy.
+As shown in supplementary fig.
+1XXX the gel consist in a space spaning fractal network with a curtoff length scale xi.
+xi corresponds to the maximum pore size of the network.
+This model fits the structure at all times but its parameters vary non monotoneously during GDL acidification.
+As shown in Fig.
+2cXXX, up to the isolectric point xi increases up to a maximum length of 7 $\mu$m due to the solvent expulsion out of the casein network [refXXX].
+This stiffen the arm of the network and is responsible for the increase of elasticity.
+Beyond the isoelectric point xi decreases down to 4$\mu$m.
+Casein regain solubility [refXXX], the strand network swell, the pore shrinks and the gel weakens.
+
+As seen on the confocal measurements, the gel is porous and is permeable to the solvant.
+This protertie is determinant as it allow the gel to absord and expel the solvant and therefore change its volume when sticky boundary are turned off.
+To quantify this effect, we have embeded the casein gel inside a capillary and deeped verticaly the capilary in an aqueous buffer [refXXX].
+In the supplementary figure XXX, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces.
+Exploitation of this experiment give the permeability of the gel: 67000 nm$^2$.
+
+Boundary conditions is the final parameter that is essential to form wrinkled patterns.
+The evolution of xi when sticky boundaries are on reflects the inner constrains of the gel:  xi increases, the stand of the nework become thinner and the gel wants to retract; then xi decreases, the stands of network swell and the gel wants to expend.
+When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detach from the top wall.
+In this situation we can monitor xi, fig.
+XXX.
+xi remains mostly constant and the major effect is repported on the volume occupied by the gel as will be repported below.
+
+%We coat the glass slides with acrylamide brushes to prevent the adhesion of the casein.
+In fig.2c, when the top slide of the optical cell is coated with acrylamide brushes, we then observe that the casein network detaches from the top slide.
+xi does not peak at the isoelectric point but remains constant
+
+%Boundary conditions is the final parameter that is essential to form wrinkled patterns.
+Once the space spanning network is formed and sticks to the container, protein attraction can only proceed by local shrinking of the protein agregates, leading to larger pores.
+
+%The difference in value of xi between the fully un-coated  and the half-coated optical cell demonstrate that the gel is subject to constrains as the pH decreases.
+The relaxation of thoses constrain is key to the wrinkling formation when both the top and bottom slide are coated with acrylamide brushes.
+
+%on a change les condition daccroche ce aui permet au contrainte de se relacher
+
+%The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation~\cite{Bremer1989} (Fig.~\ref{fig:acidification}a).
 Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
@@
 
-Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated.
-Starting from a homogeneous solution (first panel Fig.~\ref{fig:sideview}a where black margins correspond to the glass walls), the gel forms and simultaneously expels solvent on both sides.
-Since two interfaces clearly separate a protein-rich phase and a protein-poor phase, we model the resulting situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
-Wavelength being much larger than thickness, we can assume small strain in the film and lubrication in the layers.
-Relative increase in protein concentration during syn\ae{}resis can be deduced from the brighter fluorescence in the second panel of Fig.~\ref{fig:sideview}a.
+\section*{Dynamics of the wrinkling process}
+
+Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favorable to wrinkling.
+Up to 10 min the dispersion is homogeneous and fills the entire slit.
+The black margins correspond to the glass walls.
+Around 23 min, the gel forms and simultaneously becomes thiner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the situation at 10min.
+This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expeled on both the upper and lower sides.
+We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
 Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods).
+As expected from the system design, the evolution of the volume is non-monotoneous.
+After reaching a minimum around 35\% of its initial value at 23 min, the volume increases.
+Shortly after,  at 36 min, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area.
 This confirms that the instability is due to swelling.
+Around 38 min, the amplitude of the wrinkles reach the thickness of the slit.
+The size of the primary pattern is now settled Supplementary Fig.
+\ref{fig:patterns, fig:growth}.
+However swelling continues, the gel flattens on both top and bottom walls (44 min) and finally buckles back (53 min) to form the secondary pattern (1h21).
+When the excess volume saturates, the dynamics stops.
+The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we have observe the formation of higher order patterns.
 
-Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses to develop while swelling: the only situation were wrinkling was observed is when both top and bottom walls are coated.
+Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses to develop while swelling: the only situation were wrinkling was observed is when both top and bottom walls are coated with acrylamide brushes.
 We also confirmed that at any given point in time and space the thickness $h$ of the film of gel is homogeneous.
@@
 
-Longitudinally constrained swelling of the gel film is the origin and driving force of the buckling instability.
+%Longitudinally constrained swelling of the gel film is the origin and driving force of the buckling instability.
 The gel area becomes in excess compared to the available area, thus the gel must fold.
@@
 
-The characteristic length $\lambda^*$ of a wrinkling pattern is set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$~\cite{Vella2009,Kolinski2009}:
+The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem.
+Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations  refLaudauXXX,  essentially impossible to solve in analytical form.
+However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinking patern and isolate the physically meaningful ingredients.
+An elastic film buckles due to excess area with respect to its boundaries.
+Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}.
+$\lambda$ is then set by the competition between the flexural modulus of the film $B=E/(1-\nu^2)$ ($E$ Young modulus, $\nu$ Poisson ratio) and a transverse load $\sigma_\perp$.
+For small deflections of  amplitude $A$, $\lambda$ scales as~\cite{Vella2009,Kolinski2009}:
+%
 \begin{equation}
-\lambda^* \sim h \left(\frac{B}{\sigma_\perp}\right)^{1/3}.
+\lambda \sim \left( A h^3 \frac{B}{\sigma_{\perp}} \right)^{1/4}.
 \label{eq:lstar}
 \end{equation}
-For small deflections the actual wavelength $\lambda$ is related to $\lambda^*$ and to the wrinkle amplitude $A$ as $\lambda \sim A^{1/4} \lambda^{*3/4}$.
 
-% Pour Mathieu : ajouter quelque part des applications numériques pour les trois longueurs d'ondes gravité, Darcy, Poiseuille
+%L. Landau and E.M. Lifshitz, Theory of Elasticity (Pergamon, New York, 1986), 3rd ed.
+
+This framework has been successfully applied to situations where $\sigma_{\perp}$ comes from the resistance of an elastic substrate ~\cite{Gough1940, Bijlaard1946}.
+Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
+Futhermore, \ref{eq:lstar} remains valids even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+
+In our situation we now have to determine $\sigma_\perp$.
 We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
-We confirmed that the former plays no significant role by successfully repeating our experiment in a cell held vertically, i.e.
-with the weight acting longitudinally rather than transversely.
-%Since the gel is almost buoyant with the surrounding medium, we expect that the weight plays no significant role, with an elasto-gravitational length at least centimetric.
-We confirmed this insight by successfully repeating our experiment in a cell held vertically, i.e.
+We confirmed that gravity plays no significant role by successfully repeating the wrinkling patern in a cell held vertically, i.e.
 with the weight acting longitudinally rather than transversely.
@@
 
-In a first limit where the film of gel would be sitting without sticking to the bottom wall of the cell ($H_1\rightarrow0$), creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
-Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_\perp/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_\perp}{h} = \frac{dA}{dt}$.
+We consider two limit scenario that leads to wrinkling parterns.
+In a first limit, the film of gel sits without sticking to the bottom wall of the cell ($H_1\rightarrow0$).
+Creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
+Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_{\perp}/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_{\perp}}{h} = \frac{dA}{dt}$.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
@@
 This arrest may be due to bending rigidity~\cite{LeGoff2014}.
-The relevant $\tau$ corresponds to the times it takes for the amplitude to saturate (see Supplementary Method for the precise definition).
+Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate (see Supplementary Method for the precise definition).
 
@@
 There destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
-By symmetry the transverse load across the gel $\sigma_\perp$ is also the pressure difference over the wavelength.
-Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_\perp \frac{H^3}{\eta\lambda^2}$~\cite{Poiseuille1842}.
+By symmetry the transverse load across the gel $\sigma_{\perp}$ is also the pressure difference over the wavelength.
+Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_{\perp} \frac{H^3}{\eta\lambda^2}$~\cite{Poiseuille1842}.
 Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
@@
 
-To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$ the main dissipation mechanism would be the flow through the porous film, whereas for $H\gg H^*$ the flow in the viscous layers would dominate at set the wavelength.
+To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$ the main dissipation mechanism would be the flow through the porous film, whereas for $H\gg H^*$ the flow in the viscous layers would dominate and set the wavelength.
 
@@
 
-\begin{table*}
-\begin{tabular}{cSSSScSrrScSSSScSSSSSS}
-%\toprule
-&\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
-&{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-%ech2
-a & 4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
-& & & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
-b & 4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
-c & 4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
-d & 4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
-& & & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-e & 8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
-& & & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
-f & 8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
-& & & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-g & 4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
-h & 4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
-i & 4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
-\end{tabular}
-\caption{Fully characterised samples at the basis of Fig.~\ref{fig:DarcyPoiseuille}.
-Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.
-`cas', `GDL' and `gly' indicate the weight fraction of sodium caseinate, GDL and glycerol respectively.
-$\xi$ is the pore size at the end of each experiment.}
-\label{tab:data}
-\end{table*}
-
-For a typical gel sheet (Table~\ref{tab:data}b) the buoyancy after syn\ae{}resis is estimated as $\Delta\rho=\SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
-This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$.
+For a typical gel sheet (Supplementary Table~\ref{tab:data}b) the buoyancy after syn\ae{}resis is estimated as $\Delta\rho=\SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ and therefore gravity is dismissed.
 
@@
 Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
-As summarized in Table~\ref{tab:data} we systematically varied (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
+As summarized in Supplementary Table~\ref{tab:data} we systematically varied (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
 Finally, the only parameter that we had no control upon was the initial altitude $H$ of the gel before wrinkling that varied randomly.
@@
 
-\seb{\it Ajouter une conclusion.}
-%optical elements ?
-%wallpaper ?
+\section*{conclusion}
+
+In conclusion we have taken advantage of the properties of the protein casein and carefully ingineered the route to form permeable bio-gels that, in slit geometry,  produce original cascade of wrinkling patterns  when casein adhesion to the slit walls is turned off.
+We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flows of the solvent through the gel layer for short wavelengths.
+Such a model experiment, because it relies on pH induced charge stabilisation/destablisation, should be generalisable to different sort of proteins.
+As such it illustrate daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food sience refXXX or in optical devices such as diffraction gratins or fresnel lenses refXXX, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues refXXX.
+
 %role of gravity
@@
 \section*{Methods summary}
-To prevent adsorption of caseins to surfaces, microscope slides and cover slips were cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush was polymerized from methacrylate groups.
-Sodium caseinate (Firmenich) was labelled with Dylight~550 NHS ESTER (Thermo Scientific).
-Excess dye was removed by centrifugation.
-Unlabelled sodium caseinate powder was dissolved in deionised water (or water--glycerol mixture).
-To induce gelation glucono-$\delta$-lactone (GDL) in powder (Firmenich) was added to the solution.
-\SI{45}{\micro\litre} of this solution was immediately mixed with \SI{5}{\micro\litre} of labelled caseinate solution, stirred and injected in the microscopy cell, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
-3D data were collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
-We used either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) were used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}).
+
+Sodium caseinate (Firmenich)  is disolved in deionised water by stiring for 20 min  at room temperature.
+The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the begining of the experiments.
+We tipically have about 10 min to pipette the dispersion in the slit before gelation starts.
+Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific).
+Excess dye is removed by centrifugation.
+For confocal microscopy measurements, we add 10\% of fluorescently labelled caseines to the caseine/GDL dispersion while it is fluid.
+
+Home made optical cells are composed of a slide (RS France) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm).
+The optical cell is heated  on a hot plate so that the parafilm welds to the  slide and the cover slip.
+Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell between 50 et 150 $\mu$m.
+The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
+To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush is polymerized from methacrylate groups (euromedex).
+
+3D data are collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
+We use either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel.
+Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}).
 Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
-Data shown in Fig.~\ref{fig:acidification} a-b were obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
+Data shown in Fig.~\ref{fig:acidification} a-b are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
 Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}).
+The amount of free casein protein present in the gel during the GDL acidificaition process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in [refXXX]
 
@@
 Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
-The authors thank José Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
+The authors thank Jos\'e Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
 Precious experimental ideas came from Thibaut Divoux, Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
@@
 ML thanks the Region Rhône Alpes and the Programme d'Avenir Lyon - Saint Etienne (PALSE NoGELPo) for postoctoral grant. 
-TG and MN acknowledge funding from ANR XXXX. 
-ML and SM acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No. 258803.
+TG and MN acknowledge funding from the agence national de la recherche fran\c{c}aise (ANR-11-PDOC-027). 
+SM and ML acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No. 258803.
 
@@
 \section*{Supplementary material}
+
+\subsection*{Fraction of free casein proteins as a function of the pH}
+
+the fraction of free casein proteins of a 100mL, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL.
+In both cases, 0.4mL of the 100mL dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for 3 min.
+The supernatent is diluted and analysed with a UV absorption spectrometer(Ocean Optics).
+According to refXXX, we deduced the fraction of free casein proteins from the adsorption spectrum:
+%
+\begin{equation}
+x_{free}=\frac{A(\text{pH})}{A(\text{pH}=7)}\text{, with } A(\text{pH})=A_{280}-A_{320}XXX
+\end{equation}
+%
+where  $A_{280}$ is the absortion peak due to XXX and $A_{320}$ is an estimate of the absorption due to turbidity.
 
@@
 
-$\xi$ was also systematically measured in the same way at the end of each wrinkling experiment and reported in Table~\ref{tab:data} of the main text.
+$\xi$ was also systematically measured in the same way at the end of each wrinkling experiment and reported in Supplementary Table~\ref{tab:data} of the main text.
 %Final fractal dimension was always close to 2 within error bars.
@@
 	\includegraphics{patterns}
-	\caption{Patterns corresponding to the samples of Table~\ref{tab:data} of the main text.
+	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data} of the main text.
 (a-d) increasing cell thickness.
@@
 
+\subsection*{Samples characteristics}
+To test the theoretical predictions we have fully characterized 9 samples, see Supplementary Table \ref{tab:data}.
+
+\begin{table*}
+\begin{tabular}{cSSSScSrrScSSSScSSSSSS}
+%\toprule
+&\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
+\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
+&{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
+%ech2
+a & 4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
+& & & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
+b & 4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
+c & 4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
+d & 4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
+& & & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
+e & 8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
+& & & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
+f & 8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
+& & & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
+\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
+g & 4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
+h & 4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
+i & 4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
+\end{tabular}
+\caption{Fully characterised samples at the basis of Fig.~\ref{fig:DarcyPoiseuille}.
+Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.
+`cas', `GDL' and `gly' indicate the weight fraction of sodium caseinate, GDL and glycerol respectively.
+$\xi$ is the pore size at the end of each experiment.}
+\label{tab:data}
+\end{table*}
+
+\subsection*{Supplementary movies}
+
+\begin{description}
+\item[Movie 1] Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
+Top and bottom slides of the optical cell are coated with acrylamide brushes.
+The duration of the movie is 2~h~30.
+The scale bar is \SI{1}{\milli\metre}.
+
+\item[Movie 2] 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion (Supplementary Table~\ref{tab:data}c).
+Top and bottom slides of the optical cell are coated with acrylamide brushes.
+The duration of the movie is2~h~30.
+The scale bar is \SI{1}{\milli\metre}.
+
+\item[Movie 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the movie is \SI{13}{\minute}. The scale bar is XXX $\mu$m
+
+\item[Movie 4] Titration of a dispersion with 4\%w casein by 1 molar HCl. The duration of the movie is \SI{14}{\minute}. The scale bar is XXX $\mu$m
+
+\item[Movie 5] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
+1/10th of the proteins are fluorescently labelled.
+The duration of the movie is \SI{132}{\minute}.
+The scale bar is \SI{10}{\micro\metre}.
+
+\item[Movie 6] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
+1/10th of the proteins are fluorescently labelled.
+The duration of the movie is \SI{147}{\minute}.
+The scale bar is \SI{10}{\micro\metre}.
+
+\item[Movie 7] Confocal microscopy $(x,z)$ cut of the evolution of thegel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a).
+The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
+Here the gel starts close to the bottom wall, $H<H^*$.
+Darcy mechanism dominates the formation of the primary pattern.
+Total duration of the video is 145 min.
+
+\item[Movie 8] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}c).
+The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
+Here the gel starts far from either top or bottom wall, $H>H^*$.
+Poiseuille mechanism dominates the formation of the primary pattern.
+Total duration of the video is 181 min.
+
+\end{description}
+
 \end{document}

### 2015-02-12 fig_plis/article_plis.tex
Fautes d'orthographe les plus visibles Références manquantes de l'intro casein (plural or not)

@@
 \begin{abstract}
-The slow acidification of milk proteins such as casein is well known to result in the formation of gels – this is indeed one of the key steps to make cheese or yogurt.
-Here, using the casein proteins as  building blocks and the acidifier  Glucono delta-lactone, we design model thin bio-gel layers that wrinkles in cascade.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gels shrink and swell creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel.
-We have develop a hydrodynamic model that predict correctly the wrinkling wavelength.
-Our results should be universal for proteins as it is based on pH induced charge stabilasion/destablisation and control of boundary conditons.
-As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical puposes.
+The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels -- this is indeed one of the key steps to make yoghurt.
+Here, using the casein proteins as  building blocks and the acidulent Glucono-$\delta$-lactone, we design model thin biogel layers that wrinkles in cascade.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gel shrinks and swells creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel.
+We have develop an elasto-hydrodynamic model that predicts correctly the wrinkling wavelength.
+Our results should be universal for acid-induced protein gels as it is based on pH induced charge stabilisation/destabilisation and on control of boundary conditions.
+As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
 \end{abstract}
@@
 An elastic film buckles due to excess area with respect to its boundaries and wrinkles when buckling is hindered by a substrate.
-The selection of the preferred wrinkling wavelength is peculiar to the situation and is related to the mismatch of elastic properties between the surface and the substrate~\cite{Gough1940, Bijlaard1946} XXX.
+The selection of the preferred wrinkling wavelength is peculiar to the situation and is related to the mismatch of elastic properties between the surface and the substrate~\cite{Gough1940, Bijlaard1946, Biot1957, Bowden1998, Cerda2003, Genzer2006}.
 On the biology side wrinkling-controlled morphogenesis is ubiquitous.
@@
 On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear behaviour~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
-Such patterns can be triggered by temperature dilation refXXX, swelling refXXX or the removal of pre-strain refXXX.
+Such patterns can be triggered by temperature dilation~\cite{Bowden1998}, swelling~\cite{Hu1998} or the removal of pre-strain~\cite{Genzer2006}.
 
@@
 Here, we introduce model permeable biogels that produce hierarchical wrinkling as a result of their interplay between their gelation dynamics and the confinement conditions.
-In slit geometry, we slowly acidify a dispersion of casein, one of the milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Bremer1989} --this is a well-known effect used in cheese or yogurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
+In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
+As a result, not only does the casein form a gel as expected~\cite{Bremer1989} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
 Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
@@
 	\includegraphics{acidification}%
-	\caption{Acid-induced protein gel properties behave nonmonotonically with pH.
+	\caption{Acid-induced protein gel properties behave non monotonically with pH.
 (a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL.
-The horizontal line indicates isoelectric pH of casein.
+The horizontal line indicates isoelectric pH of caseins.
 (b) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator.
@@
 
-To form wrinkled patterns such as the one oberved in figure 1XXX, ones needs to take advantage of the pH sensitivity of casein to control the route to gelation, to fine tune the gel properties and the boundary conditions.
+To form wrinkled patterns such as the one oberved in figure 1XXX, ones needs to take advantage of the pH sensitivity of caseins to control the route to gelation, to fine tune the gel properties and the boundary conditions.
 
-the stability of casein dispersion strongly depend on the pH of the solution.
-Supplementary movie 3XXX shows the titration of a casein dispertion at 1\%w by 1 molar HCl.
-At pH=7, the casein is well dispersed then it floculate between pH= XXX to XXX and finaly at pH=XXX the solution become clear again.
-However as one increases the casein concentration up to 4\%w, at low pH the casein redispersion is not total, Supplementary movie 4XXX.
-The state diagram of casein as a function of the pH  is related to charge stabilization.
-At neutral pH casein are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseine (pH=4.5), caseine molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability].
-Beyond the isoelectric point casein acquire a positive net charge and therefore partialy re-disolve.
-In figure 2cXXX, we quantify this effect and measure with UV absorption spectroscopie the amount of free casein within the 4\%w dispersion as a function of pH (see Supplementary Methods).
+The stability of caseins dispersion strongly depend on the pH of the solution.
+Supplementary movie 3XXX shows the titration of a caseins dispersion at 1\%w by 1 molar HCl.
+At pH=7, the caseins are well dispersed then then flocculate between pH= XXX to XXX and finaly at pH=XXX the solution become clear again.
+However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary movie 4XXX.
+The state diagram of caseins as a function of the pH  is related to charge stabilization.
+At neutral pH caseins are charged negatively and repel each other.
+As the pH decreases to the isoelectric point of the caseins (pH=4.5), caseins molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability].
+Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve.
+In figure 2cXXX, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH (see Supplementary Methods).
 Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction is 1.
 Upon decreasing the pH of the solution to the isolectric point the free casein fraction drop to almost 0, almost all the casein molecule are aggregated.
-Further decrease of the pH of the solution lead to a partial redisolution of the casein agregate.
+Further decrease of the pH of the solution lead to a partial redissolution of the casein aggregates.
 At pH=2 the fraction of free casein is XXX.
@@
 
-To form a space spaning gel, floculation need to be prevented.
-We use of Glucono delta-lactone (GDL) a molecule  that slowly hydrlolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, fig.2aXXX.
-This GDL controled acidification permits the formation of a bulk gel phase.
+To form a space spanning gel, flocculation need to be prevented.
+We use of Glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, fig.2aXXX.
+This GDL controlled acidification permits the formation of a bulk gel phase.
 We monitor the GDL induced gelation with a rheometer in cone plate geometry.
-As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elatic and viscous modulus of the casein dispersion as a function of time.
+As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elastic and viscous modulus of the casein dispersion as a function of time.
 the mechanical properties of the gel are closely related to the fraction of free casein.
-First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finaly the elastic modulus decreases down to 100 Pa.
+First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finally the elastic modulus decreases down to 100 Pa.
 
@@
 As shown in supplementary fig.
-1XXX the gel consist in a space spaning fractal network with a curtoff length scale xi.
+1XXX the gel consist in a space spanning fractal network with a cut off length scale xi.
 xi corresponds to the maximum pore size of the network.
-This model fits the structure at all times but its parameters vary non monotoneously during GDL acidification.
+This model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
 As shown in Fig.
@@
 
-As seen on the confocal measurements, the gel is porous and is permeable to the solvant.
-This protertie is determinant as it allow the gel to absord and expel the solvant and therefore change its volume when sticky boundary are turned off.
-To quantify this effect, we have embeded the casein gel inside a capillary and deeped verticaly the capilary in an aqueous buffer [refXXX].
+As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
+This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off.
+To quantify this effect, we have embedded the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer [refXXX].
 In the supplementary figure XXX, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces.
@@
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-The evolution of xi when sticky boundaries are on reflects the inner constrains of the gel:  xi increases, the stand of the nework become thinner and the gel wants to retract; then xi decreases, the stands of network swell and the gel wants to expend.
+The evolution of xi when sticky boundaries are on reflects the inner constrains of the gel:  xi increases, the stand of the network become thinner and the gel wants to retract; then xi decreases, the stands of network swell and the gel wants to expend.
 When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detach from the top wall.
@@
 XXX.
-xi remains mostly constant and the major effect is repported on the volume occupied by the gel as will be repported below.
+xi remains mostly constant and the major effect is reported on the volume occupied by the gel as will be reported below.
 
@@
 
-Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favorable to wrinkling.
+Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favourable to wrinkling.
 Up to 10 min the dispersion is homogeneous and fills the entire slit.
 The black margins correspond to the glass walls.
-Around 23 min, the gel forms and simultaneously becomes thiner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the situation at 10min.
-This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expeled on both the upper and lower sides.
-We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thicknesses $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
+Around 23 min, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the situation at 10min.
+This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
+We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
 Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods).
-As expected from the system design, the evolution of the volume is non-monotoneous.
+As expected from the system design, the evolution of the volume is non-monotonous.
 After reaching a minimum around 35\% of its initial value at 23 min, the volume increases.
@@
 	\includegraphics{sideview}
-	\caption{Measurements and modeling.
+	\caption{Measurements and modelling.
 (a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
@@
 Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations  refLaudauXXX,  essentially impossible to solve in analytical form.
-However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinking patern and isolate the physically meaningful ingredients.
+However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients.
 An elastic film buckles due to excess area with respect to its boundaries.
@@
 We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
-We confirmed that gravity plays no significant role by successfully repeating the wrinkling patern in a cell held vertically, i.e.
+We confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
 with the weight acting longitudinally rather than transversely.
@@
 
-We consider two limit scenario that leads to wrinkling parterns.
+We consider two limit scenario that leads to wrinkling patterns.
 In a first limit, the film of gel sits without sticking to the bottom wall of the cell ($H_1\rightarrow0$).
@@
 We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flows of the solvent through the gel layer for short wavelengths.
-Such a model experiment, because it relies on pH induced charge stabilisation/destablisation, should be generalisable to different sort of proteins.
-As such it illustrate daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food sience refXXX or in optical devices such as diffraction gratins or fresnel lenses refXXX, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues refXXX.
+Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
+As such it illustrate daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses refXXX, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues refXXX.
 
@@
 
-Sodium caseinate (Firmenich)  is disolved in deionised water by stiring for 20 min  at room temperature.
-The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the begining of the experiments.
-We tipically have about 10 min to pipette the dispersion in the slit before gelation starts.
+Sodium caseinate (Firmenich)  is dissolved in deionised water by stirring for 20 min  at room temperature.
+The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the beginning of the experiments.
+We typically have about 10 min to pipette the dispersion in the slit before gelation starts.
 Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific).
 Excess dye is removed by centrifugation.
-For confocal microscopy measurements, we add 10\% of fluorescently labelled caseines to the caseine/GDL dispersion while it is fluid.
+For confocal microscopy measurements, we add 10\% of fluorescently labelled caseins to the casein/GDL dispersion while it is fluid.
 
@@
 The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
-To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush is polymerized from methacrylate groups (euromedex).
+To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush is polymerized from methacrylate groups (Euromedex).
 
@@
 Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}).
-The amount of free casein protein present in the gel during the GDL acidificaition process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in [refXXX]
+The amount of free casein protein present in the gel during the GDL acidification process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in [refXXX]
 
@@
 In both cases, 0.4mL of the 100mL dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for 3 min.
-The supernatent is diluted and analysed with a UV absorption spectrometer(Ocean Optics).
+The supernatant is diluted and analysed with a UV absorption spectrometer(Ocean Optics).
 According to refXXX, we deduced the fraction of free casein proteins from the adsorption spectrum:
@@
 
-We first denoise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
+We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
 Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
@@
 
-From these data, we obtain for each stack the volume and thickness of the gel phase, area of the midsurface and peak-to-peak amplitude.
+From these data, we obtain for each stack the volume and thickness of the gel phase, area of the mid surface and peak-to-peak amplitude.
 Velocity is obtained by time differentiation.
@@
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The duration of the movie is2~h~30.
+The duration of the movie is 2~h~30.
 The scale bar is \SI{1}{\milli\metre}.
@@
 
-\item[Movie 7] Confocal microscopy $(x,z)$ cut of the evolution of thegel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a).
+\item[Movie 7] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a).
 The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).

### 2015-02-12 fig_plis/article_plis.tex
Movie -> Video xi -> $\xi$ Fix references in System design Describe fig:acidification b and inset

@@
 We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
-As shown in Fig.~\ref{fig:dynamics} and in Supplementary Movies 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
+As shown in Fig.~\ref{fig:dynamics} and in Supplementary Videos 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
 Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates.
@@
 
-To form wrinkled patterns such as the one oberved in figure 1XXX, ones needs to take advantage of the pH sensitivity of caseins to control the route to gelation, to fine tune the gel properties and the boundary conditions.
+To form wrinkled patterns such as the one observed in figure~\ref{fig:dynamics}, ones needs to take advantage of the pH sensitivity of caseins to control the route to gelation, to fine tune the gel properties and the boundary conditions.
 
 The stability of caseins dispersion strongly depend on the pH of the solution.
-Supplementary movie 3XXX shows the titration of a caseins dispersion at 1\%w by 1 molar HCl.
-At pH=7, the caseins are well dispersed then then flocculate between pH= XXX to XXX and finaly at pH=XXX the solution become clear again.
-However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary movie 4XXX.
+Supplementary Video 3 shows the titration of a caseins dispersion at 1\%w by 1 molar HCl.
+At pH=7, the caseins are well dispersed then then flocculate between pH=5.5 to 4 and finaly at pH=3 the solution become clear again.
+However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary Video 4.
 The state diagram of caseins as a function of the pH is related to charge stabilization.
 At neutral pH caseins are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseins (pH=4.5), caseins molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability].
+As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.5$), caseins molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability].
 Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve.
-In figure 2cXXX, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH (see Supplementary Methods).
-Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction is 1.
+In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH (see Supplementary Methods).
+Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction $x_\text{free}$ is 1.
 Upon decreasing the pH of the solution to the isolectric point the free casein fraction drop to almost 0, almost all the casein molecule are aggregated.
 Further decrease of the pH of the solution lead to a partial redissolution of the casein aggregates.
-At pH=2 the fraction of free casein is XXX.
+At pH=2 the fraction of free casein is 0.35.
 
@@
 To form a space spanning gel, flocculation need to be prevented.
-We use of Glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, fig.2aXXX.
-This GDL controlled acidification permits the formation of a bulk gel phase.
+We use of Glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Fig.~\ref{fig:acidification}a.
+This GDL controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
+As shown on Figure~\ref{fig:acidification}a and b, when adding enough GDL the pH lowers beyond the isoeletric point and $x_\text{free}$ rises again.
 We monitor the GDL induced gelation with a rheometer in cone plate geometry.
-As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elastic and viscous modulus of the casein dispersion as a function of time.
-the mechanical properties of the gel are closely related to the fraction of free casein.
+As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elastic and viscous modulus of the casein dispersion as a function of time, Figure~\ref{fig:acidification}c.
+The mechanical properties of the gel are closely related to the fraction of free casein.
 First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finally the elastic modulus decreases down to 100 Pa.
+This finite value indicates that the sample is still a gel.
 
 To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy.
-As shown in supplementary fig.
-1XXX the gel consist in a space spanning fractal network with a cut off length scale xi.
-xi corresponds to the maximum pore size of the network.
+As shown in Supplementary Figure~\ref{fig:nonmonotonic} the gel consist in a space spanning fractal network with a cut off length scale $\xi$ that corresponds to the maximum pore size of the network.
 This model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
-As shown in Fig.
-2cXXX, up to the isolectric point xi increases up to a maximum length of 7 $\mu$m due to the solvent expulsion out of the casein network [refXXX].
-This stiffen the arm of the network and is responsible for the increase of elasticity.
-Beyond the isoelectric point xi decreases down to 4$\mu$m.
-Casein regain solubility [refXXX], the strand network swell, the pore shrinks and the gel weakens.
+As shown in Figure~\ref{fig:acidification}d, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network [refXXX].
+This stiffens the arms of the network and is responsible for the increase of elasticity.
+Beyond the isoelectric point $\xi$ decreases down to \SI{4}{\micro\metre}.
+As casein regain solubility, the strands of the network swell, the pores shrink and the gel weakens.
 
-As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
+XXXCe paragraphe fait-il sens ici ?XXX As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
 This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off.
 To quantify this effect, we have embedded the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer [refXXX].
-In the supplementary figure XXX, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces.
-Exploitation of this experiment give the permeability of the gel: 67000 nm$^2$.
+In the Supplementary Figure~\ref{fig:permeability}, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces.
+Exploitation of this experiment give the permeability of the gel: \SI{67000}{\square\nano\metre}.
 
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-The evolution of xi when sticky boundaries are on reflects the inner constrains of the gel:  xi increases, the stand of the network become thinner and the gel wants to retract; then xi decreases, the stands of network swell and the gel wants to expend.
+The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel:  $\xi$ increases, the stand of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of network swell and the gel wants to expend.
 When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detach from the top wall.
-In this situation we can monitor xi, fig.
-XXX.
-xi remains mostly constant and the major effect is reported on the volume occupied by the gel as will be reported below.
+In this situation we can monitor $\xi$, Fig.~\ref{fig:acidification}d.
+$\xi$ remains mostly constant and the major effect is reported on the volume occupied by the gel as will be reported below.
 
@@
 In fig.2c, when the top slide of the optical cell is coated with acrylamide brushes, we then observe that the casein network detaches from the top slide.
-xi does not peak at the isoelectric point but remains constant
+$\xi$ does not peak at the isoelectric point but remains constant
 
@@
 
-%The difference in value of xi between the fully un-coated  and the half-coated optical cell demonstrate that the gel is subject to constrains as the pH decreases.
+%The difference in value of $\xi$ between the fully un-coated  and the half-coated optical cell demonstrate that the gel is subject to constrains as the pH decreases.
 The relaxation of thoses constrain is key to the wrinkling formation when both the top and bottom slide are coated with acrylamide brushes.
@@
 
-\subsection*{Supplementary movies}
+\subsection*{Supplementary Videos}
 
 \begin{description}
-\item[Movie 1] Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
+\item[Video 1] Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The duration of the movie is 2~h~30.
+The duration of the Video is 2~h~30.
 The scale bar is \SI{1}{\milli\metre}.
 
-\item[Movie 2] 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion (Supplementary Table~\ref{tab:data}c).
+\item[Video 2] 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion (Supplementary Table~\ref{tab:data}c).
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The duration of the movie is 2~h~30.
+The duration of the Video is 2~h~30.
 The scale bar is \SI{1}{\milli\metre}.
 
-\item[Movie 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the movie is \SI{13}{\minute}. The scale bar is XXX $\mu$m
+\item[Video 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the Video is \SI{13}{\minute}. The scale bar is XXX $\mu$m
 
-\item[Movie 4] Titration of a dispersion with 4\%w casein by 1 molar HCl. The duration of the movie is \SI{14}{\minute}. The scale bar is XXX $\mu$m
+\item[Video 4] Titration of a dispersion with 4\%w casein by 1 molar HCl. The duration of the Video is \SI{14}{\minute}. The scale bar is XXX $\mu$m
 
-\item[Movie 5] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
+\item[Video 5] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
 1/10th of the proteins are fluorescently labelled.
-The duration of the movie is \SI{132}{\minute}.
+The duration of the Video is \SI{132}{\minute}.
 The scale bar is \SI{10}{\micro\metre}.
 
-\item[Movie 6] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
+\item[Video 6] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
 1/10th of the proteins are fluorescently labelled.
-The duration of the movie is \SI{147}{\minute}.
+The duration of the Video is \SI{147}{\minute}.
 The scale bar is \SI{10}{\micro\metre}.
 
-\item[Movie 7] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a).
+\item[Video 7] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a).
 The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
@@
 
-\item[Movie 8] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}c).
+\item[Video 8] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}c).
 The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).

### 2015-02-12 fig_plis/article_plis.tex
SI dans Dynamics of the wrinkling process

@@
 Beyond the isoelectric point $\xi$ decreases down to \SI{4}{\micro\metre}.
-As casein regain solubility, the strands of the network swell, the pores shrink and the gel weakens.
+As caseins regain solubility, the strands of the network swell, the pores shrink and the gel weakens.
 
@@
 
-Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favourable to wrinkling.
-Up to 10 min the dispersion is homogeneous and fills the entire slit.
+Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favourable to wrinkling (see also Supplementary Videos 7 and 8).
+Initially the dispersion is homogeneous and fills the entire slit.
 The black margins correspond to the glass walls.
-Around 23 min, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the situation at 10min.
+Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
 This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
@@
 As expected from the system design, the evolution of the volume is non-monotonous.
-After reaching a minimum around 35\% of its initial value at 23 min, the volume increases.
-Shortly after,  at 36 min, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area.
+After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
+Shortly after,  at \SI{36}{\minute}, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area.
 This confirms that the instability is due to swelling.
-Around 38 min, the amplitude of the wrinkles reach the thickness of the slit.
-The size of the primary pattern is now settled Supplementary Fig.
-\ref{fig:patterns, fig:growth}.
-However swelling continues, the gel flattens on both top and bottom walls (44 min) and finally buckles back (53 min) to form the secondary pattern (1h21).
+Around \SI{38}{\minute}, the amplitude of the wrinkles reach the thickness of the slit.
+The size of the primary pattern is now settled, Supplementary Figures~\ref{fig:patterns} and \ref{fig:growth}.
+However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (1h21).
 When the excess volume saturates, the dynamics stops.

### 2015-02-13 fig_plis/article_plis.tex
Description de fig:acidification b Corresponding Supplementary methods

@@
 Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
-On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear behaviour~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
+On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
 Such patterns can be triggered by temperature dilation~\cite{Bowden1998}, swelling~\cite{Hu1998} or the removal of pre-strain~\cite{Genzer2006}.
@@
 In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Bremer1989} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
+As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
 Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
@@
 The horizontal line indicates isoelectric pH of caseins.
-(b) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator.
-(c) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
+(b) Corresponding evolution of the fraction of free caseins.
+Symbols are direct measures upon addition of GDL.
+Continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert).
+(c) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator.
+(d) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
 allowed syn\ae{}resis and swelling (orange dashed line).}%
-	\label{fig:acidification}%
+	\label{fig:acidification}
 \end{figure}
@@
 At neutral pH caseins are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.5$), caseins molecules loose their negative net charge and tend to stick to each other [ref XXX casein stability].
+As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), caseins molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989}.
 Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve.
@@
 Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}).
-The amount of free casein protein present in the gel during the GDL acidification process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in [refXXX]
+The amount of free casein protein present in the gel during the GDL acidification process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
 
@@
 
-the fraction of free casein proteins of a 100mL, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL.
+The fraction of free casein proteins of a 100mL, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL.
 In both cases, 0.4mL of the 100mL dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for 3 min.
 The supernatant is diluted and analysed with a UV absorption spectrometer(Ocean Optics).
-According to refXXX, we deduced the fraction of free casein proteins from the adsorption spectrum:
+According to~\cite{Roefs1986}, we deduced the fraction of free casein proteins from the adsorption spectrum:
 %
-\begin{equation}
-x_{free}=\frac{A(\text{pH})}{A(\text{pH}=7)}\text{, with } A(\text{pH})=A_{280}-A_{320}XXX
-\end{equation}
+\begin{align}
+x_{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
+\text{with }A(\text{pH}) &= A_{280}-1.706 A_{320},
+\end{align}
 %
-where  $A_{280}$ is the absortion peak due to XXX and $A_{320}$ is an estimate of the absorption due to turbidity.
+where $A_{280}$ is the absorption peak due to aromatic amino acids in the casein molecules plus a turbidity background and $A_{320}$ is only due to turbidity.
+
+Above the isoelectric point, the homogeneous acidification by GDL continuously creates very small protein aggregates that are difficult to centrifuge.
+Such suspensions were too turbid to reliably measure $x_{free}$.
 

### 2015-02-13 fig_plis/article_plis.tex
references in conclusion and various typos

@@
 The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels -- this is indeed one of the key steps to make yoghurt.
-Here, using the casein proteins as  building blocks and the acidulent Glucono-$\delta$-lactone, we design model thin biogel layers that wrinkles in cascade.
+Here, using the casein proteins as  building blocks and the acidulent glucono-$\delta$-lactone, we design model thin biogel layers that wrinkles in cascade.
 Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gel shrinks and swells creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel.
@@
 
-To form a space spanning gel, flocculation need to be prevented.
-We use of Glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Fig.~\ref{fig:acidification}a.
+To form a space spanning gel, flocculation needs to be prevented.
+We use of glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Fig.~\ref{fig:acidification}a.
 This GDL controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
@@
 This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off.
-To quantify this effect, we have embedded the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer [refXXX].
+To quantify this effect, we have embedded the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}.
 In the Supplementary Figure~\ref{fig:permeability}, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces.
@@
 (b) Schematic side view of the cell.
-The constant thickness gel film (orange) is surrounded by water.
+The constant thickness gel film (orange) is surrounded by solvent.
 Yellow lines show the path of transmitted light through the gel phase.
@@
 The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem.
-Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations  refLaudauXXX,  essentially impossible to solve in analytical form.
-However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients.
+Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations ~\cite{landau1986},  essentially impossible to solve in analytical form.
+However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}.
 An elastic film buckles due to excess area with respect to its boundaries.
@@
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Futhermore, \ref{eq:lstar} remains valids even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+Futhermore, (Eq.~\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
 
@@
 
-We consider two limit scenario that leads to wrinkling patterns.
+We consider two limit scenarios that leads to wrinkling patterns.
 In a first limit, the film of gel sits without sticking to the bottom wall of the cell ($H_1\rightarrow0$).
@@
 
-In conclusion we have taken advantage of the properties of the protein casein and carefully ingineered the route to form permeable bio-gels that, in slit geometry,  produce original cascade of wrinkling patterns  when casein adhesion to the slit walls is turned off.
-We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flows of the solvent through the gel layer for short wavelengths.
+In conclusion we have taken advantage of the properties of the protein casein and carefully engineered the route to form permeable biogels that, in slit geometry, produce original cascade of wrinkling patterns when casein adhesion to the slit walls is turned off.
+We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flow of the solvent through the gel layer for short wavelengths.
 Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
-As such it illustrate daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses refXXX, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues refXXX.
+As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Chan2006a,Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013}.
 

### 2015-02-13 fig_plis/article_plis.tex
Ref yaourt : ajoute Lucey1998 et corrige Roefs1986

@@
 In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
+As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
 Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
@@
 At neutral pH caseins are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), caseins molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989}.
+As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), caseins molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998}.
 Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve.
@@
 This model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
-As shown in Figure~\ref{fig:acidification}d, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network [refXXX].
+As shown in Figure~\ref{fig:acidification}d, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998}.
 This stiffens the arms of the network and is responsible for the increase of elasticity.
@@
 Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
-As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Chan2006a,Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013}.
+As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013}.
 

### 2015-02-13 fig_plis/article_plis.tex
chasse au passé Moins de détail sur la perméabilité dans le texte (c'est expliqué en méthode et supplémentary) Explique qu'on est capable d'empècher l'adsorption grâce à l'acrylamide, mais ensuite uniquement "turn off" "turn on" Cite les Video 7 et 8 dans le texte En remet une couche pour l'interprétation de H<>H* Complète le caption de fig:nonmonotonic Protein labelling in Video description

@@
 \usepackage{hyperref}
-\usepackage{siunitx}
+\usepackage[range-units = single]{siunitx}
 %\usepackage{multirow}
 \usepackage{bm}
-\usepackage{dcolumn,ulem}
+%\usepackage{dcolumn,ulem}
 \usepackage{amsfonts}
@@
 Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gel shrinks and swells creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel.
-We have develop an elasto-hydrodynamic model that predicts correctly the wrinkling wavelength.
+We develop an elasto-hydrodynamic model that predicts correctly the wrinkling wavelength.
 Our results should be universal for acid-induced protein gels as it is based on pH induced charge stabilisation/destabilisation and on control of boundary conditions.
@@
 On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
-Such patterns can be triggered by temperature dilation~\cite{Bowden1998}, swelling~\cite{Hu1998} or the removal of pre-strain~\cite{Genzer2006}.
+Such patterns can be triggered by temperature dilation~\cite{Bowden1998}, swelling~\cite{Hu1998,Kim2010} or the removal of pre-strain~\cite{Genzer2006}.
 
@@
 Supplementary Video 3 shows the titration of a caseins dispersion at 1\%w by 1 molar HCl.
-At pH=7, the caseins are well dispersed then then flocculate between pH=5.5 to 4 and finaly at pH=3 the solution become clear again.
+At pH=7, the caseins are well dispersed then then flocculate between pH=5.5 to 4 and finally at pH=3 the solution become clear again.
 However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary Video 4.
@@
 To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy.
-As shown in Supplementary Figure~\ref{fig:nonmonotonic} the gel consist in a space spanning fractal network with a cut off length scale $\xi$ that corresponds to the maximum pore size of the network.
-This model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
-As shown in Figure~\ref{fig:acidification}d, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998}.
+As shown in Supplementary Figure~\ref{fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
+Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~\ref{fig:permeability}.
+The fractal model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
+As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998}.
 This stiffens the arms of the network and is responsible for the increase of elasticity.
@@
 
-XXXCe paragraphe fait-il sens ici ?XXX As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
+%XXXCe paragraphe fait-il sens ici ?XXX As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
 This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off.
-To quantify this effect, we have embedded the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}.
-In the Supplementary Figure~\ref{fig:permeability}, we observe that the aqueous buffer goes through the gel and rise in the tube due to capillary forces.
+To quantify this effect, we make the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}.
+In the Supplementary Figure~\ref{fig:permeability}, we observe that the aqueous buffer goes through the gel and rises in the tube due to capillary forces.
 Exploitation of this experiment give the permeability of the gel: \SI{67000}{\square\nano\metre}.
@@
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel:  $\xi$ increases, the stand of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of network swell and the gel wants to expend.
-When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detach from the top wall.
+To turn off the adsorption of the caseins on a surface, we coat it with polymer brushes (see Methods), effectively creating non adhesive boundary conditions for the gel.
+The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel: $\xi$ increases, the stands of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of the network swell and the gel wants to expend.
+When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall.
 In this situation we can monitor $\xi$, Fig.~\ref{fig:acidification}d.
-$\xi$ remains mostly constant and the major effect is reported on the volume occupied by the gel as will be reported below.
+$\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.
 
@@
 
-Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls have been coated, a configuration favourable to wrinkling (see also Supplementary Videos 7 and 8).
+Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration favourable to wrinkling (see also Supplementary Videos 7 and 8).
 Initially the dispersion is homogeneous and fills the entire slit.
@@
 This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
-We describe sample as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
+We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
 Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods).
@@
 The size of the primary pattern is now settled, Supplementary Figures~\ref{fig:patterns} and \ref{fig:growth}.
-However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (1h21).
+However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
 When the excess volume saturates, the dynamics stops.
-The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we have observe the formation of higher order patterns.
+The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
 
-Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allow subsequent planar stresses to develop while swelling: the only situation were wrinkling was observed is when both top and bottom walls are coated with acrylamide brushes.
+Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off.
 We also confirmed that at any given point in time and space the thickness $h$ of the film of gel is homogeneous.
@@
 The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem.
-Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations ~\cite{landau1986},  essentially impossible to solve in analytical form.
+Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations~\cite{landau1986}, essentially impossible to solve in analytical form.
 However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}.
@@
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Futhermore, (Eq.~\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+Furthermore, (Eq.~\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
 
@@
 We consider two limit scenarios that leads to wrinkling patterns.
-In a first limit, the film of gel sits without sticking to the bottom wall of the cell ($H_1\rightarrow0$).
+In a first limit, the film of gel sits without sticking to the bottom wall of the cell, i.e.
+$H_1\rightarrow0$, see Supplementary Video 7.
 Creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
@@
 
-The opposite limiting case is when the gel film lies in the middle of the cell ($H_1=H_2$) and impermeable ($\alpha\rightarrow 0$).
-There destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
+The opposite limiting case is when the gel film lies in the middle of the cell ($H_1=H_2$, see Supplementary Video 8) and impermeable ($\alpha\rightarrow 0$).
+Here destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
 By symmetry the transverse load across the gel $\sigma_{\perp}$ is also the pressure difference over the wavelength.
@@
 
-To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$ the main dissipation mechanism would be the flow through the porous film, whereas for $H\gg H^*$ the flow in the viscous layers would dominate and set the wavelength.
+To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$, i.e.
+slow destabilisation of a highly permeable film close to a wall, the main dissipation mechanism would be the flow through the porous film; whereas for $H\gg H^*$, i.e.
+fast wrinkling of a poorly permeable film far from any wall, the flow in the viscous layers would dominate and set the wavelength.
 
@@
 Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
-As summarized in Supplementary Table~\ref{tab:data} we systematically varied (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
-Finally, the only parameter that we had no control upon was the initial altitude $H$ of the gel before wrinkling that varied randomly.
-However we were able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
+As summarized in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
+Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly.
+However we are able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
@@
 
-Sodium caseinate (Firmenich)  is dissolved in deionised water by stirring for 20 min  at room temperature.
+Sodium caseinate (Firmenich) is dissolved in deionised water by stirring for \SI{20}{\minute} at room temperature.
 The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the beginning of the experiments.
-We typically have about 10 min to pipette the dispersion in the slit before gelation starts.
+We typically have about \SI{10}{\minute} to pipette the dispersion in the slit before gelation starts.
 Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific).
@@
 Home made optical cells are composed of a slide (RS France) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm).
-The optical cell is heated  on a hot plate so that the parafilm welds to the  slide and the cover slip.
-Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell between 50 et 150 $\mu$m.
+The optical cell is briefly heated on a hot plate so that the parafilm welds to the slide and the cover slip.
+Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell from \SIrange{50}{150}{\micro\metre}.
 The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
-To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) to which a polyacrylamide brush is polymerized from methacrylate groups (Euromedex).
+To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which a polyacrylamide brush is polymerized from methacrylate groups (Euromedex).
 
@@
 Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
-Data shown in Fig.~\ref{fig:acidification} a-b are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
+Data shown in Fig.~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
 Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}).
-The amount of free casein protein present in the gel during the GDL acidification process was measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
+The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
 
@@
 
-\subsection*{Fraction of free casein proteins as a function of the pH}
+\subsection*{Fraction of free casein as a function of the pH}
 
-The fraction of free casein proteins of a 100mL, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL.
-In both cases, 0.4mL of the 100mL dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for 3 min.
-The supernatant is diluted and analysed with a UV absorption spectrometer(Ocean Optics).
+The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL.
+In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for \SI{3}{\minute}.
+The supernatant is diluted and analysed with a UV absorption spectrometer (Ocean Optics).
 According to~\cite{Roefs1986}, we deduced the fraction of free casein proteins from the adsorption spectrum:
@@
 Above the isoelectric point, the homogeneous acidification by GDL continuously creates very small protein aggregates that are difficult to centrifuge.
-Such suspensions were too turbid to reliably measure $x_{free}$.
+Such suspensions are too turbid and too rapidly evolving to reliably measure $x_{free}$.
 
@@
 Arrows indicates the pore size.
-(b) Corresponding Fourier spectra when no adhesion on top and same times.
+(b) Corresponding Fourier spectra when no adhesion on top, same times as (a).
 Lines are fits by Eq.~\ref{eq:fractalS}.
-(c-d) Evolution of susceptibility and fractal dimension in full adhesive (black line) and no adhesion on top (orange dashed line) situations.
+The inset shows a schematic association between real space structures and wave numbers.
+(c-e) Evolution of cut-off length, susceptibility and fractal dimension in full adhesive (black line) and no adhesion on top (orange dashed line) situations.
 Symbols indicates the times of the images in (a).}%
@@
 
-To follow the evolution of the microstructure of the gel, we acquired slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:nonmonotonic}a).
-After Hann windowing we performed Fourier transform and radially averaged the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:nonmonotonic}b) that was fitted by the following fractal form:
+To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:nonmonotonic}a).
+After Hann windowing we performe Fourier transform and radially average the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:nonmonotonic}b) that is fitted by the following fractal form:
 \begin{equation}
@@
 Figure~\ref{fig:acidification}c of the main text and Supplementary Fig.~\ref{fig:nonmonotonic}c and d show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
-The later situation is a proxy to the situation where we observe wrinkling, since it was necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
-To do so, we coated only the microscope slide with acrylamide brush.
-The cover slip was cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
-Neither buckling nor wrinkling was detected in such configuration.
+The later situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
+To do so, we coat only the microscope slide with acrylamide brush.
+The cover slip is cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
+Neither buckling nor wrinkling is detected in such configuration.
 
@@
 
-$\xi$ was also systematically measured in the same way at the end of each wrinkling experiment and reported in Supplementary Table~\ref{tab:data} of the main text.
+$\xi$ is also systematically measured in the same way at the end of each wrinkling experiment and reported in Supplementary Table~\ref{tab:data}.
 %Final fractal dimension was always close to 2 within error bars.
@@
 
-At 10x magnification and depending on cell thickness, we acquired stacks 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
-Each confocal slice was $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
+At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
+Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
 At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.
@@
 
-In this framework, we recall here the well known case of an elastic film lying over an infinite elastic medium.
+In this framework, we recall here the case of an elastic film lying over an infinite elastic medium~\cite{Gough1940,Bijlaard1946,Biot1957,Cerda2003}.
 The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -q B_s w$, and thus the compression load writes
@@
 
-Minimizing $\sigma_\parallel$ as a function of $q$ yields the well-known result $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
+Minimizing $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
 \begin{equation}
@@
 
-To measure wrinkle wavelength, we used large field pictures of the final state.
-These picture were obtained either by stitching fluorescent microscopy images obtained at 10x magnification (Fig.~\ref{fig:patterns} except c and e), or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x (Fig.~\ref{fig:patterns}c and e).
+To measure wrinkle wavelength, we use large field pictures of the final state.
+These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification (Fig.~\ref{fig:patterns} except c and e), or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x (Fig.~\ref{fig:patterns}c and e).
 The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
@@
 
-In most of our samples it was possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
+In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
 However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
@@
 
-We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ was about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
-We used this approximation of $\tau$ for all our samples.
+We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
+We use this approximation of $\tau$ for all our samples.
 This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
@@
 By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
-Note that even with $H$ of the order of a few millimeters high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
+Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
 Obtaining $z_\infty$ without tube 2 would be time prohibitive.
@@
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The duration of the Video is 2~h~30.
+The duration of the Video is \SI{2}{\hour}~30.
 The scale bar is \SI{1}{\milli\metre}.
@@
 \item[Video 2] 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion (Supplementary Table~\ref{tab:data}c).
+10\% of the proteins are fluorescently labelled.
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The duration of the Video is 2~h~30.
+The duration of the Video is \SI{2}{\hour}~30.
 The scale bar is \SI{1}{\milli\metre}.
@@
 \item[Video 5] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
-1/10th of the proteins are fluorescently labelled.
+10\% of the proteins are fluorescently labelled.
 The duration of the Video is \SI{132}{\minute}.
@@
 \item[Video 6] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
-1/10th of the proteins are fluorescently labelled.
+10\% of the proteins are fluorescently labelled.
 The duration of the Video is \SI{147}{\minute}.
@@
 \item[Video 7] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a).
+10\% of the proteins are fluorescently labelled.
 The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
@@
 Darcy mechanism dominates the formation of the primary pattern.
-Total duration of the video is 145 min.
+Total duration of the video is \SI{145}{\minute}.
 
 \item[Video 8] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}c).
+10\% of the proteins are fluorescently labelled.
 The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
@@
 Poiseuille mechanism dominates the formation of the primary pattern.
-Total duration of the video is 181 min.
+Total duration of the video is \SI{181}{\minute}.
 

### 2015-02-13 fig_plis/article_plis.tex
Déplace Permeability measurements pour que les SF soient citées dans l'ordre

@@
 Around \SI{38}{\minute}, the amplitude of the wrinkles reach the thickness of the slit.
-The size of the primary pattern is now settled, Supplementary Figures~\ref{fig:patterns} and \ref{fig:growth}.
+The size of the primary pattern is now settled, Supplementary Figures~\ref{fig:generations} and \ref{fig:growth}.
 However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
@@
 We use either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:patterns} and \ref{fig:generations}).
+Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:generations} and \ref{fig:patterns}).
 Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
@@
 
-\begin{figure}[b]
+\begin{figure}
 	\includegraphics{nonmonotonic}%
@@
 Comparing $\chi$ across different setting of the microscope is not relevant.
+
+\begin{figure*}%
+	\includegraphics{permeability}%
+	\caption{Permeability measurements.
+(a) Schematic representation of the experiment.
+(b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
+The black line is the best exponential fit $Ae^{-t/\tau}$.
+(b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
+(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}%
+	\label{fig:permeability}%
+\end{figure*}
+\subsection*{Permeability measurements}
+
+To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}.
+First, the gel of interest is formed at the lower extremity of a thin glass tube.
+The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition).
+Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4.
+The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
+Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
+A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
+Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.
+
+Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
+\begin{equation}
+\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
+\end{equation}
+where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel.
+We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer.
+We can thus write
+\begin{equation}
+\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
+\end{equation}
+
+By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
+Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
+Obtaining $z_\infty$ without tube 2 would be time prohibitive.
 
@@
 
-\subsection*{Permeability measurements}
-\begin{figure*}
-	\includegraphics{permeability}
-	\caption{Permeability measurements.
-(a) Schematic representation of the experiment.
-(b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
-The black line is the best exponential fit $Ae^{-t/\tau}$.
-(b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
-(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}
-	\label{fig:permeability}
-\end{figure*}
-
-To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}.
-First, the gel of interest is formed at the lower extremity of a thin glass tube.
-The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition).
-Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4.
-The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
-Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
-A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
-Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.
-
-Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
-\begin{equation}
-\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
-\end{equation}
-where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel.
-We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer.
-We can thus write
-\begin{equation}
-\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
-\end{equation}
-
-By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
-Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
-Obtaining $z_\infty$ without tube 2 would be time prohibitive.
-
 \subsection*{Samples characteristics}

### 2015-02-13 fig_plis/article_plis.tex
Add and cite fig:notations

@@
 
+\begin{figure}
+\includegraphics{notations}%
+\caption{Schematic side view of the slit.
+The constant thickness gel film (orange) is surrounded by solvent.
+The region of interest of the mass conservation is highlighted in gray.
+Arrows are oriented as x or z to allow algebraic notations.}%
+\label{fig:notations}%
+\end{figure}
+
 The mechanical equilibrium of a plate of thickness $h$, Young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a transverse load $\sigma_\perp$ along $z$ writes
@@
 
-Mass conservation writes
+Mass conservation (on the lower gray area of Supplementary Figure~\ref{fig:notations}) writes
 \begin{equation}
@@
 considers a porous elastic film between two viscous layers of arbitrary thickness.
-In this case, the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
+In this case (see Supplementary Figure~\ref{fig:notations}), the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
 \begin{equation}

### 2015-02-14 fig_plis/article_plis.tex
Fig.->Figure Vérifie que tout les Sup sont bien notées Sup Change l'ordre des Supplementary pour suivre l'ordre du texte Equations Cite Dickinson2002 pour la suracidification micro-fingerprinting~\cite{Bae2015} plutot que food science dans la conclusion

@@
 In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
+As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
 Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
 We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
-As shown in Fig.~\ref{fig:dynamics} and in Supplementary Videos 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
+As shown in Figure~\ref{fig:dynamics} and in Supplementary Videos 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
 Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates.
-In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Fig.~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.
+In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Figure~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.
 
@@
 Successive generations of patterns are highlighted in colour to stress the absence of growth after formation.
-(b) 3D reconstruction from fluorescent confocal microscopy.
+(b) 3D reconstruction from fluorescent confocal microscopy (see Supplementary Table~\ref{tab:data}c).
 Scale bars are \SI{1}{\milli\metre}.}%
@@
 At neutral pH caseins are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), caseins molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998}.
+As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), caseins molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
 Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve.
@@
 To form a space spanning gel, flocculation needs to be prevented.
-We use of glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Fig.~\ref{fig:acidification}a.
+We use of glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Figure~\ref{fig:acidification}a.
 This GDL controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
@@
 When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall.
-In this situation we can monitor $\xi$, Fig.~\ref{fig:acidification}d.
+In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d.
 $\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.
@@
 This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
-We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Fig.~\ref{fig:sideview}b).
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Fig.~\ref{fig:sideview}c) and excess surface area of the gel film (Fig.~\ref{fig:sideview}d) (see Supplementary Methods).
+We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Figure~\ref{fig:sideview}b).
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Figure~\ref{fig:sideview}c) and excess surface area of the gel film (Figure~\ref{fig:sideview}d) (see Supplementary Methods).
 As expected from the system design, the evolution of the volume is non-monotonous.
@@
 Around \SI{38}{\minute}, the amplitude of the wrinkles reach the thickness of the slit.
-The size of the primary pattern is now settled, Supplementary Figures~\ref{fig:generations} and \ref{fig:growth}.
+The size of the primary pattern is now settled, Supplementary Figure~\ref{fig:growth}.
 However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
@@
 We also confirmed that at any given point in time and space the thickness $h$ of the film of gel is homogeneous.
-The contrast in transmission microscopy (Fig.~\ref{fig:dynamics}a) is therefore not due to thickness inhomogeneities but to altitude gradients.
-Indeed the length of the optical path along the vertical direction $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Fig.~\ref{fig:sideview}b.
+The contrast in transmission microscopy (Figure~\ref{fig:dynamics}a) is therefore not due to thickness inhomogeneities but to altitude gradients.
+Indeed the length of the optical path along the vertical direction $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Figure~\ref{fig:sideview}b.
 
@@
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Furthermore, (Eq.~\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+Furthermore, equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
 
@@
 Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_{\perp}/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_{\perp}}{h} = \frac{dA}{dt}$.
-Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
+Injecting this expression in equation~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
 \begin{equation}
@@
 This expression emphasizes the kinetic nature of the wavelength selection through Darcy flow.
-However we note that the wavelength of a generation of the pattern does not evolve with time (see Fig.~\ref{fig:dynamics}a and Sup.
-Fig.~\ref{fig:growth}) once the gel is touching both walls.
+However we note that the wavelength of a generation of the pattern does not evolve with time (see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}) once the gel is touching both walls.
 This arrest may be due to bending rigidity~\cite{LeGoff2014}.
@@
 Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_{\perp} \frac{H^3}{\eta\lambda^2}$~\cite{Poiseuille1842}.
-Injecting this expression in Eq.~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
+Injecting this expression in equation~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
 \begin{equation}
@@
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
-Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Fig.~\ref{fig:generations}.
+Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Figure~\ref{fig:generations}.
 
-Our results are displayed in Fig.~\ref{fig:DarcyPoiseuille}.
+Our results are displayed in Figure~\ref{fig:DarcyPoiseuille}.
 Although they provide the correct order of magnitude for $\lambda$, neither model alone is able to quantitatively reproduce the full data set.
@@
 Poiseuille) model fits the data well albeit with a prefactor.
-Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Fig.~\ref{fig:DarcyPoiseuille}b).
+Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b).
 Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.
@@
 Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
-As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in food science refXXX or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013}.
+As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013}.
 
@@
 Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:generations} and \ref{fig:patterns}).
-Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Fig.~\ref{fig:generations}.
-Data shown in Fig.~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
-Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Fig.~\ref{fig:permeability}).
+Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Figure~\ref{fig:generations}.
+Data shown in Figure~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
+Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Figure~\ref{fig:permeability}).
 The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
@@
 (b) Corresponding Fourier spectra when no adhesion on top, same times as (a).
-Lines are fits by Eq.~\ref{eq:fractalS}.
+Lines are fits by equation~(\ref{eq:fractalS}).
 The inset shows a schematic association between real space structures and wave numbers.
@@
 
-To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Fig.~\ref{fig:nonmonotonic}a).
-After Hann windowing we performe Fourier transform and radially average the resulting spectrum to obtain $I(q)$ (see Fig.~\ref{fig:nonmonotonic}b) that is fitted by the following fractal form:
+To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Supplementary Figure~\ref{fig:nonmonotonic}a and Supplementary Videos 5 and 6).
+After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$ (see Supplementary Figure~\ref{fig:nonmonotonic}b) that is fitted by the following fractal form:
 \begin{equation}
@@
 $\Gamma$ is the gamma function.
-Eq.~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
+Equation~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
 %En 3D on peut citer \cite{Martin1987,Chen1986}
 
-Figure~\ref{fig:acidification}c of the main text and Supplementary Fig.~\ref{fig:nonmonotonic}c and d show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
+Figure~\ref{fig:acidification}c of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c and d show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
 The later situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
@@
 In situation (i), the gel keeps its topology throughout the gelation and over acidification.
-Left of Fig.~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again.
+Left of Supplementary Figure~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again.
 Syn\ae{}resis and swelling are thus local processes.
@@
 The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
-Alongside the first tube (see Fig.~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
+Alongside the first tube (see Supplementary Figure~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
 A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
@@
 
-By fitting $z(t)-z_\infty$ to an exponential (see Fig.~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
+By fitting $z(t)-z_\infty$ to an exponential (see Supplementary Figure~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
 Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
@@
 
+\subsection*{Characteristic time measurement}
+\begin{figure}
+	\includegraphics{growth}
+	\caption{Growth of the instability.
+(a) Evolution of the gel velocity $v$ averaged over the whole surface.
+Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
+The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
+(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
+Note that the longitudinal size of the blister evolves in this regime.
+The thick black curve is the profile at much longer times and is at lower altitude due to swelling.
+Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
+	\label{fig:growth}
+\end{figure}
+
+In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
+However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
+Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
+
+We note that when both measurements were available (as in Supplementary Figure~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
+We use this approximation of $\tau$ for all our samples.
+This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Supplementary Figure~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
+
 \subsection*{Mixed Darcy--Poiseuille model}
@@
 \subsubsection*{General framework for wrinkling}
-
-\begin{figure}
-\includegraphics{notations}%
-\caption{Schematic side view of the slit.
-The constant thickness gel film (orange) is surrounded by solvent.
-The region of interest of the mass conservation is highlighted in gray.
-Arrows are oriented as x or z to allow algebraic notations.}%
-\label{fig:notations}%
-\end{figure}
 
@@
 
-We decomposed the deflection in sinusoidal modes $w(x,t) =  A \sin qx$ of wavevector $q$. Eq.~(\ref{eq:mecheq}) then reads
+We decomposed the deflection in sinusoidal modes $w(x,t) =  A \sin qx$ of wavevector $q$. Equation~(\ref{eq:mecheq}) then reads
 \begin{equation}
@@
 \subsubsection*{Elastic film on a viscous layer}
+
+\begin{figure}
+\includegraphics{notations}%
+\caption{Schematic side view of the slit.
+The constant thickness gel film (orange) is surrounded by solvent.
+The region of interest of the mass conservation is highlighted in gray.
+Arrows are oriented as x or z to allow algebraic notations.}%
+\label{fig:notations}%
+\end{figure}
+
 Let us now turn to the case of an elastic, impermeable film on a single viscous layer.
@@
 
-Combining Eqs.~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
+Combining Equations~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
 \begin{equation}
@@
 
-When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of Eq.~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness 
+When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of equation~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness
 \begin{equation}
@@
 
-Combining Eqs.~(\ref{eq:mecheqop}) and (\ref{eq:DarcyLoad}), the compression load writes
+Combining equations~(\ref{eq:mecheqop}) and (\ref{eq:DarcyLoad}), the compression load writes
 \begin{equation}
@@
 
-The mass conservation in the lower viscous layer is the same as Eq.~(\ref{eq:conservation}) with an added leak $v$ due to the porosity
+The mass conservation in the lower viscous layer is the same as equation~(\ref{eq:conservation}) with an added leak $v$ due to the porosity
 \begin{equation}
@@
 
-Combining Eqs.~(\ref{eq:pressures}, \ref{eq:conservationDarcy}, \ref{eq:Darcy}) we obtain the transverse load
+Combining equations~(\ref{eq:pressures}), (\ref{eq:conservationDarcy}) and (\ref{eq:Darcy}) we obtain the transverse load
 \begin{equation}
@@
 
-Note that we recover Eq~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and Eq.~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
+Note that we recover equation~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and equation~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
 We can recast this intuition in terms of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $q_P$ and $q_D$:
@@
 
-Analytic minimization of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome. Instead we rewrite Eq.~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
+Analytic minimization of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome.
+Instead we rewrite equation~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
 \begin{align}
@@
 and we minimize this expression numerically.
-
-\subsection*{Wavelength measurement}
-
-\begin{figure}
-	\includegraphics{generations}%
-	\caption{Fourier spectra of binary altitudes of Main Fig.~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-The scale bar is \SI{1}{\milli\metre}.}%
-	\label{fig:generations}%
-\end{figure}
-
-\begin{figure*}
-	\includegraphics{patterns}
-	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data} of the main text.
-(a-d) increasing cell thickness.
-(e-f) larger solid content.
-(g-i) increasing glycerol content.
-All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
-The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
-Length of arrows correspond to the measured primary wavelength.}
-	\label{fig:patterns}
-\end{figure*}
-
-To measure wrinkle wavelength, we use large field pictures of the final state.
-These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification (Fig.~\ref{fig:patterns} except c and e), or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x (Fig.~\ref{fig:patterns}c and e).
-The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
-White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
-It is thus possible to reconstruct a binary picture of the pattern (inset of Fig.~\ref{fig:generations}).
-Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final picture to pure first generation.
-
-Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
-When only the first generation pattern is present a peak clearly indicates the dominant mode.
-When second generation has appeared, the dominant wavelength is halved and the peak much weaker.
-This allows us to use secondary pattern is our analysis by setting their wavelength at half of the primary wavelength.
-Further generations do not seem to appear simultaneously and the peak is weak and broad.
-This could be a signature of a transition to chaos of a non-linear oscillator~\cite{Brau2011}.
-
-\subsection*{Characteristic time measurement}
-\begin{figure}
-	\includegraphics{growth}
-	\caption{Growth of the instability.
-(a) Evolution of the gel velocity $v$ averaged over the whole surface.
-Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
-The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
-(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
-Note that the longitudinal size of the blister evolves in this regime.
-The thick black curve is the profile at much longer times and is at lower altitude due to swelling.
-Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
-	\label{fig:growth}
-\end{figure}
-
-In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
-However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
-Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
-
-We note that when both measurements were available (as in Fig.~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
-We use this approximation of $\tau$ for all our samples.
-This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Fig.~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
 
@@
 \end{tabular}
-\caption{Fully characterised samples at the basis of Fig.~\ref{fig:DarcyPoiseuille}.
+\caption{Fully characterised samples at the basis of Figure~\ref{fig:DarcyPoiseuille}.
 Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.
@@
 \end{table*}
+
+\subsection*{Wavelength measurement}
+
+\begin{figure}
+	\includegraphics{generations}%
+	\caption{Fourier spectra of binary altitudes of Main Figure~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
+The scale bar is \SI{1}{\milli\metre}.}%
+	\label{fig:generations}%
+\end{figure}
+
+\begin{figure*}
+	\includegraphics{patterns}
+	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data}.
+(a-d) increasing cell thickness.
+(e-f) larger solid content.
+(g-i) increasing glycerol content.
+All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
+The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
+Length of arrows correspond to the measured primary wavelength.}
+	\label{fig:patterns}
+\end{figure*}
+
+To measure wrinkle wavelength, we use large field pictures of the final state.
+These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification (Supplementary Figure~\ref{fig:patterns} except c and e), or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x (Supplementary Figure~\ref{fig:patterns}c and e).
+The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
+White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
+It is thus possible to reconstruct a binary picture of the pattern (inset of Supplementary Figure~\ref{fig:generations}).
+Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final picture to pure first generation.
+
+Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
+When only the first generation pattern is present a peak clearly indicates the dominant mode.
+When second generation has appeared, the dominant wavelength is halved and the peak much weaker.
+This allows us to use secondary pattern is our analysis by setting their wavelength at half of the primary wavelength.
+Further generations do not seem to appear simultaneously and the peak is weak and broad.
+This could be a signature of a transition to chaos of a non-linear oscillator~\cite{Brau2011}.
 

### 2015-02-14 fig_plis/article_plis.tex
Abstract minimise, characterise (en_GB) K\'arm\'an water -> solvent Ccl: pas court vs long mais proche/loin des murs

@@
 \begin{abstract}
-The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels -- this is indeed one of the key steps to make yoghurt.
-Here, using the casein proteins as  building blocks and the acidulent glucono-$\delta$-lactone, we design model thin biogel layers that wrinkles in cascade.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during the acidification, the gel shrinks and swells creating constrains that induce Darcy and Poiseuille flows and wrinkles the gel.
-We develop an elasto-hydrodynamic model that predicts correctly the wrinkling wavelength.
-Our results should be universal for acid-induced protein gels as it is based on pH induced charge stabilisation/destabilisation and on control of boundary conditions.
+The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels, e.g.
+yoghurt.
+Less known is the partial redissolution of the gel at even lower pH.
+Here, using caseins as pH-responsive bulding blocks and the acidulent glucono-$\delta$-lactone, we design a model biogel film that spontaneously wrinkles in cascade, provided non-adhesive confinement.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing winkling.
+Taking into account both Darcy flow though the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
+Our results should be universal for acid-induced protein gels as it is based on pH-induced charge stabilisation/destabilisation.
 As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
@@
 On the biology side wrinkling-controlled morphogenesis is ubiquitous.
-Ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}.
+Ageing and the loss of elastic fibres makes our skin wrinkle~\cite{Bissett1987, Genzer2006}.
 Difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}.
-Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
+Localised cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
 On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
@@
 
-\begin{figure}[ht!]
+\begin{figure}
 	\includegraphics{dynamics}%
@@
 However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary Video 4.
-The state diagram of caseins as a function of the pH is related to charge stabilization.
+The state diagram of caseins as a function of the pH is related to charge stabilisation.
 At neutral pH caseins are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), caseins molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
-Beyond the isoelectric point casein acquire a positive net charge and therefore partially re-dissolve.
+As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), casein molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+Beyond the isoelectric point caseins acquire a positive net charge and therefore partially re-dissolve.
 In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH (see Supplementary Methods).
 Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction $x_\text{free}$ is 1.
-Upon decreasing the pH of the solution to the isolectric point the free casein fraction drop to almost 0, almost all the casein molecule are aggregated.
+Upon decreasing the pH of the solution to the isolectric point the free casein fraction drops to almost 0, almost all the casein molecule are aggregated.
 Further decrease of the pH of the solution lead to a partial redissolution of the casein aggregates.
@@
 To form a space spanning gel, flocculation needs to be prevented.
-We use of glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Figure~\ref{fig:acidification}a.
+We use glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Figure~\ref{fig:acidification}a.
 This GDL controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
@@
 After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
-Shortly after,  at \SI{36}{\minute}, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area.
+Shortly after,  at \SI{36}{\minute}, the planar film destabilises and starts to wrinkle in the $z$-direction creating excess area.
 This confirms that the instability is due to swelling.
@@
 Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off.
-We also confirmed that at any given point in time and space the thickness $h$ of the film of gel is homogeneous.
+We also confirm that at any given point in time and space the thickness $h$ of the film of gel is homogeneous.
 The contrast in transmission microscopy (Figure~\ref{fig:dynamics}a) is therefore not due to thickness inhomogeneities but to altitude gradients.
@@
 The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem.
-Large deformations of thin sheets is governed by F\"{o}ppl-von Karman equations~\cite{landau1986}, essentially impossible to solve in analytical form.
+Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, essentially impossible to solve in analytical form.
 However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}.
@@
 In our situation we now have to determine $\sigma_\perp$.
-We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding water layers.
+We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding solvent layers.
 We confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
@@
 where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
-This expression emphasizes the kinetic nature of the wavelength selection through Darcy flow.
-However we note that the wavelength of a generation of the pattern does not evolve with time (see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}) once the gel is touching both walls.
+This expression emphasises the kinetic nature of the wavelength selection through Darcy flow.
+However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls(see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}).
 This arrest may be due to bending rigidity~\cite{LeGoff2014}.
@@
 \end{equation}
-where we recognize the scaling for the wrinkles of an elastic film on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate.
+where we recognise the scaling for the wrinkles of an elastic film on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate.
 Indeed within the lubrication approximation or at low Reynolds numbers one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
@@
 Darcy and Poiseuille models yield wavelengths that are of the same order in most of the cases under study.
-Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method for details on such measurements).
-As summarized in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
+Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method on such measurements).
+As summarised in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
 Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly.
@@
 Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b).
-Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for all values of $H/H^*$.
+Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for the whole $H/H^*$ range.
 
@@
 In conclusion we have taken advantage of the properties of the protein casein and carefully engineered the route to form permeable biogels that, in slit geometry, produce original cascade of wrinkling patterns when casein adhesion to the slit walls is turned off.
-We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flow of the solvent through the gel layer for short wavelengths.
+We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall.
 Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
@@
 The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
-To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which a polyacrylamide brush is polymerized from methacrylate groups (Euromedex).
+To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which a polyacrylamide brush is polymerised from methacrylate groups (Euromedex).
 
@@
 
-Minimizing $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
+Minimising $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
 \begin{equation}
@@
 \end{equation}
-and is minimized at $q_P$ such that
+and is minimised at $q_P$ such that
 \begin{equation}
@@
 \end{equation}
-and is minimized at $q_D$ such that
+and is minimised at $q_D$ such that
 \begin{equation}
@@
 
-Analytic minimization of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome.
+Analytic minimisation of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome.
 Instead we rewrite equation~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
@@
 \end{align}
-and we minimize this expression numerically.
+and we minimise this expression numerically.
 
 \subsection*{Samples characteristics}
-To test the theoretical predictions we have fully characterized 9 samples, see Supplementary Table \ref{tab:data}.
+To test the theoretical predictions we have fully characterised 9 samples, see Supplementary Table \ref{tab:data}.
 

### 2015-02-16 fig_plis/article_plis.tex
VPS et arrested phase separation

@@
 This GDL controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
-As shown on Figure~\ref{fig:acidification}a and b, when adding enough GDL the pH lowers beyond the isoeletric point and $x_\text{free}$ rises again.
+This process is kinetically driven by diffusion limited cluster aggregation~\cite{Bremer1989}.
+As shown on Figure~\ref{fig:acidification}a and b, when adding enough GDL the pH lowers beyond the isoelectric point shifting the balance between attraction and charge repulsion at the casein level, so that casein partially dissociate from the fractal network.
+This results in the coexistence between a dilute phase $x_\text{free}\approx 5-20\%$ and a dense glassy network, a scenario reminiscent of arrested phase separation~\cite{Cardinaux2007,lu2008gelation}.
+
+Here the attraction between caseins is weakened enough so that we observe the coexistence between a dilute phase $x_\text{free}\approx 5-20\%$ and a dense glassy network, a scenario reminiscent of arrested phase separation.
 We monitor the GDL induced gelation with a rheometer in cone plate geometry.
@@
 The mechanical properties of the gel are closely related to the fraction of free casein.
-First, the dispersion solidifies then reaches its maximum elasticity (500 Pa) at the isoelectric point and finally the elastic modulus decreases down to 100 Pa.
+First, the dispersion solidifies then reaches its maximum elasticity (\SI{500}{\pascal}) at the isoelectric point and finally the elastic modulus decreases down to \SI{100}{\pascal}.
 This finite value indicates that the sample is still a gel.
@@
 The fractal model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
-As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998}.
+As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
 This stiffens the arms of the network and is responsible for the increase of elasticity.
@@
 As caseins regain solubility, the strands of the network swell, the pores shrink and the gel weakens.
+
+This process is reminiscent of arrested phase separation where a dilute phase coexist with a dense glassy network.
 
@@
 
-\section*{conclusion}
+\section*{Conclusion}
 
-In conclusion we have taken advantage of the properties of the protein casein and carefully engineered the route to form permeable biogels that, in slit geometry, produce original cascade of wrinkling patterns when casein adhesion to the slit walls is turned off.
+In conclusion we have taken advantage of the properties of the protein casein.
+In particular, we have fine-tuned the interplay between the attraction and the repulsion at the protein level to assemble dense glassy networks that shrink and swell as a result of the transition from a fractal gel to an arrested phase separation.
+This carefully engineered route to form permeable biogels is the key, with controlled adhesion of the gels to the boundaries, to produce original cascade of wrinkling patterns.
 We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall.
@@
 Data shown in Figure~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
+For every gel composition we monitor the gelation by small amplitude oscillations (0.1\%, \SI{1}{\hertz}) and estimate the flexural modulus $B$ from the maximum value of the shear modulus $G^\prime$ as $B=2G^\prime/(1-\nu)$ taking $\nu=0.3$ for the Poisson ratio, typical value for a spongy network~\cite{Greaves2011}.
 Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Figure~\ref{fig:permeability}).

### 2015-02-16 fig_plis/article_plis.tex
Fleury2015

@@
 epithelimum or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g.
-lymph, blood or mucus, a practical situation where a thin viscous substrate would set the wrinkling wavelength.
+lymph, blood or mucus, a practical situation where a thin viscous substrate would set the wrinkling wavelength~\cite{Fleury2015}.
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
@@
 Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
-As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013}.
+As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
 

### 2015-02-16 fig_plis/article_plis.tex
typo Thomas

@@
 
-To form wrinkled patterns such as the one observed in figure~\ref{fig:dynamics}, ones needs to take advantage of the pH sensitivity of caseins to control the route to gelation, to fine tune the gel properties and the boundary conditions.
+To form wrinkled patterns such as the one observed in figure~\ref{fig:dynamics}, ones needs to take advantage of the pH sensitivity of caseins, to control the route to gelation, to fine tune the gel properties and the boundary conditions.
 
@@
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-To turn off the adsorption of the caseins on a surface, we coat it with polymer brushes (see Methods), effectively creating non adhesive boundary conditions for the gel.
+To turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes (see Methods), effectively creating non adhesive boundary conditions for the gel.
 The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel: $\xi$ increases, the stands of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of the network swell and the gel wants to expend.
@@
 The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem.
-Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, essentially impossible to solve in analytical form.
+Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form.
 However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}.
@@
 The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL.
-In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of the pH, and centrifuged in eppendorf tubes at 14500 rpm for \SI{3}{\minute}.
-The supernatant is diluted and analysed with a UV absorption spectrometer (Ocean Optics).
+In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of the pH, and centrifuged (centrifuge 5424, Eppendorf) in Eppendorf tubes at 14500 rpm for \SI{3}{\minute}.
+The supernatant is diluted and analysed with a UV absorption spectrometer (USB4000, Ocean Optics).
 According to~\cite{Roefs1986}, we deduced the fraction of free casein proteins from the adsorption spectrum:

### 2015-02-16 fig_plis/cover_letter.tex
Cover letter

\documentclass[a4paper, parskip=true, firsthead=false, fromemail=true, foldmarks=false]{scrlttr2}
\usepackage[utf8]{inputenc}
\setkomavar{fromemail}{thomas.gibaud@ens-lyon.fr}
\setkomavar{signature}{Mathieu Leocmach,\\ Mathieu Nespoulous,\\ Sébastien Manneville,\\ Thomas Gibaud\\
{\small Ecole Normale Supérieure de Lyon}}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[british]{babel}

\usepackage{kpfonts}
\usepackage{tabu}

\begin{document}
\begin{letter}{From:\\
Thomas Gibaud,\\
Laboratoire de physique,\\
Ecole Normale Supérieure de Lyon,\\
\texttt{thomas.gibaud@ens-lyon.fr}
}
\opening{\bf Dear Editor,}

Please find enclosed our manuscript, entitled \emph{Hierarchical wrinkling in permeable and confined biogels}, which we are submitting for publication as an article to Nature Materials.

Supplementary video 1 arouse one's curiosity. It shows microscopy images of the biogel as a function of time in and optical cell where the adhesion of the protein to the cell wall is prohibited. How can a simple protein gel form spontaneously such complicated nested patterns?

Supplementary video 2, which shows a 3D construction of the gel based on confocal imaging, hints towards the answer. The gel shrinks then swells and wrinkles due to the excess area created during swelling. This is however not the end of the story as boundary conditions frustrate the development of the wrinkles and induce buckling within the buckling . This hierarchical pattern is a beautiful illustration of the well known Droste effect named after the cacao brand whoes trademark image displays a nurse carrying a serving tray with a cup of hot chocolate and a box with the same image. In France it is known as the    ``Vache qui rit'' effect and of course one thinks about Russian dolls.

\begin{tabu}{X[c]X[c]}
\includegraphics[height=6\baselineskip]{../Droste.jpg}&
\includegraphics[height=6\baselineskip]{../vachequirit.jpg}
\end{tabu}

It is universally recognized that confined thin surfaces may wrinkle due the growth of an excess of material. For example on the biology side different rates of growth between the gut tube and its dorsal anchoring is responsible for the vilification of guts. Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern. Such patterns can also be triggered by physical cues as temperature dilation, swelling or the removal of pre-strain.

Yet, benchmark experiments such as the one presented in Supplementary Video 1 and 2, that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line. Indeed, stability analysis of a film lying on a thin viscous substrate was studied only recently and remains theoretical because, on a free interface, gravity dominates over viscous substrate.

From a fundamental perspective, our work unifies concepts from diverse soft matter systems.
\begin{enumerate}
\item From food science we have hijacked the process of yoghurt making to form gels. Usually, when you acidify milk the caseins flocculate. To form a uniform yoghurt gel one needs to lower homogeneously the pH, as bacteria do. Here we use the acidulent GDL for finer control.

\item  Next, we exploit the concept of stability in colloidal suspensions. As the GDL acidifies the medium, caseins loose their net charge and aggregate. Upon further acidification the casein regain charges and partially desorb form the gel. This makes the gel successively shrink and swell.

\item This process also can be well understood by invoking two different theories of gelation. Diffusion limited cluster aggregation is responsible for the formation of the fractal network of the gel ; whereas arrested phase separation well describes the coexistence between a dilute phase and a dense glassy network.

\item Finally, we invoke hydrodynamics with the seminal work of Darcy and Poiseuille. In particular we demonstrate the kinetic selection of the wavelength due to two sources of viscous dissipation. We predict the wrinkling wavelength invoking Poiseuille flows of the solvent above and below the gel layer for long wavelengths and a new mechanism, Darcy flows of the solvent through the gel layer for short wavelengths.
\end{enumerate}

From a practical perspective, such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins. Controlled gelation with peculiar boundary conditions introduces a robust method to produce wrinkled surfaces that is distinct from other emerging technologies. We use a combination of complementary techniques ranging from polymer brush grafting, titration, light microscopy, rheology, permeability measurements and confocal microscopy that sustains studies to better understand the relations between all levels of hierarchy from the individual proteins to the macroscopic. Indeed, we directly measure and relate the non monotonous pH response of the proteins to the spontaneous shrinkage and swelling effect of the casein network at the micron level to the wrinkles at the millimetre scale. Playing with the geometry, the gel composition and the viscosity of the solvent, we explore a wide range of physical parameters to check the validity of our models. All of these differences enable us to assemble, to our knowledge, the first example of Darcy flow driven wrinkling.

The impact of our work extends beyond soft-matter physics. For instance, it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting or in optical devices such as diffraction gratings or Fresnel lenses, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues. Indeed during the morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.

For all these reasons we believe that our manuscript is suitable for a publication in an interdisciplinary journal such as XXX. With this manuscript we are including XXX supplementary movies, which have been mailed to your editorial office. As possible reviewers we suggest: L. Mahadevan (Harvard, USA), N. Menon (Amherst, USA), José Bico (ESPCI, France), Enrique Cerda (Univ. Santiago-Chile, Chile), Mokhtar-Adda-Bedia (ENS Paris, France), Shu Yang (U.Penn, USA), Pascal Damman (Univ. Mons, Belgium), Tian Yang (Univ. Michigan, USA), Dominic Vella (Oxford, UK) XXX.

We thank you for your time and attention.

\closing{\bf Sincerely yours,}

\end{letter} 
\end{document}

### 2015-02-16 fig_plis/cover_letter.tex
typos et raccourcit la description des images

@@
 
-Supplementary video 1 arouse one's curiosity.
-It shows microscopy images of the biogel as a function of time in and optical cell where the adhesion of the protein to the cell wall is prohibited.
-How can a simple protein gel form spontaneously such complicated nested patterns?
+Supplementary Video 1 arouses one's curiosity.
+It shows microscopy images of a thin optical cell containing a protein gel whose adhesion to the cell wall is prohibited.
+No external stimulus is applied.
+How can a simple biogel form spontaneously such complicated nested patterns?
 
-Supplementary video 2, which shows a 3D construction of the gel based on confocal imaging, hints towards the answer.
+Supplementary Video 2, which shows a 3D construction of the gel based on confocal imaging, hints towards the answer.
 The gel shrinks then swells and wrinkles due to the excess area created during swelling.
 This is however not the end of the story as boundary conditions frustrate the development of the wrinkles and induce buckling within the buckling .
-This hierarchical pattern is a beautiful illustration of the well known Droste effect named after the cacao brand whoes trademark image displays a nurse carrying a serving tray with a cup of hot chocolate and a box with the same image.
-In France it is known as the    ``Vache qui rit'' effect and of course one thinks about Russian dolls.
+This hierarchical pattern is a beautiful illustration of the well known Droste effect, known in France as ``Vache qui rit'' effect, where an image contains a smaller version of itself.
 
 \begin{tabu}{X[c]X[c]}
-\includegraphics[height=6\baselineskip]{../Droste.jpg}&
-\includegraphics[height=6\baselineskip]{../vachequirit.jpg}
+\includegraphics[height=8\baselineskip]{../Droste.jpg}&
+\includegraphics[height=8\baselineskip]{../vachequirit.jpg}
 \end{tabu}
@@
 
-Yet, benchmark experiments such as the one presented in Supplementary Video 1 and 2, that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line.
-Indeed, stability analysis of a film lying on a thin viscous substrate was studied only recently and remains theoretical because, on a free interface, gravity dominates over viscous substrate.
+Yet, benchmark experiments such as the one presented in Supplementary Videos 1 and 2, that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line.
+Indeed, stability analysis of a film lying on a thin viscous substrate was studied only recently and remains theoretical because, on a free interface, gravity dominates over viscous substrate as the wavelength selection mechanism.
 
@@
 As the GDL acidifies the medium, caseins loose their net charge and aggregate.
-Upon further acidification the casein regain charges and partially desorb form the gel.
+Upon further acidification the caseins regain charges and partially desorb from the gel.
 This makes the gel successively shrink and swell.
@@
 
-For all these reasons we believe that our manuscript is suitable for a publication in an interdisciplinary journal such as XXX.
-With this manuscript we are including XXX supplementary movies, which have been mailed to your editorial office.
+For all these reasons we believe that our manuscript is suitable for a publication in an interdisciplinary journal such as Nature Materials.
+With this manuscript we are including 8 Supplementary Videos, which have been mailed to your editorial office.
 As possible reviewers we suggest: L.

### 2015-02-17 fig_plis/article_plis.tex
début des remaques de Seb: - vire Video 3. Numérote les Vidéos avec des labels et des refs - vire Roman de la fin d'intro - HCl fluccule, pas GDL - tighter packing of caseins - Finally if top and bottom turned off

@@
 Less known is the partial redissolution of the gel at even lower pH.
-Here, using caseins as pH-responsive bulding blocks and the acidulent glucono-$\delta$-lactone, we design a model biogel film that spontaneously wrinkles in cascade, provided non-adhesive confinement.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing winkling.
-Taking into account both Darcy flow though the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
-Our results should be universal for acid-induced protein gels as it is based on pH-induced charge stabilisation/destabilisation.
-As such it could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
+Here, using caseins as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, we design a model biogel film that spontaneously wrinkles in cascade, provided non-adhesive confinement.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
+Taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
+Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilisation/destabilisation.
+As such our results could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
 \end{abstract}
@@
 
-Yet, benchmark experiments that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line, not only from a physics point of view, but also to get fundamental insights into biologically relevant situation.
-Indeed, stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002} and remains theoretical because, on a free interface, gravity dominates over viscous substrate.
-In a biological context, thin tissues, e.g.
+Yet, benchmark experiments that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line, not only from a physics point of view, but also to get fundamental insights into biologically relevant situations.
+Indeed, stability analysis of a film lying on a thin viscous substrate was performed only recently~\cite{Huang2002} and remains theoretical because, on a free interface, gravity dominates over viscous substrate.
+In a biological context, however, thin tissues, e.g.
 epithelimum or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g.
@@
 In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002} -- this is a well-known effect used in yoghurt making -- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off.
-Using a combination of rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
+As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002} --this is a well-known effect used in yoghurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off, Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{vid:transmitted} and \ref{vid:reconstructed}.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
 We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
-As shown in Figure~\ref{fig:dynamics} and in Supplementary Videos 1 and 2, transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary buckling.
-Such cascade buckling has already been observed by~\cite{Roman1999} in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates.
-In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Figure~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.
+The main goal of the present study is to unveil the physical origin of the observed patterns and to identify the selection mechanism of the wavelength.
 
@@
 	\includegraphics{dynamics}%
-	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w).
+	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w in water).
 (a) Light transmission microscopy.
-Successive generations of patterns are highlighted in colour to stress the absence of growth after formation.
-(b) 3D reconstruction from fluorescent confocal microscopy (see Supplementary Table~\ref{tab:data}c).
+Successive generations of patterns are highlighted in colour to stress the absence of coarsening after formation.
+(b) 3D reconstruction from fluorescent confocal microscopy, see Supplementary Table~\ref{tab:data}c.
 Scale bars are \SI{1}{\milli\metre}.}%
@@
 
-The main goal of the present study is thus to identify the selection mechanism of the wavelength.
 In the following we first describe the physical chemistry that allows us to form such patterns.
-Then we  model the situation and identify two possible selection mechanisms, namely porous Darcy flow or viscous Poiseuille flow.
-Comparing our experimental data to these two models we finally show that a combination of both Darcy and Poiseuille mechanisms accurately describes our data.
+We then show how transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary patterns.
+Finally we model the situation and identify two possible selection mechanisms, namely porous Darcy flow or viscous Poiseuille flow.
+Comparing our experimental data to these two models we show that a combination of both Darcy and Poiseuille mechanisms accurately describes our data.
 This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
@@
 	\caption{Acid-induced protein gel properties behave non monotonically with pH.
-(a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL.
+(a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL in water.
 The horizontal line indicates isoelectric pH of caseins.
@@
 Symbols are direct measures upon addition of GDL.
-Continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert).
-(c) Evolution of the storage modulus $G'$ measured in a rheometer with full adhesion to rotor and stator.
+The continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert).
+(c) Evolution of the elastic modulus $G'$ measured in a rheometer with full adhesion to the cone-plane geometry.
 (d) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
@@
 The stability of caseins dispersion strongly depend on the pH of the solution.
-Supplementary Video 3 shows the titration of a caseins dispersion at 1\%w by 1 molar HCl.
-At pH=7, the caseins are well dispersed then then flocculate between pH=5.5 to 4 and finally at pH=3 the solution become clear again.
-However as one increases the caseins concentration up to 4\%w, at low pH the caseins redispersion is not total, Supplementary Video 4.
-The state diagram of caseins as a function of the pH is related to charge stabilisation.
+Supplementary Video~\ref{vid:titration} shows the titration of a casein dispersion at 4\%w by 1 molar HCl.
+At pH=7, the caseins are well dispersed then they flocculate between pH=5.5 to 4 and finally at pH=3 the solution becomes clearer as caseins partially redisperse.
+The state diagram of caseins as a function of pH is thus related to charge stabilisation.
 At neutral pH caseins are charged negatively and repel each other.
@@
 Beyond the isoelectric point caseins acquire a positive net charge and therefore partially re-dissolve.
-In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH (see Supplementary Methods).
+In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH, see Supplementary Methods.
 Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction $x_\text{free}$ is 1.
-Upon decreasing the pH of the solution to the isolectric point the free casein fraction drops to almost 0, almost all the casein molecule are aggregated.
-Further decrease of the pH of the solution lead to a partial redissolution of the casein aggregates.
+Upon decreasing the pH to the isolectric point the free casein fraction drops to almost 0: almost all the casein molecules are aggregated.
+Further decrease of the pH of the solution leads to a partial redissolution of the casein aggregates.
 At pH=2 the fraction of free casein is 0.35.
@@
 
-To form a space spanning gel, flocculation needs to be prevented.
-We use glucono-$\delta$-lactone (GDL) a molecule  that slowly hydrolyses into an acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Figure~\ref{fig:acidification}a.
-This GDL controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
+When acidified by a strong acid like HCl, caseins flocculate and do not form a space-spanning gel.
+Therfore, we use glucono-$\delta$-lactone (GDL) a molecule that slowly hydrolyses into gluconic acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Figure~\ref{fig:acidification}a.
+This GDL-controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
 This process is kinetically driven by diffusion limited cluster aggregation~\cite{Bremer1989}.
-As shown on Figure~\ref{fig:acidification}a and b, when adding enough GDL the pH lowers beyond the isoelectric point shifting the balance between attraction and charge repulsion at the casein level, so that casein partially dissociate from the fractal network.
+As shown in Figure~\ref{fig:acidification}a and b, when adding enough GDL the pH decreases beyond the isoelectric point shifting the balance between attraction and charge repulsion at the casein level, so that caseins partially dissociate from the fractal network.
 This results in the coexistence between a dilute phase $x_\text{free}\approx 5-20\%$ and a dense glassy network, a scenario reminiscent of arrested phase separation~\cite{Cardinaux2007,lu2008gelation}.
 
-Here the attraction between caseins is weakened enough so that we observe the coexistence between a dilute phase $x_\text{free}\approx 5-20\%$ and a dense glassy network, a scenario reminiscent of arrested phase separation.
-We monitor the GDL induced gelation with a rheometer in cone plate geometry.
-As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elastic and viscous modulus of the casein dispersion as a function of time, Figure~\ref{fig:acidification}c.
+We monitor the GDL-induced gelation with a rheometer in cone-plate geometry.
+As casein strongly adsorbs to surfaces, the strain and stress are well transmitted to the casein dispersion and we can measure the elastic and viscous moduli of the casein dispersion as a function of time, Figure~\ref{fig:acidification}c.
 The mechanical properties of the gel are closely related to the fraction of free casein.
-First, the dispersion solidifies then reaches its maximum elasticity (\SI{500}{\pascal}) at the isoelectric point and finally the elastic modulus decreases down to \SI{100}{\pascal}.
+First, the dispersion solidifies then reaches its maximum elasticity ($\approx\SI{500}{\pascal}$) at the isoelectric point and finally the elastic modulus decreases down to $\approx\SI{100}{\pascal}$.
 This finite value indicates that the sample is still a gel.
 
-To gain further insight into the mechanical properties of the gel we turn to micro structural observations under confocal microscopy.
-As shown in Supplementary Figure~\ref{fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
+To gain further insight into the mechanical properties of the gel we turn to microstructural observations under confocal microscopy.
+As shown in Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63} and in Supplementary Figure~\ref{fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut-off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
 Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~\ref{fig:permeability}.
@@
 As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
-This stiffens the arms of the network and is responsible for the increase of elasticity.
-Beyond the isoelectric point $\xi$ decreases down to \SI{4}{\micro\metre}.
-As caseins regain solubility, the strands of the network swell, the pores shrink and the gel weakens.
-
-This process is reminiscent of arrested phase separation where a dilute phase coexist with a dense glassy network.
+This corresponds to a tighter packing of the caseins within the network strands.
+Such stiffening of the strands explains the observed increase of elasticity.
+Beyond the isoelectric point caseins partially regain solubility, the packing within the strands loosens which results in weaker, thicker strands and smaller pores as $\xi$ decreases down to \SI{4}{\micro\metre}, hence the long-term decrease of the elastic modulus.
 
@@
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-To turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes (see Methods), effectively creating non adhesive boundary conditions for the gel.
-The evolution of $\xi$ when sticky boundaries are on reflects the inner constrains of the gel: $\xi$ increases, the stands of the network become thinner and the gel wants to retract; then $\xi$ decreases, the stands of the network swell and the gel wants to expend.
+In order to turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes, see Methods, effectively creating non adhesive boundary conditions for the gel.
 When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall.
@@
 $\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.
+
+Finally, if the adhesion to both top and bottom walls is turned off, the film wrinkles as shown in Figure~\ref{fig:dynamics}.
+Such cascade buckling has already been observed in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates~\cite{Roman1999}.
+In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Figure~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.
 
@@
 
-Figure~\ref{fig:sideview}a shows confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration favourable to wrinkling (see also Supplementary Videos 7 and 8).
+Figure~\ref{fig:sideview}a shows successive side views of confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration that induces wrinkling, see also Supplementary Videos~\ref{vid:cutDarcy} and \ref{vid:cutPoiseuille}.
 Initially the dispersion is homogeneous and fills the entire slit.
 The black margins correspond to the glass walls.
-Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated - the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
+Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated --the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
 This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
-We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$ (Figure~\ref{fig:sideview}b).
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume (Figure~\ref{fig:sideview}c) and excess surface area of the gel film (Figure~\ref{fig:sideview}d) (see Supplementary Methods).
-As expected from the system design, the evolution of the volume is non-monotonous.
+We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$,  Figure~\ref{fig:sideview}b.
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}c, and excess surface area of the gel film, Figure~\ref{fig:sideview}d, see Supplementary Methods.
+As expected from the system design, the evolution of the gel volume is non-monotonous.
 After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
@@
 This confirms that the instability is due to swelling.
-Around \SI{38}{\minute}, the amplitude of the wrinkles reach the thickness of the slit.
-The size of the primary pattern is now settled, Supplementary Figure~\ref{fig:growth}.
+Around \SI{38}{\minute}, the amplitude of the wrinkles reaches the thickness of the slit.
+The size of the primary pattern is now fixed, Supplementary Figure~\ref{fig:growth}.
 However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
-When the excess volume saturates, the dynamics stops.
+When the excess area saturates, the dynamics stop (\SI{3}{\hour}).
 The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
@@
 Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off.
-We also confirm that at any given point in time and space the thickness $h$ of the film of gel is homogeneous.
-The contrast in transmission microscopy (Figure~\ref{fig:dynamics}a) is therefore not due to thickness inhomogeneities but to altitude gradients.
+We also confirm that at any given point in time the thickness $h$ of the film of gel is spatially homogeneous.
+The contrast in transmission microscopy, Figure~\ref{fig:dynamics}a, is therefore not due to thickness inhomogeneities but to altitude gradients.
 Indeed the length of the optical path along the vertical direction $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Figure~\ref{fig:sideview}b.
@@
 Yellow lines show the path of transmitted light through the gel phase.
-A region of interest is highlighted in gray.
 (c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
@@
 Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form.
-However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern and isolate the physically meaningful ingredients~\cite{Cerda2003}.
+However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern up to dimensionless prefactors and isolate the physically meaningful ingredients~\cite{Cerda2003}.
 An elastic film buckles due to excess area with respect to its boundaries.
@@
 	\includegraphics{Darcy_vs_Poiseuille}
-	\caption{Comparing model predictions with measured wavelengths.
+	\caption{Comparing model predictions $\lambda_D$, $\lambda_P$ and $\lambda_{D+P}$ with measured wavelengths $\lambda_\text{exp}$.
 Dots come from primary pattern, squares from secondary blisters.
-Lines are the best linear fit through the origin taking only into account the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
+Lines are the best linear fits through the origin taking into account only the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
 Prefactors are 0.63, 0.69 and 0.67 respectively.
-The dashed line is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all the points in (b).}
+The dashed line in (b) is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all data points.}
 	\label{fig:DarcyPoiseuille}
@@
 
-We consider two limit scenarios that leads to wrinkling patterns.
+Therefore we consider two limit scenarios that leads to wrinkling patterns.
 In a first limit, the film of gel sits without sticking to the bottom wall of the cell, i.e.
-$H_1\rightarrow0$, see Supplementary Video 7.
-Creating a 2D blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
-Darcy law~\cite{Darcy1856} relates the pressure gradient, estimated as $\sigma_{\perp}/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{\sigma_{\perp}}{h} = \frac{dA}{dt}$.
-Injecting this expression in equation~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
+$H_1\rightarrow0$, see Supplementary Video~\ref{vid:cutDarcy}.
+Creating a two dimensional blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
+Darcy law~\cite{Darcy1856} relates the pressure gradient, $(p_2-p_1)/h$, to the volume flux per unit area $v = \frac{\alpha}{\eta}\frac{p_2-p_1}{h} = \frac{dA}{dt}$.
+Identifying $\sigma_{\perp}=p_2-p_1$ and injecting this expression in equation~(\ref{eq:lstar}) yields the wavelength of the Darcy mode as
 \begin{equation}
@@
 This expression emphasises the kinetic nature of the wavelength selection through Darcy flow.
-However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls(see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}).
+However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}.
 This arrest may be due to bending rigidity~\cite{LeGoff2014}.
-Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate (see Supplementary Method for the precise definition).
+Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate, see Supplementary Methods for the precise definition.
 
-The opposite limiting case is when the gel film lies in the middle of the cell ($H_1=H_2$, see Supplementary Video 8) and impermeable ($\alpha\rightarrow 0$).
+The opposite limiting case is when the gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Supplementary Video~\ref{vid:cutPoiseuille}) and impermeable ($\alpha\rightarrow 0$).
 Here destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
-By symmetry the transverse load across the gel $\sigma_{\perp}$ is also the pressure difference over the wavelength.
-Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_{\perp} \frac{H^3}{\eta\lambda^2}$~\cite{Poiseuille1842}.
-Injecting this expression in equation~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
+By symmetry the transverse load across the gel $\sigma_{\perp}=p_2-p_1$ is also the pressure difference over the wavelength.
+Using a Poiseuille profile for the flow, $\frac{dA}{dt} \sim \sigma_{\perp} \frac{H^3}{\eta\lambda^2}$~\cite{Poiseuille1842} and injecting this expression in equation~(\ref{eq:lstar}) yields the wavelength of the Poiseuille mode as
 \begin{equation}
@@
 
-A more complete derivation of the models can be found in the Supplementary Method and yields the following prefactors:
+A more complete derivation of the models can be found in the Supplementary Methods and yields the following prefactors:
 \begin{align}
@@
 \end{align}
-where $1/H^3 = 1/H_1^3 + 1/H_2^3$. The crossover between the two models where $\lambda_D=\lambda_P$ is found for $H = H^* \equiv 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}$.
+where $H$ is redefined by $1/H^3 = 1/H_1^3 + 1/H_2^3$. The crossover between the two models where $\lambda_D=\lambda_P$ is found for $H = H^* \equiv 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}$.
 
@@
 
-For a typical gel sheet (Supplementary Table~\ref{tab:data}b) the buoyancy after syn\ae{}resis is estimated as $\Delta\rho=\SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+For a typical gel sheet, Supplementary Table~\ref{tab:data}b, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
 This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ and therefore gravity is dismissed.
@@
 Darcy and Poiseuille models yield wavelengths that are of the same order in most of the cases under study.
-Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them (see Supplementary Method on such measurements).
-As summarised in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacted the gel stiffness, permeability and $\tau$.
+Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them, see Supplementary Methods about such measurements.
+As summarised in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
 Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly.
@@
 
-In conclusion we have taken advantage of the properties of the protein casein.
-In particular, we have fine-tuned the interplay between the attraction and the repulsion at the protein level to assemble dense glassy networks that shrink and swell as a result of the transition from a fractal gel to an arrested phase separation.
-This carefully engineered route to form permeable biogels is the key, with controlled adhesion of the gels to the boundaries, to produce original cascade of wrinkling patterns.
+In conclusion we have taken advantage of the properties of the casein protein to engineer and study the wrinkling of a porous confined biogel.
+In particular, we have fine-tuned the interplay between the attraction and the repulsion at the protein level to assemble dense glassy networks that shrink and swell as a result of continuous acidification.
+Together with carefully controlled adhesion of the gels to the boundaries, this route to form permeable biogels is the key to produce original cascade of wrinkling patterns.
 We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall.
@@
 3D data are collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
-We use either a 10x (air) lens to measure the geometry of wrinkling (see Supplementary Material) or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern (Supplementary Figures~\ref{fig:generations} and \ref{fig:patterns}).
+We use either a 10x (air) lens to measure the geometry of wrinkling, see Supplementary Material, or a 63x (oil) lens to observe the microstructure of the gel.
+Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern, Supplementary Figures~\ref{fig:generations} and \ref{fig:patterns}.
 Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Figure~\ref{fig:generations}.
@@
 For every gel composition we monitor the gelation by small amplitude oscillations (0.1\%, \SI{1}{\hertz}) and estimate the flexural modulus $B$ from the maximum value of the shear modulus $G^\prime$ as $B=2G^\prime/(1-\nu)$ taking $\nu=0.3$ for the Poisson ratio, typical value for a spongy network~\cite{Greaves2011}.
-Permeability is measured using a protocol found in \cite{VanDijk1986} (see Supplementary Method and Supplementary Figure~\ref{fig:permeability}).
+Permeability is measured using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~\ref{fig:permeability}.
 The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
@@
 
-To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre} (Supplementary Figure~\ref{fig:nonmonotonic}a and Supplementary Videos 5 and 6).
-After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$ (see Supplementary Figure~\ref{fig:nonmonotonic}b) that is fitted by the following fractal form:
+To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre}, Supplementary Figure~\ref{fig:nonmonotonic}a and Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63}.
+After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$, see Supplementary Figure~\ref{fig:nonmonotonic}b that is fitted by the following fractal form:
 \begin{equation}
@@
 The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
-Alongside the first tube (see Supplementary Figure~\ref{fig:permeability} a), we dip a second identical tube with no gel in it.
+Alongside the first tube, see Supplementary Figure~\ref{fig:permeability} a, we dip a second identical tube with no gel in it.
 A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
@@
 
-By fitting $z(t)-z_\infty$ to an exponential (see Supplementary Figure~\ref{fig:permeability} b-c), we obtain $\tau$ and thus $\alpha$.
+By fitting $z(t)-z_\infty$ to an exponential, see Supplementary Figure~\ref{fig:permeability} b-c, we obtain $\tau$ and thus $\alpha$.
 Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
@@
 We use this approximation of $\tau$ for all our samples.
-This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact (see Supplementary Figure~\ref{fig:growth}b), thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
+This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact, see Supplementary Figure~\ref{fig:growth}b, thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
 
@@
 considers a porous elastic film between two viscous layers of arbitrary thickness.
-In this case (see Supplementary Figure~\ref{fig:notations}), the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
+In this case, see Supplementary Figure~\ref{fig:notations}, the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
 \begin{equation}
@@
 \cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_{min}$} & {$H_{max}$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
+\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_1$} & {$H_2$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
 &{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
@@
 `cas', `GDL' and `gly' indicate the weight fraction of sodium caseinate, GDL and glycerol respectively.
-$\xi$ is the pore size at the end of each experiment.}
+$\xi$ is the pore size at the end of each experiment.
+$h$ is the thickness of the gel film just before buckling.}
 \label{tab:data}
@@
 To measure wrinkle wavelength, we use large field pictures of the final state.
-These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification (Supplementary Figure~\ref{fig:patterns} except c and e), or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x (Supplementary Figure~\ref{fig:patterns}c and e).
+These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e.
 The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
@@
 Further generations do not seem to appear simultaneously and the peak is weak and broad.
-This could be a signature of a transition to chaos of a non-linear oscillator~\cite{Brau2011}.
 
@@
 
-\begin{description}
-\item[Video 1] Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
+\begin{enumerate}
+\renewcommand{\labelenumi}{\textbf{Video \arabic{enumi}:}}
+\itemindent 2em
+\item\label{vid:transmitted} Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
@@
 
-\item[Video 2] 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion (Supplementary Table~\ref{tab:data}c).
+\item\label{vid:reconstructed} 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion, Supplementary Table~\ref{tab:data}c.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item[Video 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the Video is \SI{13}{\minute}. The scale bar is XXX $\mu$m
+%\item[Video 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the Video is \SI{13}{\minute}. The scale bar is XXX $\mu$m
 
-\item[Video 4] Titration of a dispersion with 4\%w casein by 1 molar HCl. The duration of the Video is \SI{14}{\minute}. The scale bar is XXX $\mu$m
+\item\label{vid:titration} Titration of a dispersion with 4\%w casein by 1 molar HCl. The duration of the Video is \SI{14}{\minute}. The scale bar is XXX $\mu$m
 
-\item[Video 5] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
+\item\label{vid:stick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item[Video 6] Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
+\item\label{vid:nostick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item[Video 7] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}a).
+\item\label{vid:cutDarcy} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}a.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item[Video 8] Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes (Supplementary Table~\ref{tab:data}c).
+\item\label{vid:cutPoiseuille} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}c.
 10\% of the proteins are fluorescently labelled.
@@
 
-\end{description}
+\end{enumerate}
 

### 2015-02-18 fig_plis/article_plis.tex
Petites corrections du texte principal par Seb

@@
 
-To form wrinkled patterns such as the one observed in figure~\ref{fig:dynamics}, ones needs to take advantage of the pH sensitivity of caseins, to control the route to gelation, to fine tune the gel properties and the boundary conditions.
+To form wrinkled patterns such as the one observed in Figure~\ref{fig:dynamics}, ones needs (i) to take advantage of the pH sensitivity of caseins, (ii) to control the route to gelation, (iii) to fine tune the gel properties and the boundary conditions.
 
-The stability of caseins dispersion strongly depend on the pH of the solution.
+The stability of casein dispersions strongly depends on the pH.
 Supplementary Video~\ref{vid:titration} shows the titration of a casein dispersion at 4\%w by 1 molar HCl.
@@
 Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b).
-Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation shows a better agreement for the whole $H/H^*$ range.
+Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation provides a much better agreement for the whole $H/H^*$ range.
 
@@
 In particular, we have fine-tuned the interplay between the attraction and the repulsion at the protein level to assemble dense glassy networks that shrink and swell as a result of continuous acidification.
-Together with carefully controlled adhesion of the gels to the boundaries, this route to form permeable biogels is the key to produce original cascade of wrinkling patterns.
-We developed a model that captures the dynamical origin of the constrains exerted on the gel and predict the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall.
-Such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
+Together with carefully controlled adhesion of the gels to the boundaries, this route to form permeable biogels is the key to produce an original cascade of wrinkling patterns.
+We developed a model that captures the dynamical origin of the constraints exerted on the gel and predicts the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall.
+Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins.
 As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
@@
 Sodium caseinate (Firmenich) is dissolved in deionised water by stirring for \SI{20}{\minute} at room temperature.
-The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) marks the beginning of the experiments.
-We typically have about \SI{10}{\minute} to pipette the dispersion in the slit before gelation starts.
+The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) sets the beginning of the experiments.
+We typically have about \SI{10}{\minute} to pipet the dispersion into the slit before gelation starts.
 Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific).
@@
 
-Home made optical cells are composed of a slide (RS France) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm).
+Home made optical cells are composed of a slide (RS) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm).
 The optical cell is briefly heated on a hot plate so that the parafilm welds to the slide and the cover slip.
 Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell from \SIrange{50}{150}{\micro\metre}.
-The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation could take place.
+The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation can take place.
 To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which a polyacrylamide brush is polymerised from methacrylate groups (Euromedex).
 
-3D data are collected on a Zeiss LSM510 confocal microscope, using \SI{532}{\nano\meter} laser excitation.
+3D data are collected on a Zeiss LSM510 confocal microscope using \SI{532}{\nano\meter} laser excitation.
 We use either a 10x (air) lens to measure the geometry of wrinkling, see Supplementary Material, or a 63x (oil) lens to observe the microstructure of the gel.
 Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern, Supplementary Figures~\ref{fig:generations} and \ref{fig:patterns}.
-Secondary pattern wavelength is set to half of the primary wavelength, according to Supplementary Figure~\ref{fig:generations}.
+The secondary pattern wavelength is taken as half of the primary wavelength, according to Supplementary Figure~\ref{fig:generations}.
 Data shown in Figure~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
-For every gel composition we monitor the gelation by small amplitude oscillations (0.1\%, \SI{1}{\hertz}) and estimate the flexural modulus $B$ from the maximum value of the shear modulus $G^\prime$ as $B=2G^\prime/(1-\nu)$ taking $\nu=0.3$ for the Poisson ratio, typical value for a spongy network~\cite{Greaves2011}.
+For every gel composition we monitor the gelation by small amplitude oscillations (strain 0.1\%, frequency \SI{1}{\hertz}) and estimate the flexural modulus $B$ from the maximum value of the shear modulus $G^\prime$ as $B=2G^\prime/(1-\nu)$ taking $\nu=0.3$ for the Poisson ratio, a value typical of a spongy network~\cite{Greaves2011}.
 Permeability is measured using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~\ref{fig:permeability}.

### 2015-02-18 fig_plis/article_plis.tex
Petites corrections de la biblio et des SI par Seb

@@
 ML thanks the Region Rhône Alpes and the Programme d'Avenir Lyon - Saint Etienne (PALSE NoGELPo) for postoctoral grant. 
-TG and MN acknowledge funding from the agence national de la recherche fran\c{c}aise (ANR-11-PDOC-027). 
-SM and ML acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No. 258803.
+TG and MN acknowledge funding from the Agence Nationale de la Recherche (ANR-11-PDOC-027). 
+SM and ML acknowledge funding from the European Research Council under the European Union's Seventh Framework Program (FP7/2007-2013) / ERC grant agreement No.
+258803 and from Institut Universitaire de France.
 
@@
 
-\subsection*{Fraction of free casein as a function of the pH}
+\subsection*{Fraction of free casein as a function of pH}
 
-The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by HCl 1 molar or during the action of 4\%w GDL.
-In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of the pH, and centrifuged (centrifuge 5424, Eppendorf) in Eppendorf tubes at 14500 rpm for \SI{3}{\minute}.
+The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by 1 molar HCl, see Supplementary Video~\ref{vid:titration}, or during the action of 4\%w GDL.
+In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of pH, and centrifuged (centrifuge 5424, Eppendorf) in Eppendorf tubes at 14500 rpm for \SI{3}{\minute}.
 The supernatant is diluted and analysed with a UV absorption spectrometer (USB4000, Ocean Optics).
-According to~\cite{Roefs1986}, we deduced the fraction of free casein proteins from the adsorption spectrum:
+According to~\cite{Roefs1986}, we deduce the fraction of free casein proteins from the adsorption spectrum as:
 %
 \begin{align}
-x_{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
+x_\text{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
 \text{with }A(\text{pH}) &= A_{280}-1.706 A_{320},
@@
 %
-where $A_{280}$ is the absorption peak due to aromatic amino acids in the casein molecules plus a turbidity background and $A_{320}$ is only due to turbidity.
+where $A_{280}$ is the absorption peak at \SI{280}{\nano\metre} due to aromatic amino acids in the casein molecules plus a turbidity background whereas the absorption peak at \SI{320}{\nano\metre} is only due to turbidity.
 
 Above the isoelectric point, the homogeneous acidification by GDL continuously creates very small protein aggregates that are difficult to centrifuge.
-Such suspensions are too turbid and too rapidly evolving to reliably measure $x_{free}$.
+Such suspensions are too turbid and too rapidly evolving to reliably measure $x_\text{free}$.
 
@@
 	\includegraphics{nonmonotonic}%
-	\caption{Microstructure evolution.
+	\caption{Microstructure evolution of a 4\%w sodium caseinate solution acidified with 4\%w GDL in water.
 (a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row).
@@
 (b) Corresponding Fourier spectra when no adhesion on top, same times as (a).
-Lines are fits by equation~(\ref{eq:fractalS}).
+Lines are fits to equation~(\ref{eq:fractalS}).
 The inset shows a schematic association between real space structures and wave numbers.
-(c-e) Evolution of cut-off length, susceptibility and fractal dimension in full adhesive (black line) and no adhesion on top (orange dashed line) situations.
-Symbols indicates the times of the images in (a).}%
+(c-e) Evolution of cut-off length $\xi$, susceptibility $\chi$ and fractal dimension $d$ in full adhesive (black line) and no adhesion on top (orange dashed line) situations.
+Symbols indicate the times of the images in (a).}%
 	\label{fig:nonmonotonic}%
@@
 To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre}, Supplementary Figure~\ref{fig:nonmonotonic}a and Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63}.
-After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$, see Supplementary Figure~\ref{fig:nonmonotonic}b that is fitted by the following fractal form:
+After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$, see Supplementary Figure~\ref{fig:nonmonotonic}b that is fitted by the following fractal form~\cite{VanDijk1986}:
 \begin{equation}
@@
 \end{equation}
-where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension.
-$\Gamma$ is the gamma function.
+where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut-off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension.
+$\Gamma$ is the Gamma function.
 Equation~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
@@
 
-Figure~\ref{fig:acidification}c of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c and d show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
-The later situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
+Figure~\ref{fig:acidification}d of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c-e show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
+The latter situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
 To do so, we coat only the microscope slide with acrylamide brush.
 The cover slip is cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
-Neither buckling nor wrinkling is detected in such configuration.
+Neither buckling nor wrinkling is detected in such a configuration.
 
-In situation (i), the gel keeps its topology throughout the gelation and over acidification.
+In situation (i), the gel keeps its topology throughout the gelation and over-acidification.
 Left of Supplementary Figure~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again.
@@
 By contrast in situation (ii), the gel as a whole compresses while forming.
-Since our focal plane is a fixed depth, we cannot follow the topology of the same subset of the network, however the pore size is clearly smaller than in situation (ii).
+Since our focal plane is at a fixed depth, we cannot follow the topology of the same subset of the network, however the pore size is clearly smaller than in situation (ii).
 The isoelectric gel, at minimum gel volume, is probably isotropic or only marginally anisotropic in $z$.
@@
 	\caption{Permeability measurements.
-(a) Schematic representation of the experiment.
+(a) Schematic representation of the experimental set up.
 (b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
@@
 By fitting $z(t)-z_\infty$ to an exponential, see Supplementary Figure~\ref{fig:permeability} b-c, we obtain $\tau$ and thus $\alpha$.
-Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due both to high viscosity and low permeability.
+Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due to both high viscosity and low permeability.
 Obtaining $z_\infty$ without tube 2 would be time prohibitive.
@@
 
-From these data, we obtain for each stack the volume and thickness of the gel phase, area of the mid surface and peak-to-peak amplitude.
+From these data, we obtain for each stack the volume and thickness of the gel phase, area and peak-to-peak amplitude of the mid surface.
 Velocity is obtained by time differentiation.
@@
 (a) Evolution of the gel velocity $v$ averaged over the whole surface.
-Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scale.
+Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scales.
 The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
@@
 
-We decomposed the deflection in sinusoidal modes $w(x,t) =  A \sin qx$ of wavevector $q$. Equation~(\ref{eq:mecheq}) then reads
+We decompose the deflection in sinusoidal modes $w(x,t) =  A \sin (qx)$ of wavevector $q$. Equation~(\ref{eq:mecheq}) then reads
 \begin{equation}
@@
 The constant thickness gel film (orange) is surrounded by solvent.
-The region of interest of the mass conservation is highlighted in gray.
-Arrows are oriented as x or z to allow algebraic notations.}%
+The region of interest for mass conservation is highlighted in gray.
+Arrows are oriented along $x$ or $z$, defining positive sign.}%
 \label{fig:notations}%
@@
 Let us now turn to the case of an elastic, impermeable film on a single viscous layer.
-As \cite{Huang2002} we neglect gravity, although it is difficult to do so in practice at a free interface.
+As in \cite{Huang2002} we neglect gravity, although it would be difficult to do so in practice at a free interface.
 Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation.
@@
 \begin{equation}
-\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}
+\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}.
 \end{equation}
-
+%
 We note that the mode minimising $\sigma_\parallel$ is also maximising the growth rate
@@
 \end{equation}
-This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface gravity can indeed be neglected.
+This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface, gravity can indeed be neglected.
 
@@
 \end{equation}
-
+%
 The dominant wavelength for Darcy model is thus
@@
 \end{equation}
-
+%
 $v$ can be expressed by the Darcy law
@@
 \end{equation}
-
+%
 Note that we recover equation~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and equation~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
@@
 \subsection*{Samples characteristics}
-To test the theoretical predictions we have fully characterised 9 samples, see Supplementary Table \ref{tab:data}.
+To test the theoretical predictions we have fully characterised 9 samples. Their characteristics are gathered in Supplementary Table \ref{tab:data}.
 
@@
 \end{tabular}
-\caption{Fully characterised samples at the basis of Figure~\ref{fig:DarcyPoiseuille}.
+\caption{Characteristics of the samples used for Figure~\ref{fig:DarcyPoiseuille}.
 Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.
@@
 The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
-Length of arrows correspond to the measured primary wavelength.}
+Arrows show the measured primary wavelength.}
 	\label{fig:patterns}
@@
 
-To measure wrinkle wavelength, we use large field pictures of the final state.
+To measure the wrinkle wavelength $\lambda$, we use large field pictures of the final state.
 These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e.
@@
 White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
-It is thus possible to reconstruct a binary picture of the pattern (inset of Supplementary Figure~\ref{fig:generations}).
-Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final picture to pure first generation.
+It is thus possible to reconstruct a binary picture of the pattern, inset of Supplementary Figure~\ref{fig:generations}.
+Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final binary picture to the configuration of the primary pattern.
 
@@
 When only the first generation pattern is present a peak clearly indicates the dominant mode.
-When second generation has appeared, the dominant wavelength is halved and the peak much weaker.
-This allows us to use secondary pattern is our analysis by setting their wavelength at half of the primary wavelength.
-Further generations do not seem to appear simultaneously and the peak is weak and broad.
+When the secondary pattern has appeared, the dominant wavelength is halved and the peak much weaker.
+This allows us to use the secondary pattern is our analysis by setting their wavelength to half of the primary wavelength.
+Further generations do not seem to appear simultaneously and the corresponding peaks are weak and broad.
 
@@
 \renewcommand{\labelenumi}{\textbf{Video \arabic{enumi}:}}
-\itemindent 2em
+\itemindent 8mm
 \item\label{vid:transmitted} Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The duration of the Video is \SI{2}{\hour}~30.
+The total duration of the Video is \SI{2}{\hour}~30.
 The scale bar is \SI{1}{\milli\metre}.
@@
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The duration of the Video is \SI{2}{\hour}~30.
+The total duration of the Video is \SI{2}{\hour}~30.
 The scale bar is \SI{1}{\milli\metre}.
@@
 
-\item\label{vid:titration} Titration of a dispersion with 4\%w casein by 1 molar HCl. The duration of the Video is \SI{14}{\minute}. The scale bar is XXX $\mu$m
+\item\label{vid:titration} Titration of a dispersion with 4\%w casein by 1 molar HCl. The total duration of the Video is \SI{14}{\minute}. The Becher diameter is \SI{55}{\milli\metre}.
 
@@
 10\% of the proteins are fluorescently labelled.
-The duration of the Video is \SI{132}{\minute}.
+The total duration of the Video is \SI{132}{\minute}.
 The scale bar is \SI{10}{\micro\metre}.
@@
 10\% of the proteins are fluorescently labelled.
-The duration of the Video is \SI{147}{\minute}.
+The total duration of the Video is \SI{147}{\minute}.
 The scale bar is \SI{10}{\micro\metre}.
@@
 The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
-Here the gel starts close to the bottom wall, $H<H^*$.
+Here the gel initially lies close to the bottom wall, $H<H^*$.
 Darcy mechanism dominates the formation of the primary pattern.
-Total duration of the video is \SI{145}{\minute}.
+The total duration of the video is \SI{145}{\minute}.
 
@@
 The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
-Here the gel starts far from either top or bottom wall, $H>H^*$.
+Here the gel initially lies far from either top or bottom wall, $H>H^*$.
 Poiseuille mechanism dominates the formation of the primary pattern.
-Total duration of the video is \SI{181}{\minute}.
+The total duration of the video is \SI{181}{\minute}.
 

### 2015-02-18 fig_plis/cover_letter.tex
Vache qui rit seulement, à drote du texte Moins de referee Détail l'impact

@@
 \usepackage{tabu}
+\usepackage{wrapfig}
 
@@
 Laboratoire de physique,\\
-Ecole Normale Supérieure de Lyon,\\
+Ecole Normale Supérieure de Lyon/CNRS, France\\
 \texttt{thomas.gibaud@ens-lyon.fr}
@@
 
-Please find enclosed our manuscript, entitled \emph{Hierarchical wrinkling in permeable and confined biogels}, which we are submitting for publication as an article to Nature Materials.
+Please find enclosed our manuscript, entitled \emph{Hierarchical wrinkling in a confined permeable biogel}, which we are submitting for publication as an article to Nature Materials.
 
 Supplementary Video 1 arouses one's curiosity.
-It shows microscopy images of a thin optical cell containing a protein gel whose adhesion to the cell wall is prohibited.
+It shows microscopy images of a thin optical cell containing a protein gel whose adhesion to the lateral wall is prohibited.
 No external stimulus is applied.
-How can a simple biogel form spontaneously such complicated nested patterns?
+How can a simple biogel spontaneously form such complicated nested patterns?
 
+\begin{tabu}{@{}X[m]m{0.2\textwidth}}
+%\includegraphics[height=8.5\baselineskip]{../Droste.jpg} &
 Supplementary Video 2, which shows a 3D construction of the gel based on confocal imaging, hints towards the answer.
 The gel shrinks then swells and wrinkles due to the excess area created during swelling.
-This is however not the end of the story as boundary conditions frustrate the development of the wrinkles and induce buckling within the buckling .
-This hierarchical pattern is a beautiful illustration of the well known Droste effect, known in France as ``Vache qui rit'' effect, where an image contains a smaller version of itself.
-
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=8\baselineskip]{../Droste.jpg}&
-\includegraphics[height=8\baselineskip]{../vachequirit.jpg}
+This is however not the end of the story as boundary conditions frustrate the development of the wrinkles and induce buckling within the buckling.
+This hierarchical pattern is a beautiful and natural illustration of the well known Droste effect, known in France as ``Vache qui rit'' effect, where an image contains a smaller version of itself.&
+\includegraphics[width=0.2\textwidth]{../vachequirit.jpg}
 \end{tabu}
 
+%\includegraphics[width=0.2\textwidth]{../vachequirit.jpg}
+
+%\begin{tabu}{X[c]X[c]}
+%\includegraphics[height=8\baselineskip]{../Droste.jpg}&
+%\includegraphics[height=8\baselineskip]{../vachequirit.jpg}
+%\end{tabu}
+
 It is universally recognized that confined thin surfaces may wrinkle due the growth of an excess of material.
-For example on the biology side different rates of growth between the gut tube and its dorsal anchoring is responsible for the vilification of guts.
+For example on the biology side a difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts.
 Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern.
-Such patterns can also be triggered by physical cues as temperature dilation, swelling or the removal of pre-strain.
+Such patterns can also be triggered by physical parameters such as temperature dilation, swelling or the removal of pre-strain.
 
-Yet, benchmark experiments such as the one presented in Supplementary Videos 1 and 2, that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line.
-Indeed, stability analysis of a film lying on a thin viscous substrate was studied only recently and remains theoretical because, on a free interface, gravity dominates over viscous substrate as the wavelength selection mechanism.
+So far, laboratory experiments on synthetic systems have focused on free interfaces where gravity dominates over viscous substrates as the wavelength selection mechanism.
+However, in biology relevant situations for instance, gravity can generally be neglected and cannot be invoked as the driving mechanism for wrinkling.
+Therefore, benchmark experiments, such as the ones presented in Supplementary Videos 1 and 2, that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line.
 
@@
 \item From food science we have hijacked the process of yoghurt making to form gels.
-Usually, when you acidify milk the caseins flocculate.
+Usually, when one acidifies milk, the caseins flocculate.
 To form a uniform yoghurt gel one needs to lower homogeneously the pH, as bacteria do.
@@
 Controlled gelation with peculiar boundary conditions introduces a robust method to produce wrinkled surfaces that is distinct from other emerging technologies.
-We use a combination of complementary techniques ranging from polymer brush grafting, titration, light microscopy, rheology, permeability measurements and confocal microscopy that sustains studies to better understand the relations between all levels of hierarchy from the individual proteins to the macroscopic.
+We use a combination of complementary techniques ranging from polymer brush grafting, titration, light microscopy, rheology, permeability measurements and confocal microscopy that allow us to fully understand the relations between all levels of hierarchy from the individual proteins to the macroscopic level.
 Indeed, we directly measure and relate the non monotonous pH response of the proteins to the spontaneous shrinkage and swelling effect of the casein network at the micron level to the wrinkles at the millimetre scale.
-Playing with the geometry, the gel composition and the viscosity of the solvent, we explore a wide range of physical parameters to check the validity of our models.
-All of these differences enable us to assemble, to our knowledge, the first example of Darcy flow driven wrinkling.
+Playing with the geometry, the gel composition and the viscosity of the solvent, we explore a wide range of physical parameters to check the validity of our models and evidence the first example of wrinkling driven by Darcy flow.
 
 The impact of our work extends beyond soft-matter physics.
-For instance, it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting or in optical devices such as diffraction gratings or Fresnel lenses, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues.
+For instance, it illustrates daily life problems such as wrinkles encountered while putting up wall paper: water swells the paper and air infiltrates through the paper and produces Darcy blisters.
+It could also set a reference for potential applications.
+For example since the wrinkling pattern is unique and unreproducible it could be used as a safe key for identification.
+Moreover, the nested nature of the patterns could be use in optical devices such as diffraction gratings or Fresnel lenses.
+Finally, our experiment paves the way to explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues.
 Indeed during the morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.
@@
 For all these reasons we believe that our manuscript is suitable for a publication in an interdisciplinary journal such as Nature Materials.
-With this manuscript we are including 8 Supplementary Videos, which have been mailed to your editorial office.
+With this manuscript we are including 7 Supplementary Videos, which have been mailed to your editorial office.
 As possible reviewers we suggest: L.
 Mahadevan (Harvard, USA), N.
-Menon (Amherst, USA), José Bico (ESPCI, France), Enrique Cerda (Univ.
-Santiago-Chile, Chile), Mokhtar-Adda-Bedia (ENS Paris, France), Shu Yang (U.Penn, USA), Pascal Damman (Univ.
-Mons, Belgium), Tian Yang (Univ.
-Michigan, USA), Dominic Vella (Oxford, UK) XXX.
+Menon (Amherst, USA), Benoît Roman (ESPCI, France), Enrique Cerda (Univ.
+Santiago-Chile, Chile), Mokhtar-Adda-Bedia (ENS Paris, France), Dominic Vella (Oxford, UK), Eric Lauga (Cambridge, UK).
 
-We thank you for your time and attention.
+We thank you for your consideration, time and attention.
 

### 2015-02-18 fig_plis/article_plis.tex
Retire les références à une ligne particulière de tab:data Thibaut à sa place

@@
 Successive generations of patterns are highlighted in colour to stress the absence of coarsening after formation.
-(b) 3D reconstruction from fluorescent confocal microscopy, see Supplementary Table~\ref{tab:data}c.
+(b) 3D reconstruction from fluorescent confocal microscopy.
 Scale bars are \SI{1}{\milli\metre}.}%
@@
 
-For a typical gel sheet, Supplementary Table~\ref{tab:data}b, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+For a typical gel sheet, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
 This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ and therefore gravity is dismissed.
@@
 The authors thank Jos\'e Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
-Precious experimental ideas came from Thibaut Divoux, Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Maxime Lefranc.
-Alan Parker at Firmenich provided the casein and GDL; 
+Precious experimental ideas came from Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Thibaut Divoux.
+Alan Parker at Firmenich provided the casein and GDL. 
 Special thanks to Madame Biot who made M.A. Biot work accessible through the \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 

### 2015-02-18 fig_plis/cover_letter.tex
viscous substrate -> viscosity Typos

@@
 
-So far, laboratory experiments on synthetic systems have focused on free interfaces where gravity dominates over viscous substrates as the wavelength selection mechanism.
-However, in biology relevant situations for instance, gravity can generally be neglected and cannot be invoked as the driving mechanism for wrinkling.
+So far, laboratory experiments on synthetic systems have focused on free interfaces where gravity dominates over viscosity as the wavelength selection mechanism.
+However, in biologically relevant situations for instance, gravity can generally be neglected and cannot be invoked as the driving mechanism for wrinkling.
 Therefore, benchmark experiments, such as the ones presented in Supplementary Videos 1 and 2, that explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium are in line.
@@
 
-From a practical perspective, such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different sort of proteins.
+From a practical perspective, such a model experiment, because it relies on pH induced charge stabilisation/destabilisation, should be generalisable to different kind of proteins.
 Controlled gelation with peculiar boundary conditions introduces a robust method to produce wrinkled surfaces that is distinct from other emerging technologies.
 We use a combination of complementary techniques ranging from polymer brush grafting, titration, light microscopy, rheology, permeability measurements and confocal microscopy that allow us to fully understand the relations between all levels of hierarchy from the individual proteins to the macroscopic level.
-Indeed, we directly measure and relate the non monotonous pH response of the proteins to the spontaneous shrinkage and swelling effect of the casein network at the micron level to the wrinkles at the millimetre scale.
+Indeed, we directly measure and relate the non-monotonic pH response of the proteins to the spontaneous shrinkage and swelling effect of the casein network at the micron level to the wrinkles at the millimetre scale.
 Playing with the geometry, the gel composition and the viscosity of the solvent, we explore a wide range of physical parameters to check the validity of our models and evidence the first example of wrinkling driven by Darcy flow.
@@
 For instance, it illustrates daily life problems such as wrinkles encountered while putting up wall paper: water swells the paper and air infiltrates through the paper and produces Darcy blisters.
-It could also set a reference for potential applications.
+Our results could also set a reference for potential applications.
 For example since the wrinkling pattern is unique and unreproducible it could be used as a safe key for identification.
-Moreover, the nested nature of the patterns could be use in optical devices such as diffraction gratings or Fresnel lenses.
-Finally, our experiment paves the way to explore the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues.
-Indeed during the morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.
+Moreover, the nested nature of the patterns could be used in optical devices such as diffraction gratings or Fresnel lenses.
+Finally, our experiments pave the way for exploring the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues.
+Indeed during morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.
 
@@
 With this manuscript we are including 7 Supplementary Videos, which have been mailed to your editorial office.
-As possible reviewers we suggest: L.
+As potential reviewers we wish to suggest: L.
 Mahadevan (Harvard, USA), N.

### 2015-02-18 fig_plis/article_plis.tex
Sépare le texte principal et les supplementary materials

@@
 
-\newcommand{\seb}[1]{\textbf{\color{blue}#1}} % remarques Seb
-\newcommand{\sseb}[1]{\sout{\textbf{\color{blue}#1}}} % remarques Seb
+\usepackage{xr}
+\externaldocument[S-]{supplementary_plis}
 
@@
 In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002} --this is a well-known effect used in yoghurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off, Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{vid:transmitted} and \ref{vid:reconstructed}.
+As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002} --this is a well-known effect used in yoghurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off, Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{S-vid:transmitted} and \ref{S-vid:reconstructed}.
 Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
@@
 The stability of casein dispersions strongly depends on the pH.
-Supplementary Video~\ref{vid:titration} shows the titration of a casein dispersion at 4\%w by 1 molar HCl.
+Supplementary Video~\ref{S-vid:titration} shows the titration of a casein dispersion at 4\%w by 1 molar HCl.
 At pH=7, the caseins are well dispersed then they flocculate between pH=5.5 to 4 and finally at pH=3 the solution becomes clearer as caseins partially redisperse.
@@
 To gain further insight into the mechanical properties of the gel we turn to microstructural observations under confocal microscopy.
-As shown in Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63} and in Supplementary Figure~\ref{fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut-off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
-Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~\ref{fig:permeability}.
-The fractal model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
-As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
+As shown in Supplementary Videos~\ref{S-vid:stick63} and \ref{S-vid:nostick63} and in Supplementary Figure~\ref{S-fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut-off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
+Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~\ref{S-fig:permeability}.
+The fractal model fits the structure at all times but its parameters vary non-monotonically during GDL acidification.
+As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{S-fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
 This corresponds to a tighter packing of the caseins within the network strands.
@@
 To quantify this effect, we make the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}.
-In the Supplementary Figure~\ref{fig:permeability}, we observe that the aqueous buffer goes through the gel and rises in the tube due to capillary forces.
+In the Supplementary Figure~\ref{S-fig:permeability}, we observe that the aqueous buffer goes through the gel and rises in the tube due to capillary forces.
 Exploitation of this experiment give the permeability of the gel: \SI{67000}{\square\nano\metre}.
@@
 Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
-Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplementary Fig.~\ref{fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
-In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{fig:nonmonotonic}).
+Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplementary Fig.~\ref{S-fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
+In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{S-fig:nonmonotonic}).
 
@@
 
-Figure~\ref{fig:sideview}a shows successive side views of confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration that induces wrinkling, see also Supplementary Videos~\ref{vid:cutDarcy} and \ref{vid:cutPoiseuille}.
+Figure~\ref{fig:sideview}a shows successive side views of confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration that induces wrinkling, see also Supplementary Videos~\ref{S-vid:cutDarcy} and \ref{S-vid:cutPoiseuille}.
 Initially the dispersion is homogeneous and fills the entire slit.
@@
 Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}c, and excess surface area of the gel film, Figure~\ref{fig:sideview}d, see Supplementary Methods.
-As expected from the system design, the evolution of the gel volume is non-monotonous.
+As expected from the system design, the evolution of the gel volume is non-monotonic.
 After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
@@
 Around \SI{38}{\minute}, the amplitude of the wrinkles reaches the thickness of the slit.
-The size of the primary pattern is now fixed, Supplementary Figure~\ref{fig:growth}.
+The size of the primary pattern is now fixed, Supplementary Figure~\ref{S-fig:growth}.
 However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
@@
 In a first limit, the film of gel sits without sticking to the bottom wall of the cell, i.e.
-$H_1\rightarrow0$, see Supplementary Video~\ref{vid:cutDarcy}.
+$H_1\rightarrow0$, see Supplementary Video~\ref{S-vid:cutDarcy}.
 Creating a two dimensional blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
@@
 This expression emphasises the kinetic nature of the wavelength selection through Darcy flow.
-However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}.
+However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{S-fig:growth}.
 This arrest may be due to bending rigidity~\cite{LeGoff2014}.
@@
 
-The opposite limiting case is when the gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Supplementary Video~\ref{vid:cutPoiseuille}) and impermeable ($\alpha\rightarrow 0$).
+The opposite limiting case is when the gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Supplementary Video~\ref{S-vid:cutPoiseuille}) and impermeable ($\alpha\rightarrow 0$).
 Here destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
@@
 Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them, see Supplementary Methods about such measurements.
-As summarised in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
+As summarised in Supplementary Table~\ref{S-tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
 Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly.
@@
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
-Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Figure~\ref{fig:generations}.
+Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Figure~\ref{S-fig:generations}.
 
@@
 We use either a 10x (air) lens to measure the geometry of wrinkling, see Supplementary Material, or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern, Supplementary Figures~\ref{fig:generations} and \ref{fig:patterns}.
-The secondary pattern wavelength is taken as half of the primary wavelength, according to Supplementary Figure~\ref{fig:generations}.
+Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern, Supplementary Figures~\ref{S-fig:generations} and \ref{S-fig:patterns}.
+The secondary pattern wavelength is taken as half of the primary wavelength, according to Supplementary Figure~\ref{S-fig:generations}.
 Data shown in Figure~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
 For every gel composition we monitor the gelation by small amplitude oscillations (strain 0.1\%, frequency \SI{1}{\hertz}) and estimate the flexural modulus $B$ from the maximum value of the shear modulus $G^\prime$ as $B=2G^\prime/(1-\nu)$ taking $\nu=0.3$ for the Poisson ratio, a value typical of a spongy network~\cite{Greaves2011}.
-Permeability is measured using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~\ref{fig:permeability}.
+Permeability is measured using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~\ref{S-fig:permeability}.
 The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
@@
 
-\clearpage
-\newpage
-\setcounter{figure}{0}
-
-\section*{Supplementary material}
-
-\subsection*{Fraction of free casein as a function of pH}
-
-The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by 1 molar HCl, see Supplementary Video~\ref{vid:titration}, or during the action of 4\%w GDL.
-In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of pH, and centrifuged (centrifuge 5424, Eppendorf) in Eppendorf tubes at 14500 rpm for \SI{3}{\minute}.
-The supernatant is diluted and analysed with a UV absorption spectrometer (USB4000, Ocean Optics).
-According to~\cite{Roefs1986}, we deduce the fraction of free casein proteins from the adsorption spectrum as:
-%
-\begin{align}
-x_\text{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
-\text{with }A(\text{pH}) &= A_{280}-1.706 A_{320},
-\end{align}
-%
-where $A_{280}$ is the absorption peak at \SI{280}{\nano\metre} due to aromatic amino acids in the casein molecules plus a turbidity background whereas the absorption peak at \SI{320}{\nano\metre} is only due to turbidity.
-
-Above the isoelectric point, the homogeneous acidification by GDL continuously creates very small protein aggregates that are difficult to centrifuge.
-Such suspensions are too turbid and too rapidly evolving to reliably measure $x_\text{free}$.
-
-\subsection*{Microstructure measurement}
-
-\begin{figure}
-	\includegraphics{nonmonotonic}%
-	\caption{Microstructure evolution of a 4\%w sodium caseinate solution acidified with 4\%w GDL in water.
-(a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row).
-The scale bar is \SI{10}{\micro\metre}.
-Arrows indicates the pore size.
-(b) Corresponding Fourier spectra when no adhesion on top, same times as (a).
-Lines are fits to equation~(\ref{eq:fractalS}).
-The inset shows a schematic association between real space structures and wave numbers.
-(c-e) Evolution of cut-off length $\xi$, susceptibility $\chi$ and fractal dimension $d$ in full adhesive (black line) and no adhesion on top (orange dashed line) situations.
-Symbols indicate the times of the images in (a).}%
-	\label{fig:nonmonotonic}%
-\end{figure}
-
-To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre}, Supplementary Figure~\ref{fig:nonmonotonic}a and Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63}.
-After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$, see Supplementary Figure~\ref{fig:nonmonotonic}b that is fitted by the following fractal form~\cite{VanDijk1986}:
-\begin{equation}
-I(q) = \frac{2\chi\Gamma(d)}{\left(1+\left((d+1)\xi q\right)^2\right)^{d/2}}
-\label{eq:fractalS}
-\end{equation}
-where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut-off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension.
-$\Gamma$ is the Gamma function.
-Equation~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
-%En 3D on peut citer \cite{Martin1987,Chen1986}
-
-Figure~\ref{fig:acidification}d of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c-e show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
-The latter situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
-To do so, we coat only the microscope slide with acrylamide brush.
-The cover slip is cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens.
-Neither buckling nor wrinkling is detected in such a configuration.
-
-In situation (i), the gel keeps its topology throughout the gelation and over-acidification.
-Left of Supplementary Figure~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again.
-Syn\ae{}resis and swelling are thus local processes.
-By contrast in situation (ii), the gel as a whole compresses while forming.
-Since our focal plane is at a fixed depth, we cannot follow the topology of the same subset of the network, however the pore size is clearly smaller than in situation (ii).
-The isoelectric gel, at minimum gel volume, is probably isotropic or only marginally anisotropic in $z$.
-This explains that subsequent swelling is not occurring only along $z$ but also in plane, causing buckling.
-
-$\xi$ is also systematically measured in the same way at the end of each wrinkling experiment and reported in Supplementary Table~\ref{tab:data}.
-%Final fractal dimension was always close to 2 within error bars.
-Comparing $\chi$ across different setting of the microscope is not relevant.
-
-\begin{figure*}%
-	\includegraphics{permeability}%
-	\caption{Permeability measurements.
-(a) Schematic representation of the experimental set up.
-(b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2.
-The black line is the best exponential fit $Ae^{-t/\tau}$.
-(b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$.
-(c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}%
-	\label{fig:permeability}%
-\end{figure*}
-\subsection*{Permeability measurements}
-
-To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}.
-First, the gel of interest is formed at the lower extremity of a thin glass tube.
-The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition).
-Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4.
-The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment.
-Alongside the first tube, see Supplementary Figure~\ref{fig:permeability} a, we dip a second identical tube with no gel in it.
-A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube.
-Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.
-
-Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
-\begin{equation}
-\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
-\end{equation}
-where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel.
-We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer.
-We can thus write
-\begin{equation}
-\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
-\end{equation}
-
-By fitting $z(t)-z_\infty$ to an exponential, see Supplementary Figure~\ref{fig:permeability} b-c, we obtain $\tau$ and thus $\alpha$.
-Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due to both high viscosity and low permeability.
-Obtaining $z_\infty$ without tube 2 would be time prohibitive.
-
-\subsection*{Confocal images processing}
-
-At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
-Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
-At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.
-
-We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
-Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
-We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line.
-Since this criterion is local, we do not need to correct for background intensity variation.
-In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.
-
-From these data, we obtain for each stack the volume and thickness of the gel phase, area and peak-to-peak amplitude of the mid surface.
-Velocity is obtained by time differentiation.
-Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
-
-\subsection*{Characteristic time measurement}
-\begin{figure}
-	\includegraphics{growth}
-	\caption{Growth of the instability.
-(a) Evolution of the gel velocity $v$ averaged over the whole surface.
-Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scales.
-The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
-(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
-Note that the longitudinal size of the blister evolves in this regime.
-The thick black curve is the profile at much longer times and is at lower altitude due to swelling.
-Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
-	\label{fig:growth}
-\end{figure}
-
-In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
-However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
-Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
-
-We note that when both measurements were available (as in Supplementary Figure~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
-We use this approximation of $\tau$ for all our samples.
-This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact, see Supplementary Figure~\ref{fig:growth}b, thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
-
-\subsection*{Mixed Darcy--Poiseuille model}
-
-\subsubsection*{General framework for wrinkling}
-
-The mechanical equilibrium of a plate of thickness $h$, Young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a transverse load $\sigma_\perp$ along $z$ writes
-\begin{equation}
-\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_\parallel h \frac{\partial^2 w}{\partial x^2} = \sigma_\perp,
-\label{eq:mecheq}
-\end{equation}
-where $w(x,t)$ is the deflection of the plate along $z$~\cite{Biot1957}. Since we deal with slow rates of deformation, inertia is neglected.
-
-We decompose the deflection in sinusoidal modes $w(x,t) =  A \sin (qx)$ of wavevector $q$. Equation~(\ref{eq:mecheq}) then reads
-\begin{equation}
-\left(\frac{1}{12} B h^3 q^4 - \sigma_\parallel h q^2\right)w = \sigma_\perp,
-\label{eq:mecheqop}
-\end{equation}
-where $B=E/(1-\nu^2)$ is the modulus of the plate.
-
-Following \cite{Biot1957}, we consider the case where the compression load $\sigma_\parallel$ is increased quasistatically.
-The observed mode is the one that arises at the lowest possible $\sigma_\parallel$, i.e.
-the mode minimising $\sigma_\parallel$.
-
-\subsubsection*{Elastic film in an infinite elastic medium}
-
-In this framework, we recall here the case of an elastic film lying over an infinite elastic medium~\cite{Gough1940,Bijlaard1946,Biot1957,Cerda2003}.
-The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -q B_s w$, and thus the compression load writes
-\begin{equation}
-\sigma_\parallel = B \frac{h^2}{12} q^2 + \frac{B_s}{hq}.
-\label{eq:sigma0}
-\end{equation}
-
-Minimising $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
-\begin{equation}
-\lambda = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
-\label{eq:lambdaElEl}
-\end{equation}
-
-\subsubsection*{Elastic film on a viscous layer}
-
-\begin{figure}
-\includegraphics{notations}%
-\caption{Schematic side view of the slit.
-The constant thickness gel film (orange) is surrounded by solvent.
-The region of interest for mass conservation is highlighted in gray.
-Arrows are oriented along $x$ or $z$, defining positive sign.}%
-\label{fig:notations}%
-\end{figure}
-
-Let us now turn to the case of an elastic, impermeable film on a single viscous layer.
-As in \cite{Huang2002} we neglect gravity, although it would be difficult to do so in practice at a free interface.
-Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation.
-Further considering only small deflections, we may neglect the displacement of the plate along $x$.
-The flux along $x$ is thus of the Poiseuille form
-\begin{equation}
-Q = -\frac{H^3}{12\eta}\frac{\partial p}{\partial x}.
-\label{eq:PoiseuilleFlux}
-\end{equation}
-where $p(x)$ is the pressure in the layer of viscosity $\eta$.
-
-Mass conservation (on the lower gray area of Supplementary Figure~\ref{fig:notations}) writes
-\begin{equation}
-\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} = 0.
-\label{eq:conservation}
-\end{equation}
-
-Assuming an exponentially growing amplitude $A(t) = A_0 \exp(t/\tau)$ and keeping only linear order terms, one gets
-\begin{equation}
-\frac{12\eta}{\tau} \frac{w}{H^3q^2} = -\sigma_\perp.
-\label{eq:PoiseuilleLoad}
-\end{equation}
-
-Combining Equations~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
-\begin{equation}
-\sigma_\parallel = \frac{1}{12}B h^2 q^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h q^4}
-\label{eq:sigma0P}
-\end{equation}
-and is minimised at $q_P$ such that
-\begin{equation}
-q_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
-\label{eq:qP}
-\end{equation}
-with $\Upsilon = B\tau/\eta$ the viscoelastic factor. The dominant wavelength for the Poiseuille model is thus
-\begin{equation}
-\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}.
-\end{equation}
-%
-We note that the mode minimising $\sigma_\parallel$ is also maximising the growth rate
-\begin{equation}
-\frac{1}{\tau} = \frac{H^3q^2}{12^2\eta}\left(12\sigma_\parallel - Bh^3q^4\right).
-\end{equation}
-
-When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of equation~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness
-\begin{equation}
-\frac{1}{H^3} = \frac{1}{H_1^3}+\frac{1}{H_2^3}.
-\end{equation}
-This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface, gravity can indeed be neglected.
-
-\subsubsection*{Porous elastic film with no possible Poiseuille flow}
-Another limiting case is that of a porous elastic film immersed in a viscous medium where longitudinal resistance to flow is infinite.
-It is the case in our experiments when the film is at contact with one of the walls.
-The only way to deform the film is to flow some liquid through the film of permeability $\alpha$.
-The transverse load derives from the Darcy law:
-\begin{equation}
-\sigma_\perp = p_1-p_2 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
-\label{eq:DarcyLoad}
-\end{equation}
-
-Combining equations~(\ref{eq:mecheqop}) and (\ref{eq:DarcyLoad}), the compression load writes
-\begin{equation}
-\sigma_\parallel = \frac{1}{12}B h^2 q^2 + \frac{\eta}{\tau}\frac{1}{\alpha q^2}
-\label{eq:sigma0D}
-\end{equation}
-and is minimised at $q_D$ such that
-\begin{equation}
-q_D^4 = \frac{12}{h^2\alpha\Upsilon}.
-\label{eq:qD}
-\end{equation}
-%
-The dominant wavelength for Darcy model is thus
-\begin{equation}
-\lambda_D = 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}.
-\end{equation}
-Here also the dominant mode maximises the growth rate.
-
-\subsubsection*{Porous elastic film between two viscous layers}
-Finally the most realistic scenario mixes both Poiseuille and Darcy mechanisms, i.e.
-considers a porous elastic film between two viscous layers of arbitrary thickness.
-In this case, see Supplementary Figure~\ref{fig:notations}, the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
-\begin{equation}
-H_1^3 p_1 + H_2^3 p_2 = 0.
-\label{eq:pressures}
-\end{equation}
-
-The mass conservation in the lower viscous layer is the same as equation~(\ref{eq:conservation}) with an added leak $v$ due to the porosity
-\begin{equation}
-\frac{\partial Q_1}{\partial x} + \frac{\partial w}{\partial t} + v = 0.
-\label{eq:conservationDarcy}
-\end{equation}
-%
-$v$ can be expressed by the Darcy law
-\begin{equation}
-v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}.
-\label{eq:Darcy}
-\end{equation}
-
-Combining equations~(\ref{eq:pressures}), (\ref{eq:conservationDarcy}) and (\ref{eq:Darcy}) we obtain the transverse load
-\begin{equation}
-\sigma_\perp = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 q^2}{12} + \frac{\alpha}{h}},
-\label{eq:MixtLoad}
-\end{equation}
-%
-Note that we recover equation~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and equation~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$.
-We can recast this intuition in terms of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $q_P$ and $q_D$:
-\begin{equation}
-H^* = 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}.
-\end{equation}
-
-Analytic minimisation of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome.
-Instead we rewrite equation~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
-\begin{align}
-\frac{12\sigma_\parallel}{q_D^2 h^2 B} &= \left(\frac{q}{q_D}\right)^2 + \frac{1}{2\frac{q_D^2}{q_P^6}q^4  + \left(\frac{q}{q_D}\right)^2}\\
-&= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}
-\end{align}
-and we minimise this expression numerically.
-
-\subsection*{Samples characteristics}
-To test the theoretical predictions we have fully characterised 9 samples. Their characteristics are gathered in Supplementary Table \ref{tab:data}.
-
-\begin{table*}
-\begin{tabular}{cSSSScSrrScSSSScSSSSSS}
-%\toprule
-&\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_1$} & {$H_2$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\
-&{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-%ech2
-a & 4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
-& & & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
-b & 4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
-c & 4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
-d & 4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
-& & & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-e & 8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
-& & & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
-f & 8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
-& & & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
-\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
-g & 4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
-h & 4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
-i & 4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
-\end{tabular}
-\caption{Characteristics of the samples used for Figure~\ref{fig:DarcyPoiseuille}.
-Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.
-`cas', `GDL' and `gly' indicate the weight fraction of sodium caseinate, GDL and glycerol respectively.
-$\xi$ is the pore size at the end of each experiment.
-$h$ is the thickness of the gel film just before buckling.}
-\label{tab:data}
-\end{table*}
-
-\subsection*{Wavelength measurement}
-
-\begin{figure}
-	\includegraphics{generations}%
-	\caption{Fourier spectra of binary altitudes of Main Figure~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-The scale bar is \SI{1}{\milli\metre}.}%
-	\label{fig:generations}%
-\end{figure}
-
-\begin{figure*}
-	\includegraphics{patterns}
-	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data}.
-(a-d) increasing cell thickness.
-(e-f) larger solid content.
-(g-i) increasing glycerol content.
-All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
-The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
-Arrows show the measured primary wavelength.}
-	\label{fig:patterns}
-\end{figure*}
-
-To measure the wrinkle wavelength $\lambda$, we use large field pictures of the final state.
-These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e.
-The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
-White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
-It is thus possible to reconstruct a binary picture of the pattern, inset of Supplementary Figure~\ref{fig:generations}.
-Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final binary picture to the configuration of the primary pattern.
-
-Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
-When only the first generation pattern is present a peak clearly indicates the dominant mode.
-When the secondary pattern has appeared, the dominant wavelength is halved and the peak much weaker.
-This allows us to use the secondary pattern is our analysis by setting their wavelength to half of the primary wavelength.
-Further generations do not seem to appear simultaneously and the corresponding peaks are weak and broad.
-
-\subsection*{Supplementary Videos}
-
-\begin{enumerate}
-\renewcommand{\labelenumi}{\textbf{Video \arabic{enumi}:}}
-\itemindent 8mm
-\item\label{vid:transmitted} Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
-Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The total duration of the Video is \SI{2}{\hour}~30.
-The scale bar is \SI{1}{\milli\metre}.
-
-\item\label{vid:reconstructed} 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion, Supplementary Table~\ref{tab:data}c.
-10\% of the proteins are fluorescently labelled.
-Top and bottom slides of the optical cell are coated with acrylamide brushes.
-The total duration of the Video is \SI{2}{\hour}~30.
-The scale bar is \SI{1}{\milli\metre}.
-
-%\item[Video 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the Video is \SI{13}{\minute}. The scale bar is XXX $\mu$m
-
-\item\label{vid:titration} Titration of a dispersion with 4\%w casein by 1 molar HCl. The total duration of the Video is \SI{14}{\minute}. The Becher diameter is \SI{55}{\milli\metre}.
-
-\item\label{vid:stick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
-10\% of the proteins are fluorescently labelled.
-The total duration of the Video is \SI{132}{\minute}.
-The scale bar is \SI{10}{\micro\metre}.
-
-\item\label{vid:nostick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
-10\% of the proteins are fluorescently labelled.
-The total duration of the Video is \SI{147}{\minute}.
-The scale bar is \SI{10}{\micro\metre}.
-
-\item\label{vid:cutDarcy} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}a.
-10\% of the proteins are fluorescently labelled.
-The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
-Here the gel initially lies close to the bottom wall, $H<H^*$.
-Darcy mechanism dominates the formation of the primary pattern.
-The total duration of the video is \SI{145}{\minute}.
-
-\item\label{vid:cutPoiseuille} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}c.
-10\% of the proteins are fluorescently labelled.
-The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3).
-Here the gel initially lies far from either top or bottom wall, $H>H^*$.
-Poiseuille mechanism dominates the formation of the primary pattern.
-The total duration of the video is \SI{181}{\minute}.
-
-\end{enumerate}
-
 \end{document}

### 2015-02-18 fig_plis/supplementary_plis.tex
Sépare le texte principal et les supplementary materials

\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,
amsmath,amssymb,prl]{revtex4-1}
\usepackage[utf8]{inputenc}

\usepackage{graphicx} 
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[range-units = single]{siunitx}
%\usepackage{multirow}
\usepackage{bm}
%\usepackage{dcolumn,ulem}
\usepackage{amsfonts}

\definecolor{Main}{rgb}{1, 0.57, 0}
\definecolor{Accent1}{rgb}{1,0.28,0}
\definecolor{Accent2}{rgb}{1,0.74,0}

\usepackage{xr,xc}
\externaldocument[M-]{article_plis}
\externalcitedocument[M-]{article_plis}

\begin{document}
\title{Supplementary informations to Hierarchical wrinkling in a confined permeable biogel}
\author{Mathieu Leocmach}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{Mathieu Nespoulous}
\affiliation{Aix-Marseille Université, CNRS, MADIREL UMR 7246, Marseille, France}
\author{Sébastien Manneville}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
\author{Thomas Gibaud}
\email{thomas.gibaud@ens-lyon.fr}
\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}

\maketitle

\section*{Fraction of free casein as a function of pH}

The fraction of free casein proteins of a \SI{100}{\milli\litre}, 4\%w sodium caseinate dispersion is obtained either by titration by 1 molar HCl, see Supplementary Video~\ref{vid:titration}, or during the action of 4\%w GDL. In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of pH, and centrifuged (centrifuge 5424, Eppendorf) in Eppendorf tubes at 14500 rpm for \SI{3}{\minute}. The supernatant is diluted and analysed with a UV absorption spectrometer (USB4000, Ocean Optics). According to~\cite{M-Roefs1986}, we deduce the fraction of free casein proteins from the adsorption spectrum as:
%
\begin{align}
x_\text{free} &= \frac{A(\text{pH})}{A(\text{pH}=7)},\\
\text{with }A(\text{pH}) &= A_{280}-1.706 A_{320},
\end{align}
%
where $A_{280}$ is the absorption peak at \SI{280}{\nano\metre} due to aromatic amino acids in the casein molecules plus a turbidity background whereas the absorption peak at \SI{320}{\nano\metre} is only due to turbidity.

Above the isoelectric point, the homogeneous acidification by GDL continuously creates very small protein aggregates that are difficult to centrifuge. Such suspensions are too turbid and too rapidly evolving to reliably measure $x_\text{free}$.

\section*{Microstructure measurement}

\begin{figure}
	\includegraphics{nonmonotonic}%
	\caption{Microstructure evolution of a 4\%w sodium caseinate solution acidified with 4\%w GDL in water. (a) Details of confocal images in fully adhesive conditions (top row) and with no adhesion on top (bottom row). The scale bar is \SI{10}{\micro\metre}. Arrows indicates the pore size. (b) Corresponding Fourier spectra when no adhesion on top, same times as (a). Lines are fits to equation~(\ref{eq:fractalS}). The inset shows a schematic association between real space structures and wave numbers. (c-e) Evolution of cut-off length $\xi$, susceptibility $\chi$ and fractal dimension $d$ in full adhesive (black line) and no adhesion on top (orange dashed line) situations. Symbols indicate the times of the images in (a).}%
	\label{fig:nonmonotonic}%
\end{figure}

To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre}, Supplementary Figure~\ref{fig:nonmonotonic}a and Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63}. After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$, see Supplementary Figure~\ref{fig:nonmonotonic}b that is fitted by the following fractal form~\cite{M-VanDijk1986}:
\begin{equation}
I(q) = \frac{2\chi\Gamma(d)}{\left(1+\left((d+1)\xi q\right)^2\right)^{d/2}}
\label{eq:fractalS}
\end{equation}
where the susceptibility $\chi$ is related to the fluorescence contrast between the network and the solvent, the cut-off length $\xi$ corresponds to the largest pore size and $d$ is the fractal dimension. $\Gamma$ is the Gamma function. Equation~(\ref{eq:fractalS}) is a good approximation to the 2D Fourier transform of a fractal pair correlation function $g(r) \sim 1+ e^{-r/\xi} r^{d-2}$.
%En 3D on peut citer \cite{Martin1987,Chen1986}

Figure~\ref{M-fig:acidification}d of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c-e show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling. The latter situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope. To do so, we coat only the microscope slide with acrylamide brush. The cover slip is cleaned but not coated, thus keeping strong adhesion on the side closer to the objective lens. Neither buckling nor wrinkling is detected in such a configuration.

In situation (i), the gel keeps its topology throughout the gelation and over-acidification. Left of Supplementary Figure~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again. Syn\ae{}resis and swelling are thus local processes. By contrast in situation (ii), the gel as a whole compresses while forming. Since our focal plane is at a fixed depth, we cannot follow the topology of the same subset of the network, however the pore size is clearly smaller than in situation (ii). The isoelectric gel, at minimum gel volume, is probably isotropic or only marginally anisotropic in $z$. This explains that subsequent swelling is not occurring only along $z$ but also in plane, causing buckling.

$\xi$ is also systematically measured in the same way at the end of each wrinkling experiment and reported in Supplementary Table~\ref{tab:data}. %Final fractal dimension was always close to 2 within error bars. Comparing $\chi$ across different setting of the microscope is not relevant.

\section*{Permeability measurements}
\begin{figure*}%
	\includegraphics{permeability}%
	\caption{Permeability measurements. (a) Schematic representation of the experimental set up. (b-c) Evolution of the height of the interface in tube 1 relative to the final height in tube 2. The black line is the best exponential fit $Ae^{-t/\tau}$. (b) Gel is 4\%w casein, 4\%w GDL in water, $H=\SI{2.3}{\milli\metre}$ and $\tau=\SI{57}{\minute}$. (c) Same as (b) for a 50\%w glycerol--water mixture, $H=\SI{4}{\milli\metre}$ and $\tau=\SI{100}{\hour}$.}%
	\label{fig:permeability}%
\end{figure*}

To measure the permeability of our gels, we adapt a method from \cite{M-VanDijk1986}. First, the gel of interest is formed at the lower extremity of a thin glass tube. The tube is held vertical in a water saturated atmosphere to prevent evaporation until the pH has reached the isoelectric point ($\approx\SI{1}{\hour}$ for our most common composition). Then, we dip the tube in a bath of acetic acid/acetate buffer at pH=4. The buffer has the same glycerol content as the gel solvent, so that the viscosity is constant throughout the experiment. Alongside the first tube, see Supplementary Figure~\ref{fig:permeability} a, we dip a second identical tube with no gel in it. A webcam (Logitech Webcam Pro 9000) captures the rise of the liquid in each tube. Since the liquid height in tube 2 reaches its final value quickly, we measure directly $z_\infty-z(t)$.

Darcy law through the gel of permeability $\alpha$ and height $H$ reads 
\begin{equation}
\frac{dz}{dt} = \frac{\alpha}{\eta}\frac{\Delta P}{H}
\end{equation}
where $\Delta P=\rho g(z_\infty-z(t))$ is the hydrostatic pressure drop through the gel. We use tabulated values of water-glycerol mixtures for the density $\rho$ and viscosity $\eta$ of the buffer. We can thus write
\begin{equation}
\frac{d(z-z_\infty)}{dt} = -\frac{z-z_\infty}{\tau}\text{, with\,} \tau = \frac{\eta H}{\alpha \rho g}.
\end{equation}

By fitting $z(t)-z_\infty$ to an exponential, see Supplementary Figure~\ref{fig:permeability} b-c, we obtain $\tau$ and thus $\alpha$. Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due to both high viscosity and low permeability. Obtaining $z_\infty$ without tube 2 would be time prohibitive.

\section*{Confocal images processing}

At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling). Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}. At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.

We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$). Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$. We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line. Since this criterion is local, we do not need to correct for background intensity variation. In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.

From these data, we obtain for each stack the volume and thickness of the gel phase, area and peak-to-peak amplitude of the mid surface. Velocity is obtained by time differentiation. Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.

\section*{Characteristic time measurement}
\begin{figure}
	\includegraphics{growth}
	\caption{Growth of the instability. (a) Evolution of the gel velocity $v$ averaged over the whole surface. Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scales. The black line is the best exponential fit with $\tau=\SI{90}{\second}$. (b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown). Note that the longitudinal size of the blister evolves in this regime. The thick black curve is the profile at much longer times and is at lower altitude due to swelling. Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
	\label{fig:growth}
\end{figure}

In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements. However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface. Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.

We note that when both measurements were available (as in Supplementary Figure~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel. We use this approximation of $\tau$ for all our samples. This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact, see Supplementary Figure~\ref{fig:growth}b, thus the kinetic wavelength is set when $A(t) = A_\text{max}$.

\section*{Mixed Darcy--Poiseuille model}

\subsection*{General framework for wrinkling}

The mechanical equilibrium of a plate of thickness $h$, Young modulus $E$ and Poisson ratio $\nu$ submitted to a compression load $\sigma_\parallel$ along $x$ and a transverse load $\sigma_\perp$ along $z$ writes
\begin{equation}
\frac{E h^3}{12(1-\nu^2)}\frac{\partial^4 w}{\partial x^4} + \sigma_\parallel h \frac{\partial^2 w}{\partial x^2} = \sigma_\perp,
\label{eq:mecheq}
\end{equation}
where $w(x,t)$ is the deflection of the plate along $z$~\cite{M-Biot1957}. Since we deal with slow rates of deformation, inertia is neglected.

We decompose the deflection in sinusoidal modes $w(x,t) =  A \sin (qx)$ of wavevector $q$. Equation~(\ref{eq:mecheq}) then reads
\begin{equation}
\left(\frac{1}{12} B h^3 q^4 - \sigma_\parallel h q^2\right)w = \sigma_\perp,
\label{eq:mecheqop}
\end{equation}
where $B=E/(1-\nu^2)$ is the modulus of the plate.

Following \cite{M-Biot1957}, we consider the case where the compression load $\sigma_\parallel$ is increased quasistatically. The observed mode is the one that arises at the lowest possible $\sigma_\parallel$, i.e. the mode minimising $\sigma_\parallel$.

\subsection*{Elastic film in an infinite elastic medium}

In this framework, we recall here the case of an elastic film lying over an infinite elastic medium~\cite{M-Gough1940,M-Bijlaard1946,M-Biot1957,M-Cerda2003}. The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -q B_s w$, and thus the compression load writes
\begin{equation}
\sigma_\parallel = B \frac{h^2}{12} q^2 + \frac{B_s}{hq}.
\label{eq:sigma0}
\end{equation}

Minimising $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{M-Biot1957,M-Cerda2003} so that the dominant wavelength is 
\begin{equation}
\lambda = 2\pi h \left(\frac{B}{6B_s}\right)^{1/3}.
\label{eq:lambdaElEl}
\end{equation}

\subsection*{Elastic film on a viscous layer}

\begin{figure}
\includegraphics{notations}%
\caption{Schematic side view of the slit. The constant thickness gel film (orange) is surrounded by solvent. The region of interest for mass conservation is highlighted in gray. Arrows are oriented along $x$ or $z$, defining positive sign.}%
\label{fig:notations}%
\end{figure}

Let us now turn to the case of an elastic, impermeable film on a single viscous layer. As in \cite{M-Huang2002} we neglect gravity, although it would be difficult to do so in practice at a free interface. Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation. Further considering only small deflections, we may neglect the displacement of the plate along $x$. The flux along $x$ is thus of the Poiseuille form
\begin{equation}
Q = -\frac{H^3}{12\eta}\frac{\partial p}{\partial x}.
\label{eq:PoiseuilleFlux}
\end{equation}
where $p(x)$ is the pressure in the layer of viscosity $\eta$.

Mass conservation (on the lower gray area of Supplementary Figure~\ref{fig:notations}) writes
\begin{equation}
\frac{\partial Q}{\partial x} + \frac{\partial w}{\partial t} = 0.
\label{eq:conservation}
\end{equation}

Assuming an exponentially growing amplitude $A(t) = A_0 \exp(t/\tau)$ and keeping only linear order terms, one gets
\begin{equation}
\frac{12\eta}{\tau} \frac{w}{H^3q^2} = -\sigma_\perp.
\label{eq:PoiseuilleLoad}
\end{equation}

Combining equations~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
\begin{equation}
\sigma_\parallel = \frac{1}{12}B h^2 q^2 + 12\frac{\eta}{\tau}\frac{1}{H^3 h q^4}
\label{eq:sigma0P}
\end{equation}
and is minimised at $q_P$ such that
\begin{equation}
q_P^6 = 2\times 12^2 \frac{1}{h^3H^3\Upsilon},
\label{eq:qP}
\end{equation}
with $\Upsilon = B\tau/\eta$ the viscoelastic factor. The dominant wavelength for the Poiseuille model is thus
\begin{equation}
\lambda_P = \pi\sqrt{hH}\left(\frac{2}{9}\Upsilon\right)^{1/6}.
\end{equation}
%
We note that the mode minimising $\sigma_\parallel$ is also maximising the growth rate
\begin{equation}
\frac{1}{\tau} = \frac{H^3q^2}{12^2\eta}\left(12\sigma_\parallel - Bh^3q^4\right).
\end{equation}

When the elastic film is sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$, the linearity of equation~(\ref{eq:mecheq}) allows us to use an effective substrate of thickness 
\begin{equation}
\frac{1}{H^3} = \frac{1}{H_1^3}+\frac{1}{H_2^3}.
\end{equation}
This corresponds to the experimental case where the gel film is far from both of the cell walls. Since we have an almost buoyant gel and no more free surface, gravity can indeed be neglected.

\subsection*{Porous elastic film with no possible Poiseuille flow}
Another limiting case is that of a porous elastic film immersed in a viscous medium where longitudinal resistance to flow is infinite. It is the case in our experiments when the film is at contact with one of the walls. The only way to deform the film is to flow some liquid through the film of permeability $\alpha$. The transverse load derives from the Darcy law:
\begin{equation}
\sigma_\perp = p_1-p_2 = \frac{h\eta}{\alpha}\frac{\partial w}{\partial t}.
\label{eq:DarcyLoad}
\end{equation}

Combining equations~(\ref{eq:mecheqop}) and (\ref{eq:DarcyLoad}), the compression load writes
\begin{equation}
\sigma_\parallel = \frac{1}{12}B h^2 q^2 + \frac{\eta}{\tau}\frac{1}{\alpha q^2}
\label{eq:sigma0D}
\end{equation}
and is minimised at $q_D$ such that
\begin{equation}
q_D^4 = \frac{12}{h^2\alpha\Upsilon}.
\label{eq:qD}
\end{equation}
%
The dominant wavelength for Darcy model is thus
\begin{equation}
\lambda_D = 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}.
\end{equation}
Here also the dominant mode maximises the growth rate.

\subsection*{Porous elastic film between two viscous layers}
Finally the most realistic scenario mixes both Poiseuille and Darcy mechanisms, i.e. considers a porous elastic film between two viscous layers of arbitrary thickness. In this case, see Supplementary Figure~\ref{fig:notations}, the mass conservation $Q_1 + Q_2 = 0$ over the whole height of the cell yields
\begin{equation}
H_1^3 p_1 + H_2^3 p_2 = 0.
\label{eq:pressures}
\end{equation}

The mass conservation in the lower viscous layer is the same as equation~(\ref{eq:conservation}) with an added leak $v$ due to the porosity
\begin{equation}
\frac{\partial Q_1}{\partial x} + \frac{\partial w}{\partial t} + v = 0.
\label{eq:conservationDarcy}
\end{equation}
%
$v$ can be expressed by the Darcy law
\begin{equation}
v = \frac{\alpha}{\eta} \frac{p_1-p_2}{h}.
\label{eq:Darcy}
\end{equation}

Combining equations~(\ref{eq:pressures}), (\ref{eq:conservationDarcy}) and (\ref{eq:Darcy}) we obtain the transverse load
\begin{equation}
\sigma_\perp = p_1-p_2 = - \frac{\eta}{\tau} \frac{w}{\frac{H^3 q^2}{12} + \frac{\alpha}{h}},
\label{eq:MixtLoad}
\end{equation}
%
Note that we recover equation~(\ref{eq:PoiseuilleLoad}) when $\alpha \rightarrow 0$ and equation~(\ref{eq:DarcyLoad}) when $H \rightarrow 0$. We can recast this intuition in terms of the dimensionless number $H/H^*$ where $H^*$ is obtained by equating $q_P$ and $q_D$:
\begin{equation}
H^* = 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}.
\end{equation}

Analytic minimisation of $\sigma_\parallel$ knowing $\sigma_\perp$ is possible but cumbersome. Instead we rewrite equation~(\ref{eq:mecheqop}) in terms of previously calculated $\lambda_P$ and $\lambda_D$:
\begin{align}
\frac{12\sigma_\parallel}{q_D^2 h^2 B} &= \left(\frac{q}{q_D}\right)^2 + \frac{1}{2\frac{q_D^2}{q_P^6}q^4  + \left(\frac{q}{q_D}\right)^2}\\
&= \left(\frac{\lambda_D}{\lambda}\right)^2 + \frac{1}{2\frac{\lambda_P^6}{\lambda_D^2\lambda^4}  + \left(\frac{\lambda_D}{\lambda}\right)^2}
\end{align}
and we minimise this expression numerically.

\section*{Samples characteristics}
To test the theoretical predictions we have fully characterised 9 samples. Their characteristics are gathered in Supplementary Table \ref{tab:data}.

\begin{table*}
\begin{tabular}{cSSSScSrrScSSSScSSSSSS}
%\toprule
&\multicolumn{4}{c}{Preparation} & &  \multicolumn{4}{c}{Properties} & & \multicolumn{4}{c}{Measurements} & & \multicolumn{6}{c}{Results} \\ 
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
\# &{cas} & {GDL} & {gly} & {$e$} && {$\eta$} & \multicolumn{1}{c}{$G^\prime$} & \multicolumn{1}{c}{$\alpha$} & {$\xi$} && {$h$} & {$H_1$} & {$H_2$} & {$v_\text{max}$} && {$\Upsilon$} & {$H/H^*$} & {$\lambda_\text{exp}$} & {$\lambda_{D}$} & {$\lambda_{P}$} & {$\lambda_{D+P}$} \\ 
&{\%w} & {\%w} & {\%w} & \si{\micro\metre} && \si{\milli\pascal\second} & \multicolumn{1}{c}{\si{\pascal}} & \si{\square\nano\metre} & \si{\micro\metre} && \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre} & \si{\micro\metre/\second} && {$10^8$} &  & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} & \si{\milli\metre} \\ 
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
%ech2
a & 4 & 4 & 0 & 74.5 && 1.0 & 487 & 67000 & 4.9 && 38.3 & 6.3 & 29.3 & 0.11 && 1.62 & 0.54 & 0.89 & 1.20 & 0.88 & 1.22 \\ 
& & & & && & & & && 40.2 & 1.7 & 32.9 & 0.09 && 0.70 & 0.09 & 0.45 & 1.00 & 0.39 & 1.00\\
b & 4 & 4 & 0 & 92.4 && 1.0 & 487 & 67000 & 4.9 && 27.8 & 13.5 & 50.8 & 0.20 && 1.57 & 1.17 & 0.85 & 1.01 & 1.10 & 1.20 \\ 
c & 4 & 4 & 0 & 106.1 && 1.0 & 487 & 67000 & 4.9 && 32.7 & 30.7 & 42.9 & 0.14 && 2.54 & 2.23 & 1.52 & 1.24 & 1.85 & 1.90 \\ 
d & 4 & 4 & 0 & 138.0 && 1.0 & 487 & 67000 & 4.9 && 48.2 & 32.0 & 63.9 & 0.11 && 4.05 & 2.28 & 1.62 & 1.69 & 2.56 & 2.62 \\
& & & & && & & & && 57.4 & 2.1 & 80.9 & 0.05 && 1.25 & 0.09 & 0.81 & 1.37 & 0.6 & 1.37 \\
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
e & 8 & 8 & 0 & 85.2 && 1.0 & 1817 & 2150 & 3.1 && 49.8 & 10.0 & 26.8 & 0.10 && 6.50 & 3.79 & 1.00 & 0.82 & 1.60 & 1.61 \\ 
& & & & && & & & && 60.5 & 1.1 & 24.0 & 0.11 && 1.94 & 0.42 & 0.50 & 0.77 & 0.55 & 0.82\\
f & 8 & 8 & 0 & 91.0 && 1.0 & 1817 & 2150 & 3.1 && 23.4 & 30.2 & 47.7 & 0.10 && 12.05 & 9.70 & 1.50 & 0.66 & 2.04 & 2.04 \\ 
& & & & && & & & && 30.7 & 1.6 & 55.0 & 0.02 && 15.60 & 0.32 & 0.75 & 0.82 & 0.59 & 0.83 \\
\cline{2-5} \cline{7-10} \cline{12-15} \cline{17-22}\\[-2ex]
g & 4 & 4 & 20 & 63.7 && 1.6 & 1142 & 24600 & 2.7 && 39.1 & 1.2 & 10.9 & 0.21 && 0.86 & 0.20 & 0.49 & 0.80 & 0.36 & 0.80 \\ 
h & 4 & 4 & 35 & 82.3 && 2.7 & 1858 & 25300 & 1.8 && 58.3 & 0.9 & 12.6 & 0.02 && 10.98 & 0.13 & 0.61 & 1.06 & 0.56 & 1.06 \\ 
i & 4 & 4 & 50 & 95.3 && 5.4 & 4660 & 5700 & 1.5 && 65.1 & 7.9 & 14.4 & 0.02 && 16.30 & 1.52 & 1.23 & 1.50 & 1.85 & 1.96 \\ 
\end{tabular}
\caption{Characteristics of the samples used for Main Figure~\ref{M-fig:DarcyPoiseuille}. Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line. `cas', `GDL' and `gly' indicate the weight fraction of sodium caseinate, GDL and glycerol respectively. $\xi$ is the pore size at the end of each experiment. $h$, $H_1$ and $H_2$ are measured just before buckling.}
\label{tab:data}
\end{table*}

\section*{Wavelength measurement}

\begin{figure}
	\includegraphics{generations}%
	\caption{Fourier spectra of binary altitudes of Main Figure~\ref{M-fig:dynamics}a corresponding to primary, secondary and tertiary patterns. Inset: reconstructed binary altitudes at time \SI{2}{\hour}. The scale bar is \SI{1}{\milli\metre}.}%
	\label{fig:generations}%
\end{figure}

\begin{figure*}
	\includegraphics{patterns}
	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data}. (a-d) increasing cell thickness. (e-f) larger solid content. (g-i) increasing glycerol content. All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images. The scale is common to all panels (scale bar \SI{1}{\milli\metre}). Arrows show the measured primary wavelength.}
	\label{fig:patterns}
\end{figure*}

To measure the wrinkle wavelength $\lambda$, we use large field pictures of the final state. These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e. The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film. White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall). It is thus possible to reconstruct a binary picture of the pattern, inset of Supplementary Figure~\ref{fig:generations}. Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final binary picture to the configuration of the primary pattern.

Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times. When only the first generation pattern is present a peak clearly indicates the dominant mode. When the secondary pattern has appeared, the dominant wavelength is halved and the peak much weaker. This allows us to use the secondary pattern is our analysis by setting their wavelength to half of the primary wavelength. Further generations do not seem to appear simultaneously and the corresponding peaks are weak and broad.

\section*{Supplementary Videos}

\begin{enumerate}
\renewcommand{\labelenumi}{\textbf{Video \arabic{enumi}:}}
\itemindent 8mm
\item\label{vid:transmitted} Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion. Top and bottom slides of the optical cell are coated with acrylamide brushes. The total duration of the Video is \SI{2}{\hour}~30. The scale bar is \SI{1}{\milli\metre}.

\item\label{vid:reconstructed} 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion, Supplementary Table~\ref{tab:data}c. 10\% of the proteins are fluorescently labelled. Top and bottom slides of the optical cell are coated with acrylamide brushes. The total duration of the Video is \SI{2}{\hour}~30. The scale bar is \SI{1}{\milli\metre}.

%\item[Video 3] Titration of a dispersion with 1\%w casein by 1 molar HCl. The duration of the Video is \SI{13}{\minute}. The scale bar is XXX $\mu$m

\item\label{vid:titration} Titration of a dispersion with 4\%w casein by 1 molar HCl. The total duration of the Video is \SI{14}{\minute}. The Becher diameter is \SI{55}{\milli\metre}.

\item\label{vid:stick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on. 10\% of the proteins are fluorescently labelled. The total duration of the Video is \SI{132}{\minute}. The scale bar is \SI{10}{\micro\metre}.

\item\label{vid:nostick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky. 10\% of the proteins are fluorescently labelled. The total duration of the Video is \SI{147}{\minute}. The scale bar is \SI{10}{\micro\metre}.

\item\label{vid:cutDarcy} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}a. 10\% of the proteins are fluorescently labelled. The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3). Here the gel initially lies close to the bottom wall, $H<H^*$. Darcy mechanism dominates the formation of the primary pattern. The total duration of the video is \SI{145}{\minute}.

\item\label{vid:cutPoiseuille} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}c. 10\% of the proteins are fluorescently labelled. The scale bar is \SI{100}{\micro\metre} (with a pixel size ratio of 1.3). Here the gel initially lies far from either top or bottom wall, $H>H^*$. Poiseuille mechanism dominates the formation of the primary pattern. The total duration of the video is \SI{181}{\minute}.

\end{enumerate}

\end{document}

### 2015-02-19 fig_plis/supplementary_plis.tex
description des vidéos : sheet->film titres : measurementS

@@
 
-\section*{Microstructure measurement}
+\section*{Microstructure measurements}
 
@@
 
-\section*{Characteristic time measurement}
+\section*{Characteristic time measurements}
 \begin{figure}
@@
 
-\section*{Wavelength measurement}
+\section*{Wavelength measurements}
 
@@
 
-\item\label{vid:cutDarcy} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}a.
+\item\label{vid:cutDarcy} Confocal microscopy $(x,z)$ cut of the evolution of the gel film deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}a.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item\label{vid:cutPoiseuille} Confocal microscopy $(x,z)$ cut of the evolution of the gel sheet deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}c.
+\item\label{vid:cutPoiseuille} Confocal microscopy $(x,z)$ cut of the evolution of the gel film deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}c.
 10\% of the proteins are fluorescently labelled.

### 2015-02-25 fig_plis/article_SI.tex
Recombine article et SI pour express submission

--- 2015-02-18 fig_plis/article_plis.tex
@@
 \definecolor{Accent2}{rgb}{1,0.74,0}
-
-\newcommand{\seb}[1]{\textbf{\color{blue}#1}} % remarques Seb
-\newcommand{\sseb}[1]{\sout{\textbf{\color{blue}#1}}} % remarques Seb
 
@@
 
-\bibliographystyle{naturemag3}
+\bibliographystyle{pnas2009}
 \bibliography{Yaourt}
@@
 
-\subsection*{Microstructure measurement}
+\section*{Microstructure measurements}
 
@@
 
-\subsection*{Characteristic time measurement}
+\section*{Characteristic time measurements}
 \begin{figure}
@@
 
-\subsection*{Wavelength measurement}
+\section*{Wavelength measurements}
 

### 2015-02-25 fig_plis/article_plis.tex
documentclass sur une seule ligne pour être compatible avec texcount
PNAS version with figures at the end

@@
-\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,
-amsmath,amssymb,prl]{revtex4-1}
+%% PNAStwoS.tex
+%% Sample file to use for PNAS articles prepared in LaTeX
+%% For two column PNAS articles
+%% Version1: Apr 15, 2008
+%% Version2: Oct 04, 2013
+
+%% BASIC CLASS FILE
+\documentclass{pnastwo}
 \usepackage[utf8]{inputenc}
@@
 %\usepackage{multirow}
-\usepackage{bm}
+%\usepackage{bm}
 %\usepackage{dcolumn,ulem}
-\usepackage{amsfonts}
+%\usepackage{amsfonts}
 
@@
 \title{Hierarchical wrinkling in a confined permeable biogel}
-\author{Mathieu Leocmach}
-\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
-\author{Mathieu Nespoulous}
-\affiliation{Aix-Marseille Université, CNRS, MADIREL UMR 7246, Marseille, France}
-\author{Sébastien Manneville}
-\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
-\author{Thomas Gibaud}
-\email{thomas.gibaud@ens-lyon.fr}
-\affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
+\author{Mathieu Leocmach\affil{1}{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France},
+Mathieu Nespoulous\affil{2}{Aix-Marseille Université, CNRS, MADIREL UMR 7246, Marseille, France},
+Sébastien Manneville\affil{1}{},
+Thomas Gibaud\affil{1}{}}
+%\email{thomas.gibaud@ens-lyon.fr}
 
+\significancetext{ML and MN conducted preliminary experiments.
+ML, SM and TG designed the experiments and interpreted the results.
+ML and TG performed quantitative experiments.
+ML performed data analysis and modelling.
+ML, SM and TG wrote the manuscript.}
+
+\maketitle
+
+\begin{article}
 \begin{abstract}
@@
 
-\maketitle
+\keywords{wrinkling | porosity | gel | syn\ae{}resis}
 
-\section*{Introduction}
+\abbreviations{GDL, glucono-$\delta$-lactone}
 
-%There are many ways and reasons for a film to wrinkle.
-An elastic film buckles due to excess area with respect to its boundaries.
-Wrinkling is a buckling hindered by a substrate: the mismatch of elastic properties between the film and the substrate selects the preferred wrinkling wavelength~\cite{Gough1940, Bijlaard1946}.
-On the engineering side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
-Such patterns can be triggered by temperature dilation, swelling or the removal of pre-strain.
-On the biology side wrinkling-controlled morphogenesis is ubiquitous: ageing and the loss of elastic fibres makes our skin wrinkles~\cite{Bissett1987, Genzer2006}, difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}, localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
-On the physics side analogues of wrinkling have been found in situations where an effective substrate comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
-Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Stability analysis of a film lying on a thin viscous substrate was studied only recently~\cite{Huang2002}.
-Since in practice gravity sets the wavelength in such situation~\cite{Smoluchowski1910}, the wrinkling on a confined viscous substrate has not been studied experimentally.
-
-There are many ways and reasons for a film to wrinkle.
+\dropcap{T}here are many ways and reasons for a film to wrinkle.
 An elastic film buckles due to excess area with respect to its boundaries and wrinkles when buckling is hindered by a substrate.
@@
 Such patterns can be triggered by temperature dilation~\cite{Bowden1998}, swelling~\cite{Hu1998,Kim2010} or the removal of pre-strain~\cite{Genzer2006}.
-
-%ph swelling: hu_1998_Polymer gels with engineered environmentally and kim_2010_Dynamic display of biomolecular patterns
-%compressive stress: kim_2011_Hierarchical folding of elastic membranes under
 
@@
 
-\begin{figure}
-	\includegraphics{dynamics}%
-	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w in water).
-(a) Light transmission microscopy.
-Successive generations of patterns are highlighted in colour to stress the absence of coarsening after formation.
-(b) 3D reconstruction from fluorescent confocal microscopy.
-Scale bars are \SI{1}{\milli\metre}.}%
-	\label{fig:dynamics}%
-\end{figure}
+%fig:dynamics
 
@@
 This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
-%Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel film for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel film for short wavelengths.
 
-\section*{System design}
+\section{System design}
 
-\begin{figure}
-	\includegraphics{acidification}%
-	\caption{Acid-induced protein gel properties behave non monotonically with pH.
-(a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL in water.
-The horizontal line indicates isoelectric pH of caseins.
-(b) Corresponding evolution of the fraction of free caseins.
-Symbols are direct measures upon addition of GDL.
-The continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert).
-(c) Evolution of the elastic modulus $G'$ measured in a rheometer with full adhesion to the cone-plane geometry.
-(d) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
-allowed syn\ae{}resis and swelling (orange dashed line).}%
-	\label{fig:acidification}
-\end{figure}
+%fig:acidification
 
@@
 At pH=2 the fraction of free casein is 0.35.
-
-%On the Stability of Casein Micelles, Pieter Walstra J. of dairy science 1990
 
@@
 
-%XXXCe paragraphe fait-il sens ici ?XXX As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
-This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off.
-To quantify this effect, we make the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}.
-In the Supplementary Figure~\ref{S-fig:permeability}, we observe that the aqueous buffer goes through the gel and rises in the tube due to capillary forces.
-Exploitation of this experiment give the permeability of the gel: \SI{67000}{\square\nano\metre}.
-
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
@@
 
-%We coat the glass slides with acrylamide brushes to prevent the adhesion of the casein.
-In fig.2c, when the top slide of the optical cell is coated with acrylamide brushes, we then observe that the casein network detaches from the top slide.
-$\xi$ does not peak at the isoelectric point but remains constant
-
-%Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-Once the space spanning network is formed and sticks to the container, protein attraction can only proceed by local shrinking of the protein agregates, leading to larger pores.
-
-%The difference in value of $\xi$ between the fully un-coated  and the half-coated optical cell demonstrate that the gel is subject to constrains as the pH decreases.
-The relaxation of thoses constrain is key to the wrinkling formation when both the top and bottom slide are coated with acrylamide brushes.
-
-%on a change les condition daccroche ce aui permet au contrainte de se relacher
-
-%The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation~\cite{Bremer1989} (Fig.~\ref{fig:acidification}a).
-Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
-Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplementary Fig.~\ref{S-fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
-In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{S-fig:nonmonotonic}).
-
-\section*{Dynamics of the wrinkling process}
+\section{Dynamics of the wrinkling process}
 
@@
 
-\begin{figure}
-	\includegraphics{sideview}
-	\caption{Measurements and modelling.
-(a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
-The scale bar is \SI{100}{\micro\metre} (real size ratio).
-(b) Schematic side view of the cell.
-The constant thickness gel film (orange) is surrounded by solvent.
-Yellow lines show the path of transmitted light through the gel phase.
-(c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
-Crosses correspond to times in (a).}
-	\label{fig:sideview}
-\end{figure}
+%fig:sideview
 
-\section*{Models and scaling laws}
-
-%Longitudinally constrained swelling of the gel film is the origin and driving force of the buckling instability.
-The gel area becomes in excess compared to the available area, thus the gel must fold.
-However, this does not explains how a wavelength is selected.
+\section{Models and scaling laws}
 
@@
 
-%L. Landau and E.M. Lifshitz, Theory of Elasticity (Pergamon, New York, 1986), 3rd ed.
-
 This framework has been successfully applied to situations where $\sigma_{\perp}$ comes from the resistance of an elastic substrate ~\cite{Gough1940, Bijlaard1946}.
@@
 
-\begin{figure*}
-	\includegraphics{Darcy_vs_Poiseuille}
-	\caption{Comparing model predictions $\lambda_D$, $\lambda_P$ and $\lambda_{D+P}$ with measured wavelengths $\lambda_\text{exp}$.
-Dots come from primary pattern, squares from secondary blisters.
-Lines are the best linear fits through the origin taking into account only the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
-Prefactors are 0.63, 0.69 and 0.67 respectively.
-The dashed line in (b) is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all data points.}
-	\label{fig:DarcyPoiseuille}
-\end{figure*}
+%fig:DarcyPoiseuille
 
@@
 
-\section*{Experimental tests}
+\section{Experimental tests}
 
@@
 
-\section*{Conclusion}
+\section{Conclusion}
 
@@
 
-%role of gravity
-
-\section*{Methods summary}
+\begin{materials}
 
@@
 The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
+\end{materials}
 
-\bibliographystyle{naturemag3}
-\bibliography{Yaourt}
-
-\paragraph*{Acknowledgements}
+\begin{acknowledgments}
 Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
@@
 258803 and from Institut Universitaire de France.
+\end{acknowledgments}
 
-\paragraph*{Author Contributions}
-ML and MN conducted preliminary experiments.
-ML, SM and TG designed the experiments and interpreted the results.
-ML and TG performed quantitative experiments.
-ML performed data analysis and modelling.
-ML, SM and TG wrote the manuscript.
+\bibliographystyle{pnas2009}
+\bibliography{Yaourt}
 
-\paragraph*{Author Information} 
-The authors declare that they have no competing financial interests. 
-Correspondence and requests for materials should be addressed to TG (\href{mailto:thomas.gibaud@ens-lyon.fr}{thomas.gibaud@ens-lyon.fr}).
+\end{article}
+
+%\paragraph*{Author Information} 
+%The authors declare that they have no competing financial interests. 
+%Correspondence and requests for materials should be addressed to TG (\href{mailto:thomas.gibaud@ens-lyon.fr}{thomas.gibaud@ens-lyon.fr}).
+
+\begin{figure}
+	\includegraphics{dynamics}%
+	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w in water).
+(a) Light transmission microscopy.
+Successive generations of patterns are highlighted in colour to stress the absence of coarsening after formation.
+(b) 3D reconstruction from fluorescent confocal microscopy.
+Scale bars are \SI{1}{\milli\metre}.}%
+	\label{fig:dynamics}%
+\end{figure}
+
+\begin{figure}
+	\includegraphics{acidification}%
+	\caption{Acid-induced protein gel properties behave non monotonically with pH.
+(a) pH decrease over time in a 4\%w sodium caseinate solution acidified by 4\%w GDL in water.
+The horizontal line indicates isoelectric pH of caseins.
+(b) Corresponding evolution of the fraction of free caseins.
+Symbols are direct measures upon addition of GDL.
+The continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert).
+(c) Evolution of the elastic modulus $G'$ measured in a rheometer with full adhesion to the cone-plane geometry.
+(d) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
+allowed syn\ae{}resis and swelling (orange dashed line).}%
+	\label{fig:acidification}
+\end{figure}
+
+\begin{figure}
+	\includegraphics{sideview}
+	\caption{Measurements and modelling.
+(a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
+The scale bar is \SI{100}{\micro\metre} (real size ratio).
+(b) Schematic side view of the cell.
+The constant thickness gel film (orange) is surrounded by solvent.
+Yellow lines show the path of transmitted light through the gel phase.
+(c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
+Crosses correspond to times in (a).}
+	\label{fig:sideview}
+\end{figure}
+
+\begin{figure*}
+	\includegraphics{Darcy_vs_Poiseuille}
+	\caption{Comparing model predictions $\lambda_D$, $\lambda_P$ and $\lambda_{D+P}$ with measured wavelengths $\lambda_\text{exp}$.
+Dots come from primary pattern, squares from secondary blisters.
+Lines are the best linear fits through the origin taking into account only the points that should be (a) in Darcy mode $H<H^*$, (b) in Poiseuille mode $H>H^*$ (c) all points.
+Prefactors are 0.63, 0.69 and 0.67 respectively.
+The dashed line in (b) is the best affine fit ($\lambda_{\rm exp}=0.52\lambda_{P}+\SI{0.33}{\milli\metre}$) to all data points.}
+	\label{fig:DarcyPoiseuille}
+\end{figure*}
 

### 2015-02-26 fig_plis/article_plis.tex
Significance statement

@@
 Mathieu Nespoulous\affil{2}{Aix-Marseille Université, CNRS, MADIREL UMR 7246, Marseille, France},
-Sébastien Manneville\affil{1}{},
+Sébastien Manneville\affil{1}{} \and
 Thomas Gibaud\affil{1}{}}
@@
 
-\significancetext{ML and MN conducted preliminary experiments.
+%\contributor{Author contributions: ML and MN conducted preliminary experiments.
 ML, SM and TG designed the experiments and interpreted the results.
@@
 ML performed data analysis and modelling.
-ML, SM and TG wrote the manuscript.}
+ML, SM and TG wrote the manuscript.\\
+%The authors declare no conflict of interest.\\
+%This article is a PNAS Direct Submission.\\
+%To whom correspondence should be addressed. Email: thomas.gibaud@ens-lyon.fr}
+
+\significancetext{Confined thin surfaces may wrinkle due the growth of an excess of material.
+Here we present benchmark experiments that explore the possibility of wrinkling in biogels.
+We relate the excess area to the microstructural dynamics of the biogel.
+Usually the wrinkling wavelength is set by elasticity or gravity.
+In our case the wavelength is controlled by viscous flow, either in a lubrication layer or through the pores of the gel.
+Our results pave the way for exploring the possibility of wrinkling in confined porous soft materials immersed in a floating in a liquid such as biological tissues: during morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.}
 

### 2015-02-27 fig_plis/article_SI.tex
Significance statement in text

@@
 \maketitle
+
+\paragraph*{Significance statement}
+Confined thin surfaces may wrinkle due the growth of an excess of material.
+Here we present benchmark experiments that explore the possibility of wrinkling in biogels.
+We relate the excess area to the microstructural dynamics of the biogel.
+Usually the wrinkling wavelength is set by elasticity or gravity.
+In our case the wavelength is controlled by viscous flow, either in a lubrication layer or through the pores of the gel.
+Our results pave the way for exploring the possibility of wrinkling in confined porous soft materials immersed in a floating in a liquid such as biological tissues: during morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.
 

### 2015-02-27 fig_plis/article_SI.tex
corrige le significance statement "immersed in floating in a liquid"

@@
 In our case the wavelength is controlled by viscous flow, either in a lubrication layer or through the pores of the gel.
-Our results pave the way for exploring the possibility of wrinkling in confined porous soft materials immersed in a floating in a liquid such as biological tissues: during morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.
+Our results pave the way for exploring the possibility of wrinkling in porous soft materials floating in a liquid such as biological tissues: during morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.
 

### 2015-04-22 fig_plis/article_SI.tex
Restructure l'intro avec description de la cellule Annonce le nouveau plan Wrinkling experiment (à redispatcher)

@@
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
+The main goal of the present study is therefore threefold: to create a buoyancy-matched material that wrinkles, to unveil the physical origin of the observed patterns and to identify the selection mechanism of the wavelength.
 
 Here, we introduce model permeable biogels that produce hierarchical wrinkling as a result of their interplay between their gelation dynamics and the confinement conditions. 
-In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002} --this is a well-known effect used in yoghurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off, Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{vid:transmitted} and \ref{vid:reconstructed}.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
-We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
-The main goal of the present study is to unveil the physical origin of the observed patterns and to identify the selection mechanism of the wavelength.
+As a biogel, we use a simplified yoghurt obtained by the slow acidification of caseins, a family of milk proteins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+While still liquid, the casein dispersion is injected in a home made optical cell that is then sealed.
+This cell is designed such that the casein adhesion to both top and bottom walls is turned off and so that the gel is only anchored to the remaining four sides of the cell.
+
+Although no external stimulus is exerted on the cell, we observe the spontaneous formation of a wrinkling pattern, as shown in Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{vid:transmitted} and \ref{vid:reconstructed}.
+This pattern shows unique properties: a primary pattern appears simultaneously throughout the system with a wavelength $\lambda$  that is much smaller than the cell width $L$ and does not evolve with time.
+Furthermore, within this primary pattern, secondary and ternary patterns appear in cascade.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that the wrinkles at the millimetre scale result from the spontaneous shrinking and swelling of the casein network at the micron level upon acidification.
+
+Finally we show that the wavelength is selected by a purely viscous mechanism combining porous Darcy flow and viscous Poiseuille flow.
+We systematically checked that this combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength under a wide range of conditions.
 
@@
 	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w in water).
-(a) Light transmission microscopy.
+(a) XXX (b) Light transmission microscopy.
 Successive generations of patterns are highlighted in colour to stress the absence of coarsening after formation.
-(b) 3D reconstruction from fluorescent confocal microscopy.
-Scale bars are \SI{1}{\milli\metre}.}%
+(c) 3D reconstruction from fluorescent confocal microscopy.
+Scale bars are \SI{1}{\milli\metre}.
+XXX lambda, L, e XXX}%
 	\label{fig:dynamics}%
 \end{figure}
-
-In the following we first describe the physical chemistry that allows us to form such patterns.
-We then show how transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary patterns.
-Finally we model the situation and identify two possible selection mechanisms, namely porous Darcy flow or viscous Poiseuille flow.
-Comparing our experimental data to these two models we show that a combination of both Darcy and Poiseuille mechanisms accurately describes our data.
-This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
-%Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel film for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel film for short wavelengths.
 
@@
 $\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.
+
+\section*{Wrinkling experiment}
+
+\begin{figure*}
+\includegraphics{dilation.pdf}
+\caption{The wrinkling experiment.
+(a) Initial configuration where the sealed cell contains an homogeneous protein solution.
+(b) Around the isoelectric pH, the gel forms and immediately expels solvent leading to (c).
+If tensile stresses were not released we would observe reversible swelling back leading to (d) a flat swollen gel layer.
+If tensile stresses are released, swelling back leads either to buckling (e) or to wrinkling (f) depending whether the bending is free or hindered by a transverse load.}
+\label{fig:dilation}
+\end{figure*}
+
+The wrinkling patterns shown in Figure~\ref{fig:dynamics} are obtained in home made optical cells.
+A cell is composed of a slit and a cover slip  coated with polymer brushes and spaced by parafilm.
+The cell thickness $e$ is typically \SI{100}{\micro\metre} whereas other dimensions are $L$, typically \SI{1}{\centi\metre}.
+The cell is filled with the casein+GDL dispersion while it is fluid, and sealed with optical UV-glue, see Methods and Figure~\ref{fig:dilation}a.
+In such configuration, the adhesion to both top and bottom walls is turned off and therefore as the gel forms, it is only anchored to the remaining four sides of the cell (parafilm or cured glue).
+
+As the gel forms it expels the solvent and detaches from top and bottom wall, forming a thin film sandwiched between two solvent layers.
+Upon swelling back, the gel spontaneously wrinkles with a wavelength $\lambda\approx\SI{1}{\milli\metre}\ll L$, see first row of Figure~\ref{fig:dynamics}.
+As
+
+Let us examine possible scenarios upon shrinking and swelling of the gel.
+As the gel forms it expels the solvent and detaches from top and bottom wall, see Figure~\ref{fig:dilation}b and c.
+Then we consider three options.
+(i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat.
+Since this case is not observed it means that in this physical gel, syn\ae{}resis and swelling processes foster irreversible stress dissipation and structural changes.
+In the following, we will therefore consider that the shrunk state is stress-free.
+(ii) In Figure~\ref{fig:dilation}e, we assume that nothing hinders the buckling of the gel film.
+This simple buckling situation leads to a primary composed of a single bump of size $L$.
+As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength appear, as observed by~\cite{Roman1999}.
+That is not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
+(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom wall.
+To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load that hinders buckling on large wavelengths, as described for example by~\cite{Biot1957,Vella2009,Kolinski2009}.
 

### 2015-04-23 fig_plis/article_SI.tex
Caption of fig:dynamics L>1cm

@@
 	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w in water).
-(a) XXX (b) Light transmission microscopy.
+(a) Sketch of the cell where the adhesion on both top and bottom walls is turned off.
+The cell is sealed and the gel is only anchored to the four sides.
+Typical dimensions are $L>\SI{1}{\centi\metre}$, $e\approx\SI{100}{\micro\metre}$.
+(b) Light transmission microscopy.
 Successive generations of patterns are highlighted in colour to stress the absence of coarsening after formation.
-(c) 3D reconstruction from fluorescent confocal microscopy.
-Scale bars are \SI{1}{\milli\metre}.
-XXX lambda, L, e XXX}%
+The successive wavelengths are $\lambda=\SI{1.5}{\milli\metre}$ (yellow), $\lambda=\SI{0.75}{\milli\metre}$ (orange) and $\lambda=\SI{0.32}{\milli\metre}$ (brown).
+(c) 3D reconstruction from fluorescent confocal microscopy which highlight that patterns observed in (b) are wrinkles.
+The contrast in (b) is not due to thickness inhomogeneities but to altitude gradients as indicated by the yellow vertical lines in (a).
+Scale bars are \SI{1}{\milli\metre}.}%
 	\label{fig:dynamics}%

### 2015-04-23 fig_plis/article_SI.tex
Dynamic of the wrinkling process avant le système, avec le yaourt en intro et fig:dilation en conclusion

@@
 
-\section*{System design}
+\section*{Dynamics of the wrinkling process}
+
+To form a gel we start from an aqueous dispersion of caseins and GDL.
+GDL slowly acidify the medium and triggers gelation of the caseins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+While still liquid, this yoghurt-forming suspension is injected into an home-made optical cell, as sketched in Figure~\ref{fig:dynamics}a.
+The cell is sealed an no external stimulus is applied during gelation or pattern formation.
+Both top and bottom walls are coated by acrylamide brushes, effectively turning off the casein adhesion so that the gel is only anchored to the remaining four sides of the cell.
+
+The dynamics of the pattern formation is followed by light transmission microscopy and confocal microscopy, see Methods.
+As shown in Figure~\ref{fig:dynamics}b and Supplementary Video~\ref{vid:transmitted}, we observe the spontaneous formation of a first pattern that appear with a characteristic wavelength $\lambda\ll L$.
+The 2D shape and size of this primary pattern are set as soon as the pattern forms and do not evolve in time.
+However inside the primary pattern a secondary pattern emerges, inside which a ternary forms.
+At each step of this Russian dolls-like cascade, the wavelength is divided by two, as show in Supplementary Figure~\ref{fig:generations}.
+
+3D confocal microscopy, as shown in Figure~\ref{fig:dynamics}c and Supplementary Video~\ref{vid:reconstructed}, reveals that those patterns are actually wrinkles of a gel film.
+Figure~\ref{fig:sideview}a shows successive side views of the same confocal measurements as in  Figure~\ref{fig:dynamics}c see also Supplementary Videos~\ref{vid:cutDarcy} and \ref{vid:cutPoiseuille}.
+Initially the dispersion is homogeneous and fills the entire slit.
+The black margins in Figure~\ref{fig:sideview}a correspond to the glass walls.
+Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated --the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
+This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
+
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}c, and excess surface area of the gel film, Figure~\ref{fig:sideview}d, see Supplementary Methods.
+The evolution of the gel volume is surprisingly non-monotonous.
+After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
+Shortly after, at \SI{36}{\minute}, the planar film destabilises and starts to wrinkle in the $z$-direction creating excess area.
+This confirms that the instability leading to the primary pattern is due to swelling.
+Around \SI{38}{\minute}, the amplitude of the wrinkles reaches the thickness of the slit.
+As swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
+When the excess area saturates, the dynamics stop (\SI{3}{\hour}).
+We note that at any given point in time the thickness $h$ of the film of gel is spatially homogeneous.
+%The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
+
+%Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off.
+
+\begin{figure}
+	\includegraphics{sideview}
+	\caption{Measurements and modelling.
+(a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
+The scale bar is \SI{100}{\micro\metre} (real size ratio).
+(b) Schematic side view of the cell.
+The constant thickness gel film (orange) is surrounded by solvent.
+Yellow lines show the path of transmitted light through the gel phase.
+(c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
+Crosses correspond to times in (a).}
+	\label{fig:sideview}
+\end{figure}
+
+Based those observations, we examine possible scenarios upon shrinking and swelling of the gel.
+As the gel forms it expels the solvent and detaches from top and bottom wall, see Figure~\ref{fig:dilation}b and c.
+For the swelling back we consider three options:
+
+\noindent
+(i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat, as expected from chemical gels~\cite{Kim2010}.
+Since this case is not observed it means that in this physical gel, syn\ae{}resis and swelling processes foster irreversible stress dissipation and structural changes.
+In the following, we will therefore consider that the shrunk state is stress-free.
+
+\noindent
+(ii) In Figure~\ref{fig:dilation}e, we assume that nothing hinders the buckling of the gel film.
+This simple buckling situation leads to a primary pattern composed of a single bump of size $L$.
+As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}.
+That is not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
+
+\noindent
+(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom wall.
+To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load $\sigma_{\perp}$ that hinders buckling on large wavelengths.
+The transverse load could be gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, however we confirm that it plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
+with the weight acting longitudinally rather than transversely.
+Furthermore, the transverse load cannot be adhesion~\cite{Vella2009a} since the film is not adhering to the top and bottom walls.
+Influence of the side boundaries~\cite{Vandeparre2011a, Li2013} is also excluded since the wavelength of the pattern is uniform throughout the sample.
+Finally as the four side boundaries remain fixed $\sigma_{\perp}$ cannot come from the resistance to uniaxial stretching~\cite{Cerda2003}.
+
+At this point two points remain unclear: what triggers the the spontaneous syn\ae{}resis and swelling back processes?
+And what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
+
+\begin{figure*}
+\includegraphics{dilation.pdf}
+\caption{The wrinkling experiment.
+(a) Initial configuration where the sealed cell contains an homogeneous protein solution.
+(b) Around the isoelectric pH, the gel forms and immediately expels solvent leading to (c).
+If tensile stresses were not released we would observe reversible swelling back leading to (d) a flat swollen gel layer.
+If tensile stresses are released, swelling back leads either to buckling (e) or to wrinkling (f) depending whether the bending is free or hindered by a transverse load.}
+\label{fig:dilation}
+\end{figure*}
+
+\section*{Syn\ae{}resis and swelling back mechanism}
 
@@
 
-To form wrinkled patterns such as the one observed in Figure~\ref{fig:dynamics}, ones needs (i) to take advantage of the pH sensitivity of caseins, (ii) to control the route to gelation, (iii) to fine tune the gel properties and the boundary conditions.
+Syn\ae{}resis and swelling back mechanism depends on (i) the pH sensitivity of caseins, (ii) the route to gelation, (iii) the gel properties and (iv) the boundary conditions.
 
@@
 
-\section*{Wrinkling experiment}
-
-\begin{figure*}
-\includegraphics{dilation.pdf}
-\caption{The wrinkling experiment.
-(a) Initial configuration where the sealed cell contains an homogeneous protein solution.
-(b) Around the isoelectric pH, the gel forms and immediately expels solvent leading to (c).
-If tensile stresses were not released we would observe reversible swelling back leading to (d) a flat swollen gel layer.
-If tensile stresses are released, swelling back leads either to buckling (e) or to wrinkling (f) depending whether the bending is free or hindered by a transverse load.}
-\label{fig:dilation}
-\end{figure*}
-
-The wrinkling patterns shown in Figure~\ref{fig:dynamics} are obtained in home made optical cells.
-A cell is composed of a slit and a cover slip  coated with polymer brushes and spaced by parafilm.
-The cell thickness $e$ is typically \SI{100}{\micro\metre} whereas other dimensions are $L$, typically \SI{1}{\centi\metre}.
-The cell is filled with the casein+GDL dispersion while it is fluid, and sealed with optical UV-glue, see Methods and Figure~\ref{fig:dilation}a.
-In such configuration, the adhesion to both top and bottom walls is turned off and therefore as the gel forms, it is only anchored to the remaining four sides of the cell (parafilm or cured glue).
-
-As the gel forms it expels the solvent and detaches from top and bottom wall, forming a thin film sandwiched between two solvent layers.
-Upon swelling back, the gel spontaneously wrinkles with a wavelength $\lambda\approx\SI{1}{\milli\metre}\ll L$, see first row of Figure~\ref{fig:dynamics}.
-As
-
-Let us examine possible scenarios upon shrinking and swelling of the gel.
-As the gel forms it expels the solvent and detaches from top and bottom wall, see Figure~\ref{fig:dilation}b and c.
-Then we consider three options.
-(i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat.
-Since this case is not observed it means that in this physical gel, syn\ae{}resis and swelling processes foster irreversible stress dissipation and structural changes.
-In the following, we will therefore consider that the shrunk state is stress-free.
-(ii) In Figure~\ref{fig:dilation}e, we assume that nothing hinders the buckling of the gel film.
-This simple buckling situation leads to a primary composed of a single bump of size $L$.
-As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength appear, as observed by~\cite{Roman1999}.
-That is not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
-(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom wall.
-To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load that hinders buckling on large wavelengths, as described for example by~\cite{Biot1957,Vella2009,Kolinski2009}.
-
 Finally, if the adhesion to both top and bottom walls is turned off, the film wrinkles as shown in Figure~\ref{fig:dynamics}.
@@
 In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{fig:nonmonotonic}).
-
-\section*{Dynamics of the wrinkling process}
-
-Figure~\ref{fig:sideview}a shows successive side views of confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration that induces wrinkling, see also Supplementary Videos~\ref{vid:cutDarcy} and \ref{vid:cutPoiseuille}.
-Initially the dispersion is homogeneous and fills the entire slit.
-The black margins correspond to the glass walls.
-Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated --the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
-This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
-We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$,  Figure~\ref{fig:sideview}b.
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}c, and excess surface area of the gel film, Figure~\ref{fig:sideview}d, see Supplementary Methods.
-As expected from the system design, the evolution of the gel volume is non-monotonous.
-After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
-Shortly after,  at \SI{36}{\minute}, the planar film destabilises and starts to wrinkle in the $z$-direction creating excess area.
-This confirms that the instability is due to swelling.
-Around \SI{38}{\minute}, the amplitude of the wrinkles reaches the thickness of the slit.
-The size of the primary pattern is now fixed, Supplementary Figure~\ref{fig:growth}.
-However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
-When the excess area saturates, the dynamics stop (\SI{3}{\hour}).
-The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
-
-Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off.
-We also confirm that at any given point in time the thickness $h$ of the film of gel is spatially homogeneous.
-The contrast in transmission microscopy, Figure~\ref{fig:dynamics}a, is therefore not due to thickness inhomogeneities but to altitude gradients.
-Indeed the length of the optical path along the vertical direction $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Figure~\ref{fig:sideview}b.
-
-\begin{figure}
-	\includegraphics{sideview}
-	\caption{Measurements and modelling.
-(a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
-The scale bar is \SI{100}{\micro\metre} (real size ratio).
-(b) Schematic side view of the cell.
-The constant thickness gel film (orange) is surrounded by solvent.
-Yellow lines show the path of transmitted light through the gel phase.
-(c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
-Crosses correspond to times in (a).}
-	\label{fig:sideview}
-\end{figure}
 

### 2015-04-23 fig_plis/article_SI.tex
Conclusion de Syn\ae{}resis and swelling back mechanism Already discared

@@
 
-%XXXCe paragraphe fait-il sens ici ?XXX As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
-This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off.
-To quantify this effect, we make the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}.
-In the Supplementary Figure~\ref{fig:permeability}, we observe that the aqueous buffer goes through the gel and rises in the tube due to capillary forces.
-Exploitation of this experiment give the permeability of the gel: \SI{67000}{\square\nano\metre}.
-
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
@@
 In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d.
-$\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.
+$\xi$ remains mostly constant and variations manifests themselves at the macroscopic level, on the volume occupied by the gel.
 
-Finally, if the adhesion to both top and bottom walls is turned off, the film wrinkles as shown in Figure~\ref{fig:dynamics}.
-Such cascade buckling has already been observed in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates~\cite{Roman1999}.
-In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Figure~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.
+Therefore, the syn\ae{}resis and swelling back mechanism is related to charge stabilisation of the casein proteins with the pH.
+Such mechanism is only effective to produce wrinkle when adhesion of the caseins to both top and bottom walls is turned off.
 
-%We coat the glass slides with acrylamide brushes to prevent the adhesion of the casein.
-In fig.2c, when the top slide of the optical cell is coated with acrylamide brushes, we then observe that the casein network detaches from the top slide.
-$\xi$ does not peak at the isoelectric point but remains constant
-
-%Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-Once the space spanning network is formed and sticks to the container, protein attraction can only proceed by local shrinking of the protein agregates, leading to larger pores.
-
-%The difference in value of $\xi$ between the fully un-coated  and the half-coated optical cell demonstrate that the gel is subject to constrains as the pH decreases.
-The relaxation of thoses constrain is key to the wrinkling formation when both the top and bottom slide are coated with acrylamide brushes.
-
-%on a change les condition daccroche ce aui permet au contrainte de se relacher
-
-%The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation~\cite{Bremer1989} (Fig.~\ref{fig:acidification}a).
-Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
-Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplementary Fig.~\ref{fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
-In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{fig:nonmonotonic}).
-
-\section*{Models and scaling laws}
+\section*{Wavelength selection mechanism}
 
@@
 In our situation we now have to determine $\sigma_\perp$.
-We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding solvent layers.
-We confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
-with the weight acting longitudinally rather than transversely.
+We have already excluded uniaxial stretching~\cite{Cerda2003}, gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, boundaries~\cite{Vandeparre2011a, Li2013} and adhesion~\cite{Vella2009a}.
+Two possible candidates remains for the transverse load $\sigma_\perp$: the porous flow through the gel or the viscous flows in the surrounding solvent layers.
 
@@
 
-Therefore we consider two limit scenarios that leads to wrinkling patterns.
+Accordingly we consider two limit scenarios that leads to wrinkling patterns.
 In a first limit, the film of gel sits without sticking to the bottom wall of the cell, i.e.

### 2015-04-23 fig_plis/article_SI.tex
Commence à factoriser la discussion à propos de tau

@@
 \begin{equation}
-\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4}
+\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4},\quad\text{with } \Upsilon = \frac{B\tau}{\eta}.
 \end{equation}
-where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
-This expression emphasises the kinetic nature of the wavelength selection through Darcy flow.
-However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}.
-This arrest may be due to bending rigidity~\cite{LeGoff2014}.
-Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate, see Supplementary Methods for the precise definition.
+$\Upsilon$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
 
@@
 Indeed within the lubrication approximation or at low Reynolds numbers one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
+
+The characteristic time $\tau$ is the hallmark of the kinetic nature of the wavelength selection mechanisms and is present in the expressions of both $\lambda_D$ and $\lambda_D$.
+
+This expression emphasises the kinetic nature of the wavelength selection through Darcy flow.
+
+If the wrinkle amplitude is free, one thus expects coarsening of the pattern~\cite{Yoo2003}, i.e.
+an increase of the wavelength.
+We observe a slight increase of the wavelength between its birth and the time it touches both walls, see Supplementary Figure~\ref{fig:growth}.
+
+However we note that the wavelength does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}.
+
+However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}.
+This arrest may be due to bending rigidity~\cite{LeGoff2014}.
+Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate, see Supplementary Methods for the precise definition.
 

### 2015-04-23 fig_plis/article_SI.tex
Measure de tau lors de la description de la dynamique

@@
 
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}c, and excess surface area of the gel film, Figure~\ref{fig:sideview}d, see Supplementary Methods.
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}b, and excess surface area of the gel film, Figure~\ref{fig:sideview}c, see Supplementary Methods.
 The evolution of the gel volume is surprisingly non-monotonous.
@@
 We note that at any given point in time the thickness $h$ of the film of gel is spatially homogeneous.
-%The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
 
-%Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off.
+The pattern formation is dynamic in essence.
+In Figure XXX, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Methods for the definition.
+We observe that the velocity peaks at different times corresponding to the syn\ae{}resis and the respective growths of the primary, secondary and tertiary patterns.
+We associate each growth with a characteristic time $\tau = e/v_\text{max}$.
+For example, in the present case the $\tau$ corresponding to the primary pattern is $\SI{123}{\second}$.
 
@@
 Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Figure~\ref{fig:generations}.
+
+XXX The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
 

### 2015-04-24 fig_plis/article_SI.tex
Caption de fig:sketchDP

@@
 \section*{Wavelength selection mechanism}
+
+\begin{figure}
+\includegraphics{sketch_Darcy_Poiseuille.pdf}
+\caption{Sketch of the two limit scenarios for the wavelength selection.
+In both cases the gel film of thickness $h$ is destabilised by the excess area a give rise to wrinkles of amplitude $A(t)$ and wavelength $\lambda$.
+The value of $\lambda$ is set by the interplay between the bending rigidity $B$ and a transverse load $\sigma_\perp$ due to a pressure gradient $p_2-p_1$ in the solvent of viscosity $\eta$.
+(a) Darcy scenario: the gel initially sits without sticking to the bottom wall.
+$\sigma_\perp$ is due to the flow of the solvent through the porous gel of permeability $\alpha$ to fill the growing blister.
+(b) Poiseuille scenario: the gel film of negligible porosity is initially in the middle of the cell, separated by a distance $H$ from each wall.
+$\sigma_\perp$ is due to the lubrication flow in the top and bottom solvent layers.}
+\label{fig:sketchDP}
+\end{figure}
 

### 2015-04-24 fig_plis/article_SI.tex
Update caption of fig:sideview

@@
 The pattern formation is dynamic in essence.
-In Figure XXX, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Methods for the definition.
+In Figure~\ref{fig:sideview}d, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Methods for the definition.
 We observe that the velocity peaks at different times corresponding to the syn\ae{}resis and the respective growths of the primary, secondary and tertiary patterns.
@@
 	\includegraphics{sideview}
-	\caption{Measurements and modelling.
+	\caption{3D analysis of the wrinkling process.
 (a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
 The scale bar is \SI{100}{\micro\metre} (real size ratio).
-(b) Schematic side view of the cell.
-The constant thickness gel film (orange) is surrounded by solvent.
-Yellow lines show the path of transmitted light through the gel phase.
-(c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
+(b-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume, the excess area and the velocity along the $z$-direction.
 Crosses correspond to times in (a).}

### 2015-04-24 fig_plis/article_SI.tex
Réinsert l'infuence de la quantité de GDL Mesures de lambda et des paramèetre pour calculer lambda Usage de la croissance du patern secondaire Description de la fig:DarcyPoiseuille equation de la mesure de vitesse en supplementary

@@
 
-At this point two points remain unclear: what triggers the the spontaneous syn\ae{}resis and swelling back processes?
+Now, two points remain unclear: what triggers the the spontaneous syn\ae{}resis and swelling back processes?
 And what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
@@
 Such mechanism is only effective to produce wrinkle when adhesion of the caseins to both top and bottom walls is turned off.
+Coherently with our explanation, the amount of GDL, and thus the final pH reached by the suspension, controls the amount of swelling back.
+Therefore at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
 
@@
 \begin{align}
-\lambda_D &= 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}\\
-\lambda_P &= \pi (hH)^{1/2}\left(\frac{2}{9}\Upsilon\right)^{1/6}
+\lambda_D &= 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}\label{eq:prefDarcy}\\
+\lambda_P &= \pi (hH)^{1/2}\left(\frac{2}{9}\Upsilon\right)^{1/6}\label{eq:prefPoiseuille}
 \end{align}
@@
 Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly.
-However we are able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
+
+For each experiment we measure independently $\lambda$ and all the parameter that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}).
+$\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Methods and Supplementary methods.
+
+The beauty of this approach is that it is not only valid to model the primary wavelength, but that it describes also the growth of the secondary pattern.
+Although the wavelength $\lambda_2$ of the secondary pattern is fully determined by the primary pattern $\lambda_2 = \lambda_1/2$, Figure~\ref{fig:dynamics} and Supplementary Figure~\ref{fig:generations}, the secondary pattern has its own dynamics with its own characteristic time $\tau_2\neq\tau_1$.
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
-Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Figure~\ref{fig:generations}.
+Therefore, we are able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
 
-XXX The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
-
-Our results are displayed in Figure~\ref{fig:DarcyPoiseuille}.
-Although they provide the correct order of magnitude for $\lambda$, neither model alone is able to quantitatively reproduce the full data set.
-However considering only the points where $H<H^*$ (respectively $H>H^*$), the Darcy (resp.
-Poiseuille) model fits the data well albeit with a prefactor.
-Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b).
-Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation provides a much better agreement for the whole $H/H^*$ range.
+Despite changing the gel properties and the cell geometry we always observe wrinkling proving the robustness of this approach to form wrinkling patterns.
+We are in the position to vary $H/H^\star$ from 0.1 to 10 so that we can test the crossover between Darcy and Poiseuille mechanism.
+In Figure~\ref{fig:DarcyPoiseuille}a and b we have plotted the measure wavelength as a function of the model prediction $\lambda_D$ and $\lambda_P$ respectively.
+If the prediction are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$.
+Indeed, both model seems to apply in their domain of validity albeit with a prefactor.
+The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data.
+To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Methods.
+The results of this new model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ for the whole $H/H^*$ range.
 
@@
 However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
-Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
+Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view:
+\begin{equation}
+v(t) = \frac{\int \left|\frac{dz}{dt}\right| dxdy}{\int dxdy},
+\end{equation}
+where $z(x,y,t)$ is the altitude of the mid surface of the gel.
 
-We note that when both measurements were available (as in Supplementary Figure~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
+We note that when both measurements were available (as in Figure~\ref{fig:sideview}d of the main text and inset of Supplementary Figure~\ref{fig:growth}), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
 We use this approximation of $\tau$ for all our samples.
-This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact, see Supplementary Figure~\ref{fig:growth}b, thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
+This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact, see Supplementary Figure~\ref{fig:growth}, thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
 

### 2015-04-24 fig_plis/article_SI.tex
Insert de fig:growth remplace le (a) précédant

@@
 In Figure~\ref{fig:DarcyPoiseuille}a and b we have plotted the measure wavelength as a function of the model prediction $\lambda_D$ and $\lambda_P$ respectively.
-If the prediction are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$.
+If the predictions are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$.
 Indeed, both model seems to apply in their domain of validity albeit with a prefactor.
@@
 To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Methods.
-The results of this new model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ for the whole $H/H^*$ range.
+The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ for the whole $H/H^*$ range.
 
@@
 	\caption{Growth of the instability.
-(a) Evolution of the gel velocity $v$ averaged over the whole surface.
-Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scales.
+(a) Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scales.
 The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
 (b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
-Note that the longitudinal size of the blister evolves in this regime.
+The longitudinal size of the blister evolves in this regime.
 The thick black curve is the profile at much longer times and is at lower altitude due to swelling.
-Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
+The blister foot, highlighted by the orange vertical line, does not move after contact.}
 	\label{fig:growth}

### 2015-04-24 fig_plis/article_SI.tex
Lambda constant from birth to saturation. Lambda constant after saturation due to bending rigidity

@@
 
-The pattern formation is dynamic in essence.
+The primary, secondary and tertiary patterns follow the same path: the birth of an instability of wavelength $\lambda$ of growing amplitude until saturation.
+Throughout the whole process, $\lambda$ remains constant.
+This path is dynamic in essence.
 In Figure~\ref{fig:sideview}d, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Methods for the definition.
@@
 For example, in the present case the $\tau$ corresponding to the primary pattern is $\SI{123}{\second}$.
+
+We do not observe coarsening of the patterns as in~\cite{Yoo2003}.
+We believe the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}.
+Indeed when the amplitude saturates the gel flattens against the walls and create sharp edges.
+To accommodate those sharp edges with finite bending rigidity, the flat part of the gel actually ripples, see Supplementary Figure~\ref{fig:growth}b.
+The ripples created by opposite sides of a same flat patch interfere, creating barriers against coarsening.
 
@@
 
-The characteristic time $\tau$ is the hallmark of the kinetic nature of the wavelength selection mechanisms and is present in the expressions of both $\lambda_D$ and $\lambda_D$.
-
-This expression emphasises the kinetic nature of the wavelength selection through Darcy flow.
-
-If the wrinkle amplitude is free, one thus expects coarsening of the pattern~\cite{Yoo2003}, i.e.
-an increase of the wavelength.
-We observe a slight increase of the wavelength between its birth and the time it touches both walls, see Supplementary Figure~\ref{fig:growth}.
-
-However we note that the wavelength does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}.
-
-However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{fig:growth}.
-This arrest may be due to bending rigidity~\cite{LeGoff2014}.
-Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate, see Supplementary Methods for the precise definition.
-
 A more complete derivation of the models can be found in the Supplementary Methods and yields the following prefactors:
@@
 where $H$ is redefined by $1/H^3 = 1/H_1^3 + 1/H_2^3$. The crossover between the two models where $\lambda_D=\lambda_P$ is found for $H = H^* \equiv 2^{2/3} 3^{1/6} \alpha^{1/2} \Upsilon^{1/6}$.
+
+The characteristic time $\tau$ is the hallmark of the kinetic nature of the wavelength selection mechanisms and is present in the expressions of both $\lambda_D$ and $\lambda_P$.
 

### 2015-04-24 fig_plis/article_SI.tex
fig:ripples plus que fig:growth

@@
 
-The primary, secondary and tertiary patterns follow the same path: the birth of an instability of wavelength $\lambda$ of growing amplitude until saturation.
+The primary, secondary and tertiary patterns follow the same path: the birth of an instability of wavelength $\lambda$ of growing amplitude $A(t)$ until saturation, see Figure~\ref{fig:sideview}a.
 Throughout the whole process, $\lambda$ remains constant.
@@
 We observe that the velocity peaks at different times corresponding to the syn\ae{}resis and the respective growths of the primary, secondary and tertiary patterns.
-We associate each growth with a characteristic time $\tau = e/v_\text{max}$.
+We associate each growth with a characteristic time $\tau = A_\text{max}/v_\text{max}$.
 For example, in the present case the $\tau$ corresponding to the primary pattern is $\SI{123}{\second}$.
@@
 Indeed when the amplitude saturates the gel flattens against the walls and create sharp edges.
-To accommodate those sharp edges with finite bending rigidity, the flat part of the gel actually ripples, see Supplementary Figure~\ref{fig:growth}b.
+To accommodate those sharp edges with finite bending rigidity, the flat part of the gel actually ripples, see Supplementary Figure~\ref{fig:ripples}.
 The ripples created by opposite sides of a same flat patch interfere, creating barriers against coarsening.
@@
 
-\section*{Microstructure measurements}
+\subsection*{Microstructure measurements}
 
@@
 
-\section*{Characteristic time measurements}
-\begin{figure}
-	\includegraphics{growth}
-	\caption{Growth of the instability.
-(a) Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scales.
-The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
-(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
-The longitudinal size of the blister evolves in this regime.
-The thick black curve is the profile at much longer times and is at lower altitude due to swelling.
-The blister foot, highlighted by the orange vertical line, does not move after contact.}
-	\label{fig:growth}
-\end{figure}
+\subsection*{Characteristic time measurements}
 
@@
 where $z(x,y,t)$ is the altitude of the mid surface of the gel.
-
-We note that when both measurements were available (as in Figure~\ref{fig:sideview}d of the main text and inset of Supplementary Figure~\ref{fig:growth}), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
-We use this approximation of $\tau$ for all our samples.
-This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact, see Supplementary Figure~\ref{fig:growth}, thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
 
@@
 
-\section*{Wavelength measurements}
+\subsection*{Wavelength measurements}
 
@@
 	\label{fig:generations}%
+\end{figure}
+
+\begin{figure}
+\includegraphics{ripples.pdf}
+\caption{Evidence of the ripples generated by the junction to flat patches, see~\cite{LeGoff2014}.
+Position of the bottom surface of the gel in (a) 2D and (b) 1D.
+Only first ripple is visible, highlighted by orange rectangles.
+(c) Schematic view of the interference between ripples.}
+\label{fig:ripples}
 \end{figure}

### 2015-04-27 fig_plis/article_SI.tex
Déplace Wavelength measurements au début des SI
Réorgnise les SI Typos et bémols

@@
 
-To form a gel we start from an aqueous dispersion of caseins and GDL.
+To form a gel we start from an aqueous dispersion of caseins and glucono-$\delta$-lactone (GDL).
 GDL slowly acidify the medium and triggers gelation of the caseins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
 While still liquid, this yoghurt-forming suspension is injected into an home-made optical cell, as sketched in Figure~\ref{fig:dynamics}a.
-The cell is sealed an no external stimulus is applied during gelation or pattern formation.
+The cell is sealed and no external stimulus is applied during gelation or pattern formation.
 Both top and bottom walls are coated by acrylamide brushes, effectively turning off the casein adhesion so that the gel is only anchored to the remaining four sides of the cell.
@@
 
-We do not observe coarsening of the patterns as in~\cite{Yoo2003}.
+XXX formulation XXX We do not observe coarsening of the patterns as in~\cite{Yoo2003}.
 We believe the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}.
@@
 
-Based those observations, we examine possible scenarios upon shrinking and swelling of the gel.
+Based on those observations, we examine possible scenarios upon shrinking and swelling of the gel.
 As the gel forms it expels the solvent and detaches from top and bottom wall, see Figure~\ref{fig:dilation}b and c.
@@
 As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}.
-That is not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
+That is also not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
 
 \noindent
-(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom wall.
+(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom walls.
 To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load $\sigma_{\perp}$ that hinders buckling on large wavelengths.
@@
 Finally as the four side boundaries remain fixed $\sigma_{\perp}$ cannot come from the resistance to uniaxial stretching~\cite{Cerda2003}.
+XXX conclure plus clairement XXX
 
@@
 
+XXX repetition entre paragraphe d'avant et d'après XXX
+
 In our situation we now have to determine $\sigma_\perp$.
@@
 
-For each experiment we measure independently $\lambda$ and all the parameter that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}).
+For each experiment we measure independently $\lambda$ and all the parameters that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}).
 $\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Methods and Supplementary methods.
@@
 If the predictions are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$.
-Indeed, both model seems to apply in their domain of validity albeit with a prefactor.
+Indeed, both model seems to apply in their domain of validity albeit with a prefactor XXX Thomas dit pas clair XXX.
 The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data.
@@
 Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins.
-As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
+As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
+XXX texture food XXX
 
@@
 \section*{Supplementary material}
+
+\subsection*{Wavelength measurements}
+
+\begin{figure}
+	\includegraphics{generations}%
+	\caption{Fourier spectra of binary altitudes of Main Figure~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
+The scale bar is \SI{1}{\milli\metre}.}%
+	\label{fig:generations}%
+\end{figure}
+
+\begin{figure}
+\includegraphics{ripples.pdf}
+\caption{Evidence of the ripples generated by the junction to flat patches, see~\cite{LeGoff2014}.
+Position of the bottom surface of the gel in (a) 2D and (b) 1D.
+Only first ripple is visible, highlighted by orange rectangles.
+(c) Schematic view of the interference between ripples.}
+\label{fig:ripples}
+\end{figure}
+
+\begin{figure*}
+	\includegraphics{patterns}
+	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data}.
+(a-d) increasing cell thickness.
+(e-f) larger solid content.
+(g-i) increasing glycerol content.
+All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
+The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
+Arrows show the measured primary wavelength.}
+	\label{fig:patterns}
+\end{figure*}
+
+To measure the wrinkle wavelength $\lambda$, we use large field pictures of the final state.
+These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e.
+The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
+White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
+It is thus possible to reconstruct a binary picture of the pattern, inset of Supplementary Figure~\ref{fig:generations}.
+Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final binary picture to the configuration of the primary pattern.
+
+Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
+When only the first generation pattern is present a peak clearly indicates the dominant mode.
+When the secondary pattern has appeared, the dominant wavelength is halved and the peak much weaker.
+This allows us to use the secondary pattern is our analysis by setting their wavelength to half of the primary wavelength.
+Further generations do not seem to appear simultaneously and the corresponding peaks are weak and broad.
+
+\subsection*{Confocal images processing}
+
+At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
+Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
+At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.
+
+We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
+Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
+We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line.
+Since this criterion is local, we do not need to correct for background intensity variation.
+In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.
+
+From these data, we obtain for each stack the volume and thickness of the gel phase, area and peak-to-peak amplitude of the mid surface.
+Velocity is obtained by time differentiation.
+Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
+
+\subsection*{Characteristic time measurements}
+
+In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
+However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
+Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view:
+\begin{equation}
+v(t) = \frac{\int \left|\frac{dz}{dt}\right| dxdy}{\int dxdy},
+\end{equation}
+where $z(x,y,t)$ is the altitude of the mid surface of the gel.
 
@@
 Obtaining $z_\infty$ without tube 2 would be time prohibitive.
-
-\subsection*{Confocal images processing}
-
-At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
-Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
-At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.
-
-We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
-Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
-We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line.
-Since this criterion is local, we do not need to correct for background intensity variation.
-In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.
-
-From these data, we obtain for each stack the volume and thickness of the gel phase, area and peak-to-peak amplitude of the mid surface.
-Velocity is obtained by time differentiation.
-Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
-
-\subsection*{Characteristic time measurements}
-
-In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
-However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
-Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view:
-\begin{equation}
-v(t) = \frac{\int \left|\frac{dz}{dt}\right| dxdy}{\int dxdy},
-\end{equation}
-where $z(x,y,t)$ is the altitude of the mid surface of the gel.
 
@@
 
-\subsection*{Wavelength measurements}
-
-\begin{figure}
-	\includegraphics{generations}%
-	\caption{Fourier spectra of binary altitudes of Main Figure~\ref{fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-The scale bar is \SI{1}{\milli\metre}.}%
-	\label{fig:generations}%
-\end{figure}
-
-\begin{figure}
-\includegraphics{ripples.pdf}
-\caption{Evidence of the ripples generated by the junction to flat patches, see~\cite{LeGoff2014}.
-Position of the bottom surface of the gel in (a) 2D and (b) 1D.
-Only first ripple is visible, highlighted by orange rectangles.
-(c) Schematic view of the interference between ripples.}
-\label{fig:ripples}
-\end{figure}
-
-\begin{figure*}
-	\includegraphics{patterns}
-	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data}.
-(a-d) increasing cell thickness.
-(e-f) larger solid content.
-(g-i) increasing glycerol content.
-All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
-The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
-Arrows show the measured primary wavelength.}
-	\label{fig:patterns}
-\end{figure*}
-
-To measure the wrinkle wavelength $\lambda$, we use large field pictures of the final state.
-These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e.
-The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
-White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
-It is thus possible to reconstruct a binary picture of the pattern, inset of Supplementary Figure~\ref{fig:generations}.
-Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final binary picture to the configuration of the primary pattern.
-
-Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
-When only the first generation pattern is present a peak clearly indicates the dominant mode.
-When the secondary pattern has appeared, the dominant wavelength is halved and the peak much weaker.
-This allows us to use the secondary pattern is our analysis by setting their wavelength to half of the primary wavelength.
-Further generations do not seem to appear simultaneously and the corresponding peaks are weak and broad.
-
 \subsection*{Supplementary Videos}

### 2015-05-12 fig_plis/article_SI.tex
Corrections de Seb. Paraffine dans les méthodes uniquement. Dans les schemas Darcy et Poiseuille seule une longueur d'onde est représentée

@@
 As a biogel, we use a simplified yoghurt obtained by the slow acidification of caseins, a family of milk proteins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
-While still liquid, the casein dispersion is injected in a home made optical cell that is then sealed.
-This cell is designed such that the casein adhesion to both top and bottom walls is turned off and so that the gel is only anchored to the remaining four sides of the cell.
+While still liquid, the casein dispersion is injected in a home-made optical cell that is then sealed.
+This cell is designed such that the casein adhesion to both top and bottom walls is turned off and so that the gel is only anchored to the remaining four vertical sides of the cell.
 
-Although no external stimulus is exerted on the cell, we observe the spontaneous formation of a wrinkling pattern, as shown in Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{vid:transmitted} and \ref{vid:reconstructed}.
+Although no external stimulus is exerted on the system, we observe the spontaneous formation of a wrinkling pattern, as shown in Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{vid:transmitted} and \ref{vid:reconstructed}.
 This pattern shows unique properties: a primary pattern appears simultaneously throughout the system with a wavelength $\lambda$  that is much smaller than the cell width $L$ and does not evolve with time.
 Furthermore, within this primary pattern, secondary and ternary patterns appear in cascade.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that the wrinkles at the millimetre scale result from the spontaneous shrinking and swelling of the casein network at the micron level upon acidification.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that the wrinkles at the millimetre scale result from the spontaneous swelling of the casein network at the micron level upon acidification.
 
@@
 The successive wavelengths are $\lambda=\SI{1.5}{\milli\metre}$ (yellow), $\lambda=\SI{0.75}{\milli\metre}$ (orange) and $\lambda=\SI{0.32}{\milli\metre}$ (brown).
-(c) 3D reconstruction from fluorescent confocal microscopy which highlight that patterns observed in (b) are wrinkles.
+(c) 3D reconstruction from fluorescent confocal microscopy which highlights that patterns observed in (b) correspond to wrinkles.
 The contrast in (b) is not due to thickness inhomogeneities but to altitude gradients as indicated by the yellow vertical lines in (a).
@@
 
-To form a gel we start from an aqueous dispersion of caseins and glucono-$\delta$-lactone (GDL).
-GDL slowly acidify the medium and triggers gelation of the caseins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
-While still liquid, this yoghurt-forming suspension is injected into an home-made optical cell, as sketched in Figure~\ref{fig:dynamics}a.
+To form a biogel we start from an aqueous dispersion of caseins and glucono-$\delta$-lactone (GDL).
+GDL slowly acidifies the medium and triggers gelation of the caseins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+While still liquid, this yoghurt-forming suspension is injected into a home-made optical cell, as sketched in Figure~\ref{fig:dynamics}a.
 The cell is sealed and no external stimulus is applied during gelation or pattern formation.
-Both top and bottom walls are coated by acrylamide brushes, effectively turning off the casein adhesion so that the gel is only anchored to the remaining four sides of the cell.
+Both top and bottom walls are coated with acrylamide brushes, effectively turning off the casein adhesion, so that the gel is only anchored to the remaining four vertical sides of the cell.
 
-The dynamics of the pattern formation is followed by light transmission microscopy and confocal microscopy, see Methods.
-As shown in Figure~\ref{fig:dynamics}b and Supplementary Video~\ref{vid:transmitted}, we observe the spontaneous formation of a first pattern that appear with a characteristic wavelength $\lambda\ll L$.
+The dynamics of the pattern formation are followed by light transmission microscopy and confocal microscopy, see Methods.
+As shown in Figure~\ref{fig:dynamics}b and Supplementary Video~\ref{vid:transmitted}, we observe the spontaneous formation of a first pattern that appears with a characteristic wavelength $\lambda\ll L$.
 The 2D shape and size of this primary pattern are set as soon as the pattern forms and do not evolve in time.
 However inside the primary pattern a secondary pattern emerges, inside which a ternary forms.
-At each step of this Russian dolls-like cascade, the wavelength is divided by two, as show in Supplementary Figure~\ref{fig:generations}.
+At each step of this Russian dolls-like cascade, the wavelength is divided by two, as shown in Supplementary Figure~\ref{fig:generations}.
 
-3D confocal microscopy, as shown in Figure~\ref{fig:dynamics}c and Supplementary Video~\ref{vid:reconstructed}, reveals that those patterns are actually wrinkles of a gel film.
-Figure~\ref{fig:sideview}a shows successive side views of the same confocal measurements as in  Figure~\ref{fig:dynamics}c see also Supplementary Videos~\ref{vid:cutDarcy} and \ref{vid:cutPoiseuille}.
+3D confocal microscopy, as shown in Figure~\ref{fig:dynamics}c and Supplementary Video~\ref{vid:reconstructed}, reveals that those patterns actually correspond to wrinkles of the biogel film.
+Figure~\ref{fig:sideview}a shows successive side views of the same confocal measurements as in  Figure~\ref{fig:dynamics}c, see also Supplementary Videos~\ref{vid:cutDarcy} and \ref{vid:cutPoiseuille}.
 Initially the dispersion is homogeneous and fills the entire slit.
@@
 Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated --the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
-This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
+This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, solvent is expelled through both the upper and lower sides of the gel film.
 
@@
 
-The primary, secondary and tertiary patterns follow the same path: the birth of an instability of wavelength $\lambda$ of growing amplitude $A(t)$ until saturation, see Figure~\ref{fig:sideview}a.
+The primary, secondary and tertiary patterns follow the same path: the birth of an instability of wavelength $\lambda$ of growing amplitude $A(t)$ until saturation to $A_\text{max}$, see Figure~\ref{fig:sideview}a.
 Throughout the whole process, $\lambda$ remains constant.
@@
 In Figure~\ref{fig:sideview}d, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Methods for the definition.
-We observe that the velocity peaks at different times corresponding to the syn\ae{}resis and the respective growths of the primary, secondary and tertiary patterns.
+We observe that the velocity peaks at different times corresponding to syn\ae{}resis and to the respective growths of the primary, secondary and tertiary patterns.
 We associate each growth with a characteristic time $\tau = A_\text{max}/v_\text{max}$.
@@
 
-XXX formulation XXX We do not observe coarsening of the patterns as in~\cite{Yoo2003}.
+Therefore contrary to observations in, e.g., Ref~\cite{Yoo2003}, we do not report any coarsening of the patterns.
 We believe the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}.
-Indeed when the amplitude saturates the gel flattens against the walls and create sharp edges.
+Indeed when the amplitude saturates the gel flattens against the walls which creates sharp edges.
 To accommodate those sharp edges with finite bending rigidity, the flat part of the gel actually ripples, see Supplementary Figure~\ref{fig:ripples}.
-The ripples created by opposite sides of a same flat patch interfere, creating barriers against coarsening.
+The ripples created by opposite sides of a same flat patch interfere, creating barriers against coarsening~\cite{LeGoff2014}.
 
@@
 The scale bar is \SI{100}{\micro\metre} (real size ratio).
-(b-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume, the excess area and the velocity along the $z$-direction.
+(b-d) Confocal microscopy measurement of the evolution of the volume of the gel phase relative to cell volume, of the excess area and of the velocity along the $z$-direction.
 Crosses correspond to times in (a).}
@@
 \noindent
-(i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat, as expected from chemical gels~\cite{Kim2010}.
+(i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat, as expected for chemical gels~\cite{Kim2010}.
 Since this case is not observed it means that in this physical gel, syn\ae{}resis and swelling processes foster irreversible stress dissipation and structural changes.
@@
 As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}.
-That is also not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
+This is also not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
 
@@
 To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load $\sigma_{\perp}$ that hinders buckling on large wavelengths.
-The transverse load could be gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, however we confirm that it plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
+The transverse load could be due to gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, however we confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
 with the weight acting longitudinally rather than transversely.
 Furthermore, the transverse load cannot be adhesion~\cite{Vella2009a} since the film is not adhering to the top and bottom walls.
-Influence of the side boundaries~\cite{Vandeparre2011a, Li2013} is also excluded since the wavelength of the pattern is uniform throughout the sample.
+%
+The binding of a longitudinal edge to a rigid substrate can induce wrinkles upon compression~\cite{Vandeparre2011a, Li2013}.
+However in that case, the wavelength gets larger away from the bound edge.
+Since the wavelength of our pattern is uniform throughout the sample we exclude the influence of the side boundaries.
+%
+%Influence of the side boundaries~\cite{Vandeparre2011a, Li2013} is also excluded since the wavelength of the pattern is uniform throughout the sample. 
+%
 Finally as the four side boundaries remain fixed $\sigma_{\perp}$ cannot come from the resistance to uniaxial stretching~\cite{Cerda2003}.
-XXX conclure plus clairement XXX
 
-Now, two points remain unclear: what triggers the the spontaneous syn\ae{}resis and swelling back processes?
-And what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
+Therefore at this stage, our observations still raise two unresolved issues which we adress in the next sections: (i) what triggers the the spontaneous syn\ae{}resis and swelling back processes?
+And (ii) what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
 
@@
 \caption{The wrinkling experiment.
-(a) Initial configuration where the sealed cell contains an homogeneous protein solution.
+(a) Initial configuration where the sealed cell contains a homogeneous protein solution.
 (b) Around the isoelectric pH, the gel forms and immediately expels solvent leading to (c).
 If tensile stresses were not released we would observe reversible swelling back leading to (d) a flat swollen gel layer.
-If tensile stresses are released, swelling back leads either to buckling (e) or to wrinkling (f) depending whether the bending is free or hindered by a transverse load.}
+If tensile stresses are released, swelling back leads either to buckling (e) or to wrinkling (f) depending on whether the bending is free or hindered by a transverse load.}
 \label{fig:dilation}
@@
 Symbols are direct measures upon addition of GDL.
-The continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert).
+The continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (inset).
 (c) Evolution of the elastic modulus $G'$ measured in a rheometer with full adhesion to the cone-plane geometry.
@@
 
-Syn\ae{}resis and swelling back mechanism depends on (i) the pH sensitivity of caseins, (ii) the route to gelation, (iii) the gel properties and (iv) the boundary conditions.
+The syn\ae{}resis and swelling back mechanism depends on (i)~the pH sensitivity of caseins, (ii)~the route to gelation, (iii)~the gel properties and (iv)~the boundary conditions.
 
@@
 At neutral pH caseins are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), casein molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), casein molecules lose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
 Beyond the isoelectric point caseins acquire a positive net charge and therefore partially re-dissolve.
@@
 
-When acidified by a strong acid like HCl, caseins flocculate and do not form a space-spanning gel.
-Therfore, we use glucono-$\delta$-lactone (GDL) a molecule that slowly hydrolyses into gluconic acid and thus lowers slowly, continuously and homogeneously the pH of the medium, Figure~\ref{fig:acidification}a.
+When acidified by a strong acid-like HCl, caseins flocculate and do not form a space-spanning gel.
+Therefore, we use glucono-$\delta$-lactone (GDL) a molecule that slowly hydrolyses into gluconic acid and thus lowers slowly, continuously and homogeneously the pH of the suspension, Figure~\ref{fig:acidification}a.
 This GDL-controlled acidification enables the formation of a bulk gel phase below pH=5.3 with $x_\text{free}\approx 0$.
@@
 To gain further insight into the mechanical properties of the gel we turn to microstructural observations under confocal microscopy.
-As shown in Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63} and in Supplementary Figure~\ref{fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut-off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
+As shown in Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63} and in Supplementary Figure~\ref{fig:nonmonotonic} the gel consists in a space-spanning fractal network with a cut-off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
 Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~\ref{fig:permeability}.
@@
 
-Therefore, the syn\ae{}resis and swelling back mechanism is related to charge stabilisation of the casein proteins with the pH.
-Such mechanism is only effective to produce wrinkle when adhesion of the caseins to both top and bottom walls is turned off.
-Coherently with our explanation, the amount of GDL, and thus the final pH reached by the suspension, controls the amount of swelling back.
+Therefore, the syn\ae{}resis and swelling back mechanism is related to charge stabilisation of the casein proteins as function of the pH.
+Such a mechanism is only effective to produce wrinkles when adhesion of the caseins to both top and bottom walls is turned off.
+Consistently with our explanation, the amount of GDL, and thus the final pH reached by the suspension, controls the amount of swelling back.
 Therefore at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
@@
 \caption{Sketch of the two limit scenarios for the wavelength selection.
-In both cases the gel film of thickness $h$ is destabilised by the excess area a give rise to wrinkles of amplitude $A(t)$ and wavelength $\lambda$.
+In both cases the gel film of thickness $h$ is destabilised by the excess area which gives rise to wrinkles of amplitude $A(t)$ and wavelength $\lambda$.
+A single wavelength is represented and the longitudinal dimension $L$ of the system is much larger than $\lambda$.
 The value of $\lambda$ is set by the interplay between the bending rigidity $B$ and a transverse load $\sigma_\perp$ due to a pressure gradient $p_2-p_1$ in the solvent of viscosity $\eta$.
@@
 $\sigma_\perp$ is due to the flow of the solvent through the porous gel of permeability $\alpha$ to fill the growing blister.
-(b) Poiseuille scenario: the gel film of negligible porosity is initially in the middle of the cell, separated by a distance $H$ from each wall.
+(b) Poiseuille scenario: the gel film of negligible porosity initially lies in the middle of the cell, separated by a distance $H$ from each wall.
 $\sigma_\perp$ is due to the lubrication flow in the top and bottom solvent layers.}
@@
 The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem.
-Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form.
+Large deformations of thin sheets are governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form.
 However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern up to dimensionless prefactors and isolate the physically meaningful ingredients~\cite{Cerda2003}.
@@
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Furthermore, equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+Furthermore, it was shown that equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+However, we have excluded these various origins for $\sigma_{\perp}$ in the discussion around Figure~\ref{fig:dilation}.
 
-XXX repetition entre paragraphe d'avant et d'après XXX
-
-In our situation we now have to determine $\sigma_\perp$.
-We have already excluded uniaxial stretching~\cite{Cerda2003}, gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, boundaries~\cite{Vandeparre2011a, Li2013} and adhesion~\cite{Vella2009a}.
-Two possible candidates remains for the transverse load $\sigma_\perp$: the porous flow through the gel or the viscous flows in the surrounding solvent layers.
+The only possible explanation left for $\sigma_{\perp}$ involve hydrodynamic stresses due to viscous flow of the solvent, either through the gel porous matrix itself or in the surrounding solvent layers.
 
@@
 
-Accordingly we consider two limit scenarios that leads to wrinkling patterns.
+Accordingly we consider the two limit scenarios sketched in Figure~\ref{fig:sketchDP} that lead to wrinkling patterns.
 In a first limit, the film of gel sits without sticking to the bottom wall of the cell, i.e.
-$H_1\rightarrow0$, see Supplementary Video~\ref{vid:cutDarcy}.
+$H_1\rightarrow0$, see Figure~\ref{fig:sketchDP}a and Supplementary Video~\ref{vid:cutDarcy}.
 Creating a two dimensional blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
@@
 
-The opposite limiting case is when the gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Supplementary Video~\ref{vid:cutPoiseuille}) and impermeable ($\alpha\rightarrow 0$).
+The opposite limiting case is when an impermeable ($\alpha\rightarrow 0$) gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Figure~\ref{fig:sketchDP}b and Supplementary Video~\ref{vid:cutPoiseuille}).
 Here destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
@@
 To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$, i.e.
-slow destabilisation of a highly permeable film close to a wall, the main dissipation mechanism would be the flow through the porous film; whereas for $H\gg H^*$, i.e.
-fast wrinkling of a poorly permeable film far from any wall, the flow in the viscous layers would dominate and set the wavelength.
+slow destabilisation of a highly permeable film close to a wall, the main dissipation mechanism being the flow through the porous film; whereas for $H\gg H^*$, i.e.
+fast wrinkling of a poorly permeable film far from any wall, the flow in the viscous layers dominates and sets the wavelength.
 
@@
 For a typical gel sheet, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
-This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ and therefore gravity is dismissed.
+This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ which confirms that gravity can be neglected as already inferred from test experiments where the cell is held vertically rather than horizontally.
 
@@
 We are in the position to vary $H/H^\star$ from 0.1 to 10 so that we can test the crossover between Darcy and Poiseuille mechanism.
-In Figure~\ref{fig:DarcyPoiseuille}a and b we have plotted the measure wavelength as a function of the model prediction $\lambda_D$ and $\lambda_P$ respectively.
+In Figure~\ref{fig:DarcyPoiseuille}a and \ref{fig:DarcyPoiseuille}b we have plotted the measured wavelength as a function of the model prediction $\lambda_D$ and $\lambda_P$ respectively.
 If the predictions are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$.
 Indeed, both model seems to apply in their domain of validity albeit with a prefactor XXX Thomas dit pas clair XXX.
-The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data.
+The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data correctly.
 To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Methods.
-The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ for the whole $H/H^*$ range.
+The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ over the whole $H/H^*$ range.
 
@@
 
-In conclusion we have taken advantage of the properties of the casein protein to engineer and study the wrinkling of a porous confined biogel.
+To conclude we have taken advantage of the properties of the casein protein to engineer and study the wrinkling of a porous confined biogel.
 In particular, we have fine-tuned the interplay between the attraction and the repulsion at the protein level to assemble dense glassy networks that shrink and swell as a result of continuous acidification.
@@
 Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins.
-As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
+As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in soft optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
 XXX texture food XXX
@@
 
-Home made optical cells are composed of a slide (RS) and a cover slip (Menzel-Gl\"aser) spaced by parafilm (Parafilm).
-The optical cell is briefly heated on a hot plate so that the parafilm welds to the slide and the cover slip.
+Home made optical cells are composed of a slide (RS) and a cover slip (Menzel-Gl\"aser) spaced by paraffin film (Parafilm).
+The optical cell is briefly heated on a hot plate so that the paraffin welds to the slide and the cover slip.
 Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell from \SIrange{50}{150}{\micro\metre}.
@@
 \includegraphics{ripples.pdf}
-\caption{Evidence of the ripples generated by the junction to flat patches, see~\cite{LeGoff2014}.
-Position of the bottom surface of the gel in (a) 2D and (b) 1D.
-Only first ripple is visible, highlighted by orange rectangles.
+\caption{Evidence for the ripples generated by the junction to flat patches, see~\cite{LeGoff2014}.
+(a) Position of the bottom surface of the gel obtained by confocal images.
+(b) one dimensional cut of the region of interest highlighted in (a) using an orange box.
+Only the first ripple is visible.
 (c) Schematic view of the interference between ripples.}
@@
 \caption{Schematic side view of the slit.
+A single wavelength is represented but the longitudinal dimension $L$ of the system is much larger than $\lambda$.
 The constant thickness gel film (orange) is surrounded by solvent.

### 2015-05-15 fig_plis/article_SI.tex
Abstract reprenant les nouveaux objectifs plutôt que de se concentrer sur la caséine. Ne parle plus de la différence de préfacteur

@@
 \begin{abstract}
-The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels, e.g.
-yoghurt.
-Less known is the partial redissolution of the gel at even lower pH.
-Here, using caseins as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, we design a model biogel film that spontaneously wrinkles in cascade, provided non-adhesive confinement.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
-Taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
+Confined thin surfaces may wrinkle due the growth of an excess of material. Usually the wavelength is set by elasticity or gravity, but here we explore new selection mechanisms based on hydrodynamics.
+First, inspired by yoghurt-making processes, we use caseins, a family of milk proteins, as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, to design a porous biogel film immersed in a confined buoyancy-matched viscous medium.
+Provided specific boundary conditions the biogel film spontaneously wrinkles in cascade.
+Second, using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
+Third, taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
 Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilisation/destabilisation.
@@
 If the predictions are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$.
-Indeed, both model seems to apply in their domain of validity albeit with a prefactor XXX Thomas dit pas clair XXX.
+Indeed, both model seems to apply in their domain of validity.
 The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data correctly.
@@
 Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins.
-As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in soft optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
-XXX texture food XXX
+As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in soft optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, food texturing~\cite{Mezzenga2005} XXXGibeau Faraday discussionXXX and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
 

### 2015-05-15 fig_plis/article_SI.tex
Gibaud2012a

@@
 Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins.
-As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in soft optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, food texturing~\cite{Mezzenga2005} XXXGibeau Faraday discussionXXX and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
+As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in soft optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, food texturing~\cite{Mezzenga2005,Gibaud2012a} and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
 

### 2015-05-15 fig_plis/article_SI.tex
Typos

@@
 Therefore contrary to observations in, e.g., Ref~\cite{Yoo2003}, we do not report any coarsening of the patterns.
-We believe the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}.
+We believe that the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}.
 Indeed when the amplitude saturates the gel flattens against the walls which creates sharp edges.
@@
 
-Therefore at this stage, our observations still raise two unresolved issues which we adress in the next sections: (i) what triggers the the spontaneous syn\ae{}resis and swelling back processes?
+Therefore at this stage, our observations still raise two unresolved issues which we address in the next sections: (i) what triggers the the spontaneous syn\ae{}resis and swelling back processes?
 And (ii) what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
@@
 However, we have excluded these various origins for $\sigma_{\perp}$ in the discussion around Figure~\ref{fig:dilation}.
-
+%
 The only possible explanation left for $\sigma_{\perp}$ involve hydrodynamic stresses due to viscous flow of the solvent, either through the gel porous matrix itself or in the surrounding solvent layers.
@@
 
-\section*{Supplementary material}
+\section*{Supplementary materials}
 

### 2015-05-15 fig_plis/article_SI.tex
Results and discussion avec 4 subsections

@@
 
-\paragraph*{Significance statement}
-Confined thin surfaces may wrinkle due the growth of an excess of material.
+%\paragraph*{Significance statement}
+%Confined thin surfaces may wrinkle due the growth of an excess of material.
 Here we present benchmark experiments that explore the possibility of wrinkling in biogels.
@@
 
-\section*{Dynamics of the wrinkling process}
+\section*{Results and discussion}
+
+\subsection*{Dynamics of the wrinkling process}
 
@@
 
-\section*{Syn\ae{}resis and swelling back mechanism}
+\subsection*{Syn\ae{}resis and swelling back mechanism}
 
@@
 
-\section*{Wavelength selection mechanism}
+\subsection*{Wavelength selection mechanism}
 
@@
 
-\section*{Experimental tests}
+\subsection*{Experimental tests}
 
@@
 
-\section*{Methods summary}
+\section*{Materials and Methods}
 

### 2015-05-15 fig_plis/article_SI.tex
Typos, répétitions Methods -> Materials and Methods Supplementary Methods -> Supplementary Materials

@@
 First, inspired by yoghurt-making processes, we use caseins, a family of milk proteins, as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, to design a porous biogel film immersed in a confined buoyancy-matched viscous medium.
-Provided specific boundary conditions the biogel film spontaneously wrinkles in cascade.
+Provided specific boundary conditions, but without any external stimulus, the biogel film spontaneously wrinkles in cascade.
 Second, using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
 Third, taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
-Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilisation/destabilisation.
-As such our results could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
+Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilisation/destabilisation, and therefore could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
 \end{abstract}
@@
 This cell is designed such that the casein adhesion to both top and bottom walls is turned off and so that the gel is only anchored to the remaining four vertical sides of the cell.
-
+%
 Although no external stimulus is exerted on the system, we observe the spontaneous formation of a wrinkling pattern, as shown in Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{vid:transmitted} and \ref{vid:reconstructed}.
@@
 Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that the wrinkles at the millimetre scale result from the spontaneous swelling of the casein network at the micron level upon acidification.
-
+%
 Finally we show that the wavelength is selected by a purely viscous mechanism combining porous Darcy flow and viscous Poiseuille flow.
@@
 
-The dynamics of the pattern formation are followed by light transmission microscopy and confocal microscopy, see Methods.
+The dynamics of the pattern formation are followed by light transmission microscopy and confocal microscopy, see Materials and Methods.
 As shown in Figure~\ref{fig:dynamics}b and Supplementary Video~\ref{vid:transmitted}, we observe the spontaneous formation of a first pattern that appears with a characteristic wavelength $\lambda\ll L$.
@@
 
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}b, and excess surface area of the gel film, Figure~\ref{fig:sideview}c, see Supplementary Methods.
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}b, and excess surface area of the gel film, Figure~\ref{fig:sideview}c, see Supplementary Materials.
 The evolution of the gel volume is surprisingly non-monotonous.
@@
 This path is dynamic in essence.
-In Figure~\ref{fig:sideview}d, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Methods for the definition.
+In Figure~\ref{fig:sideview}d, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Materials for the definition.
 We observe that the velocity peaks at different times corresponding to syn\ae{}resis and to the respective growths of the primary, secondary and tertiary patterns.
@@
 Beyond the isoelectric point caseins acquire a positive net charge and therefore partially re-dissolve.
-In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH, see Supplementary Methods.
+In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH, see Supplementary Materials.
 Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction $x_\text{free}$ is 1.
@@
 Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-In order to turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes, see Methods, effectively creating non adhesive boundary conditions for the gel.
+In order to turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes, see Materials and Methods, effectively creating non adhesive boundary conditions for the gel.
 When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall.
@@
 Consistently with our explanation, the amount of GDL, and thus the final pH reached by the suspension, controls the amount of swelling back.
-Therefore at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
+Thus at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
 
@@
 
-A more complete derivation of the models can be found in the Supplementary Methods and yields the following prefactors:
+A more complete derivation of the models can be found in the Supplementary Materials and yields the following prefactors:
 \begin{align}
@@
 
-For a typical gel sheet, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+For a typical gel film, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
 This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ which confirms that gravity can be neglected as already inferred from test experiments where the cell is held vertically rather than horizontally.
@@
 Darcy and Poiseuille models yield wavelengths that are of the same order in most of the cases under study.
-Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them, see Supplementary Methods about such measurements.
+Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them, see Supplementary Materials about such measurements.
 As summarised in Supplementary Table~\ref{tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
@@
 For each experiment we measure independently $\lambda$ and all the parameters that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}).
-$\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Methods and Supplementary methods.
+$\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Materials and Methods and Supplementary Materials.
 
@@
 The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data correctly.
-To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Methods.
+To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Materials.
 The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ over the whole $H/H^*$ range.

### 2015-06-03 creep_beamer.tex
Référence du PRL en plus du Arxiv
Talk Zurich Add snapshot for creep experiments

@@
-\begin{frame}{Creep and yielding}
-Just accepted in PRL, \textit{ArXiv 1401.8234}
-
-\bigskip
-\structure{Merciless yoghurt breakers}
+\begin{frame}{Merciless yoghurt breakers}
+\textit{Leocmach et al. PRL 113, 038303 (2014), \footnotesize ArXiv 1401.8234}
 
@@
 %\tikzset{external/force remake=false}
+
+\begin{frame}{The yoghurt creep experiment}
+\includegraphics[width=\textwidth]{creep/Fig3}
+\begin{center}\tikzsetnextfilename{creep_xp_snapshot}
+\begin{tikzpicture}
+\begin{groupplot}[
+	group style={
+		group name=g, group size=2 by 1,
+		horizontal sep=4\baselineskip,
+		},
+	xmode=log,
+	ymode=log,
+	xlabel={$t$ (\si{\second})},
+	xmin=0.1,
+	%restrict x to domain=0.05:3e3,
+	]
+	\nextgroupplot[ylabel={strain rate $\dot\gamma$ (\si{\per\second})}, ymin=1e-4, ymax=2e-1]
+	\addplot[Accent1] table {Y110_300Pa.gdot}; 
+	\addplot[Accent2, only marks, mark=o] coordinates {(31, 1.6e-3) (889, 2e-4) (1442, 5.5e-4) (1509, 1.1e-2)};
+
+	\nextgroupplot[
+		ylabel={strain $\gamma$},ymin=0.16,ymax=2,
+		ytick={0.2, 0.4,0.8,1.6}, yticklabels={0.2, 0.4,0.8,1.6},]
+	\addplot[Accent1] table {Y110_300Pa.txt}; 
+	\addplot[Accent2, only marks, mark=o] coordinates {(31, 0.37481) (880, 0.70529) (1459, 0.87393) (1526,0.93622)};
+
+\end{groupplot}
+\end{tikzpicture}
+\end{center}
+\end{frame}
 

### 2015-06-03 talk_Zurich.tex
Talk Zurich Add snapshot for creep experiments

--- 2015-01-09 seminaire_Montpellier.tex
@@
 \institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
-\title[Yoghurt under stress]{Yoghurt under stress}
+\title[Yoghurt under stress]{Acid-induced protein gels: from gelation to stress-induced failure}
 \author[M. Leocmach]{Mathieu Leocmach}
-\date{20 October 2014}
+\date{10 June 2015}
 \titlegraphic{
@@
 
-\setcounter{framenumber}{0}
-
-\include{parcourt}
-
 \begin{frame}[plain]
@@
 \end{frame}
+
+\setcounter{framenumber}{0}
 
@@
 
-\section*{Wrinkling}
-
-%\tikzset{external/force remake}
-\begin{frame}{Over-acidification: Faster but weaker}
-\tikzsetnextfilename{gelation_GDL}
-\begin{tikzpicture}
-\begin{axis}[
-	height=0.8\textheight,
-	width=\textwidth,
-	xlabel={time (h)}, ylabel={$G^\prime$ (\si{\pascal})},
-	cycle list name=earthy,
-	no marks,
-	xmin=0, xmax=20,ymin=0,
-	]
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1_Y265.prise} node[anchor=base west] {1\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.25_Y277.prise} node[anchor=base west] {1.25\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL1.5_Y275.prise} node[anchor=base west] {1.5\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL2_Y268.prise} node[anchor=base west] {2\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL3_Y270.prise} node[anchor=base west] {3\%};
-	\addplot table[x expr={\thisrowno{0}/3600}]{cas4_GDL4_Y271.prise} node[anchor=base west, yshift=-0.2em] {4\%};
-\end{axis}
-\end{tikzpicture}
-
-Caseins regain some solubility at low pH.
-\end{frame}
-
-\include{plis_beamer}
-
 \end{document}

### 2015-06-03 talk_Zurich.tex
Mouth feel Ductile vs Brittle

@@
 
-\begin{frame}{Application of biogels \textup{\normalsize(physical gels
+\begin{frame}{Mouth feel}
+\begin{columns}
+\column{0.8\textwidth}
+\begin{itemize}
+\item A very complex, multidimensional sensation
+\item Food behaviour at large deformation is important
+\item How does food break into pieces ?
+\end{itemize}
+\column{0.2\textwidth}
+\includegraphics[width=\textwidth]{violon}
+\end{columns}
+
+\bigskip
+\begin{columns}[T]
+\column{0.475\textwidth}
+\begin{block}{Ductile fracture}
+\includegraphics[height=5\baselineskip]{deflated_balloon}
+
+Irreversible deformation
+\begin{itemize}
+\item rubber
+\item lead
+\item butter
+\end{itemize}
+\end{block}
+
+\column{0.475\textwidth}
+\begin{block}{Brittle fracture}
+\includegraphics[height=6\baselineskip]{spaghetti_breaking}\hfill
+\includegraphics[height=6\baselineskip]{Tea_bowl_fixed_in_the_Kintsugi_method}
+\begin{itemize}
+\item concrete
+\item asphalt
+\item glass
+\end{itemize}
+
+\end{block}
+\end{columns}
+\end{frame}
+
+\begin{frame}{What about biogels? \textup{\normalsize(physical gels
 $E_{bond}\gg k_B T$)}}
@@
 \end{frame}
-
-\begin{frame}{Fundamental issues in biogels}
-\begin{itemize}
-\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
-\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
-\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
-\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
-\end{itemize}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=6\baselineskip]{Bonn_fracture} &
-\includegraphics[height=6\baselineskip]{Villi_sq}
-\end{tabu}
-\begin{block}{Our questions}
-\begin{itemize}
-\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
-\item What is the role of the solvent in mechanical instabilities?
-\end{itemize}
-\end{block}
-\end{frame}
+%
+%
+%
+%\begin{frame}{Fundamental issues in biogels}
+%\begin{itemize}
+%\item Linear elasticity is not well understood \textit{\scriptsize Gardel et al. Science 2004}
+%\item Strain and stress stiffening \textit{\scriptsize Storm et al. Nature 2005}
+%\item Fractures \textit{\scriptsize Bonn et al. Science 1998, Baumberger et al. Nature Material 2006}
+%\item Mechanical instabilities and morphogenesis \textit{\scriptsize Shyer et al. Science 2013}
+%\end{itemize}
+%\begin{tabu}{X[c]X[c]}
+%\includegraphics[height=6\baselineskip]{Bonn_fracture} &
+%\includegraphics[height=6\baselineskip]{Villi_sq}
+%\end{tabu}
+%\begin{block}{Our questions}
+%\begin{itemize}
+%\item What is the behaviour of a biogel under shear deep into the nonlinear regime?
+%\item What is the role of the solvent in mechanical instabilities?
+%\end{itemize}
+%\end{block}
+%\end{frame}
 

### 2015-06-04 talk_Zurich.tex
Texture as important as taste

@@
 \setcounter{framenumber}{0}
+
+\begin{frame}{Texture as important as taste}
+\begin{tikzpicture}[inner sep=0, callout pointer segments=3]
+\node(ishichan) {\includegraphics[height=8\baselineskip]{ishichan}};
+\node[above=of ishichan] {\includegraphics[height=8\baselineskip]{calamar}};
+\node[cloud callout,anchor=west, draw, callout relative pointer={(-30:2\baselineskip)}] at ($(ishichan.north)+(-0.5\textwidth,0)$) {\includegraphics[height=5\baselineskip]{aldente}};
+\node[cloud callout, aspect=0.75, anchor=east, draw, callout relative pointer={(-150:2\baselineskip)}] at ($(ishichan.north)+(0.5\textwidth-10pt,0)$) {\includegraphics[height=5\baselineskip]{chewie}};
+\end{tikzpicture}
+\end{frame}
 

### 2015-06-04 talk_Zurich.tex
Behaviour of the food
Biogel, food and coauthors at the end

@@
 \item A very complex, multidimensional sensation
-\item Food behaviour at large deformation is important
-\item How does food break into pieces ?
+\item Behaviour of the food at large deformation?
+\item How does food break into pieces?
 \end{itemize}
@@
 
-\begin{frame}{What about biogels? \textup{\normalsize(physical gels
-$E_{bond}\gg k_B T$)}}
-\begin{tabu}{cX[c]c}
-\includegraphics[height=0.3\textheight]{cell_mech} &
-\includegraphics[height=0.3\textheight]{electrophoresis} &
-\includegraphics[height=0.3\textheight]{cosmetics} \\
-Cell mechanics & Electrophoresis & Cosmetics\\
-\end{tabu}
-\begin{tabu}{X[c]X[c]}
-\includegraphics[height=0.3\textheight]{bacterial_culture} &
-\includegraphics[height=0.3\textheight]{food} \\
-Bacterial culture & Food\\
-\end{tabu}
-
-\begin{footnotesize}
-\begin{block}{Sources}
-\begin{tabu}{lXl}
-Wikimedia Commons & www.madaboutscience.com & www.keautystore.com\\
-\end{tabu}
-\end{block}
-\end{footnotesize}
-\end{frame}
 %
@@
 
+\begin{frame}{Can we generalise to biogels? \textup{\normalsize(physical gels
+$E_{bond}\gg k_B T$)}}
+\begin{tabu}{cX[c]c}
+\includegraphics[height=0.3\textheight]{cell_mech} &
+\includegraphics[height=0.3\textheight]{electrophoresis} &
+\includegraphics[height=0.3\textheight]{cosmetics} \\
+Cell mechanics & Electrophoresis & Cosmetics\\
+\end{tabu}
+\begin{tabu}{X[c]X[c]}
+\includegraphics[height=0.3\textheight]{bacterial_culture} &
+\includegraphics[height=0.3\textheight]{food} \\
+Bacterial culture & Food\\
+\end{tabu}
+
+\begin{footnotesize}
+\begin{block}{Sources}
+\begin{tabu}{lXl}
+Wikimedia Commons & www.madaboutscience.com & www.keautystore.com\\
+\end{tabu}
+\end{block}
+\end{footnotesize}
+\end{frame}
+
+\begin{frame}{Merciless yoghurt breakers}
+\textit{Leocmach et al. PRL 113, 038303 (2014), \footnotesize ArXiv 1401.8234}
+
+\bigskip
+\begin{tabu}{X[c]X[c]c}
+\includegraphics[height=0.3\textheight]{Chris}&
+\includegraphics[height=0.3\textheight]{Thibaut}&
+\includegraphics[height=0.3\textheight]{Seb}\\
+Christophe Perge & Thibaut Divoux & Sebastien Manneville\\
+PhD student & CNRS researcher & Professor\\
+E.N.S. Lyon &  CRPP Bordeaux & E.N.S. Lyon\\
+\end{tabu}
+
+\bigskip
+\begin{description}[Special thanks]
+\item[Special thanks] to Alan Parker (Firmenich) for prompting this study and providing casein \& GDL
+\end{description}
+
+\end{frame}
+
 \end{document}

### 2015-06-04 creep_beamer.tex
Biogel, food and coauthors at the end
Reorganise completement suivant les conseils de Thomas. N'affiche plus de strain, que des strain rate Le but c'est d'expliquer pourquoi ça rescale Vire Monkman, Griffith, etc. Explique mieux le passage de 0.15 à -0.85 et dans le regime 3

--- 2014-10-18 creep_beamer.tex
@@
-\begin{frame}{Creep and yielding}
-Just accepted in PRL, \textit{ArXiv 1401.8234}
-
-\bigskip
-\structure{Merciless yoghurt breakers}
-
-\bigskip
-\begin{tabu}{X[c]X[c]c}
-\includegraphics[height=0.3\textheight]{Chris}&
-\includegraphics[height=0.3\textheight]{Thibaut}&
-\includegraphics[height=0.3\textheight]{Seb}\\
-Christophe Perge & Thibaut Divoux & Sebastien Manneville\\
-PhD student & CNRS researcher & Professor\\
-E.N.S. Lyon &  CRPP Bordeaux & E.N.S. Lyon\\
-\end{tabu}
-
-\bigskip
-\begin{description}[Special thanks]
-\item[Special thanks] to Alan Parker (Firmenich) for prompting this study and providing casein \& GDL
-\end{description}
-
-\end{frame}
-
 %\tikzset{external/force remake}
@@
 
+%Couette cell
+\let\mydimb\relax
+	\newlength\mydimb
+	\pgfmathsetlength{\mydimb}{6\baselineskip}
+
+\node[shape=cylinder, draw, shape border rotate=90,minimum width=\mydimb, minimum height=7\baselineskip, anchor=north, aspect=2] at (rheometer.north-|webcam.north) (Stator) {};
+
+\node[shape=cylinder, draw, shape border rotate=90,minimum width=\mydimb-1em, minimum height=6.75\baselineskip, anchor=before bottom , aspect=1.5] at ($(Stator.before bottom)+(-0.5em,0)$) (Rotor){};
+
+\draw[->, x radius=0.5\mydimb-2em, y radius=0.2\baselineskip]  (Stator.before top) ++(0.5\mydimb,+0.2\baselineskip) arc[start angle=90, end angle=270] node[right=1mm, font=\scriptsize] {$\sigma$};
+
 \begin{scope}[draw]
-	\node[below=0of spoon.south] (couette) {A Taylor-Couette cell where the yogurt is made \emph{in situ}};
-	\draw[->] (couette) -- ($(spoon.south)+(0,2em)$);
-	\node[above=0of ultrasound] {ultrasonic imaging};
-	\node[above=0of webcam, text width=0.2\textwidth, align=center] {optical imaging};
+	\node[below right=0of Stator.north east, text width=0.2\textwidth, align=center] (couette) {A Taylor-Couette cell where the yogurt is made \emph{in situ}};
+	\draw[->] ($(spoon.west)+(2em,0)$) -- (Stator);
+	\node[below=0of ultrasound] {ultrasonic velocimetry};
+	\node[below right=0of webcam.south west, text width=0.25\textwidth, align=center] {optical imaging};
 	\draw[->] (spoon.north east) +(235:2em) -- (rheometer);
@@
 
-%\tikzset{external/force remake}
 \begin{frame}{The yoghurt creep experiment}
-\movie[externalviewer]{\tikzsetnextfilename{creep_xp}\begin{tikzpicture}
-\node[inner sep=0] (a) {\includegraphics[height=16\baselineskip]{creep_experiment.jpg}};
-\node[below right=\baselineskip of a.north east, inner sep=0] (vmap) {\includegraphics[width=4\baselineskip]{jet_cmap}};
+\let\mylena\relax
+\newlength\mylena
+\pgfmathsetlength{\mylena}{0.19\textwidth}
+\includegraphics[width=\mylena]{Y110_2013-03-01_02-55-11.jpg}\hfill
+\includegraphics[width=\mylena]{Y110_2013-03-01_03-06-28.jpg}\hfill
+\includegraphics[width=\mylena]{Y110_2013-03-01_03-19-47.jpg}\hfill
+\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-10.jpg}\hfill
+%\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-18.jpg}\hfill
+\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-19.jpg}\\
+%\begin{center}
 
-\begin{scope}[font=\footnotesize]
-	\node[above=0 of vmap] {velocity map};
-	\node[below=0 of vmap.south west] {-50};
-	\node[below=0 of vmap.south] (v0) {0};
-	\node[below=0 of vmap.south east] {50};
-	\node[below=0 of v0, inner sep=0]{\si{\micro\metre\per\second}};
-\end{scope}
+\bigskip\tikzsetnextfilename{creep_xp_snapshot}
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=12\baselineskip,
+	xmode=log,
+	ymode=log,
+	xlabel={$t$ (\si{\second})},
+	xmin=0.1,
+	ylabel={strain rate $\dot\gamma$ (\si{\per\second})}, 
+	ymin=0.9e-4, ymax=2e-1,
+	axis on top,
+	]
+ 	\fill[Accent1!20]
+		(axis cs:3,1e-2) ellipse[rotate=-25, x radius=0.4\textwidth, y radius=0.04\textwidth] node[below, rotate=-25, black] {power-law creep}
+		(axis cs:0.8e3,2e-4) ellipse[x radius=0.05\textwidth, y radius=0.02\textwidth] node[below, black] {nucleation}
+		(axis cs:1577,1e-2) ellipse[x radius=0.04\textwidth, y radius=4\baselineskip] node[below, rotate=90, black] {explosive regime}
+		;
+	\addplot[thick] table {Y110_300Pa.gdot}; 
+	\addplot[thick, only marks, mark=o] coordinates {(687, 2.1e-4) (1476, 7.35e-4) (1509, 1.12e-3)};
+	\node[below] at (rel axis cs:0.5,1) {$\sigma=\SI{300}{\pascal}$};
+\end{axis}
+\end{tikzpicture}
+\end{frame}
 
-\begin{scope}[shift={(a.north west)}, shift={(5.8\baselineskip,-1.1\baselineskip)}, ultra thick]
-	\draw[Accent1, x radius=3.1\baselineskip, y radius=0.3\baselineskip] ellipse[] ++(3.1\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-3.1\baselineskip,0) ellipse[] ++(-3.1\baselineskip,0)  -- ++(0,7.4\baselineskip);
-	\shadedraw[Accent2,ball color=Accent2, x radius=2\baselineskip, y radius=0.1\baselineskip] ellipse[] ++(2\baselineskip,0)  -- ++(0,-7.4\baselineskip)  ++(-2\baselineskip,0) ellipse[] ++(-2\baselineskip,0)  -- ++(0,7.4\baselineskip);
-	\fill[Accent2, left color=black,right color=black, middle color=Accent2, ] (-0.3\baselineskip,0) rectangle +(0.6\baselineskip, \baselineskip);
-	\begin{scope}[Main]
-		\node[draw, minimum height=4.2\baselineskip, minimum width=1.5\baselineskip] at (2.5\baselineskip,-4\baselineskip) (window){};
-		\node[draw, minimum height=7.2\baselineskip, minimum width=4\baselineskip] at (8.9\baselineskip,-4.2\baselineskip) (USV){};
-		\draw(window.north east) -- (USV.north west) (window.south east) -- (USV.south west);
-		\node[right, text width=4\baselineskip] at (USV.east) {ultrasound region of interest};
-	\end{scope}
-
-\draw[<->,thin] ++(-2\baselineskip,-\baselineskip) -- +(-1.1\baselineskip,0) node[midway,above] {\SI{2}{\milli\metre}};
-\end{scope}
-\draw[->] (window.north west) ++(-1em, 0.2em) -- +(1em,0) node[midway,above,inner sep=1pt] {$r$};
-\draw[->] (window.north west) ++(-1em, 0.2em) -- +(0,-1em)  node[midway,left] {$z$};
-
-\node[text width=4\baselineskip] at (-3\baselineskip,-2.5\baselineskip) {shear rate response};
-\node at (5\baselineskip,-2.5\baselineskip) {strain response};
-\end{tikzpicture}%
-}{Yaourt141_enhanced.avi}
-\end{frame}
-%\tikzset{external/force remake=false}
-
-\begin{frame}{Creep rheology: Three regimes}
-\tikzsetnextfilename{three_regimes}
+\begin{frame}{Different (constant) stresses: universality}
+\begin{columns}
+\column{0.6\textwidth}
+\tikzsetnextfilename{gdots_not_normed}%
 \begin{tikzpicture}
@@
 	width=\textwidth,
-	height=0.8\textheight,
+	height=10\baselineskip,
 	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
-	ymin=0.2,ymax=2,
-	ytick={0.2, 0.4,0.8,1.6}, yticklabels={0.2, 0.4,0.8,1.6},
-	ylabel={strain $\gamma$},
-	extra tick style={grid=major},%
-	extra y ticks={1}, extra y tick labels={1.0},
+	ylabel={Strain rate $\dot{\gamma}$ (\si{\per\second})},
 	cycle list name=earthy,
-	axis on top,
+	no marks,
 	]
-	\fill[Accent1!20]
-		(axis cs:10,0.4) ellipse[rotate=-30, x radius=0.2\textwidth, y radius=0.12\textwidth]
-		(axis cs:2e2,1.6) ellipse[x radius=0.35\textwidth, y radius=0.05\textwidth]
-		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
-		;
-	\node[above right] at (axis cs:1e1,0.2) {primary};
-	\node[below left] at (axis cs:4e5,0.7) {secondary};
-	\node[left] at (axis cs:5,1.6) {tertiary};
-	\addplot table{Y27_200Pa_gamma_decimated.txt} node (s200){};
-	\addplot table{Y38_300Pa_gamma_decimated.txt} node (s300){};
-	\addplot table{Y25_400Pa_gamma_decimated.txt}  node (s400){};
-	\addplot table{Y32_550Pa_gamma_decimated.txt}  node (s550){};
-	\addplot table{Y39_1000Pa_gamma_decimated.txt} node (s1000){};
+% 	\fill[Accent1!20]
+% 		(axis cs:10,0.4) ellipse[rotate=-20, x radius=0.15\textwidth, y radius=0.02\textwidth]
+% 		(axis cs:2e2,1.6) ellipse[x radius=0.35\textwidth, y radius=0.05\textwidth]
+% 		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
+% 		;
+	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	%\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[sloped, pos=0.75, below left] {Primary creep};
+
 \end{loglogaxis}
-\begin{scope}[anchor=base, every node/.style={yshift=0.2em}]
+\begin{scope}[anchor=base east, every node/.style={ rotate=90}]
 	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
@@
 
-Failure at $\gamma\approx 1$ for a well defined time $\tau_f$
+\column{0.4\textwidth}
+\structure{Normalisation}
+\begin{description}[$\dot\gamma_\text{min}$]
+\item[$\tau_f$] failure time 
+\item[$\dot\gamma_\mathrm{min}$] minimum strain rate 
+\end{description}
+
+\smallskip
+\hfill$\Rightarrow$ \colorbox{Accent1}{\color{white}\Large Physical origin?}
+\end{columns}
+
+\tikzsetnextfilename{gdots_normed}%
+\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=3 by 1,
+			horizontal sep=3em,
+			},
+		width=0.35\textwidth,
+		cycle list name=earthy,
+		no marks,
+		clip mode=individual,
+		]
+	\nextgroupplot[
+		title={power-law creep},
+		xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
+		ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+		xmode=log,ymode=log,
+		xtick={1e-4, 1e-2, 1},
+		]
+		\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		%\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*x^(-0.85) + 0.187/(1-x)};
+
+	\nextgroupplot[
+		title={fracture nucleation},
+		xmin=0, xmax=1, xlabel={$t/\tau_f$},
+		ymin=0.5, ymax=4,
+		restrict y to domain=0.5:10,
+		]
+		\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		%\addplot[yellow, ultra thick, domain={0.01:0.99}] {0.378*x^(-0.85) + 0.187/(1-x)};
+
+	\nextgroupplot[
+		title={explosive regime},
+		xmin=1e-5, xmax=2, xlabel={$(\tau_f-t)/\tau_f$},
+		x dir=reverse,
+		xmode=log,ymode=log,
+		xtick={1e-4, 1e-2, 1},
+	]
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		%\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*(1-x)^(-0.85) + 0.187/x};
+	\end{groupplot}
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}{Power-law creep and plasticity}
+\begin{block}{``Andrade'' creep (1910)}
+\begin{itemize}
+\item Power-law creep known in \alert{ductile solids} for over a century
+\item Classically explained by disinclination dynamics in crystals
+\item Recently explained by plastic events in glasses \textit{\footnotesize Miguel et al., PRL 2002}
+\end{itemize}
+\end{block}
+
+Any plastic event in yoghurt?
+
+\bigskip
+\movie[externalviewer]{\structure{Ultrasound velocimetry}}{Yaourt110_primary.avi} $\Rightarrow$ local velocity maps
+\begin{itemize}
+	\item homogeneous \movie[externalviewer]{strain field}{Yaourt110_profils.avi} (no wall slip, no shear band)
+	\item if present, plastic events are below our resolution (a few \si{\micro\metre})
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Is power-law creep damaging the structure?}
+\begin{itemize}
+\item Stop creep much before ($\gamma<0.3$) fracture nucleation ($\gamma\approx 1$)
+\item Let relax $\Rightarrow$ No complete recovery
+\item Reset deformation and start again $\Rightarrow$  superimposable
+\end{itemize}
+
+\tikzsetnextfilename{recovery}
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=8\baselineskip,
+	xlabel={time (\si{\hour})},
+	ylabel={strain $\gamma$},
+	xmin=-0.15, xmax=7.5, xtick={0, 1, ..., 7},
+	axis background/.style={fill=white},
+	]
+	\addplot[black, no marks] table[x expr={\thisrowno{0}/3600}] {Y261_oscil_creep_recovery_100Pa_decimated.txt};
+	\begin{scope}[<-,Main, inner sep=0]
+	\draw (axis cs:0,0) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow \SI{100}{\pascal}$};
+	\draw (axis cs:1/6,0.2727007) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow 0$};
+	\draw (axis cs:2.388,0.0159) -- +(0,1em) node[anchor=base east]{$\gamma\leftarrow 0$};
+	\end{scope}
+\end{axis}
+\end{tikzpicture}
+
+\begin{itemize}
+\item Undamaged structure\raisebox{-0.6\baselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.\baselineskip}\right\}\Rightarrow\text{Plasticity cannot explain power law}$}
+\item Viscous dissipation
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Power-law creep as linear response}
+\begin{columns}[T]
+\column{0.45\textwidth}
+\tikzsetnextfilename{linear_rheology}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	name=a,
+	height=0.8\textwidth,
+	width=\textwidth,
+	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{$G^\prime$}, \textcolor{Accent2}{$G^{\prime\prime}$}(\si{\pascal})},
+	domain={6e-2:70},
+	ymin=10,
+	]
+	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
+	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
+	\addplot[Main, no marks]{100*x^0.15}  node[midway, below=1em] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
+
+\end{loglogaxis}
+\end{tikzpicture}
+\[ \sigma \propto  \gamma f^\alpha \xrightarrow{\mathrm{Fourier}} \sigma \propto \gamma t^{-\alpha} \]
+Constant $\sigma$
+\[ \gamma(t) \propto \sigma t^{\alpha} \xrightarrow{\frac{d}{dt}} \dot\gamma(t) \propto \sigma t^{\alpha-1}\]
+
+\column{0.55\textwidth}
+\tikzsetnextfilename{primary_rescaled}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
+	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+	cycle list name=earthy,
+	clip mode=individual,
+	]
+	\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.378*x^(-0.85)} node[Accent2,pos=0.75, below left] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
+\end{loglogaxis}
+\end{tikzpicture}
+
+Power-law creep is 
+\begin{itemize}
+\item the linear regime of the material 
+\item up to 100\% strain!
+\item explained without plasticity
+\end{itemize}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Explaining the explosive regime}
+\begin{columns}[T]
+\column{0.45\textwidth}
+
+\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-15-00.jpg}\hfill
+\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-20-00.jpg}\hfill
+\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-21-15.jpg}
+
+\structure{Fracture length $\ell$}
+
+\tikzsetnextfilename{fracture_length}
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=1e-3, xmax=1, xlabel={$(\tau_f-t)/\tau_f$},
+	xmode=log, x dir=reverse,
+	ylabel={$\ell/H$},
+	ymin=0, ymax=0.5,
+	clip mode=individual,
+	axis background/.style={fill=white},
+	]
+	\addplot[only marks, Accent1, error bars/.cd, y dir=both, y explicit] table[y error index=2]{Y110_fracturelength_log.txt};
+	\addplot[Accent2, ultra thick, no marks, domain={1e-3:1}] {- 0.06 * ln(x)} node [pos=0.25,left=1em] {$\frac{\ell}{H}\sim \ln\frac{\tau_f-t}{\tau_f}$};
+\end{axis}
+\end{tikzpicture}
+
+\column{0.55\textwidth}
+\[ \gamma \propto \ell \propto \ln\frac{\tau_f-t}{\tau_f}\]
+
+\[ \dot\gamma \propto \left(\frac{\tau_f-t}{\tau_f}\right)^{-1}\]
+
+\tikzsetnextfilename{tertiary_rescaled}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=1e-5, xmax=2, xlabel={$(\tau_f-t)/\tau_f$},
+	x dir=reverse,
+	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+	cycle list name=earthy,
+	clip mode=individual,
+	]
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.187/x} node[Accent2, pos=0.6, below right, inner sep=0, yshift=-0.5em] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{\tau_f}{\tau_f-t}\right)$};
+\end{loglogaxis}
+\end{tikzpicture}
+\end{columns}
+$\Rightarrow$ Explosive regime dominated by fracture growth
+\end{frame}
+
+\begin{frame}{Failure time \& master curve}
+\begin{columns}[T]
+\column{0.5\textwidth}
+\tikzsetnextfilename{basquin}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xlabel={$\sigma$ (\si{\pascal})},
+	ylabel={$\tau_f$ (\si{\second})},
+	xmin=100, ymin=10, ymax=1e6,
+	xtick={100, 200,..., 1000}, xticklabels={100, 200,,, 500,,,,, 1000},
+	]
+	\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
+	\addplot[no marks, black, domain={100:1000}] {4.2e17*x^(-5.45)} node [midway, above right] {$\tau_f\sim \sigma^{-5.5}$};
+	\node[above right] at (rel axis cs:0,0){No yield stress!};
+\end{loglogaxis}
+\end{tikzpicture}
+
+\begin{block}{Basquin law $\tau_f\sim \sigma^{-\beta}$}
+\begin{itemize}
+\item fatigue (oscillatory stress)
+\item heterogeneous solids
+\end{itemize}
+\end{block}
+
+\column{0.5\textwidth}
+\tikzsetnextfilename{secondary_rescaled}
+\begin{tikzpicture}
+\begin{axis}[
+	width=\textwidth,
+	height=0.8\textwidth,
+	xmin=0, xmax=1, xlabel={$t/\tau_f$}, xtick={0,0.2,...,1},
+	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+	ymin=0.5, ymax=4,
+	cycle list name=earthy,
+	restrict y to domain=0.5:10,
+	clip mode=individual,
+	]
+	\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
+	\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
+	\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
+	\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
+	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
+	\addplot[yellow, ultra thick, domain={0.01:0.99}] {0.378*x^(-0.85) + 0.187/(1-x)};
+	\draw[<-] (axis cs:0.56,1) -- +(0,1em) node[above] {$\tau_\text{min}$};
+\end{axis}
+\end{tikzpicture}
+
+\structure{Master curve}
+\[\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}} = \underbrace{\lambda \left(\frac{t}{\tau_f}\right)^{-0.85}}_\text{linear response} + \underbrace{\frac{\mu}{1 - t/\tau_f}}_\text{fractures}\]
+\end{columns}
 \end{frame}
@@
 %\tikzset{external/force remake}
-\begin{frame}{Dimensionless creep response}
-\begin{columns}
-\column{0.55\textwidth}
-\tikzsetnextfilename{primary_rescaled}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
-	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
-	cycle list name=earthy,
-	clip mode=individual,
-	]
-	\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
-	\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
-	\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
-	\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
-	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.378*x^(-0.85)} node[Accent2,pos=0.75, below left] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
-\end{loglogaxis}
-\end{tikzpicture}
-
-Linear viscoelasticity accounts for the Andrade exponent
-
-\column{0.45\textwidth}
-\tikzsetnextfilename{linear_rheology}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	height=0.8\textwidth,
-	width=\textwidth,
-	xlabel={Frequency (\si{\hertz})}, ylabel={\textcolor{Accent1}{$G^\prime$}, \textcolor{Accent2}{$G^{\prime\prime}$}(\si{\pascal})},
-	domain={6e-2:70},
-	ymin=10,
-	]
-	\addplot[Accent1, only marks, mark=*] table{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Accent2, only marks, mark=o] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt};
-	\addplot[Main, no marks]{100*x^0.15}  node[midway, below=1em] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
-
-\end{loglogaxis}
-\end{tikzpicture}
-\textcolor{Main}{
-\begin{align*}
-\Rightarrow & J(t) \equiv& \frac{\gamma(t)}{\sigma}\sim& t^{0.15}\\
-\Rightarrow & \dot{\gamma}(t)\sim& t^{0.15-1} =& t^{-0.85}
-\end{align*}
-}
-\end{columns}
-\end{frame}
 
 %\tikzset{external/force remake}
-\begin{frame}{Tertiary creep}
-\begin{columns}
-\column{0.55\textwidth}
-Time to failure
-
-\tikzsetnextfilename{tertiary_rescaled}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=1e-5, xmax=2, xlabel={$(\tau_f-t)/\tau_f$},
-	x dir=reverse,
-	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
-	cycle list name=earthy,
-	clip mode=individual,
-	]
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
-	\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.187/x} node[Accent2, pos=0.6, below right, inner sep=0] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{\tau_f}{\tau_f-t}\right)$};
-\end{loglogaxis}
-\end{tikzpicture}
-
-Close to failure: 
-\begin{itemize}
-\item finite time singularity
-\item logarithmic divergence
-\end{itemize}
-$\Rightarrow$ dominated by fracture growth
-
-\column{0.45\textwidth}
-\begin{block}{Fracture growth}
-\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-15-00.jpg}\hfill
-\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-20-00.jpg}\hfill
-\includegraphics[width=0.33\textwidth]{Y110_2013-03-01_03-21-15.jpg}
-
-\tikzsetnextfilename{fracture_length}
-\begin{tikzpicture}
-\begin{axis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=1e-3, xmax=1, xlabel={$(\tau_f-t)/\tau_f$},
-	xmode=log, x dir=reverse,
-	ylabel={$\ell/H$},
-	ymin=0, ymax=0.5,
-	clip mode=individual,
-	axis background/.style={fill=white},
-	]
-	\addplot[only marks, Accent1, error bars/.cd, y dir=both, y explicit] table[y error index=2]{Y110_fracturelength.txt};
-	\addplot[Accent2, ultra thick, no marks, domain={1e-3:1}] {- 0.06 * ln(x)} node [pos=0.25,left=1em] {$\frac{\ell}{H}\sim \ln\frac{\tau_f-t}{\tau_f}$};
-\end{axis}
-\end{tikzpicture}
-\end{block}
-\end{columns}
-\end{frame}
 

### 2015-06-04 creep_beamer.tex
Brouillon de Jagla

@@
 \[\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}} = \underbrace{\lambda \left(\frac{t}{\tau_f}\right)^{-0.85}}_\text{linear response} + \underbrace{\frac{\mu}{1 - t/\tau_f}}_\text{fractures}\]
-\end{columns}
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{Failure time}
-\begin{columns}
-\column{0.55\textwidth}
-\tikzsetnextfilename{basquin}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xlabel={$\sigma$ (\si{\pascal})},
-	ylabel={$\tau_f$ (\si{\second})},
-	xmin=100, ymin=10, ymax=1e6,
-	xtick={100, 200,..., 1000}, xticklabels={100, 200,,, 500,,,,, 1000},
-	]
-	\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
-	\addplot[no marks, black, domain={100:1000}] {4.2e17*x^(-5.45)} node [midway, above right] {$\tau_f\sim \sigma^{-5.5}$};
-\end{loglogaxis}
-\end{tikzpicture}
-
-\begin{block}{Basquin law}
-\begin{itemize}
-\item fatigue (oscillatory stress)
-\item heterogeneous solids
-\item no yield stress
-\end{itemize}
-\end{block}
-
-\column{0.45\textwidth}
-\tikzsetnextfilename{basqun_otherfits}
-\begin{tikzpicture}
-\begin{groupplot}[
-	group style={
-		group name=g, group size=1 by 2,
-		},
-	width=\textwidth,
-	height=0.7\textwidth,
-	ymode=log,
-	ylabel={$\tau_f$ (\si{\second})},
-	ymin=10, ymax=1e6,
-	]
-	\nextgroupplot[
-		xlabel={$\sigma$ (\si{\pascal})},
-		xtick={200,600,1000},
-		]
-		\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
-		\addplot[no marks, black, domain={100:1000}] {6.5e5*exp(-x/84)} node [pos=0.25, above right] {$\tau_f\nsim e^{-\frac{\sigma}{\sigma_0}}$};
-
-	\nextgroupplot[
-		xlabel={$1/\sigma$ (\si{\per\pascal})},
-		]
-		\addplot[only marks, Accent1] table[x expr=1/\thisrowno{0}, y index=3]{creep_cas4_gdl1_MCR.txt};
-		\addplot[no marks, black, domain={1e-3:7e-3}] {20.3*exp(1720*x)} node [pos=0.25, below right] {$\tau_f\nsim e^\frac{\sigma_0}{\sigma}$};
-\end{groupplot}
-\end{tikzpicture}
-
-Activated processes unlikely (Eyring, Griffith, Taylor, \ldots)
-\end{columns}
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{Primary creep}
-\begin{columns}
-\column{0.55\textwidth}
-\tikzsetnextfilename{primary_creep}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	name=g,
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
-	ylabel={Strain rate $\dot{\gamma}$ (\si{\per\second})},
-	cycle list name=earthy,
-	no marks,
-	]
-	\addplot table[y index=2]{Y27_200Pa_gdot_decimated.txt} node (s200){};
-	\addplot table[y index=2]{Y38_300Pa_gdot_decimated.txt} node (s300){};
-	\addplot table[y index=2]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
-	\addplot table[y index=2]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
-	\addplot table[y index=2]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[Accent2, ultra thick, domain={0.1:1e3}] {0.01*x^(-0.85)} node[pos=0.75, below left] {$\dot{\gamma}(t)\sim t^{-0.85}$};
-\end{loglogaxis}
-\begin{scope}[anchor=base east, every node/.style={ rotate=90}]
-	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
-	\node[red!60!black] at (s300 |- g.outer north) {300};
-	\node[red!80!black] at (s400 |- g.outer north) {400};
-	\node[red] at (s550 |- g.outer north) {550};
-	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
-\end{scope}
-\end{tikzpicture}
-
-\begin{block}{Primary creep $\Rightarrow$ ``Andrade'' creep}
-Power-law creep known in \alert{solids} for over a century
-
-Classically explained by disinclination dynamics in crystals
-\end{block}
-
-\column{0.45\textwidth}
-\includegraphics[width=\textwidth]{Andrade_1910}
-
-\bigskip
-\includegraphics[height=7\baselineskip]{Andrade_1910_fig1}%
-\tikzsetnextfilename{Andrade_historical}%
-\begin{tikzpicture}
-\node[inner sep=0] (a){\includegraphics[height=7\baselineskip]{Andrade_1910_fig2}};
-\node[below right=0 and 1em of a.north west, font=\footnotesize,fill=white] {lead \& copper};
-\node[above left=1em of a.south east, Accent2] {$\gamma-\gamma_0 \sim t^{1/3}$};
-\end{tikzpicture}
-\end{columns}
-\end{frame}
-
-%\tikzset{external/force remake}
-\begin{frame}{Primary creep}
-\begin{block}{Local insight from \movie[externalviewer]{ultrasounds}{Yaourt110_primary.avi}}
-\begin{itemize}
-	\item homogeneous \movie[externalviewer]{strain field}{Yaourt110_profils.avi} (no wall slip, no shear band)
-	\item if present, plastic events are below our resolution (a few \si{\micro\metre})
-\end{itemize}
-\end{block}
-
-\begin{block}{Partial creep and recovery}
-\tikzsetnextfilename{recovery}
-\begin{tikzpicture}
-\begin{axis}[
-	width=\textwidth,
-	height=8\baselineskip,
-	xlabel={time (\si{\hour})},
-	ylabel={strain $\gamma$},
-	xmin=-0.15, xmax=7.5, xtick={0, 1, ..., 7},
-	axis background/.style={fill=white},
-	]
-	\addplot[black, no marks] table[x expr={\thisrowno{0}/3600}] {Y261_oscil_creep_recovery_100Pa_decimated.txt};
-	\begin{scope}[<-,Main, inner sep=0]
-	\draw (axis cs:0,0) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow \SI{100}{\pascal}$};
-	\draw (axis cs:1/6,0.2727007) -- +(1em,0) node[anchor=west]{$\sigma\leftarrow 0$};
-	\draw (axis cs:2.388,0.0159) -- +(0,1em) node[anchor=base east]{$\gamma\leftarrow 0$};
-	\end{scope}
-\end{axis}
-\end{tikzpicture}
-\end{block}
-
-$\Rightarrow$ Primary creep is ``kinetically'' reversible (no detectable structure damage, but viscous dissipation)
-\end{frame}
-
-%\tikzset{external/force remake}
-
-%\tikzset{external/force remake}
-
-%\tikzset{external/force remake}
-\begin{frame}{Secondary creep: a crossover}
-\begin{columns}
-\column{0.55\textwidth}
-Creep responses in linear scales
-
-\tikzsetnextfilename{secondary_rescaled}
-\begin{tikzpicture}
-\begin{axis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xmin=0, xmax=1, xlabel={$t/\tau_f$},
-	ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
-	ymin=0.5, ymax=4,
-	cycle list name=earthy,
-	restrict y to domain=0.5:10,
-	clip mode=individual,
-	]
-	\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt} node (s200){};
-	\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt} node (s300){};
-	\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt}  node (s400){};
-	\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt}  node (s550){};
-	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[yellow, ultra thick, domain={0.01:0.99}] {0.378*x^(-0.85) + 0.187/(1-x)};
-	\draw[<-] (axis cs:0.56,1) -- +(0,1em) node[above] {$\tau_\text{min}$};
-\end{axis}
-\end{tikzpicture}
-\begin{block}{Master curve}
-\[\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}} = \underbrace{\lambda \left(\frac{t}{\tau_f}\right)^{-0.85}}_\text{Andrade} + \underbrace{\frac{\mu}{1 - t/\tau_f}}_\text{fractures}\]
-\end{block}
-
-\column{0.45\textwidth}
-\structure{Monkman-Grant relation}
-\tikzsetnextfilename{Monkman-Grant}
-\begin{tikzpicture}
-\begin{loglogaxis}[
-	width=\textwidth,
-	height=0.8\textwidth,
-	xlabel={$\tau_f$ (\si{\second})},
-	ylabel={$\tau_\text{min}$ (\si{\second})},
-	xmin=10, xmax=1e6, ymin=10, ymax=1e6,
-	]
-	\addplot[only marks, Accent1] table[x index=3, y index=4]{creep_cas4_gdl1_MCR.txt};
-	\addplot[no marks, black, domain={10:1e6}] {0.56*x)};
-\end{loglogaxis}
-\end{tikzpicture}
-\[\tau_\text{min} \approx 0.6 \tau_f\]
-``Early time'' response allows to predict failure time
+$\Rightarrow$ no room for plasticity
 \end{columns}
@@
 \column{0.35\textwidth}
-\begin{block}{Discrete dislocation dynamics (dense)}
-\textit{\footnotesize Miguel et al., PRL 2002}
-
-\hspace{1em}\includegraphics[width=\textwidth-2em]{Miguel_2002_img}
-
-\medskip
-\hspace{1em}\includegraphics[width=\textwidth-2em]{Miguel_2002_graph}
-
-predicts Andrade creep \alert{but} no fracture
-\end{block}
-\column{0.65\textwidth}
-\begin{columns}
-\column{0.5\textwidth+1em}
 \structure{Fibre bundle model~\#1}
@@
 \textit{\footnotesize Jagla et al., PRE 2011}
+
+\begin{minipage}[b][10\baselineskip][c]{1em}
+$\dot\gamma$
+
+\vspace{4\baselineskip}
+$\dot\gamma$
+\end{minipage}\includegraphics[height=10\baselineskip, clip=true, trim=7mm 0 0 6cm]{Jagla_2011_3regimes}
+
 \begin{itemize}
@@
 
-\hspace{1em}\includegraphics[width=\textwidth-1em]{Jagla_2011_gamma}
-\[\tau_f\sim\left(\sigma-\alert{\sigma_c}\right)^{-1.25}\]
+%\hspace{1em}\includegraphics[width=\textwidth-1em]{Jagla_2011_gamma}
 
-\column{0.5\textwidth-1em}
-\includegraphics[width=\textwidth]{Jagla_2011_3regimes}
+\column{0.35\textwidth}
+\includegraphics[width=\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
 \end{columns}
-predicts Andrade creep and failure \alert{but} critical stress
-\end{columns}
+predicts Andrade creep and failure \alert{but} critical stress $\tau_f\sim\left(\sigma-\alert{\sigma_c}\right)^{-1.25}$
 \end{frame}

### 2015-06-05 creep_beamer.tex
Fiber bundle model Jagla

@@
 \end{tikzpicture}
-\[ \sigma \propto  \gamma f^\alpha \xrightarrow{\mathrm{Fourier}} \sigma \propto \gamma t^{-\alpha} \]
+\[ \boxed{\sigma \propto  \gamma f^\alpha} \xrightarrow{\mathrm{Fourier}} \boxed{\sigma \propto \gamma t^{-\alpha}} \]
 Constant $\sigma$
-\[ \gamma(t) \propto \sigma t^{\alpha} \xrightarrow{\frac{d}{dt}} \dot\gamma(t) \propto \sigma t^{\alpha-1}\]
+\[ \boxed{\gamma(t) \propto \sigma t^{\alpha}} \xrightarrow{\frac{d}{dt}} \boxed{\dot\gamma(t) \propto \sigma t^{\alpha-1}}\]
 
@@
 
-\begin{frame}{What about theories and models?}
-\begin{columns}
-\column{0.35\textwidth}
-\structure{Fibre bundle model~\#1}
+\begin{frame}{Fibre bundle models}
+\tikzsetnextfilename{FBM_Jagla}
+\begin{tikzpicture}
+\let\mydima\relax
+\newlength\mydima
+\setlength{\mydima}{0.3\textwidth}
+\let\mydimb\relax
+\newlength\mydimb
+\setlength{\mydimb}{2\baselineskip}
 
+\node[fill=black, inner sep=0, minimum height=0.2em, minimum width=\mydima] at (0,\mydimb)  (top) {};
+\node[fill=black, inner sep=0, minimum height=0.2em, minimum width=\mydima] at (0,-\mydimb)  (bot) {};
+
+\foreach \x in {0.1,0.2,...,0.9}\draw[Accent1, line width=0.1em*sin(1500*\x)+0.3em] (top.south west) ++(\x\mydima, 0) -- +(0, -2\mydimb+0.2em);
+
+\node[above=0 of top, single arrow, draw, shape border rotate=90] {$\sigma$};
+\node[starburst, draw=Accent2, inner sep=0pt,  minimum width=0.1\mydima, minimum height=0.5\mydimb, starburst point height=0.15em] at (-0.1\mydima,0) {};
+\node[double arrow,draw, anchor=west, text width=0.21\textwidth, align=center] at (0.5\mydima,-0.5\mydimb) (model) {\structure{model}};
+\node[above left=\baselineskip and 0of model.east, text width=0.31\textwidth, inner sep=0] {\begin{itemize}
+	\item elastic fibres
+	\item local yield strain
+	\item local coupling
+	\end{itemize}};
+\end{tikzpicture}
+\includegraphics[height=6\baselineskip, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
+
+\bigskip
+\begin{columns}[b]
+\column{0.3\textwidth}
 \textit{\footnotesize Jagla et al., PRE 2011}
@@
 
+\column{0.3\textwidth}
+\structure{But}
 \begin{itemize}
-\item elastic fibres
-\item local yield strain
-\item local coupling
+\item yield stress
+\item built in power-law creep
+\item no master curve
 \end{itemize}
 
-%\hspace{1em}\includegraphics[width=\textwidth-1em]{Jagla_2011_gamma}
+\column{0.4\textwidth}\tikzsetnextfilename{creep_1_3}
+\begin{tikzpicture}
+	\begin{groupplot}[%
+		group style={
+			group name=g, group size=1 by 2,
+			vertical sep=2.5em,
+			},
+		height=6.5\baselineskip,
+		width=9\baselineskip,
+		cycle list name=earthy,
+		clip mode=individual,
+		ylabel={$\dot{\gamma}/\dot{\gamma}_\text{min}$},
+		xlabel shift=-0.5em,
+		]
+	\nextgroupplot[
+		xmin=1e-5, xmax=2, xlabel={$t/\tau_f$},
+		xmode=log,ymode=log,
+		xtick={1e-4, 1e-2, 1},
+		]
+		\addplot table[x index=1, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*x^(-0.85) + 0.187/(1-x)};
 
-\column{0.35\textwidth}
-\includegraphics[width=\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
+	\nextgroupplot[
+		xmin=1e-5, xmax=2, xlabel={$(\tau_f-t)/\tau_f$},
+		x dir=reverse,
+		xmode=log,ymode=log,
+		xtick={1e-4, 1e-2, 1},
+	]
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y27_200Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y38_300Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y25_400Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y32_550Pa_gdot_decimated.txt};
+		\addplot table[x expr=1-\thisrowno{1}, y index=3]{Y39_1000Pa_gdot_decimated.txt};
+		\addplot[yellow, ultra thick, samples at={1e-5,1e-4, 1e-3,1e-2,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9, 0.99, 0.999, 0.9999, 0.99999}] {0.378*(1-x)^(-0.85) + 0.187/x};
+	\end{groupplot}
+\end{tikzpicture}
 \end{columns}
-predicts Andrade creep and failure \alert{but} critical stress $\tau_f\sim\left(\sigma-\alert{\sigma_c}\right)^{-1.25}$
 \end{frame}

### 2015-06-06 creep_beamer.tex
FBM Kun; Halaz

@@
 
-\begin{frame}{What about theories and models?}
-\structure{Fibre bundle model~\#2}
+\begin{frame}{Fibre bundle models}
+	\begin{columns}[T]
+	\column{0.27\textwidth}
+	\textit{\footnotesize Kun et al.  PRL 2008}
 
-elastic fibres + local yield strain + damage accumulation
-\begin{block}{\textit{\footnotesize Kun et al. J. Stat Mech 2006, PRL 2008}}
-\begin{columns}
-\column{0.7\textwidth}
-\tikzsetnextfilename{Kun_2006_basquin}
-\begin{tikzpicture}[inner sep=0]
-\node(a){\includegraphics[height=5\baselineskip]{Kun_2006_basquin.png}};
-\node[fill=white,yshift=-0.5em] at (a) {$\tau_f \sim \sigma^{-\beta}$};
-\end{tikzpicture}\,
-\includegraphics[height=5\baselineskip]{Kun_2008_Basquin}\,
-\includegraphics[height=5\baselineskip]{asphalt}
-\column{0.3\textwidth}
-$\beta$ exponent of damage accumulation
-\[\dot{\gamma}\sim \left(\tau_f-t\right)^{-1/(1+\beta)}\]
-\end{columns}
-\end{block}
+	\includegraphics[height=6\baselineskip]{Kun_2008_Basquin}
 
-\begin{columns}
-\column{0.65\textwidth}
-\begin{block}{\textit{\footnotesize Halász et al., PRE 2012}}
-\tikzsetnextfilename{Halasz}
-\begin{tikzpicture}[inner sep=0, ultra thick, font=\footnotesize]
-\node (g) {\includegraphics[height=7\baselineskip]{Halasz_2012_Basquin}};
-\node[draw=green!50!black, below right=0 and \baselineskip of g.north east] (c) {\includegraphics[height=4\baselineskip]{Halasz_2012_crack}};
-\node[draw=red, above right=0 and 6\baselineskip of g.south east] (d) {\includegraphics[height=4\baselineskip]{Halasz_2012_disordered}};
-\draw[->,green!50!black] (c.west) -- ++(-2\baselineskip,0) node[above] {$\beta=5$} node[below right=0 of c.north east, text width=5\baselineskip]{single crack propagation};
-\draw[->, red] (d.south west) -- +(-5\baselineskip,0) --($(g.south east) +(0, \baselineskip)$) node[below=1em] {$\beta=1$} node[above left=0 of d.south west] {diffusive damage};
+	\textit{\footnotesize Halász et al., PRE 2012}
+
+	\includegraphics[height=6\baselineskip]{Halasz_2012_crack}
+
+	\column{0.35\textwidth}
+	\begin{itemize}
+		\item elastic fibres
+		\item local yield strain
+		\item local coupling
+		\item[+] damage accumulation
+		\end{itemize}
+
+	\tikzsetnextfilename{model_arrow}
+	\begin{tikzpicture}
+	\node[double arrow,draw, anchor=west, text width=\textwidth-1\baselineskip, align=center] (model) {\structure{model}};
 	\end{tikzpicture}
-\end{block}
-\column{0.35\textwidth}
-Predicts Basquin law \& fractures \alert{but} no Andrade creep \& too slow divergence
+
+	\bigskip
+	\structure{But}
+	\begin{itemize}
+	\item no power-law creep 
+	\item too slow divergence
+	\item 2D fractures
+	\end{itemize}
+
+	\column{0.38\textwidth}
+	\tikzsetnextfilename{basquin_fracture}
+	\begin{tikzpicture}
+	\begin{loglogaxis}[
+		name=a,
+		width=\textwidth,
+		height=0.8\textwidth,
+		xlabel={$\sigma$ (\si{\pascal})},
+		ylabel={$\tau_f$ (\si{\second})},
+		xmin=100, ymin=10, ymax=1e6,
+		xtick={100, 200,..., 1000}, xticklabels={100, 200,,, 500,,,,, 1000},
+		]
+		\addplot[only marks, Accent1] table[y index=3]{creep_cas4_gdl1_MCR.txt};
+		\addplot[no marks, black, domain={100:1000}] {4.2e17*x^(-5.45)};
+	\end{loglogaxis}
+	\node[below = of a]{\includegraphics[height=6\baselineskip]{Y110_2013-03-01_03-21-15.jpg}};
+	\end{tikzpicture}
+
 	\end{columns}

### 2015-06-08 talk_Zurich.tex
Fiber bundle model Jagla
Strain in appendix Gallot 2013 The simplest yoghurt

@@
 \begin{frame}{Texture as important as taste}
+\tikzsetnextfilename{texture}%
 \begin{tikzpicture}[inner sep=0, callout pointer segments=3]
@@
 %\tikzset{external/force remake}
-\begin{frame}{Acid-set protein gel (Yoghurt)}
-\begin{block}{Recipe: take your time}
+\begin{frame}{The simplest yoghurt}
+\begin{block}{Acid-set protein gel}
 \begin{itemize}
@@
 
+\appendix
+\newcounter{finalframe}
+\setcounter{finalframe}{\value{framenumber}}
+
+\begin{frame}[plain]
+\end{frame}
+
+\begin{frame}{Creep rheology: Three regimes}
+\tikzsetnextfilename{three_regimes}
+\begin{tikzpicture}
+\begin{loglogaxis}[
+   name=g,
+	width=\textwidth,
+	height=0.8\textheight,
+	xmin=2e-2, xmax=4e5, xlabel={time (\si{\second})},
+	ymin=0.2,ymax=2,
+	ytick={0.2, 0.4,0.8,1.6}, yticklabels={0.2, 0.4,0.8,1.6},
+	ylabel={strain $\gamma$},
+	extra tick style={grid=major},%
+	extra y ticks={1}, extra y tick labels={1.0},
+	cycle list name=earthy,
+	axis on top,
+	]
+	\fill[Accent1!20]
+		(axis cs:10,0.4) ellipse[rotate=-30, x radius=0.2\textwidth, y radius=0.12\textwidth]
+		(axis cs:2e2,1.6) ellipse[x radius=0.35\textwidth, y radius=0.05\textwidth]
+		(axis cs:3e3,0.85) ellipse[rotate=-7.5, x radius=0.3\textwidth, y radius=0.07\textwidth]
+		;
+	\node[above right] at (axis cs:1e1,0.2) {primary};
+	\node[below left] at (axis cs:4e5,0.7) {secondary};
+	\node[left] at (axis cs:5,1.6) {tertiary};
+	\addplot table{Y27_200Pa_gamma_decimated.txt} node (s200){};
+	\addplot table{Y38_300Pa_gamma_decimated.txt} node (s300){};
+	\addplot table{Y25_400Pa_gamma_decimated.txt}  node (s400){};
+	\addplot table{Y32_550Pa_gamma_decimated.txt}  node (s550){};
+	\addplot table{Y39_1000Pa_gamma_decimated.txt} node (s1000){};
+\end{loglogaxis}
+\begin{scope}[anchor=base, every node/.style={yshift=0.2em}]
+	\node[red!40!black] at (s200 |- g.outer north) {\SI{200}{\pascal}};
+	\node[red!60!black] at (s300 |- g.outer north) {300};
+	\node[red!80!black] at (s400 |- g.outer north) {400};
+	\node[red] at (s550 |- g.outer north) {550};
+	\node[red!80!yellow] at (s1000 |- g.outer north) {1000};
+\end{scope}
+\end{tikzpicture}
+
+Failure at $\gamma\approx 1$ for a well defined time $\tau_f$
+\end{frame}
+
+\setcounter{framenumber}{\value{finalframe}}
 \end{document}

### 2015-07-20 fig_plis/article_SI.tex
Use Science citation style
Response to Review

@@
-\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,
-amsmath,amssymb,prl]{revtex4-1}
+\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers, amsmath,amssymb,prl]{revtex4-1}
 \usepackage[utf8]{inputenc}
@@
 Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated --the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
-This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, solvent is expelled through both the upper and lower sides of the gel film.
+This effect is the fingerprint of syneresis: as gelation proceeds, solvent is expelled through both the upper and lower sides of the gel film.
 
@@
 In Figure~\ref{fig:sideview}d, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Materials for the definition.
-We observe that the velocity peaks at different times corresponding to syn\ae{}resis and to the respective growths of the primary, secondary and tertiary patterns.
+We observe that the velocity peaks at different times corresponding to syneresis and to the respective growths of the primary, secondary and tertiary patterns.
 We associate each growth with a characteristic time $\tau = A_\text{max}/v_\text{max}$.
@@
 	\caption{3D analysis of the wrinkling process.
-(a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
+(a) Confocal $(x,z)$ cuts showing syneresis, swelling, wrinkling and cascade buckling.
 The scale bar is \SI{100}{\micro\metre} (real size ratio).
@@
 (i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat, as expected for chemical gels~\cite{Kim2010}.
-Since this case is not observed it means that in this physical gel, syn\ae{}resis and swelling processes foster irreversible stress dissipation and structural changes.
+Since this case is not observed it means that in this physical gel, syneresis and swelling processes foster irreversible stress dissipation and structural changes.
 In the following, we will therefore consider that the shrunk state is stress-free.
@@
 This simple buckling situation leads to a primary pattern composed of a single bump of size $L$.
-As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}.
+As the bump amplitude grows larger \textcolor{blue}{and} saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}.
 This is also not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
@@
 
-Therefore at this stage, our observations still raise two unresolved issues which we address in the next sections: (i) what triggers the the spontaneous syn\ae{}resis and swelling back processes?
+Therefore at this stage, our observations still raise two unresolved issues which we address in the next sections: (i) what triggers the the spontaneous syneresis and swelling back processes?
 And (ii) what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
@@
 
-\subsection*{Syn\ae{}resis and swelling back mechanism}
+\subsection*{Syneresis and swelling back mechanism}
 
@@
 (d) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
-allowed syn\ae{}resis and swelling (orange dashed line).}%
+allowed syneresis and swelling (orange dashed line).}%
 	\label{fig:acidification}
@@
 
-The syn\ae{}resis and swelling back mechanism depends on (i)~the pH sensitivity of caseins, (ii)~the route to gelation, (iii)~the gel properties and (iv)~the boundary conditions.
+The syneresis and swelling back mechanism depends on (i)~the pH sensitivity of caseins, (ii)~the route to gelation, (iii)~the gel properties and (iv)~the boundary conditions.
 
@@
 In order to turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes, see Materials and Methods, effectively creating non adhesive boundary conditions for the gel.
-When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall.
-In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d.
+\textcolor{blue}{When the casein adhesion to the upper boundary is turned off, the gel detaches from the top wall.
+However the gel does not wrinkle because it remains attached to the bottom wall.} In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d.
 $\xi$ remains mostly constant and variations manifests themselves at the macroscopic level, on the volume occupied by the gel.
 
-Therefore, the syn\ae{}resis and swelling back mechanism is related to charge stabilisation of the casein proteins as function of the pH.
-Such a mechanism is only effective to produce wrinkles when adhesion of the caseins to both top and bottom walls is turned off.
+Therefore, the syneresis and swelling back mechanism is related to charge stabilisation of the casein proteins as function of the pH.
+\textcolor{blue}{In particular, beyond the isoelectric point, the casein are charged and repel each other, leading to a swelling at every scale: proteins are further from each other, the strands become larger and longer and the gel sheet swells.} Such a mechanism is only effective to produce wrinkles when adhesion of the caseins to both top and bottom walls is turned off.
 Consistently with our explanation, the amount of GDL, and thus the final pH reached by the suspension, controls the amount of swelling back.
@@
 Large deformations of thin sheets are governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form.
-However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern up to dimensionless prefactors and isolate the physically meaningful ingredients~\cite{Cerda2003}.
+However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern up to dimensionless prefactors and isolate the physically meaningful ingredients~\cite{Cerda2003,Davidovitch2011}.
 An elastic film buckles due to excess area with respect to its boundaries.
@@
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Furthermore, it was shown that equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+Furthermore, it was shown that equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from capillary forces~\cite{Huang2007}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
 However, we have excluded these various origins for $\sigma_{\perp}$ in the discussion around Figure~\ref{fig:dilation}.
@@
 
-For a typical gel film, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+For a typical gel film, the buoyancy after syneresis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
 This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ which confirms that gravity can be neglected as already inferred from test experiments where the cell is held vertically rather than horizontally.
@@
 The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ over the whole $H/H^*$ range.
+
+\textcolor{blue}{Despite a very good agreement of the scaling, the exact prefactors of all models disagree by a factor 0.63-0.69.
+We suggest that this discrepancy comes from the fact that our model is 2D, i.e.
+uniaxial compression and deflection, whereas our system is 3D, i.e.
+biaxial compression and deflection.
+Since our system is isotropic in the plane, this does not change the scaling but might account for a dimensional factor.}
 
@@
 
-\bibliographystyle{pnas2009}
+\bibliographystyle{Science}
 \bibliography{Yaourt}
@@
 
-Figure~\ref{fig:acidification}d of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c-e show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
+Figure~\ref{fig:acidification}d of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c-e show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syneresis and swelling; (ii) complete adhesion except on the top wall, allowing syneresis and swelling.
 The latter situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
@@
 Left of Supplementary Figure~\ref{fig:nonmonotonic}a shows that the arms of the gel first shrink laterally while keeping their length, and then swell again.
-Syn\ae{}resis and swelling are thus local processes.
+Syneresis and swelling are thus local processes.
 By contrast in situation (ii), the gel as a whole compresses while forming.

### 2015-07-20 fig_plis/article_plis.tex
Split supplementary

--- 2015-02-18 fig_plis/article_plis.tex
@@
 \begin{abstract}
-The slow acidification of caseins, a family of milk proteins, is well known to result in the formation of gels, e.g.
-yoghurt.
-Less known is the partial redissolution of the gel at even lower pH.
-Here, using caseins as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, we design a model biogel film that spontaneously wrinkles in cascade, provided non-adhesive confinement.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
-Taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
-Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilisation/destabilisation.
-As such our results could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
+Confined thin surfaces may wrinkle due the growth of an excess of material. Usually the wavelength is set by elasticity or gravity, but here we explore new selection mechanisms based on hydrodynamics.
+First, inspired by yoghurt-making processes, we use caseins, a family of milk proteins, as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, to design a porous biogel film immersed in a confined buoyancy-matched viscous medium.
+Provided specific boundary conditions, but without any external stimulus, the biogel film spontaneously wrinkles in cascade.
+Second, using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
+Third, taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
+Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilisation/destabilisation, and therefore could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
 \end{abstract}
@@
 \maketitle
+
+%\paragraph*{Significance statement}
+%Confined thin surfaces may wrinkle due the growth of an excess of material.
+Here we present benchmark experiments that explore the possibility of wrinkling in biogels.
+We relate the excess area to the microstructural dynamics of the biogel.
+Usually the wrinkling wavelength is set by elasticity or gravity.
+In our case the wavelength is controlled by viscous flow, either in a lubrication layer or through the pores of the gel.
+Our results pave the way for exploring the possibility of wrinkling in porous soft materials floating in a liquid such as biological tissues: during morphogenesis of the embryo the development of confined permeable and wrinkled surface is ubiquitous.
 
@@
 Moreover, most biological films are porous, and little is known on the interplay between permeability and wrinkling~\cite{Ma2004,Longley2013}.
+The main goal of the present study is therefore threefold: to create a buoyancy-matched material that wrinkles, to unveil the physical origin of the observed patterns and to identify the selection mechanism of the wavelength.
 
 Here, we introduce model permeable biogels that produce hierarchical wrinkling as a result of their interplay between their gelation dynamics and the confinement conditions. 
-In slit geometry, we slowly acidify a dispersion of caseins, a family of milk proteins.
-As a result, not only does the casein form a gel as expected~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002} --this is a well-known effect used in yoghurt making-- but the gel starts to produce wrinkling patterns in cascade when casein adhesion to the slit walls is turned off, Figure~\ref{fig:dynamics} and Supplementary Videos~\ref{S-vid:transmitted} and \ref{S-vid:reconstructed}.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that upon acidification the casein network spontaneously shrinks then swells at the micron level and wrinkles at the millimetre scale.
-We systematically investigate such swelling-induced wrinkling of a porous gel film sandwiched between two viscous layers, the whole system being confined between two glass plates.
-The main goal of the present study is to unveil the physical origin of the observed patterns and to identify the selection mechanism of the wavelength.
+As a biogel, we use a simplified yoghurt obtained by the slow acidification of caseins, a family of milk proteins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+While still liquid, the casein dispersion is injected in a home-made optical cell that is then sealed.
+This cell is designed such that the casein adhesion to both top and bottom walls is turned off and so that the gel is only anchored to the remaining four vertical sides of the cell.
+%
+Although no external stimulus is exerted on the system, we observe the spontaneous formation of a wrinkling pattern, as shown in Figure~\ref{fig:dynamics} and Supplementary Videos~S\ref{S-vid:transmitted} and S\ref{S-vid:reconstructed}.
+This pattern shows unique properties: a primary pattern appears simultaneously throughout the system with a wavelength $\lambda$  that is much smaller than the cell width $L$ and does not evolve with time.
+Furthermore, within this primary pattern, secondary and ternary patterns appear in cascade.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that the wrinkles at the millimetre scale result from the spontaneous swelling of the casein network at the micron level upon acidification.
+%
+Finally we show that the wavelength is selected by a purely viscous mechanism combining porous Darcy flow and viscous Poiseuille flow.
+We systematically checked that this combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength under a wide range of conditions.
 
@@
 	\caption{Dynamics of pattern formation in a confined film of casein gel (caseinate 4\%w, GDL 4\%w in water).
-(a) Light transmission microscopy.
+(a) Sketch of the cell where the adhesion on both top and bottom walls is turned off.
+The cell is sealed and the gel is only anchored to the four sides.
+Typical dimensions are $L>\SI{1}{\centi\metre}$, $e\approx\SI{100}{\micro\metre}$.
+(b) Light transmission microscopy.
 Successive generations of patterns are highlighted in colour to stress the absence of coarsening after formation.
-(b) 3D reconstruction from fluorescent confocal microscopy.
+The successive wavelengths are $\lambda=\SI{1.5}{\milli\metre}$ (yellow), $\lambda=\SI{0.75}{\milli\metre}$ (orange) and $\lambda=\SI{0.32}{\milli\metre}$ (brown).
+(c) 3D reconstruction from fluorescent confocal microscopy which highlights that patterns observed in (b) correspond to wrinkles.
+The contrast in (b) is not due to thickness inhomogeneities but to altitude gradients as indicated by the yellow vertical lines in (a).
 Scale bars are \SI{1}{\milli\metre}.}%
@@
 
-In the following we first describe the physical chemistry that allows us to form such patterns.
-We then show how transverse confinement leads to saturation of the amplitude on both the top and bottom walls followed by a cascade of secondary and ternary patterns.
-Finally we model the situation and identify two possible selection mechanisms, namely porous Darcy flow or viscous Poiseuille flow.
-Comparing our experimental data to these two models we show that a combination of both Darcy and Poiseuille mechanisms accurately describes our data.
-This combined model pins down the dynamical origin of the constraints exerted on the gel and nicely predicts the wrinkling wavelength.
-%Poiseuille~\cite{Poiseuille1842} flows of the solvent above and below the gel film for long wavelengths and a new mechanism, Darcy~\cite{Darcy1856} flows of the solvent through the gel film for short wavelengths.
+\section*{Results and discussion}
 
-\section*{System design}
+\subsection*{Dynamics of the wrinkling process}
+
+To form a biogel we start from an aqueous dispersion of caseins and glucono-$\delta$-lactone (GDL).
+GDL slowly acidifies the medium and triggers gelation of the caseins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+While still liquid, this yoghurt-forming suspension is injected into a home-made optical cell, as sketched in Figure~\ref{fig:dynamics}a.
+The cell is sealed and no external stimulus is applied during gelation or pattern formation.
+Both top and bottom walls are coated with acrylamide brushes, effectively turning off the casein adhesion, so that the gel is only anchored to the remaining four vertical sides of the cell.
+
+The dynamics of the pattern formation are followed by light transmission microscopy and confocal microscopy, see Materials and Methods.
+As shown in Figure~\ref{fig:dynamics}b and Supplementary Video~S\ref{S-vid:transmitted}, we observe the spontaneous formation of a first pattern that appears with a characteristic wavelength $\lambda\ll L$.
+The 2D shape and size of this primary pattern are set as soon as the pattern forms and do not evolve in time.
+However inside the primary pattern a secondary pattern emerges, inside which a ternary forms.
+At each step of this Russian dolls-like cascade, the wavelength is divided by two, as shown in Supplementary Figure~S\ref{S-fig:generations}.
+
+3D confocal microscopy, as shown in Figure~\ref{fig:dynamics}c and Supplementary Video~S\ref{S-vid:reconstructed}, reveals that those patterns actually correspond to wrinkles of the biogel film.
+Figure~\ref{fig:sideview}a shows successive side views of the same confocal measurements as in  Figure~\ref{fig:dynamics}c, see also Supplementary Videos~S\ref{S-vid:cutDarcy} and S\ref{S-vid:cutPoiseuille}.
+Initially the dispersion is homogeneous and fills the entire slit.
+The black margins in Figure~\ref{fig:sideview}a correspond to the glass walls.
+Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated --the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
+This effect is the fingerprint of syneresis: as gelation proceeds, solvent is expelled through both the upper and lower sides of the gel film.
+
+Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}b, and excess surface area of the gel film, Figure~\ref{fig:sideview}c, see Supplementary Materials.
+The evolution of the gel volume is surprisingly non-monotonous.
+After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
+Shortly after, at \SI{36}{\minute}, the planar film destabilises and starts to wrinkle in the $z$-direction creating excess area.
+This confirms that the instability leading to the primary pattern is due to swelling.
+Around \SI{38}{\minute}, the amplitude of the wrinkles reaches the thickness of the slit.
+As swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
+When the excess area saturates, the dynamics stop (\SI{3}{\hour}).
+We note that at any given point in time the thickness $h$ of the film of gel is spatially homogeneous.
+
+The primary, secondary and tertiary patterns follow the same path: the birth of an instability of wavelength $\lambda$ of growing amplitude $A(t)$ until saturation to $A_\text{max}$, see Figure~\ref{fig:sideview}a.
+Throughout the whole process, $\lambda$ remains constant.
+This path is dynamic in essence.
+In Figure~\ref{fig:sideview}d, we thus report the velocity $v(t)$ of the wrinkles as it develops in the $z$-direction, see Supplementary Materials for the definition.
+We observe that the velocity peaks at different times corresponding to syneresis and to the respective growths of the primary, secondary and tertiary patterns.
+We associate each growth with a characteristic time $\tau = A_\text{max}/v_\text{max}$.
+For example, in the present case the $\tau$ corresponding to the primary pattern is $\SI{123}{\second}$.
+
+Therefore contrary to observations in, e.g., Ref~\cite{Yoo2003}, we do not report any coarsening of the patterns.
+We believe that the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}.
+Indeed when the amplitude saturates the gel flattens against the walls which creates sharp edges.
+To accommodate those sharp edges with finite bending rigidity, the flat part of the gel actually ripples, see Supplementary Figure~S\ref{S-fig:ripples}.
+The ripples created by opposite sides of a same flat patch interfere, creating barriers against coarsening~\cite{LeGoff2014}.
+
+\begin{figure}
+	\includegraphics{sideview}
+	\caption{3D analysis of the wrinkling process.
+(a) Confocal $(x,z)$ cuts showing syneresis, swelling, wrinkling and cascade buckling.
+The scale bar is \SI{100}{\micro\metre} (real size ratio).
+(b-d) Confocal microscopy measurement of the evolution of the volume of the gel phase relative to cell volume, of the excess area and of the velocity along the $z$-direction.
+Crosses correspond to times in (a).}
+	\label{fig:sideview}
+\end{figure}
+
+Based on those observations, we examine possible scenarios upon shrinking and swelling of the gel.
+As the gel forms it expels the solvent and detaches from top and bottom wall, see Figure~\ref{fig:dilation}b and c.
+For the swelling back we consider three options:
+
+\noindent
+(i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat, as expected for chemical gels~\cite{Kim2010}.
+Since this case is not observed it means that in this physical gel, syneresis and swelling processes foster irreversible stress dissipation and structural changes.
+In the following, we will therefore consider that the shrunk state is stress-free.
+
+\noindent
+(ii) In Figure~\ref{fig:dilation}e, we assume that nothing hinders the buckling of the gel film.
+This simple buckling situation leads to a primary pattern composed of a single bump of size $L$.
+As the bump amplitude grows larger \textcolor{blue}{and} saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}.
+This is also not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.
+
+\noindent
+(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom walls.
+To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load $\sigma_{\perp}$ that hinders buckling on large wavelengths.
+The transverse load could be due to gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, however we confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
+with the weight acting longitudinally rather than transversely.
+Furthermore, the transverse load cannot be adhesion~\cite{Vella2009a} since the film is not adhering to the top and bottom walls.
+%
+The binding of a longitudinal edge to a rigid substrate can induce wrinkles upon compression~\cite{Vandeparre2011a, Li2013}.
+However in that case, the wavelength gets larger away from the bound edge.
+Since the wavelength of our pattern is uniform throughout the sample we exclude the influence of the side boundaries.
+%
+%Influence of the side boundaries~\cite{Vandeparre2011a, Li2013} is also excluded since the wavelength of the pattern is uniform throughout the sample. 
+%
+Finally as the four side boundaries remain fixed $\sigma_{\perp}$ cannot come from the resistance to uniaxial stretching~\cite{Cerda2003}.
+
+Therefore at this stage, our observations still raise two unresolved issues which we address in the next sections: (i) what triggers the the spontaneous syneresis and swelling back processes?
+And (ii) what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
+
+\begin{figure*}
+\includegraphics{dilation.pdf}
+\caption{The wrinkling experiment.
+(a) Initial configuration where the sealed cell contains a homogeneous protein solution.
+(b) Around the isoelectric pH, the gel forms and immediately expels solvent leading to (c).
+If tensile stresses were not released we would observe reversible swelling back leading to (d) a flat swollen gel layer.
+If tensile stresses are released, swelling back leads either to buckling (e) or to wrinkling (f) depending on whether the bending is free or hindered by a transverse load.}
+\label{fig:dilation}
+\end{figure*}
+
+\subsection*{Syneresis and swelling back mechanism}
 
@@
 Symbols are direct measures upon addition of GDL.
-The continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (insert).
+The continuous line is deduced from $x_\text{free}$ at the same pH obtained by addition of HCl 1 molar and waiting for equilibration (inset).
 (c) Evolution of the elastic modulus $G'$ measured in a rheometer with full adhesion to the cone-plane geometry.
 (d) Evolution of pore size measured by confocal microscopy in a slit geometry with either adhesion to all walls (black line) or no adhesion to top wall, i.e.
-allowed syn\ae{}resis and swelling (orange dashed line).}%
+allowed syneresis and swelling (orange dashed line).}%
 	\label{fig:acidification}
@@
 
-To form wrinkled patterns such as the one observed in Figure~\ref{fig:dynamics}, ones needs (i) to take advantage of the pH sensitivity of caseins, (ii) to control the route to gelation, (iii) to fine tune the gel properties and the boundary conditions.
+The syneresis and swelling back mechanism depends on (i)~the pH sensitivity of caseins, (ii)~the route to gelation, (iii)~the gel properties and (iv)~the boundary conditions.
 
 The stability of casein dispersions strongly depends on the pH.
-Supplementary Video~\ref{S-vid:titration} shows the titration of a casein dispersion at 4\%w by 1 molar HCl.
+Supplementary Video~S\ref{S-vid:titration} shows the titration of a casein dispersion at 4\%w by 1 molar HCl.
 At pH=7, the caseins are well dispersed then they flocculate between pH=5.5 to 4 and finally at pH=3 the solution becomes clearer as caseins partially redisperse.
@@
 At neutral pH caseins are charged negatively and repel each other.
-As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), casein molecules loose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
+As the pH decreases to the isoelectric point of the caseins (pH$\approx 4.6$), casein molecules lose their negative net charge and tend to stick to each other~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}.
 Beyond the isoelectric point caseins acquire a positive net charge and therefore partially re-dissolve.
-In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH, see Supplementary Methods.
+In the inset of Figure~\ref{fig:acidification}b, we quantify this effect and measure with UV absorption spectroscopy the amount of free caseins within the 4\%w dispersion as a function of pH, see Supplementary Materials.
 Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction $x_\text{free}$ is 1.
@@
 To gain further insight into the mechanical properties of the gel we turn to microstructural observations under confocal microscopy.
-As shown in Supplementary Videos~\ref{S-vid:stick63} and \ref{S-vid:nostick63} and in Supplementary Figure~\ref{S-fig:nonmonotonic} the gel consists in a space spanning fractal network with a cut-off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
-Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~\ref{S-fig:permeability}.
-The fractal model fits the structure at all times but its parameters vary non-monotonically during GDL acidification.
-As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~\ref{S-fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
+As shown in Supplementary Videos~S\ref{S-vid:stick63} and S\ref{S-vid:nostick63} and in Supplementary Figure~S\ref{S-fig:nonmonotonic} the gel consists in a space-spanning fractal network with a cut-off length scale $\xi$ that corresponds to the maximum pore size~\cite{VanDijk1986}.
+Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~S\ref{S-fig:permeability}.
+The fractal model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
+As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~S\ref{S-fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
 This corresponds to a tighter packing of the caseins within the network strands.
@@
 
-%XXXCe paragraphe fait-il sens ici ?XXX As seen on the confocal measurements, the gel is porous and is permeable to the solvent.
-This property is determinant as it allow the gel to absorb and expel the solvent and therefore change its volume when sticky boundary are turned off.
-To quantify this effect, we make the casein gel inside a capillary and dip vertically the capillary in an aqueous buffer~\cite{VanDijk1986}.
-In the Supplementary Figure~\ref{S-fig:permeability}, we observe that the aqueous buffer goes through the gel and rises in the tube due to capillary forces.
-Exploitation of this experiment give the permeability of the gel: \SI{67000}{\square\nano\metre}.
+Boundary conditions is the final parameter that is essential to form wrinkled patterns.
+In order to turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes, see Materials and Methods, effectively creating non adhesive boundary conditions for the gel.
+\textcolor{blue}{When the casein adhesion to the upper boundary is turned off, the gel detaches from the top wall.
+However the gel does not wrinkle because it remains attached to the bottom wall.} In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d.
+$\xi$ remains mostly constant and variations manifests themselves at the macroscopic level, on the volume occupied by the gel.
 
-Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-In order to turn off the adsorption of the caseins on a surface, we coat the surface with polymer brushes, see Methods, effectively creating non adhesive boundary conditions for the gel.
-When the casein adhesion to the upper boundary is turned off, the gel still not wrinkles because it remains attached to the bottom wall, but it detaches from the top wall.
-In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d.
-$\xi$ remains mostly constant but the volume occupied by the gel varies, as will be reported below.
+Therefore, the syneresis and swelling back mechanism is related to charge stabilisation of the casein proteins as function of the pH.
+\textcolor{blue}{In particular, beyond the isoelectric point, the casein are charged and repel each other, leading to a swelling at every scale: proteins are further from each other, the strands become larger and longer and the gel sheet swells.} Such a mechanism is only effective to produce wrinkles when adhesion of the caseins to both top and bottom walls is turned off.
+Consistently with our explanation, the amount of GDL, and thus the final pH reached by the suspension, controls the amount of swelling back.
+Thus at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
 
-Finally, if the adhesion to both top and bottom walls is turned off, the film wrinkles as shown in Figure~\ref{fig:dynamics}.
-Such cascade buckling has already been observed in a one-dimensional macroscopic system, a sheet of steel forced to buckle between two solid plates~\cite{Roman1999}.
-In addition one has to note that the primary wavelength is much smaller than the size of the system, implying that a simple buckling scenario cannot explain the observations in Figure~\ref{fig:dynamics} and that only a wrinkling instability can be invoked.
-
-%We coat the glass slides with acrylamide brushes to prevent the adhesion of the casein.
-In fig.2c, when the top slide of the optical cell is coated with acrylamide brushes, we then observe that the casein network detaches from the top slide.
-$\xi$ does not peak at the isoelectric point but remains constant
-
-%Boundary conditions is the final parameter that is essential to form wrinkled patterns.
-Once the space spanning network is formed and sticks to the container, protein attraction can only proceed by local shrinking of the protein agregates, leading to larger pores.
-
-%The difference in value of $\xi$ between the fully un-coated  and the half-coated optical cell demonstrate that the gel is subject to constrains as the pH decreases.
-The relaxation of thoses constrain is key to the wrinkling formation when both the top and bottom slide are coated with acrylamide brushes.
-
-%on a change les condition daccroche ce aui permet au contrainte de se relacher
-
-%The slow acidification of a protein (sodium caseinate) solution down to the isoelectric pH induces gelation~\cite{Bremer1989} (Fig.~\ref{fig:acidification}a).
-Since caseins strongly adsorb to interfaces, the resulting gel sticks to the container walls, allowing rheological measurements (Fig.~\ref{fig:acidification}b).
-Further decrease of the pH beyond the isoelectric point makes the casein aggregates swell (Supplementary Fig.~\ref{S-fig:nonmonotonic}), thus weakening the gel~\cite{Braga2006}.
-In a microscopy cell of slit geometry coated so that the casein cannot adsorb on the top wall, this microscopic nonmonotonic behaviour translates into little variation of the pore size $\xi$ but to solvent expulsion out of the gel phase (syn\ae{}resis) followed by global swelling (Fig.~\ref{fig:acidification}c and Supplementary Fig.~\ref{S-fig:nonmonotonic}).
-
-\section*{Dynamics of the wrinkling process}
-
-Figure~\ref{fig:sideview}a shows successive side views of confocal measurements in the case where both top and bottom walls are coated, adhesion turned off, a configuration that induces wrinkling, see also Supplementary Videos~\ref{S-vid:cutDarcy} and \ref{S-vid:cutPoiseuille}.
-Initially the dispersion is homogeneous and fills the entire slit.
-The black margins correspond to the glass walls.
-Around \SI{23}{\minute}, the gel forms and simultaneously becomes thinner and more concentrated --the fluorescence signal which is proportional to the casein concentration is brighter compared to the initial situation.
-This effect is the fingerprint of syn\ae{}resis: as gelation proceeds, the solvent is expelled on both the upper and lower sides.
-We describe this situation as a porous elastic film of thickness $h$ sandwiched between two viscous layers of respective thickness $H_1$ and $H_2$ in a cell of thickness $e=H_1+h+H_2$,  Figure~\ref{fig:sideview}b.
-Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}c, and excess surface area of the gel film, Figure~\ref{fig:sideview}d, see Supplementary Methods.
-As expected from the system design, the evolution of the gel volume is non-monotonic.
-After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
-Shortly after,  at \SI{36}{\minute}, the planar film destabilises and starts to wrinkle in the $z$-direction creating excess area.
-This confirms that the instability is due to swelling.
-Around \SI{38}{\minute}, the amplitude of the wrinkles reaches the thickness of the slit.
-The size of the primary pattern is now fixed, Supplementary Figure~\ref{S-fig:growth}.
-However swelling continues, the gel flattens on both top and bottom walls (\SI{44}{\minute}) and finally buckles back (\SI{53}{\minute}) to form the secondary pattern (\SI{1}{\hour}21).
-When the excess area saturates, the dynamics stop (\SI{3}{\hour}).
-The amount of GDL controls the ability of the gel to swell: at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
-
-Note that adhesion to side walls is necessary to prevent total syn\ae{}resis and thus allows subsequent planar stresses to develop while swelling: the only situation where we observe wrinkling is when the adhesion on both top and bottom wall is turned off.
-We also confirm that at any given point in time the thickness $h$ of the film of gel is spatially homogeneous.
-The contrast in transmission microscopy, Figure~\ref{fig:dynamics}a, is therefore not due to thickness inhomogeneities but to altitude gradients.
-Indeed the length of the optical path along the vertical direction $z$ through the gel phase is larger where the gel makes a slope compared to a flat situation, as indicated by the yellow vertical lines in Figure~\ref{fig:sideview}b.
+\subsection*{Wavelength selection mechanism}
 
 \begin{figure}
-	\includegraphics{sideview}
-	\caption{Measurements and modelling.
-(a) Confocal $(x,z)$ cuts showing syn\ae{}resis, swelling, wrinkling and cascade buckling.
-The scale bar is \SI{100}{\micro\metre} (real size ratio).
-(b) Schematic side view of the cell.
-The constant thickness gel film (orange) is surrounded by solvent.
-Yellow lines show the path of transmitted light through the gel phase.
-(c-d) Confocal microscopy measurement of the volume of the gel phase relative to cell volume and excess area.
-Crosses correspond to times in (a).}
-	\label{fig:sideview}
+\includegraphics{sketch_Darcy_Poiseuille.pdf}
+\caption{Sketch of the two limit scenarios for the wavelength selection.
+In both cases the gel film of thickness $h$ is destabilised by the excess area which gives rise to wrinkles of amplitude $A(t)$ and wavelength $\lambda$.
+A single wavelength is represented and the longitudinal dimension $L$ of the system is much larger than $\lambda$.
+The value of $\lambda$ is set by the interplay between the bending rigidity $B$ and a transverse load $\sigma_\perp$ due to a pressure gradient $p_2-p_1$ in the solvent of viscosity $\eta$.
+(a) Darcy scenario: the gel initially sits without sticking to the bottom wall.
+$\sigma_\perp$ is due to the flow of the solvent through the porous gel of permeability $\alpha$ to fill the growing blister.
+(b) Poiseuille scenario: the gel film of negligible porosity initially lies in the middle of the cell, separated by a distance $H$ from each wall.
+$\sigma_\perp$ is due to the lubrication flow in the top and bottom solvent layers.}
+\label{fig:sketchDP}
 \end{figure}
-
-\section*{Models and scaling laws}
 
@@
 The determination of the characteristic length $\lambda$ of a wrinkling pattern is an old problem.
-Large deformations of thin sheets is governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form.
-However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern up to dimensionless prefactors and isolate the physically meaningful ingredients~\cite{Cerda2003}.
+Large deformations of thin sheets are governed by F\"{o}ppl-von K\'arm\'an equations~\cite{landau1986}, which are essentially impossible to solve in analytical form.
+However using scaling and asymptotic arguments one can obtain the dominant wavelength of the wrinkling pattern up to dimensionless prefactors and isolate the physically meaningful ingredients~\cite{Cerda2003,Davidovitch2011}.
 An elastic film buckles due to excess area with respect to its boundaries.
@@
 Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
-Furthermore, equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
-
-In our situation we now have to determine $\sigma_\perp$.
-We identify three possible candidates for the transverse load $\sigma_\perp$: the weight or buoyancy of the gel, the porous flow through the gel or the viscous flows in the surrounding solvent layers.
-We confirmed that gravity plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e.
-with the weight acting longitudinally rather than transversely.
+Furthermore, it was shown that equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from capillary forces~\cite{Huang2007}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
+However, we have excluded these various origins for $\sigma_{\perp}$ in the discussion around Figure~\ref{fig:dilation}.
+%
+The only possible explanation left for $\sigma_{\perp}$ involve hydrodynamic stresses due to viscous flow of the solvent, either through the gel porous matrix itself or in the surrounding solvent layers.
 
@@
 
-Therefore we consider two limit scenarios that leads to wrinkling patterns.
+Accordingly we consider the two limit scenarios sketched in Figure~\ref{fig:sketchDP} that lead to wrinkling patterns.
 In a first limit, the film of gel sits without sticking to the bottom wall of the cell, i.e.
-$H_1\rightarrow0$, see Supplementary Video~\ref{S-vid:cutDarcy}.
+$H_1\rightarrow0$, see Figure~\ref{fig:sketchDP}a and Supplementary Video~S\ref{S-vid:cutDarcy}.
 Creating a two dimensional blister requires flowing solvent of viscosity $\eta$ through the gel of permeability $\alpha$.
@@
 \begin{equation}
-\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4}
+\lambda_D \sim h^{1/2} \alpha^{1/4} \Upsilon^{1/4},\quad\text{with } \Upsilon = \frac{B\tau}{\eta}.
 \end{equation}
-where $\Upsilon = B\tau/\eta$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
-This expression emphasises the kinetic nature of the wavelength selection through Darcy flow.
-However we note that the wavelength of a generation of the pattern does not evolve with time once the gel is touching both walls, see Figure~\ref{fig:dynamics}a and Supplementary Figure~\ref{S-fig:growth}.
-This arrest may be due to bending rigidity~\cite{LeGoff2014}.
-Here the relevant time scale $\tau$ corresponds to the times it takes for the wrinkle amplitude to saturate, see Supplementary Methods for the precise definition.
+$\Upsilon$ is a dimensionless viscoelastic factor quantifying the relative stiffness of the film and the solvent at a characteristic time $\tau$.
 
-The opposite limiting case is when the gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Supplementary Video~\ref{S-vid:cutPoiseuille}) and impermeable ($\alpha\rightarrow 0$).
+The opposite limiting case is when an impermeable ($\alpha\rightarrow 0$) gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Figure~\ref{fig:sketchDP}b and Supplementary Video~S\ref{S-vid:cutPoiseuille}).
 Here destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
@@
 
-A more complete derivation of the models can be found in the Supplementary Methods and yields the following prefactors:
+A more complete derivation of the models can be found in the Supplementary Materials and yields the following prefactors:
 \begin{align}
-\lambda_D &= 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}\\
-\lambda_P &= \pi (hH)^{1/2}\left(\frac{2}{9}\Upsilon\right)^{1/6}
+\lambda_D &= 2\pi h^{1/2}\alpha^{1/4}\left(\frac{\Upsilon}{12}\right)^{1/4}\label{eq:prefDarcy}\\
+\lambda_P &= \pi (hH)^{1/2}\left(\frac{2}{9}\Upsilon\right)^{1/6}\label{eq:prefPoiseuille}
 \end{align}
@@
 
+The characteristic time $\tau$ is the hallmark of the kinetic nature of the wavelength selection mechanisms and is present in the expressions of both $\lambda_D$ and $\lambda_P$.
+
 To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$, i.e.
-slow destabilisation of a highly permeable film close to a wall, the main dissipation mechanism would be the flow through the porous film; whereas for $H\gg H^*$, i.e.
-fast wrinkling of a poorly permeable film far from any wall, the flow in the viscous layers would dominate and set the wavelength.
+slow destabilisation of a highly permeable film close to a wall, the main dissipation mechanism being the flow through the porous film; whereas for $H\gg H^*$, i.e.
+fast wrinkling of a poorly permeable film far from any wall, the flow in the viscous layers dominates and sets the wavelength.
 
-\section*{Experimental tests}
+\subsection*{Experimental tests}
 
-For a typical gel sheet, the buoyancy after syn\ae{}resis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
-This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ and therefore gravity is dismissed.
+For a typical gel film, the buoyancy after syneresis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ which confirms that gravity can be neglected as already inferred from test experiments where the cell is held vertically rather than horizontally.
 
 Darcy and Poiseuille models yield wavelengths that are of the same order in most of the cases under study.
-Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them, see Supplementary Methods about such measurements.
-As summarised in Supplementary Table~\ref{S-tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
+Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them, see Supplementary Materials about such measurements.
+As summarised in Supplementary Table~S\ref{S-tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
 Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly.
-However we are able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
+
+For each experiment we measure independently $\lambda$ and all the parameters that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}).
+$\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Materials and Methods and Supplementary Materials.
+
+The beauty of this approach is that it is not only valid to model the primary wavelength, but that it describes also the growth of the secondary pattern.
+Although the wavelength $\lambda_2$ of the secondary pattern is fully determined by the primary pattern $\lambda_2 = \lambda_1/2$, Figure~\ref{fig:dynamics} and Supplementary Figure~S\ref{S-fig:generations}, the secondary pattern has its own dynamics with its own characteristic time $\tau_2\neq\tau_1$.
 Since the saturation of the primary pattern flattens the gel on the top and bottom walls, the secondary blisters always appear in a Darcy situation ($H\ll H^*$).
-Here the wavelength associated to the secondary blister is taken as half of that of the primary pattern as shown in Supplementary Figure~\ref{S-fig:generations}.
+Therefore, we are able to obtain more data at small $H/H^*$ ratios by considering not only the primary patterns but also the secondary ones.
 
-Our results are displayed in Figure~\ref{fig:DarcyPoiseuille}.
-Although they provide the correct order of magnitude for $\lambda$, neither model alone is able to quantitatively reproduce the full data set.
-However considering only the points where $H<H^*$ (respectively $H>H^*$), the Darcy (resp.
-Poiseuille) model fits the data well albeit with a prefactor.
-Note that it would be easy to miss out on the failure of the Poiseuille model at small wavelengths by adding an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b).
-Nonetheless, a complete model (derived in Supplementary Methods), summing up the influence of both Poiseuille and Darcy dissipation provides a much better agreement for the whole $H/H^*$ range.
+Despite changing the gel properties and the cell geometry we always observe wrinkling proving the robustness of this approach to form wrinkling patterns.
+We are in the position to vary $H/H^\star$ from 0.1 to 10 so that we can test the crossover between Darcy and Poiseuille mechanism.
+In Figure~\ref{fig:DarcyPoiseuille}a and \ref{fig:DarcyPoiseuille}b we have plotted the measured wavelength as a function of the model prediction $\lambda_D$ and $\lambda_P$ respectively.
+If the predictions are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$.
+Indeed, both model seems to apply in their domain of validity.
+The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data correctly.
+To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Materials.
+The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ over the whole $H/H^*$ range.
+
+\textcolor{blue}{Despite a very good agreement of the scaling, the exact prefactors of all models disagree by a factor 0.63-0.69.
+We suggest that this discrepancy comes from the fact that our model is 2D, i.e.
+uniaxial compression and deflection, whereas our system is 3D, i.e.
+biaxial compression and deflection.
+Since our system is isotropic in the plane, this does not change the scaling but might account for a dimensional factor.}
 
@@
 Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins.
-As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
+As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in soft optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, food texturing~\cite{Mezzenga2005,Gibaud2012a} and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
 
@@
 We use either a 10x (air) lens to measure the geometry of wrinkling, see Supplementary Material, or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern, Supplementary Figures~\ref{S-fig:generations} and \ref{S-fig:patterns}.
-The secondary pattern wavelength is taken as half of the primary wavelength, according to Supplementary Figure~\ref{S-fig:generations}.
+Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern, Supplementary Figures~S\ref{S-fig:generations} and S\ref{S-fig:patterns}.
+The secondary pattern wavelength is taken as half of the primary wavelength, according to Supplementary Figure~S\ref{S-fig:generations}.
 Data shown in Figure~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
 For every gel composition we monitor the gelation by small amplitude oscillations (strain 0.1\%, frequency \SI{1}{\hertz}) and estimate the flexural modulus $B$ from the maximum value of the shear modulus $G^\prime$ as $B=2G^\prime/(1-\nu)$ taking $\nu=0.3$ for the Poisson ratio, a value typical of a spongy network~\cite{Greaves2011}.
-Permeability is measured using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~\ref{S-fig:permeability}.
+Permeability is measured using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~S\ref{S-fig:permeability}.
 The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
 
-\bibliographystyle{naturemag3}
+\bibliographystyle{Science}
 \bibliography{Yaourt}
@@
 Correspondence and requests for materials should be addressed to TG (\href{mailto:thomas.gibaud@ens-lyon.fr}{thomas.gibaud@ens-lyon.fr}).
-
 \end{document}

### 2015-07-20 fig_plis/supplementary_plis.tex
Split supplementary

@@
 \externaldocument[M-]{article_plis}
-\externalcitedocument[M-]{article_plis}
+\externalcitedocument{article_plis}
+
+% Hack for making figures Say \figurename S\thefigure, e.g. Figure S1:
+\makeatletter
+\makeatletter \renewcommand{\fnum@figure}
+{\textbf{Fig.~S\thefigure}}
+\makeatother
+
+\renewcommand{\thetable}{\arabic{table}}
+\makeatletter
+\makeatletter \renewcommand{\fnum@table}
+{\textbf{Tab.~S\thetable}}
+\makeatother
 
@@
 
-\section*{Fraction of free casein as a function of pH}
+\subsection*{Wavelength measurements}
+
+\begin{figure}
+	\includegraphics{generations}%
+	\caption{Fourier spectra of binary altitudes of Main Figure~\ref{M-fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
+Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
+The scale bar is \SI{1}{\milli\metre}.}%
+	\label{fig:generations}%
+\end{figure}
+
+\begin{figure}
+\includegraphics{ripples.pdf}
+\caption{Evidence for the ripples generated by the junction to flat patches, see~\cite{LeGoff2014}.
+(a) Position of the bottom surface of the gel obtained by confocal images.
+(b) one dimensional cut of the region of interest highlighted in (a) using an orange box.
+Only the first ripple is visible.
+(c) Schematic view of the interference between ripples.}
+\label{fig:ripples}
+\end{figure}
+
+\begin{figure*}
+	\includegraphics{patterns}
+	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data}.
+(a-d) increasing cell thickness.
+(e-f) larger solid content.
+(g-i) increasing glycerol content.
+All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
+The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
+Arrows show the measured primary wavelength.}
+	\label{fig:patterns}
+\end{figure*}
+
+To measure the wrinkle wavelength $\lambda$, we use large field pictures of the final state.
+These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e.
+The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
+White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
+It is thus possible to reconstruct a binary picture of the pattern, inset of Supplementary Figure~\ref{fig:generations}.
+Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final binary picture to the configuration of the primary pattern.
+
+Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
+When only the first generation pattern is present a peak clearly indicates the dominant mode.
+When the secondary pattern has appeared, the dominant wavelength is halved and the peak much weaker.
+This allows us to use the secondary pattern is our analysis by setting their wavelength to half of the primary wavelength.
+Further generations do not seem to appear simultaneously and the corresponding peaks are weak and broad.
+
+\subsection*{Confocal images processing}
+
+At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
+Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
+At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.
+
+We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
+Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
+We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line.
+Since this criterion is local, we do not need to correct for background intensity variation.
+In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.
+
+From these data, we obtain for each stack the volume and thickness of the gel phase, area and peak-to-peak amplitude of the mid surface.
+Velocity is obtained by time differentiation.
+Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
+
+\subsection*{Characteristic time measurements}
+
+In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
+However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
+Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view:
+\begin{equation}
+v(t) = \frac{\int \left|\frac{dz}{dt}\right| dxdy}{\int dxdy},
+\end{equation}
+where $z(x,y,t)$ is the altitude of the mid surface of the gel.
+
+\subsection*{Fraction of free casein as a function of pH}
 
@@
 The supernatant is diluted and analysed with a UV absorption spectrometer (USB4000, Ocean Optics).
-According to~\cite{M-Roefs1986}, we deduce the fraction of free casein proteins from the adsorption spectrum as:
+According to~\cite{Roefs1986}, we deduce the fraction of free casein proteins from the adsorption spectrum as:
 %
@@
 To follow the evolution of the microstructure of the gel, we acquire slices of $(\SI{2048}{px})^2$ with pixels of \SI{98}{\nano\metre}, Supplementary Figure~\ref{fig:nonmonotonic}a and Supplementary Videos~\ref{vid:stick63} and \ref{vid:nostick63}.
-After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$, see Supplementary Figure~\ref{fig:nonmonotonic}b that is fitted by the following fractal form~\cite{M-VanDijk1986}:
+After Hann windowing we perform Fourier transform and radially average the resulting spectrum to obtain $I(q)$, see Supplementary Figure~\ref{fig:nonmonotonic}b that is fitted by the following fractal form~\cite{VanDijk1986}:
 \begin{equation}
@@
 
-Figure~\ref{M-fig:acidification}d of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c-e show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syn\ae{}resis and swelling; (ii) complete adhesion except on the top wall, allowing syn\ae{}resis and swelling.
+Figure~\ref{M-fig:acidification}d of the main text and Supplementary Figure~\ref{fig:nonmonotonic}c-e show the evolution of the fitting parameters in two different sets of boundary conditions: (i) complete adhesion of the gel on all walls, preventing syneresis and swelling; (ii) complete adhesion except on the top wall, allowing syneresis and swelling.
 The latter situation is a proxy to the situation where we observe wrinkling, since it is necessary to keep the gel into the thin focal plane of the 63x objective lens of the confocal microscope.
@@
 
-\section*{Permeability measurements}
 \begin{figure*}%
@@
 \end{figure*}
+\subsection*{Permeability measurements}
 
-To measure the permeability of our gels, we adapt a method from \cite{M-VanDijk1986}.
+To measure the permeability of our gels, we adapt a method from \cite{VanDijk1986}.
 First, the gel of interest is formed at the lower extremity of a thin glass tube.
@@
 
-\section*{Confocal images processing}
+\subsection*{Mixed Darcy--Poiseuille model}
 
-At 10x magnification and depending on cell thickness, we acquire stacks of 20 to 50 images spaced by \SI{6.69}{\micro\metre} (Nyquist sampling).
-Each confocal slice is $(\SI{256}{px})^2$ with pixels of \SI{4.97}{\micro\metre}.
-At such low resolution details of the gel microstructure are undistinguishable as $\xi$ is smaller than the pixel size for all compositions and the gel appears as a continuous medium.
-
-We first de-noise each plane by a short-range Gaussian blur ($\sigma=\SI{2}{px}\ll\lambda$).
-Then, for each $(x,y)$, we focus on the z-dependent intensity profile $I(z)$.
-We define the position of the top surface of the gel $z_\text{top}$ as the point where $I(z)$ first crosses half of its maximum value over the given vertical line.
-Since this criterion is local, we do not need to correct for background intensity variation.
-In this way we obtain $z_\text{top}(x,y)$ with subpixel resolution ($\pm\SI{0.1}{px}=\SI{0.67}{\micro\metre}$), and $z_\text{bottom}(x,y)$ in a symmetrical way.
-
-From these data, we obtain for each stack the volume and thickness of the gel phase, area and peak-to-peak amplitude of the mid surface.
-Velocity is obtained by time differentiation.
-Once the instability has saturated on the top and bottom walls of the cell, peaks in the histograms of $z_\text{top}$ and $z_\text{bottom}$ yield the measure of the positions of both walls, thus of the cell thickness $e$ and retrospectively of $H_1$ and $H_2$.
-
-\section*{Characteristic time measurements}
-\begin{figure}
-	\includegraphics{growth}
-	\caption{Growth of the instability.
-(a) Evolution of the gel velocity $v$ averaged over the whole surface.
-Inset: Evolution of the peak-to-peak amplitude $A(t)$ in semi logarithmic scales.
-The black line is the best exponential fit with $\tau=\SI{90}{\second}$.
-(b) Evolution of the bottom profile of the gel from onset of instability (yellow) to contact to both walls (brown).
-Note that the longitudinal size of the blister evolves in this regime.
-The thick black curve is the profile at much longer times and is at lower altitude due to swelling.
-Note that the blister foot, highlighted by the orange vertical line, has not moved since contact.}
-	\label{fig:growth}
-\end{figure}
-
-In most of our samples it is possible to measure directly the peak-to-peak amplitude $A(t)$ from our confocal measurements.
-However when the wavelength is too large compared to the field of view, this approach breaks down since we cannot observe both the highest and lowest points of the surface.
-Even when valid, such a local measurement is rather noisy and we prefer to focus on the measure of the vertical velocity $v$ of the gel that can be integrated over the whole field of view.
-
-We note that when both measurements were available (as in Supplementary Figure~\ref{fig:growth}a), the characteristic time $\tau$ obtained from the fit of the exponential growth of $A(t)$ is about $4A_\text{max}/v_\text{max}$, where $A_\text{max} = e-h$ is the amplitude when the gel comes in contact with both walls and $v_\text{max}$ is the maximum velocity of the gel.
-We use this approximation of $\tau$ for all our samples.
-This is consistent with the fact that the blister size ($\propto\lambda$) does not evolve after contact, see Supplementary Figure~\ref{fig:growth}b, thus the kinetic wavelength is set when $A(t) = A_\text{max}$.
-
-\section*{Mixed Darcy--Poiseuille model}
-
-\subsection*{General framework for wrinkling}
+\subsubsection*{General framework for wrinkling}
 
@@
 \end{equation}
-where $w(x,t)$ is the deflection of the plate along $z$~\cite{M-Biot1957}. Since we deal with slow rates of deformation, inertia is neglected.
+where $w(x,t)$ is the deflection of the plate along $z$~\cite{Biot1957}. Since we deal with slow rates of deformation, inertia is neglected.
 
@@
 
-Following \cite{M-Biot1957}, we consider the case where the compression load $\sigma_\parallel$ is increased quasistatically.
+Following \cite{Biot1957}, we consider the case where the compression load $\sigma_\parallel$ is increased quasistatically.
 The observed mode is the one that arises at the lowest possible $\sigma_\parallel$, i.e.
@@
 
-\subsection*{Elastic film in an infinite elastic medium}
+\subsubsection*{Elastic film in an infinite elastic medium}
 
-In this framework, we recall here the case of an elastic film lying over an infinite elastic medium~\cite{M-Gough1940,M-Bijlaard1946,M-Biot1957,M-Cerda2003}.
+In this framework, we recall here the case of an elastic film lying over an infinite elastic medium~\cite{Gough1940,Bijlaard1946,Biot1957,Cerda2003}.
 The deformation of the substrate of modulus $B_s$ gives the transverse load as $\sigma_\perp = -q B_s w$, and thus the compression load writes
@@
 
-Minimising $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{M-Biot1957,M-Cerda2003} so that the dominant wavelength is 
+Minimising $\sigma_\parallel$ as a function of $q$ yields $hq = \left(6B_s/B\right)^{1/3}$ \cite{Biot1957,Cerda2003} so that the dominant wavelength is 
 \begin{equation}
@@
 
-\subsection*{Elastic film on a viscous layer}
+\subsubsection*{Elastic film on a viscous layer}
 
@@
 Let us now turn to the case of an elastic, impermeable film on a single viscous layer.
-As in \cite{M-Huang2002} we neglect gravity, although it would be difficult to do so in practice at a free interface.
+As in \cite{Huang2002} we neglect gravity, although it would be difficult to do so in practice at a free interface.
 Assuming the layer thickness $H$ to be much smaller than $\lambda_d$, we can use the lubrication approximation.
@@
 
-Combining equations~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
+Combining Equations~(\ref{eq:mecheqop}) and (\ref{eq:PoiseuilleLoad}), the compression load writes
 \begin{equation}
@@
 
-\subsection*{Porous elastic film with no possible Poiseuille flow}
+\subsubsection*{Porous elastic film with no possible Poiseuille flow}
 Another limiting case is that of a porous elastic film immersed in a viscous medium where longitudinal resistance to flow is infinite.
@@
 
-\subsection*{Porous elastic film between two viscous layers}
+\subsubsection*{Porous elastic film between two viscous layers}
 Finally the most realistic scenario mixes both Poiseuille and Darcy mechanisms, i.e.
@@
 
-\section*{Samples characteristics}
+\subsection*{Samples characteristics}
 To test the theoretical predictions we have fully characterised 9 samples. Their characteristics are gathered in Supplementary Table \ref{tab:data}.
@@
 \end{tabular}
-\caption{Characteristics of the samples used for Main Figure~\ref{M-fig:DarcyPoiseuille}.
+\caption{Characteristics of the samples used for Figure~\ref{M-fig:DarcyPoiseuille}.
 Lines where preparation and properties are left blank correspond to the average of the secondary blisters of the previous line.
@@
 $\xi$ is the pore size at the end of each experiment.
-$h$, $H_1$ and $H_2$ are measured just before buckling.}
+$h$ is the thickness of the gel film just before buckling.}
 \label{tab:data}
@@
 
-\section*{Wavelength measurements}
-
-\begin{figure}
-	\includegraphics{generations}%
-	\caption{Fourier spectra of binary altitudes of Main Figure~\ref{M-fig:dynamics}a corresponding to primary, secondary and tertiary patterns.
-Inset: reconstructed binary altitudes at time \SI{2}{\hour}.
-The scale bar is \SI{1}{\milli\metre}.}%
-	\label{fig:generations}%
-\end{figure}
-
-\begin{figure*}
-	\includegraphics{patterns}
-	\caption{Patterns corresponding to the samples of Supplementary Table~\ref{tab:data}.
-(a-d) increasing cell thickness.
-(e-f) larger solid content.
-(g-i) increasing glycerol content.
-All pictures are stitching of fluorescent microscopy images except (c,e) which are details of reflected light macroscope images.
-The scale is common to all panels (scale bar \SI{1}{\milli\metre}).
-Arrows show the measured primary wavelength.}
-	\label{fig:patterns}
-\end{figure*}
-
-To measure the wrinkle wavelength $\lambda$, we use large field pictures of the final state.
-These picture are obtained either by stitching fluorescent microscopy images obtained at 10x magnification, Supplementary Figure~\ref{fig:patterns} except c and e, or by transmitted or reflected light macroscope at magnifications ranging from 1x to 6x, Supplementary Figure~\ref{fig:patterns}c and e.
-The contrast in such pictures is a measure of the magnitude of the altitude gradient of the gel film.
-White lines in fluorescence and reflection (respectively black lines in transmission) denote a sharp transition between two extreme altitudes (on top wall or on bottom wall).
-It is thus possible to reconstruct a binary picture of the pattern, inset of Supplementary Figure~\ref{fig:generations}.
-Furthermore since the shape and size of a given generation of pattern do not change after formation, it is possible to revert the final binary picture to the configuration of the primary pattern.
-
-Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
-When only the first generation pattern is present a peak clearly indicates the dominant mode.
-When the secondary pattern has appeared, the dominant wavelength is halved and the peak much weaker.
-This allows us to use the secondary pattern is our analysis by setting their wavelength to half of the primary wavelength.
-Further generations do not seem to appear simultaneously and the corresponding peaks are weak and broad.
-
-\section*{Supplementary Videos}
+\subsection*{Supplementary Videos}
 

### 2015-07-20 fig_plis/article_plis.tex
Passive voice -> active voice when possible

@@
 \begin{abstract}
-Confined thin surfaces may wrinkle due the growth of an excess of material. Usually the wavelength is set by elasticity or gravity, but here we explore new selection mechanisms based on hydrodynamics.
+Confined thin surfaces may wrinkle due the growth of an excess of material. Usually elasticity or gravity set the wavelength, but here we explore new selection mechanisms based on hydrodynamics.
 First, inspired by yoghurt-making processes, we use caseins, a family of milk proteins, as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, to design a porous biogel film immersed in a confined buoyancy-matched viscous medium.
@@
 There are many ways and reasons for a film to wrinkle.
-An elastic film buckles due to excess area with respect to its boundaries and wrinkles when buckling is hindered by a substrate.
+An elastic film buckles due to excess area with respect to its boundaries and wrinkles when a substrate hinders buckling.
 The selection of the preferred wrinkling wavelength is peculiar to the situation and is related to the mismatch of elastic properties between the surface and the substrate~\cite{Gough1940, Bijlaard1946, Biot1957, Bowden1998, Cerda2003, Genzer2006}.
@@
 While still liquid, the casein dispersion is injected in a home-made optical cell that is then sealed.
-This cell is designed such that the casein adhesion to both top and bottom walls is turned off and so that the gel is only anchored to the remaining four vertical sides of the cell.
+We design this cell such that the casein adhesion to both top and bottom walls is turned off and so that the gel is only anchored to the remaining four vertical sides of the cell.
 %
@@
 
-The dynamics of the pattern formation are followed by light transmission microscopy and confocal microscopy, see Materials and Methods.
+We follow the dynamics of the pattern formation by light transmission microscopy and confocal microscopy, see Materials and Methods.
 As shown in Figure~\ref{fig:dynamics}b and Supplementary Video~S\ref{S-vid:transmitted}, we observe the spontaneous formation of a first pattern that appears with a characteristic wavelength $\lambda\ll L$.
@@
 (i) In Figure~\ref{fig:dilation}d, the swelling back is reversible, only releasing tensile stresses accumulated during shrinking, and the layer remains flat, as expected for chemical gels~\cite{Kim2010}.
-Since this case is not observed it means that in this physical gel, syneresis and swelling processes foster irreversible stress dissipation and structural changes.
+Since we do not observe this case, it means that in this physical gel, syneresis and swelling processes foster irreversible stress dissipation and structural changes.
 In the following, we will therefore consider that the shrunk state is stress-free.
@@
 
-For a typical gel film, the buoyancy after syneresis is estimated from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
+For a typical gel film, we estimate the buoyancy after syneresis from the initial sodium caseinate mass concentration $C_\text{cas}$ as $\Delta\rho=C_\text{cas} e/h = \SI{130}{\kilo\gram\per\cubic\metre}$ leading to a gravity stress $\sigma_\perp^G \approx \SI{0.04}{\pascal}$.
 This is much lower than typical Darcy or Poiseuille stresses $\sigma_\perp^D \approx \sigma_\perp^P \approx \SI{0.15}{\pascal}$ which confirms that gravity can be neglected as already inferred from test experiments where the cell is held vertically rather than horizontally.
@@
 For each experiment we measure independently $\lambda$ and all the parameters that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}).
-$\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Materials and Methods and Supplementary Materials.
+We measure $\lambda$ by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ by confocal microscopy; $\alpha$ by capillary rise experiment and $B$ by rheology, see Materials and Methods and Supplementary Materials.
 
@@
 
-Sodium caseinate (Firmenich) is dissolved in deionised water by stirring for \SI{20}{\minute} at room temperature.
+We dissolve sodium caseinate (Firmenich) in deionised water by stirring for \SI{20}{\minute} at room temperature.
 The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) sets the beginning of the experiments.
 We typically have about \SI{10}{\minute} to pipet the dispersion into the slit before gelation starts.
-Sodium caseinate (Firmenich) is labelled with Dylight~550 NHS ESTER (Thermo Scientific).
+We label sodium caseinate (Firmenich) with Dylight~550 NHS ESTER (Thermo Scientific).
 Excess dye is removed by centrifugation.
@@
 Home made optical cells are composed of a slide (RS) and a cover slip (Menzel-Gl\"aser) spaced by paraffin film (Parafilm).
-The optical cell is briefly heated on a hot plate so that the paraffin welds to the slide and the cover slip.
-Gently applying a pressure on the cover slip while being heated allows us to tune the thickness of the cell from \SIrange{50}{150}{\micro\metre}.
-The slit is filled with the casein dispersion, immediately sealed using ultraviolet-cured glue (Norland Optical) and placed under the microscope before gelation can take place.
-To prevent adsorption of caseins to surfaces, microscope slides and cover slips used to make microscopy cells are cleaned and subsequently coated with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which a polyacrylamide brush is polymerised from methacrylate groups (Euromedex).
+We heat briefly the optical cell on a hot plate so that the paraffin welds to the slide and the cover slip.
+We tune the thickness of the cell from \SIrange{50}{150}{\micro\metre} by gently applying a pressure on the cover slip during heating.
+We fill the slit with the casein dispersion, and immediately seal it using ultraviolet-cured glue (Norland Optical) and place it under the microscope before gelation can take place.
+To prevent adsorption of caseins to surfaces, we clean microscope slides and cover slips used to make microscopy cells and subsequently coat them with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which we polymerise a polyacrylamide brush from methacrylate groups (Euromedex).
 
-3D data are collected on a Zeiss LSM510 confocal microscope using \SI{532}{\nano\meter} laser excitation.
+We collect 3D data on a Zeiss LSM510 confocal microscope using \SI{532}{\nano\meter} laser excitation.
 We use either a 10x (air) lens to measure the geometry of wrinkling, see Supplementary Material, or a 63x (oil) lens to observe the microstructure of the gel.
-Larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000) are used to measure the wavelength of the primary pattern, Supplementary Figures~S\ref{S-fig:generations} and S\ref{S-fig:patterns}.
-The secondary pattern wavelength is taken as half of the primary wavelength, according to Supplementary Figure~S\ref{S-fig:generations}.
-Data shown in Figure~\ref{fig:acidification}a and c are obtained with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
+We use larger scope pictures, obtained either by stitching fluorescent microscopy images (Nikon Eclipse Ti) or by transmitted/reflected light macroscope (Nikon SMZ745T/Leica DMS1000), to measure the wavelength of the primary pattern, Supplementary Figures~S\ref{S-fig:generations} and S\ref{S-fig:patterns}.
+We take the secondary pattern wavelength as half of the primary wavelength, according to Supplementary Figure~S\ref{S-fig:generations}.
+We obtain data shown in Figure~\ref{fig:acidification}a with a SevenCompact pH-meter (Mettler Toledo) and an MCR 301 rheometer (Anton Paar).
 For every gel composition we monitor the gelation by small amplitude oscillations (strain 0.1\%, frequency \SI{1}{\hertz}) and estimate the flexural modulus $B$ from the maximum value of the shear modulus $G^\prime$ as $B=2G^\prime/(1-\nu)$ taking $\nu=0.3$ for the Poisson ratio, a value typical of a spongy network~\cite{Greaves2011}.
-Permeability is measured using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~S\ref{S-fig:permeability}.
-The amount of free casein protein present in the gel during the GDL acidification process is measured by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
+We measure permeability using a protocol found in \cite{VanDijk1986}, see Supplementary Method and Supplementary Figure~S\ref{S-fig:permeability}.
+We measure the amount of free casein protein present in the gel during the GDL acidification process by titration and UV-spectrometry (Ocean Optics) according to the protocol found in~\cite{Roefs1986} and detailed in Supplementary Method.
 

### 2015-07-22 fig_plis/article_plis.tex
Science style citations in text (\it 1)
page numbers in footer
Modifs Seb American English

@@
 \usepackage{amsfonts}
+
+\pagestyle{plain}
+
+%Science style citations
+\setcitestyle{round}
+\renewcommand{\citenumfont}{\textit}
 
@@
 \author{Mathieu Nespoulous}
-\affiliation{Aix-Marseille Université, CNRS, MADIREL UMR 7246, Marseille, France}
-\author{Sébastien Manneville}
+\affiliation{Aix-Marseille Universit\'e, CNRS, MADIREL UMR 7246, Marseille, France}
+\author{S\'ebastien Manneville}
 \affiliation{Universit\'e de Lyon, Laboratoire de Physique, \'Ecole Normale Sup\'erieure de Lyon, CNRS UMR 5672, 46 All\'ee d'Italie, 69364 Lyon cedex 07, France}
@@
 \begin{abstract}
-Confined thin surfaces may wrinkle due the growth of an excess of material. Usually elasticity or gravity set the wavelength, but here we explore new selection mechanisms based on hydrodynamics.
+Confined thin surfaces may wrinkle due to the growth of an excess of material. Usually elasticity or gravity sets the wavelength. Here we explore new selection mechanisms based on hydrodynamics.
 First, inspired by yoghurt-making processes, we use caseins, a family of milk proteins, as pH-responsive building blocks and the acidulent glucono-$\delta$-lactone, to design a porous biogel film immersed in a confined buoyancy-matched viscous medium.
-Provided specific boundary conditions, but without any external stimulus, the biogel film spontaneously wrinkles in cascade.
-Second, using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that, during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
-Third, taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that predicts correctly the wrinkling wavelength.
-Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilisation/destabilisation, and therefore could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
+Under specific boundary conditions, yet without any external stimulus, the biogel film spontaneously wrinkles in cascade.
+Second, using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that during continuous acidification, the gel first shrinks and then swells, inducing wrinkling.
+Third, taking into account both Darcy flow through the gel and Poiseuille flow in the surrounding solvent, we develop a model that correctly predicts the wrinkling wavelength.
+Our results should be universal for acid-induced protein gels as they are based on pH-induced charge stabilization/destabilization, and therefore could set a benchmark to gain fundamental insights into wrinkled biological tissues, to texture food or to design surfaces for optical purposes.
 \end{abstract}
@@
 On the biology side wrinkling-controlled morphogenesis is ubiquitous.
-Ageing and the loss of elastic fibres makes our skin wrinkle~\cite{Bissett1987, Genzer2006}.
+Aging and the loss of elastic fibers makes our skin wrinkle~\cite{Bissett1987, Genzer2006}.
 Difference in growth rates between the gut tube and its dorsal anchoring is responsible for the vilification of guts~\cite{Savin2011, Ciarletta2014, Shyer2013}.
-Localised cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
+Localized cell death in biofilms focuses mechanical forces and initiates 3D labyrinth pattern~\cite{Trejo2013, Asally2012}.
 On the physics side the last two decades have seen a bloom of methods to obtain well-controlled patterns via linear~\cite{Bowden1998, Genzer2006, Hu1998, Kim2010, Vandeparre2011, Li2013} or nonlinear wrinkling~\cite{Efimenko2005, Guvendiren2010, Kim2011a, Brau2011}.
@@
 In a biological context, however, thin tissues, e.g.
-epithelimum or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g.
+epithelium or endothelium, are naturally confined and immersed in a nearly buoyant medium, e.g.
 lymph, blood or mucus, a practical situation where a thin viscous substrate would set the wrinkling wavelength~\cite{Fleury2015}.
@@
 Furthermore, within this primary pattern, secondary and ternary patterns appear in cascade.
-Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that the wrinkles at the millimetre scale result from the spontaneous swelling of the casein network at the micron level upon acidification.
+Using a combination of titration, rheology, light microscopy and confocal microscopy, we demonstrate that the wrinkles at the millimeter scale result from the spontaneous swelling of the casein network at the micron level upon acidification.
 %
@@
 Extracting the top and bottom surface of the gel phase allows for quantitative measurements of the relative volume, Figure~\ref{fig:sideview}b, and excess surface area of the gel film, Figure~\ref{fig:sideview}c, see Supplementary Materials.
-The evolution of the gel volume is surprisingly non-monotonous.
+The evolution of the gel volume is surprisingly non-monotonic.
 After reaching a minimum around 35\% of its initial value at \SI{23}{\minute}, the volume increases.
-Shortly after, at \SI{36}{\minute}, the planar film destabilises and starts to wrinkle in the $z$-direction creating excess area.
+Shortly after, at \SI{36}{\minute}, the planar film destabilizes and starts to wrinkle in the $z$-direction creating excess area.
 This confirms that the instability leading to the primary pattern is due to swelling.
@@
 
-Therefore at this stage, our observations still raise two unresolved issues which we address in the next sections: (i) what triggers the the spontaneous syneresis and swelling back processes?
+Therefore at this stage, our observations still raise two unresolved issues which we address in the next sections: (i) what triggers the spontaneous syneresis and swelling back processes?
 And (ii) what is the nature of $\sigma_{\perp}$ responsible for the selection of the primary wavelength?
@@
 At pH=7, the caseins are well dispersed then they flocculate between pH=5.5 to 4 and finally at pH=3 the solution becomes clearer as caseins partially redisperse.
-The state diagram of caseins as a function of pH is thus related to charge stabilisation.
+The state diagram of caseins as a function of pH is thus related to charge stabilization.
 At neutral pH caseins are charged negatively and repel each other.
@@
 Starting at pH=6.8 all the casein molecules are well dispersed and the free casein fraction $x_\text{free}$ is 1.
-Upon decreasing the pH to the isolectric point the free casein fraction drops to almost 0: almost all the casein molecules are aggregated.
+Upon decreasing the pH to the isoelectric point the free casein fraction drops to almost 0: almost all the casein molecules are aggregated.
 Further decrease of the pH of the solution leads to a partial redissolution of the casein aggregates.
@@
 Permeability $\alpha$ is of the order of \SI{67000}{\square\nano\metre}, see Supplementary Figure~S\ref{S-fig:permeability}.
-The fractal model fits the structure at all times but its parameters vary non monotonously during GDL acidification.
-As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~S\ref{S-fig:nonmonotonic}, when reaching the isolectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
+The fractal model fits the structure at all times but its parameters vary non monotonically during GDL acidification.
+As shown in Figure~\ref{fig:acidification}d and Supplementary Figure~S\ref{S-fig:nonmonotonic}, when reaching the isoelectric point $\xi$ increases up to a maximum length of \SI{7}{\micro\metre} due to the solvent expulsion out of the casein network~\cite{Lucey1998} like in viscoelastic phase separation~\cite{Tanaka2000h}.
 This corresponds to a tighter packing of the caseins within the network strands.
@@
 However the gel does not wrinkle because it remains attached to the bottom wall.} In this situation we can monitor $\xi$, Figure~\ref{fig:acidification}d.
-$\xi$ remains mostly constant and variations manifests themselves at the macroscopic level, on the volume occupied by the gel.
+$\xi$ remains mostly constant and variations manifest themselves at the macroscopic level, on the volume occupied by the gel.
 
-Therefore, the syneresis and swelling back mechanism is related to charge stabilisation of the casein proteins as function of the pH.
-\textcolor{blue}{In particular, beyond the isoelectric point, the casein are charged and repel each other, leading to a swelling at every scale: proteins are further from each other, the strands become larger and longer and the gel sheet swells.} Such a mechanism is only effective to produce wrinkles when adhesion of the caseins to both top and bottom walls is turned off.
+Therefore, the syneresis and swelling back mechanism is related to charge stabilization of the casein proteins as a function of the pH.
+\textcolor{blue}{In particular, beyond the isoelectric point, caseins are charged and repel each other, leading to swelling at every scale: proteins get farther apart from each other, the strands become larger and longer and the gel sheet swells.} Such a mechanism is only effective to produce wrinkles when adhesion of the caseins to both top and bottom walls is turned off.
 Consistently with our explanation, the amount of GDL, and thus the final pH reached by the suspension, controls the amount of swelling back.
-Thus at low GDL concentrations we only observe the primary pattern, at high GDL concentrations we observe the formation of higher order patterns.
+Thus at low GDL concentrations we only observe the primary pattern, while at high GDL concentrations we observe the formation of higher order patterns.
 
@@
 \caption{Sketch of the two limit scenarios for the wavelength selection.
-In both cases the gel film of thickness $h$ is destabilised by the excess area which gives rise to wrinkles of amplitude $A(t)$ and wavelength $\lambda$.
+In both cases the gel film of thickness $h$ is destabilized by the excess area which gives rise to wrinkles of amplitude $A(t)$ and wavelength $\lambda$.
 A single wavelength is represented and the longitudinal dimension $L$ of the system is much larger than $\lambda$.
@@
 This framework has been successfully applied to situations where $\sigma_{\perp}$ comes from the resistance of an elastic substrate ~\cite{Gough1940, Bijlaard1946}.
-Dealing only with (semi)infinite substrates, Biot generalised the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
+Dealing only with (semi)infinite substrates, Biot generalized the elasticity of both the film and the substrate to viscoelasticity~\cite{Biot1957} or poroelasticity~\cite{Biot1964}, leading to time-dependent kinetic wavelength selection.
 Furthermore, it was shown that equation~(\ref{eq:lstar}) remains valid even when $\sigma_{\perp}$ comes from the resistance to uniaxial stretching~\cite{Cerda2003}, from gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, from capillary forces~\cite{Huang2007}, from boundaries~\cite{Vandeparre2011a, Li2013} or from adhesion~\cite{Vella2009a}.
@@
 The opposite limiting case is when an impermeable ($\alpha\rightarrow 0$) gel film lies in the middle of the cell ($H\equiv H_1=H_2$, see Figure~\ref{fig:sketchDP}b and Supplementary Video~S\ref{S-vid:cutPoiseuille}).
-Here destabilisation over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
+Here destabilization over a wavelength $\lambda$ creates a lubrication ($H\ll\lambda$) flow in the viscous layers.
 By symmetry the transverse load across the gel $\sigma_{\perp}=p_2-p_1$ is also the pressure difference over the wavelength.
@@
 \end{equation}
-where we recognise the scaling for the wrinkles of an elastic film on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate.
+where we recognize the scaling for the wrinkles of an elastic film on a thin ($H\ll\lambda$) elastic substrate~\cite{Cerda2003} where $\Upsilon$ would be the ratio between the flexural modulus of the film and that of the substrate.
 Indeed within the lubrication approximation or at low Reynolds numbers one can consider a viscous film as elastic with effective Young modulus $3\eta/\tau$~\cite{Biot1957,Boudaoud2001}.
@@
 To sum up, we expect two regimes of viscosity-dominated wrinkling: for $H\ll H^*$, i.e.
-slow destabilisation of a highly permeable film close to a wall, the main dissipation mechanism being the flow through the porous film; whereas for $H\gg H^*$, i.e.
+slow destabilization of a highly permeable film close to a wall, the main dissipation mechanism being the flow through the porous film; whereas for $H\gg H^*$, i.e.
 fast wrinkling of a poorly permeable film far from any wall, the flow in the viscous layers dominates and sets the wavelength.
@@
 Thus only systematic measurements of the experimental $\lambda$ together with estimates for all the parameters involved in either model shall allow one to discriminate between them, see Supplementary Materials about such measurements.
-As summarised in Supplementary Table~S\ref{S-tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
+As summarized in Supplementary Table~S\ref{S-tab:data} we systematically vary (i) the thickness of the cell, impacting $h$ and the maximal amplitude, (ii) the gel composition and thus both its stiffness and its permeability, (iii) the viscosity of the solvent by adding glycerol, which also impacts the gel stiffness, permeability and $\tau$.
 Finally, the only parameter that we have no control upon is the initial altitude $H$ of the gel before wrinkling that varies randomly.
@@
 \textcolor{blue}{Despite a very good agreement of the scaling, the exact prefactors of all models disagree by a factor 0.63-0.69.
-We suggest that this discrepancy comes from the fact that our model is 2D, i.e.
-uniaxial compression and deflection, whereas our system is 3D, i.e.
+We suggest that this discrepancy comes from the fact that our model is two-dimensional, i.e.
+uniaxial compression and deflection, whereas our system is three-dimensional, i.e.
 biaxial compression and deflection.
-Since our system is isotropic in the plane, this does not change the scaling but might account for a dimensional factor.}
+Since our system is isotropic in the horizontal plane, this does not change the scaling but it could account for a dimensional factor.}
 
@@
 We developed a model that captures the dynamical origin of the constraints exerted on the gel and predicts the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall.
-Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins.
+Such a model experiment, because it relies on pH-induced charge stabilization/destabilization, should be generalizable to different kinds of proteins.
 As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in soft optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, food texturing~\cite{Mezzenga2005,Gibaud2012a} and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.
@@
 
-We dissolve sodium caseinate (Firmenich) in deionised water by stirring for \SI{20}{\minute} at room temperature.
-The mixing of the caseine dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) sets the beginning of the experiments.
+We dissolve sodium caseinate (Firmenich) in deionized water by stirring for \SI{20}{\minute} at room temperature.
+The mixing of the casein dispersion with glucono-$\delta$-lactone (GDL) in powder (Firmenich) sets the beginning of the experiments.
 We typically have about \SI{10}{\minute} to pipet the dispersion into the slit before gelation starts.
@@
 We fill the slit with the casein dispersion, and immediately seal it using ultraviolet-cured glue (Norland Optical) and place it under the microscope before gelation can take place.
-To prevent adsorption of caseins to surfaces, we clean microscope slides and cover slips used to make microscopy cells and subsequently coat them with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which we polymerise a polyacrylamide brush from methacrylate groups (Euromedex).
+To prevent adsorption of caseins to surfaces, we clean microscope slides and cover slips used to make microscopy cells and subsequently coat them with a silane agent (3-(trimethoxysilyl) propyl methacrylate, Sigma Aldrich) from which we polymerize a polyacrylamide brush from methacrylate groups (Euromedex).
 
@@
 
+\paragraph*{Competing Interests}
+The authors declare that they have no competing interests.
+
 \paragraph*{Acknowledgements}
 Confocal experiments were conducted at SFR BioSciences Gerland - Lyon Sud (US8 / UMS3444).
-The authors thank Jos\'e Bico, Arezeki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
+The authors thank Jos\'e Bico, Arezki Boudaoud, Cyprien Gay, L. Mahadevan and Olivier Pierre-Louis for theoretical insights at various stages of the research. 
 Precious experimental ideas came from Catherine Barentin, Denis Bartolo, Elisabeth Bouchaud and Thibaut Divoux.
@@
 Special thanks to Madame Biot who made M.A. Biot work accessible through the \href{http://www.olemiss.edu/sciencenet/poronet/}{poronet} website. 
-ML thanks the Region Rhône Alpes and the Programme d'Avenir Lyon - Saint Etienne (PALSE NoGELPo) for postoctoral grant. 
+ML thanks the Region Rh{\^o}ne Alpes and the Programme d'Avenir Lyon - Saint Etienne (PALSE NoGELPo) for postdoctoral grant. 
 TG and MN acknowledge funding from the Agence Nationale de la Recherche (ANR-11-PDOC-027). 

### 2015-07-22 fig_plis/supplementary_plis.tex
Modifs Seb American English

@@
 
-Figure~\ref{fig:generations} shows Fourier spectra of binarised altitudes at different times.
+Figure~\ref{fig:generations} shows Fourier spectra of binarized altitudes at different times.
 When only the first generation pattern is present a peak clearly indicates the dominant mode.
@@
 In both cases, \SI{0.4}{\milli\litre} of the \SI{100}{\milli\litre} dispersion is sampled regularly as a function of pH, and centrifuged (centrifuge 5424, Eppendorf) in Eppendorf tubes at 14500 rpm for \SI{3}{\minute}.
-The supernatant is diluted and analysed with a UV absorption spectrometer (USB4000, Ocean Optics).
+The supernatant is diluted and analyzed with a UV absorption spectrometer (USB4000, Ocean Optics).
 According to~\cite{Roefs1986}, we deduce the fraction of free casein proteins from the adsorption spectrum as:
@@
 By fitting $z(t)-z_\infty$ to an exponential, see Supplementary Figure~\ref{fig:permeability} b-c, we obtain $\tau$ and thus $\alpha$.
-Note that even with $H$ of the order of a few millimetres high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due to both high viscosity and low permeability.
+Note that even with $H$ of the order of a few millimeters high glycerol contents yield $\tau\approx \SI{100}{\hour}$ due to both high viscosity and low permeability.
 Obtaining $z_\infty$ without tube 2 would be time prohibitive.
@@
 \end{equation}
-Here also the dominant mode maximises the growth rate.
+Here also the dominant mode maximizes the growth rate.
 

### 2015-09-02 fig_plis/article_SI.tex
Add links to youtube videos

@@
 \itemindent 8mm
-\item\label{vid:transmitted} Transmitted light microscopy images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
+\item\label{vid:transmitted} \href{https://youtu.be/KsSRDUxmmbo}{Transmitted light microscopy} images of the pattern formation in a 4\%w casein and 4\%w GDL aqueous dispersion.
 Top and bottom slides of the optical cell are coated with acrylamide brushes.
@@
 
-\item\label{vid:reconstructed} 3D reconstruction of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion, Supplementary Table~\ref{tab:data}c.
+\item\label{vid:reconstructed} \href{https://youtu.be/EI4K2akmhL8}{3D reconstruction} of the pattern formation using confocal microscopy in a 4\%w casein and 4\%w GDL aqueous dispersion, Supplementary Table~\ref{tab:data}c.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item\label{vid:titration} Titration of a dispersion with 4\%w casein by 1 molar HCl. The total duration of the Video is \SI{14}{\minute}. The Becher diameter is \SI{55}{\milli\metre}.
+\item\label{vid:titration} \href{https://youtu.be/t6_vnQmzais}{Titration} of a dispersion with 4\%w casein by 1 molar HCl.
+The total duration of the Video is \SI{14}{\minute}.
+The Becher diameter is \SI{55}{\milli\metre}.
 
-\item\label{vid:stick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when sticky boundary are on.
+\item\label{vid:stick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when \href{https://youtu.be/64ntew44Uy4}{sticky boundary are on}.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item\label{vid:nostick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when the top wall is coated with acrylamide brushes and the bottom wall remains sticky.
+\item\label{vid:nostick63} Confocal microscopy $(x,y)$ cut of the evolution of the local structure of a 4\%w casein and 4\%w GDL aqueous dispersion when \href{https://youtu.be/3UFrgvDk2eM}{the top wall is coated with acrylamide brushes and the bottom wall remains sticky}.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item\label{vid:cutDarcy} Confocal microscopy $(x,z)$ cut of the evolution of the gel film deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}a.
+\item\label{vid:cutDarcy} \href{https://youtu.be/Z2wOHFEpfhc}{Confocal microscopy $(x,z)$ cut} of the evolution of the gel film deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}a.
 10\% of the proteins are fluorescently labelled.
@@
 
-\item\label{vid:cutPoiseuille} Confocal microscopy $(x,z)$ cut of the evolution of the gel film deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}c.
+\item\label{vid:cutPoiseuille} \href{https://youtu.be/fkJeP_VpbdM}{Confocal microscopy $(x,z)$ cut} of the evolution of the gel film deflection of a 4\%w casein and 4\%w GDL aqueous dispersion when both top and bottom walls are coated with acrylamide brushes, Supplementary Table~\ref{tab:data}c.
 10\% of the proteins are fluorescently labelled.

### 2015-10-16 fig_plis/supplementary_plis.tex
Talk ontpellier Ajoute dosage

@@
 
+Combining equations~(\ref{eq:sigma0P}) and (\ref{eq:qP}) we obtain the minimal stress to observe buckling
+\begin{equation}
+\sigma_\parallel = \frac{h}{H}  B  \left(\frac{24}{\Upsilon}\right)^{1/3}.
+\end{equation}
+
+Let us compare $\sigma_\parallel$ to the stress induced by swelling $\sigma_s$ for a sample abiding to Poiseuille model: sample c in Table~\ref{tab:data}.
+From the previous equation, we obtain $\sigma_\parallel \approx \SI{9}{\pascal}$.
+On the other hand, we estimate $\sigma_s = B\gamma$ where $\gamma$ is the strain induced by swelling.
+For the same sample, Figure \ref{M-fig:sideview}c gives us the excess area $\epsilon$ as a function of time.
+The first reliable measure of $\gamma = \sqrt{\epsilon}$ after buckling is 2.5\% at $t=\SI{35}{\minute}$.
+Therefore, the stress induced by swelling is $\sigma_s \approx \SI{35}{\pascal}$, indeed sufficient to cause buckling.
+
 \subsubsection*{Porous elastic film with no possible Poiseuille flow}

### 2015-10-16 talk_Montpellier.tex
Talk ontpellier Ajoute dosage

--- 2015-06-08 talk_Zurich.tex
@@
 
-\institute[E.N.S. Lyon]{Laboratoire de physique, Ecole Normale Sup\'{e}rieure de Lyon}
+\institute[iLM]{Institut Lumière Matière, Lyon, France}
 \title[Yoghurt under stress]{Acid-induced protein gels: from gelation to stress-induced failure}
@@
 \titlegraphic{
-	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}\quad
+	\begin{tabu}{X[c]X[c]X[c]X[c]}
+		\includegraphics[height=3\baselineskip]{Chris}&
+		\includegraphics[height=3\baselineskip]{Thomas}&
+		\includegraphics[height=3\baselineskip]{Thibaut}&
+		\includegraphics[height=3\baselineskip]{Seb}\\
+		Christophe Perge & Thomas\linebreak Gibaud & Thibaut\linebreak Divoux & Sebastien Manneville\\
+	\end{tabu}
+
+	\vfill
+	\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
 	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
 	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
-	\includegraphics[height=2\baselineskip]{logo_ens-lyon}
+	\includegraphics[height=2\baselineskip]{CRPP}\quad
+	\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}
 	}
@@
 \item Sodium caseinate (milk protein) 4\%
-\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
+\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right(\}$ stable solution}
 \item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow homogeneous acidification
@@
 
-\begin{frame}{Merciless yoghurt breakers}
-\textit{Leocmach et al. PRL 113, 038303 (2014), \footnotesize ArXiv 1401.8234}
+\section*{Beyond critical gels}
 
-\bigskip
-\begin{tabu}{X[c]X[c]c}
-\includegraphics[height=0.3\textheight]{Chris}&
-\includegraphics[height=0.3\textheight]{Thibaut}&
-\includegraphics[height=0.3\textheight]{Seb}\\
-Christophe Perge & Thibaut Divoux & Sebastien Manneville\\
-PhD student & CNRS researcher & Professor\\
-E.N.S. Lyon &  CRPP Bordeaux & E.N.S. Lyon\\
-\end{tabu}
+\begin{frame}{Acidifying casein solutions with \textup{HCl}}
+\tikzsetnextfilename{dosage}
+\input{fig_presentation/dosage.pgf}
+\end{frame}
 
-\bigskip
-\begin{description}[Special thanks]
-\item[Special thanks] to Alan Parker (Firmenich) for prompting this study and providing casein \& GDL
-\end{description}
-
-\end{frame}
+%\begin{frame}{Merciless yoghurt breakers}
+%\textit{Leocmach et al. PRL 113, 038303 (2014), \footnotesize ArXiv 1401.8234}
+%
+%\bigskip
+%\begin{tabu}{X[c]X[c]c}
+%\includegraphics[height=0.3\textheight]{Chris}&
+%\includegraphics[height=0.3\textheight]{Thibaut}&
+%\includegraphics[height=0.3\textheight]{Seb}\\
+%Christophe Perge & Thibaut Divoux & Sebastien Manneville\\
+%PhD student & CNRS researcher & Professor\\
+%E.N.S. Lyon &  CRPP Bordeaux & E.N.S. Lyon\\
+%\end{tabu}
+%
+%
+%\bigskip
+%\begin{description}[Special thanks]
+%\item[Special thanks] to Alan Parker (Firmenich) for prompting this study and providing casein \& GDL
+%\end{description} 
+%
+%\end{frame}
 

### 2015-10-16 talk_Montpellier.tex
append Soundbite palavas allow pgfplottable relax tables after use

@@
 \usepackage{pgfplots}
+\usepackage{pgfplotstable}
 \pgfplotsset{compat=1.9}
@@
 
+\begin{frame}{Over-acidified protein gels: beyond isoelectric pH}
+	\tikzsetnextfilename{prise_cas4}
+	\input{fig_soundbite_palavas/prise_cas4.pgf}
+	\begin{itemize}
+	\item[$\Rightarrow$] still solid
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Linear rheology, from power-law to soft glassy}
+	\tikzsetnextfilename{sweep_cas4}
+	\input{fig_soundbite_palavas/sweep_cas4.pgf}
+	\begin{itemize}
+	\item Near isoelectric, power law rheology $G^\prime, G^{\prime\prime}$ typical of gels 
+	\item Over-acidification lowers exponent and creates a minimum in $G^{\prime\prime}$
+	\item Transition to soft-glassy behaviour \\\textit{\footnotesize Sollich, PRE (1998); Mason, Curr. Opin. Colloid Interface Sci. (1999)}
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Questions}
+\begin{description}
+\item[What] is the change in microstructure?
+\item[How] does it affects the failure mechanism?
+\end{description}
+\end{frame}
+
+\begin{frame}{Microscructure: fluorescent confocal microscopy}
+	\tikzsetnextfilename{structure_confocal}
+	\input{fig_soundbite_palavas/structure_confocal.pgf}
+	\begin{itemize}
+	\item Maximum $G^\prime$ corresponds to maximum contrast between phases
+	\item Characteristic size $\xi$ increases during over-acidification
+	\item Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Three creep regimes + softness\hfill\textsc{gdl} 4\%}
+\vfill
+	\tikzsetnextfilename{creep_cas4_gdl4}
+	\input{fig_soundbite_palavas/creep_cas4_gdl4.pgf}
+\begin{itemize}
+%\item Failure scenario almost conserved: ``fibre bundle'' network
+%\item Other dissipation due to softness: overcompressed gel
+\item More difficult to predict failure time from fractures nucleation time
+\end{itemize}
+
+\end{frame}
+
 %\begin{frame}{Merciless yoghurt breakers}

### 2015-10-26 talk_Montpellier.tex
critical vs super-critical gel story
conclusion of over-acidification

@@
 
-\begin{frame}{Acidifying casein solutions with \textup{HCl}}
-\tikzsetnextfilename{dosage}
-\input{fig_presentation/dosage.pgf}
-\end{frame}
-
 \begin{frame}{Over-acidified protein gels: beyond isoelectric pH}
@@
 	\item[$\Rightarrow$] still solid
+	\item From critical to super-critical gels?
 	\end{itemize}
@@
 	\begin{itemize}
-	\item Near isoelectric, power law rheology $G^\prime, G^{\prime\prime}$ typical of gels 
+	\item Near isoelectric, power law rheology $G^\prime, G^{\prime\prime}$ typical of critical gels 
 	\item Over-acidification lowers exponent and creates a minimum in $G^{\prime\prime}$
@@
 
+\end{frame}
+
+\begin{frame}{Over-acidification summary}
+\begin{itemize}
+\item The critical network is still there but over-compressed
+\item Transition to soft glassy rheology
+\item Creep regimes 1 and 2 still valid but messy in the middle
+\item[$\Rightarrow$] Rupture scenario of critical gel only partially valid for other gels.
+\end{itemize}
 \end{frame}

### 2015-10-26 talk_Montpellier.tex
Only gdot, modeled on cas4gdl1

@@
 	\input{fig_soundbite_palavas/structure_confocal.pgf}
+	\vspace{-1em}
 	\begin{itemize}
 	\item Maximum $G^\prime$ corresponds to maximum contrast between phases
-	\item Characteristic size $\xi$ increases during over-acidification
-	\item Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy
+	\item Pore size $\xi$ decrease during over-acidification
+	\item<2> Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy
 	\end{itemize}
@@
 	\tikzsetnextfilename{creep_cas4_gdl4}
-	\input{fig_soundbite_palavas/creep_cas4_gdl4.pgf}
+	\input{fig_presentation/creep_cas4_gdl4.pgf}
 \begin{itemize}
-%\item Failure scenario almost conserved: ``fibre bundle'' network
-%\item Other dissipation due to softness: overcompressed gel
-\item More difficult to predict failure time from fractures nucleation time
+\item reversible, homogeneous creep $\rightarrow$ irreversible fracture growth
+\item nucleation regime \alert{NOT universal}
+\begin{itemize}
+	\item possible to reform bond?
+	\item friction between swollen strands?
+	\item glassy dissipation?
+\end{itemize}
 \end{itemize}
@@
 \item Transition to soft glassy rheology
-\item Creep regimes 1 and 2 still valid but messy in the middle
+\item Power-law creep and fracture growth still present but do not rule everything
 \item[$\Rightarrow$] Rupture scenario of critical gel only partially valid for other gels.

### 2016-05-09 abstract_fracture.tex
Abstract for the fracture story, uncluding picture

\documentclass[onecolumn,12pt,tightenlines, a4paper,superscriptaddress,notitlepage]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{kpfonts}
\usepackage{siunitx}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.9}

%\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.5,0.5,0.5}%{0.76,0.36,0.13}
%\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\begin{document}

\title{Yoghurt under stress}
%\date{}

\author{\bf Mathieu Leocmach}
\email{mathieu.leocmach@univ-lyon1.fr}
\affiliation{Institut Lumière Matière, Lyon}
\author{C.~Perge}
\affiliation{Laboratoire de Physique, ENS de Lyon}
\author{T.~Divoux}
%\email{divoux@crpp-bordeaux.cnrs.fr}
\affiliation{CRPP, Bordeaux, France}
\author{S.~Manneville}
\affiliation{Laboratoire de Physique, ENS de Lyon}

\maketitle

Mouthfeel is a complex, multidimensional  sensory experience, whose design is critical to the success of a dish or a food product. Here we investigate the mechanical behaviour of a model yoghurt under large deformation. Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids and to rupture under an external load. Combining optical and ultrasonic imaging to shear rheology we show [Leocmach et al., Phys. Rev. Lett., 2014, 113, 038303] that the failure scenario of an acid-induced sodium caseinate gel is reminiscent of brittle solids: after a primary creep regime characterized by a power-law behavior whose exponent is fully accounted for by linear viscoelasticity, fractures nucleate and grow logarithmically perpendicularly to shear, up to the sudden rupture of the gel. A single equation accounting for those two successive processes nicely captures the full rheological response. The failure time follows a decreasing power law with the applied shear stress, similar to the Basquin law of fatigue for solids. These results are in excellent agreement with recent fibre-bundle models that include damage accumulation on elastic fibres and exemplify protein gels as model, brittlelike soft solids.

\vfill
\begin{center}
%\begin{figure}
\let\mylena\relax
\newlength\mylena
\pgfmathsetlength{\mylena}{0.19\textwidth}
\includegraphics[width=\mylena]{Y110_2013-03-01_02-55-11.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-06-28.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-19-47.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-10.jpg}\hfill
%\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-18.jpg}\hfill
\includegraphics[width=\mylena]{Y110_2013-03-01_03-21-19.jpg}\\

\vspace{0.2em}
\hfill\begin{tikzpicture}
\begin{axis}[
	width=0.975\textwidth,
	height=12\baselineskip,
	xmode=log,
	ymode=log,
	xlabel={$t$ (\si{\second})},
	xmin=0.1,
	ylabel={strain rate $\dot\gamma$ (\si{\per\second})}, 
	ymin=0.9e-4, ymax=2e-1,
	axis on top,
	]
 	\fill[Accent1!20]
		(axis cs:3,1e-2) ellipse[rotate=-17.5, x radius=0.4\textwidth, y radius=0.04\textwidth] node[below, rotate=-17.5, black] {power-law creep}
		(axis cs:0.8e3,2e-4) ellipse[x radius=0.05\textwidth, y radius=0.02\textwidth] node[below, black] {nucleation}
		(axis cs:1577,1e-2) ellipse[x radius=0.04\textwidth, y radius=4\baselineskip] node[below, rotate=90, black] {explosive regime}
		;
	\addplot[thick] table {Y110_300Pa.gdot}; 
	\addplot[thick, only marks, mark=o] coordinates {(687, 2.1e-4) (1476, 7.35e-4) (1509, 1.12e-3)};
	\node[below] at (rel axis cs:0.5,1) {Couette cell $\sigma=\SI{300}{\pascal}$};
\end{axis}
\end{tikzpicture}%
%\caption{Optical visualisation of fractures in a casein gel confined between two concentric cylinders submitted to a constant shear stress (right) with the corresponding local strain field (in \si{\micro\metre}) measured by ultrasound echography in a vertical slice. Scale bars are in millimetres.}
%\end{figure}
\end{center}

%We investigate the robustness of this scenario with varying acidulent concentrations, paying special attention to the case of over-acidified gels previously overlooked in the literature. Qualitative differences are explained by microstructural changes below isoelectri pH that we monitor throughout gelation both by rheology and confocal microscopy. As a whole, our results highlight protein gels as a versatile and model system for the study of plasticity in soft amorphous materials.
\end{document}

### 2016-06-08 creep_beamer.tex
Strain in appendix Gallot 2013 The simplest yoghurt
Adapt to iLM template

@@
 
-\node[shape=cylinder, draw, shape border rotate=90,minimum width=\mydimb, minimum height=7\baselineskip, anchor=north, aspect=2] at (rheometer.north-|webcam.north) (Stator) {};
+\node[shape=cylinder, draw, shape border rotate=90,minimum width=\mydimb, minimum height=7\baselineskip, anchor=north, aspect=2, yshift=-1.5em] at (rheometer.north-|webcam.north) (Stator) {};
 
@@
 \begin{frame}{The yoghurt creep experiment}
+\vspace{1.25\baselineskip}
 \let\mylena\relax
@@
 
-\begin{frame}{Different (constant) stresses: universality}
+\begin{frame}{Different stresses: universality}
 \begin{columns}
+\column{0.4\textwidth}
+\textit{\footnotesize constant stresses}\\
+\vspace{\baselineskip}
+\structure{Normalisation}
+\begin{description}[$\dot\gamma_\text{min}$]
+\item[$\tau_f$] failure time 
+\item[$\dot\gamma_\mathrm{min}$] minimum strain rate 
+\end{description}
+
+\smallskip
+\hfill\begin{beamercolorbox}[rounded=true]{block title}\Large $\Rightarrow$Physical origin?\end{beamercolorbox}
+
 \column{0.6\textwidth}
@@
 \end{tikzpicture}
-
-\column{0.4\textwidth}
-\structure{Normalisation}
-\begin{description}[$\dot\gamma_\text{min}$]
-\item[$\tau_f$] failure time 
-\item[$\dot\gamma_\mathrm{min}$] minimum strain rate 
-\end{description}
-
-\smallskip
-\hfill$\Rightarrow$ \colorbox{Accent1}{\color{white}\Large Physical origin?}
 \end{columns}
@@
 \item Classically explained by disinclination dynamics in crystals
-\item Recently explained by plastic events in glasses \textit{\footnotesize Miguel et al., PRL 2002}
+\item Recently explained by plastic events in glasses\\\hfill \textit{\footnotesize Miguel et al., PRL 2002}
 \end{itemize}
@@
 \bigskip
-\movie[externalviewer]{\structure{Ultrasound velocimetry}}{Yaourt110_primary.avi} $\Rightarrow$ local velocity maps
+\movie[externalviewer]{\structure{Ultrasound velocimetry}}{Yaourt110_primary.avi} $\Rightarrow$ local velocity maps\\\hfill\textit{\footnotesize T. Gallot et al., Rev. Sci. Instrum. 84, 045107 (2013)}
 \begin{itemize}
 	\item homogeneous \movie[externalviewer]{strain field}{Yaourt110_profils.avi} (no wall slip, no shear band)
-	\item if present, plastic events are below our resolution (a few \si{\micro\metre})
+	\item if present, plastic events are below our resolution {\small (a~few~\si{\micro\metre})}
 \end{itemize}
@@
 
-\begin{frame}{Is power-law creep damaging the structure?}
+\begin{frame}{Is creep damaging the structure?}
+\vspace{0.25\baselineskip}
 \begin{itemize}
@@
 \begin{itemize}
-\item Undamaged structure\raisebox{-0.6\baselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.\baselineskip}\right\}\Rightarrow\text{Plasticity cannot explain power law}$}
+\item Undamaged structure\raisebox{-0.6\baselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.\baselineskip}\right\}\Rightarrow\begin{array}{c}\text{Plasticity cannot explain}\\ \text{power law}\end{array}$}
 \item Viscous dissipation
@@
 \begin{frame}{Fibre bundle models}
-	\tikzsetnextfilename{FBM_Jagla}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{FBM_Jagla}%
 	\begin{tikzpicture}
@@
 	\node[starburst, draw=Accent2, inner sep=0pt,  minimum width=0.1\mydima, minimum height=0.5\mydimb, starburst point height=0.15em] at (-0.1\mydima,0) {};
-	\node[double arrow,draw, anchor=west, text width=0.21\textwidth, align=center] at (0.5\mydima,-0.5\mydimb) (model) {\structure{model}};
-	\node[above left=\baselineskip and 0of model.east, text width=0.31\textwidth, inner sep=0] {\begin{itemize}
+		\node[double arrow,draw, anchor=west, text width=0.25\textwidth, align=center] at (0.5\mydima,-0.5\mydimb) (model) {\structure{model}};
+	\node[above left=\baselineskip and 0of model.east, text width=0.35\textwidth, inner sep=0] {\begin{itemize}
 		\item elastic fibres
@@
 	\end{tikzpicture}
-	\includegraphics[height=6\baselineskip, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
+	\includegraphics[width=0.35\textwidth, clip=true, trim=0 0 0 10cm]{MEB_cas4_gdl1_22}
 
@@
 
-	\column{0.3\textwidth}
+	\column{0.34\textwidth}
 	\structure{But}
@@
 
-	\column{0.4\textwidth}\tikzsetnextfilename{creep_1_3}
+	\column{0.4\textwidth}\tikzsetnextfilename{creep_1_3}%
 	\begin{tikzpicture}
@@
 \begin{frame}{Fibre bundle models}
-	\begin{columns}[T]
+	\begin{columns}
 	\column{0.27\textwidth}
-	\textit{\footnotesize Kun et al.  PRL 2008}
+	\vspace{0.5\baselineskip}
+	\textit{\scriptsize Kun et al.  PRL 2008}
 
@@
 
-	\textit{\footnotesize Halász et al., PRE 2012}
+	\textit{\scriptsize Halász et al. PRE 2012}
 
@@
 
-	\tikzsetnextfilename{model_arrow}
+	\tikzsetnextfilename{model_arrow}%
 	\begin{tikzpicture}
@@
 	\column{0.38\textwidth}
-	\tikzsetnextfilename{basquin_fracture}
+	\tikzsetnextfilename{basquin_fracture}%
 	\begin{tikzpicture}
@@
 \item failure involves a single time scale $\tau_f \sim \sigma^{-5.5}$ Basquin law
-\item reversible, homogeneous creep $\rightarrow$ irreversible fracture growth
+\item \begin{minipage}{0.4\textwidth}\center reversible,\\ homogeneous creep\end{minipage} \hfill\textcolor{ilmorange}{$\rightarrow$}\hfill \begin{minipage}{0.4\textwidth}\center irreversible fracture growth\end{minipage}
 \item a single expression captures all the global rheological response
@@
 			group name=g, group size=3 by 1,
-			horizontal sep=3em,
+			horizontal sep=2.5em,
 			},
-		width=0.35\textwidth,
+		width=0.375\textwidth,
 		cycle list name=earthy,

### 2016-06-08 seminaire_PIMM.tex
Adapt to iLM template

--- 2015-10-26 talk_Montpellier.tex
@@
 %\usecolortheme{crane}
-\usefonttheme{structuresmallcapsserif}
+%\usefonttheme{structuresmallcapsserif}
 \setbeamertemplate{navigation symbols}{}
@@
 
-\usecolortheme{rose}
-%\useinnertheme[shadow]{circles}
-\usecolortheme{whale}
-%\useoutertheme{infolines}
+\mode<presentation>{\usetheme{ilm}}
 
-\usecolortheme[named=Accent1]{structure}
-
-\setbeamercolor{alerted text}{fg=Accent2}
+%\usecolortheme{rose}
+%%\useinnertheme[shadow]{circles}
+%\usecolortheme{whale}
+%%\useoutertheme{infolines}
+%
+%\usecolortheme[named=Accent1]{structure}
+%
+%\setbeamercolor{alerted text}{fg=Accent2}
 %\setbeamercolor{palette primary}{fg=white}
-%\setbeamercolor{palette secondary}{bg=Accent1}
+\setbeamercolor{palette secondary}{bg=ilmcolor!10!white}
 %\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}
 
-\setbeamerfont{page number in head/foot}{size=\large}
-\setbeamercolor{page number in head/foot}{fg=Main}
-% page/total
-%\setbeamertemplate{footline}[frame number]
-% pas de total
-\setbeamertemplate{footline}{%
-    	\hfill%
-	\usebeamercolor[fg]{page number in head/foot}%
-	\usebeamerfont{page number in head/foot}%
-	\insertframenumber\kern1em\vskip2pt%
-}
-
-\setbeamersize{text margin left=1em}
-\setbeamersize{text margin right=1em}
+%\setbeamerfont{page number in head/foot}{size=\large}
+%\setbeamercolor{page number in head/foot}{fg=Main}
+%% page/total
+%%\setbeamertemplate{footline}[frame number]
+%% pas de total
+%\setbeamertemplate{footline}{%
+%    	\hfill%
+%	\usebeamercolor[fg]{page number in head/foot}%
+%	\usebeamerfont{page number in head/foot}%
+%	\insertframenumber\kern1em\vskip2pt%
+%}
+%
+%\setbeamersize{text margin left=1em}
+%\setbeamersize{text margin right=1em}
 
@@
 \usepackage[T1]{fontenc}
-\usepackage[oldstylenums]{kpfonts}
+\usepackage{times}
 
@@
 \tikzexternalize
-\tikzsetexternalprefix{fig_presentation/}
+\tikzsetexternalprefix{fig_iLM/}
 %\tikzset{external/optimize=false}
@@
 
-\institute[iLM]{Institut Lumière Matière, Lyon, France}
+\institute[Lyon]{iLM, Univ. Claude Bernard Lyon 1 and CNRS, France}
 \title[Yoghurt under stress]{Acid-induced protein gels: from gelation to stress-induced failure}
 \author[M. Leocmach]{Mathieu Leocmach}
-\date{10 June 2015}
+\date{9 June 2016}
 \titlegraphic{
@@
 
-\begin{frame}[plain]
+\begin{frame}{\pgfuseimage{cnrs-logo}\hspace*{0.3cm}\pgfuseimage{ucbl-logo}}%[plain]
 	\titlepage
@@
 \begin{frame}{Texture as important as taste}
+\vspace{1.25\baselineskip}
 \tikzsetnextfilename{texture}%
@@
 \node(ishichan) {\includegraphics[height=8\baselineskip]{ishichan}};
-\node[above=of ishichan] {\includegraphics[height=8\baselineskip]{calamar}};
-\node[cloud callout,anchor=west, draw, callout relative pointer={(-30:2\baselineskip)}] at ($(ishichan.north)+(-0.5\textwidth,0)$) {\includegraphics[height=5\baselineskip]{aldente}};
+\node[above=1em of ishichan] {\includegraphics[height=8\baselineskip]{calamar}};
+\node[cloud callout,anchor=west, draw, callout relative pointer={(-30:2\baselineskip)}] at ($(ishichan.north)+(-0.5\textwidth-10pt,0)$) {\includegraphics[height=5\baselineskip]{aldente}};
 \node[cloud callout, aspect=0.75, anchor=east, draw, callout relative pointer={(-150:2\baselineskip)}] at ($(ishichan.north)+(0.5\textwidth-10pt,0)$) {\includegraphics[height=5\baselineskip]{chewie}};
@@
 
-\column{0.475\textwidth}
+\column{0.5\textwidth}
 \begin{block}{Brittle fracture}
@@
 \item Sodium caseinate (milk protein) 4\%
-\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right(\}$ stable solution}
+\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ stable solution}
 \item Glucono-$\delta$-lactone (GDL) 1\% $\Rightarrow$ slow homogeneous acidification
@@
 
-\begin{frame}{Linear rheology: Power law (visco)elastic solid}
+\begin{frame}{Power law linear rheology}
 \[\text{stress}\rightarrow \sigma = G \gamma\leftarrow\text{strain}\]
@@
 	\addplot[Main, no marks]{300*x^0.15}  node[midway, below right] {$G^\prime\sim G^{\prime\prime}\sim f^{0.15}$};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y265_cas4_GDL1.txt} node[anchor=base east] {1\%};
-%	\addplot{590*x^0.14};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y277_cas4_GDL1.25.txt} node[anchor=base east] {1.25\%};
-%	\addplot{350*x^0.12};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y275_cas4_GDL1.5.txt} node[anchor=base east] {1.5\%};
-%	\addplot{246*x^0.1};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y268_cas4_GDL2.txt} node[anchor=base east] {2\%};
-%	\addplot{163*x^0.067};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y270_cas4_GDL3.txt} node[anchor=base east] {3\%};
-%	\addplot{109*x^0.044};
-%	\addplot+[only marks, mark=*, forget plot] table{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot+[only marks, mark=o, forget plot] table[y=LossModulus]{freqsweep_Y271_cas4_GDL4.txt} node[anchor=base east] {4\%};
-%	\addplot{76*x^0.035};
-
+	\node[below right] at (rel axis cs:0.1,0.9) {(visco)elastic solid};
 \end{loglogaxis}
@@
 
-\begin{frame}{Can we generalise to biogels? \textup{\normalsize(physical gels
-$E_{bond}\gg k_B T$)}}
+\begin{frame}{Can we generalise to biogels?}
+\hfill\textup{\footnotesize\usebeamercolor[fg]{titlelike}(physical gels
+$E_{bond}\gg k_B T$)}
 \begin{tabu}{cX[c]c}
@@
 
-\begin{frame}{Over-acidified protein gels: beyond isoelectric pH}
+\begin{frame}{Acidifying casein solutions with \textup{HCl}}
+\vspace{1.25\baselineskip}
+\tikzsetnextfilename{dosage}%
+\input{fig_presentation/dosage.pgf}
+\end{frame}
+
+\begin{frame}{Over-acidified protein gels}
+	\hfill beyond isoelectric pH
 	\tikzsetnextfilename{prise_cas4}
@@
 
-\begin{frame}{Linear rheology, from power-law to soft glassy}
-	\tikzsetnextfilename{sweep_cas4}
+\begin{frame}{Linear rheology $\rightarrow$ to soft glassy}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{sweep_cas4}%
 \input{fig_soundbite_palavas/sweep_cas4.pgf}
 	\begin{itemize}
-	\item Near isoelectric, power law rheology $G^\prime, G^{\prime\prime}$ typical of critical gels 
-	\item Over-acidification lowers exponent and creates a minimum in $G^{\prime\prime}$
-	\item Transition to soft-glassy behaviour \\\textit{\footnotesize Sollich, PRE (1998); Mason, Curr. Opin. Colloid Interface Sci. (1999)}
+	\item Near isoelectric, power law rheology typical of critical gels 
+	\item Over-acidification: exponent $\searrow$ \hfill\textcolor{ilmorange}{\&}\hfill minimum in $G^{\prime\prime}$
+	\item Transition from power-law to soft-glassy behaviour \\\textit{\footnotesize Sollich, PRE (1998); Mason, Curr. Opin. Colloid Interface Sci. (1999)}
 	\end{itemize}
@@
 
-\begin{frame}{Microscructure: fluorescent confocal microscopy}
-	\tikzsetnextfilename{structure_confocal}
+\begin{frame}{Microscructure: confocal microscopy}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{structure_confocal}%
 \input{fig_soundbite_palavas/structure_confocal.pgf}
@@
 	\begin{itemize}
-	\item Maximum $G^\prime$ corresponds to maximum contrast between phases
+	\item Maximum $G^\prime$\hfill\textcolor{ilmorange}{$\leftrightarrow$}\hfill Maximum contrast between phases
 	\item Pore size $\xi$ decrease during over-acidification
-	\item<2> Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy
+	\item<2> Casein aggregates swell $\sim$ over compressed $\Rightarrow$ soft glassy
 	\end{itemize}
@@
 
-\begin{frame}{Three creep regimes + softness\hfill\textsc{gdl} 4\%}
+\begin{frame}{3 creep regimes + softness\textsc{gdl} 4\%}
 \vfill
@@
 \end{itemize}
+
+\bigskip
+\begin{center}
+\structure{\huge Thank you for your attention}
+\end{center}
 \end{frame}
-
-%\begin{frame}{Merciless yoghurt breakers}
-%\textit{Leocmach et al. PRL 113, 038303 (2014), \footnotesize ArXiv 1401.8234}
-%
-%\bigskip
-%\begin{tabu}{X[c]X[c]c}
-%\includegraphics[height=0.3\textheight]{Chris}&
-%\includegraphics[height=0.3\textheight]{Thibaut}&
-%\includegraphics[height=0.3\textheight]{Seb}\\
-%Christophe Perge & Thibaut Divoux & Sebastien Manneville\\
-%PhD student & CNRS researcher & Professor\\
-%E.N.S. Lyon &  CRPP Bordeaux & E.N.S. Lyon\\
-%\end{tabu}
-%
-%
-%\bigskip
-%\begin{description}[Special thanks]
-%\item[Special thanks] to Alan Parker (Firmenich) for prompting this study and providing casein \& GDL
-%\end{description} 
-%
-%\end{frame}
 
@@
 
+\begin{frame}{Basquin law rescaling}
+	\vspace{0.5\baselineskip}
+	\tikzsetnextfilename{basquin_rescale}%
+	\input{fig_soundbite_palavas/basquin.pgf}
+\end{frame}
+
+\begin{frame}{Monkman-Grant}
+	\vspace{0.5\baselineskip}
+	\hfill Predict failure from nucleation
+	\tikzsetnextfilename{MonkmanGrant}
+	\input{fig_soundbite_palavas/MonkmanGrant.pgf}
+\end{frame}
+
+\begin{frame}{Wavelength}
+	\vspace{1.25\baselineskip}
+	\tikzsetnextfilename{lambda}%
+	\input{fig_soundbite_palavas/lambda.pgf}
+	\textcolor{Accent1}{$\blacktriangleright$} $\lambda\propto e$ compatible with nucleation rate
+
+\end{frame}
+
 \setcounter{framenumber}{\value{finalframe}}

### 2016-06-08 abstract_fracture.tex
Add DOI link to PRL

@@
 Biomaterials such as protein or polysaccharide gels are known to behave qualitatively as soft solids and to rupture under an external load.
-Combining optical and ultrasonic imaging to shear rheology we show [Leocmach et al., Phys.
+Combining optical and ultrasonic imaging to shear rheology we show [\href{http://link.aps.org/doi/10.1103/PhysRevLett.113.038303}{Leocmach et al., Phys.
 Rev.
-Lett., 2014, 113, 038303] that the failure scenario of an acid-induced sodium caseinate gel is reminiscent of brittle solids: after a primary creep regime characterized by a power-law behavior whose exponent is fully accounted for by linear viscoelasticity, fractures nucleate and grow logarithmically perpendicularly to shear, up to the sudden rupture of the gel.
+Lett., 2014, 113, 038303}] that the failure scenario of an acid-induced sodium caseinate gel is reminiscent of brittle solids: after a primary creep regime characterized by a power-law behavior whose exponent is fully accounted for by linear viscoelasticity, fractures nucleate and grow logarithmically perpendicularly to shear, up to the sudden rupture of the gel.
 A single equation accounting for those two successive processes nicely captures the full rheological response.

### 2016-06-13 creep_beamer.tex
changes made in the train
shorter talk for CECAM workshop Only PRL story. No confocal, so neither Thomas nor PLATIM.

@@
 	\addplot table[x index=1, y index=3]{Y39_1000Pa_gdot_decimated.txt} node (s1000){};
-	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.378*x^(-0.85)} node[Accent2,pos=0.75, below left] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
+	\addplot[yellow, ultra thick, domain={1e-4:0.1}] {0.378*x^(-0.85)} node[Accent2,pos=0.75, below left, font=\footnotesize] {$\frac{\dot{\gamma}(t)}{\dot{\gamma}_\text{min}}\sim \left(\frac{t}{\tau_f}\right)^{-0.85}$};
 \end{loglogaxis}
@@
 \item failure involves a single time scale $\tau_f \sim \sigma^{-5.5}$ Basquin law
-\item \begin{minipage}{0.4\textwidth}\center reversible,\\ homogeneous creep\end{minipage} \hfill\textcolor{ilmorange}{$\rightarrow$}\hfill \begin{minipage}{0.4\textwidth}\center irreversible fracture growth\end{minipage}
+\item $\left.\begin{minipage}{0.4\textwidth}\center reversible,\\ homogeneous creep\end{minipage}\right\rbrace$ \hfill\textcolor{ilmorange}{$\rightarrow$}\hfill $\left\lbrace\begin{minipage}{0.4\textwidth}\center irreversible fracture growth\end{minipage}\right.$
 \item a single expression captures all the global rheological response

### 2016-06-13 talk_CECAM.tex
shorter talk for CECAM workshop Only PRL story. No confocal, so neither Thomas nor PLATIM.

--- 2016-06-08 seminaire_PIMM.tex
@@
 %\setbeamercolor{palette primary}{fg=white}
-\setbeamercolor{palette secondary}{bg=ilmcolor!10!white}
+%\setbeamercolor{palette secondary}{bg=ilmcolor!10!white}
 %\setbeamercolor{palette tertiary}{bg=Accent2,fg=white}
@@
 \institute[Lyon]{iLM, Univ. Claude Bernard Lyon 1 and CNRS, France}
-\title[Yoghurt under stress]{Acid-induced protein gels: from gelation to stress-induced failure}
+\title{Yoghurt under stress}%[Yoghurt under stress]{Acid-induced protein gels: from gelation to stress-induced failure}
 \author[M. Leocmach]{Mathieu Leocmach}
-\date{9 June 2016}
+\date{15 June 2016}
 \titlegraphic{
-	\begin{tabu}{X[c]X[c]X[c]X[c]}
+	\begin{tabu}{X[c]X[c]X[c]}
 		\includegraphics[height=3\baselineskip]{Chris}&
-		\includegraphics[height=3\baselineskip]{Thomas}&
 		\includegraphics[height=3\baselineskip]{Thibaut}&
 		\includegraphics[height=3\baselineskip]{Seb}\\
-		Christophe Perge & Thomas\linebreak Gibaud & Thibaut\linebreak Divoux & Sebastien Manneville\\
+		Christophe\linebreak Perge & Thibaut\linebreak Divoux & Sebastien Manneville\\
 	\end{tabu}
@@
 	\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
-	\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
+	%\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
 	\includegraphics[height=2\baselineskip]{CNRSfilaire-Q}\quad
@@
 
+\appendix
+\newcounter{finalframe}
+\setcounter{finalframe}{\value{framenumber}}
+
+\begin{frame}[plain]
+\vspace{6\baselineskip}
+\begin{center}
+\structure{\huge Thank you for your attention}
+\end{center}
+\vspace{6\baselineskip}
+\textit{\footnotesize Leocmach, Perge, Divoux \& Manneville. Creep and fracture of a protein gel under stress. Phys Rev Lett. 2014;113(3):038303.}
+\end{frame}
+
 \begin{frame}{Can we generalise to biogels?}
@@
 
-\appendix
-\newcounter{finalframe}
-\setcounter{finalframe}{\value{framenumber}}
-
-\begin{frame}[plain]
-\end{frame}
-
 \begin{frame}{Creep rheology: Three regimes}
